### 2022-09-26 paper/paper.tex
Create repo for paper and code

\documentclass[runningheads]{llncs}

\usepackage{graphicx}
\usepackage{amsmath,amssymb}
\usepackage[labelfont=bf]{caption}
\usepackage{subcaption}
\captionsetup{compatibility=false}
\usepackage[table]{xcolor}
\usepackage{tikz}
\usepackage{siunitx}
\usepackage{hyperref}
\usepackage{tabularx}
\usepackage{pifont}
\usepackage{tablefootnote}
\usepackage{array}

\newcommand{\cmark}{\ding{51}}
\newcommand{\xmark}{\ding{55}}

\renewcommand{\arraystretch}{1.5}
%\newcolumntype{L}[1]{>{\raggedright\arraybackslash}p{#1}}
\setlength{\textfloatsep}{12pt}

\usepackage[ruled,vlined,linesnumbered,norelsize]{algorithm2e}
\DontPrintSemicolon
\def\NlSty#1{\textnormal{\fontsize{8}{10}\selectfont{}#1}}
\SetKwSty{texttt}
\SetCommentSty{emph}
\def\listalgorithmcfname{List of Algorithms}
\def\algorithmautorefname{Algorithm}
\let\chapter=\section % repariert ein Problem mit algorithm2e

\begin{document}

\title{Active-Learning for SAT-Solver Benchmarking}
\author{Tobias Fuchs}
\institute{Institute of Theoretical Informatics, \\ Karlsruhe Institute of Technology, Germany \\
\email{info@tobiasfuchs.de}}

\maketitle

\begin{abstract}
  Experimentation is one of the main phases in algorithm engineering, including the archetypal $\mathcal{NP}$-complete propositional-sat\-is\-fia\-bi\-li\-ty problem (SAT).
  While the field of SAT-solver benchmarking is well established, traditional benchmark-set-selection approaches do not optimize for benchmark runtime.
  Benchmark-set selection is the process of choosing a subset of instances out of a pool to be representative and reliable in discriminating SAT solvers.
  This paper presents an active-learning-based benchmark-set-selection approach that solves the new-solver problem:
  Given a new solver, we want to find its ranking amidst its competitors with as little runtime as possible while still giving accurate ranking predictions.
  To do so, we use well-known techniques from the domain of active learning.
  We can determine a new solver's rank regarding the SAT Competition 2022 Anniversary Track instances in about \SI{10}{\%} of the time it would take to run the solver on all instances of this dataset.
  We achieve this while still maintaining a prediction accuracy of about \SI{92}{\%} regarding the PAR-2 ranking of this solver if it were evaluated on the complete dataset.
  Instance-selection decisions made by our approach also allow for discussing instance-family importance in the selection process. 
  Our tool provides the engineer of propositional-satisfiability solvers with a faster way of determining a new SAT-solver's performance in comparison to sampling the benchmark set by hand.
\end{abstract}

\section{Introduction}
\label{sec:intro}
One of the main phases of algorithm engineering is experimentation.
This is al\-so true for the archetypal $\mathcal{NP}$-complete pro\-po\-si\-tio\-nal-sat\-is\-fia\-bi\-li\-ty problem (SAT).
While state-of-the-art solvers embark upon more and more problem instances, SAT-solver evaluation is steadily getting more costly.
Usually, submissions to competitions like the annual SAT~Competition~\cite{FroleyksHIJS21} themselves precede heavy experimentation and are the remnant of a multitude of ideas and considerations.
While the field of SAT-solver benchmarking is well established, traditional benchmark-set-selection approaches~\cite{Gelder11,HoosKSS13} do not optimize for benchmark runtime.
Their goal is to select a representative and robust benchmark set for discriminating solvers reliably.
Within our work, we show that we can arrive at the same conclusions with high accuracy in only a fraction of the time that would be needed to evaluate the full benchmark set.
To be more specific, we address the following problem.

\textbf{New solver problem.} Given a pool of instances $\mathcal{I}$ and solvers $\mathcal{A}$ with already known runtimes $r\!: \mathcal{A} \times \mathcal{I} \rightarrow \left[0, \tau\right] \cup \left\lbrace \perp \right\rbrace$ (timeout $\tau$ usually \SI{5000}{s}~\cite{FroleyksHIJS21}), the problem is to decide with as little resources (CPU time) as possible how a new solver $\hat{a} \notin \mathcal{A}$ performs in comparison to the existing ones by incrementally selecting a benchmark set $\tilde{\mathcal{I}}$ on which to evaluate $\hat{a}$.
We want to find a scoring function $\operatorname{score}\!: \mathcal{A} \cup \left\lbrace \hat{a} \right\rbrace \rightarrow \mathbb{R}$ assigning each solver a score.
Having a set of solvers with their respective scores induces a ranking.
This allows for both comparing the solvers as well as identifying solvers with similar behavior.
While the annual SAT~Competitions~\cite{FroleyksHIJS21} rely on the PAR-2 score ranking, i.e, the average runtime with a penalty timeout-term of $2\tau$, which we want to predict, this work examines a model-based way of ranking solvers.

This current article is proposing a novel active-learning-based benchmark-set-selection approach.
Our algorithm is unique in that it is the only one, to the best of our knowledge, to address all of the following features (also refer to Tab.~\ref{tab:requirements}).
Our benchmarking process outputs a \textit{ranking} to compare the new solver against the existing set of solvers as described above.
We optimize the \textit{runtime} that our strategy needs to arrive at its conclusion.
We use problem-instance and runtime \textit{features} $\varphi\!: \mathcal{A} \times \mathcal{I} \rightarrow \mathbb{R}^n$ since it already has proven to be helpful in classic benchmark-set-selection approaches~\cite{HoosKSS13} as well as in portfolio selection~\cite{CollauttiMMO13,NgokoCT19}.
Moreover, we select instances \textit{non-randomly} and \textit{incrementally}.
By doing so, we have a greater impact on the properties of the benchmarking approach such as its required runtime.
Sampling instances incrementally means taking runtime information from already done experiments into account when choosing the next.
Rather than solely outputting a binary classification, i.e, the new solver is worse than solver A, we provide a \textit{scoring function} that shows by which margin a solver is worse and how similar it is to existing approaches (\textit{interpretability}).
Finally, our strategy's cost \textit{scales linearly} with the cardinality of the set of existing solvers $\mathcal{A}$ to compare against and the instances $\mathcal{I}$ within our pool of possible benchmark instances.
We avoid comparing pairs of solvers.
This would scale quadratically.

To sum up our contribution, we take an existing benchmark set, e.g., the SAT Competition~2022 Anniversary Track~\cite{sat2022}, and predict a new solver's PAR2 score on this dataset with only a fraction of the runtime that would be needed to evaluate this solver on the complete set of instances.

For reproducibility, our source code, all experiments, and data are available on GitHub\footnote{\url{https://github.com/mathefuchs/al-sat-benchmarking-experiments}}.

\section{Related Work}
This section discusses work in several related fields.
First, we take a look at traditional SAT-solver benchmarking.
Second, we explore how SAT-solver configuration systems deal with navigating possible configurations.
Third, we discuss an existing benchmark-set-selection approach that optimizes for benchmark runtime.
Finally, we show how active-learning techniques can help to tackle the aforementioned \textit{new-solver} problem.

Tab.~\ref{tab:requirements} shows a concise overview of existing approaches with a related goal.
While traditional benchmark-set-selection~\cite{HoosKSS13} lacks the aspect of runtime optimization, related work within algorithm configuration~\cite{HutterHL11} is concerned with finding better configurations fast.
However, algorithm configuration is not suited for ranking.
There is already existing work employing active learning for SAT-solver discrimination~\cite{MatriconAFSH21}.
This approach has some major shortcomings though.
Finally, our approach seeks to provide all stated features.
Note that all approaches already make use of problem-instance features as it showed to be beneficial in past work~\cite{HoosKSS13,CollauttiMMO13,NgokoCT19}.

\subsubsection{Traditional benchmark-set-selection.}
Recent studies~\cite{abs-2107-07002,NiesslHWCB22} show that selecting a benchmark set in general is a difficult endeavor for multiple reasons.
Dehghani~et.~al.~\cite{abs-2107-07002} describe how a biased benchmarking set can easily lead to fallacious interpretations.
Moreover, Nießl~et.~al.~\cite{NiesslHWCB22} state that benchmark-set design has many movable parts, e.g., performance measures, aggregation, and handling of missing values.
Questionable research practices might modify those design elements a-posteriori to fit expectations, leading to biased results.

Hoos et. al.~\cite{HoosKSS13} discuss which properties are most desirable when selecting a set of SAT benchmark instances, for instance for a competition, e.g., the SAT Competition~2020~\cite{FroleyksHIJS21}.
The selection strategy presented is static, i.e., it does not depend on the solvers to distinguish.
Desirable properties are a variety of instances (avoids over-fitting of solver approaches), adapted instance hardness (not too easy but also not too hard; solvers cannot be distinguished if too many time-out or finish immediately), and no duplicates (having duplicates gives those instances more weight).
To avoid having too similar instances, they make use of a distance-based approach using the SATzilla features~\cite{XuHHL08,features}.
Those features comprise general instance-size features and graph-representation features among others.
Their approach \textit{ranks} solvers by providing its PAR-2 score, is \textit{feature-based}, and is \textit{scalable} in the sense that the evaluation time scales linearly with more instances and solvers.
Also, their approach is \textit{interpretable} in the sense that the obtained PAR-2 score induces a distance between solvers.
However, they do not optimize for benchmark \textit{runtime} and select instances \textit{randomly}\footnote{Random with constraints on the instance hardness and feature distance.}.

%Our approach should optimally also have these three properties, i.e., a variety of instances without duplicates that are neither too easy nor too hard.
%While this selection strategy is useful for creating diverse benchmark sets for competitions, it does not optimize for runtime.
%It is expected that dynamic selection strategies outperform static ones regarding the runtime of experiments. 
%Dynamic strategies chose instances incrementally, one at a time, by integrating runtime data (of the instances that the solver has been already evaluated on) to make subsequent selection decisions rather than completely choosing the benchmark set upfront.

% In terms of our desired features (also refer to Tab.~\ref{tab:requirements}),

%The aforementioned SATzilla features~\cite{XuHHL08,features} are provided within the GBD meta-data database~\cite{IserS18}.
%Finding a good representation of instances is quintessential for our work since any feature-based benchmark-set-selection approach is only as good as the quality of the employed features.

\begin{table}[tbp]
  \centering
  \caption{Support for desired features that are introduced in Sec.~\ref{sec:intro}. The shown approaches include traditional benchmark-set-selection techniques~\cite{HoosKSS13}, an algorithm configuration system~\cite{HutterHL11}, an existing active-learning approach~\cite{MatriconAFSH21}, as well as our approach.}
  \label{tab:requirements}
  \vspace{0.2cm}
  \begin{tabular}{
    m{0.2\textwidth}
    >{\centering\arraybackslash}m{0.15\textwidth}
    >{\centering\arraybackslash}m{0.15\textwidth}
    >{\centering\arraybackslash}m{0.15\textwidth}
    >{\centering\arraybackslash}m{0.15\textwidth}
  }
    \hline
    Feature & Hoos \mbox{et. al.~\cite{HoosKSS13}} & SMAC~\cite{HutterHL11} & Matricon \mbox{et. al.~\cite{MatriconAFSH21}} & Our approach \\
    \hline
    Ranking & \cmark & \xmark & \cmark & \cmark \\
    Runtime \mbox{optimization} & \xmark & \cmark & \cmark & \cmark \\
    %Feature-based & \cmark & \cmark & \cmark & \cmark \\
    Incremental \mbox{non-random} \mbox{sampling} & \xmark & \xmark & \cmark & \cmark \\
    % \tablefootnote{While using a model-based selection strategy to select the next configuration to evaluate, the instance sampling is done randomly.}
    Interpretability & \cmark & \xmark & \xmark & \cmark \\
    % \tablefootnote{While they provide a confidence measure, the strategy's output is still only binary, i.e., whether a solver A is better or worse than solver B.}
    Scalability & \cmark & \cmark & \xmark & \cmark \\
    \hline
  \end{tabular}
\end{table}

\subsubsection{SAT-solver configuration.}
Another interesting approach can be found within algorithm-configuration systems~\cite{HutterHL11,HoosHL21,Stutzle0P22}.
The goal is to tune SAT-solvers for a given sub-domain of problem instances.

SMAC~\cite{HutterHL11} tries to find a close-to-optimal configuration for a given SAT-solver on a given set of instances.
Although this task is different from our goal, e.g., we do not need to navigate configuration space, they use a similar framework.
There are mainly two similarities.
First, both approaches are model-based.
While SMAC tries to gradually build a model of the solver's performance as a function of the configuration parameters, we want to gradually build a model of a new solver's performance based on already known runtimes of instances and runtime information on already known solvers.
Second, both approaches employ some kind of selection strategy that is based upon the previously mentioned model.
Theirs selects configurations rather than instances though.
Furthermore, Hutter et. al.~\cite{HutterHL11} use an approach that makes use of exploitation (what configuration parameters are already good) and exploration (parameter space with little data points) to select the next configuration for which to acquire runtime information. SMAC's configuration sampling has also proven to be better than random sampling in all tested scenarios and has even proven to be significantly better in most cases~\cite{HutterHL11}.

While this approach considers the configuration's \textit{runtime}, is \textit{feature-based}, and \textit{scales linearly} by having a fixed time budget, it is not capable of the other three features listed in Tab.~\ref{tab:requirements}.
First, an algorithm configurator cannot be used to \textit{rank} a new solver since it only looks at promising solvers/configurations. If the new solver does not belong to this category (new ideas are seldomly the best approach right from the start, i.e., they need some tweaking), we do not have any information on how this solver performed.
Second, while using a model-based selection strategy to sample configurations, instance sampling is done \textit{randomly}.
And third, their approach is not \textit{interpretable} regarding the new solver performance since we are already missing the feature to \textit{rank} the new solver among existing ones.

\subsubsection{Incremental benchmark-set-selection.}
Matricon et. al.~\cite{MatriconAFSH21} present an incremental benchmark-set-selection approach.
Given a pair of SAT solvers, they want to iteratively select a subset of instances to decide with high confidence which of the two solvers is better.
The selection of instances is dependent on the choice of the solvers to distinguish.
They calculate a scoring metric for all instances (that have not been selected yet), run the experiment with the highest score, and update the confidence. If the confidence exceeds a threshold (e.g., 95\%), they stop iterating.
Otherwise, the aforementioned steps are repeated until the desired confidence level is reached or the pool of instances is exhausted.
Most of the proposed strategies in~\cite{MatriconAFSH21} are instance-based methods, i.e., no model of the solver runtimes is built.

For the scoring metric, they discuss (1) a random strategy as a baseline, (2) a discrimination strategy based on the work of Gent et. al.~\cite{GentHJKMNN14}, (3) a variance-based strategy (select instance with the highest variance in its distribution of running times), (4) an information-gain-based method, and (5) a feature-distance-based strategy.
To update the confidence level after each experiment, they discuss using a fixed subset size, a Wilcoxon test, and a distribution-based method. Their experiments show that random instance-selection with a Wilcoxon test as the stopping criterion performs consistently well across different data sets in the sense that it belongs to the Pareto-front or is close to it in all cases.

Regarding our desired features of adaptive benchmark-set-selection (Tab.~\ref{tab:requirements}), the aforementioned approach already ticks off the first four features, i.e., it \textit{ranks} solvers, optimizes for \textit{runtime}, is \textit{feature-based}, and uses \textit{incremental non-random} sampling.
Although they provide a confidence metric for their prediction, which makes this approach partially \textit{interpretable}, it is not clear how similar two given solvers are or on which instances they behave similarly, as the prediction is solely binary.
Moreover, the major shortcoming lies within it being not \textit{scalable} with the number solvers.
As they compare only pairs of solvers, scaling with more and more solvers and ideas becomes an issue as there is a quadratic amount of pairs to distinguish. Since a new solver idea is seldomly best at first try but requires more or less tweaking, it is desired to rank a new solver in a way that scales linearly with the number of existing approaches.
In this way, a new solver idea can be discarded if it performs poorly across the board or may be further tweaked if it shows promising results at least in some cases.

\subsubsection{Active learning (AL).}
The posed \textit{new solver problem} has stark similarities to the well-studied field of active learning within recommender systems, especially the \textit{new user problem}~\cite{RubensESK15}.
The classic exploration-exploitation dilemma arises since, on the one hand, we want to maximize the utility an instance provides to our model and, on the other hand, minimize the cost (CPU time) that is associated with its acquisition.

\begin{figure}[tbp!]
  \centering
  \begin{tabular}[c]{ccc}
  \begin{subfigure}[b]{0.26\textwidth}
  \centering
  \resizebox{!}{2.85cm}{
  \begin{tikzpicture}
  \draw[fill=black]  (-2.8,3.7) ellipse (0.1 and 0.1);
  \draw[fill=black]  (-2.6,3.4) arc (0:180:0.2);
  \node[right] at (-5.9,3.6) {Agent (Solver)};
  \draw[fill=white]  (-3.3,2.8) rectangle (-2.5,2.2);
  \draw[fill=white]  (-3.2,2.7) rectangle (-2.4,2.1);
  \draw[fill=white]  (-3.1,2.6) rectangle (-2.3,2);
  \node[right] at (-5.9,2.4) {Training Data};
  \node (v1) at (-2.8,3.4) {};
  \node (v2) at (-2.8,2.8) {};
  \draw[->]  (v1) edge (v2);
  \node[right] at (-2.8,3.1) {$x, y$};
  \node (v3) at (-2.8,2) {};
  \node (v4) at (-2.8,1.4) {};
  \draw[->]  (v3) edge (v4);
  \draw  (-3.1,1.4) rectangle (-2.5,1);
  \draw  (-2.4,3.2) rectangle (-2.4,3.2);
  \node[right] at (-5.9,1.2) {Learning Alg.};
  \node (v5) at (-2.8,1) {};
  \node (v6) at (-2.8,0.4) {};
  \draw[->]  (v5) edge (v6);
  \node[right] at (-5.9,0.25) {Approx. Function};
  \node at (-2.8,0.25) {$f$};
  \end{tikzpicture}
  }
  \caption{Passive.}
  \label{fig:passive}
  \end{subfigure}
  &
  \begin{subfigure}[b]{0.36\textwidth}
  \centering
  \resizebox{!}{2.85cm}{
  \begin{tikzpicture}
  \draw[fill=black]  (-2.8,3.7) ellipse (0.1 and 0.1);
  \draw[fill=black]  (-2.6,3.4) arc (0:180:0.2);
  \node[right] at (-5.9,3.6) {Agent (Solver)};
  \draw[fill=white]  (-3.3,2.8) rectangle (-2.5,2.2);
  \draw[fill=white]  (-3.2,2.7) rectangle (-2.4,2.1);
  \draw[fill=white]  (-3.1,2.6) rectangle (-2.3,2);
  \node[right] at (-5.9,2.4) {Training Data};
  \node (v1) at (-2.8,3.4) {};
  \node (v2) at (-2.8,2.8) {};
  \draw[->]  (v1) edge (v2);
  \node[right] at (-2.8,3.1) {$x, y$};
  \node (v3) at (-2.8,2) {};
  \node (v4) at (-2.8,1.4) {};
  \draw[->]  (v3) edge (v4);
  \draw  (-3.1,1.4) rectangle (-2.5,1);
  \draw  (-2.4,3.2) rectangle (-2.4,3.2);
  \node[right] at (-5.9,1.2) {Learning Alg.};
  \node (v5) at (-2.8,1) {};
  \node (v6) at (-2.8,0.4) {};
  \draw[->]  (v5) edge (v6);
  \node[right] at (-5.9,0.25) {Approx. Function};
  \node at (-2.8,0.25) {$f$};
  \node[right] at (-1.2,2.4) {Pool};
  \draw[->, densely dashed] (-2.4,1.2) arc (-90:0:0.8);
  \draw[->, densely dashed] (-1.6,2.8) arc (0:90:0.8);
  \node[right] at (-1.6,3.135) {$x, ?$};
  \draw  (-1.6,2.4) ellipse (0.4 and 0.3);
  \draw[fill=black]  (-1.725,2.5) ellipse (0.07 and 0.07);
  \draw[fill=black]  (-1.625,2.25) ellipse (0.07 and 0.07);
  \draw[fill=black]  (-1.425,2.4) ellipse (0.07 and 0.07);
  \end{tikzpicture}
  }
  \caption{Active Pool-based.}
  \label{fig:activepool}
  \end{subfigure}
  &
  \begin{subfigure}[b]{0.36\textwidth}
  \centering
  \resizebox{!}{2.85cm}{
  \begin{tikzpicture}
  \draw[fill=black]  (-2.8,3.7) ellipse (0.1 and 0.1);
  \draw[fill=black]  (-2.6,3.4) arc (0:180:0.2);
  \node[right] at (-5.9,3.6) {Agent (Solver)};
  \draw[fill=white]  (-3.3,2.8) rectangle (-2.5,2.2);
  \draw[fill=white]  (-3.2,2.7) rectangle (-2.4,2.1);
  \draw[fill=white]  (-3.1,2.6) rectangle (-2.3,2);
  \node[right] at (-5.9,2.4) {Training Data};
  \node (v1) at (-2.8,3.4) {};
  \node (v2) at (-2.8,2.8) {};
  \draw[->]  (v1) edge (v2);
  \node[right] at (-2.8,3.1) {$x, y$};
  \node (v3) at (-2.8,2) {};
  \node (v4) at (-2.8,1.4) {};
  \draw[->]  (v3) edge (v4);
  \draw  (-3.1,1.4) rectangle (-2.5,1);
  \draw  (-2.4,3.2) rectangle (-2.4,3.2);
  \node[right] at (-5.9,1.2) {Learning Alg.};
  \node (v5) at (-2.8,1) {};
  \node (v6) at (-2.8,0.4) {};
  \draw[->]  (v5) edge (v6);
  \node[right] at (-5.9,0.25) {Approx. Function};
  \node at (-2.8,0.25) {$f$};
  \node[right] at (-1.1,2.4) {GAN};
  \node[right] at (-1.3,3.135) {$x, ?$};
  \draw[->, densely dashed] (-2.4,1.2) arc (-90:90:1.2);
  \end{tikzpicture}
  }
  \caption{Active Synthesis-based.}
  \label{fig:activesynth}
  \end{subfigure}
  \end{tabular}

  \caption{Types of Learning (depiction inspired by Rubens~et.al.~\cite{RubensESK15}).}
  \label{fig:learning}
\end{figure}

% Starting with only unlabeled data, e.g., a new user has not yet rated any content, an active-learning algorithm iteratively shows the items that are best suited to refine the user's model based on his previous interaction.
% In contrast to traditional passive machine-learning methods (Fig.~\ref{fig:passive}), active-learning approaches can actively select for which items to acquire labels~\cite{RubensESK15}.
% Since the potential runtime of experiments is by magnitudes larger than the model's training time, we only consider incrementing our benchmark set by one instance at a time rather than using batches which is also proposed in current active-learning advances~\cite{SinhaED19,TranDRC19}.

AL algorithms can be categorized into \textit{synthesis-bas\-ed} \cite{0001AEMN22,GarzonMG22,2019gaal} and \textit{pool-bas\-ed} approaches \cite{GolbandiKL11,distribAL,HarpaleY08,KapoorGUD07,KorenBV09,KornerW06,MelvilleM04,SinhaED19,TongK01}.
While synthesis-based methods (Fig. \ref{fig:activesynth}) generate instances that might be interesting to label, pool-based methods (Fig. \ref{fig:activepool}) rely on a pool of unlabeled instances from which to draw items.

Recent synthesis-based methods within the field of satisfiability solving~\cite{0001AEMN22,GarzonMG22} show how to generate problem instances with desired properties.
This goal is, however, orthogonal to ours.
While those approaches want to generate problem instances on which a solver is good or bad, we want to predict whether a solver is good or bad on an existing benchmark set.

Pool-based methods can be further refined into \textit{Error-based} \cite{GolbandiKL11,KorenBV09}, \textit{Un\-cer\-tain\-ty-based} \cite{HarpaleY08,KapoorGUD07,TongK01}, \textit{Distribution-based} methods \cite{distribAL,SinhaED19} and \textit{Ensembles} \cite{KornerW06,MelvilleM04}.
Error-based strategies pick instances that have the most influence on the model parameters.
Uncertainty-based methods sample instances close to their perceived decision boundary.
Distribution-based approaches sample instances whose features are unaccounted for amongst the already sampled instances.
Finally, ensembles are a meta-strategy that combines the predictions of several different active-learning models into a selection choice.

\subsubsection{Summary.}
To sum up this section, we have shown both traditional and recent benchmark-set-selection approaches. Moreover, we have briefly outlined how active-learning techniques can help to incrementally select benchmark sets. In the following sections, we present an active-learning-based benchmark-set-selection approach that solves the \textit{new solver problem} while optimizing for runtime and being both interpretable and scalable.

\section{Active-Learning for SAT-Solver Benchmarking}
\label{sec:main}
Our algorithm follows a simple three-step schema (Alg.~\ref{algALBenchmark}). After initializing the model $\mathcal{M}$, we incrementally \textit{select} a problem instance from the pool of instances, one at a time, evaluate the new solver $\hat{a}$ on it, and \textit{update} the model with the acquired result. We \textit{stop} when meeting some stopping condition.
Since the potential runtime of experiments is by magnitudes larger than the model's update time, we only consider incrementing our benchmark set by one instance at a time rather than using batches.
% For now, we also restrict ourselves to the aforementioned sequential one-instance-at-a-time process.
At the core of the model $\mathcal{M}$ is a prediction function $f\!: \left(\mathcal{A} \cup \left\lbrace \hat{a} \right\rbrace\right) \times \mathcal{I} \rightarrow \mathcal{V}$ that powers the decisions within the previously mentioned three steps.
Note that the prediction target values $\mathcal{V}$ are unspecified to allow for more flexibility within our particular approaches.
Refer to the next sections for more details.

\begin{algorithm}
  \caption{Simple incremental benchmark-set-selection schema.}
  \label{algALBenchmark}

  \KwIn{A pool of instances $\mathcal{I}$, a set of known solvers $\mathcal{A}$ with known runtimes $r\!: \mathcal{A} \times \mathcal{I} \rightarrow \left[0, \tau\right] \cup \left\lbrace \perp \right\rbrace$, and a new solver $\hat{a} \notin \mathcal{A}$.}

  $\mathcal{M} \leftarrow \textrm{init model with $\mathcal{A}$, $\mathcal{I}$, $r$, and $\hat{a}$}$

  \While{$\operatorname{not} \, \operatorname{stop}\!\left(\mathcal{M}\right)$}{
    $s \leftarrow \operatorname{selectNextInstance}\!\left(\mathcal{M}\right)$

    $t \leftarrow \operatorname{runExperiment}\!\left(\hat{a},\,  s\right)$  \tcp*{Runs $\hat{a}$ on $s$ with time-out $\tau$}

    $\operatorname{updateModel}\!\left(\mathcal{M},\, s,\, t\right)$
  }

  $\operatorname{score} \leftarrow \operatorname{predictScores}\!\left(\mathcal{M}\right)$

  \KwOut{Return scoring function $\operatorname{score}\!: \mathcal{A} \cup \left\lbrace \hat{a} \right\rbrace \rightarrow \mathbb{R}$ and the fraction of sampled instances and runtime needed to provide it.}
\end{algorithm}

The following section (Sec.~\ref{sec:main1}) discusses how we may derive a solver ranking from our model (line~6).
Thereafter, we look at possible sampling criteria (line~3) in Sec.~\ref{sec:main2} and, finally, possible stopping conditions (line~2) in Sec.~\ref{sec:main3}.

\subsection{Predicting Rankings}
\label{sec:main1}
In order to predict a ranking of solver $\hat{a}$ amidst its competitors, we need to infer its performance on all instances, even the ones on which we have not evaluated it.
To do so, there are generally two notions of this problem, i.e., regression and classification.
A regression model tries to predict the \mbox{PAR-2} runtime of $\hat{a}$ on all instances based on the samples that have been selected.
Then, the ranking is induced by the average predictions.
Alternatively, a classification model can be used to predict some discrete runtime labels of $\hat{a}$ on all instances.
Similarly, the average runtime label determines the ranking.

Preliminary experiments, as well as research in portfolio solvers~\cite{NgokoCT19,CollauttiMMO13}, have shown however that directly predicting runtimes does not work well in practice.
In our experiments, the mean squared error was within the same magnitude as the values to be predicted.
In contrast to that, discretizing runtimes and using those labels in the model's training and prediction yields quite good results (cf.~Sec.~\ref{sec:eval}).
The following paragraphs motivate why it makes sense to discretize runtimes.

By discretizing runtimes into $k$ clusters, we mean using a function $\gamma\!: {\mathcal{A} \times \mathcal{I}} \rightarrow \left\lbrace 1, \dots, k \right\rbrace$ to transform runtime observations.
Note that this transformation is done per instance to adapt to different instance difficulties. 
Also, we define ${\gamma\!\left(a, i\right) := k}$ iff $r\!\left(a, i\right) = \,\perp$ for all $a \in \mathcal{A}$ and $i \in \mathcal{I}$.
To determine a ranking based on those labels, we define
\begin{equation}
  \operatorname{score}\!\left(a\right) := \frac{1}{|\mathcal{I}|} \sum_{i \in \mathcal{I}} w_{a,i} \cdot \left(\gamma\!\left(a, i\right) - 1\right) \enspace \textrm{,}
  \label{eq:rankingeq}
\end{equation}
with $w_{a,i} := 2$ (PAR-\textit{2} penalty factor) if $\gamma\!\left(a, i\right) = k$ else $w_{a,i} := 1$. To be applicable to $\hat{a}$, we extend $\operatorname{score}$ by replacing $\gamma\!\left(a, i\right)$ by the model's predictions $f\!: \left(\mathcal{A} \cup \left\lbrace \hat{a} \right\rbrace\right) \times \mathcal{I} \rightarrow \left\lbrace 1, \dots, k \right\rbrace$.

Turning a regression into a classification problem, discretization is a useful tool.
However, it does not provide much value if the resulting cluster labels are not predictive of the PAR-2 ranking we seek to predict.
In our work, we use a hierarchical-clustering approach with a single-link criterion for $k = 3$.
For the SAT Competition~2022 Anniversary Track instances~\cite{sat2022}, this ranking approach is capable of deciding almost all pairs of solvers correctly (about \SI{97.45}{\%}; $\sigma = \SI{3.68}{\%}$).
Moreover, it decides all solvers correctly if the difference in the cluster-label-ranking scores between two solvers is at least \SI{0.11}{}.
Refer to Tab.~\ref{tab:discreteranking} for a comparison with the PAR-2 ranking.
The Spearman correlation~\cite{de2016comparing} of cluster-label and PAR-2 ranking is about \SI{0.988} (in comparison to a perfect value of \SI{1}). This shows that monotonicity and score distances between ranks align quite well between both ranking methods.

% In that sense, we also experiment with a combination of runtime-label ranking and the PAR-2 ranking of the subsample if the difference in label ranking scores drops below a certain threshold.
% Besides that, we additionally consider using the past $l$ model predictions in a voting approach for stability.

\begin{table}[tbp]
  \centering
  \caption{PAR-2 ranking vs. ranking based on discretized runtimes for the SAT Competition~2022 Anniversary Track instances~\cite{sat2022} ($k = 3$; hierarchical clustering).}
  \label{tab:discreteranking}
  \vspace{0.2cm}
  \begin{tabular}{
    c
    >{\centering\arraybackslash}p{2.2cm}p{3.6cm}
    >{\centering\arraybackslash}p{1.8cm}p{3.3cm}
  }
    \hline
    Rank & \multicolumn{2}{c}{PAR-2 Ranking} & \multicolumn{2}{c}{Cluster-Label Ranking} \\
    \hline
1   & 2806.400 & Kissat\_MAB\_ESA           & 1.172 & Kissat\_MAB\_ESA \\
2   & 2810.959 & kissat-sc2022-bulky        & 1.183 & kissat-sc2022-bulky \\
3   & 2830.143 & ekissat-mab-gb-db          & 1.185 & ekissat-mab-gb-db \\
\cellcolor{gray!15} 4   & \cellcolor{gray!15} 2833.081 & \cellcolor{gray!15} Kissat\_MAB\_UCB           & \cellcolor{gray!15} 1.186 & \cellcolor{gray!15} kissat\_inc \\
\cellcolor{gray!15} 5   & \cellcolor{gray!15} 2833.397 & \cellcolor{gray!15} kissat\_inc                & \cellcolor{gray!15} 1.186 & \cellcolor{gray!15} Kissat\_MAB\_UCB \\
6   & 2841.734 & ekissat-mab-db-v1          & 1.192 & ekissat-mab-db-v1 \\
7   & 2842.666 & Kissat\_MAB\_MOSS          & 1.192 & Kissat\_MAB\_MOSS \\
8   & 2857.219 & Kissat\_MAB-HyWalk         & 1.195 & Kissat\_MAB-HyWalk \\
9   & 2866.960 & kissat-sc2022-light        & 1.120 & kissat-sc2022-light \\
10  & 2896.371 & kissat-els-v2              & 1.216 & kissat-els-v2 \\
11  & 2950.179 & hKis-unsat                 & 1.229 & hKis-unsat \\
12  & 2964.729 & Kissat\_adaptive\_restart  & 1.234 & Kissat\_adaptive\_restart \\
13  & 2974.990 & SeqFROST-NoExtend          & 1.247 & SeqFROST-NoExtend \\
14  & 3010.642 & Cadical\_ESA               & 1.248 & Cadical\_ESA \\
\cellcolor{gray!15} 15  & \cellcolor{gray!15} 3012.152 & \cellcolor{gray!15} kissat-els-v1              & \cellcolor{gray!15} 1.258 & \cellcolor{gray!15} CadicalReorder \\
\cellcolor{gray!15} 16  & \cellcolor{gray!15} 3028.922 & \cellcolor{gray!15} CadicalReorder             & \cellcolor{gray!15} 1.264 & \cellcolor{gray!15} kissat-els-v1 \\
\cellcolor{gray!15} 17  & \cellcolor{gray!15} 3042.540 & \cellcolor{gray!15} cadical\_rel\_Scavel       & \cellcolor{gray!15} 1.278 & \cellcolor{gray!15} CaDiCaL\_DVDL\_V1 \\
\cellcolor{gray!15} 18  & \cellcolor{gray!15} 3082.293 & \cellcolor{gray!15} kissat\_relaxed            & \cellcolor{gray!15} 1.283 & \cellcolor{gray!15} CaDiCaL\_DVDL\_V2 \\
\cellcolor{gray!15} 19  & \cellcolor{gray!15} 3089.304 & \cellcolor{gray!15} CaDiCaL\_DVDL\_V1          & \cellcolor{gray!15} 1.298 & \cellcolor{gray!15} kissat\_relaxed \\
\cellcolor{gray!15} 20  & \cellcolor{gray!15} 3095.118 & \cellcolor{gray!15} CaDiCaL\_DVDL\_V2          & \cellcolor{gray!15} 1.361 & \cellcolor{gray!15} cadical\_rel\_Scavel \\
21  & 3272.674 & glucose-reboot             & 1.379 & glucose-reboot \\
\cellcolor{gray!15} 22  & \cellcolor{gray!15} 3287.322 & \cellcolor{gray!15} LStech-Maple-HyWalk        & \cellcolor{gray!15} 1.414 & \cellcolor{gray!15} hCaD\_V1-psids \\
23  & 3288.798 & LSTech\_Maple              & 1.446 & LSTech\_Maple \\
\cellcolor{gray!15} 24  & \cellcolor{gray!15} 3399.281 & \cellcolor{gray!15} SLIME-SC-2022-beta         & \cellcolor{gray!15} 1.469 & \cellcolor{gray!15} LStech-Maple-HyWalk \\
\cellcolor{gray!15} 25  & \cellcolor{gray!15} 3410.345 & \cellcolor{gray!15} SLIME-SC-2022              & \cellcolor{gray!15} 1.470 & \cellcolor{gray!15} SLIME-SC-2022-beta \\
\cellcolor{gray!15} 26  & \cellcolor{gray!15} 3430.403 & \cellcolor{gray!15} hCaD\_V1-psids             & \cellcolor{gray!15} 1.523 & \cellcolor{gray!15} SLIME-SC-2022 \\
27  & 3504.244 & MapleLCMDistChrBt-DL-v3    & 1.542 & MapleLCMDistChrBt-DL-v3 \\
28  & 4750.808 & IsaSAT                     & 2.062 & IsaSAT \\
    \hline
  \end{tabular}
\end{table}

That discretized runtimes are sufficient for distinguishing solvers can also be shown by comparing the fraction of solver pairs whose runtimes are significantly different.
Within the SAT~Competition~2022 Anniversary Track, about \SI{89.95}{\%} of solver pairs are significantly different (according to a Wilcoxon-signed-rank test with $\alpha = 0.05$).
While only time-out labels (solver times out or not on a particular instance) are sufficient to significantly distinguish about \SI{87.04}{\%} of solver pairs, discrete-runtime labels\footnote{Based on hierarchical clustering with a log-single-link criterion for $k = 3$.} are sufficient to distinguish about \SI{87.83}{\%} of solver pairs.
This means that we do not lose much discriminatory power using discrete runtime labels.
The number of significant comparisons between solvers \textit{alone}, though, does only offer limited insights as we might achieve significant differences in all pairs by assigning some randomly skewed values as labels.
However, the predictiveness of the PAR-2 score, the labels' discriminatory power, as well as that it enables us to use classification at the same time, show the clear advantage of runtime discretization.

\subsection{Instance Sampling}
\label{sec:main2}
Our instance-sampling strategies (also refer to Alg.~\ref{algALBenchmark}; line 3) can be divided into two categories.
First, we present possible baseline approaches including existing work on how to pick \textit{interesting} instances~\cite{Bossek021a,GentHJKMNN14,MatriconAFSH21}.
Second, we show promising machine-learning-based sampling strategies that make use of the model's label prediction function $f$ and are inspired by existing work within the realms of active learning~\cite{settles2009active}.

\subsubsection{Baselines and existing approaches.}
\label{sec:sampling1}
Apart from obvious baseline instance-selection strategies, i.e., random sampling and neighborhood-aware sampling (randomly choosing among the instances with the highest amount of non-sampled $k$ neighbors), we also implement existing instance-selection strategies.

Bossek and Wagner~\cite{Bossek021a} introduce a custom ranking-score metric incorporating the number of \textit{good} pairs and \textit{bad} pairs.
A good solver pair, regarding a particular instance, are solvers that perform according to their global ranking.
So if solver $a$ is better than solver $b$ on instance $i$ and $a$ is also in total better than $b$, it is a \textit{good} solver pair on $i$.
Otherwise, it is a \textit{bad} pair.
Based on this classification, they define a 3-tuple for each instance consisting of the number of good pairs, the accumulated performance differences of bad pairs, and the accumulated performance differences of good pairs.
By selecting instances with the highest score according to the lexicographical order of the tuples, they try to favor instances that are predictive of the final ranking.
This assumes though that the new solver $\hat{a}$ behaves similarly to the existing solvers.

Gent~et.~al.~\cite{GentHJKMNN14} make use of a concept known from racing for automated model selection.
Given a particular instance and a parameter $\rho$, a solver is $\rho$-dominated if it is at least $\rho$ times worse than the best solver on this problem instance.
By selecting instances with the highest amount of $\rho$-dominated solvers, they hope to sample instances that provide the best discriminatory power to distinguish the runtimes of a new contestant.

Finally, we also implement some of the selection approaches described by Matricon~et.al.~\cite{MatriconAFSH21}.
Besides uniformly random sampling and the aforementioned discrimination-based method by Gent~et.~al.~\cite{GentHJKMNN14}, they also describe a variance-based selection approach.
The variance-based strategy selects instances with the highest runtime variance per average runtime. We also implement a logarithmized version of this, i.e., the highest log-runtime variance per average log-runtime.

\subsubsection{Active-learning-based instance selection.}
\label{sec:sampling2}
For conciseness, we only present two of our active-learning-based instance-selection approaches that show the most promising results.
We implement a model-uncertainty-based approach and a model-based information-gain sampling strategy.
%, and a voting-based committee-disagreement approach.

These methods require the model's predictions to also include probabilities for the $k$ possible runtime labels, i.e., 
\begin{equation}
  f'\!: \left(\mathcal{A} \cup \left\lbrace \hat{a} \right\rbrace\right) \times \mathcal{I} \rightarrow \left[0, 1\right]^k  \enspace \textrm{.}
\end{equation}
The model-uncertainty-based approach simply selects the instance that is closest to the model's decision boundary and, of course, not already selected, i.e.,
\begin{equation}
  \underset{i \in \mathcal{I} \setminus \tilde{\mathcal{I}}}{\arg\min} \left\lvert \frac{1}{k} - \max_{n \in \left\lbrace 1, \dots, k \right\rbrace} f'\!\left(\hat{a}, i\right)_{n} \right\rvert \enspace \textrm{.}
\end{equation}

The model-based information-gain sampling strategy selects the instance with the highest expected information-entropy reduction, assuming that the model's runtime-label probabilities for an instance are close to being correct.
To be more specific, we select the instance $i$ that maximizes
\begin{equation}
  \operatorname{IG}\!\left(i\right) := \operatorname{H}\!\left(i\right) - \sum_{n = 1}^{k} f'\!\left(\hat{a}, i\right)_{n} \operatorname{H}\!\left(i, n\right) \enspace \textrm{,}
\end{equation}
whereby $\operatorname{H}\!\left(i\right)$ denotes the entropy of the runtime labels $\gamma\!\left(a, i\right)$ for all $a \in \mathcal{A}$ and $\operatorname{H}\!\left(i, n\right)$ denotes the entropy of the runtime labels including the runtime label of $\hat{a}$ which is $n$.
The term $\operatorname{H}\!\left(i, n\right)$ is computed for every possible runtime label $n$.

%Finally, the voting-based committee-disagreement approach requires a committee of $c$ models.
%Rather than having a single prediction function $f$, there is a separate function $f_j$ for all committee members $j \in \left\lbrace 1, \dots, c  \right\rbrace$.
%We then select the instance $i$ whose entropy of runtime-label votes by the different ensemble members is the highest.

\subsection{Stopping Criteria}
\label{sec:main3}
Similar to the instance-selection strategies, we can also divide stopping criteria into baseline and active-learning-based approaches.
Also, refer to~Alg.~\ref{algALBenchmark}, line~2.
While a fixed subset-size criterion, e.g., only sample \SI{20}{\%} of all available instances, is an example of the former, a stopping criterion based on a Wilcoxon signed-rank test or the convergence of the model's predictions is an example for the latter.

The Wilcoxon stopping criterion stops the iterative active-learning algorithm when we are on average certain enough that the predicted runtime labels are sufficiently different in comparison to the other solvers.
Note that this is a rather heuristic criterion, it shows promising results nevertheless.
To be more specific, we stop when
\begin{equation}
  W_{\hat{a}} := \frac{1}{\left\lvert \mathcal{A} \right\rvert} \sum_{a \in \mathcal{A}} \operatorname{wilcoxon}_{\alpha}\!\left(\gamma\!\left(a, \mathcal{I}\right), f\!\left(\hat{a}, \mathcal{I}\right) \right)
\end{equation}
drops below a certain threshold.
Thereby, \textit{wilcoxon} denotes the $p$-value of a Wilcoxon signed-rank test of the given runtime-label distributions and $\gamma\!\left(a, \mathcal{I}\right)$ and $f\!\left(\hat{a}, \mathcal{I}\right)$ are the expected set extensions on $\gamma$ and $f$ respectively.
To improve the stability of this criterion, we use an exponential moving average
\begin{equation}
  W_{\exp}^{\left(m\right)} := \beta W_{\hat{a}} + \left(1 - \beta\right) W_{\exp}^{\left(m - 1\right)} \enspace \textrm{,}
\end{equation}
whereby $m$ denotes the count of iterations within Alg.~\ref{algALBenchmark} and $W_{\exp}^{\left(0\right)} := 1$.

The model-convergence stopping criterion is less sophisticated in comparison.
It simply stops the active-learning process (Alg.~\ref{algALBenchmark}) if within the last $l$ iterations the ranking that is induced by the model's predictions (Eq.~\ref{eq:rankingeq}) remains unchanged.
Note that the concrete values of $\operatorname{score}\!\left(a\right)$ might change.
We are solemnly interested in the induced ranking in this case.

\section{Experimental Design}
Given all the previously presented instantiations for Alg.~\ref{algALBenchmark}, this section briefly outlines our experimental design, including our optimization goal, evaluation framework, used data sets, hyper-parameter choices, as well as implementation details.
% hyper-parameter values that we experiment with

\subsection{Optimization Goal}
\label{sec:goal}
As already stated in the introductory section, this work addresses the \textit{new solver problem}.
By running Alg.~\ref{algALBenchmark}, we obtain a prediction model $\mathcal{M}$ that provides us with a scoring function \textit{score}.
Also, it outputs the fraction of runtime that is needed to arrive at its conclusion.
First and foremost, our goal is to provide the engineer of new SAT solvers with an accurate ranking.
We define the ranking accuracy $O_{\operatorname{acc}} \in \left[0, 1\right]$ (higher is better) by the fraction of pairs $\left(\hat{a}, a\right)$ for all $a \in \mathcal{A}$ that are decided correctly by the given ranking.
For now, we exclude the equality of solvers since data shows that the PAR-2 scores of all our solvers are different.
So, possible ranking decisions may be, for example, solver $a$ is better than $b$ or $b$ is better than $a$.
The ranking accuracy is affected by whether this decision is correctly made.
Matricon et. al.~\cite{MatriconAFSH21} also evaluate their approach by the fraction of solver pairs that is correctly decided.
Second, we also have to optimize for runtime.
The fraction of runtime that the algorithm needs to arrive at its conclusion is denoted by $O_{\operatorname{rt}} \in \left[0, 1\right]$ (lower is better).
In that sense, we want to find an approach that maximizes
\begin{equation}
  O_\delta := \delta O_{\operatorname{acc}} + \left(1 - \delta\right) \left(1 - O_{\operatorname{rt}}\right) \enspace \textrm{,}
  \label{eq:opt}
\end{equation} 
whereby $\delta \in \left[0, 1\right]$ allows for linear weighting between the two optimization goals $O_{\operatorname{acc}}$ and $O_{\operatorname{rt}}$.
Plotting the approaches that maximize $O_\delta$ for all $\delta \in \left[0, 1\right]$ on a $O_{\operatorname{rt}}$-$O_{\operatorname{acc}}$-diagram provides us with a Pareto front of the best approaches for different optimization-goal weightings.

\subsection{Evaluation Framework}
\label{sec:evalframe}
To evaluate a concrete instantiation of Alg.~\ref{algALBenchmark} (a concrete choice for all the sub-routines), we perform cross-validation on our set of solvers.
That means that each solver once plays the role of the new solver.
Alg.~\ref{alg:eval} shows this.
Note that the \textit{new} solver in each iteration is excluded from the set of solvers $\mathcal{A}$ to avoid data leakage.
Also, the runtime function $r$ is restricted to $\mathcal{A} \times \mathcal{I}$.

\vspace{-0.2cm}
\begin{algorithm}[htbp]
  \caption{Evaluation framework.}
  \label{alg:eval}

  \KwIn{A pool of instances $\mathcal{I}$ and a set of solvers $\mathcal{S}$ with known runtimes $r\!: \mathcal{S} \times \mathcal{I} \rightarrow \left[0, \tau\right] \cup \left\lbrace \perp \right\rbrace$.}

  $res \leftarrow \emptyset$

  \For{$\hat{a} \in \mathcal{S}$}{
    $\mathcal{A} \leftarrow \mathcal{S} \setminus \left\lbrace \hat{a} \right\rbrace$

    $\left( ranking,\, runtime \right) \leftarrow \operatorname{runALAlgorithm}\!\left(\mathcal{I}, \mathcal{A}, \left.r\right|_{\mathcal{A} \times \mathcal{I}}, \hat{a}\right)$ \tcp*{Refer to Alg.~\ref{algALBenchmark}}

    $O_{\operatorname{acc}} \leftarrow \operatorname{computeRankingAccuracy}\!\left( ranking,\, \mathcal{I},\, \mathcal{S},\, r \right)$ \tcp*{Refer to Sec.~\ref{sec:goal}}

    $O_{\operatorname{rt}} \leftarrow \operatorname{computeRuntimeFractionUsed}\!\left( runtime,\, \mathcal{I},\, \mathcal{S},\, r \right)$

    $res \leftarrow res \cup \left\lbrace \left( O_{\operatorname{acc}},\, O_{\operatorname{rt}} \right) \right\rbrace$
  }

  $\left( \bar{O}_{\operatorname{acc}},\, \bar{O}_{\operatorname{rt}} \right) \leftarrow \operatorname{mean}\!\left( res \right)$

  \KwOut{Return the average ranking accuracy $\bar{O}_{\operatorname{acc}}$ and the average fraction of runtime needed $\bar{O}_{\operatorname{rt}}$.}
\end{algorithm}
\vspace{-0.2cm}

\subsection{Data}
In our experiments, we work with the SAT~Competition~2022 Anniversary Track instances~\cite{sat2022}.
The dataset consists of 5355 instances with respective runtime data.
It has complete runtimes of 28 solvers.
We use the GBD metadata database~\cite{IserS18} to provide us with the aforementioned instance features, problem instances, and solver runtimes.
Thereby, all our approaches make use of the 56 base features that are maintained in the \textit{base} database of GBD~\cite{IserS18}.
They are inspired by the SATzilla features~\cite{features}.
Those features comprise general instance-size features and graph-representation features among others.
All features are numeric and fortunately free of any missing values.
We drop 10 out of 56 features because of zero variance.
For hyper-parameter tuning, we randomly sample \SI{10}{\%} of the complete set of 5355 instances with stratification regarding the instance's family.
All instance families that are too \textit{small}, i.e., \SI{10}{\%} of it is less than one, are put into one meta-family for stratification.
This smaller dataset allows for a more extensive exploration of hyper-parameter space.

\subsection{Hyper-parameters}
\label{sec:hyper}
Given Alg.~\ref{algALBenchmark}, there are several possible instantiations for the three phases, i.e., \textit{selection}, \textit{stopping}, and \textit{ranking}.
Also, there are different choices for the runtime-label prediction model.
Note that we are not considering all previously listed approaches since a grid search of all combinations would be infeasible. Rather, we filter approaches based on preliminary experimental results (cf. Sec.~\ref{sec:evalprel}) and do the main end-to-end experiments only with a subset.
The end-to-end experiment configurations are given below.

\subsubsection{Ranking.}
Regarding \textit{ranking} (cf. Sec.~\ref{sec:main1}), we experiment with the following approaches, including our used hyper-parameter values:
\begin{itemize}
  \item Observed PAR-2 ranking of already sampled instances
  \item Predicted ranking induced by runtime-label predictions
  \begin{itemize}
    \item History size: consider the latest 1, 10, 20, 30, or 40 predictions within a voting approach for stability.
    \item Fallback threshold: if the difference of ranking scores drops below \SI{0.01}, \SI{0.05}, or \SI{0.1}, use the partially observed PAR-2 ranking as a tie-breaker.
  \end{itemize}
\end{itemize}

\subsubsection{Selection.}
For \textit{selection} (cf. Sec.~\ref{sec:sampling1}), we experiment with the following methods, including our used hyper-parameter values:
\begin{itemize}
  \item Random sampling 
  \item Model-based uncertainty sampling
  \begin{itemize}
    \item Fallback threshold: use random sampling for the first \SI{0}{\%}, \SI{5}{\%}, \SI{10}{\%}, \SI{15}{\%}, or \SI{20}{\%} of instances to explore the instance space.
    \item Runtime scaling: prefer instances with the greatest uncertainty per average runtime (\texttt{True} or \texttt{False}).
  \end{itemize}

  \item Model-based information-gain sampling
  \begin{itemize}
    \item Fallback threshold: use random sampling for the first \SI{0}{\%}, \SI{5}{\%}, \SI{10}{\%}, \SI{15}{\%}, or \SI{20}{\%} of instances to explore the instance space.
    \item Runtime scaling: prefer instances with the greatest information-gain per average runtime (\texttt{True} or \texttt{False}).
  \end{itemize}
\end{itemize}

\subsubsection{Stopping.}
For \textit{stopping} decisions (cf. Sec.~\ref{sec:main3}), we experiment with the following criteria, including our used hyper-parameter values:
\begin{itemize}
  \item Fixed subset size of \SI{10}{\%} or \SI{20}{\%} of instances
  \item Ranking convergence criterion
  \begin{itemize}
    \item Minimum amount: sample at least \SI{2}{\%}, \SI{8}{\%}, \SI{10}{\%}, or \SI{12}{\%} of instances before applying the criterion.
    \item Convergence duration: stop if the predicted ranking stays the same for the last \SI{1}{\%} or \SI{2}{\%} of sampled instances.
  \end{itemize}

  \item Wilcoxon criterion
  \begin{itemize}
    \item Minimum amount: sample at least \SI{2}{\%}, \SI{8}{\%}, \SI{10}{\%}, or \SI{12}{\%} of instances before applying the criterion.
    \item Average of $p$-values to drop below: \SI{5}{\%}.
    \item Exponential-moving average: incorporate previous significance values by using an EMA with $\beta = 0.1$ or $\beta = 0.7$.
  \end{itemize}
\end{itemize}

\subsubsection{Runtime-label prediction model.}
As a runtime-label prediction model, we only make use of one method that has proven to be successful in preliminary experiments (cf.~Sec.~\ref{sec:evalprel}).
Our model of choice is an ensemble, stacking a quadratic-discriminant analysis\footnote{With a regularization parameter of zero and a singular value threshold of zero} onto a random forest model\footnote{With \textit{entropy} splitting criterion and \textit{balanced} class weights} using a decision tree\footnote{With \textit{GINI} impurity splitting criterion; choosing the best split for each branching decision and maximum tree depth of 5 levels} to weight its members.
If not stated explicitly, all other model parameters default to the presets of \textit{scikit-learn}\footnote{\url{https://scikit-learn.org/stable/index.html}; Version 1.0.2}~\cite{scikit-learn}.

%the proposed runtime label ranking, (2) the PAR-2 subsample ranking, and (3) the runtime label ranking that uses the PAR-2 subsample ranking as a fallback when performance differences drop below a threshold. Possible values are \SI{0.01}{}, \SI{0.05}{}, and \SI{0.1}{}.
%Also, we experiment with different lengths of the prediction history that is used for stabilizing the ranking. Possible values are 10, 20, 30, 40, and 50.

\subsection{Implementation Details}
For reproducibility, our source code, all experiments, and data are available on GitHub\footnote{\url{https://github.com/mathefuchs/al-sat-benchmarking-experiments}}.
It is implemented in \textsc{Python} using \textit{scikit-learn}~\cite{scikit-learn} for making predictions and \textit{gbd-tools}~\cite{IserS18} for SAT-instance retrieval.

\section{Evaluation}
\label{sec:eval}
Now that we are all set, we want to, first, discuss preliminary results including the choice of the runtime-discretization method and the choice of the machine-learning predictor.
Thereafter, we look at end-to-end results and instance-family importance.

\subsection{Preliminary Results}
\label{sec:evalprel}
An exhaustive grid search of all approaches outlined in Sec.~\ref{sec:main} is infeasible.
We are, therefore, narrowing down the search by filtering based on preliminary results.
Since all relevant decisions, i.e., \textit{selection}, \textit{stopping}, and \textit{ranking}, revolve around the choice of prediction model, we are first looking at the \textit{Matthews-Correlation coefficient} (MCC) scores~\cite{gorodkin2004comparing,matthews1975comparison} that a particular prediction model achieves.
MCC scores are great in dealing with class imbalances in contrast to conventional metrics like accuracy.

\begin{table}
  \centering
  \caption{Time-out prediction Matthews-Correlation coefficient performances. Results are obtained using cross-validation. We provide \SI{64}{\%} of ground-truth for training, use \SI{16}{\%} as a validation set, and \SI{20}{\%} as a hold-out test set (5-fold cross-validation).}
  \label{tab:mcc}
  \begin{tabular}{lc}
    Prediction Model & Avg. MCC (Std.) \\
    \hline
    Stacking ensemble (QDA, RF)           & 0.9527 (0.0292) \\
    Quadratic-Discriminant analysis (QDA) & 0.9290 (0.0339) \\
    Random forest (RF)                    & 0.8530 (0.0479) \\
    AdaBoost with decision trees          & 0.8384 (0.0444) \\
    AdaBoost with logistic regression     & 0.8158 (0.0985) \\
    Decision tree                         & 0.8059 (0.0707) \\
    Logistic regression                   & 0.8052 (0.1018) \\
    kNN-Classifier                        & 0.7885 (0.1521) \\
    MLP-Classifier                        & 0.7760 (0.1408) \\
    Support-Vector classifier             & 0.7757 (0.2149) \\
    Naive Bayes                           & 0.7306 (0.1394) \\
    \hline
  \end{tabular}
\end{table}

% Stacking MCC
% 0.888221577313471
% 0.04753979056930393

To obtain the results in Tab.~\ref{tab:mcc}, we perform cross-validation among all solvers: we assume that all other solver runtimes are known and provide \SI{64}{\%} of ground-truth runtime labels of the target solver for training, use \SI{16}{\%} as a validation set for hyper-parameter tuning, and \SI{20}{\%} as a hold-out test set (5-fold cross-validation).
This is repeated for each solver once.
Note that we perform cross-validation on two levels, i.e., among the solvers and the target solver runtimes.
We report the average performances on the hold-out test sets.
Since the differences in the MCC-performance distributions of the best and second-best approaches are statistically significant regarding a Wilcoxon signed-rank test with $\alpha = \SI{5}{\%}$, we use the former for all further experiments as a prediction model.

As a runtime discretization technique, we use hierarchical clustering with $k = 3$ using a log-single-link criterion for all further experiments.
Each non-time-out runtime starts in a separate interval.
We then gradually merge intervals whose single-link logarithmic distance is the smallest until the desired number of partitions is reached.
As already mentioned in Sec.~\ref{sec:main1}, this approach has good properties regarding the predictiveness of the PAR-2 score ranking and its discriminatory power.
Also, preliminary experiments showed that our stacking ensemble performed particularly well for this kind of discretization method.
Other clustering approaches that we have tried include hierarchical clustering with mean-, median- and complete-link criteria, as well as $k$-means and spectral clustering.
Predicting runtime labels in a similar experiment as above produced an MCC score of on average \SI{88.82}{\%} ($\sigma = \SI{4.75}{\%}$; in comparison to \SI{95.27}{\%} for predicting time-outs).

\subsection{Experimental Results}
\subsubsection{Tuning our algorithm.}
Our main end-to-end experiments follow the evaluation framework that has been introduced in Sec.~\ref{sec:evalframe}.
Fig.~\ref{fig:e2eallsolvers} reveals the performance of the approaches listed in Sec.~\ref{sec:hyper} on $O_{\operatorname{rt}}$-$O_{\operatorname{acc}}$-diagrams for the smaller hyper-parameter tuning dataset.
The plotted lines represent the Pareto-front for each possible choice of ranking approaches (Fig.~\ref{fig:annitraincolorranking}), selection approaches (Fig.~\ref{fig:annitraincolorselection}), and stopping criteria (Fig.~\ref{fig:annitraincolorstopping}).
Colors indicate the different hyper-parameter instantiations.
To be more specific, they represent the Pareto-front of all configurations with the respective instantiations.

Regarding the ranking approach (Fig.~\ref{fig:annitraincolorranking}), using the ranking that is induced by our runtime-label predictions consistently performs better than the partially observed PAR-2 ranking for each possible value of $\delta$.
%This was expected since a possible bias in the selection decisions of the model is accounted for when using the model's predictions for ranking.
This is expected since selection decisions are not random.
For example, we might sample more instances of one family if it benefits discrimination of solvers.
While the partially observed PAR-2 score is skewed, the prediction model can account for this.

Regarding the selection approaches (Fig.~\ref{fig:annitraincolorselection}), our model-based uncertainty strategy performs best in most cases.
However, our model-based information-gain sampling is beneficial if runtime is strongly favored (very small $\delta$; runtime fraction less than \SI{5}{\%}).

Regarding the stopping criterion (Fig.~\ref{fig:annitraincolorstopping}), the ranking convergence criterion performs the most consistently well in most cases.
If PAR-2 accuracy is strongly favored (very high $\delta$), the Wilcoxon stopping criterion performs better.

\begin{figure}[tbp!]
  \centering
  \begin{subfigure}{1.0\textwidth}
    \resizebox{\textwidth}{!}{
      \graphicspath{{../plots/}}
      \input{../plots/anni_train_color_ranking.pgf}
    }
    \caption{$O_{\operatorname{rt}}$-$O_{\operatorname{acc}}$-diagram showing Pareto-fronts for configurations with the given ranking approaches. The lines show the best performances of approaches using runtime-label-prediction ranking and partially observed PAR-2 ranking respectively.}
    \label{fig:annitraincolorranking}
  \end{subfigure}
  \\
  \vspace{0.2cm}
  \begin{subfigure}{1.0\textwidth}
    \resizebox{\textwidth}{!}{
      \graphicspath{{../plots/}}
      \input{../plots/anni_train_color_selection.pgf}
    }
    \caption{$O_{\operatorname{rt}}$-$O_{\operatorname{acc}}$-diagram showing Pareto-fronts for configurations with the given instance-selection approaches. The lines show the best performances of approaches using uncertainty-based, information-gain-based, and random instance-selection respectively.}
    \label{fig:annitraincolorselection}
  \end{subfigure}
  \\
  \vspace{0.2cm}
  \begin{subfigure}{1.0\textwidth}
    \resizebox{\textwidth}{!}{
      \graphicspath{{../plots/}}
      \input{../plots/anni_train_color_stopping.pgf}
    }
    \caption{$O_{\operatorname{rt}}$-$O_{\operatorname{acc}}$-diagram showing Pareto-fronts for configurations with the given stopping criteria. The lines show the best performances of approaches using the ranking-convergence criterion, Wilcoxon criterion, and a fixed subset size respectively.}
    \label{fig:annitraincolorstopping}
  \end{subfigure}

  \caption{$O_{\operatorname{rt}}$-$O_{\operatorname{acc}}$-diagrams showing the performance of different hyper-parameter instantiations. The x-axis shows the fraction of runtime in comparison to the time needed to evaluate a solver on all instances. The y-axis shows the accuracy of the predicted rank in comparison to the true PAR-2 rank. Each line entails the front of Pareto-optimal configurations with the respective instantiation.}
  \label{fig:e2eallsolvers}
\end{figure}

% One of the main observations is that our model-uncertainty-based sampling method (blue points) consistently outperformed random selection (red points) for almost any value of $\delta$ within Eq.~\ref{eq:opt}.
% Moreover, our ranking-convergence criterion (dot-shaped points) is particularly suited to achieve a good ranking accuracy fast.
% In comparison, our Wilcoxon-based criterion is suited better for high-accuracy predictions.
% This comes at a high runtime cost though.
% A good trade-off is achieved for $\delta = 0.85$ with an average PAR-2 ranking performance of \SI{91.83}{\%} and a runtime requirement of on average \SI{23.05}{\%} of the time it would take to evaluate the complete dataset.

\begin{table}[tb!]
  \centering
  \caption{Pareto-optimal approaches for different choices of $\delta$ regarding our smaller hyper-parameter tuning dataset. Rows with no changes in comparison to their predecessor are hidden. Ranking approaches are either predicted cluster-label ranking (Pred.) or partially observed PAR-2 ranking (PAR-2). Selection approaches are either uncertainty-based selection (Uncert.) or information-gain-based selection (IG). Stopping criteria are either ranking convergence (Conv.) or Wilcoxon-based (Wilcoxon).}
  \label{tab:paretotable}
  \vspace{0.2cm}
  \begin{tabular}{
    >{\centering\arraybackslash}m{0.4cm}
    >{\centering\arraybackslash}m{1.6cm}
    >{\centering\arraybackslash}m{1.6cm}
    >{\centering\arraybackslash}m{1.6cm}
    >{\centering\arraybackslash}m{1.5cm}
    >{\centering\arraybackslash}m{1.5cm}
    >{\centering\arraybackslash}m{1.5cm}
    >{\centering\arraybackslash}m{1.5cm}
  }
    \hline
    $\delta$ & Ranking & Selection & Stopping & Fraction Runtime ($O_{\operatorname{rt}}$) & Fraction Instances & $O_{\operatorname{acc}}$ & $O_{\delta}$ \\
    \hline
    0.00 & PAR-2 & Uncert. & Conv. & 0.021 & 0.025 & 0.593 & 0.979 \\
    0.05 & Pred. & Uncert. & Conv. & 0.024 & 0.158 & 0.765 & 0.966 \\
    0.10 & Pred. & Uncert. & Conv. & 0.025 & 0.128 & 0.779 & 0.955 \\
    0.15 & Pred. & IG & Conv. & 0.028 & 0.112 & 0.802 & 0.946 \\
    $\vdots$ & $\vdots$ & $\vdots$ & $\vdots$ & $\vdots$ & $\vdots$ & $\vdots$ & $\vdots$ \\
    0.50 & Pred. & IG & Conv. & 0.028 & 0.112 & 0.802 & 0.887 \\
    0.55 & Pred. & Uncert. & Conv. & 0.053 & 0.142 & 0.824 & 0.880 \\
    0.60 & Pred. & Uncert. & Conv. & 0.146 & 0.082 & 0.897 & 0.880 \\
    $\vdots$ & $\vdots$ & $\vdots$ & $\vdots$ & $\vdots$ & $\vdots$ & $\vdots$ & $\vdots$ \\
    0.75 & Pred. & Uncert. & Conv. & 0.146 & 0.082 & 0.897 & 0.886 \\
    0.80 & Pred. & Uncert. & Conv. & 0.219 & 0.128 & 0.917 & 0.890 \\
    \hline
    0.85 & Pred. & Uncert. & Conv. & 0.272 & 0.171 & 0.927 & 0.897 \\
    \hline
    0.90 & Pred. & Uncert. & Wilcoxon & 0.894 & 0.888 & 0.996 & 0.907 \\
    0.95 & Pred. & Uncert. & Wilcoxon & 0.931 & 0.924 & 1.000 & 0.953 \\
    1.00 & Pred. & Uncert. & Wilcoxon & 0.989 & 0.981 & 1.000 & 1.000 \\
    \hline
  \end{tabular}
\end{table}

Tab.~\ref{tab:paretotable} also shows the performance of particular hyper-parameter choices for different values of $\delta$.
Thereby, we observe that for small $\delta$ (runtime more important) our approaches preferably use easy instances to get a good sense of the PAR-2 ranking fast, e.g., row with $\delta = 0.15$.
In contrast to that, harder instances are preferred for higher values of $\delta$, i.e., the sampled fraction of runtime is greater than the fraction of instances.
In our opinion, the row with $\delta = 0.85$ represents the best trade-off between runtime and PAR-2 accuracy.

\subsubsection{Full dataset evaluation.}
Having selected the most promising hyper-parameters, we run our end-to-end active-learning experiments on the complete Anniversary Track dataset (5355 instances).
The best-performing approach for $\delta = 0.85$ (since this was the most promising trade-off on the tuning set) uses runtime-label prediction ranking, model-based uncertainty sampling, and a ranking-convergence stopping criterion.
Thereby, the sampled instances account for about \SI{5.24}{\%} of instances and about \SI{10.35}{\%} of runtime.
Moreover, the PAR-2 accuracy is about \SI{92.33}{\%}, the average Spearman correlation between the PAR-2 ranking and the predicted ranking scores about \SI{0.9572}{}, and the weighted objective $O_{\delta} = 0.9193$.

\subsubsection{Instance-family importance.}
For all configurations of our approach that sample on average less than \SI{10}{\%} of all 5355 instances, we look at the particular instances and families that are selected by our active-learning approach.
Thereby, the average benchmark set contains about \SI{192.34}{} instances.
Tab.~\ref{tab:familiesa} lists the ten most frequent families that are chosen. For example, \textit{planning} instances account for on average \SI{19.14}{} of the \SI{192.34}{} benchmark instances.
In contrast to that, Tab.~\ref{tab:familiesb} shows the families that are sampled the most often relative to their class sizes in the underlying dataset.
There are, for example, only nine \textit{circuit-multiplier} instances in the complete dataset.
However, we sample on average \SI{1.59}{} of them.

\begin{table}[tb!]
  \centering
  \caption{Top-10 most \textit{important} families regarding the average occurrences and the fraction of selected occurrences regarding our AL approach.}
  \label{tab:families}
  \hspace{-0.58cm}
  \begin{subtable}{0.45\textwidth}
    \centering
    \caption{Sorted by the average number of occurrences.}
    \label{tab:familiesa}

    \begin{tabular}{
      >{\centering\arraybackslash}m{1.6cm}
      >{\centering\arraybackslash}m{1.3cm}
      >{\centering\arraybackslash}m{1.3cm}
      >{\centering\arraybackslash}m{1.3cm}
    }
      \hline
      Family & Avg. Occ. & In Dataset & Fraction (\%) \\
      \hline
      planning              & 19.14 &  333 &  5.7 \\
      subgraph-isomorphism  & 16.27 &  175 &  9.3 \\
      crypto-graphy         & 13.11 &  311 &  4.2 \\
      pigeon-hole           &  7.55 &  151 &  5.0 \\
      fpga-routing          &  6.16 &  73  &  8.4 \\
      rbsat                 &  5.93 &  113 &  5.2 \\
      quasigroup-completion &  5.23 &  210 &  2.5 \\
      unknown               &  4.93 &  169 &  2.9 \\
      miter                 &  4.84 &  199 &  2.4 \\
      bitvector             &  4.73 &  214 &  2.2 \\
      % clique-coloring       &  4.21 &  64  &  0.066 \\
      % mosoi-289             &  3.75 &  31  &  0.121 \\
      % hardware-verification &  3.70 &  344 &  0.011 \\
      % hardware-bmc          &  2.77 &  78  &  0.035 \\
      % graph-based           &  2.77 &  134 &  0.021 \\
      % coloring              &  2.66 &  41  &  0.065 \\
      % waerden               &  2.55 &  43  &  0.059 \\
      % graph-coloring        &  2.50 &  87  &  0.029 \\
      % auto-correlation      &  2.39 &  46  &  0.052 \\
      % knights-problem       &  2.25 &  17  &  0.132 \\
      %\hline
      Remaining families    & 104.45 & 3407 & 3.1 \\
      \hline
      Sum                   & 192.34 & 5355 & 3.6 \\
      \hline
    \end{tabular}
  \end{subtable}
  \hspace{0.55cm}
  \begin{subtable}{0.45\textwidth}
    \centering
    \caption{Sorted by the fraction of sampled occurrences.}
    \label{tab:familiesb}

    \begin{tabular}{
      >{\centering\arraybackslash}m{1.6cm}
      >{\centering\arraybackslash}m{1.3cm}
      >{\centering\arraybackslash}m{1.3cm}
      >{\centering\arraybackslash}m{1.3cm}
    }
      \hline
      Family & Avg. Occ. & In Dataset & Fraction (\%) \\
      \hline
      circuit-multiplier          & 1.59    &      9      &     17.7 \\
      subset-cardinality          & 1.55    &      9      &     17.3 \\
      ssp-0                       & 0.34    &      2      &     17.0 \\
      sliding-puzzle              & 1.52    &      11     &     13.8 \\
      knights-problem             & 2.25    &      17     &     13.2 \\
      genurq                      & 1.55    &      12     &     12.9 \\
      stone                       & 1.39    &      11     &     12.7 \\
      mosoi-289                   & 3.75    &      31     &     12.1 \\
      anti-bandwidth              & 1.52    &      13     &     11.7 \\
      random                      & 0.91    &      8      &     11.4 \\
      % spectrum-repacking          & 1.30    &      12     &     0.108631 \\
      % social-golfer               & 0.95    &      9      &     0.105159 \\
      % product-configuration       & 0.52    &      5      &     0.103571 \\
      % minimal-disagreement-parity & 1.54    &      16     &     0.095982 \\
      % subgraph-isomorphism        & 16.27   &       175   &     0.092959 \\
      % sgen-balanced               & 0.27    &      3      &     0.089286 \\
      % pythagorean-triples         & 1.82    &      21     &     0.086735 \\
      % fpga-routing                & 6.16    &      73     &     0.084393 \\
      % petrinet-concurrency        & 2.11    &      25     &     0.084286 \\
      % core-based-generator        & 1.00    &      13     &     0.076923 \\
      %\hline
      Remaining families    & 175.97 & 5232 & 3.4 \\
      \hline
      Sum                   & 192.34 & 5355 & 3.6 \\
      \hline
    \end{tabular}
  \end{subtable}
\end{table}

%\vspace{10cm}
\section{Conclusion}
In this work, we have discussed possible solutions to the \textit{new-solver} problem:
Given a new solver, we want to find its ranking amidst its competitors.
Our approaches provide accurate ranking predictions while only needing a fraction of the runtime resources that a full evaluation on all benchmarking instances would need.
We use a runtime discretization technique as this enables us to transform the regression problem of directly predicting runtimes into the much simpler notion of classification.
We have shown that, albeit being more simple, the classification of discrete runtime labels produces good results.
We have evaluated several ranking algorithms, instance-selection approaches, and stopping criteria within our sequential active-learning process.
A model-uncertainty-based selection approach in combination with runtime-label-prediction ranking and a ranking-convergence stopping criterion showed the consistently best results.
We also took a brief look at which instance families are the most prevalent when sampling instances.

\subsection{Future Work}
In future work, it may be of interest to compare further ranking algorithms, instance-selection approaches, and stopping criteria.
Furthermore, it is possible to formulate the runtime discretization as an optimization problem.
Given a dataset, optimization could select the discretization technique with the best discriminatory power rather than selecting one approach upfront by hand.

A major shortcoming is the lack of parallelizability.
Our current approach selects instances one at a time.
Running benchmarks on a cluster with $n$ cores requires a batch of $n$ instances at a time.
This is, however, not trivial since, on the one hand, a higher $n$ leads to less \textit{active learning} (because of bigger batch sizes) and, on the other hand, it is not clear how to synchronize the model update and instance selection without a barrier lock, which wastes a lot of runtime resources.

On a more general note, it would make sense to generalize this evaluation framework for arbitrary $\mathcal{NP}$-complete problems.
Those problems share most of the relevant properties, i.e., there are established problem-instance features, a full benchmark run takes days, and creating new problem solvers traditionally requires \textit{expert} knowledge to hand-select instances of interest.

%
% ---- Bibliography ----
%
\bibliographystyle{splncs04}
\bibliography{literature}

\end{document}

### 2022-09-26 run.sh

@@ -1,0 +1,3 @@
+set -x
+export PYTHONPATH="./src"
+python src/al_experiments.py

### 2022-09-26 run_debug.sh

@@ -1,0 +1,3 @@
+set -x
+export PYTHONPATH="./src"
+DEBUG=1 python src/al_experiments.py

### 2022-09-26 src/al_experiments.py

@@ -1,0 +1,454 @@
+import numpy as np
+import pandas as pd
+import logging
+import sys
+import time
+import random
+import os
+import pickle
+import warnings
+
+from typing import List, Tuple
+from joblib import Parallel, delayed
+from scipy.stats import spearmanr
+from sklearn.tree import DecisionTreeClassifier
+from sklearn.discriminant_analysis import QuadraticDiscriminantAnalysis
+from sklearn.ensemble import StackingClassifier, RandomForestClassifier
+
+from al_experiments.config import DEBUG
+from al_experiments.clustering import merge_closest_intervals, assign_labels_from_label_boundaries
+from al_experiments.evaluation import (
+    total_amount_runtime_target_solver,
+    mean_wilcoxon_pvalue_predicted_labels,
+)
+from al_experiments.final_experiments import all_experiments
+from al_experiments.experiment import Experiment
+from al_experiments.helper import push_notification
+
+
+logging.basicConfig(
+    format="%(asctime)s %(levelname)s: %(message)s",
+    datefmt="[%Y-%m-%d %H:%M:%S]",
+    level=logging.INFO,
+    handlers=[
+        logging.FileHandler("./run.log"),
+        logging.StreamHandler(sys.stdout)
+    ]
+)
+
+
+def run_e2e_experiment(
+    i: int, experiment: Experiment, target_solver: str
+) -> Tuple[Tuple[int, int, float, float, float, float], np.ndarray]:
+    """ Runs a single end-to-end experiment.
+
+    Args:
+        i (int): The number of the experiment.
+        experiment (Experiment): The experiment configuration.
+        target_solver (str): The target solver.
+
+    Returns:
+        Tuple[int, int, float, float, float, float]: The performance summary.
+    """
+
+    # Features
+    with open("./pickled-data/base_features_df.pkl", "rb") as file:
+        base_features_df: pd.DataFrame = pickle.load(file).copy()
+
+    if experiment.only_hashes:
+        # Load full runtimes of 16 solvers and filter to competition instances
+        with open("./pickled-data/runtimes_df.pkl", "rb") as file:
+            runtimes_df: pd.DataFrame = pickle.load(file).copy()
+        with open(experiment.instance_filter, "rb") as file:
+            instance_filter: pd.Series = pickle.load(file).copy()
+    else:
+        # Load dedicated competition dataset with more solvers
+        with open(f"./pickled-data/{experiment.instance_filter_prefix}_df.pkl", "rb") as file:
[…]

### 2022-09-26 src/al_experiments/clustering.py

@@ -1,0 +1,152 @@
+from typing import List, Tuple
+
+import numpy as np
+import pandas as pd
+
+
+def min_log_dist(i: List[float], j: List[float]) -> float:
+    """ Single-link logarithmic distance.
+
+    Args:
+        i (float): The first interval.
+        j (float): The second interval.
+
+    Returns:
+        float: The single-link logarithmic distance.
+    """
+
+    fst = np.log1p(i)
+    snd = np.log1p(j)
+    return np.min(snd) - np.max(fst)
+
+
+def merge_closest_intervals(
+    train_solvers: List[str], runtimes_df: pd.DataFrame, n_intervals=2
+) -> Tuple[pd.DataFrame, pd.DataFrame]:
+    """ Hierarchical merging of the runtimes into clusters.
+
+    Args:
+        train_solvers (List[str]): The list of known solvers.
+        runtimes_df (pd.DataFrame): The runtimes to cluster.
+        n_intervals (int, optional): The number of labels to
+        cluster the runtime into. Defaults to 2.
+
+    Returns:
+        Tuple[pd.DataFrame, pd.DataFrame]: A tuple of cluster labels
+        and the boundaries between the labels.
+    """
+
+    merge_runtimes_df = runtimes_df[train_solvers].replace(
+        [-np.inf, np.inf, np.nan], 0).astype(dtype=np.int8).copy()
+    boundaries = {}
+
+    for idx, (pd_index, instance) in enumerate(list(runtimes_df[train_solvers].iterrows())):
+        if np.all(np.isinf(instance)):
+            merge_runtimes_df.iloc[idx, :] = n_intervals
+
+            # All solvers have time-out
+            decision_boundaries = [
+                np.nan,
+                np.nan,
+            ]
+        else:
+            actual_n_intervals = min(n_intervals, np.unique(instance).shape[0])
+            unique_vals = list(np.sort(np.unique(instance)))
+            if unique_vals[-1] == np.inf:
+                del unique_vals[-1]
+            intervals = [[i] for i in unique_vals]
+
+            # Function to find the smallest distance between neighboring intervals
+            def find_smallest_dist_neighbors(intervals):
+                min_val = -1
+                min_idx = -1
+                idx = 0
+                for i, j in zip(intervals[:-1], intervals[1:]):
+                    val = min_log_dist(i, j)
+                    if min_val == -1 or val < min_val:
+                        min_val = val
+                        min_idx = idx
+                    idx += 1
+                return min_idx, min_val
+
[…]

### 2022-09-26 src/al_experiments/config.py

@@ -1,0 +1,4 @@
+import os
+
+
+DEBUG = bool(int(os.getenv("DEBUG", "0")))

### 2022-09-26 src/al_experiments/evaluation.py

@@ -1,0 +1,101 @@
+import warnings
+import numpy as np
+import pandas as pd
+
+from scipy.stats import wilcoxon
+from sklearn.metrics import mean_squared_error
+
+from al_experiments.experiment import Experiment
+from al_experiments.stopping import WilcoxonSignificance
+
+
+def total_amount_runtime_target_solver(
+    y_sampled: np.ndarray, runtimes_df: pd.DataFrame,
+    cluster_boundaries: pd.DataFrame, target_solver: str,
+    experiment: Experiment
+) -> float:
+    """ Returns the total amount of runtime that has been used so far within the selected experiments.
+
+    Args:
+        y_sampled (np.ndarray): Which instances are sampled.
+        runtimes_df (pd.DataFrame): The runtimes.
+        cluster_boundaries (pd.DataFrame): The cluster boundaries.
+        target_solver (str): The target solver.
+        experiment (Experiment): The experiment configuration.
+
+    Returns:
+        float: The amount of runtime sampled.
+    """
+
+    return (
+        np.sum(runtimes_df.loc[
+            y_sampled == 1,
+            target_solver
+        ].replace(np.inf, 5000))
+        / np.sum(runtimes_df.loc[:, target_solver].replace(np.inf, 5000))
+    )
+
+
+def mean_wilcoxon_pvalue_predicted_labels(
+    clustering: pd.DataFrame, pred: np.ndarray,
+    y: np.ndarray, y_sampled: np.ndarray,
+    target_solver: str, experiment: Experiment,
+) -> float:
+    """ Returns the mean Wilcoxon p-value of the predicted
+    labels of the target solvers with each other known solver.
+
+    Args:
+        clustering (pd.DataFrame): The cluster labels.
+        pred (np.ndarray): The predictions.
+        y (np.ndarray): The ground-truth labels.
+        y_sampled (np.ndarray): Which instances are sampled.
+
+    Returns:
+        float: The mean Wilcoxon p-value.
+    """
+
+    if isinstance(experiment.stopping, WilcoxonSignificance):
+        t = experiment.stopping.observed_thresh
+    else:
+        t = 0.2
+
+    if np.count_nonzero(y_sampled) / y_sampled.shape[0] <= t:
+        other_solver_labels = clustering.iloc[
+            y_sampled == 1,
+            clustering.columns != target_solver
+        ]
+        target_labels = y[y_sampled == 1]
+    else:
+        other_solver_labels = clustering.iloc[
+            :, clustering.columns != target_solver
+        ]
[…]

### 2022-09-26 src/al_experiments/experiment.py

@@ -1,0 +1,504 @@
+from al_experiments.selection import (
+    SelectionFunctor,
+    RandomSampling,
+    ModelUncertaintySampling,
+    ModelBasedInformationGainSampling,
+    VotingEntropyBasedCommitteeDisagreementSampling,
+    EntropyBasedUncertaintySampling,
+    FastRuntimeSampling,
+    VarianceRuntimePerRuntimeSampling,
+)
+from al_experiments.stopping import StoppingPredicate, FixedSubsetSize
+from al_experiments.ranking import (
+    RankingFunctor,
+    PartialObservationBasedRanking,
+    PredictedClusterLabelRanking,
+    PartialObservedLabelRanking,
+    PartialObservedLabelAndPredictionRanking,
+)
+
+
+class Experiment:
+    """ Configuration class for a single experiment. """
+
+    def __init__(
+        self, key: str, description: str,
+        selection: SelectionFunctor, stopping: StoppingPredicate, ranking: RankingFunctor,
+        repetitions: int = 1, par: float = 2,
+        instance_filter: str = "sat_202x_submission_hashes", only_hashes: bool = True,
+    ):
+        """ Initializes an experiment configuration.
+
+        Args:
+            key (str): The key of the experiment.
+            description (str): The description.
+            selection (SelectionFunctor): The selection functor.
+            stopping (StoppingPredicate): The stopping predicate.
+            ranking (RankingFunctor): The ranking functor.
+            repetitions (int, optional): The number of repetitions. Defaults to 1.
+            par (float, optional): The PAR value. Defaults to 2.
+            instance_filter (str, optional): Which instances to use. Defaults to "sat_202x_submission_hashes".
+        """
+
+        self.key = key
+        self.description = description
+        self.selection = selection
+        self.stopping = stopping
+        self.ranking = ranking
+        self.results_location = f"./pickled-data/end_to_end/{key}.csv"
+        self.repetitions = repetitions
+        self.par = par
+        self.instance_filter_prefix = instance_filter
+        self.instance_filter = f"./pickled-data/instance_filters/{instance_filter}.pkl"
+        self.label_ranking_timeout_weight = 2.0
+        self.only_hashes = only_hashes
+
+    @classmethod
+    def simple_baseline(
+        cls,
+        instance_filter: str,
+        repetitions: int,
+    ) -> "Experiment":
+        """ Simple baseline configuration.
+
+        Args:
+            instance_filter (str): Which instances to use.
[…]

### 2022-09-26 src/al_experiments/final_experiments.py

@@ -1,0 +1,171 @@
+from typing import List
+
+from al_experiments.experiment import Experiment
+from al_experiments.selection import (
+    RandomSampling,
+    SelectionFunctor,
+    ModelUncertaintySampling,
+    ModelBasedInformationGainSampling,
+)
+from al_experiments.ranking import (
+    RankingFunctor,
+    PartialObservationBasedRanking,
+    PredictedClusterLabelRanking,
+)
+from al_experiments.stopping import (
+    StoppingPredicate,
+    FixedSubsetSize,
+    RankingConverged,
+    WilcoxonSignificance,
+)
+
+
+# # Selection strategies (9)
+# selection_strategies: List[SelectionFunctor] = []
+# selection_strategies.append(RandomSampling())
+# for f in [0.0, 0.05, 0.1, 0.15, 0.2]:
+#     selection_strategies.append(ModelUncertaintySampling(
+#         f, True
+#     ))
+#     selection_strategies.append(ModelUncertaintySampling(
+#         f, False
+#     ))
+#     selection_strategies.append(ModelBasedInformationGainSampling(
+#         f, True,
+#     ))
+#     selection_strategies.append(ModelBasedInformationGainSampling(
+#         f, False,
+#     ))
+
+# # Ranking strategies (15)
+# ranking_strategies: List[RankingFunctor] = []
+# ranking_strategies.append(PartialObservationBasedRanking())
+# for fallback_thresh in [0.01, 0.05, 0.1]:
+#     # ranking_strategies.append(PartialObservedLabelRanking(fallback_thresh))
+#     for num_hist in [1, 10, 20, 30, 40]:
+#         ranking_strategies.append(PredictedClusterLabelRanking(
+#             True, 0.0, num_hist, fallback_thresh
+#         ))
+#         # for obs in [0.0, 0.1, 0.2]:
+#         #     # ranking_strategies.append(PredictedClusterLabelRanking(
+#         #     #     False, obs, 1, fallback_thresh
+#         #     # ))
+#         #     # ranking_strategies.append(PartialObservedLabelAndPredictionRanking(
+#         #     #     False, obs, 1, fallback_thresh
+#         #     # ))
+#         #     ranking_strategies.append(PartialObservedLabelAndPredictionRanking(
+#         #         True, obs, num_hist, fallback_thresh
[…]

### 2022-09-26 src/al_experiments/helper.py

@@ -1,0 +1,29 @@
+import logging
+import requests
+
+from al_experiments.config import DEBUG
+
+
+def push_notification(text: str = "Execution finished.", is_error: bool = False) -> None:
+    """ Helper method for push notifications.
+
+    Args:
+        text (str, optional): The text of the message. Defaults to "Execution finished.".
+        is_error (bool, optional): Whether it is an error. Defaults to False.
+    """
+
+    if not is_error:
+        logging.info(text)
+    else:
+        logging.error(text)
+
+    if not DEBUG:
+        with open("./secret", "r") as file:
+            api_secret = str(file.read()).strip()
+        headers = {"x-api-key": api_secret}
+        message = {"title": "Jupyter Notebook", "body": text}
+        requests.post(
+            "https://push.techulus.com/api/v1/notify",
+            data=message, headers=headers
+        )
+        del api_secret

### 2022-09-26 src/al_experiments/ranking.py

@@ -1,0 +1,340 @@
+from abc import ABC, abstractmethod
+from typing import List
+
+import numpy as np
+import pandas as pd
+
+from sklearn.ensemble import StackingClassifier
+
+
+class RankingFunctor(ABC):
+    """ Base class for ranking decision making. """
+
+    def __init__(self, needs_predictions) -> None:
+        self.needs_predictions = needs_predictions
+
+    @abstractmethod
+    def get_params(self) -> str:
+        return ""
+
+    @abstractmethod
+    def __call__(
+        self, y_sampled: np.ndarray, runtimes_df: pd.DataFrame,
+        clustering: pd.DataFrame, cluster_boundaries: pd.DataFrame,
+        timeout_clf: StackingClassifier, label_clf: StackingClassifier,
+        x: np.ndarray, y: np.ndarray, target_solver: str, pred: np.ndarray,
+        experiment, pred_history: List[np.ndarray]
+    ) -> np.ndarray:
+        """ Calculates how the target solver compares against known solvers.
+
+        Args:
+            y_sampled (np.ndarray): The already sampled instances.
+            runtimes_df (pd.DataFrame): The runtimes.
+            clustering (pd.DataFrame): The labels of known solvers.
+            cluster_boundaries (pd.DataFrame): The cluster boundaries.
+            timeout_clf (StackingClassifier): The time-out predictor.
+            label_clf (StackingClassifier): The non-time-out label predictor.
+            x (np.ndarray): The input features.
+            y (np.ndarray): The target labels.
+            target_solver (str): The target solver.
+            pred (np.ndarray): The predictions.
+            experiment (Experiment): The experiment configuration.
+
+        Returns:
+            np.ndarray: Binary vector indicating whether the
+            target solver is better than the respective solver
+        """
+
+        pass
+
+
+class PartialObservationBasedRanking(RankingFunctor):
+    """ Ranking based on partially observed runtimes. """
+
+    def __init__(self) -> None:
+        super().__init__(False)
+
+    def get_params(self) -> str:
+        return "par2_obs"
+
+    def __call__(
+        self, y_sampled: np.ndarray, runtimes_df: pd.DataFrame,
+        clustering: pd.DataFrame, cluster_boundaries: pd.DataFrame,
+        timeout_clf: StackingClassifier, label_clf: StackingClassifier,
+        x: np.ndarray, y: np.ndarray, target_solver: str, pred: np.ndarray,
+        experiment, pred_history: List[np.ndarray]
+    ) -> np.ndarray:
[…]

### 2022-09-26 src/al_experiments/selection.py

@@ -1,0 +1,551 @@
+from abc import ABC, abstractmethod
+
+import numpy as np
+import pandas as pd
+
+from scipy.stats import entropy
+from sklearn.ensemble import StackingClassifier
+
+
+class SelectionFunctor(ABC):
+    """ Base class for instance selection making. """
+
+    def __init__(self, model_based_selection_threshold: float) -> None:
+        self.model_based_selection_threshold = model_based_selection_threshold
+
+    @abstractmethod
+    def get_params(self) -> str:
+        return ""
+
+    @abstractmethod
+    def __call__(
+        self, y_sampled: np.ndarray, runtimes_df: pd.DataFrame,
+        clustering: pd.DataFrame, cluster_boundaries: pd.DataFrame,
+        timeout_clf: StackingClassifier, label_clf: StackingClassifier,
+        x: np.ndarray, y: np.ndarray, target_solver: str, experiment
+    ) -> None:
+        """ Selects another instance in 'y_sampled'.
+
+        Args:
+            y_sampled (np.ndarray): The already sampled instances.
+            runtimes_df (pd.DataFrame): The runtimes.
+            clustering (pd.DataFrame): The labels of known solvers.
+            cluster_boundaries (pd.DataFrame): The cluster boundaries.
+            timeout_clf (StackingClassifier): The time-out predictor.
+            label_clf (StackingClassifier): The non-time-out label predictor.
+            x (np.ndarray): The input features.
+            y (np.ndarray): The target labels.
+            target_solver (str): The target solver.
+            experiment (Experiment): The experiment configuration.
+        """
+
+        pass
+
+
+class RandomSampling(SelectionFunctor):
+    """ Random sampling. """
+
+    def __init__(self, batch: int = 1) -> None:
+        super().__init__(0.0)
+        self.batch = batch
+
+    def get_params(self) -> str:
+        return f"random_{self.batch}"
+
+    def __call__(
+        self, y_sampled: np.ndarray, runtimes_df: pd.DataFrame,
+        clustering: pd.DataFrame, cluster_boundaries: pd.DataFrame,
+        timeout_clf: StackingClassifier, label_clf: StackingClassifier,
+        x: np.ndarray, y: np.ndarray, target_solver: str, experiment
+    ) -> None:
+        """ Selects another instance in 'y_sampled'.
+
+        Args:
+            y_sampled (np.ndarray): The already sampled instances.
+            runtimes_df (pd.DataFrame): The runtimes.
+            clustering (pd.DataFrame): The labels of known solvers.
[…]

### 2022-09-26 src/al_experiments/stopping.py

@@ -1,0 +1,183 @@
+from abc import ABC, abstractmethod
+from typing import List
+
+import numpy as np
+import pandas as pd
+
+from sklearn.ensemble import StackingClassifier
+
+
+class StoppingPredicate(ABC):
+    """ Base class representing the stopping criterion. """
+
+    @abstractmethod
+    def get_params(self) -> str:
+        return ""
+
+    @abstractmethod
+    def __call__(
+        self, y_sampled: np.ndarray, runtimes_df: pd.DataFrame,
+        clustering: pd.DataFrame, cluster_boundaries: pd.DataFrame,
+        timeout_clf: StackingClassifier, label_clf: StackingClassifier,
+        x: np.ndarray, y: np.ndarray, target_solver: str, experiment,
+        ranking_history: List[np.ndarray], wilcoxon_history: List[float],
+    ) -> bool:
+        """ Determines when to stop iterating.
+
+        Args:
+            y_sampled (np.ndarray): The already sampled instances.
+            runtimes_df (pd.DataFrame): The runtimes.
+            clustering (pd.DataFrame): The labels of known solvers.
+            cluster_boundaries (pd.DataFrame): The cluster boundaries.
+            timeout_clf (StackingClassifier): The time-out predictor.
+            label_clf (StackingClassifier): The non-time-out label predictor.
+            x (np.ndarray): The input features.
+            y (np.ndarray): The target labels.
+            target_solver (str): The target solver.
+            experiment (Experiment): The experiment configuration.
+
+        Returns:
+            bool: Whether to stop adding more instances.
+        """
+
+        pass
+
+
+class FixedSubsetSize(StoppingPredicate):
+    """ Stop after a fixed amount of instances is in the subset. """
+
+    def __init__(self, subset_size: float) -> None:
+        """ Initializes this stopping predicate.
+
+        Args:
+            subset_size (float): The desired subset size.
+        """
+
+        self.subset_size = subset_size
+
+    def get_params(self) -> str:
+        return f"fixed_{self.subset_size:.2f}"
+
+    def __call__(
+        self, y_sampled: np.ndarray, runtimes_df: pd.DataFrame,
+        clustering: pd.DataFrame, cluster_boundaries: pd.DataFrame,
+        timeout_clf: StackingClassifier, label_clf: StackingClassifier,
+        x: np.ndarray, y: np.ndarray, target_solver: str, experiment,
+        ranking_history: List[np.ndarray], wilcoxon_history: List[float],
+    ) -> bool:
[…]

### 2022-09-26 paper/paper.tex
Adapt title and authors
- remove (first) hyphen in title
- add further authors and re-format author block

@@
 \usepackage{array}
+\usepackage[misc]{ifsym} % letter symbol to mark corresponding author
 
@@
 
-\title{Active-Learning for SAT-Solver Benchmarking}
-\author{Tobias Fuchs}
-\institute{Institute of Theoretical Informatics, \\ Karlsruhe Institute of Technology, Germany \\
-\email{info@tobiasfuchs.de}}
+\title{Active Learning for SAT-Solver Benchmarking}
+
+\author{
+	Tobias Fuchs (\Letter) \and
+	Jakob Bach\orcidID{0000-0003-0301-2798} \and
+	Markus Iser\orcidID{0000-0003-2904-232X}
+}
+
+\institute{
+	Karlsruhe Institute of Technology (KIT), Karlsruhe, Germany \\
+	\email{info@tobiasfuchs.de, jakob.bach@kit.edu, markus.iser@kit.edu}
+}
+
+\authorrunning{T. Fuchs et al.} % according to sample paper of template: first names are abbreviated; if there are more than two authors, 'et al.' is used.
 
@@
 
-\section{Active-Learning for SAT-Solver Benchmarking}
+\section{Active Learning for SAT-Solver Benchmarking}
 \label{sec:main}

### 2022-09-26 paper/paper.tex
Add license and orcid
Conduct minor adaptations to template
... based on the LNCS template's instructions and sample file:

- add font encoding
- avoid re-defining basic layout settings
- color links in blue
- add keywords
- use dedicated environment for definition (though we only have one)
- don't end captions with period if not full sentences

@@
 
+\usepackage[T1]{fontenc} % recommended by LNCS template
 \usepackage{graphicx}
@@
 
-\renewcommand{\arraystretch}{1.5}
-%\newcolumntype{L}[1]{>{\raggedright\arraybackslash}p{#1}}
-\setlength{\textfloatsep}{12pt}
-
 \usepackage[ruled,vlined,linesnumbered,norelsize]{algorithm2e}
@@
 \SetCommentSty{emph}
-\def\listalgorithmcfname{List of Algorithms}
-\def\algorithmautorefname{Algorithm}
-\let\chapter=\section % repariert ein Problem mit algorithm2e
+
+\renewcommand\UrlFont{\color{blue}\rmfamily} % link coloring according to LNCS template
 
@@
 \author{
-	Tobias Fuchs (\Letter) \and
-	Jakob Bach\orcidID{0000-0003-0301-2798} \and
+	Tobias Fuchs\orcidID{0000-0001-9727-2878} (\Letter) \and \\
+	Jakob Bach\orcidID{0000-0003-0301-2798} \and \\
 	Markus Iser\orcidID{0000-0003-2904-232X}
@@
   Our tool provides the engineer of propositional-satisfiability solvers with a faster way of determining a new SAT-solver's performance in comparison to sampling the benchmark set by hand.
+
+  \keywords{Propositional satisfiability \and Benchmarking \and Active learning}
 \end{abstract}
@@
 
-\textbf{New solver problem.} Given a pool of instances $\mathcal{I}$ and solvers $\mathcal{A}$ with already known runtimes $r\!: \mathcal{A} \times \mathcal{I} \rightarrow \left[0, \tau\right] \cup \left\lbrace \perp \right\rbrace$ (timeout $\tau$ usually \SI{5000}{s}~\cite{FroleyksHIJS21}), the problem is to decide with as little resources (CPU time) as possible how a new solver $\hat{a} \notin \mathcal{A}$ performs in comparison to the existing ones by incrementally selecting a benchmark set $\tilde{\mathcal{I}}$ on which to evaluate $\hat{a}$.
+\begin{definition}[New solver problem]
+	Given a pool of instances $\mathcal{I}$ and solvers $\mathcal{A}$ with already known runtimes $r\!: \mathcal{A} \times \mathcal{I} \rightarrow \left[0, \tau\right] \cup \left\lbrace \perp \right\rbrace$ (timeout $\tau$ usually \SI{5000}{s}~\cite{FroleyksHIJS21}), the problem is to decide with as little resources (CPU time) as possible how a new solver $\hat{a} \notin \mathcal{A}$ performs in comparison to the existing ones by incrementally selecting a benchmark set $\tilde{\mathcal{I}}$ on which to evaluate $\hat{a}$.
+\end{definition}
+
 We want to find a scoring function $\operatorname{score}\!: \mathcal{A} \cup \left\lbrace \hat{a} \right\rbrace \rightarrow \mathbb{R}$ assigning each solver a score.
@@
   \begin{tabular}{
-    m{0.2\textwidth}
+    m{0.3\textwidth}
     >{\centering\arraybackslash}m{0.15\textwidth}
@@
     Ranking & \cmark & \xmark & \cmark & \cmark \\
-    Runtime \mbox{optimization} & \xmark & \cmark & \cmark & \cmark \\
+    Runtime optimization & \xmark & \cmark & \cmark & \cmark \\
     %Feature-based & \cmark & \cmark & \cmark & \cmark \\
-    Incremental \mbox{non-random} \mbox{sampling} & \xmark & \xmark & \cmark & \cmark \\
+    Incremental non-random sampling & \xmark & \xmark & \cmark & \cmark \\
     % \tablefootnote{While using a model-based selection strategy to select the next configuration to evaluate, the instance sampling is done randomly.}
@@
   }
-  \caption{Passive.}
+  \caption{Passive}
   \label{fig:passive}
@@
   }
-  \caption{Active Pool-based.}
+  \caption{Active Pool-based}
   \label{fig:activepool}
@@
   }
-  \caption{Active Synthesis-based.}
+  \caption{Active Synthesis-based}
   \label{fig:activesynth}
@@
 
-  \caption{Types of Learning (depiction inspired by Rubens~et.al.~\cite{RubensESK15}).}
+  \caption{Types of Learning (depiction inspired by Rubens~et.al.~\cite{RubensESK15})}
+  %JB: slightly larger font size in figure (at least having the size of the caption) would be nice
   \label{fig:learning}
@@
 \begin{algorithm}
-  \caption{Simple incremental benchmark-set-selection schema.}
+  \caption{Simple incremental benchmark-set-selection schema}
   \label{algALBenchmark}
@@
   \centering
-  \caption{PAR-2 ranking vs. ranking based on discretized runtimes for the SAT Competition~2022 Anniversary Track instances~\cite{sat2022} ($k = 3$; hierarchical clustering).}
+  \caption{PAR-2 ranking vs. ranking based on discretized runtimes for the SAT Competition~2022 Anniversary Track instances~\cite{sat2022} ($k = 3$; hierarchical clustering)}
   \label{tab:discreteranking}
@@
 \begin{algorithm}[htbp]
-  \caption{Evaluation framework.}
+  \caption{Evaluation framework}
   \label{alg:eval}
@@
 Each line entails the front of Pareto-optimal configurations with the respective instantiation.}
+  %JB: not sure if the color palette in the plots is grayscale-/color-blind-friendly; consider changing the palette (see https://colorbrewer2.org/#type=qualitative&scheme=Set3&n=3 ) and using different line types
   \label{fig:e2eallsolvers}
@@
   \centering
-  \caption{Top-10 most \textit{important} families regarding the average occurrences and the fraction of selected occurrences regarding our AL approach.}
+  \caption{Top-10 most \textit{important} families regarding the average occurrences and the fraction of selected occurrences regarding our AL approach}
   \label{tab:families}
@@
     \centering
-    \caption{Sorted by the average number of occurrences.}
+    \caption{Sorted by the average number of occurrences}
     \label{tab:familiesa}
@@
     \centering
-    \caption{Sorted by the fraction of sampled occurrences.}
+    \caption{Sorted by the fraction of sampled occurrences}
     \label{tab:familiesb}

### 2022-09-26 src/al_experiments.py
Make experiments reproducible

@@ -52,18 +52,18 @@ def run_e2e_experiment(
     """
 
     # Features
-    with open("./pickled-data/base_features_df.pkl", "rb") as file:
+    with open("../al-for-sat-solver-benchmarking-data/pickled-data/base_features_df.pkl", "rb") as file:
         base_features_df: pd.DataFrame = pickle.load(file).copy()
 
     if experiment.only_hashes:
         # Load full runtimes of 16 solvers and filter to competition instances
-        with open("./pickled-data/runtimes_df.pkl", "rb") as file:
+        with open("../al-for-sat-solver-benchmarking-data/pickled-data/runtimes_df.pkl", "rb") as file:
             runtimes_df: pd.DataFrame = pickle.load(file).copy()
         with open(experiment.instance_filter, "rb") as file:
             instance_filter: pd.Series = pickle.load(file).copy()
     else:
         # Load dedicated competition dataset with more solvers
-        with open(f"./pickled-data/{experiment.instance_filter_prefix}_df.pkl", "rb") as file:
+        with open(f"../al-for-sat-solver-benchmarking-data/pickled-data/{experiment.instance_filter_prefix}_df.pkl", "rb") as file:
             runtimes_df = pickle.load(file).copy()
         instance_filter = runtimes_df.index.copy()
 
@@ -396,22 +396,23 @@ def run_e2e_experiment(
 if __name__ == "__main__":
     # Notify experiment start
     push_notification("Starting experiments.")
+    rand_int = random.randint(1, 100)
 
     for i_exp, experiment in enumerate(all_experiments):
         # Skip if present
         time.sleep(random.random() * 5)
-        if os.path.exists(experiment.results_location):
+        if os.path.exists(f"{experiment.results_location}_{rand_int:03}.csv"):
             continue
         else:
-            with open(experiment.results_location, "w") as lock_file:
+            with open(f"{experiment.results_location}_{rand_int:03}.csv", "w") as lock_file:
                 lock_file.write("lock")
 
         # Retrieve column list
         if experiment.only_hashes:
-            with open("./pickled-data/runtimes_df.pkl", "rb") as file:
+            with open("../al-for-sat-solver-benchmarking-data/pickled-data/runtimes_df.pkl", "rb") as file:
                 runtimes_df = pickle.load(file).copy()
         else:
[…]

### 2022-09-26 src/al_experiments/experiment.py

@@ -45,11 +45,11 @@ class Experiment:
         self.selection = selection
         self.stopping = stopping
         self.ranking = ranking
-        self.results_location = f"./pickled-data/end_to_end/{key}.csv"
+        self.results_location = f"../al-for-sat-solver-benchmarking-data/pickled-data/end_to_end/{key}"
         self.repetitions = repetitions
         self.par = par
         self.instance_filter_prefix = instance_filter
-        self.instance_filter = f"./pickled-data/instance_filters/{instance_filter}.pkl"
+        self.instance_filter = f"../al-for-sat-solver-benchmarking-data/pickled-data/instance_filters/{instance_filter}.pkl"
         self.label_ranking_timeout_weight = 2.0
         self.only_hashes = only_hashes
 

### 2022-09-26 src/al_experiments/helper.py

@@ -1,5 +1,4 @@
 import logging
-import requests
 
 from al_experiments.config import DEBUG
 
@@ -17,13 +16,13 @@ def push_notification(text: str = "Execution finished.", is_error: bool = False)
     else:
         logging.error(text)
 
-    if not DEBUG:
-        with open("./secret", "r") as file:
-            api_secret = str(file.read()).strip()
-        headers = {"x-api-key": api_secret}
-        message = {"title": "Jupyter Notebook", "body": text}
-        requests.post(
-            "https://push.techulus.com/api/v1/notify",
-            data=message, headers=headers
-        )
-        del api_secret
+    # if not DEBUG:
+    #     with open("./secret", "r") as file:
+    #         api_secret = str(file.read()).strip()
+    #     headers = {"x-api-key": api_secret}
+    #     message = {"title": "Jupyter Notebook", "body": text}
+    #     requests.post(
+    #         "https://push.techulus.com/api/v1/notify",
+    #         data=message, headers=headers
+    #     )
+    #     del api_secret

### 2022-09-26 paper/paper.tex
Update github links

@@
 
-For reproducibility, our source code, all experiments, and data are available on GitHub\footnote{\url{https://github.com/mathefuchs/al-sat-benchmarking-experiments}}.
+For reproducibility, our source code, all experiments, and data are available on GitHub\footnote{\url{https://github.com/mathefuchs/al-for-sat-solver-benchmarking}}.
 
@@
 \subsection{Implementation Details}
-For reproducibility, our source code, all experiments, and data are available on GitHub\footnote{\url{https://github.com/mathefuchs/al-sat-benchmarking-experiments}}.
+For reproducibility, our source code, all experiments, and data are available on GitHub\footnote{\url{https://github.com/mathefuchs/al-for-sat-solver-benchmarking}}.
 It is implemented in \textsc{Python} using \textit{scikit-learn}~\cite{scikit-learn} for making predictions and \textit{gbd-tools}~\cite{IserS18} for SAT-instance retrieval.

### 2022-09-26 paper/paper.tex
Sort LaTeX imports and remove unused ones
Just for fun.
Fix overfull boxes in plots/tables
... but not the overfull boxes in main text, as the latter is very likely
to change.

@@
 
-\usepackage[T1]{fontenc} % recommended by LNCS template
+\usepackage[ruled,vlined,linesnumbered,norelsize]{algorithm2e}
+\usepackage{amsmath,amssymb}
+\usepackage[T1]{fontenc} % font encoding; recommended by LNCS template
 \usepackage{graphicx}
-\usepackage{amsmath,amssymb}
-\usepackage[labelfont=bf]{caption}
+\usepackage{hyperref}
+\usepackage[misc]{ifsym} % letter symbol to mark corresponding author
+\usepackage{pifont}
+\usepackage{siunitx}
 \usepackage{subcaption}
-\captionsetup{compatibility=false}
+\usepackage{tablefootnote}
+\usepackage{tabularx}
 \usepackage[table]{xcolor}
-\usepackage{tikz}
-\usepackage{siunitx}
-\usepackage{hyperref}
-\usepackage{tabularx}
-\usepackage{pifont}
-\usepackage{tablefootnote}
-\usepackage{array}
-\usepackage[misc]{ifsym} % letter symbol to mark corresponding author
-
-\newcommand{\cmark}{\ding{51}}
+\usepackage{tikz} % has to be loaded after "xcolor" here, else "option clash"
+
+\newcommand{\cmark}{\ding{51}} % symbols for table comparing benchmarking approaches
 \newcommand{\xmark}{\ding{55}}
 
-\usepackage[ruled,vlined,linesnumbered,norelsize]{algorithm2e}
-\DontPrintSemicolon
+\DontPrintSemicolon % configuration for algorithm2e
 \def\NlSty#1{\textnormal{\fontsize{8}{10}\selectfont{}#1}}
@@
   \begin{tabular}[c]{ccc}
-  \begin{subfigure}[b]{0.26\textwidth}
-  \centering
-  \resizebox{!}{2.85cm}{
+  \begin{subfigure}[b]{0.25\textwidth}
+  \centering
+  \resizebox{!}{2.75cm}{
   \begin{tikzpicture}
@@
   &
-  \begin{subfigure}[b]{0.36\textwidth}
-  \centering
-  \resizebox{!}{2.85cm}{
+  \begin{subfigure}[b]{0.35\textwidth}
+  \centering
+  \resizebox{!}{2.75cm}{
   \begin{tikzpicture}
@@
   \label{tab:discreteranking}
-  \vspace{0.2cm}
   \begin{tabular}{
     c
-    >{\centering\arraybackslash}p{2.2cm}p{3.6cm}
-    >{\centering\arraybackslash}p{1.8cm}p{3.3cm}
+    >{\centering\arraybackslash}p{1.8cm}p{3.6cm}
+    >{\centering\arraybackslash}p{1.8cm}p{3.6cm}
   }
@@
   \begin{tabular}{
-    >{\centering\arraybackslash}m{0.4cm}
+    >{\centering\arraybackslash}m{0.6cm}
     >{\centering\arraybackslash}m{1.6cm}
@@
   \label{tab:families}
-  \hspace{-0.58cm}
-  \begin{subtable}{0.45\textwidth}
+  \begin{subtable}[t]{0.48\textwidth}
     \centering
@@
     \begin{tabular}{
-      >{\centering\arraybackslash}m{1.6cm}
-      >{\centering\arraybackslash}m{1.3cm}
-      >{\centering\arraybackslash}m{1.3cm}
-      >{\centering\arraybackslash}m{1.3cm}
+      >{\centering\arraybackslash}m{1.8cm}
+      >{\centering\arraybackslash}m{1.2cm}
+      >{\centering\arraybackslash}m{1.2cm}
+      >{\centering\arraybackslash}m{1.2cm}
     }
@@
   \end{subtable}
-  \hspace{0.55cm}
-  \begin{subtable}{0.45\textwidth}
+  \hfill
+  \begin{subtable}[t]{0.48\textwidth}
     \centering

### 2022-09-26 paper/paper.tex
Add comments from PdF paper, part 1 (Sections 1-3)
... and slightly adapt text in some places (comments I directly resolved or
other stuff I noticed).

@@
 \begin{abstract}
-  Experimentation is one of the main phases in algorithm engineering, including the archetypal $\mathcal{NP}$-complete propositional-sat\-is\-fia\-bi\-li\-ty problem (SAT).
-  While the field of SAT-solver benchmarking is well established, traditional benchmark-set-selection approaches do not optimize for benchmark runtime.
+  Experimentation is one of the main phases in algorithm engineering, including the propositional-sat\-is\-fia\-bi\-li\-ty problem (SAT).
+  While the field of SAT-solver benchmarking is well established, traditional bench\-mark-set-selection approaches do not optimize for benchmark runtime.
   Benchmark-set selection is the process of choosing a subset of instances out of a pool to be representative and reliable in discriminating SAT solvers.
-  This paper presents an active-learning-based benchmark-set-selection approach that solves the new-solver problem:
+  This paper presents an active-learning-based benchmark-set-selec\-tion approach that solves the new-solver problem:
   Given a new solver, we want to find its ranking amidst its competitors with as little runtime as possible while still giving accurate ranking predictions.
-  To do so, we use well-known techniques from the domain of active learning.
   We can determine a new solver's rank regarding the SAT Competition 2022 Anniversary Track instances in about \SI{10}{\%} of the time it would take to run the solver on all instances of this dataset.
   We achieve this while still maintaining a prediction accuracy of about \SI{92}{\%} regarding the PAR-2 ranking of this solver if it were evaluated on the complete dataset.
-  Instance-selection decisions made by our approach also allow for discussing instance-family importance in the selection process. 
+  Additionally, we discuss the importance of instance families in the selection process. 
   Our tool provides the engineer of propositional-satisfiability solvers with a faster way of determining a new SAT-solver's performance in comparison to sampling the benchmark set by hand.
@@
 One of the main phases of algorithm engineering is experimentation.
-This is al\-so true for the archetypal $\mathcal{NP}$-complete pro\-po\-si\-tio\-nal-sat\-is\-fia\-bi\-li\-ty problem (SAT).
+This is also true for the pro\-po\-si\-tio\-nal-sat\-is\-fia\-bi\-li\-ty problem (SAT), the archetypal $\mathcal{NP}$-complete problem.
 While state-of-the-art solvers embark upon more and more problem instances, SAT-solver evaluation is steadily getting more costly.
-Usually, submissions to competitions like the annual SAT~Competition~\cite{FroleyksHIJS21} themselves precede heavy experimentation and are the remnant of a multitude of ideas and considerations.
+Usually, solvers submitted to competitions like the annual SAT~Competition~\cite{FroleyksHIJS21,sat2022} themselves precede heavy experimentation and are the remnant of a multitude of ideas and considerations.
 While the field of SAT-solver benchmarking is well established, traditional benchmark-set-selection approaches~\cite{Gelder11,HoosKSS13} do not optimize for benchmark runtime.
@@
 Within our work, we show that we can arrive at the same conclusions with high accuracy in only a fraction of the time that would be needed to evaluate the full benchmark set.
-To be more specific, we address the following problem.
+To be more specific, we address the following problem:
 
@@
 We want to find a scoring function $\operatorname{score}\!: \mathcal{A} \cup \left\lbrace \hat{a} \right\rbrace \rightarrow \mathbb{R}$ assigning each solver a score.
-Having a set of solvers with their respective scores induces a ranking.
+SAT~Competitions~\cite{FroleyksHIJS21} typically rely on the PAR-2 score, i.e, the average runtime with a penalty term of $2\tau$ for timeouts.
+We predict the solver ranking induced by this scoring, without predicting exact solver runtimes.
 This allows for both comparing the solvers as well as identifying solvers with similar behavior.
-While the annual SAT~Competitions~\cite{FroleyksHIJS21} rely on the PAR-2 score ranking, i.e, the average runtime with a penalty timeout-term of $2\tau$, which we want to predict, this work examines a model-based way of ranking solvers.
+%JB: is there any further behavior we can deduce about a solver scoring than a performance ranking?
+similar behavior could be something like two solvers being fast/slow on the same instances, but that cannot be deduced from overall score
 
 This current article is proposing a novel active-learning-based benchmark-set-selection approach.
-Our algorithm is unique in that it is the only one, to the best of our knowledge, to address all of the following features (also refer to Tab.~\ref{tab:requirements}).
+Our algorithm is unique in that it is the only one, to the best of our knowledge, that has all of the following features (also refer to Tab.~\ref{tab:requirements}).
 Our benchmarking process outputs a \textit{ranking} to compare the new solver against the existing set of solvers as described above.
 We optimize the \textit{runtime} that our strategy needs to arrive at its conclusion.
-We use problem-instance and runtime \textit{features} $\varphi\!: \mathcal{A} \times \mathcal{I} \rightarrow \mathbb{R}^n$ since it already has proven to be helpful in classic benchmark-set-selection approaches~\cite{HoosKSS13} as well as in portfolio selection~\cite{CollauttiMMO13,NgokoCT19}.
+We use problem-instance and runtime \textit{features} $\varphi\!: \mathcal{A} \times \mathcal{I} \rightarrow \mathbb{R}^n$ since such features already have proven to be helpful in classic benchmark-set-selection approaches~\cite{HoosKSS13} as well as in portfolio selection~\cite{CollauttiMMO13,NgokoCT19}.
+%JB: I think we can remove this criterion, as it also is satisified by all competitors (and is commented out in table; if we keep criterion, we should also show it in table)
 Moreover, we select instances \textit{non-randomly} and \textit{incrementally}.
@@
 Rather than solely outputting a binary classification, i.e, the new solver is worse than solver A, we provide a \textit{scoring function} that shows by which margin a solver is worse and how similar it is to existing approaches (\textit{interpretability}).
+%JB: we should rename "interpretability" criterion, term can have lots of different meanings and our specific meaning is not very intuitive
 Finally, our strategy's cost \textit{scales linearly} with the cardinality of the set of existing solvers $\mathcal{A}$ to compare against and the instances $\mathcal{I}$ within our pool of possible benchmark instances.
@@
 This would scale quadratically.
-
-To sum up our contribution, we take an existing benchmark set, e.g., the SAT Competition~2022 Anniversary Track~\cite{sat2022}, and predict a new solver's PAR2 score on this dataset with only a fraction of the runtime that would be needed to evaluate this solver on the complete set of instances.
-
-For reproducibility, our source code, all experiments, and data are available on GitHub\footnote{\url{https://github.com/mathefuchs/al-for-sat-solver-benchmarking}}.
+%JB: need to discuss this; log-linear effort might be sufficient if binary-search-/recursive-sorting-like procedure used
+
+To sum up our contribution, we take an existing benchmark set, i.e., the SAT Competition~2022 Anniversary Track~\cite{sat2022}, and predict a new solver's PAR2 score on this dataset with only a fraction of the runtime that would be needed to evaluate this solver on the complete set of instances.
+%JB: I think we should expand a bit more on experimental design (how many instances/solvers, what do we compare), and the key results; I think it's okay if the intro becomes a bit longer (and in turn, the overview of the approach's features can be cut a bit)
+
+For reproducibility, all our source code\footnote{temporary, anonymized version for review: \url{xxx}} and and data are available on GitHub\footnote{temporary, anonymized version for review: \url{xxx}}.
 
@@
 Tab.~\ref{tab:requirements} shows a concise overview of existing approaches with a related goal.
+%JB: we can probably cut this part a bit, as we discuss weakness of individual approaches later anyway
 While traditional benchmark-set-selection~\cite{HoosKSS13} lacks the aspect of runtime optimization, related work within algorithm configuration~\cite{HutterHL11} is concerned with finding better configurations fast.
@@
 Note that all approaches already make use of problem-instance features as it showed to be beneficial in past work~\cite{HoosKSS13,CollauttiMMO13,NgokoCT19}.
+%JB: as commented above, we should either remove this criterion or show it in the table
 
 \subsubsection{Traditional benchmark-set-selection.}
+%JB: I think we should do some further research on static benchmark-set selection approaches (maybe in SMT, CP, etc.)
 Recent studies~\cite{abs-2107-07002,NiesslHWCB22} show that selecting a benchmark set in general is a difficult endeavor for multiple reasons.
@@
 
-Hoos et.
-al.~\cite{HoosKSS13} discuss which properties are most desirable when selecting a set of SAT benchmark instances, for instance for a competition, e.g., the SAT Competition~2020~\cite{FroleyksHIJS21}.
+Hoos et. al.~\cite{HoosKSS13} discuss which properties are most desirable when selecting a set of SAT benchmark instances, for instance for a competition.
 The selection strategy presented is static, i.e., it does not depend on the solvers to distinguish.
@@
 Those features comprise general instance-size features and graph-representation features among others.
+%JB: we should probably cut the discussion of each approach's criteria to save space, maybe only mention the unsatisifed ones
 Their approach \textit{ranks} solvers by providing its PAR-2 score, is \textit{feature-based}, and is \textit{scalable} in the sense that the evaluation time scales linearly with more instances and solvers.
 Also, their approach is \textit{interpretable} in the sense that the obtained PAR-2 score induces a distance between solvers.
-However, they do not optimize for benchmark \textit{runtime} and select instances \textit{randomly}\footnote{Random with constraints on the instance hardness and feature distance.}.
+However, they do not optimize for benchmark \textit{runtime} and select instances \textit{randomly}, apart from constraints on the instance hardness and feature distance.
 
@@
 SMAC~\cite{HutterHL11} tries to find a close-to-optimal configuration for a given SAT-solver on a given set of instances.
+%JB: discussion of SMAC is too long, we should cut it a bit
 Although this task is different from our goal, e.g., we do not need to navigate configuration space, they use a similar framework.
@@
 While this approach considers the configuration's \textit{runtime}, is \textit{feature-based}, and \textit{scales linearly} by having a fixed time budget, it is not capable of the other three features listed in Tab.~\ref{tab:requirements}.
-First, an algorithm configurator cannot be used to \textit{rank} a new solver since it only looks at promising solvers/configurations.
+First, an algorithm configurator cannot be used to \textit{rank} a new solver since it only looks at promising solvers/configurations rather than the overall average performance.
 If the new solver does not belong to this category (new ideas are seldomly the best approach right from the start, i.e., they need some tweaking), we do not have any information on how this solver performed.
-Second, while using a model-based selection strategy to sample configurations, instance sampling is done \textit{randomly}.
+%JB: "this category" = category of most promising solvers? not 100% clear
+Second, while using a model-based selection strategy to sample configurations, instance sampling is done \textit{randomly}, i.e., without building a model over instances.
 And third, their approach is not \textit{interpretable} regarding the new solver performance since we are already missing the feature to \textit{rank} the new solver among existing ones.
+%JB: well, the internal meta-model of the configurator allows to estimate the solver performance for particular configurations (instead of comparing solvers), which allows a specific performance assessment, so the approach is interpretable in that direction (though for comparison to other solvers, only top performance should be interesting)
 
@@
 Matricon et. al.~\cite{MatriconAFSH21} present an incremental benchmark-set-selection approach.
+%JB: again, I think there are too many details for describing this single approach
 Given a pair of SAT solvers, they want to iteratively select a subset of instances to decide with high confidence which of the two solvers is better.
@@
 To update the confidence level after each experiment, they discuss using a fixed subset size, a Wilcoxon test, and a distribution-based method.
-Their experiments show that random instance-selection with a Wilcoxon test as the stopping criterion performs consistently well across different data sets in the sense that it belongs to the Pareto-front or is close to it in all cases.
+Their experiments show that random instance-selection with a Wilcoxon test as the stopping criterion performs consistently well across different data sets in the sense that it belongs to the Pareto front or is close to it in all cases.
+%JB: in our evaluation, we should state whether our results are consistent to this or not
 
@@
 As they compare only pairs of solvers, scaling with more and more solvers and ideas becomes an issue as there is a quadratic amount of pairs to distinguish.
+%JB: maybe rather log-linear than quadratic, see a prior comment
 Since a new solver idea is seldomly best at first try but requires more or less tweaking, it is desired to rank a new solver in a way that scales linearly with the number of existing approaches.
+%JB: this general motivation should go there where linearity is introduced as criterion, not here
 In this way, a new solver idea can be discarded if it performs poorly across the board or may be further tweaked if it shows promising results at least in some cases.
@@
 \subsubsection{Active learning (AL).}
+%JB: section does not make clear which of the following approaches we will use (reads more like fundamentals than related work)
 The posed \textit{new solver problem} has stark similarities to the well-studied field of active learning within recommender systems, especially the \textit{new user problem}~\cite{RubensESK15}.
 The classic exploration-exploitation dilemma arises since, on the one hand, we want to maximize the utility an instance provides to our model and, on the other hand, minimize the cost (CPU time) that is associated with its acquisition.
+%JB: no sure it is arises here as well; dilemma is typical for searching one optimal solution in continuous space; however, if we find an interesting/useful instance here, I'm not sure if we would continue picking similar instances ("exploting" instead of "exploring") just because we expect them to be useful as well
 
@@
   \node[right] at (-1.1,2.4) {GAN};
+  %JB: does the model need to be a GAN or can it also be a different type of model?
   \node[right] at (-1.3,3.135) {$x, ?$};
@@
 AL algorithms can be categorized into \textit{synthesis-bas\-ed} \cite{0001AEMN22,GarzonMG22,2019gaal} and \textit{pool-bas\-ed} approaches \cite{GolbandiKL11,distribAL,HarpaleY08,KapoorGUD07,KorenBV09,KornerW06,MelvilleM04,SinhaED19,TongK01}.
+%JB: we can probably cut some citations, the three most important should be sufficient
 While synthesis-based methods (Fig.
@@
 Recent synthesis-based methods within the field of satisfiability solving~\cite{0001AEMN22,GarzonMG22} show how to generate problem instances with desired properties.
+%JB: we should conduct some further research if there are other approaches using AL for SAT (or SMT, CP, etc.) benchmarking
 This goal is, however, orthogonal to ours.
@@
 Our algorithm follows a simple three-step schema (Alg.~\ref{algALBenchmark}).
+%JB: as it is rather simple and has different parts to be plugged in, I think we should rebrand it as "framework" rater than algorithm
 After initializing the model $\mathcal{M}$, we incrementally \textit{select} a problem instance from the pool of instances, one at a time, evaluate the new solver $\hat{a}$ on it, and \textit{update} the model with the acquired result.
@@
 \begin{algorithm}
-  \caption{Simple incremental benchmark-set-selection schema}
+  \caption{Incremental benchmark-set-selection schema}
   \label{algALBenchmark}
@@
   $\operatorname{score} \leftarrow \operatorname{predictScores}\!\left(\mathcal{M}\right)$
+  %JB: bit confusing that this is not a numerical score, but a function (see next line); maybe call it "score()"?
+or return a tuple with all estimated scores (one per solver)?; also, one could expect to get a benchmark set (see title of algorithm), but as instance selection is solver-specific, this is not case
 
   \KwOut{Return scoring function $\operatorname{score}\!: \mathcal{A} \cup \left\lbrace \hat{a} \right\rbrace \rightarrow \mathbb{R}$ and the fraction of sampled instances and runtime needed to provide it.}
+  %JB: we should formally specify the tuple rather than only describing it
 \end{algorithm}
@@
 Preliminary experiments, as well as research in portfolio solvers~\cite{NgokoCT19,CollauttiMMO13}, have shown however that directly predicting runtimes does not work well in practice.
+%JB: Collauti at el. still predict actual (though normalized) runtimes rather than discrete labels, they only heavily discretize runtimes when using them as features
 In our experiments, the mean squared error was within the same magnitude as the values to be predicted.
-In contrast to that, discretizing runtimes and using those labels in the model's training and prediction yields quite good results (cf.~Sec.~\ref{sec:eval}).
+In contrast to that, discretizing runtimes and using those labels in the model's training and prediction yields good results (cf.~Sec.~\ref{sec:eval}).
 The following paragraphs motivate why it makes sense to discretize runtimes.
-
-By discretizing runtimes into $k$ clusters, we mean using a function $\gamma\!: {\mathcal{A} \times \mathcal{I}} \rightarrow \left\lbrace 1, \dots, k \right\rbrace$ to transform runtime observations.
+%JB: we probably should cut this a bit
+
+We use a function $\gamma\!: {\mathcal{A} \times \mathcal{I}} \rightarrow \left\lbrace 1, \dots, k \right\rbrace$ to transform runtime observations.
+In our experiments, this function is a clustering algorithm.
 Note that this transformation is done per instance to adapt to different instance difficulties. 
@@
   \operatorname{score}\!\left(a\right) := \frac{1}{|\mathcal{I}|} \sum_{i \in \mathcal{I}} w_{a,i} \cdot \left(\gamma\!\left(a, i\right) - 1\right) \enspace \textrm{,}
+  %JB: does the "-1" provide any benefits, or does it only shift scores?
   \label{eq:rankingeq}
@@
 To be applicable to $\hat{a}$, we extend $\operatorname{score}$ by replacing $\gamma\!\left(a, i\right)$ by the model's predictions $f\!: \left(\mathcal{A} \cup \left\lbrace \hat{a} \right\rbrace\right) \times \mathcal{I} \rightarrow \left\lbrace 1, \dots, k \right\rbrace$.
+%JB: as "$\operatorname{score}$" is not visible, we should use some other way of highlighting
 
@@
 In our work, we use a hierarchical-clustering approach with a single-link criterion for $k = 3$.
-For the SAT Competition~2022 Anniversary Track instances~\cite{sat2022}, this ranking approach is capable of deciding almost all pairs of solvers correctly (about \SI{97.45}{\%}; $\sigma = \SI{3.68}{\%}$).
+For the SAT Competition~2022 Anniversary Track instances~\cite{sat2022}, this ranking approach is capable of correctly deciding for almost all pairs of solvers which one is faster (about \SI{97.45}{\%}; $\sigma = \SI{3.68}{\%}$).
 Moreover, it decides all solvers correctly if the difference in the cluster-label-ranking scores between two solvers is at least \SI{0.11}{}.
-Refer to Tab.~\ref{tab:discreteranking} for a comparison with the PAR-2 ranking.
+%Refer to Tab.~\ref{tab:discreteranking} for a comparison with the PAR-2 ranking.
 The Spearman correlation~\cite{de2016comparing} of cluster-label and PAR-2 ranking is about \SI{0.988} (in comparison to a perfect value of \SI{1}).
@@
 
-\begin{table}[tbp]
-  \centering
-  \caption{PAR-2 ranking vs. ranking based on discretized runtimes for the SAT Competition~2022 Anniversary Track instances~\cite{sat2022} ($k = 3$; hierarchical clustering)}
-  \label{tab:discreteranking}
-  \begin{tabular}{
-    c
-    >{\centering\arraybackslash}p{1.8cm}p{3.6cm}
-    >{\centering\arraybackslash}p{1.8cm}p{3.6cm}
-  }
-    \hline
-    Rank & \multicolumn{2}{c}{PAR-2 Ranking} & \multicolumn{2}{c}{Cluster-Label Ranking} \\
-    \hline
-1   & 2806.400 & Kissat\_MAB\_ESA           & 1.172 & Kissat\_MAB\_ESA \\
-2   & 2810.959 & kissat-sc2022-bulky        & 1.183 & kissat-sc2022-bulky \\
-3   & 2830.143 & ekissat-mab-gb-db          & 1.185 & ekissat-mab-gb-db \\
-\cellcolor{gray!15} 4   & \cellcolor{gray!15} 2833.081 & \cellcolor{gray!15} Kissat\_MAB\_UCB           & \cellcolor{gray!15} 1.186 & \cellcolor{gray!15} kissat\_inc \\
-\cellcolor{gray!15} 5   & \cellcolor{gray!15} 2833.397 & \cellcolor{gray!15} kissat\_inc                & \cellcolor{gray!15} 1.186 & \cellcolor{gray!15} Kissat\_MAB\_UCB \\
-6   & 2841.734 & ekissat-mab-db-v1          & 1.192 & ekissat-mab-db-v1 \\
-7   & 2842.666 & Kissat\_MAB\_MOSS          & 1.192 & Kissat\_MAB\_MOSS \\
-8   & 2857.219 & Kissat\_MAB-HyWalk         & 1.195 & Kissat\_MAB-HyWalk \\
-9   & 2866.960 & kissat-sc2022-light        & 1.120 & kissat-sc2022-light \\
-10  & 2896.371 & kissat-els-v2              & 1.216 & kissat-els-v2 \\
-11  & 2950.179 & hKis-unsat                 & 1.229 & hKis-unsat \\
-12  & 2964.729 & Kissat\_adaptive\_restart  & 1.234 & Kissat\_adaptive\_restart \\
-13  & 2974.990 & SeqFROST-NoExtend          & 1.247 & SeqFROST-NoExtend \\
-14  & 3010.642 & Cadical\_ESA               & 1.248 & Cadical\_ESA \\
-\cellcolor{gray!15} 15  & \cellcolor{gray!15} 3012.152 & \cellcolor{gray!15} kissat-els-v1              & \cellcolor{gray!15} 1.258 & \cellcolor{gray!15} CadicalReorder \\
-\cellcolor{gray!15} 16  & \cellcolor{gray!15} 3028.922 & \cellcolor{gray!15} CadicalReorder             & \cellcolor{gray!15} 1.264 & \cellcolor{gray!15} kissat-els-v1 \\
-\cellcolor{gray!15} 17  & \cellcolor{gray!15} 3042.540 & \cellcolor{gray!15} cadical\_rel\_Scavel       & \cellcolor{gray!15} 1.278 & \cellcolor{gray!15} CaDiCaL\_DVDL\_V1 \\
-\cellcolor{gray!15} 18  & \cellcolor{gray!15} 3082.293 & \cellcolor{gray!15} kissat\_relaxed            & \cellcolor{gray!15} 1.283 & \cellcolor{gray!15} CaDiCaL\_DVDL\_V2 \\
-\cellcolor{gray!15} 19  & \cellcolor{gray!15} 3089.304 & \cellcolor{gray!15} CaDiCaL\_DVDL\_V1          & \cellcolor{gray!15} 1.298 & \cellcolor{gray!15} kissat\_relaxed \\
-\cellcolor{gray!15} 20  & \cellcolor{gray!15} 3095.118 & \cellcolor{gray!15} CaDiCaL\_DVDL\_V2          & \cellcolor{gray!15} 1.361 & \cellcolor{gray!15} cadical\_rel\_Scavel \\
-21  & 3272.674 & glucose-reboot             & 1.379 & glucose-reboot \\
-\cellcolor{gray!15} 22  & \cellcolor{gray!15} 3287.322 & \cellcolor{gray!15} LStech-Maple-HyWalk        & \cellcolor{gray!15} 1.414 & \cellcolor{gray!15} hCaD\_V1-psids \\
-23  & 3288.798 & LSTech\_Maple              & 1.446 & LSTech\_Maple \\
-\cellcolor{gray!15} 24  & \cellcolor{gray!15} 3399.281 & \cellcolor{gray!15} SLIME-SC-2022-beta         & \cellcolor{gray!15} 1.469 & \cellcolor{gray!15} LStech-Maple-HyWalk \\
-\cellcolor{gray!15} 25  & \cellcolor{gray!15} 3410.345 & \cellcolor{gray!15} SLIME-SC-2022              & \cellcolor{gray!15} 1.470 & \cellcolor{gray!15} SLIME-SC-2022-beta \\
-\cellcolor{gray!15} 26  & \cellcolor{gray!15} 3430.403 & \cellcolor{gray!15} hCaD\_V1-psids             & \cellcolor{gray!15} 1.523 & \cellcolor{gray!15} SLIME-SC-2022 \\
-27  & 3504.244 & MapleLCMDistChrBt-DL-v3    & 1.542 & MapleLCMDistChrBt-DL-v3 \\
-28  & 4750.808 & IsaSAT                     & 2.062 & IsaSAT \\
-    \hline
-  \end{tabular}
-\end{table}
+%\begin{table}[tbp]
+%  \centering
+%  \caption{PAR-2 ranking vs.
+discretized-runtime ranking for the SAT Competition~2022 Anniversary Track instances~\cite{sat2022} ($k = 3$; hierarchical clustering).
+Differences are highlighted with a gray background.}
+%  \label{tab:discreteranking}
+%  \begin{tabular}{
+%    c
+%    >{\centering\arraybackslash}p{1.8cm}p{3.6cm}
+%    >{\centering\arraybackslash}p{1.8cm}p{3.6cm}
+%  }
+%    \hline
+%    Rank & \multicolumn{2}{c}{PAR-2 Ranking} & \multicolumn{2}{c}{Cluster-Label Ranking} \\
+%    \hline
+%1   & 2806.400 & Kissat\_MAB\_ESA           & 1.172 & Kissat\_MAB\_ESA \\
+%2   & 2810.959 & kissat-sc2022-bulky        & 1.183 & kissat-sc2022-bulky \\
+%3   & 2830.143 & ekissat-mab-gb-db          & 1.185 & ekissat-mab-gb-db \\
+%\cellcolor{gray!15} 4   & \cellcolor{gray!15} 2833.081 & \cellcolor{gray!15} Kissat\_MAB\_UCB           & \cellcolor{gray!15} 1.186 & \cellcolor{gray!15} kissat\_inc \\
+%\cellcolor{gray!15} 5   & \cellcolor{gray!15} 2833.397 & \cellcolor{gray!15} kissat\_inc                & \cellcolor{gray!15} 1.186 & \cellcolor{gray!15} Kissat\_MAB\_UCB \\
+%6   & 2841.734 & ekissat-mab-db-v1          & 1.192 & ekissat-mab-db-v1 \\
+%7   & 2842.666 & Kissat\_MAB\_MOSS          & 1.192 & Kissat\_MAB\_MOSS \\
+%8   & 2857.219 & Kissat\_MAB-HyWalk         & 1.195 & Kissat\_MAB-HyWalk \\
+%9   & 2866.960 & kissat-sc2022-light        & 1.120 & kissat-sc2022-light \\
+%10  & 2896.371 & kissat-els-v2              & 1.216 & kissat-els-v2 \\
+%11  & 2950.179 & hKis-unsat                 & 1.229 & hKis-unsat \\
+%12  & 2964.729 & Kissat\_adaptive\_restart  & 1.234 & Kissat\_adaptive\_restart \\
+%13  & 2974.990 & SeqFROST-NoExtend          & 1.247 & SeqFROST-NoExtend \\
+%14  & 3010.642 & Cadical\_ESA               & 1.248 & Cadical\_ESA \\
+%\cellcolor{gray!15} 15  & \cellcolor{gray!15} 3012.152 & \cellcolor{gray!15} kissat-els-v1              & \cellcolor{gray!15} 1.258 & \cellcolor{gray!15} CadicalReorder \\
+%\cellcolor{gray!15} 16  & \cellcolor{gray!15} 3028.922 & \cellcolor{gray!15} CadicalReorder             & \cellcolor{gray!15} 1.264 & \cellcolor{gray!15} kissat-els-v1 \\
+%\cellcolor{gray!15} 17  & \cellcolor{gray!15} 3042.540 & \cellcolor{gray!15} cadical\_rel\_Scavel       & \cellcolor{gray!15} 1.278 & \cellcolor{gray!15} CaDiCaL\_DVDL\_V1 \\
+%\cellcolor{gray!15} 18  & \cellcolor{gray!15} 3082.293 & \cellcolor{gray!15} kissat\_relaxed            & \cellcolor{gray!15} 1.283 & \cellcolor{gray!15} CaDiCaL\_DVDL\_V2 \\
+%\cellcolor{gray!15} 19  & \cellcolor{gray!15} 3089.304 & \cellcolor{gray!15} CaDiCaL\_DVDL\_V1          & \cellcolor{gray!15} 1.298 & \cellcolor{gray!15} kissat\_relaxed \\
+%\cellcolor{gray!15} 20  & \cellcolor{gray!15} 3095.118 & \cellcolor{gray!15} CaDiCaL\_DVDL\_V2          & \cellcolor{gray!15} 1.361 & \cellcolor{gray!15} cadical\_rel\_Scavel \\
+%21  & 3272.674 & glucose-reboot             & 1.379 & glucose-reboot \\
+%\cellcolor{gray!15} 22  & \cellcolor{gray!15} 3287.322 & \cellcolor{gray!15} LStech-Maple-HyWalk        & \cellcolor{gray!15} 1.414 & \cellcolor{gray!15} hCaD\_V1-psids \\
+%23  & 3288.798 & LSTech\_Maple              & 1.446 & LSTech\_Maple \\
+%\cellcolor{gray!15} 24  & \cellcolor{gray!15} 3399.281 & \cellcolor{gray!15} SLIME-SC-2022-beta         & \cellcolor{gray!15} 1.469 & \cellcolor{gray!15} LStech-Maple-HyWalk \\
+%\cellcolor{gray!15} 25  & \cellcolor{gray!15} 3410.345 & \cellcolor{gray!15} SLIME-SC-2022              & \cellcolor{gray!15} 1.470 & \cellcolor{gray!15} SLIME-SC-2022-beta \\
+%\cellcolor{gray!15} 26  & \cellcolor{gray!15} 3430.403 & \cellcolor{gray!15} hCaD\_V1-psids             & \cellcolor{gray!15} 1.523 & \cellcolor{gray!15} SLIME-SC-2022 \\
+%27  & 3504.244 & MapleLCMDistChrBt-DL-v3    & 1.542 & MapleLCMDistChrBt-DL-v3 \\
+%28  & 4750.808 & IsaSAT                     & 2.062 & IsaSAT \\
+%    \hline
+%  \end{tabular}
+%\end{table}
 
@@
 Within the SAT~Competition~2022 Anniversary Track, about \SI{89.95}{\%} of solver pairs are significantly different (according to a Wilcoxon-signed-rank test with $\alpha = 0.05$).
-While only time-out labels (solver times out or not on a particular instance) are sufficient to significantly distinguish about \SI{87.04}{\%} of solver pairs, discrete-runtime labels\footnote{Based on hierarchical clustering with a log-single-link criterion for $k = 3$.} are sufficient to distinguish about \SI{87.83}{\%} of solver pairs.
+While only time-out labels (solver times out or not on a particular instance) are sufficient to significantly distinguish about \SI{87.04}{\%} of solver pairs, discrete-runtime labels are sufficient to distinguish about \SI{87.83}{\%} of solver pairs.
+%\footnote{Based on hierarchical clustering with a log-single-link criterion for $k = 3$.}
 This means that we do not lose much discriminatory power using discrete runtime labels.
 The number of significant comparisons between solvers \textit{alone}, though, does only offer limited insights as we might achieve significant differences in all pairs by assigning some randomly skewed values as labels.
+%JB: thus, we should reorder section to mention significant difference (this paragraph) first and deciding solver pairs (previous paragraph, stronger argument) second (rather than giving the stronger argument first, making the second argument superfluous)
 However, the predictiveness of the PAR-2 score, the labels' discriminatory power, as well as that it enables us to use classification at the same time, show the clear advantage of runtime discretization.
+%JB: at the same time with what? with the other benefits?
 
@@
 First, we present possible baseline approaches including existing work on how to pick \textit{interesting} instances~\cite{Bossek021a,GentHJKMNN14,MatriconAFSH21}.
-Second, we show promising machine-learning-based sampling strategies that make use of the model's label prediction function $f$ and are inspired by existing work within the realms of active learning~\cite{settles2009active}.
+%JB: approaches under "Second" are also at least "inspired by existing work" and probably also pick "interesting" instances, distinction between the two categories not entirely clear
+Second, we show promising machine-learning-based sampling strategies that make use of the model's label-prediction function $f$ and are inspired by existing work within the realms of active learning~\cite{settles2009active}.
 
@@
 Bossek and Wagner~\cite{Bossek021a} introduce a custom ranking-score metric incorporating the number of \textit{good} pairs and \textit{bad} pairs.
+%JB: some selection methods, like this one, are not mentioned later any more (in particular, not in experimental design), so we don't need to describe them in detail (maybe mention shortly if used in preliminary experiments, but turned out to be not useful)
 A good solver pair, regarding a particular instance, are solvers that perform according to their global ranking.
@@
 By selecting instances with the highest score according to the lexicographical order of the tuples, they try to favor instances that are predictive of the final ranking.
-This assumes though that the new solver $\hat{a}$ behaves similarly to the existing solvers.
+This assumes, though, that the new solver $\hat{a}$ behaves similarly to the existing solvers.
 
@@
 \label{sec:sampling2}
-For conciseness, we only present two of our active-learning-based instance-selection approaches that show the most promising results.
+For conciseness, we only present two of our active-learning-based instance-selection approaches, which show the most promising results.
 We implement a model-uncertainty-based approach and a model-based information-gain sampling strategy.
@@
 
-The model-based information-gain sampling strategy selects the instance with the highest expected information-entropy reduction, assuming that the model's runtime-label probabilities for an instance are close to being correct.
+The model-based information-gain sampling strategy selects the instance with the highest expected entropy reduction, assuming that the model's runtime-label probabilities for an instance are close to being correct.
+%JB: not sure what "close to being correct" means excactly: does strategy assume correct predictions, but implicitly has a tolerance for wrong predictions, or does it explicitly account for errors somehow?
 To be more specific, we select the instance $i$ that maximizes
@@
 The Wilcoxon stopping criterion stops the iterative active-learning algorithm when we are on average certain enough that the predicted runtime labels are sufficiently different in comparison to the other solvers.
-Note that this is a rather heuristic criterion, it shows promising results nevertheless.
 To be more specific, we stop when
@@
 drops below a certain threshold.
-Thereby, \textit{wilcoxon} denotes the $p$-value of a Wilcoxon signed-rank test of the given runtime-label distributions and $\gamma\!\left(a, \mathcal{I}\right)$ and $f\!\left(\hat{a}, \mathcal{I}\right)$ are the expected set extensions on $\gamma$ and $f$ respectively.
+%JB: if this is info gain regarding runtimes, doesn't Eq.
+(4) mean that we would prefer instances with low entropy of runtime distribution after considering the new solver, i.e., instances that are not very discriminatory?
+Thereby, \textit{wilcoxon} denotes the $p$-value of a Wilcoxon signed-rank test of the runtime-label distributions and $\gamma\!\left(a, \mathcal{I}\right)$ and $f\!\left(\hat{a}, \mathcal{I}\right)$ are the expected set extensions on $\gamma$ and $f$ respectively.
+%JB: "set extension" might be rather technical and hard to understand; also, why "expected"? in the sense of "straightforward"?
 To improve the stability of this criterion, we use an exponential moving average
@@
 It simply stops the active-learning process (Alg.~\ref{algALBenchmark}) if within the last $l$ iterations the ranking that is induced by the model's predictions (Eq.~\ref{eq:rankingeq}) remains unchanged.
-Note that the concrete values of $\operatorname{score}\!\left(a\right)$ might change.
+Note that the concrete values of $\operatorname{score}\!\left(a\right)$ might still change.
 We are solemnly interested in the induced ranking in this case.
@@
 Fig.~\ref{fig:e2eallsolvers} reveals the performance of the approaches listed in Sec.~\ref{sec:hyper} on $O_{\operatorname{rt}}$-$O_{\operatorname{acc}}$-diagrams for the smaller hyper-parameter tuning dataset.
-The plotted lines represent the Pareto-front for each possible choice of ranking approaches (Fig.~\ref{fig:annitraincolorranking}), selection approaches (Fig.~\ref{fig:annitraincolorselection}), and stopping criteria (Fig.~\ref{fig:annitraincolorstopping}).
+The plotted lines represent the Pareto front for each possible choice of ranking approaches (Fig.~\ref{fig:annitraincolorranking}), selection approaches (Fig.~\ref{fig:annitraincolorselection}), and stopping criteria (Fig.~\ref{fig:annitraincolorstopping}).
 Colors indicate the different hyper-parameter instantiations.
-To be more specific, they represent the Pareto-front of all configurations with the respective instantiations.
+To be more specific, they represent the Pareto front of all configurations with the respective instantiations.
 
@@
     }
-    \caption{$O_{\operatorname{rt}}$-$O_{\operatorname{acc}}$-diagram showing Pareto-fronts for configurations with the given ranking approaches.
+    \caption{$O_{\operatorname{rt}}$-$O_{\operatorname{acc}}$-diagram showing Pareto fronts for configurations with the given ranking approaches.
 The lines show the best performances of approaches using runtime-label-prediction ranking and partially observed PAR-2 ranking respectively.}
@@
     }
-    \caption{$O_{\operatorname{rt}}$-$O_{\operatorname{acc}}$-diagram showing Pareto-fronts for configurations with the given instance-selection approaches.
+    \caption{$O_{\operatorname{rt}}$-$O_{\operatorname{acc}}$-diagram showing Pareto fronts for configurations with the given instance-selection approaches.
 The lines show the best performances of approaches using uncertainty-based, information-gain-based, and random instance-selection respectively.}
@@
     }
-    \caption{$O_{\operatorname{rt}}$-$O_{\operatorname{acc}}$-diagram showing Pareto-fronts for configurations with the given stopping criteria.
+    \caption{$O_{\operatorname{rt}}$-$O_{\operatorname{acc}}$-diagram showing Pareto fronts for configurations with the given stopping criteria.
 The lines show the best performances of approaches using the ranking-convergence criterion, Wilcoxon criterion, and a fixed subset size respectively.}

### 2022-09-27 paper/paper.tex
Add comments from PdF paper, part 2 (Sections 4-6)
... and slightly adapt text in some places (comments I directly resolved or
other stuff I noticed). Sections 1-3 are affected as well for consistency
reasons (e.g., add further line breaks between sentences and harmonize
usage of term "New-Solver Problem").

@@
   Benchmark-set selection is the process of choosing a subset of instances out of a pool to be representative and reliable in discriminating SAT solvers.
-  This paper presents an active-learning-based benchmark-set-selec\-tion approach that solves the new-solver problem:
+  This paper presents an active-learning-based benchmark-set-selec\-tion approach that solves the \textit{New-Solver Problem}:
   Given a new solver, we want to find its ranking amidst its competitors with as little runtime as possible while still giving accurate ranking predictions.
@@
 
-\begin{definition}[New solver problem]
+\begin{definition}[New-Solver Problem]
 	Given a pool of instances $\mathcal{I}$ and solvers $\mathcal{A}$ with already known runtimes $r\!: \mathcal{A} \times \mathcal{I} \rightarrow \left[0, \tau\right] \cup \left\lbrace \perp \right\rbrace$ (timeout $\tau$ usually \SI{5000}{s}~\cite{FroleyksHIJS21}), the problem is to decide with as little resources (CPU time) as possible how a new solver $\hat{a} \notin \mathcal{A}$ performs in comparison to the existing ones by incrementally selecting a benchmark set $\tilde{\mathcal{I}}$ on which to evaluate $\hat{a}$.
+	\label{def:new-solver-problem}
 \end{definition}
@@
 
-To sum up our contribution, we take an existing benchmark set, i.e., the SAT Competition~2022 Anniversary Track~\cite{sat2022}, and predict a new solver's PAR2 score on this dataset with only a fraction of the runtime that would be needed to evaluate this solver on the complete set of instances.
+To sum up our contribution, we take an existing benchmark set, i.e., the SAT Competition~2022 Anniversary Track~\cite{sat2022}, and predict a new solver's PAR-2 score on this dataset with only a fraction of the runtime that would be needed to evaluate this solver on the complete set of instances.
 %JB: I think we should expand a bit more on experimental design (how many instances/solvers, what do we compare), and the key results; I think it's okay if the intro becomes a bit longer (and in turn, the overview of the approach's features can be cut a bit)
@@
 Third, we discuss an existing benchmark-set-selection approach that optimizes for benchmark runtime.
-Finally, we show how active-learning techniques can help to tackle the aforementioned \textit{new-solver} problem.
+Finally, we show how active-learning techniques can help to tackle the aforementioned \textit{New-Solver Problem}.
 
@@
   \centering
-  \caption{Support for desired features that are introduced in Sec.~\ref{sec:intro}.
-The shown approaches include traditional benchmark-set-selection techniques~\cite{HoosKSS13}, an algorithm configuration system~\cite{HutterHL11}, an existing active-learning approach~\cite{MatriconAFSH21}, as well as our approach.}
+  \caption{
+  	Support for desired features that are introduced in Sec.~\ref{sec:intro}.
+  	The shown approaches include traditional benchmark-set-selection techniques~\cite{HoosKSS13}, an algorithm configuration system~\cite{HutterHL11}, an existing active-learning approach~\cite{MatriconAFSH21}, as well as our approach.
+  }
   \label{tab:requirements}
@@
 %JB: section does not make clear which of the following approaches we will use (reads more like fundamentals than related work)
-The posed \textit{new solver problem} has stark similarities to the well-studied field of active learning within recommender systems, especially the \textit{new user problem}~\cite{RubensESK15}.
+The posed \textit{New-Solver Problem} has stark similarities to the well-studied field of active learning within recommender systems, especially the \textit{new user problem}~\cite{RubensESK15}.
 The classic exploration-exploitation dilemma arises since, on the one hand, we want to maximize the utility an instance provides to our model and, on the other hand, minimize the cost (CPU time) that is associated with its acquisition.
@@
 Moreover, we have briefly outlined how active-learning techniques can help to incrementally select benchmark sets.
-In the following sections, we present an active-learning-based benchmark-set-selection approach that solves the \textit{new solver problem} while optimizing for runtime and being both interpretable and scalable.
+In the following sections, we present an active-learning-based benchmark-set-selection approach that solves the \textit{New-Solver Problem} while optimizing for runtime and being both interpretable and scalable.
 
@@
 %  \centering
-%  \caption{PAR-2 ranking vs.
-discretized-runtime ranking for the SAT Competition~2022 Anniversary Track instances~\cite{sat2022} ($k = 3$; hierarchical clustering).
-Differences are highlighted with a gray background.}
+%   \caption{
+%     PAR-2 ranking vs. discretized-runtime ranking for the SAT Competition~2022 Anniversary Track instances~\cite{sat2022} ($k = 3$; hierarchical clustering).
+%	  Differences are highlighted with a gray background.
+%	}
 %  \label{tab:discreteranking}
@@
 \label{sec:goal}
-As already stated in the introductory section, this work addresses the \textit{new solver problem}.
+As already stated in the introductory section, this work addresses the \textit{New-Solver Problem} (cf.~Def.~\ref{def:new-solver-problem}).
 By running Alg.~\ref{algALBenchmark}, we obtain a prediction model $\mathcal{M}$ that provides us with a scoring function \textit{score}.
@@
 First and foremost, our goal is to provide the engineer of new SAT solvers with an accurate ranking.
-We define the ranking accuracy $O_{\operatorname{acc}} \in \left[0, 1\right]$ (higher is better) by the fraction of pairs $\left(\hat{a}, a\right)$ for all $a \in \mathcal{A}$ that are decided correctly by the given ranking.
+We define the \textit{ranking accuracy} $O_{\operatorname{acc}} \in \left[0, 1\right]$ (higher is better) by the fraction of pairs $\left(\hat{a}, a\right)$ for all $a \in \mathcal{A}$ that are decided correctly by the given ranking.
 For now, we exclude the equality of solvers since data shows that the PAR-2 scores of all our solvers are different.
@@
 Second, we also have to optimize for runtime.
-The fraction of runtime that the algorithm needs to arrive at its conclusion is denoted by $O_{\operatorname{rt}} \in \left[0, 1\right]$ (lower is better).
-In that sense, we want to find an approach that maximizes
+The \textit{fraction of runtime} that the algorithm needs to arrive at its conclusion is denoted by $O_{\operatorname{rt}} \in \left[0, 1\right]$ (lower is better).
+This metric puts the runtime summed over the sampled instances in relation to the runtime summed over all instances in the dataset.
+Overall, we want to find an approach that maximizes
 \begin{equation}
@@
     $O_{\operatorname{rt}} \leftarrow \operatorname{computeRuntimeFractionUsed}\!\left( runtime,\, \mathcal{I},\, \mathcal{S},\, r \right)$
+    %JB: do we need to call a function here or can't we directly write how this is computed (define a variable starting at 0 to which the runtime fraction is added each iteration)
 
@@
 For hyper-parameter tuning, we randomly sample \SI{10}{\%} of the complete set of 5355 instances with stratification regarding the instance's family.
-All instance families that are too \textit{small}, i.e., \SI{10}{\%} of it is less than one, are put into one meta-family for stratification.
+All instance families that are too \textit{small}, i.e., \SI{10}{\%} of them corresponds to less than one instance, are put into one meta-family for stratification.
 This smaller dataset allows for a more extensive exploration of hyper-parameter space.
@@
 Note that we are not considering all previously listed approaches since a grid search of all combinations would be infeasible.
-Rather, we filter approaches based on preliminary experimental results (cf.
-Sec.~\ref{sec:evalprel}) and do the main end-to-end experiments only with a subset.
+Rather, we filter approaches based on preliminary experimental results (cf. Sec.~\ref{sec:evalprel}) and do the main end-to-end experiments only with a subset.
 The end-to-end experiment configurations are given below.
+%JB: still interesting to see which approaches were considered in preliminary experiments; could also list them here and indicate somehow which methods were only used preliminarily (e.g., by putting "(preliminray experiments only)" or some shorter label/sign behind them)
 
@@
 Regarding \textit{ranking} (cf. Sec.~\ref{sec:main1}), we experiment with the following approaches, including our used hyper-parameter values:
+%JB: as these approaches are mostly new (Section 3.1 only explains prediction-based ranking in general, neither observed ranking nor history nor fallback), I'm not sure if we should provide further motivation/explanation here); also applies to other components of experimental design
 \begin{itemize}
@@
     \item Fallback threshold: if the difference of ranking scores drops below \SI{0.01}, \SI{0.05}, or \SI{0.1}, use the partially observed PAR-2 ranking as a tie-breaker.
+    %JB: "difference of ranking" between a pair of new solver and another solver? or has this phrase a different meaning?
   \end{itemize}
@@
 For \textit{selection} (cf. Sec.~\ref{sec:sampling1}), we experiment with the following methods, including our used hyper-parameter values:
+%JB: sampling section describes more methods than the ones listed below, I guess some got dropped after preliminary experiments
 \begin{itemize}
@@
     \item Runtime scaling: prefer instances with the greatest uncertainty per average runtime (\texttt{True} or \texttt{False}).
+    %JB: how does this scaling work exactly? divide runtimes by average runtime of all solvers on that instance?
   \end{itemize}
@@
     \item Convergence duration: stop if the predicted ranking stays the same for the last \SI{1}{\%} or \SI{2}{\%} of sampled instances.
+    %JB: why this fraction and not an absolute number of instances?
+the definition here might make stopping in early iterations rather likely, as just very few consecutive instances are needed (which might happen by coincidence; e.g., for ~5k instances, 2% minimum amount-> ~100 instances, 1% convergence duration -> ~1 instance), while the criterion becomes harder in later iterations
   \end{itemize}
@@
 %the proposed runtime label ranking, (2) the PAR-2 subsample ranking, and (3) the runtime label ranking that uses the PAR-2 subsample ranking as a fallback when performance differences drop below a threshold.
-Possible values are \SI{0.01}{}, \SI{0.05}{}, and \SI{0.1}{}.
-%Also, we experiment with different lengths of the prediction history that is used for stabilizing the ranking. Possible values are 10, 20, 30, 40, and 50.
+%Possible values are \SI{0.01}{}, \SI{0.05}{}, and \SI{0.1}{}.
+%Also, we experiment with different lengths of the prediction history that is used for stabilizing the ranking.
+%Possible values are 10, 20, 30, 40, and 50.
 
@@
 For reproducibility, our source code, all experiments, and data are available on GitHub\footnote{\url{https://github.com/mathefuchs/al-for-sat-solver-benchmarking}}.
-It is implemented in \textsc{Python} using \textit{scikit-learn}~\cite{scikit-learn} for making predictions and \textit{gbd-tools}~\cite{IserS18} for SAT-instance retrieval.
+Our code is implemented in \textsc{Python} using \textit{scikit-learn}~\cite{scikit-learn} for making predictions and \textit{gbd-tools}~\cite{IserS18} for SAT-instance retrieval.
 
@@
 Thereafter, we look at end-to-end results and instance-family importance.
+%JB: instance families not really motivated till now, only appear in abstract and in text regarding stratification
 
@@
 \label{sec:evalprel}
+%JB: if we dedicate a whole section to these results, we should probably rename this section (not "preliminary" in the typical sense); for TACAS, I guess this can be cut a bit
 An exhaustive grid search of all approaches outlined in Sec.~\ref{sec:main} is infeasible.
 We are, therefore, narrowing down the search by filtering based on preliminary results.
-Since all relevant decisions, i.e., \textit{selection}, \textit{stopping}, and \textit{ranking}, revolve around the choice of prediction model, we are first looking at the \textit{Matthews-Correlation coefficient} (MCC) scores~\cite{gorodkin2004comparing,matthews1975comparison} that a particular prediction model achieves.
+Since all relevant decisions, i.e., \textit{selection}, \textit{stopping}, and \textit{ranking}, revolve around the choice of prediction model, we are first looking at the \textit{Matthews Correlation Coefficient} (MCC) scores~\cite{gorodkin2004comparing,matthews1975comparison} that a particular prediction model achieves.
 MCC scores are great in dealing with class imbalances in contrast to conventional metrics like accuracy.
@@
   \centering
-  \caption{Time-out prediction Matthews-Correlation coefficient performances.
+  \caption{
+  	Time-out prediction Matthews Correlation Coefficient performances.
   	Results are obtained using cross-validation.
-We provide \SI{64}{\%} of ground-truth for training, use \SI{16}{\%} as a validation set, and \SI{20}{\%} as a hold-out test set (5-fold cross-validation).}
+  	We provide \SI{64}{\%} of ground-truth for training, use \SI{16}{\%} as a validation set, and \SI{20}{\%} as a test set (5-fold cross-validation).
+  }
   \label{tab:mcc}
@@
     Stacking ensemble (QDA, RF)           & 0.9527 (0.0292) \\
-    Quadratic-Discriminant analysis (QDA) & 0.9290 (0.0339) \\
+    Quadratic-discriminant analysis (QDA) & 0.9290 (0.0339) \\
     Random forest (RF)                    & 0.8530 (0.0479) \\
@@
     Logistic regression                   & 0.8052 (0.1018) \\
-    kNN-Classifier                        & 0.7885 (0.1521) \\
-    MLP-Classifier                        & 0.7760 (0.1408) \\
-    Support-Vector classifier             & 0.7757 (0.2149) \\
+    kNN classifier                        & 0.7885 (0.1521) \\
+    MLP classifier                        & 0.7760 (0.1408) \\
+    Support-vector classifier             & 0.7757 (0.2149) \\
     Naive Bayes                           & 0.7306 (0.1394) \\
@@
 
-To obtain the results in Tab.~\ref{tab:mcc}, we perform cross-validation among all solvers: we assume that all other solver runtimes are known and provide \SI{64}{\%} of ground-truth runtime labels of the target solver for training, use \SI{16}{\%} as a validation set for hyper-parameter tuning, and \SI{20}{\%} as a hold-out test set (5-fold cross-validation).
+To obtain the results in Tab.~\ref{tab:mcc}, we perform cross-validation among all solvers: we assume that all other solver runtimes are known and provide \SI{64}{\%} of ground-truth runtime labels of the target solver for training, use \SI{16}{\%} as a validation set for hyper-parameter tuning, and \SI{20}{\%} as a test set (5-fold cross-validation).
 This is repeated for each solver once.
 Note that we perform cross-validation on two levels, i.e., among the solvers and the target solver runtimes.
-We report the average performances on the hold-out test sets.
-Since the differences in the MCC-performance distributions of the best and second-best approaches are statistically significant regarding a Wilcoxon signed-rank test with $\alpha = \SI{5}{\%}$, we use the former for all further experiments as a prediction model.
+We report the average performances on the test sets.
+Since the differences in the MCC-performance distributions of the best and second-best approaches are statistically significant regarding a Wilcoxon signed-rank test with $\alpha = \SI{5}{\%}$, we only use the best approach for all further experiments as a prediction model.
 
@@
 Predicting runtime labels in a similar experiment as above produced an MCC score of on average \SI{88.82}{\%} ($\sigma = \SI{4.75}{\%}$; in comparison to \SI{95.27}{\%} for predicting time-outs).
+%: JB: " similar experiment as above" - reference unclear; what do you want to say?
 
@@
 Our main end-to-end experiments follow the evaluation framework that has been introduced in Sec.~\ref{sec:evalframe}.
-Fig.~\ref{fig:e2eallsolvers} reveals the performance of the approaches listed in Sec.~\ref{sec:hyper} on $O_{\operatorname{rt}}$-$O_{\operatorname{acc}}$-diagrams for the smaller hyper-parameter tuning dataset.
+Fig.~\ref{fig:e2eallsolvers} shows the performance of the approaches listed in Sec.~\ref{sec:hyper} on $O_{\operatorname{rt}}$-$O_{\operatorname{acc}}$-diagrams for the smaller hyper-parameter-tuning dataset.
 The plotted lines represent the Pareto front for each possible choice of ranking approaches (Fig.~\ref{fig:annitraincolorranking}), selection approaches (Fig.~\ref{fig:annitraincolorselection}), and stopping criteria (Fig.~\ref{fig:annitraincolorstopping}).
+%JB: as far as I understand it, the lines only fully compare some top-level hyperparameters, but for each of these hyperparemeters, take the best out of some more fine-grained settings (the hyperparameters of the hyperparameters, so to speak), so it's no really "each possible choice"
 Colors indicate the different hyper-parameter instantiations.
@@
 Regarding the ranking approach (Fig.~\ref{fig:annitraincolorranking}), using the ranking that is induced by our runtime-label predictions consistently performs better than the partially observed PAR-2 ranking for each possible value of $\delta$.
+%JB: consistent outperformance not guaranteed in case the lines are not individual approaches, but Pareto fronts within the compared categories (which means that different parts of the line might be related to different hyperparameter settings)
 %This was expected since a possible bias in the selection decisions of the model is accounted for when using the model's predictions for ranking.
@@
 
-Regarding the selection approaches (Fig.~\ref{fig:annitraincolorselection}), our model-based uncertainty strategy performs best in most cases.
-However, our model-based information-gain sampling is beneficial if runtime is strongly favored (very small $\delta$; runtime fraction less than \SI{5}{\%}).
-
-Regarding the stopping criterion (Fig.~\ref{fig:annitraincolorstopping}), the ranking convergence criterion performs the most consistently well in most cases.
+Regarding the selection approaches (Fig.~\ref{fig:annitraincolorselection}), the model-based uncertainty strategy performs best in most cases.
+However, the model-based information-gain sampling is beneficial if runtime is strongly favored (very small $\delta$; runtime fraction less than \SI{5}{\%}).
+%JB: we should also describe performance of random sampling
+
+Regarding the stopping criterion (Fig.~\ref{fig:annitraincolorstopping}), the ranking convergence criterion performs the most consistently well.
 If PAR-2 accuracy is strongly favored (very high $\delta$), the Wilcoxon stopping criterion performs better.
@@
     }
-    \caption{$O_{\operatorname{rt}}$-$O_{\operatorname{acc}}$-diagram showing Pareto fronts for configurations with the given ranking approaches.
-The lines show the best performances of approaches using runtime-label-prediction ranking and partially observed PAR-2 ranking respectively.}
+    \caption{
+    	$O_{\operatorname{rt}}$-$O_{\operatorname{acc}}$-diagram showing Pareto fronts for configurations with the given ranking approaches.
+    	The lines show the best performances of approaches using runtime-label-prediction ranking and partially observed PAR-2 ranking respectively.
+    }
     \label{fig:annitraincolorranking}
@@
     }
-    \caption{$O_{\operatorname{rt}}$-$O_{\operatorname{acc}}$-diagram showing Pareto fronts for configurations with the given instance-selection approaches.
-The lines show the best performances of approaches using uncertainty-based, information-gain-based, and random instance-selection respectively.}
+    \caption{
+    	$O_{\operatorname{rt}}$-$O_{\operatorname{acc}}$-diagram showing Pareto fronts for configurations with the given instance-selection approaches.
+    	The lines show the best performances of approaches using uncertainty-based, information-gain-based, and random instance selection respectively.
+    }
     \label{fig:annitraincolorselection}
@@
     }
-    \caption{$O_{\operatorname{rt}}$-$O_{\operatorname{acc}}$-diagram showing Pareto fronts for configurations with the given stopping criteria.
-The lines show the best performances of approaches using the ranking-convergence criterion, Wilcoxon criterion, and a fixed subset size respectively.}
+    \caption{
+    	$O_{\operatorname{rt}}$-$O_{\operatorname{acc}}$-diagram showing Pareto fronts for configurations with the given stopping criteria.
+    	The lines show the best performances of approaches using the ranking-convergence criterion, Wilcoxon criterion, and a fixed subset size respectively.
+    }
     \label{fig:annitraincolorstopping}
@@
 
-  \caption{$O_{\operatorname{rt}}$-$O_{\operatorname{acc}}$-diagrams showing the performance of different hyper-parameter instantiations.
+  \caption{
+  	$O_{\operatorname{rt}}$-$O_{\operatorname{acc}}$-diagrams showing the performance of different hyper-parameter instantiations.
   	The x-axis shows the fraction of runtime in comparison to the time needed to evaluate a solver on all instances.
   	The y-axis shows the accuracy of the predicted rank in comparison to the true PAR-2 rank.
-Each line entails the front of Pareto-optimal configurations with the respective instantiation.}
+  	Each line entails the front of Pareto-optimal configurations with the respective instantiation.
+  }
   %JB: not sure if the color palette in the plots is grayscale-/color-blind-friendly; consider changing the palette (see https://colorbrewer2.org/#type=qualitative&scheme=Set3&n=3 ) and using different line types
@@
 \begin{table}[tb!]
+  %JB: really hard to extract take-aways by looking at this table (several settings and several metrics)
   \centering
-  \caption{Pareto-optimal approaches for different choices of $\delta$ regarding our smaller hyper-parameter tuning dataset.
+  \caption{
+  	Pareto-optimal approaches for different choices of $\delta$ regarding our smaller hyper-parameter tuning dataset.
   	Rows with no changes in comparison to their predecessor are hidden.
+  	%JB: no change in what? all columns except delta?
   	Ranking approaches are either predicted cluster-label ranking (Pred.) or partially observed PAR-2 ranking (PAR-2).
   	Selection approaches are either uncertainty-based selection (Uncert.) or information-gain-based selection (IG).
-Stopping criteria are either ranking convergence (Conv.) or Wilcoxon-based (Wilcoxon).}
+  	Stopping criteria are either ranking convergence (Conv.) or Wilcoxon-based (Wilcoxon).
+  	%JB: How are tthe values for ranking/selection/stopping determined? whatever is best for a particular delta?
+  }
   \label{tab:paretotable}
@@
 Tab.~\ref{tab:paretotable} also shows the performance of particular hyper-parameter choices for different values of $\delta$.
+%JB: some plot for comparison would be better than a truncated table showing a variety hyperparameter configs
 Thereby, we observe that for small $\delta$ (runtime more important) our approaches preferably use easy instances to get a good sense of the PAR-2 ranking fast, e.g., row with $\delta = 0.15$.
+%JB: the "e.g." part needs more explanation
 In contrast to that, harder instances are preferred for higher values of $\delta$, i.e., the sampled fraction of runtime is greater than the fraction of instances.
 In our opinion, the row with $\delta = 0.85$ represents the best trade-off between runtime and PAR-2 accuracy.
+%JB: this is a scientific article, not an opinion piece :-); all deltas represent different trade-offs, so I would not claim to have found a best trade-off (probably depends on user what is most desired)
 
@@
 Moreover, the PAR-2 accuracy is about \SI{92.33}{\%}, the average Spearman correlation between the PAR-2 ranking and the predicted ranking scores about \SI{0.9572}{}, and the weighted objective $O_{\delta} = 0.9193$.
+%JB: many numbers, but some qualitative (and positively sounding, i.e., saying that our approach indeed is useful) interpretation would be nice
 
@@
 Thereby, the average benchmark set contains about \SI{192.34}{} instances.
-Tab.~\ref{tab:familiesa} lists the ten most frequent families that are chosen. For example, \textit{planning} instances account for on average \SI{19.14}{} of the \SI{192.34}{} benchmark instances.
-In contrast to that, Tab.~\ref{tab:familiesb} shows the families that are sampled the most often relative to their class sizes in the underlying dataset.
+Tab.~\ref{tab:familiesa} lists the ten most frequent families that are chosen.
+For example, \textit{planning} instances account for on average \SI{19.14}{} of the \SI{192.34}{} benchmark instances.
+In contrast to that, Tab.~\ref{tab:familiesb} shows the families that are sampled the most often relative to their occurrences in the underlying dataset.
 There are, for example, only nine \textit{circuit-multiplier} instances in the complete dataset.
 However, we sample on average \SI{1.59}{} of them.
+%JB: there should be some (qualitative) interpretation (based on our results, is the sample interesting family-wise or not?) rather than just dropping a table and explaining its structure
 
 \begin{table}[tb!]
+  %JB: might be interesting to have scatterplot with fraction in sample / fraction in dataset (does not allow to show all family labels, but allows to show all families)
   \centering
   \caption{Top-10 most \textit{important} families regarding the average occurrences and the fraction of selected occurrences regarding our AL approach}
+  %JB: also interesting: distribution of families in sample vs. full dataset (can even correlate these two vectors); "fraction" here only draws family-specific comparison of sample to full dataset
   \label{tab:families}
@@
     \centering
-    \caption{Sorted by the average number of occurrences}
+    \caption{Sorted by the average number of occurrences in the sample}
     \label{tab:familiesa}
@@
       \hline
-      Family & Avg. Occ. & In Dataset & Fraction (\%) \\
+      Family & In Sample & In Dataset & Fraction (\%) \\
       \hline
@@
 \section{Conclusion}
-In this work, we have discussed possible solutions to the \textit{new-solver} problem:
+In this work, we have discussed possible solutions to the \textit{New-Solver Problem}:
 Given a new solver, we want to find its ranking amidst its competitors.
@@
 A model-uncertainty-based selection approach in combination with runtime-label-prediction ranking and a ranking-convergence stopping criterion showed the consistently best results.
+%JB: "consistently best" -- it's the best approach for one delta, but not sure regarding other deltas nor its sensitivity to hyperparameters
 We also took a brief look at which instance families are the most prevalent when sampling instances.
@@
 Our current approach selects instances one at a time.
-Running benchmarks on a cluster with $n$ cores requires a batch of $n$ instances at a time.
+Running benchmarks on a computing cluster with $n$ cores benefits from having a batch of $n$ instances at a time.
 This is, however, not trivial since, on the one hand, a higher $n$ leads to less \textit{active learning} (because of bigger batch sizes) and, on the other hand, it is not clear how to synchronize the model update and instance selection without a barrier lock, which wastes a lot of runtime resources.
@@
 On a more general note, it would make sense to generalize this evaluation framework for arbitrary $\mathcal{NP}$-complete problems.
-Those problems share most of the relevant properties, i.e., there are established problem-instance features, a full benchmark run takes days, and creating new problem solvers traditionally requires \textit{expert} knowledge to hand-select instances of interest.
+%JB: is there methodological generalization needed or can't we just apply the same framework (instantiating the instance-feature part differently) to these use cases?
+Those problems share most of the relevant properties of SAT solving, i.e., there are established problem-instance features, a full benchmark run takes days, and creating new problem solvers traditionally requires expert knowledge to hand-select instances of interest.
 

### 2022-09-27 paper/paper.tex
Replace \textit with \emph
... for highlighting.

@@
   Benchmark-set selection is the process of choosing a subset of instances out of a pool to be representative and reliable in discriminating SAT solvers.
-  This paper presents an active-learning-based benchmark-set-selec\-tion approach that solves the \textit{New-Solver Problem}:
+  This paper presents an active-learning-based benchmark-set-selec\-tion approach that solves the \emph{New-Solver Problem}:
   Given a new solver, we want to find its ranking amidst its competitors with as little runtime as possible while still giving accurate ranking predictions.
@@
 Our algorithm is unique in that it is the only one, to the best of our knowledge, that has all of the following features (also refer to Tab.~\ref{tab:requirements}).
-Our benchmarking process outputs a \textit{ranking} to compare the new solver against the existing set of solvers as described above.
-We optimize the \textit{runtime} that our strategy needs to arrive at its conclusion.
-We use problem-instance and runtime \textit{features} $\varphi\!: \mathcal{A} \times \mathcal{I} \rightarrow \mathbb{R}^n$ since such features already have proven to be helpful in classic benchmark-set-selection approaches~\cite{HoosKSS13} as well as in portfolio selection~\cite{CollauttiMMO13,NgokoCT19}.
+Our benchmarking process outputs a \emph{ranking} to compare the new solver against the existing set of solvers as described above.
+We optimize the \emph{runtime} that our strategy needs to arrive at its conclusion.
+We use problem-instance and runtime \emph{features} $\varphi\!: \mathcal{A} \times \mathcal{I} \rightarrow \mathbb{R}^n$ since such features already have proven to be helpful in classic benchmark-set-selection approaches~\cite{HoosKSS13} as well as in portfolio selection~\cite{CollauttiMMO13,NgokoCT19}.
 %JB: I think we can remove this criterion, as it also is satisified by all competitors (and is commented out in table; if we keep criterion, we should also show it in table)
-Moreover, we select instances \textit{non-randomly} and \textit{incrementally}.
+Moreover, we select instances \emph{non-randomly} and \emph{incrementally}.
 By doing so, we have a greater impact on the properties of the benchmarking approach such as its required runtime.
 Sampling instances incrementally means taking runtime information from already done experiments into account when choosing the next.
-Rather than solely outputting a binary classification, i.e, the new solver is worse than solver A, we provide a \textit{scoring function} that shows by which margin a solver is worse and how similar it is to existing approaches (\textit{interpretability}).
+Rather than solely outputting a binary classification, i.e, the new solver is worse than solver A, we provide a \emph{scoring function} that shows by which margin a solver is worse and how similar it is to existing approaches (\emph{interpretability}).
 %JB: we should rename "interpretability" criterion, term can have lots of different meanings and our specific meaning is not very intuitive
-Finally, our strategy's cost \textit{scales linearly} with the cardinality of the set of existing solvers $\mathcal{A}$ to compare against and the instances $\mathcal{I}$ within our pool of possible benchmark instances.
+Finally, our strategy's cost \emph{scales linearly} with the cardinality of the set of existing solvers $\mathcal{A}$ to compare against and the instances $\mathcal{I}$ within our pool of possible benchmark instances.
 We avoid comparing pairs of solvers.
@@
 Third, we discuss an existing benchmark-set-selection approach that optimizes for benchmark runtime.
-Finally, we show how active-learning techniques can help to tackle the aforementioned \textit{New-Solver Problem}.
+Finally, we show how active-learning techniques can help to tackle the aforementioned \emph{New-Solver Problem}.
 
@@
 %JB: we should probably cut the discussion of each approach's criteria to save space, maybe only mention the unsatisifed ones
-Their approach \textit{ranks} solvers by providing its PAR-2 score, is \textit{feature-based}, and is \textit{scalable} in the sense that the evaluation time scales linearly with more instances and solvers.
-Also, their approach is \textit{interpretable} in the sense that the obtained PAR-2 score induces a distance between solvers.
-However, they do not optimize for benchmark \textit{runtime} and select instances \textit{randomly}, apart from constraints on the instance hardness and feature distance.
+Their approach \emph{ranks} solvers by providing its PAR-2 score, is \emph{feature-based}, and is \emph{scalable} in the sense that the evaluation time scales linearly with more instances and solvers.
+Also, their approach is \emph{interpretable} in the sense that the obtained PAR-2 score induces a distance between solvers.
+However, they do not optimize for benchmark \emph{runtime} and select instances \emph{randomly}, apart from constraints on the instance hardness and feature distance.
 
@@
 
-While this approach considers the configuration's \textit{runtime}, is \textit{feature-based}, and \textit{scales linearly} by having a fixed time budget, it is not capable of the other three features listed in Tab.~\ref{tab:requirements}.
-First, an algorithm configurator cannot be used to \textit{rank} a new solver since it only looks at promising solvers/configurations rather than the overall average performance.
+While this approach considers the configuration's \emph{runtime}, is \emph{feature-based}, and \emph{scales linearly} by having a fixed time budget, it is not capable of the other three features listed in Tab.~\ref{tab:requirements}.
+First, an algorithm configurator cannot be used to \emph{rank} a new solver since it only looks at promising solvers/configurations rather than the overall average performance.
 If the new solver does not belong to this category (new ideas are seldomly the best approach right from the start, i.e., they need some tweaking), we do not have any information on how this solver performed.
 %JB: "this category" = category of most promising solvers? not 100% clear
-Second, while using a model-based selection strategy to sample configurations, instance sampling is done \textit{randomly}, i.e., without building a model over instances.
-And third, their approach is not \textit{interpretable} regarding the new solver performance since we are already missing the feature to \textit{rank} the new solver among existing ones.
+Second, while using a model-based selection strategy to sample configurations, instance sampling is done \emph{randomly}, i.e., without building a model over instances.
+And third, their approach is not \emph{interpretable} regarding the new solver performance since we are already missing the feature to \emph{rank} the new solver among existing ones.
 %JB: well, the internal meta-model of the configurator allows to estimate the solver performance for particular configurations (instead of comparing solvers), which allows a specific performance assessment, so the approach is interpretable in that direction (though for comparison to other solvers, only top performance should be interesting)
@@
 
-Regarding our desired features of adaptive benchmark-set-selection (Tab.~\ref{tab:requirements}), the aforementioned approach already ticks off the first four features, i.e., it \textit{ranks} solvers, optimizes for \textit{runtime}, is \textit{feature-based}, and uses \textit{incremental non-random} sampling.
-Although they provide a confidence metric for their prediction, which makes this approach partially \textit{interpretable}, it is not clear how similar two given solvers are or on which instances they behave similarly, as the prediction is solely binary.
-Moreover, the major shortcoming lies within it being not \textit{scalable} with the number solvers.
+Regarding our desired features of adaptive benchmark-set-selection (Tab.~\ref{tab:requirements}), the aforementioned approach already ticks off the first four features, i.e., it \emph{ranks} solvers, optimizes for \emph{runtime}, is \emph{feature-based}, and uses \emph{incremental non-random} sampling.
+Although they provide a confidence metric for their prediction, which makes this approach partially \emph{interpretable}, it is not clear how similar two given solvers are or on which instances they behave similarly, as the prediction is solely binary.
+Moreover, the major shortcoming lies within it being not \emph{scalable} with the number solvers.
 As they compare only pairs of solvers, scaling with more and more solvers and ideas becomes an issue as there is a quadratic amount of pairs to distinguish.
@@
 %JB: section does not make clear which of the following approaches we will use (reads more like fundamentals than related work)
-The posed \textit{New-Solver Problem} has stark similarities to the well-studied field of active learning within recommender systems, especially the \textit{new user problem}~\cite{RubensESK15}.
+The posed \emph{New-Solver Problem} has stark similarities to the well-studied field of active learning within recommender systems, especially the \emph{new-user problem}~\cite{RubensESK15}.
 The classic exploration-exploitation dilemma arises since, on the one hand, we want to maximize the utility an instance provides to our model and, on the other hand, minimize the cost (CPU time) that is associated with its acquisition.
@@
 
-AL algorithms can be categorized into \textit{synthesis-bas\-ed} \cite{0001AEMN22,GarzonMG22,2019gaal} and \textit{pool-bas\-ed} approaches \cite{GolbandiKL11,distribAL,HarpaleY08,KapoorGUD07,KorenBV09,KornerW06,MelvilleM04,SinhaED19,TongK01}.
+AL algorithms can be categorized into \emph{synthesis-bas\-ed} \cite{0001AEMN22,GarzonMG22,2019gaal} and \emph{pool-bas\-ed} approaches \cite{GolbandiKL11,distribAL,HarpaleY08,KapoorGUD07,KorenBV09,KornerW06,MelvilleM04,SinhaED19,TongK01}.
 %JB: we can probably cut some citations, the three most important should be sufficient
@@
 
-Pool-based methods can be further refined into \textit{Error-based} \cite{GolbandiKL11,KorenBV09}, \textit{Un\-cer\-tain\-ty-based} \cite{HarpaleY08,KapoorGUD07,TongK01}, \textit{Distribution-based} methods \cite{distribAL,SinhaED19} and \textit{Ensembles} \cite{KornerW06,MelvilleM04}.
+Pool-based methods can be further refined into \emph{Error-based} \cite{GolbandiKL11,KorenBV09}, \emph{Un\-cer\-tain\-ty-based} \cite{HarpaleY08,KapoorGUD07,TongK01}, \emph{Distribution-based} methods \cite{distribAL,SinhaED19} and \emph{Ensembles} \cite{KornerW06,MelvilleM04}.
 Error-based strategies pick instances that have the most influence on the model parameters.
@@
 Moreover, we have briefly outlined how active-learning techniques can help to incrementally select benchmark sets.
-In the following sections, we present an active-learning-based benchmark-set-selection approach that solves the \textit{New-Solver Problem} while optimizing for runtime and being both interpretable and scalable.
+In the following sections, we present an active-learning-based benchmark-set-selection approach that solves the \emph{New-Solver Problem} while optimizing for runtime and being both interpretable and scalable.
 
@@
 %JB: as it is rather simple and has different parts to be plugged in, I think we should rebrand it as "framework" rater than algorithm
-After initializing the model $\mathcal{M}$, we incrementally \textit{select} a problem instance from the pool of instances, one at a time, evaluate the new solver $\hat{a}$ on it, and \textit{update} the model with the acquired result.
-We \textit{stop} when meeting some stopping condition.
+After initializing the model $\mathcal{M}$, we incrementally \emph{select} a problem instance from the pool of instances, one at a time, evaluate the new solver $\hat{a}$ on it, and \emph{update} the model with the acquired result.
+We \emph{stop} when meeting some stopping condition.
 Since the potential runtime of experiments is by magnitudes larger than the model's update time, we only consider incrementing our benchmark set by one instance at a time rather than using batches.
@@
 \end{equation}
-with $w_{a,i} := 2$ (PAR-\textit{2} penalty factor) if $\gamma\!\left(a, i\right) = k$ else $w_{a,i} := 1$.
+with $w_{a,i} := 2$ (PAR-\emph{2} penalty factor) if $\gamma\!\left(a, i\right) = k$ else $w_{a,i} := 1$.
 To be applicable to $\hat{a}$, we extend $\operatorname{score}$ by replacing $\gamma\!\left(a, i\right)$ by the model's predictions $f\!: \left(\mathcal{A} \cup \left\lbrace \hat{a} \right\rbrace\right) \times \mathcal{I} \rightarrow \left\lbrace 1, \dots, k \right\rbrace$.
@@
 This means that we do not lose much discriminatory power using discrete runtime labels.
-The number of significant comparisons between solvers \textit{alone}, though, does only offer limited insights as we might achieve significant differences in all pairs by assigning some randomly skewed values as labels.
+The number of significant comparisons between solvers \emph{alone}, though, does only offer limited insights as we might achieve significant differences in all pairs by assigning some randomly skewed values as labels.
 %JB: thus, we should reorder section to mention significant difference (this paragraph) first and deciding solver pairs (previous paragraph, stronger argument) second (rather than giving the stronger argument first, making the second argument superfluous)
@@
 Our instance-sampling strategies (also refer to Alg.~\ref{algALBenchmark}; line 3) can be divided into two categories.
-First, we present possible baseline approaches including existing work on how to pick \textit{interesting} instances~\cite{Bossek021a,GentHJKMNN14,MatriconAFSH21}.
+First, we present possible baseline approaches including existing work on how to pick \emph{interesting} instances~\cite{Bossek021a,GentHJKMNN14,MatriconAFSH21}.
 %JB: approaches under "Second" are also at least "inspired by existing work" and probably also pick "interesting" instances, distinction between the two categories not entirely clear
@@
 
-Bossek and Wagner~\cite{Bossek021a} introduce a custom ranking-score metric incorporating the number of \textit{good} pairs and \textit{bad} pairs.
+Bossek and Wagner~\cite{Bossek021a} introduce a custom ranking-score metric incorporating the number of \emph{good} pairs and \emph{bad} pairs.
 %JB: some selection methods, like this one, are not mentioned later any more (in particular, not in experimental design), so we don't need to describe them in detail (maybe mention shortly if used in preliminary experiments, but turned out to be not useful)
 A good solver pair, regarding a particular instance, are solvers that perform according to their global ranking.
-So if solver $a$ is better than solver $b$ on instance $i$ and $a$ is also in total better than $b$, it is a \textit{good} solver pair on $i$.
-Otherwise, it is a \textit{bad} pair.
+So if solver $a$ is better than solver $b$ on instance $i$ and $a$ is also in total better than $b$, it is a \emph{good} solver pair on $i$.
+Otherwise, it is a \emph{bad} pair.
 Based on this classification, they define a 3-tuple for each instance consisting of the number of good pairs, the accumulated performance differences of bad pairs, and the accumulated performance differences of good pairs.
@@
 (4) mean that we would prefer instances with low entropy of runtime distribution after considering the new solver, i.e., instances that are not very discriminatory?
-Thereby, \textit{wilcoxon} denotes the $p$-value of a Wilcoxon signed-rank test of the runtime-label distributions and $\gamma\!\left(a, \mathcal{I}\right)$ and $f\!\left(\hat{a}, \mathcal{I}\right)$ are the expected set extensions on $\gamma$ and $f$ respectively.
+Thereby, \emph{wilcoxon} denotes the $p$-value of a Wilcoxon signed-rank test of the runtime-label distributions and $\gamma\!\left(a, \mathcal{I}\right)$ and $f\!\left(\hat{a}, \mathcal{I}\right)$ are the expected set extensions on $\gamma$ and $f$ respectively.
 %JB: "set extension" might be rather technical and hard to understand; also, why "expected"? in the sense of "straightforward"?
@@
 \label{sec:goal}
-As already stated in the introductory section, this work addresses the \textit{New-Solver Problem} (cf.~Def.~\ref{def:new-solver-problem}).
-By running Alg.~\ref{algALBenchmark}, we obtain a prediction model $\mathcal{M}$ that provides us with a scoring function \textit{score}.
+As already stated in the introductory section, this work addresses the \emph{New-Solver Problem} (cf.~Def.~\ref{def:new-solver-problem}).
+By running Alg.~\ref{algALBenchmark}, we obtain a prediction model $\mathcal{M}$ that provides us with a scoring function \emph{score}.
 Also, it outputs the fraction of runtime that is needed to arrive at its conclusion.
 First and foremost, our goal is to provide the engineer of new SAT solvers with an accurate ranking.
-We define the \textit{ranking accuracy} $O_{\operatorname{acc}} \in \left[0, 1\right]$ (higher is better) by the fraction of pairs $\left(\hat{a}, a\right)$ for all $a \in \mathcal{A}$ that are decided correctly by the given ranking.
+We define the \emph{ranking accuracy} $O_{\operatorname{acc}} \in \left[0, 1\right]$ (higher is better) by the fraction of pairs $\left(\hat{a}, a\right)$ for all $a \in \mathcal{A}$ that are decided correctly by the given ranking.
 For now, we exclude the equality of solvers since data shows that the PAR-2 scores of all our solvers are different.
@@
 Second, we also have to optimize for runtime.
-The \textit{fraction of runtime} that the algorithm needs to arrive at its conclusion is denoted by $O_{\operatorname{rt}} \in \left[0, 1\right]$ (lower is better).
+The \emph{fraction of runtime} that the algorithm needs to arrive at its conclusion is denoted by $O_{\operatorname{rt}} \in \left[0, 1\right]$ (lower is better).
 This metric puts the runtime summed over the sampled instances in relation to the runtime summed over all instances in the dataset.
@@
 Alg.~\ref{alg:eval} shows this.
-Note that the \textit{new} solver in each iteration is excluded from the set of solvers $\mathcal{A}$ to avoid data leakage.
+Note that the \emph{new} solver in each iteration is excluded from the set of solvers $\mathcal{A}$ to avoid data leakage.
 Also, the runtime function $r$ is restricted to $\mathcal{A} \times \mathcal{I}$.
@@
 We use the GBD metadata database~\cite{IserS18} to provide us with the aforementioned instance features, problem instances, and solver runtimes.
-Thereby, all our approaches make use of the 56 base features that are maintained in the \textit{base} database of GBD~\cite{IserS18}.
+Thereby, all our approaches make use of the 56 base features that are maintained in the \emph{base} database of GBD~\cite{IserS18}.
 They are inspired by the SATzilla features~\cite{features}.
@@
 For hyper-parameter tuning, we randomly sample \SI{10}{\%} of the complete set of 5355 instances with stratification regarding the instance's family.
-All instance families that are too \textit{small}, i.e., \SI{10}{\%} of them corresponds to less than one instance, are put into one meta-family for stratification.
+All instance families that are too \emph{small}, i.e., \SI{10}{\%} of them corresponds to less than one instance, are put into one meta-family for stratification.
 This smaller dataset allows for a more extensive exploration of hyper-parameter space.
@@
 \label{sec:hyper}
-Given Alg.~\ref{algALBenchmark}, there are several possible instantiations for the three phases, i.e., \textit{selection}, \textit{stopping}, and \textit{ranking}.
+Given Alg.~\ref{algALBenchmark}, there are several possible instantiations for the three phases, i.e., \emph{selection}, \emph{stopping}, and \emph{ranking}.
 Also, there are different choices for the runtime-label prediction model.
@@
 \subsubsection{Ranking.}
-Regarding \textit{ranking} (cf. Sec.~\ref{sec:main1}), we experiment with the following approaches, including our used hyper-parameter values:
+Regarding \emph{ranking} (cf. Sec.~\ref{sec:main1}), we experiment with the following approaches, including our used hyper-parameter values:
 %JB: as these approaches are mostly new (Section 3.1 only explains prediction-based ranking in general, neither observed ranking nor history nor fallback), I'm not sure if we should provide further motivation/explanation here); also applies to other components of experimental design
@@
 \subsubsection{Selection.}
-For \textit{selection} (cf. Sec.~\ref{sec:sampling1}), we experiment with the following methods, including our used hyper-parameter values:
+For \emph{selection} (cf. Sec.~\ref{sec:sampling1}), we experiment with the following methods, including our used hyper-parameter values:
 %JB: sampling section describes more methods than the ones listed below, I guess some got dropped after preliminary experiments
@@
 \subsubsection{Stopping.}
-For \textit{stopping} decisions (cf. Sec.~\ref{sec:main3}), we experiment with the following criteria, including our used hyper-parameter values:
+For \emph{stopping} decisions (cf. Sec.~\ref{sec:main3}), we experiment with the following criteria, including our used hyper-parameter values:
 \begin{itemize}
@@
 As a runtime-label prediction model, we only make use of one method that has proven to be successful in preliminary experiments (cf.~Sec.~\ref{sec:evalprel}).
-Our model of choice is an ensemble, stacking a quadratic-discriminant analysis\footnote{With a regularization parameter of zero and a singular value threshold of zero} onto a random forest model\footnote{With \textit{entropy} splitting criterion and \textit{balanced} class weights} using a decision tree\footnote{With \textit{GINI} impurity splitting criterion; choosing the best split for each branching decision and maximum tree depth of 5 levels} to weight its members.
-If not stated explicitly, all other model parameters default to the presets of \textit{scikit-learn}\footnote{\url{https://scikit-learn.org/stable/index.html}; Version 1.0.2}~\cite{scikit-learn}.
+Our model of choice is an ensemble, stacking a quadratic-discriminant analysis\footnote{With a regularization parameter of zero and a singular value threshold of zero} onto a random forest model\footnote{With \emph{entropy} splitting criterion and \emph{balanced} class weights} using a decision tree\footnote{With \emph{Gini} impurity splitting criterion; choosing the best split for each branching decision and maximum tree depth of 5 levels} to weight its members.
+If not stated explicitly, all other model parameters default to the presets of \emph{scikit-learn}\footnote{\url{https://scikit-learn.org/stable/index.html}; Version 1.0.2}~\cite{scikit-learn}.
 
@@
 For reproducibility, our source code, all experiments, and data are available on GitHub\footnote{\url{https://github.com/mathefuchs/al-for-sat-solver-benchmarking}}.
-Our code is implemented in \textsc{Python} using \textit{scikit-learn}~\cite{scikit-learn} for making predictions and \textit{gbd-tools}~\cite{IserS18} for SAT-instance retrieval.
+Our code is implemented in \textsc{Python} using \emph{scikit-learn}~\cite{scikit-learn} for making predictions and \emph{gbd-tools}~\cite{IserS18} for SAT-instance retrieval.
 
@@
 We are, therefore, narrowing down the search by filtering based on preliminary results.
-Since all relevant decisions, i.e., \textit{selection}, \textit{stopping}, and \textit{ranking}, revolve around the choice of prediction model, we are first looking at the \textit{Matthews Correlation Coefficient} (MCC) scores~\cite{gorodkin2004comparing,matthews1975comparison} that a particular prediction model achieves.
+Since all relevant decisions, i.e., \emph{selection}, \emph{stopping}, and \emph{ranking}, revolve around the choice of prediction model, we are first looking at the \emph{Matthews Correlation Coefficient} (MCC) scores~\cite{gorodkin2004comparing,matthews1975comparison} that a particular prediction model achieves.
 MCC scores are great in dealing with class imbalances in contrast to conventional metrics like accuracy.
@@
 Tab.~\ref{tab:familiesa} lists the ten most frequent families that are chosen.
-For example, \textit{planning} instances account for on average \SI{19.14}{} of the \SI{192.34}{} benchmark instances.
+For example, \emph{planning} instances account for on average \SI{19.14}{} of the \SI{192.34}{} benchmark instances.
 In contrast to that, Tab.~\ref{tab:familiesb} shows the families that are sampled the most often relative to their occurrences in the underlying dataset.
-There are, for example, only nine \textit{circuit-multiplier} instances in the complete dataset.
+There are, for example, only nine \emph{circuit-multiplier} instances in the complete dataset.
 However, we sample on average \SI{1.59}{} of them.
@@
   \centering
-  \caption{Top-10 most \textit{important} families regarding the average occurrences and the fraction of selected occurrences regarding our AL approach}
+  \caption{Top-10 most \emph{important} families regarding the average occurrences and the fraction of selected occurrences regarding our AL approach}
   %JB: also interesting: distribution of families in sample vs. full dataset (can even correlate these two vectors); "fraction" here only draws family-specific comparison of sample to full dataset
@@
 \section{Conclusion}
-In this work, we have discussed possible solutions to the \textit{New-Solver Problem}:
+In this work, we have discussed possible solutions to the \emph{New-Solver Problem}:
 Given a new solver, we want to find its ranking amidst its competitors.
@@
 Running benchmarks on a computing cluster with $n$ cores benefits from having a batch of $n$ instances at a time.
-This is, however, not trivial since, on the one hand, a higher $n$ leads to less \textit{active learning} (because of bigger batch sizes) and, on the other hand, it is not clear how to synchronize the model update and instance selection without a barrier lock, which wastes a lot of runtime resources.
+This is, however, not trivial since, on the one hand, a higher $n$ leads to less \emph{active learning} (because of bigger batch sizes) and, on the other hand, it is not clear how to synchronize the model update and instance selection without a barrier lock, which wastes a lot of runtime resources.
 

### 2022-10-04 paper/paper.tex
Add more details to some comments in paper
Also, make a few suggestions for further related work.
Add some related work
... in all four subsections of that section. Not many relevant works found.

@@
 
+Note that we know the runtimes of all solvers, except the new one, on all instances.
+There also is work on collaborative filtering, where runtimes are only partially known, for algorithm selection~\cite{misir2017alors,misir2017data}.
 We want to find a scoring function $\operatorname{score}\!: \mathcal{A} \cup \left\lbrace \hat{a} \right\rbrace \rightarrow \mathbb{R}$ assigning each solver a score.
@@
 This allows for both comparing the solvers as well as identifying solvers with similar behavior.
-%JB: is there any further behavior we can deduce about a solver scoring than a performance ranking?
-similar behavior could be something like two solvers being fast/slow on the same instances, but that cannot be deduced from overall score
+%JB: is there any further behavior we can deduce about a solver scoring apart from a performance ranking?
+"similar behavior" could IMHO be something like two solvers being fast/slow on the same instances, but that cannot be deduced from overall score
 
@@
 Rather than solely outputting a binary classification, i.e, the new solver is worse than solver A, we provide a \emph{scoring function} that shows by which margin a solver is worse and how similar it is to existing approaches (\emph{interpretability}).
-%JB: we should rename "interpretability" criterion, term can have lots of different meanings and our specific meaning is not very intuitive
+%JB: we should rename "interpretability" criterion (make it more specific), term "interpretability" can have lots of different meanings and our specific meaning is not very intuitive
 Finally, our strategy's cost \emph{scales linearly} with the cardinality of the set of existing solvers $\mathcal{A}$ to compare against and the instances $\mathcal{I}$ within our pool of possible benchmark instances.
-We avoid comparing pairs of solvers.
-This would scale quadratically.
-%JB: need to discuss this; log-linear effort might be sufficient if binary-search-/recursive-sorting-like procedure used
+%JB: for what do we measure cost here?
+% (1) solving new-solver-problem once?
+-> apart from more runtimes to be acquired beforehand, is running our framework affected by increasing the set of existing solvers, and how does this depend on the configuration of our framework, i.e., sampling/selection/stopping?
+(e.g., Wilcoxon stopping criterion iterates over set of existing solvers, showing linear complexity, while fixed-subset stopping is independent from number of existing solvers)
+% (2) solving new-solver-problem for each solver out of a set of solvers?
+-> is a rather theoretical scenario, because our algorithm itself assumes we know runtimes of all but one solver, so in practice we cannot throw in a bunch of new solvers and get a ranking over all of them
+We avoid comparing pairs of solvers, which would scale quadratically.
+%JB: need to discuss this; logarithmic (compare one-vs rest) / log-linear (compare each pair) effort might be sufficient if binary-search-/recursive-sorting-like procedure used
+Since a new solver idea is rarely best at first try but requires more or less tweaking, it is desired to obtain a solver ranking fast.
+In this way, a new solver idea can be discarded if it performs poorly across the board or may be further tweaked if it shows promising results at least in some cases.
 
@@
 Tab.~\ref{tab:requirements} shows a concise overview of existing approaches with a related goal.
-%JB: we can probably cut this part a bit, as we discuss weakness of individual approaches later anyway
+%JB: we can probably cut this part a bit, as we discuss weaknesses of individual approaches later anyway
 While traditional benchmark-set-selection~\cite{HoosKSS13} lacks the aspect of runtime optimization, related work within algorithm configuration~\cite{HutterHL11} is concerned with finding better configurations fast.
@@
 \subsubsection{Traditional benchmark-set-selection.}
-%JB: I think we should do some further research on static benchmark-set selection approaches (maybe in SMT, CP, etc.)
 Recent studies~\cite{abs-2107-07002,NiesslHWCB22} show that selecting a benchmark set in general is a difficult endeavor for multiple reasons.
@@
 
-Hoos et. al.~\cite{HoosKSS13} discuss which properties are most desirable when selecting a set of SAT benchmark instances, for instance for a competition.
-The selection strategy presented is static, i.e., it does not depend on the solvers to distinguish.
+Benchmark-set selection is an important issue for solver competitions, e.g., the SAT Competition.
+In such competitions, the organizers define the rules for composing the corresponding benchmarks.
+Balint et al.~\cite{balint2015overview} provide an overview of benchmark-selection criteria in different solver competitions.
+%JB: not sure if we should provide some details regarding SAT Competition 2022 \cite{sat2022}, previous competitions (if different procedure used; Related-Work section of Matricon et al.
+(2021) states that) or other competitions (e.g., those in https://www.floc2022.org/floc-olympic-games)
+Manthey and Möhle~\cite{manthey2016better} find that competition benchmarks might contain redundant instances, and propose a feature-based approach to remove redundancy.
+M{\i}s{\i}r~\cite{misir2021benchmark} presents a feature-based approach to reduce benchmark sets by matrix factorization and clustering.
+
+Hoos et. al.~\cite{HoosKSS13} discuss which properties are most desirable when selecting a set of SAT benchmark instances.
+The selection strategy presented is static, i.e., it does not depend on particular solvers to distinguish.
 Desirable properties are a variety of instances (avoids over-fitting of solver approaches), adapted instance hardness (not too easy but also not too hard; solvers cannot be distinguished if too many time-out or finish immediately), and no duplicates (having duplicates gives those instances more weight).
@@
 Those features comprise general instance-size features and graph-representation features among others.
-%JB: we should probably cut the discussion of each approach's criteria to save space, maybe only mention the unsatisifed ones
+%JB: we should probably cut the discussion of each approach's criteria to save space, maybe only mention the unsatisifed criteria
 Their approach \emph{ranks} solvers by providing its PAR-2 score, is \emph{feature-based}, and is \emph{scalable} in the sense that the evaluation time scales linearly with more instances and solvers.
@@
     %Feature-based & \cmark & \cmark & \cmark & \cmark \\
-    Incremental non-random sampling & \xmark & \xmark & \cmark & \cmark \\
+    Incremental non-random\\ sampling & \xmark & \xmark & \cmark & \cmark \\
     % \tablefootnote{While using a model-based selection strategy to select the next configuration to evaluate, the instance sampling is done randomly.}
@@
 First, an algorithm configurator cannot be used to \emph{rank} a new solver since it only looks at promising solvers/configurations rather than the overall average performance.
-If the new solver does not belong to this category (new ideas are seldomly the best approach right from the start, i.e., they need some tweaking), we do not have any information on how this solver performed.
+If the new solver does not belong to this category (new ideas are rarely the best approach right from the start, i.e., they need some tweaking), we do not have any information on how this solver performed.
 %JB: "this category" = category of most promising solvers? not 100% clear
 Second, while using a model-based selection strategy to sample configurations, instance sampling is done \emph{randomly}, i.e., without building a model over instances.
+In fact, tuning algorithms to a small or non-homogeneous set of instances might lead to over-tuning, i.e., lack of generalization~\cite{eggensperger2019pitfalls}.
 And third, their approach is not \emph{interpretable} regarding the new solver performance since we are already missing the feature to \emph{rank} the new solver among existing ones.
-%JB: well, the internal meta-model of the configurator allows to estimate the solver performance for particular configurations (instead of comparing solvers), which allows a specific performance assessment, so the approach is interpretable in that direction (though for comparison to other solvers, only top performance should be interesting)
+%JB: well, the internal meta-model of the configurator allows to estimate the solver performance for particular configurations (rather than solvers), so the approach is kind of interpretable, though regarding configurations (while for comparison to other solvers, only top-configuration performance should be interesting, so the approach is indeed not interpretable regarding solvers)
 
@@
 Matricon et. al.~\cite{MatriconAFSH21} present an incremental benchmark-set-selection approach.
+Their \emph{per-set efficient algorithm selection problem} (PSEAS) is similar to our \emph{New-Solver Problem} (cf.~Definition~\ref{def:new-solver-problem}).
+The main difference is that Matricon et. al. compare a solver to another one, while we compare to a set of existing solvers.
+To the best of our knowledge, we are the first to tackle the latter problem.
 %JB: again, I think there are too many details for describing this single approach
-Given a pair of SAT solvers, they want to iteratively select a subset of instances to decide with high confidence which of the two solvers is better.
+
+Given a pair of SAT solvers, Matricon et. al. want to iteratively select a subset of instances to decide with high confidence which of the two solvers is better.
 The selection of instances is dependent on the choice of the solvers to distinguish.
@@
 As they compare only pairs of solvers, scaling with more and more solvers and ideas becomes an issue as there is a quadratic amount of pairs to distinguish.
-%JB: maybe rather log-linear than quadratic, see a prior comment
-Since a new solver idea is seldomly best at first try but requires more or less tweaking, it is desired to rank a new solver in a way that scales linearly with the number of existing approaches.
-%JB: this general motivation should go there where linearity is introduced as criterion, not here
-In this way, a new solver idea can be discarded if it performs poorly across the board or may be further tweaked if it shows promising results at least in some cases.
+%JB: maybe rather log-linear than quadratic, see a prior comment in intro
 
 \subsubsection{Active learning (AL).}
-%JB: section does not make clear which of the following approaches we will use (reads more like fundamentals than related work)
+%JB: section does not make clear which of the following approaches we will use, i.e., relationship to our work not clear (reads more like fundamentals than related work)
 The posed \emph{New-Solver Problem} has stark similarities to the well-studied field of active learning within recommender systems, especially the \emph{new-user problem}~\cite{RubensESK15}.
 The classic exploration-exploitation dilemma arises since, on the one hand, we want to maximize the utility an instance provides to our model and, on the other hand, minimize the cost (CPU time) that is associated with its acquisition.
-%JB: no sure it is arises here as well; dilemma is typical for searching one optimal solution in continuous space; however, if we find an interesting/useful instance here, I'm not sure if we would continue picking similar instances ("exploting" instead of "exploring") just because we expect them to be useful as well
+%JB: no sure if this dilemma arises here as well; dilemma is typical for searching one optimal solution in continuous space, where there is an incentive to exploit neighborhood in promising regions; however, if we find an interesting/useful (= good for assessing solver performance) instance in our approach, I'm not sure if we would continue picking similar instances ("exploiting" instead of "exploring") just because we expect them to be useful as well
 
@@
 Recent synthesis-based methods within the field of satisfiability solving~\cite{0001AEMN22,GarzonMG22} show how to generate problem instances with desired properties.
-%JB: we should conduct some further research if there are other approaches using AL for SAT (or SMT, CP, etc.) benchmarking
 This goal is, however, orthogonal to ours.
@@
 
+In the field of SAT-solving, Volpato and Guangyan~\cite{volpato2019active} use uncertainty-sampling-based AL to learn an instance-specific algorithm selector.
+Rather than benchmarking a solver's overall performance, the goal is to recommend the best solver out of a set of solvers for each SAT instance.
+
 \subsubsection{Summary.}
@@
 Our algorithm follows a simple three-step schema (Alg.~\ref{algALBenchmark}).
-%JB: as it is rather simple and has different parts to be plugged in, I think we should rebrand it as "framework" rater than algorithm
+%JB: as the algorithm is rather simple and has different parts to be plugged in, I think we should rebrand it as "framework" rather than algorithm (and change the term consistently in text)
 After initializing the model $\mathcal{M}$, we incrementally \emph{select} a problem instance from the pool of instances, one at a time, evaluate the new solver $\hat{a}$ on it, and \emph{update} the model with the acquired result.
@@
   %JB: bit confusing that this is not a numerical score, but a function (see next line); maybe call it "score()"?
-or return a tuple with all estimated scores (one per solver)?; also, one could expect to get a benchmark set (see title of algorithm), but as instance selection is solver-specific, this is not case
+or return a tuple with all estimated scores (one per solver?); also, one could expect to get a benchmark set as result (see title of algorithm), but since instance selection is solver-specific, this is not case (it's rather a benchmarking schema)
 
   \KwOut{Return scoring function $\operatorname{score}\!: \mathcal{A} \cup \left\lbrace \hat{a} \right\rbrace \rightarrow \mathbb{R}$ and the fraction of sampled instances and runtime needed to provide it.}
-  %JB: we should formally specify the tuple rather than only describing it
+  %JB: we should formally specify the tuple rather than only describing it in words
 \end{algorithm}
@@
 The following paragraphs motivate why it makes sense to discretize runtimes.
-%JB: we probably should cut this a bit
+%JB: we should probably cut this a bit
 
@@
 To be applicable to $\hat{a}$, we extend $\operatorname{score}$ by replacing $\gamma\!\left(a, i\right)$ by the model's predictions $f\!: \left(\mathcal{A} \cup \left\lbrace \hat{a} \right\rbrace\right) \times \mathcal{I} \rightarrow \left\lbrace 1, \dots, k \right\rbrace$.
-%JB: as "$\operatorname{score}$" is not visible, we should use some other way of highlighting
+%JB: as "$\operatorname{score}$" is not visibly different to rest of text, we should use some other way of highlighting
 
@@
 The number of significant comparisons between solvers \emph{alone}, though, does only offer limited insights as we might achieve significant differences in all pairs by assigning some randomly skewed values as labels.
-%JB: thus, we should reorder section to mention significant difference (this paragraph) first and deciding solver pairs (previous paragraph, stronger argument) second (rather than giving the stronger argument first, making the second argument superfluous)
+%JB: thus, we should reorder this section to mention argument "significant differenc"e (this paragraph) first and "deciding solver pairs" (previous paragraph, stronger argument) second (rather than giving the stronger argument first, making the second argument superfluous)
 However, the predictiveness of the PAR-2 score, the labels' discriminatory power, as well as that it enables us to use classification at the same time, show the clear advantage of runtime discretization.
@@
 First, we present possible baseline approaches including existing work on how to pick \emph{interesting} instances~\cite{Bossek021a,GentHJKMNN14,MatriconAFSH21}.
-%JB: approaches under "Second" are also at least "inspired by existing work" and probably also pick "interesting" instances, distinction between the two categories not entirely clear
+%JB: approaches under "Second" are also at least "inspired by existing work" and probably also pick "interesting" instances, so distinction between the two categories of approaches not entirely clear
 Second, we show promising machine-learning-based sampling strategies that make use of the model's label-prediction function $f$ and are inspired by existing work within the realms of active learning~\cite{settles2009active}.
@@
     $O_{\operatorname{rt}} \leftarrow \operatorname{computeRuntimeFractionUsed}\!\left( runtime,\, \mathcal{I},\, \mathcal{S},\, r \right)$
-    %JB: do we need to call a function here or can't we directly write how this is computed (define a variable starting at 0 to which the runtime fraction is added each iteration)
+    %JB: do we need to call a function here or can't we directly write how this is computed (define a variable starting at 0 to which the runtime fraction is added each iteration)?
 
@@
 Regarding \emph{ranking} (cf. Sec.~\ref{sec:main1}), we experiment with the following approaches, including our used hyper-parameter values:
-%JB: as these approaches are mostly new (Section 3.1 only explains prediction-based ranking in general, neither observed ranking nor history nor fallback), I'm not sure if we should provide further motivation/explanation here); also applies to other components of experimental design
+%JB: as parts of these approaches are mostly new to readers (Section 3.1 only explains prediction-based ranking in general, neither observed ranking nor history nor fallback), I'm not sure if we should provide further motivation/explanation here); also applies to other components of experimental design
 \begin{itemize}
@@
 For \emph{selection} (cf. Sec.~\ref{sec:sampling1}), we experiment with the following methods, including our used hyper-parameter values:
-%JB: sampling section describes more methods than the ones listed below, I guess some got dropped after preliminary experiments
+%JB: the referenced sampling section describes more methods than the ones listed below, so I guess some methods got dropped after preliminary experiments
 \begin{itemize}
@@
 Predicting runtime labels in a similar experiment as above produced an MCC score of on average \SI{88.82}{\%} ($\sigma = \SI{4.75}{\%}$; in comparison to \SI{95.27}{\%} for predicting time-outs).
-%: JB: " similar experiment as above" - reference unclear; what do you want to say?
+%JB: "similar experiment as above" - reference unclear; what do you want to say?
 
@@
 The plotted lines represent the Pareto front for each possible choice of ranking approaches (Fig.~\ref{fig:annitraincolorranking}), selection approaches (Fig.~\ref{fig:annitraincolorselection}), and stopping criteria (Fig.~\ref{fig:annitraincolorstopping}).
-%JB: as far as I understand it, the lines only fully compare some top-level hyperparameters, but for each of these hyperparemeters, take the best out of some more fine-grained settings (the hyperparameters of the hyperparameters, so to speak), so it's no really "each possible choice"
+%JB: as far as I understand it, the lines only compare some top-level hyperparameters, but for each of these hyperparemeters, take the best (= Pareto front) out of some more fine-grained settings (the hyperparameters of the hyperparameters, so to speak), so it's no really "each possible choice" in the sense of actual, individual configurations, but rather a combination of configurations
 Colors indicate the different hyper-parameter instantiations.
@@
 Regarding the ranking approach (Fig.~\ref{fig:annitraincolorranking}), using the ranking that is induced by our runtime-label predictions consistently performs better than the partially observed PAR-2 ranking for each possible value of $\delta$.
-%JB: consistent outperformance not guaranteed in case the lines are not individual approaches, but Pareto fronts within the compared categories (which means that different parts of the line might be related to different hyperparameter settings)
+%JB: as in previous comment: consistent outperformance not guaranteed in case the lines are not individual approaches, but Pareto fronts within the compared categories (which means that different parts of the line might be related to different hyperparameter settings)
 %This was expected since a possible bias in the selection decisions of the model is accounted for when using the model's predictions for ranking.
@@
 However, the model-based information-gain sampling is beneficial if runtime is strongly favored (very small $\delta$; runtime fraction less than \SI{5}{\%}).
-%JB: we should also describe performance of random sampling
+%JB: we should also describe performance of random sampling, which is visible in plot as well
 
@@
   }
-  %JB: not sure if the color palette in the plots is grayscale-/color-blind-friendly; consider changing the palette (see https://colorbrewer2.org/#type=qualitative&scheme=Set3&n=3 ) and using different line types
+  %JB: not sure if the color palette in the plots is grayscale-/color-blind-friendly; consider changing the palette (see https://colorbrewer2.org/#type=qualitative&scheme=Set3&n=3 ) and using different line types for each color
   \label{fig:e2eallsolvers}
@@
 \begin{table}[tb!]
-  %JB: really hard to extract take-aways by looking at this table (several settings and several metrics)
+  %JB: really hard to extract take-aways by looking at this table (several hyperprameter settings and several metrics, the hyperparameter settings of our approach change from row to row)
   \centering
@@
   	Stopping criteria are either ranking convergence (Conv.) or Wilcoxon-based (Wilcoxon).
-  	%JB: How are tthe values for ranking/selection/stopping determined? whatever is best for a particular delta?
+  	%JB: How are the values for ranking/selection/stopping determined -- whatever is best for a particular delta?
   }
@@
 Thereby, we observe that for small $\delta$ (runtime more important) our approaches preferably use easy instances to get a good sense of the PAR-2 ranking fast, e.g., row with $\delta = 0.15$.
-%JB: the "e.g." part needs more explanation
+%JB: the "e.g." part needs more explanation (how can we see that easy instances used? -> fraction runtime vs. fraction instances)
 In contrast to that, harder instances are preferred for higher values of $\delta$, i.e., the sampled fraction of runtime is greater than the fraction of instances.
 In our opinion, the row with $\delta = 0.85$ represents the best trade-off between runtime and PAR-2 accuracy.
-%JB: this is a scientific article, not an opinion piece :-); all deltas represent different trade-offs, so I would not claim to have found a best trade-off (probably depends on user what is most desired)
+%JB: this is a scientific article, not an opinion piece :-); all deltas represent different trade-offs, so I would not claim to have found a best trade-off (probably depends on user which kind of trade-off is most desired)
 
@@
 However, we sample on average \SI{1.59}{} of them.
-%JB: there should be some (qualitative) interpretation (based on our results, is the sample interesting family-wise or not?) rather than just dropping a table and explaining its structure
+%JB: there should be some (qualitative) interpretation (based on our results, is the sample interesting/special family-wise or not?) rather than just dropping a table into the paper and mainly explaining its structure
 
 \begin{table}[tb!]
-  %JB: might be interesting to have scatterplot with fraction in sample / fraction in dataset (does not allow to show all family labels, but allows to show all families)
+  %JB: "fraction" here only draws family-specific comparison of sample vs.
+full dataset, but does not compare overall distribution of families in sample/dataset; might be interesting to have scatterplot with "(relative) frequency of family in sample" vs.
+"(relative) frequency of family in dataset" (does not allow to show all family labels (though we probably could manually label some families), but allows to show all families)
+  %JB: could also correlate "(relative) frequency of family in sample" with "(relative) frequency of family in dataset"
   \centering
   \caption{Top-10 most \emph{important} families regarding the average occurrences and the fraction of selected occurrences regarding our AL approach}
-  %JB: also interesting: distribution of families in sample vs. full dataset (can even correlate these two vectors); "fraction" here only draws family-specific comparison of sample to full dataset
   \label{tab:families}
@@
 On a more general note, it would make sense to generalize this evaluation framework for arbitrary $\mathcal{NP}$-complete problems.
-%JB: is there methodological generalization needed or can't we just apply the same framework (instantiating the instance-feature part differently) to these use cases?
+%JB: is there some methodological generalization needed or can't we just apply the same framework (instantiating the instance-feature part differently) to these use cases?
 Those problems share most of the relevant properties of SAT solving, i.e., there are established problem-instance features, a full benchmark run takes days, and creating new problem solvers traditionally requires expert knowledge to hand-select instances of interest.

### 2022-10-06 paper/paper.tex
Address comments in sections 1 to 4

@@
 We predict the solver ranking induced by this scoring, without predicting exact solver runtimes.
-This allows for both comparing the solvers as well as identifying solvers with similar behavior.
+This allows for both comparing the solvers as well as identifying solvers with similar behavior if looking at individual predictions per instance.
 %JB: is there any further behavior we can deduce about a solver scoring apart from a performance ranking?
 "similar behavior" could IMHO be something like two solvers being fast/slow on the same instances, but that cannot be deduced from overall score
+%TF: made it clearer
 
@@
 %JB: I think we can remove this criterion, as it also is satisified by all competitors (and is commented out in table; if we keep criterion, we should also show it in table)
+%TF: I still find it worth mentioning. So, I added it back to the table.
 Moreover, we select instances \emph{non-randomly} and \emph{incrementally}.
@@
 Sampling instances incrementally means taking runtime information from already done experiments into account when choosing the next.
-Rather than solely outputting a binary classification, i.e, the new solver is worse than solver A, we provide a \emph{scoring function} that shows by which margin a solver is worse and how similar it is to existing approaches (\emph{interpretability}).
-%JB: we should rename "interpretability" criterion (make it more specific), term "interpretability" can have lots of different meanings and our specific meaning is not very intuitive
-Finally, our strategy's cost \emph{scales linearly} with the cardinality of the set of existing solvers $\mathcal{A}$ to compare against and the instances $\mathcal{I}$ within our pool of possible benchmark instances.
+Rather than solely outputting a binary classification, i.e, the new solver is worse than solver A, we provide a \emph{scoring function} that shows by which margin a solver is worse and how similar it is to existing approaches (\emph{comparability}).
+%JB: we should rename the "interpretability" criterion (make it more specific), term "interpretability" can have lots of different meanings and our specific meaning is not very intuitive
+%TF: Renamed it to "comparability"
+Finally, our strategy incrementally selects only a single benchmark set to compare one new solver $\hat{a}$ with the set of existing solvers $\mathcal{A}$ whose cardinality can be any number (\emph{scalability}).
 %JB: for what do we measure cost here?
-% (1) solving new-solver-problem once?
--> apart from more runtimes to be acquired beforehand, is running our framework affected by increasing the set of existing solvers, and how does this depend on the configuration of our framework, i.e., sampling/selection/stopping?
-(e.g., Wilcoxon stopping criterion iterates over set of existing solvers, showing linear complexity, while fixed-subset stopping is independent from number of existing solvers)
-% (2) solving new-solver-problem for each solver out of a set of solvers?
--> is a rather theoretical scenario, because our algorithm itself assumes we know runtimes of all but one solver, so in practice we cannot throw in a bunch of new solvers and get a ranking over all of them
-We avoid comparing pairs of solvers, which would scale quadratically.
-%JB: need to discuss this; logarithmic (compare one-vs rest) / log-linear (compare each pair) effort might be sufficient if binary-search-/recursive-sorting-like procedure used
-Since a new solver idea is rarely best at first try but requires more or less tweaking, it is desired to obtain a solver ranking fast.
+%TF: made it clearer
+We avoid comparing pairs of solvers, which requires a non-constant number of benchmark sets.
+%JB: need to discuss this; logarithmic (compare one-vs rest) / log-linear (compare each pair) effort might be sufficient if the binary-search-/recursive-sorting-like procedure used
+%TF: "non-constant number of benchmark sets"
+Since a new solver idea is rarely best on the first try but requires more or less tweaking, it is desired to obtain a solver ranking fast.
 In this way, a new solver idea can be discarded if it performs poorly across the board or may be further tweaked if it shows promising results at least in some cases.
 
-To sum up our contribution, we take an existing benchmark set, i.e., the SAT Competition~2022 Anniversary Track~\cite{sat2022}, and predict a new solver's PAR-2 score on this dataset with only a fraction of the runtime that would be needed to evaluate this solver on the complete set of instances.
+To sum up our contribution, we take the SAT Competition~2022 Anniversary Track benchmark~\cite{sat2022} (5355 instances, full runtimes of 28 solvers), perform cross-validation by treating each solver as new once, and predict these solvers' PAR-2 rank with about \SI{92}{\%} accuracy with only about \SI{10}{\%} of the runtime that would be needed to evaluate these solvers on the complete set of instances.
+Both results are averages across all cross-validation iterations.
 %JB: I think we should expand a bit more on experimental design (how many instances/solvers, what do we compare), and the key results; I think it's okay if the intro becomes a bit longer (and in turn, the overview of the approach's features can be cut a bit)
-
-For reproducibility, all our source code\footnote{temporary, anonymized version for review: \url{xxx}} and and data are available on GitHub\footnote{temporary, anonymized version for review: \url{xxx}}.
+%TF: added a bit more information
+
+For reproducibility, all our source code\footnote{temporary, anonymized version for review: \url{xxx}} and and data\footnote{temporary, anonymized version for review: \url{xxx}} are available on GitHub.
 
@@
 %JB: we can probably cut this part a bit, as we discuss weaknesses of individual approaches later anyway
+%TF: it's supposed to be a short overview of what is to come
 While traditional benchmark-set-selection~\cite{HoosKSS13} lacks the aspect of runtime optimization, related work within algorithm configuration~\cite{HutterHL11} is concerned with finding better configurations fast.
@@
 %JB: as commented above, we should either remove this criterion or show it in the table
+%TF: added it back to the table
 
@@
 (2021) states that) or other competitions (e.g., those in https://www.floc2022.org/floc-olympic-games)
+%TF: @Markus maybe you are better at briefly discussing the  SAT Competition 2022 and previous competitions
 Manthey and Möhle~\cite{manthey2016better} find that competition benchmarks might contain redundant instances, and propose a feature-based approach to remove redundancy.
@@
 Those features comprise general instance-size features and graph-representation features among others.
-%JB: we should probably cut the discussion of each approach's criteria to save space, maybe only mention the unsatisifed criteria
+%JB: we should probably cut the discussion of each approach's criteria to save space, maybe only mention the unsatisfied criteria
+%TF: I still find it useful though to discuss everything in the table. Let's cut space in the end.
 Their approach \emph{ranks} solvers by providing its PAR-2 score, is \emph{feature-based}, and is \emph{scalable} in the sense that the evaluation time scales linearly with more instances and solvers.
-Also, their approach is \emph{interpretable} in the sense that the obtained PAR-2 score induces a distance between solvers.
+Also, solvers are \emph{comparable} within their approach in the sense that the obtained PAR-2 score induces a distance between solvers.
 However, they do not optimize for benchmark \emph{runtime} and select instances \emph{randomly}, apart from constraints on the instance hardness and feature distance.
@@
     Runtime optimization & \xmark & \cmark & \cmark & \cmark \\
-    %Feature-based & \cmark & \cmark & \cmark & \cmark \\
+    Feature-based & \cmark & \cmark & \cmark & \cmark \\
     Incremental non-random\\ sampling & \xmark & \xmark & \cmark & \cmark \\
     % \tablefootnote{While using a model-based selection strategy to select the next configuration to evaluate, the instance sampling is done randomly.}
-    Interpretability & \cmark & \xmark & \xmark & \cmark \\
+    Comparability & \cmark & \xmark & \xmark & \cmark \\
     % \tablefootnote{While they provide a confidence measure, the strategy's output is still only binary, i.e., whether a solver A is better or worse than solver B.}
@@
 Another interesting approach can be found within algorithm-configuration systems~\cite{HutterHL11,HoosHL21,Stutzle0P22}.
-The goal is to tune SAT-solvers for a given sub-domain of problem instances.
+The goal is to tune SAT solvers for a given sub-domain of problem instances.
 
@@
 %JB: discussion of SMAC is too long, we should cut it a bit
+%TF: Agreed. Not sure however where. I find it useful to discuss the similarities and differences between algorithm configuration and our problem.
 Although this task is different from our goal, e.g., we do not need to navigate configuration space, they use a similar framework.
@@
 First, both approaches are model-based.
-While SMAC tries to gradually build a model of the solver's performance as a function of the configuration parameters, we want to gradually build a model of a new solver's performance based on already known runtimes of instances and runtime information on already known solvers.
-Second, both approaches employ some kind of selection strategy that is based upon the previously mentioned model.
+While SMAC tries to gradually build a model of the solver's performance as a function of the configuration parameters, we want to gradually build a model of a new solver's performance based on already-known runtimes of instances and runtime information on already-known solvers.
+Second, both approaches employ some kind of selection strategy that is based on the previously mentioned model.
 Theirs selects configurations rather than instances though.
 Furthermore, Hutter et.
-al.~\cite{HutterHL11} use an approach that makes use of exploitation (what configuration parameters are already good) and exploration (parameter space with little data points) to select the next configuration for which to acquire runtime information.
+al.~\cite{HutterHL11} use an approach that makes use of exploitation (what configuration parameters are already good) and exploration (parameter space with few data points) to select the next configuration for which to acquire runtime information.
 SMAC's configuration sampling has also proven to be better than random sampling in all tested scenarios and has even proven to be significantly better in most cases~\cite{HutterHL11}.
@@
 First, an algorithm configurator cannot be used to \emph{rank} a new solver since it only looks at promising solvers/configurations rather than the overall average performance.
-If the new solver does not belong to this category (new ideas are rarely the best approach right from the start, i.e., they need some tweaking), we do not have any information on how this solver performed.
+If the new solver does not belong to the best-performing solvers (new ideas are rarely the best approach right from the start, i.e., they need some tweaking), we do not have any information on how this solver performed.
 %JB: "this category" = category of most promising solvers? not 100% clear
+%TF: made it clearer
 Second, while using a model-based selection strategy to sample configurations, instance sampling is done \emph{randomly}, i.e., without building a model over instances.
 In fact, tuning algorithms to a small or non-homogeneous set of instances might lead to over-tuning, i.e., lack of generalization~\cite{eggensperger2019pitfalls}.
-And third, their approach is not \emph{interpretable} regarding the new solver performance since we are already missing the feature to \emph{rank} the new solver among existing ones.
+And third, configurations within their approach are not \emph{comparable} since algorithm configuration seeks to find solemnly the best performing configuration.
 %JB: well, the internal meta-model of the configurator allows to estimate the solver performance for particular configurations (rather than solvers), so the approach is kind of interpretable, though regarding configurations (while for comparison to other solvers, only top-configuration performance should be interesting, so the approach is indeed not interpretable regarding solvers)
+%TF: changed
 
 \subsubsection{Incremental benchmark-set-selection.}
-Matricon et. al.~\cite{MatriconAFSH21} present an incremental benchmark-set-selection approach.
+Matricon et.~al.~\cite{MatriconAFSH21} present an incremental benchmark-set-selection approach.
 Their \emph{per-set efficient algorithm selection problem} (PSEAS) is similar to our \emph{New-Solver Problem} (cf.~Definition~\ref{def:new-solver-problem}).
-The main difference is that Matricon et. al. compare a solver to another one, while we compare to a set of existing solvers.
+The main difference is that Matricon et.~al. compare a solver to another one, while we compare it to a set of existing solvers.
 To the best of our knowledge, we are the first to tackle the latter problem.
 %JB: again, I think there are too many details for describing this single approach
-
-Given a pair of SAT solvers, Matricon et. al. want to iteratively select a subset of instances to decide with high confidence which of the two solvers is better.
+%TF: unsure what to cut as the concrete approaches are also used later in the main section
+
+Given a pair of SAT solvers, Matricon et.~al. want to iteratively select a subset of instances to decide with high confidence which of the two solvers is better.
 The selection of instances is dependent on the choice of the solvers to distinguish.
@@
 If the confidence exceeds a threshold (e.g., 95\%), they stop iterating.
-Otherwise, the aforementioned steps are repeated until the desired confidence level is reached or the pool of instances is exhausted.
+%Otherwise, the aforementioned steps are repeated until the desired confidence level is reached or the pool of instances is exhausted.
 Most of the proposed strategies in~\cite{MatriconAFSH21} are instance-based methods, i.e., no model of the solver runtimes is built.
@@
 al.~\cite{GentHJKMNN14}, (3) a variance-based strategy (select instance with the highest variance in its distribution of running times), (4) an information-gain-based method, and (5) a feature-distance-based strategy.
-To update the confidence level after each experiment, they discuss using a fixed subset size, a Wilcoxon test, and a distribution-based method.
-Their experiments show that random instance-selection with a Wilcoxon test as the stopping criterion performs consistently well across different data sets in the sense that it belongs to the Pareto front or is close to it in all cases.
-%JB: in our evaluation, we should state whether our results are consistent to this or not
+To update the confidence level after each experiment, they discuss using (1) a fixed subset size, (2) a Wilcoxon test, and (3) a distribution-based method.
+Their experiments show that random instance selection with a Wilcoxon test as the stopping criterion performs consistently well across different data sets in the sense that it belongs to the Pareto front or is close to it in all cases.
+%JB: in our evaluation, we should state whether our results are consistent with this or not
+%TF: will do
 
 Regarding our desired features of adaptive benchmark-set-selection (Tab.~\ref{tab:requirements}), the aforementioned approach already ticks off the first four features, i.e., it \emph{ranks} solvers, optimizes for \emph{runtime}, is \emph{feature-based}, and uses \emph{incremental non-random} sampling.
-Although they provide a confidence metric for their prediction, which makes this approach partially \emph{interpretable}, it is not clear how similar two given solvers are or on which instances they behave similarly, as the prediction is solely binary.
+Although they provide a confidence metric for their prediction, which makes solvers partially \emph{comparable}, it is not clear how similar two given solvers are or on which instances they behave similarly, as the prediction is solely binary.
 Moreover, the major shortcoming lies within it being not \emph{scalable} with the number solvers.
-As they compare only pairs of solvers, scaling with more and more solvers and ideas becomes an issue as there is a quadratic amount of pairs to distinguish.
+As they compare only pairs of solvers, evaluating a new solver requires a non-constant number of benchmark sets, which in turn requires more runtime.
 %JB: maybe rather log-linear than quadratic, see a prior comment in intro
+%TF: changed
 
@@
 The posed \emph{New-Solver Problem} has stark similarities to the well-studied field of active learning within recommender systems, especially the \emph{new-user problem}~\cite{RubensESK15}.
-The classic exploration-exploitation dilemma arises since, on the one hand, we want to maximize the utility an instance provides to our model and, on the other hand, minimize the cost (CPU time) that is associated with its acquisition.
-%JB: no sure if this dilemma arises here as well; dilemma is typical for searching one optimal solution in continuous space, where there is an incentive to exploit neighborhood in promising regions; however, if we find an interesting/useful (= good for assessing solver performance) instance in our approach, I'm not sure if we would continue picking similar instances ("exploiting" instead of "exploring") just because we expect them to be useful as well
+On the one hand, we want to maximize the utility an instance provides to our model and, on the other hand, minimize the cost (CPU time) that is associated with its acquisition.
+%JB: not sure if this dilemma arises here as well; a dilemma is typical for searching for one optimal solution in continuous space, where there is an incentive to exploit neighborhoods in promising regions; however, if we find an interesting/useful (= good for assessing solver performance) instance in our approach, I'm not sure if we would continue picking similar instances ("exploiting" instead of "exploring") just because we expect them to be useful as well
+%TF: removed the exploration-exploitation dilemma
 
@@
   \draw[fill=black]  (-2.6,3.4) arc (0:180:0.2);
-  \node[right] at (-5.9,3.6) {Agent (Solver)};
+  \node[right] at (-5.9,3.6) {\large Agent (Solver)};
   \draw[fill=white]  (-3.3,2.8) rectangle (-2.5,2.2);
@@
   \draw[fill=white]  (-3.1,2.6) rectangle (-2.3,2);
-  \node[right] at (-5.9,2.4) {Training Data};
+  \node[right] at (-5.9,2.4) {\large Train. Data};
   \node (v1) at (-2.8,3.4) {};
@@
   \draw[->]  (v1) edge (v2);
-  \node[right] at (-2.8,3.1) {$x, y$};
+  \node[right] at (-2.8,3.1) {\large $x, y$};
   \node (v3) at (-2.8,2) {};
@@
   \draw  (-2.4,3.2) rectangle (-2.4,3.2);
-  \node[right] at (-5.9,1.2) {Learning Alg.};
+  \node[right] at (-5.9,1.2) {\large Learning Alg.};
   \node (v5) at (-2.8,1) {};
@@
   \draw[->]  (v5) edge (v6);
-  \node[right] at (-5.9,0.25) {Approx. Function};
-  \node at (-2.8,0.25) {$f$};
+  \node[right] at (-5.9,0.25) {\large Approx. Func.};
+  \node at (-2.8,0.25) {\large $f$};
   \end{tikzpicture}
@@
   \draw[->]  (v5) edge (v6);
-  \node[right] at (-5.9,0.25) {Approx. Function};
-  \node at (-2.8,0.25) {$f$};
-  \node[right] at (-1.2,2.4) {Pool};
+  \node[right] at (-5.9,0.25) {\large Approx. Func.};
+  \node at (-2.8,0.25) {\large $f$};
+  \node[right] at (-1.2,2.4) {\large Pool};
   \draw[->, densely dashed] (-2.4,1.2) arc (-90:0:0.8);
   \draw[->, densely dashed] (-1.6,2.8) arc (0:90:0.8);
-  \node[right] at (-1.6,3.135) {$x, ?$};
+  \node[right] at (-1.6,3.135) {\large $x, ?$};
   \draw  (-1.6,2.4) ellipse (0.4 and 0.3);
@@
   \draw[->]  (v5) edge (v6);
-  \node[right] at (-5.9,0.25) {Approx. Function};
-  \node at (-2.8,0.25) {$f$};
-  \node[right] at (-1.1,2.4) {GAN};
+  \node[right] at (-5.9,0.25) {\large Approx. Func.};
+  \node at (-2.8,0.25) {\large $f$};
+  \node[right, label={[align=left]\large Gen.\\\large Model}] at (-0.6,1.85) {};
   %JB: does the model need to be a GAN or can it also be a different type of model?
-  \node[right] at (-1.3,3.135) {$x, ?$};
+  %TF: not necessarily, changed it
+  \node[right] at (-1.3,3.135) {\large $x, ?$};
   \draw[->, densely dashed] (-2.4,1.2) arc (-90:90:1.2);
@@
   \caption{Types of Learning (depiction inspired by Rubens~et.al.~\cite{RubensESK15})}
-  %JB: slightly larger font size in figure (at least having the size of the caption) would be nice
+  %JB: slightly larger font size in the figure (at least having the size of the caption) would be nice
+  %TF: made it larger
   \label{fig:learning}
@@
 
-AL algorithms can be categorized into \emph{synthesis-bas\-ed} \cite{0001AEMN22,GarzonMG22,2019gaal} and \emph{pool-bas\-ed} approaches \cite{GolbandiKL11,distribAL,HarpaleY08,KapoorGUD07,KorenBV09,KornerW06,MelvilleM04,SinhaED19,TongK01}.
+AL algorithms can be categorized into \emph{synthesis-bas\-ed} \cite{0001AEMN22,GarzonMG22,2019gaal} and \emph{pool-bas\-ed} approaches \cite{GolbandiKL11,HarpaleY08,KorenBV09}.%SinhaED19,distribAL,TongK01,KapoorGUD07,KornerW06,MelvilleM04}.
 %JB: we can probably cut some citations, the three most important should be sufficient
+%TF: done
 While synthesis-based methods (Fig.
@@
 Moreover, we have briefly outlined how active-learning techniques can help to incrementally select benchmark sets.
-In the following sections, we present an active-learning-based benchmark-set-selection approach that solves the \emph{New-Solver Problem} while optimizing for runtime and being both interpretable and scalable.
+In the following sections, we present an active-learning-based benchmark-set-selection approach that solves the \emph{New-Solver Problem} while optimizing for runtime and fulfilling both comparability and scalability.
 
@@
 \label{sec:main}
-Our algorithm follows a simple three-step schema (Alg.~\ref{algALBenchmark}).
+Our framework follows a simple three-step schema (Alg.~\ref{algALBenchmark}).
 %JB: as the algorithm is rather simple and has different parts to be plugged in, I think we should rebrand it as "framework" rather than algorithm (and change the term consistently in text)
+%TF: changed
 After initializing the model $\mathcal{M}$, we incrementally \emph{select} a problem instance from the pool of instances, one at a time, evaluate the new solver $\hat{a}$ on it, and \emph{update} the model with the acquired result.
@@
 % For now, we also restrict ourselves to the aforementioned sequential one-instance-at-a-time process.
-At the core of the model $\mathcal{M}$ is a prediction function $f\!: \left(\mathcal{A} \cup \left\lbrace \hat{a} \right\rbrace\right) \times \mathcal{I} \rightarrow \mathcal{V}$ that powers the decisions within the previously mentioned three steps.
-Note that the prediction target values $\mathcal{V}$ are unspecified to allow for more flexibility within our particular approaches.
-Refer to the next sections for more details.
+At the core of the model $\mathcal{M}$ is a prediction function $f\!: \left(\mathcal{A} \cup \left\lbrace \hat{a} \right\rbrace\right) \times \mathcal{I} \rightarrow \mathbb{R}$ that powers the decisions within the previously mentioned three steps.
+%Note that the prediction target values may be also a subset of $\mathbb{R}$.
+%Refer to the next sections for more details.
 
 \begin{algorithm}
-  \caption{Incremental benchmark-set-selection schema}
+  \caption{Incremental benchmarking framework}
   \label{algALBenchmark}
@@
   $\mathcal{M} \leftarrow \textrm{init model with $\mathcal{A}$, $\mathcal{I}$, $r$, and $\hat{a}$}$
+
+  $\tilde{\mathcal{I}} \leftarrow \emptyset$
 
@@
 
+    $\tilde{\mathcal{I}} \leftarrow \tilde{\mathcal{I}} \cup \left\lbrace s \right\rbrace$
+
     $\operatorname{updateModel}\!\left(\mathcal{M},\, s,\, t\right)$
@@
 
-  $\operatorname{score} \leftarrow \operatorname{predictScores}\!\left(\mathcal{M}\right)$
+  $\operatorname{score}\!\left(\right) \leftarrow \operatorname{predictScores}\!\left(\mathcal{M}\right)$
   %JB: bit confusing that this is not a numerical score, but a function (see next line); maybe call it "score()"?
 or return a tuple with all estimated scores (one per solver?); also, one could expect to get a benchmark set as result (see title of algorithm), but since instance selection is solver-specific, this is not case (it's rather a benchmarking schema)
-
-  \KwOut{Return scoring function $\operatorname{score}\!: \mathcal{A} \cup \left\lbrace \hat{a} \right\rbrace \rightarrow \mathbb{R}$ and the fraction of sampled instances and runtime needed to provide it.}
+  %TF: renamed it to benchmarking framework and score()
+
+  \KwOut{Return tuple of scoring function $\operatorname{score}\!: \mathcal{A} \cup \left\lbrace \hat{a} \right\rbrace \rightarrow \mathbb{R}$ and the selected instances $\tilde{\mathcal{I}}$.}
   %JB: we should formally specify the tuple rather than only describing it in words
+  %TF: made it clearer
 \end{algorithm}
 
-The following section (Sec.~\ref{sec:main1}) discusses how we may derive a solver ranking from our model (line~6).
-Thereafter, we look at possible sampling criteria (line~3) in Sec.~\ref{sec:main2} and, finally, possible stopping conditions (line~2) in Sec.~\ref{sec:main3}.
+The following section (Sec.~\ref{sec:main1}) discusses how we may derive a solver ranking from our model (line~8).
+Thereafter, we look at possible sampling criteria (line~4) in Sec.~\ref{sec:main2} and, finally, possible stopping conditions (line~3) in Sec.~\ref{sec:main3}.
 
@@
 
-Preliminary experiments, as well as research in portfolio solvers~\cite{NgokoCT19,CollauttiMMO13}, have shown however that directly predicting runtimes does not work well in practice.
+Preliminary experiments, as well as research in portfolio solvers~\cite{NgokoCT19,CollauttiMMO13}, have shown however that directly predicting runtimes without any transformation does not work well in practice.
 %JB: Collauti at el. still predict actual (though normalized) runtimes rather than discrete labels, they only heavily discretize runtimes when using them as features
-In our experiments, the mean squared error was within the same magnitude as the values to be predicted.
+%TF: added "without any transformation"
+%In our experiments, the mean squared error was within the same magnitude as the values to be predicted.
 In contrast to that, discretizing runtimes and using those labels in the model's training and prediction yields good results (cf.~Sec.~\ref{sec:eval}).
@@
 %JB: we should probably cut this a bit
+%TF: I've commented out a few sentences but we probably need to cut more.
 
@@
   %JB: does the "-1" provide any benefits, or does it only shift scores?
+  %TF: it is used to resemble the 0-indexed implementation. I'm also not happy with this, but using \lbrace 0, \dots, k - 1 \rbrace does also look ugly.
   \label{eq:rankingeq}
@@
 with $w_{a,i} := 2$ (PAR-\emph{2} penalty factor) if $\gamma\!\left(a, i\right) = k$ else $w_{a,i} := 1$.
-To be applicable to $\hat{a}$, we extend $\operatorname{score}$ by replacing $\gamma\!\left(a, i\right)$ by the model's predictions $f\!: \left(\mathcal{A} \cup \left\lbrace \hat{a} \right\rbrace\right) \times \mathcal{I} \rightarrow \left\lbrace 1, \dots, k \right\rbrace$.
-%JB: as "$\operatorname{score}$" is not visibly different to rest of text, we should use some other way of highlighting
+To be applicable to $\hat{a}$, we extend \emph{score} by replacing $\gamma\!\left(a, i\right)$ by the model's predictions $f\!: \left(\mathcal{A} \cup \left\lbrace \hat{a} \right\rbrace\right) \times \mathcal{I} \rightarrow \left\lbrace 1, \dots, k \right\rbrace$.
+%JB: as "$\operatorname{score}$" is not visibly different to the rest of the text, we should use some other way of highlighting
+%TF: replaced it with emph{score} for now
 
@@
 This means that we do not lose much discriminatory power using discrete runtime labels.
-The number of significant comparisons between solvers \emph{alone}, though, does only offer limited insights as we might achieve significant differences in all pairs by assigning some randomly skewed values as labels.
-%JB: thus, we should reorder this section to mention argument "significant differenc"e (this paragraph) first and "deciding solver pairs" (previous paragraph, stronger argument) second (rather than giving the stronger argument first, making the second argument superfluous)
-However, the predictiveness of the PAR-2 score, the labels' discriminatory power, as well as that it enables us to use classification at the same time, show the clear advantage of runtime discretization.
+%The number of significant comparisons between solvers \emph{alone}, though, does only offer limited insights as we might achieve significant differences in all pairs by assigning some randomly skewed values as labels.
+%JB: thus, we should reorder this section to mention the argument "significant difference" (this paragraph) first and "deciding solver pairs" (previous paragraph, stronger argument) second (rather than giving the stronger argument first, making the second argument superfluous)
+%TF: Removed the previous sentence
+The predictiveness of the PAR-2 score, the labels' discriminatory power, as well as that it enables us to use classification, show the clear advantage of runtime discretization.
 %JB: at the same time with what? with the other benefits?
+%TF: removed redundant phrase
 
@@
 \label{sec:main2}
-Our instance-sampling strategies (also refer to Alg.~\ref{algALBenchmark}; line 3) can be divided into two categories.
-First, we present possible baseline approaches including existing work on how to pick \emph{interesting} instances~\cite{Bossek021a,GentHJKMNN14,MatriconAFSH21}.
+Our instance-sampling strategies (also refer to Alg.~\ref{algALBenchmark}; line 4) can be divided into two categories.
+First, we present possible baseline approaches including existing work on instance-based selection strategies~\cite{Bossek021a,GentHJKMNN14,MatriconAFSH21}.
 %JB: approaches under "Second" are also at least "inspired by existing work" and probably also pick "interesting" instances, so distinction between the two categories of approaches not entirely clear
+%TF: made it clearer. First = instance-based methods, second = model-based methods
 Second, we show promising machine-learning-based sampling strategies that make use of the model's label-prediction function $f$ and are inspired by existing work within the realms of active learning~\cite{settles2009active}.
@@
 Bossek and Wagner~\cite{Bossek021a} introduce a custom ranking-score metric incorporating the number of \emph{good} pairs and \emph{bad} pairs.
-%JB: some selection methods, like this one, are not mentioned later any more (in particular, not in experimental design), so we don't need to describe them in detail (maybe mention shortly if used in preliminary experiments, but turned out to be not useful)
+%JB: some selection methods, like this one, are not mentioned later anymore (in particular, not in experimental design), so we don't need to describe them in detail (maybe mention them shortly if used in preliminary experiments, but turned out to be not useful)
+%TF: I've cut the section down a bit
 A good solver pair, regarding a particular instance, are solvers that perform according to their global ranking.
-So if solver $a$ is better than solver $b$ on instance $i$ and $a$ is also in total better than $b$, it is a \emph{good} solver pair on $i$.
+%So if solver $a$ is better than solver $b$ on instance $i$ and $a$ is also in total better than $b$, it is a \emph{good} solver pair on $i$.
 Otherwise, it is a \emph{bad} pair.
-Based on this classification, they define a 3-tuple for each instance consisting of the number of good pairs, the accumulated performance differences of bad pairs, and the accumulated performance differences of good pairs.
-By selecting instances with the highest score according to the lexicographical order of the tuples, they try to favor instances that are predictive of the final ranking.
-This assumes, though, that the new solver $\hat{a}$ behaves similarly to the existing solvers.
+Based on this classification, they assign a custom score to each instance.
+By selecting instances with the highest score, they try to favor instances that are predictive of the final ranking.
+%This assumes, though, that the new solver $\hat{a}$ behaves similarly to the existing solvers.
 
@@
 The variance-based strategy selects instances with the highest runtime variance per average runtime.
-We also implement a logarithmized version of this, i.e., the highest log-runtime variance per average log-runtime.
+We also implement a logarithmized version of this, i.e., the highest log-runtime variance per average log runtime.
 
@@
 
-The model-based information-gain sampling strategy selects the instance with the highest expected entropy reduction, assuming that the model's runtime-label probabilities for an instance are close to being correct.
+The model-based information-gain sampling strategy selects the instance with the highest expected entropy reduction.
+%This implicitly assumes that the model's runtime-label probabilities for an instance are close to being correct.
 %JB: not sure what "close to being correct" means excactly: does strategy assume correct predictions, but implicitly has a tolerance for wrong predictions, or does it explicitly account for errors somehow?
+%TF: removed the previous sentence as it probably confuses readers more and is not essential for the explanation
 To be more specific, we select the instance $i$ that maximizes
@@
 Similar to the instance-selection strategies, we can also divide stopping criteria into baseline and active-learning-based approaches.
-Also, refer to~Alg.~\ref{algALBenchmark}, line~2.
+Also, refer to~Alg.~\ref{algALBenchmark}, line~3.
 While a fixed subset-size criterion, e.g., only sample \SI{20}{\%} of all available instances, is an example of the former, a stopping criterion based on a Wilcoxon signed-rank test or the convergence of the model's predictions is an example for the latter.
@@
 (4) mean that we would prefer instances with low entropy of runtime distribution after considering the new solver, i.e., instances that are not very discriminatory?
-Thereby, \emph{wilcoxon} denotes the $p$-value of a Wilcoxon signed-rank test of the runtime-label distributions and $\gamma\!\left(a, \mathcal{I}\right)$ and $f\!\left(\hat{a}, \mathcal{I}\right)$ are the expected set extensions on $\gamma$ and $f$ respectively.
+%TF: I don't think so. We want to maximize the reduction in entropy of the runtime labels. Low entropy afterward -> we know that the solver behaves similarly to existing ones.
+Thereby, \emph{wilcoxon} denotes the $p$-value of a Wilcoxon signed-rank test of the runtime-label distributions and $\gamma\!\left(a, \mathcal{I}\right)$ and $f\!\left(\hat{a}, \mathcal{I}\right)$ are the natural set extensions on $\gamma$ and $f$ respectively.
 %JB: "set extension" might be rather technical and hard to understand; also, why "expected"? in the sense of "straightforward"?
+%TF: I already saw this expression in other papers. I'm not sure how to say the same thing clearer in the same amount of space.
 To improve the stability of this criterion, we use an exponential moving average
@@
 First and foremost, our goal is to provide the engineer of new SAT solvers with an accurate ranking.
-We define the \emph{ranking accuracy} $O_{\operatorname{acc}} \in \left[0, 1\right]$ (higher is better) by the fraction of pairs $\left(\hat{a}, a\right)$ for all $a \in \mathcal{A}$ that are decided correctly by the given ranking.
+We define the \emph{ranking accuracy} $O_{\operatorname{acc}} \in \left[0, 1\right]$ (higher is better) by the fraction of pairs $\left(\hat{a}, a\right)$ for all $a \in \mathcal{A}$ that are decided correctly by the given ranking (cf.
+Alg.~\ref{alg:eval}).
 For now, we exclude the equality of solvers since data shows that the PAR-2 scores of all our solvers are different.
@@
 The \emph{fraction of runtime} that the algorithm needs to arrive at its conclusion is denoted by $O_{\operatorname{rt}} \in \left[0, 1\right]$ (lower is better).
-This metric puts the runtime summed over the sampled instances in relation to the runtime summed over all instances in the dataset.
+This metric puts the runtime summed over the sampled instances in relation to the runtime summed over all instances in the dataset (cf. Alg.~\ref{alg:eval}).
 Overall, we want to find an approach that maximizes
@@
 
-  \KwIn{A pool of instances $\mathcal{I}$ and a set of solvers $\mathcal{S}$ with known runtimes $r\!: \mathcal{S} \times \mathcal{I} \rightarrow \left[0, \tau\right] \cup \left\lbrace \perp \right\rbrace$.}
+  \KwIn{A pool of instances $\mathcal{I}$, a set of solvers $\mathcal{S}$ with known runtimes $r\!: \mathcal{S} \times \mathcal{I} \rightarrow \left[0, \tau\right] \cup \left\lbrace \perp \right\rbrace$, and PAR-2 scores $par2\!: \mathcal{S} \rightarrow \mathbb{R}$ for each solver.}
 
@@
 
-    $\left( ranking,\, runtime \right) \leftarrow \operatorname{runALAlgorithm}\!\left(\mathcal{I}, \mathcal{A}, \left.r\right|_{\mathcal{A} \times \mathcal{I}}, \hat{a}\right)$ \tcp*{Refer to Alg.~\ref{algALBenchmark}}
-
-    $O_{\operatorname{acc}} \leftarrow \operatorname{computeRankingAccuracy}\!\left( ranking,\, \mathcal{I},\, \mathcal{S},\, r \right)$ \tcp*{Refer to Sec.~\ref{sec:goal}}
-
-    $O_{\operatorname{rt}} \leftarrow \operatorname{computeRuntimeFractionUsed}\!\left( runtime,\, \mathcal{I},\, \mathcal{S},\, r \right)$
+    $\left( score,\, \tilde{\mathcal{I}} \right) \leftarrow \operatorname{runALAlgorithm}\!\left(\mathcal{I}, \mathcal{A}, \left.r\right|_{\mathcal{A} \times \mathcal{I}}, \hat{a}\right)$ \tcp*{Refer to Alg.~\ref{algALBenchmark}}
+
+    $O_{\operatorname{acc}} \leftarrow \left(\sum_{a \in \mathcal{A}} \left[\operatorname{sign}\!\left(score\!\left(a\right) - score\!\left(\hat{a}\right)\right) = \operatorname{sign}\!\left(par2\!\left(a\right) - par2\!\left(\hat{a}\right)\right) \right]\right) / |\mathcal{A}|$
+
+    \tcp*{Iverson-bracket notation}
+
+    $O_{\operatorname{rt}} \leftarrow \left(\sum_{i \in \tilde{\mathcal{I}}} r\!\left(i, \hat{a}\right)\right) / \left(\sum_{i \in \mathcal{I}} r\!\left(i, \hat{a}\right)\right)$
     %JB: do we need to call a function here or can't we directly write how this is computed (define a variable starting at 0 to which the runtime fraction is added each iteration)?
+    %TF: fixed
 
@@
 The end-to-end experiment configurations are given below.
-%JB: still interesting to see which approaches were considered in preliminary experiments; could also list them here and indicate somehow which methods were only used preliminarily (e.g., by putting "(preliminray experiments only)" or some shorter label/sign behind them)
+%JB: still interesting to see which approaches were considered in preliminary experiments; could also list them here and indicate somehow which methods were only used preliminarily (e.g., by putting "(preliminary experiments only)" or some shorter label/sign behind them)
+%TF: Added some selection methods that are mentioned.
 
@@
 %JB: as parts of these approaches are mostly new to readers (Section 3.1 only explains prediction-based ranking in general, neither observed ranking nor history nor fallback), I'm not sure if we should provide further motivation/explanation here); also applies to other components of experimental design
+%TF: I've tried to add some more information. 
 \begin{itemize}
@@
   \begin{itemize}
-    \item History size: consider the latest 1, 10, 20, 30, or 40 predictions within a voting approach for stability.
-    \item Fallback threshold: if the difference of ranking scores drops below \SI{0.01}, \SI{0.05}, or \SI{0.1}, use the partially observed PAR-2 ranking as a tie-breaker.
+    \item History size: consider the latest 1, 10, 20, 30, or 40 predictions within a voting approach for stability. Given a particular ranking decision, the latest $x$ predictions vote on the winner.
+    \item Fallback threshold: if the difference of scores between two solvers drops below \SI{0.01}, \SI{0.05}, or \SI{0.1}, use the partially observed PAR-2 ranking as a tie-breaker.
+As discussed in Sec.~\ref{sec:main1}, ranking decisions are all correct if the difference in scores between two solvers exceeds a certain value.
     %JB: "difference of ranking" between a pair of new solver and another solver? or has this phrase a different meaning?
+    %TF: tried to clarify
   \end{itemize}
@@
 %JB: the referenced sampling section describes more methods than the ones listed below, so I guess some methods got dropped after preliminary experiments
+%TF: added them
 \begin{itemize}
@@
     %JB: how does this scaling work exactly? divide runtimes by average runtime of all solvers on that instance?
+    %TF: yes
   \end{itemize}
@@
   \end{itemize}
+
+  \item Neighborhood-aware random sampling (preliminary experiments only); randomly chooses among the instances with the highest amount of non-sampled $k$ neighbors
+  \item Ranking-based sampling~\cite{Bossek021a} (preliminary experiments only)
+  \item Discrimination-based sampling~\cite{GentHJKMNN14} (preliminary experiments only)
+  \item Variance-based sampling~\cite{MatriconAFSH21} (preliminary experiments only)
 \end{itemize}
@@
 the definition here might make stopping in early iterations rather likely, as just very few consecutive instances are needed (which might happen by coincidence; e.g., for ~5k instances, 2% minimum amount-> ~100 instances, 1% convergence duration -> ~1 instance), while the criterion becomes harder in later iterations
+    %TF: for that reason, it is combined with the "minimum amount". Using fractions makes the approach more useful when adapting it to benchmarks of different sizes.
   \end{itemize}

### 2022-10-06 paper/paper.tex
Replace tables in evaluation with plots

@@
 
-\begin{table}[tb!]
-  %JB: really hard to extract take-aways by looking at this table (several hyperprameter settings and several metrics, the hyperparameter settings of our approach change from row to row)
+% \begin{table}[tb!]
+%   %JB: really hard to extract take-aways by looking at this table (several hyperprameter settings and several metrics, the hyperparameter settings of our approach change from row to row)
+%   \centering
+%   \caption{
+%   	Pareto-optimal approaches for different choices of $\delta$ regarding our smaller hyper-parameter tuning dataset.
+%   	Rows with no changes in comparison to their predecessor are hidden.
+%   	%JB: no change in what? all columns except delta?
+%   	Ranking approaches are either predicted cluster-label ranking (Pred.) or partially observed PAR-2 ranking (PAR-2).
+%   	Selection approaches are either uncertainty-based selection (Uncert.) or information-gain-based selection (IG).
+%   	Stopping criteria are either ranking convergence (Conv.) or Wilcoxon-based (Wilcoxon).
+%   	%JB: How are the values for ranking/selection/stopping determined -- whatever is best for a particular delta?
+%   }
+%   \label{tab:paretotable}
+%   \vspace{0.2cm}
+%   \begin{tabular}{
+%     >{\centering\arraybackslash}m{0.6cm}
+%     >{\centering\arraybackslash}m{1.6cm}
+%     >{\centering\arraybackslash}m{1.6cm}
+%     >{\centering\arraybackslash}m{1.6cm}
+%     >{\centering\arraybackslash}m{1.5cm}
+%     >{\centering\arraybackslash}m{1.5cm}
+%     >{\centering\arraybackslash}m{1.5cm}
+%     >{\centering\arraybackslash}m{1.5cm}
+%   }
+%     \hline
+%     $\delta$ & Ranking & Selection & Stopping & Fraction Runtime ($O_{\operatorname{rt}}$) & Fraction Instances & $O_{\operatorname{acc}}$ & $O_{\delta}$ \\
+%     \hline
+%     0.00 & PAR-2 & Uncert. & Conv. & 0.021 & 0.025 & 0.593 & 0.979 \\
+%     0.05 & Pred. & Uncert. & Conv. & 0.024 & 0.158 & 0.765 & 0.966 \\
+%     0.10 & Pred. & Uncert. & Conv. & 0.025 & 0.128 & 0.779 & 0.955 \\
+%     0.15 & Pred. & IG & Conv. & 0.028 & 0.112 & 0.802 & 0.946 \\
+%     $\vdots$ & $\vdots$ & $\vdots$ & $\vdots$ & $\vdots$ & $\vdots$ & $\vdots$ & $\vdots$ \\
+%     0.50 & Pred. & IG & Conv. & 0.028 & 0.112 & 0.802 & 0.887 \\
+%     0.55 & Pred. & Uncert. & Conv. & 0.053 & 0.142 & 0.824 & 0.880 \\
+%     0.60 & Pred. & Uncert. & Conv. & 0.146 & 0.082 & 0.897 & 0.880 \\
+%     $\vdots$ & $\vdots$ & $\vdots$ & $\vdots$ & $\vdots$ & $\vdots$ & $\vdots$ & $\vdots$ \\
+%     0.75 & Pred. & Uncert. & Conv. & 0.146 & 0.082 & 0.897 & 0.886 \\
+%     0.80 & Pred. & Uncert. & Conv. & 0.219 & 0.128 & 0.917 & 0.890 \\
+%     \hline
+%     0.85 & Pred. & Uncert. & Conv. & 0.272 & 0.171 & 0.927 & 0.897 \\
+%     \hline
+%     0.90 & Pred. & Uncert. & Wilcoxon & 0.894 & 0.888 & 0.996 & 0.907 \\
+%     0.95 & Pred. & Uncert. & Wilcoxon & 0.931 & 0.924 & 1.000 & 0.953 \\
+%     1.00 & Pred. & Uncert. & Wilcoxon & 0.989 & 0.981 & 1.000 & 1.000 \\
+%     \hline
+%   \end{tabular}
+% \end{table}
+
+\begin{figure}[tb!]
   \centering
+  \resizebox{0.78\textwidth}{!}{
+    \graphicspath{{../plots/}}
+    \input{../plots/anni_train_optimization_goal.pgf}
+  }
   \caption{
-  	Pareto-optimal approaches for different choices of $\delta$ regarding our smaller hyper-parameter tuning dataset.
-  	Rows with no changes in comparison to their predecessor are hidden.
-  	%JB: no change in what? all columns except delta?
-  	Ranking approaches are either predicted cluster-label ranking (Pred.) or partially observed PAR-2 ranking (PAR-2).
-  	Selection approaches are either uncertainty-based selection (Uncert.) or information-gain-based selection (IG).
-  	Stopping criteria are either ranking convergence (Conv.) or Wilcoxon-based (Wilcoxon).
-  	%JB: How are the values for ranking/selection/stopping determined -- whatever is best for a particular delta?
-  }
-  \label{tab:paretotable}
-  \vspace{0.2cm}
-  \begin{tabular}{
-    >{\centering\arraybackslash}m{0.6cm}
-    >{\centering\arraybackslash}m{1.6cm}
-    >{\centering\arraybackslash}m{1.6cm}
-    >{\centering\arraybackslash}m{1.6cm}
-    >{\centering\arraybackslash}m{1.5cm}
-    >{\centering\arraybackslash}m{1.5cm}
-    >{\centering\arraybackslash}m{1.5cm}
-    >{\centering\arraybackslash}m{1.5cm}
-  }
-    \hline
-    $\delta$ & Ranking & Selection & Stopping & Fraction Runtime ($O_{\operatorname{rt}}$) & Fraction Instances & $O_{\operatorname{acc}}$ & $O_{\delta}$ \\
-    \hline
-    0.00 & PAR-2 & Uncert. & Conv. & 0.021 & 0.025 & 0.593 & 0.979 \\
-    0.05 & Pred. & Uncert. & Conv. & 0.024 & 0.158 & 0.765 & 0.966 \\
-    0.10 & Pred. & Uncert. & Conv. & 0.025 & 0.128 & 0.779 & 0.955 \\
-    0.15 & Pred. & IG & Conv. & 0.028 & 0.112 & 0.802 & 0.946 \\
-    $\vdots$ & $\vdots$ & $\vdots$ & $\vdots$ & $\vdots$ & $\vdots$ & $\vdots$ & $\vdots$ \\
-    0.50 & Pred. & IG & Conv. & 0.028 & 0.112 & 0.802 & 0.887 \\
-    0.55 & Pred. & Uncert. & Conv. & 0.053 & 0.142 & 0.824 & 0.880 \\
-    0.60 & Pred. & Uncert. & Conv. & 0.146 & 0.082 & 0.897 & 0.880 \\
-    $\vdots$ & $\vdots$ & $\vdots$ & $\vdots$ & $\vdots$ & $\vdots$ & $\vdots$ & $\vdots$ \\
-    0.75 & Pred. & Uncert. & Conv. & 0.146 & 0.082 & 0.897 & 0.886 \\
-    0.80 & Pred. & Uncert. & Conv. & 0.219 & 0.128 & 0.917 & 0.890 \\
-    \hline
-    0.85 & Pred. & Uncert. & Conv. & 0.272 & 0.171 & 0.927 & 0.897 \\
-    \hline
-    0.90 & Pred. & Uncert. & Wilcoxon & 0.894 & 0.888 & 0.996 & 0.907 \\
-    0.95 & Pred. & Uncert. & Wilcoxon & 0.931 & 0.924 & 1.000 & 0.953 \\
-    1.00 & Pred. & Uncert. & Wilcoxon & 0.989 & 0.981 & 1.000 & 1.000 \\
-    \hline
-  \end{tabular}
-\end{table}
+    Scatter plot showing different instantiations of our active-learning strategy for different $\delta$ on our smaller hyper-parameter-tuning dataset.
+    The x-axis shows the fraction of runtime $O_{\operatorname{rt}}$ that a particular instantiation needs.
+    The y-axis shows the fraction of instances that a particular instantiation selects to arrive at its conclusion.
+    The color indicates the weighting between different optimization goals $\delta \in \left[0, 1\right]$.
+    A small $\delta$ refers to favoring runtime over accuracy and a big $\delta$ refers to favoring accuracy over runtime.
+  }
+  \label{fig:annitrainoptgoal}
+\end{figure}
 
@@
 
-\begin{table}[tb!]
-  %JB: "fraction" here only draws family-specific comparison of sample vs.
+% \begin{table}[tb!]
+%   %JB: "fraction" here only draws family-specific comparison of sample vs.
 full dataset, but does not compare overall distribution of families in sample/dataset; might be interesting to have scatterplot with "(relative) frequency of family in sample" vs.
 "(relative) frequency of family in dataset" (does not allow to show all family labels (though we probably could manually label some families), but allows to show all families)
-  %JB: could also correlate "(relative) frequency of family in sample" with "(relative) frequency of family in dataset"
+%   %JB: could also correlate "(relative) frequency of family in sample" with "(relative) frequency of family in dataset"
+%   \centering
+%   \caption{Top-10 most \emph{important} families regarding the average occurrences and the fraction of selected occurrences regarding our AL approach}
+%   \label{tab:families}
+%   \begin{subtable}[t]{0.48\textwidth}
+%     \centering
+%     \caption{Sorted by the average number of occurrences in the sample}
+%     \label{tab:familiesa}
+
+%     \begin{tabular}{
+%       >{\centering\arraybackslash}m{1.8cm}
+%       >{\centering\arraybackslash}m{1.2cm}
+%       >{\centering\arraybackslash}m{1.2cm}
+%       >{\centering\arraybackslash}m{1.2cm}
+%     }
+%       \hline
+%       Family & In Sample & In Dataset & Fraction (\%) \\
+%       \hline
+%       planning              & 19.14 &  333 &  5.7 \\
+%       subgraph-isomorphism  & 16.27 &  175 &  9.3 \\
+%       crypto-graphy         & 13.11 &  311 &  4.2 \\
+%       pigeon-hole           &  7.55 &  151 &  5.0 \\
+%       fpga-routing          &  6.16 &  73  &  8.4 \\
+%       rbsat                 &  5.93 &  113 &  5.2 \\
+%       quasigroup-completion &  5.23 &  210 &  2.5 \\
+%       unknown               &  4.93 &  169 &  2.9 \\
+%       miter                 &  4.84 &  199 &  2.4 \\
+%       bitvector             &  4.73 &  214 &  2.2 \\
+%       % clique-coloring       &  4.21 &  64  &  0.066 \\
+%       % mosoi-289             &  3.75 &  31  &  0.121 \\
+%       % hardware-verification &  3.70 &  344 &  0.011 \\
+%       % hardware-bmc          &  2.77 &  78  &  0.035 \\
+%       % graph-based           &  2.77 &  134 &  0.021 \\
+%       % coloring              &  2.66 &  41  &  0.065 \\
+%       % waerden               &  2.55 &  43  &  0.059 \\
+%       % graph-coloring        &  2.50 &  87  &  0.029 \\
+%       % auto-correlation      &  2.39 &  46  &  0.052 \\
+%       % knights-problem       &  2.25 &  17  &  0.132 \\
+%       %\hline
+%       Remaining families    & 104.45 & 3407 & 3.1 \\
+%       \hline
+%       Sum                   & 192.34 & 5355 & 3.6 \\
+%       \hline
+%     \end{tabular}
+%   \end{subtable}
+%   \hfill
+%   \begin{subtable}[t]{0.48\textwidth}
+%     \centering
+%     \caption{Sorted by the fraction of sampled occurrences}
+%     \label{tab:familiesb}
+
+%     \begin{tabular}{
+%       >{\centering\arraybackslash}m{1.8cm}
+%       >{\centering\arraybackslash}m{1.2cm}
+%       >{\centering\arraybackslash}m{1.2cm}
+%       >{\centering\arraybackslash}m{1.2cm}
+%     }
+%       \hline
+%       Family & In Sample & In Dataset & Fraction (\%) \\
+%       \hline
+%       circuit-multiplier          & 1.59    &      9      &     17.7 \\
+%       subset-cardinality          & 1.55    &      9      &     17.3 \\
+%       ssp-0                       & 0.34    &      2      &     17.0 \\
+%       sliding-puzzle              & 1.52    &      11     &     13.8 \\
+%       knights-problem             & 2.25    &      17     &     13.2 \\
+%       genurq                      & 1.55    &      12     &     12.9 \\
+%       stone                       & 1.39    &      11     &     12.7 \\
+%       mosoi-289                   & 3.75    &      31     &     12.1 \\
+%       anti-bandwidth              & 1.52    &      13     &     11.7 \\
+%       random                      & 0.91    &      8      &     11.4 \\
+%       % spectrum-repacking          & 1.30    &      12     &     0.108631 \\
+%       % social-golfer               & 0.95    &      9      &     0.105159 \\
+%       % product-configuration       & 0.52    &      5      &     0.103571 \\
+%       % minimal-disagreement-parity & 1.54    &      16     &     0.095982 \\
+%       % subgraph-isomorphism        & 16.27   &       175   &     0.092959 \\
+%       % sgen-balanced               & 0.27    &      3      &     0.089286 \\
+%       % pythagorean-triples         & 1.82    &      21     &     0.086735 \\
+%       % fpga-routing                & 6.16    &      73     &     0.084393 \\
+%       % petrinet-concurrency        & 2.11    &      25     &     0.084286 \\
+%       % core-based-generator        & 1.00    &      13     &     0.076923 \\
+%       %\hline
+%       Remaining families    & 175.97 & 5232 & 3.4 \\
+%       \hline
+%       Sum                   & 192.34 & 5355 & 3.6 \\
+%       \hline
+%     \end{tabular}
+%   \end{subtable}
+% \end{table}
+
+\begin{figure}[tb!]
   \centering
-  \caption{Top-10 most \emph{important} families regarding the average occurrences and the fraction of selected occurrences regarding our AL approach}
-  \label{tab:families}
-  \begin{subtable}[t]{0.48\textwidth}
-    \centering
-    \caption{Sorted by the average number of occurrences in the sample}
-    \label{tab:familiesa}
-
-    \begin{tabular}{
-      >{\centering\arraybackslash}m{1.8cm}
-      >{\centering\arraybackslash}m{1.2cm}
-      >{\centering\arraybackslash}m{1.2cm}
-      >{\centering\arraybackslash}m{1.2cm}
-    }
-      \hline
-      Family & In Sample & In Dataset & Fraction (\%) \\
-      \hline
-      planning              & 19.14 &  333 &  5.7 \\
-      subgraph-isomorphism  & 16.27 &  175 &  9.3 \\
-      crypto-graphy         & 13.11 &  311 &  4.2 \\
-      pigeon-hole           &  7.55 &  151 &  5.0 \\
-      fpga-routing          &  6.16 &  73  &  8.4 \\
-      rbsat                 &  5.93 &  113 &  5.2 \\
-      quasigroup-completion &  5.23 &  210 &  2.5 \\
-      unknown               &  4.93 &  169 &  2.9 \\
-      miter                 &  4.84 &  199 &  2.4 \\
-      bitvector             &  4.73 &  214 &  2.2 \\
-      % clique-coloring       &  4.21 &  64  &  0.066 \\
-      % mosoi-289             &  3.75 &  31  &  0.121 \\
-      % hardware-verification &  3.70 &  344 &  0.011 \\
-      % hardware-bmc          &  2.77 &  78  &  0.035 \\
-      % graph-based           &  2.77 &  134 &  0.021 \\
-      % coloring              &  2.66 &  41  &  0.065 \\
-      % waerden               &  2.55 &  43  &  0.059 \\
-      % graph-coloring        &  2.50 &  87  &  0.029 \\
-      % auto-correlation      &  2.39 &  46  &  0.052 \\
-      % knights-problem       &  2.25 &  17  &  0.132 \\
-      %\hline
-      Remaining families    & 104.45 & 3407 & 3.1 \\
-      \hline
-      Sum                   & 192.34 & 5355 & 3.6 \\
-      \hline
-    \end{tabular}
-  \end{subtable}
-  \hfill
-  \begin{subtable}[t]{0.48\textwidth}
-    \centering
-    \caption{Sorted by the fraction of sampled occurrences}
-    \label{tab:familiesb}
-
-    \begin{tabular}{
-      >{\centering\arraybackslash}m{1.8cm}
-      >{\centering\arraybackslash}m{1.2cm}
-      >{\centering\arraybackslash}m{1.2cm}
-      >{\centering\arraybackslash}m{1.2cm}
-    }
-      \hline
-      Family & In Sample & In Dataset & Fraction (\%) \\
-      \hline
-      circuit-multiplier          & 1.59    &      9      &     17.7 \\
-      subset-cardinality          & 1.55    &      9      &     17.3 \\
-      ssp-0                       & 0.34    &      2      &     17.0 \\
-      sliding-puzzle              & 1.52    &      11     &     13.8 \\
-      knights-problem             & 2.25    &      17     &     13.2 \\
-      genurq                      & 1.55    &      12     &     12.9 \\
-      stone                       & 1.39    &      11     &     12.7 \\
-      mosoi-289                   & 3.75    &      31     &     12.1 \\
-      anti-bandwidth              & 1.52    &      13     &     11.7 \\
-      random                      & 0.91    &      8      &     11.4 \\
-      % spectrum-repacking          & 1.30    &      12     &     0.108631 \\
-      % social-golfer               & 0.95    &      9      &     0.105159 \\
-      % product-configuration       & 0.52    &      5      &     0.103571 \\
-      % minimal-disagreement-parity & 1.54    &      16     &     0.095982 \\
-      % subgraph-isomorphism        & 16.27   &       175   &     0.092959 \\
-      % sgen-balanced               & 0.27    &      3      &     0.089286 \\
-      % pythagorean-triples         & 1.82    &      21     &     0.086735 \\
-      % fpga-routing                & 6.16    &      73     &     0.084393 \\
-      % petrinet-concurrency        & 2.11    &      25     &     0.084286 \\
-      % core-based-generator        & 1.00    &      13     &     0.076923 \\
-      %\hline
-      Remaining families    & 175.97 & 5232 & 3.4 \\
-      \hline
-      Sum                   & 192.34 & 5355 & 3.6 \\
-      \hline
-    \end{tabular}
-  \end{subtable}
-\end{table}
-
-%\vspace{10cm}
+  \resizebox{0.85\textwidth}{!}{
+    \graphicspath{{../plots/}}
+    \input{../plots/anni_final_families.pgf}
+  }
+  \caption{
+    Scatter plot showing the \emph{importance} of different instance families.
+    The x-axis shows the fraction of instances within a particular family in the complete SAT Competition~2022 Anniversary Track benchmark set~\cite{sat2022}.
+    The y-axis shows the average fraction of instances within a particular family that is selected by our active-learning strategies.
+    The dashed line represents families that are represented with the same fraction in, both, our dataset and the incremental benchmark sets $\tilde{\mathcal{I}}$ that are chosen by our approach.
+  }
+  \label{fig:annifinalfamilies}
+\end{figure}
+
 \section{Conclusion}

### 2022-10-07 paper/paper.tex
abstract

@@
 \begin{abstract}
-  Experimentation is one of the main phases in algorithm engineering, including the propositional-sat\-is\-fia\-bi\-li\-ty problem (SAT).
-  While the field of SAT-solver benchmarking is well established, traditional bench\-mark-set-selection approaches do not optimize for benchmark runtime.
-  Benchmark-set selection is the process of choosing a subset of instances out of a pool to be representative and reliable in discriminating SAT solvers.
-  This paper presents an active-learning-based benchmark-set-selec\-tion approach that solves the \emph{New-Solver Problem}:
-  Given a new solver, we want to find its ranking amidst its competitors with as little runtime as possible while still giving accurate ranking predictions.
-  We can determine a new solver's rank regarding the SAT Competition 2022 Anniversary Track instances in about \SI{10}{\%} of the time it would take to run the solver on all instances of this dataset.
-  We achieve this while still maintaining a prediction accuracy of about \SI{92}{\%} regarding the PAR-2 ranking of this solver if it were evaluated on the complete dataset.
-  Additionally, we discuss the importance of instance families in the selection process. 
-  Our tool provides the engineer of propositional-satisfiability solvers with a faster way of determining a new SAT-solver's performance in comparison to sampling the benchmark set by hand.
+  Experimentation, in particular the evaluation of the runtime of a concrete implementation, is one of the most important phases in the development of algorithms.
+  This also applies to solutions of the SAT problem (propositional sat\-is\-fia\-bi\-li\-ty).
+  While the field of SAT solver benchmarking is well established, traditional benchmark selection approaches do not optimize for benchmark runtime.
+  Benchmark compilation selects representative instances from a pool of instances such that they reliably discriminate SAT solvers based on their runtime.
+  In this paper, we present an active learning based dynamic benchmark selection approach that predicts the rank of a new solver among its competitors with minimum running time and maximum rank prediction accuracy.
+  We evaluated this approach using the Anniversary track dataset from the 2022 SAT Competition. 
+  Our dynamic instance selection approach can predict the rank of a new solver after about \SI{10}{\%} of the time it would take to run the solver on all instances of this dataset, with a prediction accuracy of about \SI{92}{\%}.
+  In this paper, we also discuss the importance of instance families in the selection process. 
+  Our tool provides a reliable way for SAT solver engineers to efficiently determine the performance of a new SAT solver.
 

### 2022-10-07 paper/paper.tex
intro1

@@
 
-\title{Active Learning for SAT-Solver Benchmarking}
+\title{Active Learning for SAT solver Benchmarking}
 
@@
 \label{sec:intro}
-One of the main phases of algorithm engineering is experimentation.
-This is also true for the pro\-po\-si\-tio\-nal-sat\-is\-fia\-bi\-li\-ty problem (SAT), the archetypal $\mathcal{NP}$-complete problem.
-While state-of-the-art solvers embark upon more and more problem instances, SAT-solver evaluation is steadily getting more costly.
-Usually, solvers submitted to competitions like the annual SAT~Competition~\cite{FroleyksHIJS21,sat2022} themselves precede heavy experimentation and are the remnant of a multitude of ideas and considerations.
-While the field of SAT-solver benchmarking is well established, traditional benchmark-set-selection approaches~\cite{Gelder11,HoosKSS13} do not optimize for benchmark runtime.
-Their goal is to select a representative and robust benchmark set for discriminating solvers reliably.
-Within our work, we show that we can arrive at the same conclusions with high accuracy in only a fraction of the time that would be needed to evaluate the full benchmark set.
-To be more specific, we address the following problem:
+One of the main phases of algorithm engineering is benchmarking.
+This is also true for the pro\-po\-si\-tio\-nal sat\-is\-fia\-bi\-li\-ty problem (SAT), the archetypal $\mathcal{NP}$-complete problem.
+While state-of-the-art solvers embark upon more and more problem instances, SAT solver evaluation is steadily getting more costly.
+Competitive SAT solvers are the result of extensive experimentation and a variety of ideas and considerations~\cite{FroleyksHIJS21,sat2022}.
+While the field of SAT solver benchmarking is well established, traditional benchmark selection approaches do not optimize for benchmark runtime.
+The goal of traditional approaches is to select a representative set of instances for ranking solvers based on a scoring scheme~\cite{Gelder11,HoosKSS13}.
+SAT~Competitions typically use the widely adopted PAR-2 score, i.e, the average runtime with a penalty term of $2 \tau$ for timeouts with time-limit $\tau$~\cite{FroleyksHIJS21}.
+
+In this paper, we present a novel active learning based benchmark selection approach which is able to predict the rank of a new solver with high accuracy in only a fraction of the time needed to evaluate the full benchmark.
+The problem we solve is specified in Definition~\ref{def:new-solver-problem}.
 
 \begin{definition}[New-Solver Problem]
-	Given a pool of instances $\mathcal{I}$ and solvers $\mathcal{A}$ with already known runtimes $r\!: \mathcal{A} \times \mathcal{I} \rightarrow \left[0, \tau\right] \cup \left\lbrace \perp \right\rbrace$ (timeout $\tau$ usually \SI{5000}{s}~\cite{FroleyksHIJS21}), the problem is to decide with as little resources (CPU time) as possible how a new solver $\hat{a} \notin \mathcal{A}$ performs in comparison to the existing ones by incrementally selecting a benchmark set $\tilde{\mathcal{I}}$ on which to evaluate $\hat{a}$.
+	Given a pool of instances~$\mathcal{I}$, solvers~$\mathcal{A}$, and runtimes~$r\!: \mathcal{A} \times \mathcal{I} \rightarrow \left[0, \tau\right]$ with time limit $\tau$, maximize the confidence in predicting the rank of a new solver $\hat{a} \notin \mathcal{A}$ while minimizing the total benchmark runtime by incrementally selecting instances in $\mathcal{I}$ to evaluate $\hat{a}$.
 	\label{def:new-solver-problem}
@@
 
-Note that we know the runtimes of all solvers, except the new one, on all instances.
-There also is work on collaborative filtering, where runtimes are only partially known, for algorithm selection~\cite{misir2017alors,misir2017data}.
-We want to find a scoring function $\operatorname{score}\!: \mathcal{A} \cup \left\lbrace \hat{a} \right\rbrace \rightarrow \mathbb{R}$ assigning each solver a score.
-SAT~Competitions~\cite{FroleyksHIJS21} typically rely on the PAR-2 score, i.e, the average runtime with a penalty term of $2\tau$ for timeouts.
-We predict the solver ranking induced by this scoring, without predicting exact solver runtimes.
-This allows for both comparing the solvers as well as identifying solvers with similar behavior if looking at individual predictions per instance.
+%Note that we know the runtimes of all solvers, except the new one, on all instances.
+%There also is work on collaborative filtering, where runtimes are only partially known, for algorithm selection~\cite{misir2017alors,misir2017data}.
+%MI: this belongs to discussion of related work
+
+%We predict the solver ranking induced by this scoring, without predicting exact solver runtimes.
+
+%This allows for both comparing the solvers as well as identifying solvers with similar behavior if looking at individual predictions per instance.
 %JB: is there any further behavior we can deduce about a solver scoring apart from a performance ranking?
@@
 %TF: made it clearer
-
-This current article is proposing a novel active-learning-based benchmark-set-selection approach.
+%MI: makes no sense in the context
+
 Our algorithm is unique in that it is the only one, to the best of our knowledge, that has all of the following features (also refer to Tab.~\ref{tab:requirements}).
@@
 This section discusses work in several related fields.
-First, we take a look at traditional SAT-solver benchmarking.
-Second, we explore how SAT-solver configuration systems deal with navigating possible configurations.
+First, we take a look at traditional SAT solver benchmarking.
+Second, we explore how SAT solver configuration systems deal with navigating possible configurations.
 Third, we discuss an existing benchmark-set-selection approach that optimizes for benchmark runtime.
@@
 However, algorithm configuration is not suited for ranking.
-There is already existing work employing active learning for SAT-solver discrimination~\cite{MatriconAFSH21}.
+There is already existing work employing active learning for SAT solver discrimination~\cite{MatriconAFSH21}.
 This approach has some major shortcomings though.
@@
 
-\subsubsection{SAT-solver configuration.}
+\subsubsection{SAT solver configuration.}
 Another interesting approach can be found within algorithm-configuration systems~\cite{HutterHL11,HoosHL21,Stutzle0P22}.
@@
 
-SMAC~\cite{HutterHL11} tries to find a close-to-optimal configuration for a given SAT-solver on a given set of instances.
+SMAC~\cite{HutterHL11} tries to find a close-to-optimal configuration for a given SAT solver on a given set of instances.
 %JB: discussion of SMAC is too long, we should cut it a bit
@@
 
-\section{Active Learning for SAT-Solver Benchmarking}
+\section{Active Learning for SAT solver Benchmarking}
 \label{sec:main}
@@
 \subsection{Implementation Details}
-For reproducibility, our source code, all experiments, and data are available on GitHub\footnote{\url{https://github.com/mathefuchs/al-for-sat-solver-benchmarking}}.
+For reproducibility, our source code, all experiments, and data are available on GitHub\footnote{\url{https://github.com/mathefuchs/al-for-SAT solver-benchmarking}}.
 Our code is implemented in \textsc{Python} using \emph{scikit-learn}~\cite{scikit-learn} for making predictions and \emph{gbd-tools}~\cite{IserS18} for SAT-instance retrieval.

### 2022-10-07 paper/paper.tex
intro1

@@
 
-Our algorithm is unique in that it is the only one, to the best of our knowledge, that has all of the following features (also refer to Tab.~\ref{tab:requirements}).
+% The following belongs to Related Work discussion... already references the table there... will fuse / rework tomorrow
+Our algorithm is unique in that it is the only one, to the best of our knowledge, that has all of the following features (also refer to Table~\ref{tab:requirements}).
 Our benchmarking process outputs a \emph{ranking} to compare the new solver against the existing set of solvers as described above.
@@
 
-Tab.~\ref{tab:requirements} shows a concise overview of existing approaches with a related goal.
+Table~\ref{tab:requirements} shows a concise overview of existing approaches with a related goal.
 %JB: we can probably cut this part a bit, as we discuss weaknesses of individual approaches later anyway
@@
 
-% In terms of our desired features (also refer to Tab.~\ref{tab:requirements}),
+% In terms of our desired features (also refer to Table~\ref{tab:requirements}),
 
@@
 
-While this approach considers the configuration's \emph{runtime}, is \emph{feature-based}, and \emph{scales linearly} by having a fixed time budget, it is not capable of the other three features listed in Tab.~\ref{tab:requirements}.
+While this approach considers the configuration's \emph{runtime}, is \emph{feature-based}, and \emph{scales linearly} by having a fixed time budget, it is not capable of the other three features listed in Table~\ref{tab:requirements}.
 First, an algorithm configurator cannot be used to \emph{rank} a new solver since it only looks at promising solvers/configurations rather than the overall average performance.
@@
 
-Regarding our desired features of adaptive benchmark-set-selection (Tab.~\ref{tab:requirements}), the aforementioned approach already ticks off the first four features, i.e., it \emph{ranks} solvers, optimizes for \emph{runtime}, is \emph{feature-based}, and uses \emph{incremental non-random} sampling.
+Regarding our desired features of adaptive benchmark-set-selection (Table~\ref{tab:requirements}), the aforementioned approach already ticks off the first four features, i.e., it \emph{ranks} solvers, optimizes for \emph{runtime}, is \emph{feature-based}, and uses \emph{incremental non-random} sampling.
 Although they provide a confidence metric for their prediction, which makes solvers partially \emph{comparable}, it is not clear how similar two given solvers are or on which instances they behave similarly, as the prediction is solely binary.
@@
 Moreover, it decides all solvers correctly if the difference in the cluster-label-ranking scores between two solvers is at least \SI{0.11}{}.
-%Refer to Tab.~\ref{tab:discreteranking} for a comparison with the PAR-2 ranking.
+%Refer to Table~\ref{tab:discreteranking} for a comparison with the PAR-2 ranking.
 The Spearman correlation~\cite{de2016comparing} of cluster-label and PAR-2 ranking is about \SI{0.988} (in comparison to a perfect value of \SI{1}).
@@
 
-To obtain the results in Tab.~\ref{tab:mcc}, we perform cross-validation among all solvers: we assume that all other solver runtimes are known and provide \SI{64}{\%} of ground-truth runtime labels of the target solver for training, use \SI{16}{\%} as a validation set for hyper-parameter tuning, and \SI{20}{\%} as a test set (5-fold cross-validation).
+To obtain the results in Table~\ref{tab:mcc}, we perform cross-validation among all solvers: we assume that all other solver runtimes are known and provide \SI{64}{\%} of ground-truth runtime labels of the target solver for training, use \SI{16}{\%} as a validation set for hyper-parameter tuning, and \SI{20}{\%} as a test set (5-fold cross-validation).
 This is repeated for each solver once.
@@
 
-Tab.~\ref{tab:paretotable} also shows the performance of particular hyper-parameter choices for different values of $\delta$.
+Table~\ref{tab:paretotable} also shows the performance of particular hyper-parameter choices for different values of $\delta$.
 %JB: some plot for comparison would be better than a truncated table showing a variety hyperparameter configs
@@
 Thereby, the average benchmark set contains about \SI{192.34}{} instances.
-Tab.~\ref{tab:familiesa} lists the ten most frequent families that are chosen.
+Table~\ref{tab:familiesa} lists the ten most frequent families that are chosen.
 For example, \emph{planning} instances account for on average \SI{19.14}{} of the \SI{192.34}{} benchmark instances.
-In contrast to that, Tab.~\ref{tab:familiesb} shows the families that are sampled the most often relative to their occurrences in the underlying dataset.
+In contrast to that, Table~\ref{tab:familiesb} shows the families that are sampled the most often relative to their occurrences in the underlying dataset.
 There are, for example, only nine \emph{circuit-multiplier} instances in the complete dataset.

### 2022-10-08 paper/paper.tex
minor-fixes

@@
 
-\title{Active Learning for SAT solver Benchmarking}
+\title{Active Learning for SAT Solver Benchmarking}
 
@@
 \begin{abstract}
-  Experimentation, in particular the evaluation of the runtime of a concrete implementation, is one of the most important phases in the development of algorithms.
+  Benchmarking is one of the most important phases in the development of algorithms.
   This also applies to solutions of the SAT problem (propositional sat\-is\-fia\-bi\-li\-ty).
@@
 
-\subsubsection{Traditional benchmark-set-selection.}
+\subsubsection{Traditional Benchmark Selection.}
 Recent studies~\cite{abs-2107-07002,NiesslHWCB22} show that selecting a benchmark set in general is a difficult endeavor for multiple reasons.
@@
 
-\section{Active Learning for SAT solver Benchmarking}
+\section{Active Learning for SAT Solver Benchmarking}
 \label{sec:main}
@@
 \begin{algorithm}
-  \caption{Incremental benchmarking framework}
+  \caption{Incremental Benchmarking Framework}
   \label{algALBenchmark}
 
-  \KwIn{A pool of instances $\mathcal{I}$, a set of known solvers $\mathcal{A}$ with known runtimes $r\!: \mathcal{A} \times \mathcal{I} \rightarrow \left[0, \tau\right] \cup \left\lbrace \perp \right\rbrace$, and a new solver $\hat{a} \notin \mathcal{A}$.}
+  \KwIn{Instances $\mathcal{I}$, Solvers $\mathcal{A}$, Runtimes $r\!: \mathcal{A} \times \mathcal{I} \rightarrow \left[0, \tau\right]$}
+  \KwIn{New Solver $\hat{a} \notin \mathcal{A}$}
 

### 2022-10-08 paper/paper.tex
intro2

@@
 
-% The following belongs to Related Work discussion... already references the table there... will fuse / rework tomorrow
-Our algorithm is unique in that it is the only one, to the best of our knowledge, that has all of the following features (also refer to Table~\ref{tab:requirements}).
-Our benchmarking process outputs a \emph{ranking} to compare the new solver against the existing set of solvers as described above.
+%MI: In the following I stripped out stuff which belongs to related work section. Stick to advertising own approach here.
+Our benchmarking process outputs a \emph{ranking} to compare the new solver against the existing set of solvers.
 We optimize the \emph{runtime} that our strategy needs to arrive at its conclusion.
-We use problem-instance and runtime \emph{features} $\varphi\!: \mathcal{A} \times \mathcal{I} \rightarrow \mathbb{R}^n$ since such features already have proven to be helpful in classic benchmark-set-selection approaches~\cite{HoosKSS13} as well as in portfolio selection~\cite{CollauttiMMO13,NgokoCT19}.
-%JB: I think we can remove this criterion, as it also is satisified by all competitors (and is commented out in table; if we keep criterion, we should also show it in table)
-%TF: I still find it worth mentioning. So, I added it back to the table.
+We use problem-instance and runtime \emph{features}.
 Moreover, we select instances \emph{non-randomly} and \emph{incrementally}.
@@
 Sampling instances incrementally means taking runtime information from already done experiments into account when choosing the next.
-Rather than solely outputting a binary classification, i.e, the new solver is worse than solver A, we provide a \emph{scoring function} that shows by which margin a solver is worse and how similar it is to existing approaches (\emph{comparability}).
+Rather than solely outputting a binary classification, i.e, the new solver is worse than solver A, we provide a \emph{scoring} function that shows by which margin a solver is worse and how similar it is to existing approaches.
 %JB: we should rename the "interpretability" criterion (make it more specific), term "interpretability" can have lots of different meanings and our specific meaning is not very intuitive
 %TF: Renamed it to "comparability"
-Finally, our strategy incrementally selects only a single benchmark set to compare one new solver $\hat{a}$ with the set of existing solvers $\mathcal{A}$ whose cardinality can be any number (\emph{scalability}).
+%MI: Renamed it to "scoring"
+Our approach is \emph{scalable} in that it ranks a new solver $\hat{a}$ among any number of known solvers $\mathcal{A}$.
 %JB: for what do we measure cost here?
@@
   \centering
-  \caption{
-  	Support for desired features that are introduced in Sec.~\ref{sec:intro}.
-  	The shown approaches include traditional benchmark-set-selection techniques~\cite{HoosKSS13}, an algorithm configuration system~\cite{HutterHL11}, an existing active-learning approach~\cite{MatriconAFSH21}, as well as our approach.
-  }
-  \label{tab:requirements}
-  \vspace{0.2cm}
   \begin{tabular}{
@@
     \hline
-    Feature & Hoos \mbox{et. al.~\cite{HoosKSS13}} & SMAC~\cite{HutterHL11} & Matricon \mbox{et. al.~\cite{MatriconAFSH21}} & Our approach \\
+    Feature & Hoos~\cite{HoosKSS13} & SMAC~\cite{HutterHL11} & Matricon~\cite{MatriconAFSH21} & Our approach \\
     \hline
     Ranking & \cmark & \xmark & \cmark & \cmark \\
-    Runtime optimization & \xmark & \cmark & \cmark & \cmark \\
-    Feature-based & \cmark & \cmark & \cmark & \cmark \\
-    Incremental non-random\\ sampling & \xmark & \xmark & \cmark & \cmark \\
+    Runtime Minimization & \xmark & \cmark & \cmark & \cmark \\
+    % Feature-based & \cmark & \cmark & \cmark & \cmark \\
+    Incremental Sampling & \xmark & \xmark & \cmark & \cmark \\
     % \tablefootnote{While using a model-based selection strategy to select the next configuration to evaluate, the instance sampling is done randomly.}
-    Comparability & \cmark & \xmark & \xmark & \cmark \\
+    Scoring & \cmark & \xmark & \xmark & \cmark \\
     % \tablefootnote{While they provide a confidence measure, the strategy's output is still only binary, i.e., whether a solver A is better or worse than solver B.}
@@
   \end{tabular}
+  ~\\[1em]
+  \caption{Support for desired features that are introduced in Section~\ref{sec:intro}.
+	The shown approaches include traditional benchmark-set-selection techniques~\cite{HoosKSS13}, an algorithm configuration system~\cite{HutterHL11}, an existing active-learning approach~\cite{MatriconAFSH21}, as well as our approach.
+  }
+  \label{tab:requirements}
 \end{table}
@@
 
-The following section (Sec.~\ref{sec:main1}) discusses how we may derive a solver ranking from our model (line~8).
-Thereafter, we look at possible sampling criteria (line~4) in Sec.~\ref{sec:main2} and, finally, possible stopping conditions (line~3) in Sec.~\ref{sec:main3}.
+The following section (Section~\ref{sec:main1}) discusses how we may derive a solver ranking from our model (line~8).
+Thereafter, we look at possible sampling criteria (line~4) in Section~\ref{sec:main2} and, finally, possible stopping conditions (line~3) in Section~\ref{sec:main3}.
 
@@
 %In our experiments, the mean squared error was within the same magnitude as the values to be predicted.
-In contrast to that, discretizing runtimes and using those labels in the model's training and prediction yields good results (cf.~Sec.~\ref{sec:eval}).
+In contrast to that, discretizing runtimes and using those labels in the model's training and prediction yields good results (cf.~Section~\ref{sec:eval}).
 The following paragraphs motivate why it makes sense to discretize runtimes.
@@
 \begin{algorithm}[htbp]
-  \caption{Evaluation framework}
+  \caption{Evaluation Framework}
   \label{alg:eval}
@@
 Note that we are not considering all previously listed approaches since a grid search of all combinations would be infeasible.
-Rather, we filter approaches based on preliminary experimental results (cf. Sec.~\ref{sec:evalprel}) and do the main end-to-end experiments only with a subset.
+Rather, we filter approaches based on preliminary experimental results (cf. Section~\ref{sec:evalprel}) and do the main end-to-end experiments only with a subset.
 The end-to-end experiment configurations are given below.
@@
 \subsubsection{Ranking.}
-Regarding \emph{ranking} (cf. Sec.~\ref{sec:main1}), we experiment with the following approaches, including our used hyper-parameter values:
+Regarding \emph{ranking} (cf. Section~\ref{sec:main1}), we experiment with the following approaches, including our used hyper-parameter values:
 %JB: as parts of these approaches are mostly new to readers (Section 3.1 only explains prediction-based ranking in general, neither observed ranking nor history nor fallback), I'm not sure if we should provide further motivation/explanation here); also applies to other components of experimental design
@@
     \item Fallback threshold: if the difference of scores between two solvers drops below \SI{0.01}, \SI{0.05}, or \SI{0.1}, use the partially observed PAR-2 ranking as a tie-breaker.
-As discussed in Sec.~\ref{sec:main1}, ranking decisions are all correct if the difference in scores between two solvers exceeds a certain value.
+As discussed in Section~\ref{sec:main1}, ranking decisions are all correct if the difference in scores between two solvers exceeds a certain value.
     %JB: "difference of ranking" between a pair of new solver and another solver? or has this phrase a different meaning?
@@
 \subsubsection{Selection.}
-For \emph{selection} (cf. Sec.~\ref{sec:sampling1}), we experiment with the following methods, including our used hyper-parameter values:
+For \emph{selection} (cf. Section~\ref{sec:sampling1}), we experiment with the following methods, including our used hyper-parameter values:
 %JB: the referenced sampling section describes more methods than the ones listed below, so I guess some methods got dropped after preliminary experiments
@@
 \subsubsection{Stopping.}
-For \emph{stopping} decisions (cf. Sec.~\ref{sec:main3}), we experiment with the following criteria, including our used hyper-parameter values:
+For \emph{stopping} decisions (cf. Section~\ref{sec:main3}), we experiment with the following criteria, including our used hyper-parameter values:
 \begin{itemize}
@@
 \subsubsection{Runtime-label prediction model.}
-As a runtime-label prediction model, we only make use of one method that has proven to be successful in preliminary experiments (cf.~Sec.~\ref{sec:evalprel}).
+As a runtime-label prediction model, we only make use of one method that has proven to be successful in preliminary experiments (cf.~Section~\ref{sec:evalprel}).
 Our model of choice is an ensemble, stacking a quadratic-discriminant analysis\footnote{With a regularization parameter of zero and a singular value threshold of zero} onto a random forest model\footnote{With \emph{entropy} splitting criterion and \emph{balanced} class weights} using a decision tree\footnote{With \emph{Gini} impurity splitting criterion; choosing the best split for each branching decision and maximum tree depth of 5 levels} to weight its members.
@@
 %JB: if we dedicate a whole section to these results, we should probably rename this section (not "preliminary" in the typical sense); for TACAS, I guess this can be cut a bit
-An exhaustive grid search of all approaches outlined in Sec.~\ref{sec:main} is infeasible.
+An exhaustive grid search of all approaches outlined in Section~\ref{sec:main} is infeasible.
 We are, therefore, narrowing down the search by filtering based on preliminary results.
@@
 We then gradually merge intervals whose single-link logarithmic distance is the smallest until the desired number of partitions is reached.
-As already mentioned in Sec.~\ref{sec:main1}, this approach has good properties regarding the predictiveness of the PAR-2 score ranking and its discriminatory power.
+As already mentioned in Section~\ref{sec:main1}, this approach has good properties regarding the predictiveness of the PAR-2 score ranking and its discriminatory power.
 Also, preliminary experiments showed that our stacking ensemble performed particularly well for this kind of discretization method.
@@
 \subsubsection{Tuning our algorithm.}
-Our main end-to-end experiments follow the evaluation framework that has been introduced in Sec.~\ref{sec:evalframe}.
-Fig.~\ref{fig:e2eallsolvers} shows the performance of the approaches listed in Sec.~\ref{sec:hyper} on $O_{\operatorname{rt}}$-$O_{\operatorname{acc}}$-diagrams for the smaller hyper-parameter-tuning dataset.
+Our main end-to-end experiments follow the evaluation framework that has been introduced in Section~\ref{sec:evalframe}.
+Fig.~\ref{fig:e2eallsolvers} shows the performance of the approaches listed in Section~\ref{sec:hyper} on $O_{\operatorname{rt}}$-$O_{\operatorname{acc}}$-diagrams for the smaller hyper-parameter-tuning dataset.
 The plotted lines represent the Pareto front for each possible choice of ranking approaches (Fig.~\ref{fig:annitraincolorranking}), selection approaches (Fig.~\ref{fig:annitraincolorselection}), and stopping criteria (Fig.~\ref{fig:annitraincolorstopping}).

### 2022-10-08 paper/paper.tex
related1

@@
 \subsubsection{Traditional Benchmark Selection.}
-Recent studies~\cite{abs-2107-07002,NiesslHWCB22} show that selecting a benchmark set in general is a difficult endeavor for multiple reasons.
-Dehghani~et.~al.~\cite{abs-2107-07002} describe how a biased benchmarking set can easily lead to fallacious interpretations.
-Moreover, Nießl~et.~al.~\cite{NiesslHWCB22} state that benchmark-set design has many movable parts, e.g., performance measures, aggregation, and handling of missing values.
-Questionable research practices might modify those design elements a-posteriori to fit expectations, leading to biased results.
-
-Benchmark-set selection is an important issue for solver competitions, e.g., the SAT Competition.
+Recent studies show that selecting a benchmark set in general is a difficult endeavor for multiple reasons.
+Dehghani~et~al. describe how a biased benchmarking set can easily lead to fallacious interpretations~\cite{abs-2107-07002}.
+Moreover, Nießl~et~al. state that benchmark-set design has many movable parts, e.g., performance measures, aggregation, and handling of missing values.
+Questionable research practices might modify those design elements a-posteriori to fit expectations, leading to biased results~\cite{NiesslHWCB22}.
+
+Benchmark selection is an important issue for solver competitions, e.g., the SAT Competition.
 In such competitions, the organizers define the rules for composing the corresponding benchmarks.
-Balint et al.~\cite{balint2015overview} provide an overview of benchmark-selection criteria in different solver competitions.
+Balint et al. provide an overview of benchmark-selection criteria in different solver competitions~\cite{balint2015overview}.
 %JB: not sure if we should provide some details regarding SAT Competition 2022 \cite{sat2022}, previous competitions (if different procedure used; Related-Work section of Matricon et al.
@@
 %TF: @Markus maybe you are better at briefly discussing the  SAT Competition 2022 and previous competitions
-Manthey and Möhle~\cite{manthey2016better} find that competition benchmarks might contain redundant instances, and propose a feature-based approach to remove redundancy.
-M{\i}s{\i}r~\cite{misir2021benchmark} presents a feature-based approach to reduce benchmark sets by matrix factorization and clustering.
-
-Hoos et. al.~\cite{HoosKSS13} discuss which properties are most desirable when selecting a set of SAT benchmark instances.
+Manthey and Möhle find that competition benchmarks might contain redundant instances, and propose a feature-based approach to remove redundancy~\cite{manthey2016better}.
+M{\i}s{\i}r presents a feature-based approach to reduce benchmark sets by matrix factorization and clustering~\cite{misir2021benchmark}.
+
+Hoos et al. discuss which properties are most desirable when selecting a set of SAT benchmark instances~\cite{HoosKSS13}.
 The selection strategy presented is static, i.e., it does not depend on particular solvers to distinguish.
@@
 \subsubsection{SAT solver configuration.}
-Another interesting approach can be found within algorithm-configuration systems~\cite{HutterHL11,HoosHL21,Stutzle0P22}.
+Another interesting approach can be found within algorithm configuration systems~\cite{HutterHL11,HoosHL21,Stutzle0P22}.
 The goal is to tune SAT solvers for a given sub-domain of problem instances.
 
-SMAC~\cite{HutterHL11} tries to find a close-to-optimal configuration for a given SAT solver on a given set of instances.
+SMAC tries to find a close-to-optimal configuration for a given SAT solver on a given set of instances~\cite{HutterHL11}.
 %JB: discussion of SMAC is too long, we should cut it a bit
@@
 \subsubsection{Incremental benchmark-set-selection.}
-Matricon et.~al.~\cite{MatriconAFSH21} present an incremental benchmark-set-selection approach.
+Matricon et~al.~\cite{MatriconAFSH21} present an incremental benchmark-set-selection approach.
 Their \emph{per-set efficient algorithm selection problem} (PSEAS) is similar to our \emph{New-Solver Problem} (cf.~Definition~\ref{def:new-solver-problem}).
-The main difference is that Matricon et.~al. compare a solver to another one, while we compare it to a set of existing solvers.
+The main difference is that Matricon et~al. compare a solver to another one, while we compare it to a set of existing solvers.
 To the best of our knowledge, we are the first to tackle the latter problem.
@@
 
-Given a pair of SAT solvers, Matricon et.~al. want to iteratively select a subset of instances to decide with high confidence which of the two solvers is better.
+Given a pair of SAT solvers, Matricon et~al. want to iteratively select a subset of instances to decide with high confidence which of the two solvers is better.
 The selection of instances is dependent on the choice of the solvers to distinguish.
@@
 
-For the scoring metric, they discuss (1) a random strategy as a baseline, (2) a discrimination strategy based on the work of Gent et.
-al.~\cite{GentHJKMNN14}, (3) a variance-based strategy (select instance with the highest variance in its distribution of running times), (4) an information-gain-based method, and (5) a feature-distance-based strategy.
+For the scoring metric, they discuss (1) a random strategy as a baseline, (2) a discrimination strategy based on the work of Gent et al.~\cite{GentHJKMNN14}, (3) a variance-based strategy (select instance with the highest variance in its distribution of running times), (4) an information-gain-based method, and (5) a feature-distance-based strategy.
 To update the confidence level after each experiment, they discuss using (1) a fixed subset size, (2) a Wilcoxon test, and (3) a distribution-based method.
@@
 
-Regarding our desired features of adaptive benchmark-set-selection (Table~\ref{tab:requirements}), the aforementioned approach already ticks off the first four features, i.e., it \emph{ranks} solvers, optimizes for \emph{runtime}, is \emph{feature-based}, and uses \emph{incremental non-random} sampling.
+Regarding our desired features of adaptive benchmark-set-selection (Table~\ref{tab:requirements}), the aforementioned approach already ticks off the first four features, i.e, it \emph{ranks} solvers, optimizes for \emph{runtime}, is \emph{feature-based}, and uses \emph{incremental non-random} sampling.
 Although they provide a confidence metric for their prediction, which makes solvers partially \emph{comparable}, it is not clear how similar two given solvers are or on which instances they behave similarly, as the prediction is solely binary.
@@
 
-Recent synthesis-based methods within the field of satisfiability solving~\cite{0001AEMN22,GarzonMG22} show how to generate problem instances with desired properties.
-This goal is, however, orthogonal to ours.
+Recent synthesis-based methods within the field of satisfiability solving show how to generate problem instances with desired properties.
+This goal is, however, orthogonal to ours~\cite{0001AEMN22,GarzonMG22}.
 While those approaches want to generate problem instances on which a solver is good or bad, we want to predict whether a solver is good or bad on an existing benchmark set.
@@
 
-In the field of SAT-solving, Volpato and Guangyan~\cite{volpato2019active} use uncertainty-sampling-based AL to learn an instance-specific algorithm selector.
+In the field of SAT-solving, Volpato and Guangyan use uncertainty-sampling-based AL to learn an instance-specific algorithm selector~\cite{volpato2019active}.
 Rather than benchmarking a solver's overall performance, the goal is to recommend the best solver out of a set of solvers for each SAT instance.
-
-\subsubsection{Summary.}
-To sum up this section, we have shown both traditional and recent benchmark-set-selection approaches.
-Moreover, we have briefly outlined how active-learning techniques can help to incrementally select benchmark sets.
-In the following sections, we present an active-learning-based benchmark-set-selection approach that solves the \emph{New-Solver Problem} while optimizing for runtime and fulfilling both comparability and scalability.
 
@@
 
-Gent~et.~al.~\cite{GentHJKMNN14} make use of a concept known from racing for automated model selection.
+Gent~et~al.~\cite{GentHJKMNN14} make use of a concept known from racing for automated model selection.
 Given a particular instance and a parameter $\rho$, a solver is $\rho$-dominated if it is at least $\rho$ times worse than the best solver on this problem instance.
@@
 Finally, we also implement some of the selection approaches described by Matricon~et.al.~\cite{MatriconAFSH21}.
-Besides uniformly random sampling and the aforementioned discrimination-based method by Gent~et.~al.~\cite{GentHJKMNN14}, they also describe a variance-based selection approach.
+Besides uniformly random sampling and the aforementioned discrimination-based method by Gent~et~al.~\cite{GentHJKMNN14}, they also describe a variance-based selection approach.
 The variance-based strategy selects instances with the highest runtime variance per average runtime.

### 2022-10-08 paper/paper.tex
related2

@@
 \section{Related Work}
-This section discusses work in several related fields.
-First, we take a look at traditional SAT solver benchmarking.
-Second, we explore how SAT solver configuration systems deal with navigating possible configurations.
-Third, we discuss an existing benchmark-set-selection approach that optimizes for benchmark runtime.
-Finally, we show how active-learning techniques can help to tackle the aforementioned \emph{New-Solver Problem}.
-
-Table~\ref{tab:requirements} shows a concise overview of existing approaches with a related goal.
-%JB: we can probably cut this part a bit, as we discuss weaknesses of individual approaches later anyway
-%TF: it's supposed to be a short overview of what is to come
-While traditional benchmark-set-selection~\cite{HoosKSS13} lacks the aspect of runtime optimization, related work within algorithm configuration~\cite{HutterHL11} is concerned with finding better configurations fast.
-However, algorithm configuration is not suited for ranking.
-There is already existing work employing active learning for SAT solver discrimination~\cite{MatriconAFSH21}.
-This approach has some major shortcomings though.
-Finally, our approach seeks to provide all stated features.
-Note that all approaches already make use of problem-instance features as it showed to be beneficial in past work~\cite{HoosKSS13,CollauttiMMO13,NgokoCT19}.
-%JB: as commented above, we should either remove this criterion or show it in the table
-%TF: added it back to the table
-
-\subsubsection{Traditional Benchmark Selection.}
-Recent studies show that selecting a benchmark set in general is a difficult endeavor for multiple reasons.
-Dehghani~et~al. describe how a biased benchmarking set can easily lead to fallacious interpretations~\cite{abs-2107-07002}.
-Moreover, Nießl~et~al. state that benchmark-set design has many movable parts, e.g., performance measures, aggregation, and handling of missing values.
+
+Being of high interest in many fields, benchmarking is an active area of research on its own. 
+Recent studies show that benchmark compilation is a difficult endeavor for multiple reasons.
+Dehghani~et~al. describe how a biased benchmark can easily lead to fallacious interpretations~\cite{abs-2107-07002}.
+Nießl~et~al. state that benchmark compilation has many movable parts, e.g, performance measures, aggregation, and handling of missing values.
 Questionable research practices might modify those design elements a-posteriori to fit expectations, leading to biased results~\cite{NiesslHWCB22}.
 
-Benchmark selection is an important issue for solver competitions, e.g., the SAT Competition.
+Benchmark compilation is an important issue for competitions, e.g., the SAT Competition.
 In such competitions, the organizers define the rules for composing the corresponding benchmarks.
@@
 
+%MI: one paragraph for Hoos
 Hoos et al. discuss which properties are most desirable when selecting a set of SAT benchmark instances~\cite{HoosKSS13}.
 The selection strategy presented is static, i.e., it does not depend on particular solvers to distinguish.
-Desirable properties are a variety of instances (avoids over-fitting of solver approaches), adapted instance hardness (not too easy but also not too hard; solvers cannot be distinguished if too many time-out or finish immediately), and no duplicates (having duplicates gives those instances more weight).
+Desirable properties are a variety of instances to avoid over-fitting, adapted instance hardness (not too easy but also not too hard), and avoidance of duplicate instances.
 To avoid having too similar instances, they make use of a distance-based approach using the SATzilla features~\cite{XuHHL08,features}.
-Those features comprise general instance-size features and graph-representation features among others.
-%JB: we should probably cut the discussion of each approach's criteria to save space, maybe only mention the unsatisfied criteria
-%TF: I still find it useful though to discuss everything in the table. Let's cut space in the end.
-Their approach \emph{ranks} solvers by providing its PAR-2 score, is \emph{feature-based}, and is \emph{scalable} in the sense that the evaluation time scales linearly with more instances and solvers.
+Their approach \emph{ranks} solvers, is \emph{feature-based}, and \emph{scalable} in the sense that the evaluation time scales linearly with more instances and solvers.
 Also, solvers are \emph{comparable} within their approach in the sense that the obtained PAR-2 score induces a distance between solvers.
@@
 %Dynamic strategies chose instances incrementally, one at a time, by integrating runtime data (of the instances that the solver has been already evaluated on) to make subsequent selection decisions rather than completely choosing the benchmark set upfront.
-
-% In terms of our desired features (also refer to Table~\ref{tab:requirements}),
-
-%The aforementioned SATzilla features~\cite{XuHHL08,features} are provided within the GBD meta-data database~\cite{IserS18}.
-%Finding a good representation of instances is quintessential for our work since any feature-based benchmark-set-selection approach is only as good as the quality of the employed features.
 
@@
 
-\subsubsection{SAT solver configuration.}
+% MI: one paragraph for SMAC
 Another interesting approach can be found within algorithm configuration systems~\cite{HutterHL11,HoosHL21,Stutzle0P22}.
 The goal is to tune SAT solvers for a given sub-domain of problem instances.
-
-SMAC tries to find a close-to-optimal configuration for a given SAT solver on a given set of instances~\cite{HutterHL11}.
-%JB: discussion of SMAC is too long, we should cut it a bit
-%TF: Agreed. Not sure however where. I find it useful to discuss the similarities and differences between algorithm configuration and our problem.
 Although this task is different from our goal, e.g., we do not need to navigate configuration space, they use a similar framework.
@@
 SMAC's configuration sampling has also proven to be better than random sampling in all tested scenarios and has even proven to be significantly better in most cases~\cite{HutterHL11}.
-
 While this approach considers the configuration's \emph{runtime}, is \emph{feature-based}, and \emph{scales linearly} by having a fixed time budget, it is not capable of the other three features listed in Table~\ref{tab:requirements}.
 First, an algorithm configurator cannot be used to \emph{rank} a new solver since it only looks at promising solvers/configurations rather than the overall average performance.
-If the new solver does not belong to the best-performing solvers (new ideas are rarely the best approach right from the start, i.e., they need some tweaking), we do not have any information on how this solver performed.
-%JB: "this category" = category of most promising solvers? not 100% clear
-%TF: made it clearer
 Second, while using a model-based selection strategy to sample configurations, instance sampling is done \emph{randomly}, i.e., without building a model over instances.
-In fact, tuning algorithms to a small or non-homogeneous set of instances might lead to over-tuning, i.e., lack of generalization~\cite{eggensperger2019pitfalls}.
 And third, configurations within their approach are not \emph{comparable} since algorithm configuration seeks to find solemnly the best performing configuration.
@@
 
-\subsubsection{Incremental benchmark-set-selection.}
-Matricon et~al.~\cite{MatriconAFSH21} present an incremental benchmark-set-selection approach.
+% MI: one paragraph for Matricon
+Matricon~et~al. present an incremental benchmark instance selection approach~\cite{MatriconAFSH21}.
 Their \emph{per-set efficient algorithm selection problem} (PSEAS) is similar to our \emph{New-Solver Problem} (cf.~Definition~\ref{def:new-solver-problem}).
-The main difference is that Matricon et~al. compare a solver to another one, while we compare it to a set of existing solvers.
-To the best of our knowledge, we are the first to tackle the latter problem.
-%JB: again, I think there are too many details for describing this single approach
-%TF: unsure what to cut as the concrete approaches are also used later in the main section
-
-Given a pair of SAT solvers, Matricon et~al. want to iteratively select a subset of instances to decide with high confidence which of the two solvers is better.
+Given a pair of SAT solvers, Matricon et~al. iteratively select a subset of instances until the desired confidence level is reached to decide with high confidence which of the two solvers is better.
 The selection of instances is dependent on the choice of the solvers to distinguish.
 They calculate a scoring metric for all instances (that have not been selected yet), run the experiment with the highest score, and update the confidence.
-If the confidence exceeds a threshold (e.g., 95\%), they stop iterating.
-%Otherwise, the aforementioned steps are repeated until the desired confidence level is reached or the pool of instances is exhausted.
 Most of the proposed strategies in~\cite{MatriconAFSH21} are instance-based methods, i.e., no model of the solver runtimes is built.
-
-For the scoring metric, they discuss (1) a random strategy as a baseline, (2) a discrimination strategy based on the work of Gent et al.~\cite{GentHJKMNN14}, (3) a variance-based strategy (select instance with the highest variance in its distribution of running times), (4) an information-gain-based method, and (5) a feature-distance-based strategy.
-To update the confidence level after each experiment, they discuss using (1) a fixed subset size, (2) a Wilcoxon test, and (3) a distribution-based method.
-Their experiments show that random instance selection with a Wilcoxon test as the stopping criterion performs consistently well across different data sets in the sense that it belongs to the Pareto front or is close to it in all cases.
-%JB: in our evaluation, we should state whether our results are consistent with this or not
-%TF: will do
-
+The main difference is that Matricon et~al. compare a solver to another one, while we compare it to a set of existing solvers.
 Regarding our desired features of adaptive benchmark-set-selection (Table~\ref{tab:requirements}), the aforementioned approach already ticks off the first four features, i.e, it \emph{ranks} solvers, optimizes for \emph{runtime}, is \emph{feature-based}, and uses \emph{incremental non-random} sampling.
@@
 Moreover, the major shortcoming lies within it being not \emph{scalable} with the number solvers.
-As they compare only pairs of solvers, evaluating a new solver requires a non-constant number of benchmark sets, which in turn requires more runtime.
-%JB: maybe rather log-linear than quadratic, see a prior comment in intro
-%TF: changed
+As they compare only pairs of solvers, evaluating a new solver requires a non-constant number of benchmarks, which in turn requires more runtime.
+
+% MI: move to approach if needed there:
+%For the scoring metric, they discuss (1) a random strategy as a baseline, (2) a discrimination strategy based on the work of Gent et al.~\cite{GentHJKMNN14}, (3) a variance-based strategy (select instance with the highest variance in its distribution of running times), (4) an information-gain-based method, and (5) a feature-distance-based strategy.
+%To update the confidence level after each experiment, they discuss using (1) a fixed subset size, (2) a Wilcoxon test, and (3) a distribution-based method.
+%Their experiments show that random instance selection with a Wilcoxon test as the stopping criterion performs consistently well across different data sets in the sense that it belongs to the Pareto front or is close to it in all cases.
+%JB: in our evaluation, we should state whether our results are consistent with this or not
+%TF: will do
 
 \subsubsection{Active learning (AL).}
-%JB: section does not make clear which of the following approaches we will use, i.e., relationship to our work not clear (reads more like fundamentals than related work)
-The posed \emph{New-Solver Problem} has stark similarities to the well-studied field of active learning within recommender systems, especially the \emph{new-user problem}~\cite{RubensESK15}.
-On the one hand, we want to maximize the utility an instance provides to our model and, on the other hand, minimize the cost (CPU time) that is associated with its acquisition.
-%JB: not sure if this dilemma arises here as well; a dilemma is typical for searching for one optimal solution in continuous space, where there is an incentive to exploit neighborhoods in promising regions; however, if we find an interesting/useful (= good for assessing solver performance) instance in our approach, I'm not sure if we would continue picking similar instances ("exploiting" instead of "exploring") just because we expect them to be useful as well
-%TF: removed the exploration-exploitation dilemma
 
@@
 
-  \caption{Types of Learning (depiction inspired by Rubens~et.al.~\cite{RubensESK15})}
+  \caption{Types of Learning (Inspired by Rubens~et.al.~\cite{RubensESK15})}
   %JB: slightly larger font size in the figure (at least having the size of the caption) would be nice
@@
 \end{figure}
+
+%JB: section does not make clear which of the following approaches we will use, i.e., relationship to our work not clear (reads more like fundamentals than related work)
+The posed \emph{New-Solver Problem} has stark similarities to the well-studied field of active learning within recommender systems, especially the \emph{new-user problem}~\cite{RubensESK15}.
+On the one hand, we want to maximize the utility an instance provides to our model and, on the other hand, minimize the cost (CPU time) that is associated with its acquisition.
+%JB: not sure if this dilemma arises here as well; a dilemma is typical for searching for one optimal solution in continuous space, where there is an incentive to exploit neighborhoods in promising regions; however, if we find an interesting/useful (= good for assessing solver performance) instance in our approach, I'm not sure if we would continue picking similar instances ("exploiting" instead of "exploring") just because we expect them to be useful as well
+%TF: removed the exploration-exploitation dilemma
 
@@
 
+% MI: here we should limit this to the methods needed for presenting related work for AL in benchmarking:
 Pool-based methods can be further refined into \emph{Error-based} \cite{GolbandiKL11,KorenBV09}, \emph{Un\-cer\-tain\-ty-based} \cite{HarpaleY08,KapoorGUD07,TongK01}, \emph{Distribution-based} methods \cite{distribAL,SinhaED19} and \emph{Ensembles} \cite{KornerW06,MelvilleM04}.

### 2022-10-08 paper/paper.tex
Revise Introduction
... after changes by Tobias and Markus.

- remove resolved JB comments
- adapt wording

@@
 \label{sec:intro}
+
 One of the main phases of algorithm engineering is benchmarking.
@@
 While state-of-the-art solvers embark upon more and more problem instances, SAT solver evaluation is steadily getting more costly.
+%JB: not sure about the "while" here -- is the message that (1) evaluation gets costly *since* solvers have to solve more instances or is it that (2) solvers are threoretically able to solve more instances, but evaluation gets costlier anyway
 Competitive SAT solvers are the result of extensive experimentation and a variety of ideas and considerations~\cite{FroleyksHIJS21,sat2022}.
 While the field of SAT solver benchmarking is well established, traditional benchmark selection approaches do not optimize for benchmark runtime.
-The goal of traditional approaches is to select a representative set of instances for ranking solvers based on a scoring scheme~\cite{Gelder11,HoosKSS13}.
-SAT~Competitions typically use the widely adopted PAR-2 score, i.e, the average runtime with a penalty term of $2 \tau$ for timeouts with time-limit $\tau$~\cite{FroleyksHIJS21}.
-
-In this paper, we present a novel active learning based benchmark selection approach which is able to predict the rank of a new solver with high accuracy in only a fraction of the time needed to evaluate the full benchmark.
+Instead, the primary goal of traditional approaches is to select a representative set of instances for ranking solvers with a scoring scheme~\cite{Gelder11,HoosKSS13}.
+SAT~Competitions typically use the widely adopted PAR-2 score, i.e, the average runtime with a penalty of $2 \tau$ for timeouts with time limit~$\tau$~\cite{FroleyksHIJS21}.
+
+In this paper, we present a novel benchmark selection approach based on active learning.
+Our approach is able to predict the rank of a new solver with high accuracy in only a fraction of the time needed to evaluate the full benchmark.
 The problem we solve is specified in Definition~\ref{def:new-solver-problem}.
@@
 
-%We predict the solver ranking induced by this scoring, without predicting exact solver runtimes.
-
-%This allows for both comparing the solvers as well as identifying solvers with similar behavior if looking at individual predictions per instance.
-%JB: is there any further behavior we can deduce about a solver scoring apart from a performance ranking?
-"similar behavior" could IMHO be something like two solvers being fast/slow on the same instances, but that cannot be deduced from overall score
-%TF: made it clearer
-%MI: makes no sense in the context
-
 %MI: In the following I stripped out stuff which belongs to related work section. Stick to advertising own approach here.
-Our benchmarking process outputs a \emph{ranking} to compare the new solver against the existing set of solvers.
+Our benchmarking process outputs a \emph{ranking} to compare the new solver against the existing set of solvers, without predicting exact solver runtimes.
 We optimize the \emph{runtime} that our strategy needs to arrive at its conclusion.
@@
 Moreover, we select instances \emph{non-randomly} and \emph{incrementally}.
-By doing so, we have a greater impact on the properties of the benchmarking approach such as its required runtime.
-Sampling instances incrementally means taking runtime information from already done experiments into account when choosing the next.
-Rather than solely outputting a binary classification, i.e, the new solver is worse than solver A, we provide a \emph{scoring} function that shows by which margin a solver is worse and how similar it is to existing approaches.
-%JB: we should rename the "interpretability" criterion (make it more specific), term "interpretability" can have lots of different meanings and our specific meaning is not very intuitive
-%TF: Renamed it to "comparability"
-%MI: Renamed it to "scoring"
+In particular, we take runtime information from already done experiments into account when choosing the next.
+By doing so, we can control the properties of the benchmarking approach such as its required runtime.
+Rather than solely outputting a binary classification, i.e, the new solver is worse than existing solver A, we provide a \emph{scoring} function that shows by which margin a solver is worse and how similar it is to existing solvers.
 Our approach is \emph{scalable} in that it ranks a new solver $\hat{a}$ among any number of known solvers $\mathcal{A}$.
-%JB: for what do we measure cost here?
-%TF: made it clearer
-We avoid comparing pairs of solvers, which requires a non-constant number of benchmark sets.
-%JB: need to discuss this; logarithmic (compare one-vs rest) / log-linear (compare each pair) effort might be sufficient if the binary-search-/recursive-sorting-like procedure used
-%TF: "non-constant number of benchmark sets"
-Since a new solver idea is rarely best on the first try but requires more or less tweaking, it is desired to obtain a solver ranking fast.
+In particular, we only subsample the benchmark set once instead of comparing pairwise against each other solver~\cite{MatriconAFSH21}.
+Since a new solver idea is rarely best on the first try, it is desired to obtain a solver ranking fast.
 In this way, a new solver idea can be discarded if it performs poorly across the board or may be further tweaked if it shows promising results at least in some cases.
 
-To sum up our contribution, we take the SAT Competition~2022 Anniversary Track benchmark~\cite{sat2022} (5355 instances, full runtimes of 28 solvers), perform cross-validation by treating each solver as new once, and predict these solvers' PAR-2 rank with about \SI{92}{\%} accuracy with only about \SI{10}{\%} of the runtime that would be needed to evaluate these solvers on the complete set of instances.
-Both results are averages across all cross-validation iterations.
-%JB: I think we should expand a bit more on experimental design (how many instances/solvers, what do we compare), and the key results; I think it's okay if the intro becomes a bit longer (and in turn, the overview of the approach's features can be cut a bit)
-%TF: added a bit more information
-
-For reproducibility, all our source code\footnote{temporary, anonymized version for review: \url{xxx}} and and data\footnote{temporary, anonymized version for review: \url{xxx}} are available on GitHub.
+We evaluate our approach with the SAT Competition~2022 Anniversary Track benchmark~\cite{sat2022}, consisting of 5355~instances and full runtimes of 28~solvers.
+We perform cross-validation by treating each solver as new once, and predict these solvers' PAR-2 rank.
+On average, our predictions reach about \SI{92}{\%} accuracy with only about \SI{10}{\%} of the runtime that would be needed to evaluate these solvers on the complete set of instances.
+
+All our source code\footnote{temporary, anonymized version for review: \url{xxx}} and and data\footnote{temporary, anonymized version for review: \url{xxx}} are available on GitHub.
 

### 2022-10-08 paper/paper.tex
Revise Abstract
Only language. Also, get rid off the last line only containing one word.

@@
 \begin{abstract}
-  Benchmarking is one of the most important phases in the development of algorithms.
-  This also applies to solutions of the SAT problem (propositional sat\-is\-fia\-bi\-li\-ty).
+  Benchmarking is one of the most important phases when developing algorithms.
+  This also applies to solutions of the SAT (propositional satisfiability) problem.
   While the field of SAT solver benchmarking is well established, traditional benchmark selection approaches do not optimize for benchmark runtime.
-  Benchmark compilation selects representative instances from a pool of instances such that they reliably discriminate SAT solvers based on their runtime.
-  In this paper, we present an active learning based dynamic benchmark selection approach that predicts the rank of a new solver among its competitors with minimum running time and maximum rank prediction accuracy.
-  We evaluated this approach using the Anniversary track dataset from the 2022 SAT Competition. 
-  Our dynamic instance selection approach can predict the rank of a new solver after about \SI{10}{\%} of the time it would take to run the solver on all instances of this dataset, with a prediction accuracy of about \SI{92}{\%}.
+  Benchmark selection chooses representative instances from a pool of instances such that they reliably discriminate SAT solvers based on their runtime.
+  In this paper, we present a dynamic benchmark selection approach based on active learning.
+  Our approach predicts the rank of a new solver among its competitors with minimum running time and maximum rank prediction accuracy.
+  We evaluated this approach using the Anniversary Track dataset from the 2022 SAT Competition.
+  Our selection approach can predict the rank of a new solver after about \SI{10}{\%} of the time it would take to run the solver on all instances of this dataset, with a prediction accuracy of about \SI{92}{\%}.
   In this paper, we also discuss the importance of instance families in the selection process. 
-  Our tool provides a reliable way for SAT solver engineers to efficiently determine the performance of a new SAT solver.
+  Overall, our tool provides a reliable way for SAT solver engineers to efficiently determine the performance of a new SAT solver.
 

### 2022-10-08 paper/paper.tex
added two sections 3.1 and 3.2

@@
 \label{sec:main}
-Our framework follows a simple three-step schema (Alg.~\ref{algALBenchmark}).
-%JB: as the algorithm is rather simple and has different parts to be plugged in, I think we should rebrand it as "framework" rather than algorithm (and change the term consistently in text)
-%TF: changed
-After initializing the model $\mathcal{M}$, we incrementally \emph{select} a problem instance from the pool of instances, one at a time, evaluate the new solver $\hat{a}$ on it, and \emph{update} the model with the acquired result.
-We \emph{stop} when meeting some stopping condition.
-Since the potential runtime of experiments is by magnitudes larger than the model's update time, we only consider incrementing our benchmark set by one instance at a time rather than using batches.
-% For now, we also restrict ourselves to the aforementioned sequential one-instance-at-a-time process.
-At the core of the model $\mathcal{M}$ is a prediction function $f\!: \left(\mathcal{A} \cup \left\lbrace \hat{a} \right\rbrace\right) \times \mathcal{I} \rightarrow \mathbb{R}$ that powers the decisions within the previously mentioned three steps.
-%Note that the prediction target values may be also a subset of $\mathbb{R}$.
-%Refer to the next sections for more details.
+
+In this section, we present the details of our approach.
+Section~\ref{sec:main-framework} describes the general procedure.
+Section~\ref{sec:main-model} specifies the details of the underlying prediction model.
+Section~\ref{sec:main1} discusses how we may derive a solver ranking from that model.
+Possible criteria for sampling instances are discussed in Section~\ref{sec:main2}.
+Section~\ref{sec:main3} concludes with possible stopping conditions.
+
+\subsection{Incremental Benchmarking Framework}
+\label{sec:main-framework}
+
+Our framework follows a simple three-step schema.
+Algorithm~\ref{algALBenchmark} outlines the procedure. 
+While the stopping condition is not met (line~1), we select \emph{one} problem instance from the pool of instances (line~2) and evaluate the new solver $\hat{a}$ on it (line~3).
+We then use the acquired result to update the model (line~4).
+When the stopping criterion is met, we return the predicted score of the given solver $\hat{a}$.
 
 \begin{algorithm}
+% MI: comment in to rename to framework and remove numbering:
+%\SetAlgorithmName{Framework}{framework}{List of Frameworks}
+%\renewcommand{\thealgocf}{}
   \caption{Incremental Benchmarking Framework}
@@
 
-  \KwIn{Instances $\mathcal{I}$, Solvers $\mathcal{A}$, Runtimes $r\!: \mathcal{A} \times \mathcal{I} \rightarrow \left[0, \tau\right]$}
-  \KwIn{New Solver $\hat{a} \notin \mathcal{A}$}
-
-  $\mathcal{M} \leftarrow \textrm{init model with $\mathcal{A}$, $\mathcal{I}$, $r$, and $\hat{a}$}$
-
-  $\tilde{\mathcal{I}} \leftarrow \emptyset$
+  %\KwIn{Instances $\mathcal{I}$, Solvers $\mathcal{A}$, Runtimes $r\!: \mathcal{A} \times \mathcal{I} \rightarrow \left[0, \tau\right]$}
+  \KwIn{Model $\mathcal{M}$,  Solver $\hat{a}$}
+  \KwOut{Score of $\hat{a}$}
+
+  \BlankLine
+
+  %$\mathcal{M} \leftarrow \textrm{init model with $\mathcal{A}$, $\mathcal{I}$, $r$, and $\hat{a}$}$
+  %$\tilde{\mathcal{I}} \leftarrow \emptyset$
 
@@
 
-    $\tilde{\mathcal{I}} \leftarrow \tilde{\mathcal{I}} \cup \left\lbrace s \right\rbrace$
+    %$\tilde{\mathcal{I}} \leftarrow \tilde{\mathcal{I}} \cup \left\lbrace s \right\rbrace$
 
@@
 
-  $\operatorname{score}\!\left(\right) \leftarrow \operatorname{predictScores}\!\left(\mathcal{M}\right)$
-  %JB: bit confusing that this is not a numerical score, but a function (see next line); maybe call it "score()"?
-or return a tuple with all estimated scores (one per solver?); also, one could expect to get a benchmark set as result (see title of algorithm), but since instance selection is solver-specific, this is not case (it's rather a benchmarking schema)
-  %TF: renamed it to benchmarking framework and score()
-
-  \KwOut{Return tuple of scoring function $\operatorname{score}\!: \mathcal{A} \cup \left\lbrace \hat{a} \right\rbrace \rightarrow \mathbb{R}$ and the selected instances $\tilde{\mathcal{I}}$.}
-  %JB: we should formally specify the tuple rather than only describing it in words
-  %TF: made it clearer
+  \Return $\operatorname{predictScore}\!\left(\mathcal{M}, \hat{a}\right)$
+
+  %$\operatorname{score}\!\left(\right) \leftarrow \operatorname{predictScores}\!\left(\mathcal{M}\right)$
+  %\KwOut{Return tuple of scoring function $\operatorname{score}\!: \mathcal{A} \cup \left\lbrace \hat{a} \right\rbrace \rightarrow \mathbb{R}$ and the selected instances $\tilde{\mathcal{I}}$.}
 \end{algorithm}
 
-The following section (Section~\ref{sec:main1}) discusses how we may derive a solver ranking from our model (line~8).
-Thereafter, we look at possible sampling criteria (line~4) in Section~\ref{sec:main2} and, finally, possible stopping conditions (line~3) in Section~\ref{sec:main3}.
+\subsection{Solver Model}
+\label{sec:main-model}
+
+% MI: describe initialization and update of model
+
+Since the potential runtime of experiments is by magnitudes larger than the model's update time, we only consider incrementing our benchmark set by one instance at a time rather than using batches.
+At the core of the model $\mathcal{M}$ is a prediction function $f\!: \left(\mathcal{A} \cup \left\lbrace \hat{a} \right\rbrace\right) \times \mathcal{I} \rightarrow \mathbb{R}$ that powers the decisions within the previously mentioned three steps.
 

### 2022-10-08 paper/paper.tex
section3.3

@@
 
+Let benchmark instances $\mathcal{I}$, SAT solvers $\mathcal{A}$, and runtimes $r\!: \mathcal{A} \times \mathcal{I} \rightarrow \left[0, \tau\right]$ be given.
+
 Our framework follows a simple three-step schema.
@@
 
-\subsection{Predicting Rankings}
+\subsection{Runtime Preprocessing and Rank Prediction}
 \label{sec:main1}
-In order to predict a ranking of solver $\hat{a}$ amidst its competitors, we need to infer its performance on all instances, even the ones on which we have not evaluated it.
-To do so, there are generally two notions of this problem, i.e., regression and classification.
-A regression model tries to predict the \mbox{PAR-2} runtime of $\hat{a}$ on all instances based on the samples that have been selected.
-Then, the ranking is induced by the average predictions.
-Alternatively, a classification model can be used to predict some discrete runtime labels of $\hat{a}$ on all instances.
-Similarly, the average runtime label determines the ranking.
-
-Preliminary experiments, as well as research in portfolio solvers~\cite{NgokoCT19,CollauttiMMO13}, have shown however that directly predicting runtimes without any transformation does not work well in practice.
+
+%In order to predict a ranking of solver $\hat{a}$ amidst its competitors, we need to infer its performance on all instances, even the ones on which we have not evaluated it.
+%To do so, there are generally two notions of this problem, i.e, regression and classification.
+%A regression model tries to predict the \mbox{PAR-2} runtime of $\hat{a}$ on all instances based on the samples that have been selected.
+%Then, the ranking is induced by the average predictions.
+%Alternatively, a classification model can be used to predict some discrete runtime labels of $\hat{a}$ on all instances.
+%Similarly, the average runtime label determines the ranking.
+% MI: skipping this elaboration, since we do not use regression
+
+Preliminary experiments, as well as research in portfolio solvers, have shown however that directly predicting runtimes without any transformation does not work well in practice~\cite{NgokoCT19,CollauttiMMO13}.
 %JB: Collauti at el. still predict actual (though normalized) runtimes rather than discrete labels, they only heavily discretize runtimes when using them as features
 %TF: added "without any transformation"
-%In our experiments, the mean squared error was within the same magnitude as the values to be predicted.
+In our experiments, the mean squared error was within the same magnitude as the values to be predicted.
 In contrast to that, discretizing runtimes and using those labels in the model's training and prediction yields good results (cf.~Section~\ref{sec:eval}).
-The following paragraphs motivate why it makes sense to discretize runtimes.
+%The following paragraphs motivate why it makes sense to discretize runtimes.%MI: no further motivation found, commented out
 %JB: we should probably cut this a bit
@@
   %TF: it is used to resemble the 0-indexed implementation. I'm also not happy with this, but using \lbrace 0, \dots, k - 1 \rbrace does also look ugly.
+  %MI: Why is it so important to have this zero-based?
   \label{eq:rankingeq}
 \end{equation}
-with $w_{a,i} := 2$ (PAR-\emph{2} penalty factor) if $\gamma\!\left(a, i\right) = k$ else $w_{a,i} := 1$.
+with $w_{a,i} := 2$ if $\gamma\!\left(a, i\right) = k$ else $w_{a,i} := 1$.
 To be applicable to $\hat{a}$, we extend \emph{score} by replacing $\gamma\!\left(a, i\right)$ by the model's predictions $f\!: \left(\mathcal{A} \cup \left\lbrace \hat{a} \right\rbrace\right) \times \mathcal{I} \rightarrow \left\lbrace 1, \dots, k \right\rbrace$.
@@
 
-Turning a regression into a classification problem, discretization is a useful tool.
-However, it does not provide much value if the resulting cluster labels are not predictive of the PAR-2 ranking we seek to predict.
+Discretization would not provide much value if the resulting cluster labels were not predictive of the PAR-2 ranking we seek to predict.
 In our work, we use a hierarchical-clustering approach with a single-link criterion for $k = 3$.
@@
 % Besides that, we additionally consider using the past $l$ model predictions in a voting approach for stability.
+
+That discretized runtimes are sufficient for distinguishing solvers can also be shown by comparing the fraction of solver pairs whose runtimes are significantly different.
+Within the SAT~Competition~2022 Anniversary Track, about \SI{89.95}{\%} of solver pairs are significantly different (according to a Wilcoxon-signed-rank test with $\alpha = 0.05$).
+While only time-out labels (solver times out or not on a particular instance) are sufficient to significantly distinguish about \SI{87.04}{\%} of solver pairs, discrete-runtime labels are sufficient to distinguish about \SI{87.83}{\%} of solver pairs.
+%\footnote{Based on hierarchical clustering with a log-single-link criterion for $k = 3$.}
+This means that we do not lose much discriminatory power using discrete runtime labels.
+%The number of significant comparisons between solvers \emph{alone}, though, does only offer limited insights as we might achieve significant differences in all pairs by assigning some randomly skewed values as labels.
+%JB: thus, we should reorder this section to mention the argument "significant difference" (this paragraph) first and "deciding solver pairs" (previous paragraph, stronger argument) second (rather than giving the stronger argument first, making the second argument superfluous)
+%TF: Removed the previous sentence
+The predictiveness of the PAR-2 score, the labels' discriminatory power, as well as that it enables us to use classification, show the clear advantage of runtime discretization.
+%JB: at the same time with what? with the other benefits?
+%TF: removed redundant phrase
 
@@
 
-That discretized runtimes are sufficient for distinguishing solvers can also be shown by comparing the fraction of solver pairs whose runtimes are significantly different.
-Within the SAT~Competition~2022 Anniversary Track, about \SI{89.95}{\%} of solver pairs are significantly different (according to a Wilcoxon-signed-rank test with $\alpha = 0.05$).
-While only time-out labels (solver times out or not on a particular instance) are sufficient to significantly distinguish about \SI{87.04}{\%} of solver pairs, discrete-runtime labels are sufficient to distinguish about \SI{87.83}{\%} of solver pairs.
-%\footnote{Based on hierarchical clustering with a log-single-link criterion for $k = 3$.}
-This means that we do not lose much discriminatory power using discrete runtime labels.
-%The number of significant comparisons between solvers \emph{alone}, though, does only offer limited insights as we might achieve significant differences in all pairs by assigning some randomly skewed values as labels.
-%JB: thus, we should reorder this section to mention the argument "significant difference" (this paragraph) first and "deciding solver pairs" (previous paragraph, stronger argument) second (rather than giving the stronger argument first, making the second argument superfluous)
-%TF: Removed the previous sentence
-The predictiveness of the PAR-2 score, the labels' discriminatory power, as well as that it enables us to use classification, show the clear advantage of runtime discretization.
-%JB: at the same time with what? with the other benefits?
-%TF: removed redundant phrase
-
 \subsection{Instance Sampling}
 \label{sec:main2}
-Our instance-sampling strategies (also refer to Alg.~\ref{algALBenchmark}; line 4) can be divided into two categories.
+Our instance-sampling strategies (also refer to Algorithm~\ref{algALBenchmark}; line 4) can be divided into two categories.
 First, we present possible baseline approaches including existing work on instance-based selection strategies~\cite{Bossek021a,GentHJKMNN14,MatriconAFSH21}.
@@
 Similar to the instance-selection strategies, we can also divide stopping criteria into baseline and active-learning-based approaches.
-Also, refer to~Alg.~\ref{algALBenchmark}, line~3.
+Also, refer to~Algorithm~\ref{algALBenchmark}, line~3.
 While a fixed subset-size criterion, e.g., only sample \SI{20}{\%} of all available instances, is an example of the former, a stopping criterion based on a Wilcoxon signed-rank test or the convergence of the model's predictions is an example for the latter.
@@
 \end{equation}
-whereby $m$ denotes the count of iterations within Alg.~\ref{algALBenchmark} and $W_{\exp}^{\left(0\right)} := 1$.
+whereby $m$ denotes the count of iterations within Algorithm~\ref{algALBenchmark} and $W_{\exp}^{\left(0\right)} := 1$.
 
@@
 \label{sec:goal}
-As already stated in the introductory section, this work addresses the \emph{New-Solver Problem} (cf.~Def.~\ref{def:new-solver-problem}).
-By running Alg.~\ref{algALBenchmark}, we obtain a prediction model $\mathcal{M}$ that provides us with a scoring function \emph{score}.
+As already stated in the introductory section, this work addresses the \emph{New-Solver Problem} (cf.~Definition~\ref{def:new-solver-problem}).
+By running Algorithm~\ref{algALBenchmark}, we obtain a prediction model $\mathcal{M}$ that provides us with a scoring function \emph{score}.
 Also, it outputs the fraction of runtime that is needed to arrive at its conclusion.
@@
 \label{sec:evalframe}
-To evaluate a concrete instantiation of Alg.~\ref{algALBenchmark} (a concrete choice for all the sub-routines), we perform cross-validation on our set of solvers.
+To evaluate a concrete instantiation of Algorithm~\ref{algALBenchmark} (a concrete choice for all the sub-routines), we perform cross-validation on our set of solvers.
 That means that each solver once plays the role of the new solver.

### 2022-10-08 paper/paper.tex
Review Related Work
- remove resolved JB comments
- adapt language
- cut text (describe some approaches more shortly, remove some AL basics)

@@
 
-%Note that we know the runtimes of all solvers, except the new one, on all instances.
-%There also is work on collaborative filtering, where runtimes are only partially known, for algorithm selection~\cite{misir2017alors,misir2017data}.
-%MI: this belongs to discussion of related work
-
-%MI: In the following I stripped out stuff which belongs to related work section. Stick to advertising own approach here.
-Our benchmarking process outputs a \emph{ranking} to compare the new solver against the existing set of solvers, without predicting exact solver runtimes.
+Note that our scenario assumes that we know the runtimes of all solvers, except the new one, on all instances.
+One could also imagine a collaborative filtering scenario, where runtimes are only partially known~\cite{misir2017data,misir2017alors}.
+
+Our approach satisfies several desirable criteria for benchmarking:
+It outputs a \emph{ranking} to compare the new solver against the existing set of solvers.
+For this ranking, we do not even need to predict exact solver runtimes, which is trickier.
 We optimize the \emph{runtime} that our strategy needs to arrive at its conclusion.
@@
 Our approach is \emph{scalable} in that it ranks a new solver $\hat{a}$ among any number of known solvers $\mathcal{A}$.
-In particular, we only subsample the benchmark set once instead of comparing pairwise against each other solver~\cite{MatriconAFSH21}.
+In particular, we only subsample the benchmark once instead of comparing pairwise against each other solver~\cite{MatriconAFSH21}.
 Since a new solver idea is rarely best on the first try, it is desired to obtain a solver ranking fast.
@@
 
-Being of high interest in many fields, benchmarking is an active area of research on its own. 
-Recent studies show that benchmark compilation is a difficult endeavor for multiple reasons.
-Dehghani~et~al. describe how a biased benchmark can easily lead to fallacious interpretations~\cite{abs-2107-07002}.
-Nießl~et~al. state that benchmark compilation has many movable parts, e.g, performance measures, aggregation, and handling of missing values.
-Questionable research practices might modify those design elements a-posteriori to fit expectations, leading to biased results~\cite{NiesslHWCB22}.
-
-Benchmark compilation is an important issue for competitions, e.g., the SAT Competition.
+Benchmarking is not only of high interest in many fields but also an active research area on its own.
+Recent studies show that benchmark selection is a difficult endeavor for multiple reasons.
+Biased benchmarks can easily lead to fallacious interpretations~\cite{abs-2107-07002}.
+Also, benchmark selection has many movable parts, e.g., performance measures, aggregation, and handling of missing values.
+Questionable research practices might modify these elements a-posteriori to fit expectations, leading to biased results~\cite{NiesslHWCB22}.
+
+\subsubsection{Static benchmark selection.}
+
+Benchmark selection is an important issue for competitions, e.g., the SAT Competition.
 In such competitions, the organizers define the rules for composing the corresponding benchmarks.
@@
 Manthey and Möhle find that competition benchmarks might contain redundant instances, and propose a feature-based approach to remove redundancy~\cite{manthey2016better}.
-M{\i}s{\i}r presents a feature-based approach to reduce benchmark sets by matrix factorization and clustering~\cite{misir2021benchmark}.
-
-%MI: one paragraph for Hoos
-Hoos et al. discuss which properties are most desirable when selecting a set of SAT benchmark instances~\cite{HoosKSS13}.
+M{\i}s{\i}r presents a feature-based approach to reduce benchmarks by matrix factorization and clustering~\cite{misir2021benchmark}.
+
+Hoos et al. discuss which properties are most desirable when selecting SAT benchmark instances~\cite{HoosKSS13}.
 The selection strategy presented is static, i.e., it does not depend on particular solvers to distinguish.
-Desirable properties are a variety of instances to avoid over-fitting, adapted instance hardness (not too easy but also not too hard), and avoidance of duplicate instances.
-To avoid having too similar instances, they make use of a distance-based approach using the SATzilla features~\cite{XuHHL08,features}.
+Selection criteria are a instance variety to avoid over-fitting, adapted instance hardness (not too easy but also not too hard), and avoiding duplicate instances.
+To filter too similar instances, they use a distance-based approach with the SATzilla features~\cite{XuHHL08,features}.
 Their approach \emph{ranks} solvers, is \emph{feature-based}, and \emph{scalable} in the sense that the evaluation time scales linearly with more instances and solvers.
-Also, solvers are \emph{comparable} within their approach in the sense that the obtained PAR-2 score induces a distance between solvers.
-However, they do not optimize for benchmark \emph{runtime} and select instances \emph{randomly}, apart from constraints on the instance hardness and feature distance.
-
-%Our approach should optimally also have these three properties, i.e., a variety of instances without duplicates that are neither too easy nor too hard.
-%While this selection strategy is useful for creating diverse benchmark sets for competitions, it does not optimize for runtime.
-%It is expected that dynamic selection strategies outperform static ones regarding the runtime of experiments. 
-%Dynamic strategies chose instances incrementally, one at a time, by integrating runtime data (of the instances that the solver has been already evaluated on) to make subsequent selection decisions rather than completely choosing the benchmark set upfront.
+Also, it allows \emph{scoring} and thereby comparing solvers.
+However, the approach does not optimize for benchmark \emph{runtime} and selects instances \emph{randomly}, apart from constraints on the instance hardness and feature distance.
 
@@
   \begin{tabular}{
-    m{0.3\textwidth}
-    >{\centering\arraybackslash}m{0.15\textwidth}
-    >{\centering\arraybackslash}m{0.15\textwidth}
-    >{\centering\arraybackslash}m{0.15\textwidth}
-    >{\centering\arraybackslash}m{0.15\textwidth}
+    m{0.32\textwidth}
+    >{\centering\arraybackslash}m{0.16\textwidth}
+    >{\centering\arraybackslash}m{0.16\textwidth}
+    >{\centering\arraybackslash}m{0.16\textwidth}
+    >{\centering\arraybackslash}m{0.16\textwidth}
   }
@@
     % Feature-based & \cmark & \cmark & \cmark & \cmark \\
-    Incremental Sampling & \xmark & \xmark & \cmark & \cmark \\
+    Incremental/Non-Random & \xmark & \xmark & \cmark & \cmark \\
     % \tablefootnote{While using a model-based selection strategy to select the next configuration to evaluate, the instance sampling is done randomly.}
@@
   ~\\[1em]
-  \caption{Support for desired features that are introduced in Section~\ref{sec:intro}.
-	The shown approaches include traditional benchmark-set-selection techniques~\cite{HoosKSS13}, an algorithm configuration system~\cite{HutterHL11}, an existing active-learning approach~\cite{MatriconAFSH21}, as well as our approach.
+  \caption{Support for desired criteria from Section~\ref{sec:intro}.
+	The table compares a static benchmark-selection approach~\cite{HoosKSS13}, an algorithm configuration system~\cite{HutterHL11}, an existing active-learning approach~\cite{MatriconAFSH21}, as well as our approach.
   }
@@
 
-% MI: one paragraph for SMAC
-Another interesting approach can be found within algorithm configuration systems~\cite{HutterHL11,HoosHL21,Stutzle0P22}.
-The goal is to tune SAT solvers for a given sub-domain of problem instances.
-Although this task is different from our goal, e.g., we do not need to navigate configuration space, they use a similar framework.
-There are mainly two similarities.
-First, both approaches are model-based.
-While SMAC tries to gradually build a model of the solver's performance as a function of the configuration parameters, we want to gradually build a model of a new solver's performance based on already-known runtimes of instances and runtime information on already-known solvers.
-Second, both approaches employ some kind of selection strategy that is based on the previously mentioned model.
-Theirs selects configurations rather than instances though.
-Furthermore, Hutter et.
-al.~\cite{HutterHL11} use an approach that makes use of exploitation (what configuration parameters are already good) and exploration (parameter space with few data points) to select the next configuration for which to acquire runtime information.
-SMAC's configuration sampling has also proven to be better than random sampling in all tested scenarios and has even proven to be significantly better in most cases~\cite{HutterHL11}.
-While this approach considers the configuration's \emph{runtime}, is \emph{feature-based}, and \emph{scales linearly} by having a fixed time budget, it is not capable of the other three features listed in Table~\ref{tab:requirements}.
+\subsubsection{SAT solver configuration.}
+
+Further related work can be found within the field of algorithm configuration~\cite{HoosHL21,Stutzle0P22}, e.g., the configuration system SMAC~\cite{HutterHL11}.
+There, the goal is to tune SAT solvers for a given sub-domain of problem instances.
+Although this task is different from our goal, e.g., we do not need to navigate configuration space, there are similarities to our approach as well.
+For example, SMAC also employs an iterative, model-based selection procedure, though for configurations rather than instances.
+SMAC's configuration sampling has also proven to be better than random sampling in all tested scenarios and has even proven to be significantly better in most cases.
+While this approach considers the configuration's \emph{runtime}, is \emph{feature-based}, and \emph{scales} by having a fixed time budget, it does not satisfy the other three criteria listed in Table~\ref{tab:requirements}.
 First, an algorithm configurator cannot be used to \emph{rank} a new solver since it only looks at promising solvers/configurations rather than the overall average performance.
 Second, while using a model-based selection strategy to sample configurations, instance sampling is done \emph{randomly}, i.e., without building a model over instances.
-And third, configurations within their approach are not \emph{comparable} since algorithm configuration seeks to find solemnly the best performing configuration.
-%JB: well, the internal meta-model of the configurator allows to estimate the solver performance for particular configurations (rather than solvers), so the approach is kind of interpretable, though regarding configurations (while for comparison to other solvers, only top-configuration performance should be interesting, so the approach is indeed not interpretable regarding solvers)
-%TF: changed
-
-% MI: one paragraph for Matricon
-Matricon~et~al. present an incremental benchmark instance selection approach~\cite{MatriconAFSH21}.
+And third, only few configurations within their approach are \emph{scored} since algorithm configuration seeks to find solemnly the best performing configuration.
+
+\subsubsection{Incremental benchmark selection.}
+
+Matricon~et~al. present an incremental benchmark selection approach~\cite{MatriconAFSH21}.
 Their \emph{per-set efficient algorithm selection problem} (PSEAS) is similar to our \emph{New-Solver Problem} (cf.~Definition~\ref{def:new-solver-problem}).
-Given a pair of SAT solvers, Matricon et~al. iteratively select a subset of instances until the desired confidence level is reached to decide with high confidence which of the two solvers is better.
+Given a pair of SAT solvers, they iteratively select a subset of instances until the desired confidence level is reached to decide which of the two solvers is better.
 The selection of instances is dependent on the choice of the solvers to distinguish.
-They calculate a scoring metric for all instances (that have not been selected yet), run the experiment with the highest score, and update the confidence.
-Most of the proposed strategies in~\cite{MatriconAFSH21} are instance-based methods, i.e., no model of the solver runtimes is built.
-The main difference is that Matricon et~al. compare a solver to another one, while we compare it to a set of existing solvers.
-Regarding our desired features of adaptive benchmark-set-selection (Table~\ref{tab:requirements}), the aforementioned approach already ticks off the first four features, i.e, it \emph{ranks} solvers, optimizes for \emph{runtime}, is \emph{feature-based}, and uses \emph{incremental non-random} sampling.
-Although they provide a confidence metric for their prediction, which makes solvers partially \emph{comparable}, it is not clear how similar two given solvers are or on which instances they behave similarly, as the prediction is solely binary.
-Moreover, the major shortcoming lies within it being not \emph{scalable} with the number solvers.
-As they compare only pairs of solvers, evaluating a new solver requires a non-constant number of benchmarks, which in turn requires more runtime.
+They calculate a scoring metric for all unselected instances, run the experiment with the highest score, and update the confidence.
+Most of the proposed strategies in~\cite{MatriconAFSH21} are instance-based, i.e., no model of the solver runtimes is built.
+Regarding our desired criteria of benchmark selection (Table~\ref{tab:requirements}), the aforementioned approach \emph{ranks} solvers, optimizes for \emph{runtime}, is \emph{feature-based}, and uses \emph{incremental non-random} sampling.
+However, the approach only compares solvers binarily rather than providing a \emph{scoring}.
+Thus, it is not clear how similar two given solvers are or on which instances they behave similarly.
+Moreover, a major shortcoming lies is the lacking \emph{scalability} with the number solvers.
+Comparing only pairs of solvers, evaluating a new solver requires sampling several benchmarks.
+In contrast, our approach allows comparing a new solver against a set of existing solvers with one benchmark.
 
@@
 
-\begin{figure}[tbp!]
-  \centering
-  \begin{tabular}[c]{ccc}
-  \begin{subfigure}[b]{0.25\textwidth}
-  \centering
-  \resizebox{!}{2.75cm}{
-  \begin{tikzpicture}
-  \draw[fill=black]  (-2.8,3.7) ellipse (0.1 and 0.1);
-  \draw[fill=black]  (-2.6,3.4) arc (0:180:0.2);
-  \node[right] at (-5.9,3.6) {\large Agent (Solver)};
-  \draw[fill=white]  (-3.3,2.8) rectangle (-2.5,2.2);
-  \draw[fill=white]  (-3.2,2.7) rectangle (-2.4,2.1);
-  \draw[fill=white]  (-3.1,2.6) rectangle (-2.3,2);
-  \node[right] at (-5.9,2.4) {\large Train. Data};
-  \node (v1) at (-2.8,3.4) {};
-  \node (v2) at (-2.8,2.8) {};
-  \draw[->]  (v1) edge (v2);
-  \node[right] at (-2.8,3.1) {\large $x, y$};
-  \node (v3) at (-2.8,2) {};
-  \node (v4) at (-2.8,1.4) {};
-  \draw[->]  (v3) edge (v4);
-  \draw  (-3.1,1.4) rectangle (-2.5,1);
-  \draw  (-2.4,3.2) rectangle (-2.4,3.2);
-  \node[right] at (-5.9,1.2) {\large Learning Alg.};
-  \node (v5) at (-2.8,1) {};
-  \node (v6) at (-2.8,0.4) {};
-  \draw[->]  (v5) edge (v6);
-  \node[right] at (-5.9,0.25) {\large Approx. Func.};
-  \node at (-2.8,0.25) {\large $f$};
-  \end{tikzpicture}
-  }
-  \caption{Passive}
-  \label{fig:passive}
-  \end{subfigure}
-  &
-  \begin{subfigure}[b]{0.35\textwidth}
-  \centering
-  \resizebox{!}{2.75cm}{
-  \begin{tikzpicture}
-  \draw[fill=black]  (-2.8,3.7) ellipse (0.1 and 0.1);
-  \draw[fill=black]  (-2.6,3.4) arc (0:180:0.2);
-  \node[right] at (-5.9,3.6) {\large Agent (Solver)};
-  \draw[fill=white]  (-3.3,2.8) rectangle (-2.5,2.2);
-  \draw[fill=white]  (-3.2,2.7) rectangle (-2.4,2.1);
-  \draw[fill=white]  (-3.1,2.6) rectangle (-2.3,2);
-  \node[right] at (-5.9,2.4) {\large Train. Data};
-  \node (v1) at (-2.8,3.4) {};
-  \node (v2) at (-2.8,2.8) {};
-  \draw[->]  (v1) edge (v2);
-  \node[right] at (-2.8,3.1) {\large $x, y$};
-  \node (v3) at (-2.8,2) {};
-  \node (v4) at (-2.8,1.4) {};
-  \draw[->]  (v3) edge (v4);
-  \draw  (-3.1,1.4) rectangle (-2.5,1);
-  \draw  (-2.4,3.2) rectangle (-2.4,3.2);
-  \node[right] at (-5.9,1.2) {\large Learning Alg.};
-  \node (v5) at (-2.8,1) {};
-  \node (v6) at (-2.8,0.4) {};
-  \draw[->]  (v5) edge (v6);
-  \node[right] at (-5.9,0.25) {\large Approx. Func.};
-  \node at (-2.8,0.25) {\large $f$};
-  \node[right] at (-1.2,2.4) {\large Pool};
-  \draw[->, densely dashed] (-2.4,1.2) arc (-90:0:0.8);
-  \draw[->, densely dashed] (-1.6,2.8) arc (0:90:0.8);
-  \node[right] at (-1.6,3.135) {\large $x, ?$};
-  \draw  (-1.6,2.4) ellipse (0.4 and 0.3);
-  \draw[fill=black]  (-1.725,2.5) ellipse (0.07 and 0.07);
-  \draw[fill=black]  (-1.625,2.25) ellipse (0.07 and 0.07);
-  \draw[fill=black]  (-1.425,2.4) ellipse (0.07 and 0.07);
-  \end{tikzpicture}
-  }
-  \caption{Active Pool-based}
-  \label{fig:activepool}
-  \end{subfigure}
-  &
-  \begin{subfigure}[b]{0.35\textwidth}
-  \centering
-  \resizebox{!}{2.75cm}{
-  \begin{tikzpicture}
-  \draw[fill=black]  (-2.8,3.7) ellipse (0.1 and 0.1);
-  \draw[fill=black]  (-2.6,3.4) arc (0:180:0.2);
-  \node[right] at (-5.9,3.6) {\large Agent (Solver)};
-  \draw[fill=white]  (-3.3,2.8) rectangle (-2.5,2.2);
-  \draw[fill=white]  (-3.2,2.7) rectangle (-2.4,2.1);
-  \draw[fill=white]  (-3.1,2.6) rectangle (-2.3,2);
-  \node[right] at (-5.9,2.4) {\large Train. Data};
-  \node (v1) at (-2.8,3.4) {};
-  \node (v2) at (-2.8,2.8) {};
-  \draw[->]  (v1) edge (v2);
-  \node[right] at (-2.8,3.1) {\large $x, y$};
-  \node (v3) at (-2.8,2) {};
-  \node (v4) at (-2.8,1.4) {};
-  \draw[->]  (v3) edge (v4);
-  \draw  (-3.1,1.4) rectangle (-2.5,1);
-  \draw  (-2.4,3.2) rectangle (-2.4,3.2);
-  \node[right] at (-5.9,1.2) {\large Learning Alg.};
-  \node (v5) at (-2.8,1) {};
-  \node (v6) at (-2.8,0.4) {};
-  \draw[->]  (v5) edge (v6);
-  \node[right] at (-5.9,0.25) {\large Approx. Func.};
-  \node at (-2.8,0.25) {\large $f$};
-  \node[right, label={[align=left]\large Gen.\\\large Model}] at (-0.6,1.85) {};
-  %JB: does the model need to be a GAN or can it also be a different type of model?
-  %TF: not necessarily, changed it
-  \node[right] at (-1.3,3.135) {\large $x, ?$};
-  \draw[->, densely dashed] (-2.4,1.2) arc (-90:90:1.2);
-  \end{tikzpicture}
-  }
-  \caption{Active Synthesis-based}
-  \label{fig:activesynth}
-  \end{subfigure}
-  \end{tabular}
-
-  \caption{Types of Learning (Inspired by Rubens~et.al.~\cite{RubensESK15})}
-  %JB: slightly larger font size in the figure (at least having the size of the caption) would be nice
-  %TF: made it larger
-  \label{fig:learning}
-\end{figure}
-
-%JB: section does not make clear which of the following approaches we will use, i.e., relationship to our work not clear (reads more like fundamentals than related work)
-The posed \emph{New-Solver Problem} has stark similarities to the well-studied field of active learning within recommender systems, especially the \emph{new-user problem}~\cite{RubensESK15}.
+%\begin{figure}[tbp!]
+%  \centering
+%  \begin{tabular}[c]{ccc}
+%  \begin{subfigure}[b]{0.25\textwidth}
+%  \centering
+%  \resizebox{!}{2.75cm}{
+%  \begin{tikzpicture}
+%  \draw[fill=black]  (-2.8,3.7) ellipse (0.1 and 0.1);
+%  \draw[fill=black]  (-2.6,3.4) arc (0:180:0.2);
+%  \node[right] at (-5.9,3.6) {\large Agent (Solver)};
+%  \draw[fill=white]  (-3.3,2.8) rectangle (-2.5,2.2);
+%  \draw[fill=white]  (-3.2,2.7) rectangle (-2.4,2.1);
+%  \draw[fill=white]  (-3.1,2.6) rectangle (-2.3,2);
+%  \node[right] at (-5.9,2.4) {\large Train. Data};
+%  \node (v1) at (-2.8,3.4) {};
+%  \node (v2) at (-2.8,2.8) {};
+%  \draw[->]  (v1) edge (v2);
+%  \node[right] at (-2.8,3.1) {\large $x, y$};
+%  \node (v3) at (-2.8,2) {};
+%  \node (v4) at (-2.8,1.4) {};
+%  \draw[->]  (v3) edge (v4);
+%  \draw  (-3.1,1.4) rectangle (-2.5,1);
+%  \draw  (-2.4,3.2) rectangle (-2.4,3.2);
+%  \node[right] at (-5.9,1.2) {\large Learning Alg.};
+%  \node (v5) at (-2.8,1) {};
+%  \node (v6) at (-2.8,0.4) {};
+%  \draw[->]  (v5) edge (v6);
+%  \node[right] at (-5.9,0.25) {\large Approx. Func.};
+%  \node at (-2.8,0.25) {\large $f$};
+%  \end{tikzpicture}
+%  }
+%  \caption{Passive}
+%  \label{fig:passive}
+%  \end{subfigure}
+%  &
+%  \begin{subfigure}[b]{0.35\textwidth}
+%  \centering
+%  \resizebox{!}{2.75cm}{
+%  \begin{tikzpicture}
+%  \draw[fill=black]  (-2.8,3.7) ellipse (0.1 and 0.1);
+%  \draw[fill=black]  (-2.6,3.4) arc (0:180:0.2);
+%  \node[right] at (-5.9,3.6) {\large Agent (Solver)};
+%  \draw[fill=white]  (-3.3,2.8) rectangle (-2.5,2.2);
+%  \draw[fill=white]  (-3.2,2.7) rectangle (-2.4,2.1);
+%  \draw[fill=white]  (-3.1,2.6) rectangle (-2.3,2);
+%  \node[right] at (-5.9,2.4) {\large Train. Data};
+%  \node (v1) at (-2.8,3.4) {};
+%  \node (v2) at (-2.8,2.8) {};
+%  \draw[->]  (v1) edge (v2);
+%  \node[right] at (-2.8,3.1) {\large $x, y$};
+%  \node (v3) at (-2.8,2) {};
+%  \node (v4) at (-2.8,1.4) {};
+%  \draw[->]  (v3) edge (v4);
+%  \draw  (-3.1,1.4) rectangle (-2.5,1);
+%  \draw  (-2.4,3.2) rectangle (-2.4,3.2);
+%  \node[right] at (-5.9,1.2) {\large Learning Alg.};
+%  \node (v5) at (-2.8,1) {};
+%  \node (v6) at (-2.8,0.4) {};
+%  \draw[->]  (v5) edge (v6);
+%  \node[right] at (-5.9,0.25) {\large Approx. Func.};
+%  \node at (-2.8,0.25) {\large $f$};
+%  \node[right] at (-1.2,2.4) {\large Pool};
+%  \draw[->, densely dashed] (-2.4,1.2) arc (-90:0:0.8);
+%  \draw[->, densely dashed] (-1.6,2.8) arc (0:90:0.8);
+%  \node[right] at (-1.6,3.135) {\large $x, ?$};
+%  \draw  (-1.6,2.4) ellipse (0.4 and 0.3);
+%  \draw[fill=black]  (-1.725,2.5) ellipse (0.07 and 0.07);
+%  \draw[fill=black]  (-1.625,2.25) ellipse (0.07 and 0.07);
+%  \draw[fill=black]  (-1.425,2.4) ellipse (0.07 and 0.07);
+%  \end{tikzpicture}
+%  }
+%  \caption{Active Pool-based}
+%  \label{fig:activepool}
+%  \end{subfigure}
+%  &
+%  \begin{subfigure}[b]{0.35\textwidth}
+%  \centering
+%  \resizebox{!}{2.75cm}{
+%  \begin{tikzpicture}
+%  \draw[fill=black]  (-2.8,3.7) ellipse (0.1 and 0.1);
+%  \draw[fill=black]  (-2.6,3.4) arc (0:180:0.2);
+%  \node[right] at (-5.9,3.6) {\large Agent (Solver)};
+%  \draw[fill=white]  (-3.3,2.8) rectangle (-2.5,2.2);
+%  \draw[fill=white]  (-3.2,2.7) rectangle (-2.4,2.1);
+%  \draw[fill=white]  (-3.1,2.6) rectangle (-2.3,2);
+%  \node[right] at (-5.9,2.4) {\large Train. Data};
+%  \node (v1) at (-2.8,3.4) {};
+%  \node (v2) at (-2.8,2.8) {};
+%  \draw[->]  (v1) edge (v2);
+%  \node[right] at (-2.8,3.1) {\large $x, y$};
+%  \node (v3) at (-2.8,2) {};
+%  \node (v4) at (-2.8,1.4) {};
+%  \draw[->]  (v3) edge (v4);
+%  \draw  (-3.1,1.4) rectangle (-2.5,1);
+%  \draw  (-2.4,3.2) rectangle (-2.4,3.2);
+%  \node[right] at (-5.9,1.2) {\large Learning Alg.};
+%  \node (v5) at (-2.8,1) {};
+%  \node (v6) at (-2.8,0.4) {};
+%  \draw[->]  (v5) edge (v6);
+%  \node[right] at (-5.9,0.25) {\large Approx. Func.};
+%  \node at (-2.8,0.25) {\large $f$};
+%  \node[right, label={[align=left]\large Gen.\\\large Model}] at (-0.6,1.85) {};
+%  \node[right] at (-1.3,3.135) {\large $x, ?$};
+%  \draw[->, densely dashed] (-2.4,1.2) arc (-90:90:1.2);
+%  \end{tikzpicture}
+%  }
+%  \caption{Active Synthesis-based}
+%  \label{fig:activesynth}
+%  \end{subfigure}
+%  \end{tabular}
+%  
+%  \caption{Types of Learning (Inspired by Rubens~et.al.~\cite{RubensESK15})}
+%  \label{fig:learning}
+%\end{figure}
+
+The posed \emph{New-Solver Problem} has stark similarities to the well-studied field of active learning AL within recommender systems, especially the \emph{new-user problem}~\cite{RubensESK15}.
 On the one hand, we want to maximize the utility an instance provides to our model and, on the other hand, minimize the cost (CPU time) that is associated with its acquisition.
-%JB: not sure if this dilemma arises here as well; a dilemma is typical for searching for one optimal solution in continuous space, where there is an incentive to exploit neighborhoods in promising regions; however, if we find an interesting/useful (= good for assessing solver performance) instance in our approach, I'm not sure if we would continue picking similar instances ("exploiting" instead of "exploring") just because we expect them to be useful as well
-%TF: removed the exploration-exploitation dilemma
-
-% Starting with only unlabeled data, e.g., a new user has not yet rated any content, an active-learning algorithm iteratively shows the items that are best suited to refine the user's model based on his previous interaction.
-% In contrast to traditional passive machine-learning methods (Fig.~\ref{fig:passive}), active-learning approaches can actively select for which items to acquire labels~\cite{RubensESK15}.
-% Since the potential runtime of experiments is by magnitudes larger than the model's training time, we only consider incrementing our benchmark set by one instance at a time rather than using batches which is also proposed in current active-learning advances~\cite{SinhaED19,TranDRC19}.
-
+In contrast to traditional passive machine-learning methods with given instance labels, active learning allows selecting instances for which to acquire labels.
 AL algorithms can be categorized into \emph{synthesis-bas\-ed} \cite{0001AEMN22,GarzonMG22,2019gaal} and \emph{pool-bas\-ed} approaches \cite{GolbandiKL11,HarpaleY08,KorenBV09}.%SinhaED19,distribAL,TongK01,KapoorGUD07,KornerW06,MelvilleM04}.
-%JB: we can probably cut some citations, the three most important should be sufficient
-%TF: done
-While synthesis-based methods (Fig.
-\ref{fig:activesynth}) generate instances that might be interesting to label, pool-based methods (Fig.
-\ref{fig:activepool}) rely on a pool of unlabeled instances from which to draw items.
-
-Recent synthesis-based methods within the field of satisfiability solving show how to generate problem instances with desired properties.
+While synthesis-based methods generate instances for labeling, pool-based methods rely on a fixed set of unlabeled instances from which to sample.
+
+Recent synthesis-based methods within the field of SAT solving show how to generate problem instances with desired properties.
 This goal is, however, orthogonal to ours~\cite{0001AEMN22,GarzonMG22}.
-While those approaches want to generate problem instances on which a solver is good or bad, we want to predict whether a solver is good or bad on an existing benchmark set.
+While those approaches want to generate problem instances on which a solver is good or bad, we want to predict whether a solver is good or bad on an existing benchmark.
+Volpato and Guangyan use pool-based AL to learn an instance-specific algorithm selector~\cite{volpato2019active}.
+Rather than benchmarking a solver's overall performance, the goal is to recommend the best solver out of a set of solvers for each SAT instance.
 
 % MI: here we should limit this to the methods needed for presenting related work for AL in benchmarking:
-Pool-based methods can be further refined into \emph{Error-based} \cite{GolbandiKL11,KorenBV09}, \emph{Un\-cer\-tain\-ty-based} \cite{HarpaleY08,KapoorGUD07,TongK01}, \emph{Distribution-based} methods \cite{distribAL,SinhaED19} and \emph{Ensembles} \cite{KornerW06,MelvilleM04}.
-Error-based strategies pick instances that have the most influence on the model parameters.
-Uncertainty-based methods sample instances close to their perceived decision boundary.
-Distribution-based approaches sample instances whose features are unaccounted for amongst the already sampled instances.
-Finally, ensembles are a meta-strategy that combines the predictions of several different active-learning models into a selection choice.
-
-In the field of SAT-solving, Volpato and Guangyan use uncertainty-sampling-based AL to learn an instance-specific algorithm selector~\cite{volpato2019active}.
-Rather than benchmarking a solver's overall performance, the goal is to recommend the best solver out of a set of solvers for each SAT instance.
+%JB: should probably just put the citation there where we explain the strategies (in Section 3)
+%Pool-based methods can be further refined into \emph{Error-based} \cite{GolbandiKL11,KorenBV09}, \emph{Un\-cer\-tain\-ty-based} \cite{HarpaleY08,KapoorGUD07,TongK01}, \emph{Distribution-based} methods \cite{distribAL,SinhaED19} and \emph{Ensembles} \cite{KornerW06,MelvilleM04}.
+%Error-based strategies pick instances that have the most influence on the model parameters.
+%Uncertainty-based methods sample instances close to their perceived decision boundary.
+%Distribution-based approaches sample instances whose features are unaccounted for amongst the already sampled instances.
+%Finally, ensembles are a meta-strategy that combines the predictions of several different active-learning models into a selection choice.
 
@@
 
-Since the potential runtime of experiments is by magnitudes larger than the model's update time, we only consider incrementing our benchmark set by one instance at a time rather than using batches.
+Since the potential runtime of experiments is by magnitudes larger than the model's update time, we only consider incrementing our benchmark by one instance at a time rather than using batches which is also proposed in current active-learning advances~\cite{SinhaED19,2019gaal}.
+
 At the core of the model $\mathcal{M}$ is a prediction function $f\!: \left(\mathcal{A} \cup \left\lbrace \hat{a} \right\rbrace\right) \times \mathcal{I} \rightarrow \mathbb{R}$ that powers the decisions within the previously mentioned three steps.

### 2022-10-09 paper/paper.tex
Remove some resolved JB comments in Sections 3 & 4
No impact on visible output.
Fix grammar and spelling in sections 1 to 4

@@
 By doing so, we can control the properties of the benchmarking approach such as its required runtime.
-Rather than solely outputting a binary classification, i.e, the new solver is worse than existing solver A, we provide a \emph{scoring} function that shows by which margin a solver is worse and how similar it is to existing solvers.
+Rather than solely outputting a binary classification, i.e, the new solver is worse than an existing solver, we provide a \emph{scoring} function that shows by which margin a solver is worse and how similar it is to existing solvers.
 Our approach is \emph{scalable} in that it ranks a new solver $\hat{a}$ among any number of known solvers $\mathcal{A}$.
@@
 We evaluate our approach with the SAT Competition~2022 Anniversary Track benchmark~\cite{sat2022}, consisting of 5355~instances and full runtimes of 28~solvers.
-We perform cross-validation by treating each solver as new once, and predict these solvers' PAR-2 rank.
+We perform cross-validation by treating each solver as new once and predicting these solvers' PAR-2 rank.
 On average, our predictions reach about \SI{92}{\%} accuracy with only about \SI{10}{\%} of the runtime that would be needed to evaluate these solvers on the complete set of instances.
@@
 The selection strategy presented is static, i.e., it does not depend on particular solvers to distinguish.
-Selection criteria are a instance variety to avoid over-fitting, adapted instance hardness (not too easy but also not too hard), and avoiding duplicate instances.
+Selection criteria are instance variety to avoid over-fitting, adapted instance hardness (not too easy but also not too hard), and avoiding duplicate instances.
 To filter too similar instances, they use a distance-based approach with the SATzilla features~\cite{XuHHL08,features}.
@@
 Thus, it is not clear how similar two given solvers are or on which instances they behave similarly.
-Moreover, a major shortcoming lies is the lacking \emph{scalability} with the number solvers.
+Moreover, a major shortcoming is the lacking \emph{scalability} with the number of solvers.
 Comparing only pairs of solvers, evaluating a new solver requires sampling several benchmarks.
@@
 On the one hand, we want to maximize the utility an instance provides to our model and, on the other hand, minimize the cost (CPU time) that is associated with its acquisition.
-In contrast to traditional passive machine-learning methods with given instance labels, active learning allows selecting instances for which to acquire labels.
+In contrast to traditional passive machine-learning methods with given instance labels, active learning allows for selecting instances for which to acquire labels.
 AL algorithms can be categorized into \emph{synthesis-bas\-ed} \cite{0001AEMN22,GarzonMG22,2019gaal} and \emph{pool-bas\-ed} approaches \cite{GolbandiKL11,HarpaleY08,KorenBV09}.%SinhaED19,distribAL,TongK01,KapoorGUD07,KornerW06,MelvilleM04}.
@@
 Preliminary experiments, as well as research in portfolio solvers, have shown however that directly predicting runtimes without any transformation does not work well in practice~\cite{NgokoCT19,CollauttiMMO13}.
-%JB: Collauti at el. still predict actual (though normalized) runtimes rather than discrete labels, they only heavily discretize runtimes when using them as features
-%TF: added "without any transformation"
 In our experiments, the mean squared error was within the same magnitude as the values to be predicted.
@@
 To be applicable to $\hat{a}$, we extend \emph{score} by replacing $\gamma\!\left(a, i\right)$ by the model's predictions $f\!: \left(\mathcal{A} \cup \left\lbrace \hat{a} \right\rbrace\right) \times \mathcal{I} \rightarrow \left\lbrace 1, \dots, k \right\rbrace$.
-%JB: as "$\operatorname{score}$" is not visibly different to the rest of the text, we should use some other way of highlighting
-%TF: replaced it with emph{score} for now
 
@@
 This means that we do not lose much discriminatory power using discrete runtime labels.
-%The number of significant comparisons between solvers \emph{alone}, though, does only offer limited insights as we might achieve significant differences in all pairs by assigning some randomly skewed values as labels.
-%JB: thus, we should reorder this section to mention the argument "significant difference" (this paragraph) first and "deciding solver pairs" (previous paragraph, stronger argument) second (rather than giving the stronger argument first, making the second argument superfluous)
-%TF: Removed the previous sentence
 The predictiveness of the PAR-2 score, the labels' discriminatory power, as well as that it enables us to use classification, show the clear advantage of runtime discretization.
-%JB: at the same time with what? with the other benefits?
-%TF: removed redundant phrase
 
@@
 First, we present possible baseline approaches including existing work on instance-based selection strategies~\cite{Bossek021a,GentHJKMNN14,MatriconAFSH21}.
-%JB: approaches under "Second" are also at least "inspired by existing work" and probably also pick "interesting" instances, so distinction between the two categories of approaches not entirely clear
-%TF: made it clearer. First = instance-based methods, second = model-based methods
 Second, we show promising machine-learning-based sampling strategies that make use of the model's label-prediction function $f$ and are inspired by existing work within the realms of active learning~\cite{settles2009active}.
@@
 The model-based information-gain sampling strategy selects the instance with the highest expected entropy reduction.
-%This implicitly assumes that the model's runtime-label probabilities for an instance are close to being correct.
-%JB: not sure what "close to being correct" means excactly: does strategy assume correct predictions, but implicitly has a tolerance for wrong predictions, or does it explicitly account for errors somehow?
-%TF: removed the previous sentence as it probably confuses readers more and is not essential for the explanation
 To be more specific, we select the instance $i$ that maximizes
@@
 Thereby, \emph{wilcoxon} denotes the $p$-value of a Wilcoxon signed-rank test of the runtime-label distributions and $\gamma\!\left(a, \mathcal{I}\right)$ and $f\!\left(\hat{a}, \mathcal{I}\right)$ are the natural set extensions on $\gamma$ and $f$ respectively.
-%JB: "set extension" might be rather technical and hard to understand; also, why "expected"? in the sense of "straightforward"?
-%TF: I already saw this expression in other papers. I'm not sure how to say the same thing clearer in the same amount of space.
 To improve the stability of this criterion, we use an exponential moving average
@@
     $O_{\operatorname{rt}} \leftarrow \left(\sum_{i \in \tilde{\mathcal{I}}} r\!\left(i, \hat{a}\right)\right) / \left(\sum_{i \in \mathcal{I}} r\!\left(i, \hat{a}\right)\right)$
-    %JB: do we need to call a function here or can't we directly write how this is computed (define a variable starting at 0 to which the runtime fraction is added each iteration)?
-    %TF: fixed
 
@@
 The end-to-end experiment configurations are given below.
-%JB: still interesting to see which approaches were considered in preliminary experiments; could also list them here and indicate somehow which methods were only used preliminarily (e.g., by putting "(preliminary experiments only)" or some shorter label/sign behind them)
-%TF: Added some selection methods that are mentioned.
 

### 2022-10-10 paper/paper.tex
elaborate on runtime transformation function

@@
 Section~\ref{sec:main-framework} describes the general procedure.
-Section~\ref{sec:main-model} specifies the details of the underlying prediction model.
-Section~\ref{sec:main1} discusses how we may derive a solver ranking from that model.
+Section~\ref{sec:main-model} specifies the details of the underlying prediction model and discusses how we may derive a solver ranking from that model.
 Possible criteria for sampling instances are discussed in Section~\ref{sec:main2}.
@@
 \label{sec:main-framework}
-
-Let benchmark instances $\mathcal{I}$, SAT solvers $\mathcal{A}$, and runtimes $r\!: \mathcal{A} \times \mathcal{I} \rightarrow \left[0, \tau\right]$ be given.
 
@@
 When the stopping criterion is met, we return the predicted score of the given solver $\hat{a}$.
+
+Since the potential runtime of experiments is by magnitudes larger than the model's update time, we only consider incrementing our benchmark by one instance at a time rather than using batches which is also proposed in current active-learning advances~\cite{SinhaED19,2019gaal}.
 
@@
 
-Since the potential runtime of experiments is by magnitudes larger than the model's update time, we only consider incrementing our benchmark by one instance at a time rather than using batches which is also proposed in current active-learning advances~\cite{SinhaED19,2019gaal}.
-
-At the core of the model $\mathcal{M}$ is a prediction function $f\!: \left(\mathcal{A} \cup \left\lbrace \hat{a} \right\rbrace\right) \times \mathcal{I} \rightarrow \mathbb{R}$ that powers the decisions within the previously mentioned three steps.
-
-\subsection{Runtime Preprocessing and Rank Prediction}
-\label{sec:main1}
-
-%In order to predict a ranking of solver $\hat{a}$ amidst its competitors, we need to infer its performance on all instances, even the ones on which we have not evaluated it.
-%To do so, there are generally two notions of this problem, i.e, regression and classification.
-%A regression model tries to predict the \mbox{PAR-2} runtime of $\hat{a}$ on all instances based on the samples that have been selected.
-%Then, the ranking is induced by the average predictions.
-%Alternatively, a classification model can be used to predict some discrete runtime labels of $\hat{a}$ on all instances.
-%Similarly, the average runtime label determines the ranking.
-% MI: skipping this elaboration, since we do not use regression
-
-Preliminary experiments, as well as research in portfolio solvers, have shown however that directly predicting runtimes without any transformation does not work well in practice~\cite{NgokoCT19,CollauttiMMO13}.
-In our experiments, the mean squared error was within the same magnitude as the values to be predicted.
-In contrast to that, discretizing runtimes and using those labels in the model's training and prediction yields good results (cf.~Section~\ref{sec:eval}).
-%The following paragraphs motivate why it makes sense to discretize runtimes.%MI: no further motivation found, commented out
-%JB: we should probably cut this a bit
-%TF: I've commented out a few sentences but we probably need to cut more.
-
-We use a function $\gamma\!: {\mathcal{A} \times \mathcal{I}} \rightarrow \left\lbrace 1, \dots, k \right\rbrace$ to transform runtime observations.
-In our experiments, this function is a clustering algorithm.
-Note that this transformation is done per instance to adapt to different instance difficulties. 
-Also, we define ${\gamma\!\left(a, i\right) := k}$ iff $r\!\left(a, i\right) = \,\perp$ for all $a \in \mathcal{A}$ and $i \in \mathcal{I}$.
+Let benchmark instances $\mathcal{I}$, SAT solvers $\mathcal{A}$, and runtimes $r : \mathcal{A} \times \mathcal{I} \rightarrow \left[0, \tau\right]$ be given. 
+At the core of the model $\mathcal{M}$ is a prediction function $f : \left(\mathcal{A} \cup \left\lbrace \hat{a} \right\rbrace\right) \times \mathcal{I} \rightarrow \mathbb{R}$ that powers the decisions within the previously mentioned three steps.
+
+\subsubsection{Runtime Transformation.}
+
+We transform the runtimes into discrete runtime classes which are determined on a per instance basis.
+For each instance $e \in \mathcal{I}$, we use a hierarchial clustering algorithm to assign the runtimes in $\{ r(a, e) \mid a \in A \}$ to one of the $k$ clusters $C_1, \dots, C_k$.
+Without loss of generality, the fastest runtimes for instance $e$ are in cluster $C_1$ and the slowest are in cluster $C_k$.
+For instances with many timeouts, we manually fix the clustering results such that all timeouts $\tau$ are in cluster $C_k$.
+The runtime transformation function $\gamma_k : {\mathcal{A} \times \mathcal{I}} \rightarrow \left\lbrace 1, \dots, k \right\rbrace$ is then specified as follows.
+$$\gamma_k(a, e) = j ~\Leftrightarrow~ r(a, e) \in C_j$$
+
+Given an instance $e \in \mathcal{I}$, a solver $a \in A$ belongs to the $\gamma_k(a, e)$-fastest solvers on instance $e$. 
+In our preliminary experiments, we achieved higher accuracy with prediction of such discrete runtime labels than with prediction of raw runtimes. 
+The mean squared error of raw runtime prediction was within the same magnitude as the values to be predicted. 
+%MI: how much more accurate are we with discrete runtime labels
+Research in portfolio solvers has also shown that directly predicting runtimes without any transformation does not work well in practice~\cite{NgokoCT19,CollauttiMMO13}.
+
 To determine a ranking based on those labels, we define
 \begin{equation}
-  \operatorname{score}\!\left(a\right) := \frac{1}{|\mathcal{I}|} \sum_{i \in \mathcal{I}} w_{a,i} \cdot \left(\gamma\!\left(a, i\right) - 1\right) \enspace \textrm{,}
-  %JB: does the "-1" provide any benefits, or does it only shift scores?
-  %TF: it is used to resemble the 0-indexed implementation. I'm also not happy with this, but using \lbrace 0, \dots, k - 1 \rbrace does also look ugly.
-  %MI: Why is it so important to have this zero-based?
+  \operatorname{score}\!\left(a\right) := \frac{1}{|\mathcal{I}|} \sum_{e \in \mathcal{I}} w_{a,i} \cdot \gamma(a, i)
   \label{eq:rankingeq}
@@
 
-\subsubsection{Baselines and existing approaches.}
-\label{sec:sampling1}
-Apart from obvious baseline instance-selection strategies, i.e., random sampling and neighborhood-aware sampling (randomly choosing among the instances with the highest amount of non-sampled $k$ neighbors), we also implement existing instance-selection strategies.
-
-Bossek and Wagner~\cite{Bossek021a} introduce a custom ranking-score metric incorporating the number of \emph{good} pairs and \emph{bad} pairs.
+% MI: The following belongs either to Related Work or to Experimental Design (commented out)
+
+%\subsubsection{Baselines and existing approaches.}
+%\label{sec:sampling1}
+%Apart from obvious baseline instance-selection strategies, i.e., random sampling and neighborhood-aware sampling (randomly choosing among the instances with the highest amount of non-sampled $k$ neighbors), we also implement existing instance-selection strategies.
+
+%Bossek and Wagner~\cite{Bossek021a} introduce a custom ranking-score metric incorporating the number of \emph{good} pairs and \emph{bad} pairs.
 %JB: some selection methods, like this one, are not mentioned later anymore (in particular, not in experimental design), so we don't need to describe them in detail (maybe mention them shortly if used in preliminary experiments, but turned out to be not useful)
 %TF: I've cut the section down a bit
-A good solver pair, regarding a particular instance, are solvers that perform according to their global ranking.
-%So if solver $a$ is better than solver $b$ on instance $i$ and $a$ is also in total better than $b$, it is a \emph{good} solver pair on $i$.
-Otherwise, it is a \emph{bad} pair.
-Based on this classification, they assign a custom score to each instance.
-By selecting instances with the highest score, they try to favor instances that are predictive of the final ranking.
-%This assumes, though, that the new solver $\hat{a}$ behaves similarly to the existing solvers.
-
-Gent~et~al.~\cite{GentHJKMNN14} make use of a concept known from racing for automated model selection.
-Given a particular instance and a parameter $\rho$, a solver is $\rho$-dominated if it is at least $\rho$ times worse than the best solver on this problem instance.
-By selecting instances with the highest amount of $\rho$-dominated solvers, they hope to sample instances that provide the best discriminatory power to distinguish the runtimes of a new contestant.
-
-Finally, we also implement some of the selection approaches described by Matricon~et.al.~\cite{MatriconAFSH21}.
-Besides uniformly random sampling and the aforementioned discrimination-based method by Gent~et~al.~\cite{GentHJKMNN14}, they also describe a variance-based selection approach.
-The variance-based strategy selects instances with the highest runtime variance per average runtime.
-We also implement a logarithmized version of this, i.e., the highest log-runtime variance per average log runtime.
+%A good solver pair, regarding a particular instance, are solvers that perform according to their global ranking.
+%Otherwise, it is a \emph{bad} pair.
+%Based on this classification, they assign a custom score to each instance.
+%By selecting instances with the highest score, they try to favor instances that are predictive of the final ranking.
+
+%Gent~et~al.~\cite{GentHJKMNN14} make use of a concept known from racing for automated model selection.
+%Given a particular instance and a parameter $\rho$, a solver is $\rho$-dominated if it is at least $\rho$ times worse than the best solver on this problem instance.
+%By selecting instances with the highest amount of $\rho$-dominated solvers, they hope to sample instances that provide the best discriminatory power to distinguish the runtimes of a new contestant.
+
+%Finally, we also implement some of the selection approaches described by Matricon~et.al.~\cite{MatriconAFSH21}.
+%Besides uniformly random sampling and the aforementioned discrimination-based method by Gent~et~al.~\cite{GentHJKMNN14}, they also describe a variance-based selection approach.
+%The variance-based strategy selects instances with the highest runtime variance per average runtime.
+%We also implement a logarithmized version of this, i.e., the highest log-runtime variance per average log runtime.
 
@@
 \begin{equation}
-  \underset{i \in \mathcal{I} \setminus \tilde{\mathcal{I}}}{\arg\min} \left\lvert \frac{1}{k} - \max_{n \in \left\lbrace 1, \dots, k \right\rbrace} f'\!\left(\hat{a}, i\right)_{n} \right\rvert \enspace \textrm{.}
+  \underset{e \in \mathcal{I} \setminus \tilde{\mathcal{I}}}{\arg\min} \left\lvert \frac{1}{k} - \max_{n \in \left\lbrace 1, \dots, k \right\rbrace} f'\!\left(\hat{a}, i\right)_{n} \right\rvert \enspace \textrm{.}
 \end{equation}

### 2022-10-10 paper/paper.tex
Fix and Adapt Section 3's labels

@@
 Questionable research practices might modify these elements a-posteriori to fit expectations, leading to biased results~\cite{NiesslHWCB22}.
+In the following, we discuss related work from the areas of static benchmark selection, algorithm configuration, incremental benchmark selection, and active learning.
 
@@
 
-\subsubsection{SAT solver configuration.}
+\subsubsection{Algorithm configuration.}
 
@@
 Section~\ref{sec:main-model} specifies the details of the underlying prediction model and discusses how we may derive a solver ranking from that model.
-Possible criteria for sampling instances are discussed in Section~\ref{sec:main2}.
-Section~\ref{sec:main3} concludes with possible stopping conditions.
+Possible criteria for sampling instances are discussed in Section~\ref{sec:main-sampling}.
+Section~\ref{sec:main-stopping} concludes with possible stopping conditions.
 
@@
 \subsection{Instance Sampling}
-\label{sec:main2}
+\label{sec:main-sampling}
 Our instance-sampling strategies (also refer to Algorithm~\ref{algALBenchmark}; line 4) can be divided into two categories.
@@
 \subsection{Stopping Criteria}
-\label{sec:main3}
+\label{sec:main-stopping}
 Similar to the instance-selection strategies, we can also divide stopping criteria into baseline and active-learning-based approaches.
@@
 \subsubsection{Ranking.}
-Regarding \emph{ranking} (cf. Section~\ref{sec:main1}), we experiment with the following approaches, including our used hyper-parameter values:
+Regarding \emph{ranking} (cf. Section~\ref{sec:main-model}), we experiment with the following approaches, including our used hyper-parameter values:
 %JB: as parts of these approaches are mostly new to readers (Section 3.1 only explains prediction-based ranking in general, neither observed ranking nor history nor fallback), I'm not sure if we should provide further motivation/explanation here); also applies to other components of experimental design
@@
     \item Fallback threshold: if the difference of scores between two solvers drops below \SI{0.01}, \SI{0.05}, or \SI{0.1}, use the partially observed PAR-2 ranking as a tie-breaker.
-As discussed in Section~\ref{sec:main1}, ranking decisions are all correct if the difference in scores between two solvers exceeds a certain value.
+As discussed in Section~\ref{sec:main-model}, ranking decisions are all correct if the difference in scores between two solvers exceeds a certain value.
     %JB: "difference of ranking" between a pair of new solver and another solver? or has this phrase a different meaning?
@@
 \subsubsection{Selection.}
-For \emph{selection} (cf. Section~\ref{sec:sampling1}), we experiment with the following methods, including our used hyper-parameter values:
+For \emph{selection} (cf. Section~\ref{sec:main-sampling}), we experiment with the following methods, including our used hyper-parameter values:
 %JB: the referenced sampling section describes more methods than the ones listed below, so I guess some methods got dropped after preliminary experiments
@@
 \subsubsection{Stopping.}
-For \emph{stopping} decisions (cf. Section~\ref{sec:main3}), we experiment with the following criteria, including our used hyper-parameter values:
+For \emph{stopping} decisions (cf. Section~\ref{sec:main-stopping}), we experiment with the following criteria, including our used hyper-parameter values:
 \begin{itemize}
@@
 We then gradually merge intervals whose single-link logarithmic distance is the smallest until the desired number of partitions is reached.
-As already mentioned in Section~\ref{sec:main1}, this approach has good properties regarding the predictiveness of the PAR-2 score ranking and its discriminatory power.
+As already mentioned in Section~\ref{sec:main-model}, this approach has good properties regarding the predictiveness of the PAR-2 score ranking and its discriminatory power.
 Also, preliminary experiments showed that our stacking ensemble performed particularly well for this kind of discretization method.

### 2022-10-10 paper/paper.tex
Revise Section 3.1 (Incremental Benchmarking Framework)
Rewording for clarity and language.

@@
 In this section, we present the details of our approach.
-Section~\ref{sec:main-framework} describes the general procedure.
-Section~\ref{sec:main-model} specifies the details of the underlying prediction model and discusses how we may derive a solver ranking from that model.
-Possible criteria for sampling instances are discussed in Section~\ref{sec:main-sampling}.
+Section~\ref{sec:main-framework} describes the general framework.
+Section~\ref{sec:main-model} specifies the details of the underlying prediction model and describes how we may derive a solver ranking from that model.
+We discuss possible criteria for sampling instances in Section~\ref{sec:main-sampling}.
 Section~\ref{sec:main-stopping} concludes with possible stopping conditions.
@@
 
-Our framework follows a simple three-step schema.
-Algorithm~\ref{algALBenchmark} outlines the procedure. 
-While the stopping condition is not met (line~1), we select \emph{one} problem instance from the pool of instances (line~2) and evaluate the new solver $\hat{a}$ on it (line~3).
-We then use the acquired result to update the model (line~4).
-When the stopping criterion is met, we return the predicted score of the given solver $\hat{a}$.
-
-Since the potential runtime of experiments is by magnitudes larger than the model's update time, we only consider incrementing our benchmark by one instance at a time rather than using batches which is also proposed in current active-learning advances~\cite{SinhaED19,2019gaal}.
+Our framework to score a new solver~$\hat{a}$ follows a simple three-step procedure, outlined in Algorithm~\ref{algALBenchmark}:
+While the stopping condition is not met (line~1), we select \emph{one} problem instance from the pool of instances with the help of a prediction model~$\mathcal{M}$ (line~2) and evaluate the new solver~$\hat{a}$ on this instance (line~3).
+We then use the acquired result to update the prediction model~$\mathcal{M}$ (line~4).
+When the stopping criterion is met, we return the predicted score of the new solver~$\hat{a}$ (line~5).
+
+Since the potential runtime of experiments is by magnitudes larger than the model's update time, we only consider incrementing our benchmark by one instance at a time rather than using batches, which is also proposed in current active-learning advances~\cite{SinhaED19,2019gaal}.
 
@@
 
-  \While{$\operatorname{not} \, \operatorname{stop}\!\left(\mathcal{M}\right)$}{
-    $s \leftarrow \operatorname{selectNextInstance}\!\left(\mathcal{M}\right)$
+  \While(\tcp*[f]{cf. Section~\ref{sec:main-stopping}}){$\operatorname{not} \, \operatorname{stop}\!\left(\mathcal{M}\right)$}{
+    $s \leftarrow \operatorname{selectNextInstance}\!\left(\mathcal{M}\right)$ \tcp*{cf. Section~\ref{sec:main-sampling}}
 
@@
 
-    $\operatorname{updateModel}\!\left(\mathcal{M},\, s,\, t\right)$
-  }
-
-  \Return $\operatorname{predictScore}\!\left(\mathcal{M}, \hat{a}\right)$
+    $\operatorname{updateModel}\!\left(\mathcal{M},\, s,\, t\right)$ \tcp*{cf. Section~\ref{sec:main-model}}
+  }
+
+  \Return $\operatorname{predictScore}\!\left(\mathcal{M}\right)$ \tcp*{cf. Section~\ref{sec:main-model}}
 

### 2022-10-10 paper/paper.tex
section3.2

@@
 
+To be applicable to $\hat{a}$, we extend \emph{score} by replacing $\gamma\!\left(a, i\right)$ by the model's predictions $f\!: \left(\mathcal{A} \cup 
+\left\lbrace \hat{a} \right\rbrace\right) \times \mathcal{I} \rightarrow \left\lbrace 1, \dots, k \right\rbrace$.
+
 \subsubsection{Runtime Transformation.}
@@
 We transform the runtimes into discrete runtime classes which are determined on a per instance basis.
-For each instance $e \in \mathcal{I}$, we use a hierarchial clustering algorithm to assign the runtimes in $\{ r(a, e) \mid a \in A \}$ to one of the $k$ clusters $C_1, \dots, C_k$.
+For each instance $e \in \mathcal{I}$, we use a single-link hierarchial clustering algorithm to assign the runtimes in $\{ r(a, e) \mid a \in A \}$ to one of the $k$ clusters $C_1, \dots, C_k$.
 Without loss of generality, the fastest runtimes for instance $e$ are in cluster $C_1$ and the slowest are in cluster $C_k$.
-For instances with many timeouts, we manually fix the clustering results such that all timeouts $\tau$ are in cluster $C_k$.
+For instances with many timeouts, we have to manually fix the clustering results such that all timeouts $\tau$ are in cluster $C_k$.
 The runtime transformation function $\gamma_k : {\mathcal{A} \times \mathcal{I}} \rightarrow \left\lbrace 1, \dots, k \right\rbrace$ is then specified as follows.
@@
 
-To determine a ranking based on those labels, we define
-\begin{equation}
-  \operatorname{score}\!\left(a\right) := \frac{1}{|\mathcal{I}|} \sum_{e \in \mathcal{I}} w_{a,i} \cdot \gamma(a, i)
+To determine solver ranks, we define the adapted scoring function $s : A \rightarrow [1, 2 \cdot k]$ as follows.
+\begin{align*}
+  \operatorname{s}(a) &:= \frac{1}{|\mathcal{I}|} \sum_{e \in \mathcal{I}} \gamma'_k(a, e)
+  &&
+  \gamma'_k(a, e) &:= \begin{cases}
+  	2 \cdot \gamma_k(a, e) 	& \text{if } \gamma_k(a, e) = k\\
+	\gamma_k(a, e)	& \text{otherwise}
+  \end{cases}
   \label{eq:rankingeq}
-\end{equation}
-with $w_{a,i} := 2$ if $\gamma\!\left(a, i\right) = k$ else $w_{a,i} := 1$.
-To be applicable to $\hat{a}$, we extend \emph{score} by replacing $\gamma\!\left(a, i\right)$ by the model's predictions $f\!: \left(\mathcal{A} \cup \left\lbrace \hat{a} \right\rbrace\right) \times \mathcal{I} \rightarrow \left\lbrace 1, \dots, k \right\rbrace$.
+\end{align*}
 
 Discretization would not provide much value if the resulting cluster labels were not predictive of the PAR-2 ranking we seek to predict.
-In our work, we use a hierarchical-clustering approach with a single-link criterion for $k = 3$.
 For the SAT Competition~2022 Anniversary Track instances~\cite{sat2022}, this ranking approach is capable of correctly deciding for almost all pairs of solvers which one is faster (about \SI{97.45}{\%}; $\sigma = \SI{3.68}{\%}$).

### 2022-10-10 paper/paper.tex
section2.3
section3.2

@@
 
-To determine solver ranks, we define the adapted scoring function $s : A \rightarrow [1, 2 \cdot k]$ as follows.
+To determine solver ranks, we define the adapted scoring function $s_k : A \rightarrow [1, 2 \cdot k]$ as follows.
 \begin{align*}
-  \operatorname{s}(a) &:= \frac{1}{|\mathcal{I}|} \sum_{e \in \mathcal{I}} \gamma'_k(a, e)
+  s_k(a) := \frac{1}{|\mathcal{I}|} \sum_{e \in \mathcal{I}} \gamma'_k(a, e)
   &&
-  \gamma'_k(a, e) &:= \begin{cases}
+  \gamma'_k(a, e) := \begin{cases}
   	2 \cdot \gamma_k(a, e) 	& \text{if } \gamma_k(a, e) = k\\
@@
 
-Discretization would not provide much value if the resulting cluster labels were not predictive of the PAR-2 ranking we seek to predict.
-For the SAT Competition~2022 Anniversary Track instances~\cite{sat2022}, this ranking approach is capable of correctly deciding for almost all pairs of solvers which one is faster (about \SI{97.45}{\%}; $\sigma = \SI{3.68}{\%}$).
-Moreover, it decides all solvers correctly if the difference in the cluster-label-ranking scores between two solvers is at least \SI{0.11}{}.
-%Refer to Table~\ref{tab:discreteranking} for a comparison with the PAR-2 ranking.
-The Spearman correlation~\cite{de2016comparing} of cluster-label and PAR-2 ranking is about \SI{0.988} (in comparison to a perfect value of \SI{1}).
+The scoring function $s_k$ induces a ranking among solvers. 
+In preliminary experiments, we compared the ranking induced by $s_3$ to the ranking induced by the PAR-2 score which is commonly used in SAT competitions~\cite{FroleyksHIJS21}. 
+The Spearman correlation of $s_3$ and PAR-2 ranking is about \SI{0.988} which is very close to the optimal value of 1~\cite{de2016comparing}.
+
+That discretized runtimes are sufficient for discriminating solvers can also be shown by comparing the fraction of solver pairs whose runtimes are significantly different. 
+In preliminary experiments, we used a Wilcoxon signed rank test with $\alpha = 0.05$ to compare solver pairs by their runtime vectors.
+Within the SAT~Competition~2022 Anniversary track, about \SI{89.95}{\%} of solver pairs are significantly different according to that measure.
+The strongly discretized runtimes $\gamma_2$ are sufficient to significantly distinguish about \SI{87.04}{\%} of solver pairs, while $\gamma_3$ is sufficient to distinguish about \SI{87.83}{\%} of solver pairs.
+
+% MI: Enough justification, commented out the following
+%For the SAT Competition~2022 Anniversary track instances, this ranking approach is capable of correctly deciding for almost all pairs of solvers which one is faster (about \SI{97.45}{\%}; $\sigma = \SI{3.68}{\%}$).
+%Moreover, it decides all solvers correctly if the difference in the cluster-label-ranking scores between two solvers is at least \SI{0.11}{}.
+
 This shows that monotonicity and score distances between ranks align quite well between both ranking methods.
-
-% In that sense, we also experiment with a combination of runtime-label ranking and the PAR-2 ranking of the subsample if the difference in label ranking scores drops below a certain threshold.
-% Besides that, we additionally consider using the past $l$ model predictions in a voting approach for stability.
-
-That discretized runtimes are sufficient for distinguishing solvers can also be shown by comparing the fraction of solver pairs whose runtimes are significantly different.
-Within the SAT~Competition~2022 Anniversary Track, about \SI{89.95}{\%} of solver pairs are significantly different (according to a Wilcoxon-signed-rank test with $\alpha = 0.05$).
-While only time-out labels (solver times out or not on a particular instance) are sufficient to significantly distinguish about \SI{87.04}{\%} of solver pairs, discrete-runtime labels are sufficient to distinguish about \SI{87.83}{\%} of solver pairs.
-%\footnote{Based on hierarchical clustering with a log-single-link criterion for $k = 3$.}
-This means that we do not lose much discriminatory power using discrete runtime labels.
-The predictiveness of the PAR-2 score, the labels' discriminatory power, as well as that it enables us to use classification, show the clear advantage of runtime discretization.
+The correlation between $s_k$-induced and PAR-2 induced ranks, the discriminatory power of $\gamma_k$, as well as that it enables us to use classification, show the clear advantage of runtime discretization.
 

### 2022-10-10 paper/paper.tex
Minor revision of Experimental Design
- "Alg." -> "Algorithm"
- whitespace
- comments

@@
 The model-convergence stopping criterion is less sophisticated in comparison.
-It simply stops the active-learning process (Alg.~\ref{algALBenchmark}) if within the last $l$ iterations the ranking that is induced by the model's predictions (Eq.~\ref{eq:rankingeq}) remains unchanged.
+It simply stops the active-learning process (Algorithm~\ref{algALBenchmark}) if within the last $l$ iterations the ranking that is induced by the model's predictions (Eq.~\ref{eq:rankingeq}) remains unchanged.
 Note that the concrete values of $\operatorname{score}\!\left(a\right)$ might still change.
@@
 \section{Experimental Design}
-Given all the previously presented instantiations for Alg.~\ref{algALBenchmark}, this section briefly outlines our experimental design, including our optimization goal, evaluation framework, used data sets, hyper-parameter choices, as well as implementation details.
+
+Given all the previously presented instantiations for Algorithm~\ref{algALBenchmark}, this section briefly outlines our experimental design, including our optimization goal, evaluation framework, used data sets, hyper-parameter choices, as well as implementation details.
 % hyper-parameter values that we experiment with
@@
 \label{sec:goal}
+
 As already stated in the introductory section, this work addresses the \emph{New-Solver Problem} (cf.~Definition~\ref{def:new-solver-problem}).
@@
 We define the \emph{ranking accuracy} $O_{\operatorname{acc}} \in \left[0, 1\right]$ (higher is better) by the fraction of pairs $\left(\hat{a}, a\right)$ for all $a \in \mathcal{A}$ that are decided correctly by the given ranking (cf.
-Alg.~\ref{alg:eval}).
+Algorithm~\ref{alg:eval}).
 For now, we exclude the equality of solvers since data shows that the PAR-2 scores of all our solvers are different.
@@
 The \emph{fraction of runtime} that the algorithm needs to arrive at its conclusion is denoted by $O_{\operatorname{rt}} \in \left[0, 1\right]$ (lower is better).
-This metric puts the runtime summed over the sampled instances in relation to the runtime summed over all instances in the dataset (cf. Alg.~\ref{alg:eval}).
+This metric puts the runtime summed over the sampled instances in relation to the runtime summed over all instances in the dataset (cf. Algorithm~\ref{alg:eval}).
 Overall, we want to find an approach that maximizes
+%
 \begin{equation}
@@
 \end{equation} 
+%
 whereby $\delta \in \left[0, 1\right]$ allows for linear weighting between the two optimization goals $O_{\operatorname{acc}}$ and $O_{\operatorname{rt}}$.
@@
 \label{sec:evalframe}
+
 To evaluate a concrete instantiation of Algorithm~\ref{algALBenchmark} (a concrete choice for all the sub-routines), we perform cross-validation on our set of solvers.
 That means that each solver once plays the role of the new solver.
-Alg.~\ref{alg:eval} shows this.
+Algorithm~\ref{alg:eval} shows this.
 Note that the \emph{new} solver in each iteration is excluded from the set of solvers $\mathcal{A}$ to avoid data leakage.
@@
 
-\vspace{-0.2cm}
 \begin{algorithm}[htbp]
@@
 
-    $\left( score,\, \tilde{\mathcal{I}} \right) \leftarrow \operatorname{runALAlgorithm}\!\left(\mathcal{I}, \mathcal{A}, \left.r\right|_{\mathcal{A} \times \mathcal{I}}, \hat{a}\right)$ \tcp*{Refer to Alg.~\ref{algALBenchmark}}
+    $\left( score,\, \tilde{\mathcal{I}} \right) \leftarrow \operatorname{runALAlgorithm}\!\left(\mathcal{I}, \mathcal{A}, \left.r\right|_{\mathcal{A} \times \mathcal{I}}, \hat{a}\right)$ \tcp*{Refer to Algorithm~\ref{algALBenchmark}}
 
@@
     \tcp*{Iverson-bracket notation}
+    %JB: maybe we can simplify the expresson, since readers might be unfamiliar with that notation (I had to look it up)
 
@@
 \end{algorithm}
-\vspace{-0.2cm}
 
 \subsection{Data}
+
 In our experiments, we work with the SAT~Competition~2022 Anniversary Track instances~\cite{sat2022}.
@@
 \label{sec:hyper}
-Given Alg.~\ref{algALBenchmark}, there are several possible instantiations for the three phases, i.e., \emph{selection}, \emph{stopping}, and \emph{ranking}.
+
+Given Algorithm~\ref{algALBenchmark}, there are several possible instantiations for the three phases, i.e., \emph{selection}, \emph{stopping}, and \emph{ranking}.
 Also, there are different choices for the runtime-label prediction model.
@@
 \subsubsection{Ranking.}
+
 Regarding \emph{ranking} (cf. Section~\ref{sec:main-model}), we experiment with the following approaches, including our used hyper-parameter values:
-%JB: as parts of these approaches are mostly new to readers (Section 3.1 only explains prediction-based ranking in general, neither observed ranking nor history nor fallback), I'm not sure if we should provide further motivation/explanation here); also applies to other components of experimental design
-%TF: I've tried to add some more information. 
+
 \begin{itemize}
@@
   \begin{itemize}
-    \item History size: consider the latest 1, 10, 20, 30, or 40 predictions within a voting approach for stability. Given a particular ranking decision, the latest $x$ predictions vote on the winner.
-    \item Fallback threshold: if the difference of scores between two solvers drops below \SI{0.01}, \SI{0.05}, or \SI{0.1}, use the partially observed PAR-2 ranking as a tie-breaker.
+    \item
+    History size: consider the latest 1, 10, 20, 30, or 40 predictions within a voting approach for stability.
+    Given a particular ranking decision, the latest $x$ predictions vote on the winner.
+    %JB: confuses me a bit as it sounds we rank first and then vote on a winner; isn't it the other way around?
+    \item
+    Fallback threshold: if the difference of scores between two solvers drops below \SI{0.01}, \SI{0.05}, or \SI{0.1}, use the partially observed PAR-2 ranking as a tie-breaker.
+    %JB: as clarification: one these two solvers has to be the new solver, right? if yes, would write it explicitly
     As discussed in Section~\ref{sec:main-model}, ranking decisions are all correct if the difference in scores between two solvers exceeds a certain value.
-    %JB: "difference of ranking" between a pair of new solver and another solver? or has this phrase a different meaning?
-    %TF: tried to clarify
   \end{itemize}
@@
 \subsubsection{Selection.}
+
 For \emph{selection} (cf. Section~\ref{sec:main-sampling}), we experiment with the following methods, including our used hyper-parameter values:
-%JB: the referenced sampling section describes more methods than the ones listed below, so I guess some methods got dropped after preliminary experiments
-%TF: added them
+
 \begin{itemize}
@@
     \item Fallback threshold: use random sampling for the first \SI{0}{\%}, \SI{5}{\%}, \SI{10}{\%}, \SI{15}{\%}, or \SI{20}{\%} of instances to explore the instance space.
-    \item Runtime scaling: prefer instances with the greatest uncertainty per average runtime (\texttt{True} or \texttt{False}).
-    %JB: how does this scaling work exactly? divide runtimes by average runtime of all solvers on that instance?
-    %TF: yes
+    \item Runtime scaling: prefer instances with the greatest uncertainty per average (over all solvers) runtime (\texttt{True} or \texttt{False}).
   \end{itemize}
@@
 \subsubsection{Stopping.}
+
 For \emph{stopping} decisions (cf. Section~\ref{sec:main-stopping}), we experiment with the following criteria, including our used hyper-parameter values:
+
 \begin{itemize}
@@
     %TF: for that reason, it is combined with the "minimum amount". Using fractions makes the approach more useful when adapting it to benchmarks of different sizes.
+    %JB: yes, but (1) even combination with "minimum amount" can make stopping very easy (see my example) and (2) fraction for convergence could also be relative to absolute number of instances (which is constant over AL run) rather than fraction of currently sampled instances (which increases over AL run)
   \end{itemize}
@@
 \subsubsection{Runtime-label prediction model.}
+
 As a runtime-label prediction model, we only make use of one method that has proven to be successful in preliminary experiments (cf.~Section~\ref{sec:evalprel}).
@@
 \subsection{Implementation Details}
+
 For reproducibility, our source code, all experiments, and data are available on GitHub\footnote{\url{https://github.com/mathefuchs/al-for-SAT solver-benchmarking}}.

### 2022-10-10 paper/paper.tex
section3.2

@@
 We transform the runtimes into discrete runtime classes which are determined on a per instance basis.
-For each instance $e \in \mathcal{I}$, we use a single-link hierarchial clustering algorithm to assign the runtimes in $\{ r(a, e) \mid a \in A \}$ to one of the $k$ clusters $C_1, \dots, C_k$.
-Without loss of generality, the fastest runtimes for instance $e$ are in cluster $C_1$ and the slowest are in cluster $C_k$.
+For each instance $e \in \mathcal{I}$, we use a single-link hierarchial clustering algorithm to assign the runtimes in $\bigl\{ r(a, e) \mid a \in A \bigr\}$ to one of the $k$ clusters $C_1, \dots, C_k$ such that the fastest runtimes for instance $e$ are in cluster $C_1$ and the slowest are in cluster $C_k$.
 For instances with many timeouts, we have to manually fix the clustering results such that all timeouts $\tau$ are in cluster $C_k$.
@@
 
-To determine solver ranks, we define the adapted scoring function $s_k : A \rightarrow [1, 2 \cdot k]$ as follows.
+To determine solver ranks, we use the transformed runtimes $\gamma_k(a, e)$ in the adapted scoring function $s_k : A \rightarrow [1, 2 \cdot k]$ as follows.
 \begin{align*}
@@
 \label{sec:main-sampling}
+
 Our instance-sampling strategies (also refer to Algorithm~\ref{algALBenchmark}; line 4) can be divided into two categories.

### 2022-10-10 paper/paper.tex
section3

@@
 To determine solver ranks, we use the transformed runtimes $\gamma_k(a, e)$ in the adapted scoring function $s_k : A \rightarrow [1, 2 \cdot k]$ as follows.
-\begin{align*}
+\begin{align}
   s_k(a) := \frac{1}{|\mathcal{I}|} \sum_{e \in \mathcal{I}} \gamma'_k(a, e)
@@
   \label{eq:rankingeq}
-\end{align*}
+\end{align}
 
@@
 
-Our instance-sampling strategies (also refer to Algorithm~\ref{algALBenchmark}; line 4) can be divided into two categories.
-First, we present possible baseline approaches including existing work on instance-based selection strategies~\cite{Bossek021a,GentHJKMNN14,MatriconAFSH21}.
-Second, we show promising machine-learning-based sampling strategies that make use of the model's label-prediction function $f$ and are inspired by existing work within the realms of active learning~\cite{settles2009active}.
+Selecting an instance based on the model is the core functionality of our method (cf. Algorithm~\ref{algALBenchmark}, Line 2). 
+In this section, we introduce our sampling strategies that make use of the model's label-prediction function $f$ which are inspired by existing work within the realms of active learning~\cite{settles2009active}.
+We implement a model-uncertainty based approach and a model-based information-gain sampling strategy.
+%, and a voting-based committee-disagreement approach.
+
+These methods require the model's predictions to also include probabilities for the $k$ possible runtime labels, i.e., 
+\begin{equation*}
+  f'\!: \left(\mathcal{A} \cup \left\lbrace \hat{a} \right\rbrace\right) \times \mathcal{I} \rightarrow \left[0, 1\right]^k
+\end{equation*}
+
+\subsubsection{Model Uncertainty Based Sampling.}
+
+The model-uncertainty-based approach simply selects the instance that is closest to the model's decision boundary and, of course, not already selected, i.e.,
+\begin{equation*}
+  \underset{e \in \mathcal{I} \setminus \tilde{\mathcal{I}}}{\arg\min} \left\lvert \frac{1}{k} - \max_{n \in \left\lbrace 1, \dots, k \right\rbrace} f'\!\left(\hat{a}, e\right)_{n} \right\rvert
+\end{equation*}
+
+\subsubsection{Information Gain Based Sampling.}
+
+The model-based information-gain sampling strategy selects the instance with the highest expected entropy reduction.
+To be more specific, we select the instance $e$ that maximizes
+\begin{equation*}
+  \operatorname{IG}(e) := \operatorname{H}(e) - \sum_{n = 1}^{k} f'(\hat{a}, e)_{n} \operatorname{H}(e, n)
+\end{equation*}
+whereby $\operatorname{H}\!\left(i\right)$ denotes the entropy of the runtime labels $\gamma\!\left(a, i\right)$ for all $a \in \mathcal{A}$ and $\operatorname{H}\!\left(i, n\right)$ denotes the entropy of the runtime labels including the runtime label of $\hat{a}$ which is $n$.
+The term $\operatorname{H}\!\left(i, n\right)$ is computed for every possible runtime label $n$.
+
+%Finally, the voting-based committee-disagreement approach requires a committee of $c$ models.
+%Rather than having a single prediction function $f$, there is a separate function $f_j$ for all committee members $j \in \left\lbrace 1, \dots, c  \right\rbrace$.
+%We then select the instance $i$ whose entropy of runtime-label votes by the different ensemble members is the highest.
 
 % MI: The following belongs either to Related Work or to Experimental Design (commented out)
-
 %\subsubsection{Baselines and existing approaches.}
@@
 
-\subsubsection{Active-learning-based instance selection.}
-\label{sec:sampling2}
-For conciseness, we only present two of our active-learning-based instance-selection approaches, which show the most promising results.
-We implement a model-uncertainty-based approach and a model-based information-gain sampling strategy.
-%, and a voting-based committee-disagreement approach.
-
-These methods require the model's predictions to also include probabilities for the $k$ possible runtime labels, i.e., 
-\begin{equation}
-  f'\!: \left(\mathcal{A} \cup \left\lbrace \hat{a} \right\rbrace\right) \times \mathcal{I} \rightarrow \left[0, 1\right]^k  \enspace \textrm{.}
-\end{equation}
-The model-uncertainty-based approach simply selects the instance that is closest to the model's decision boundary and, of course, not already selected, i.e.,
-\begin{equation}
-  \underset{e \in \mathcal{I} \setminus \tilde{\mathcal{I}}}{\arg\min} \left\lvert \frac{1}{k} - \max_{n \in \left\lbrace 1, \dots, k \right\rbrace} f'\!\left(\hat{a}, i\right)_{n} \right\rvert \enspace \textrm{.}
-\end{equation}
-
-The model-based information-gain sampling strategy selects the instance with the highest expected entropy reduction.
-To be more specific, we select the instance $i$ that maximizes
-\begin{equation}
-  \operatorname{IG}\!\left(i\right) := \operatorname{H}\!\left(i\right) - \sum_{n = 1}^{k} f'\!\left(\hat{a}, i\right)_{n} \operatorname{H}\!\left(i, n\right) \enspace \textrm{,}
-\end{equation}
-whereby $\operatorname{H}\!\left(i\right)$ denotes the entropy of the runtime labels $\gamma\!\left(a, i\right)$ for all $a \in \mathcal{A}$ and $\operatorname{H}\!\left(i, n\right)$ denotes the entropy of the runtime labels including the runtime label of $\hat{a}$ which is $n$.
-The term $\operatorname{H}\!\left(i, n\right)$ is computed for every possible runtime label $n$.
-
-%Finally, the voting-based committee-disagreement approach requires a committee of $c$ models.
-%Rather than having a single prediction function $f$, there is a separate function $f_j$ for all committee members $j \in \left\lbrace 1, \dots, c  \right\rbrace$.
-%We then select the instance $i$ whose entropy of runtime-label votes by the different ensemble members is the highest.
-
 \subsection{Stopping Criteria}
 \label{sec:main-stopping}
-Similar to the instance-selection strategies, we can also divide stopping criteria into baseline and active-learning-based approaches.
-Also, refer to~Algorithm~\ref{algALBenchmark}, line~3.
-While a fixed subset-size criterion, e.g., only sample \SI{20}{\%} of all available instances, is an example of the former, a stopping criterion based on a Wilcoxon signed-rank test or the convergence of the model's predictions is an example for the latter.
-
+
+%Similar to the instance-selection strategies, we can also divide stopping criteria into baseline and active-learning-based approaches.
+%Also, refer to~Algorithm~\ref{algALBenchmark}, line~3.
+%While a fixed subset-size criterion, e.g., only sample \SI{20}{\%} of all available instances, is an example of the former, a stopping criterion based on a Wilcoxon signed-rank test or the convergence of the model's predictions is an example for the latter.
+
+In this section, we present two dynamic stopping criteria, the Wilcoxon and the Model Uncertainty stopping criteria (cf. Algorithm~\ref{algALBenchmark}, Line~1).
+
+\subsubsection{Wilcoxon Stopping Criterion}
 The Wilcoxon stopping criterion stops the iterative active-learning algorithm when we are on average certain enough that the predicted runtime labels are sufficiently different in comparison to the other solvers.
@@
 To improve the stability of this criterion, we use an exponential moving average
+
 \begin{equation}
@@
 \end{equation}
+
 whereby $m$ denotes the count of iterations within Algorithm~\ref{algALBenchmark} and $W_{\exp}^{\left(0\right)} := 1$.
 
+\subsubsection{Model Convergence Stopping Criterion}
+
 The model-convergence stopping criterion is less sophisticated in comparison.
-It simply stops the active-learning process (Algorithm~\ref{algALBenchmark}) if within the last $l$ iterations the ranking that is induced by the model's predictions (Eq.~\ref{eq:rankingeq}) remains unchanged.
-Note that the concrete values of $\operatorname{score}\!\left(a\right)$ might still change.
+It simply stops the active-learning process if within the last $l$ iterations the ranking that is induced by the model's predictions (Equation~\ref{eq:rankingeq}) remains unchanged.
+Note that the concrete values of $s_k(a)$ might still change.
 We are solemnly interested in the induced ranking in this case.

### 2022-10-10 paper/paper.tex
introduced \hat big-A

@@
 Let benchmark instances $\mathcal{I}$, SAT solvers $\mathcal{A}$, and runtimes $r : \mathcal{A} \times \mathcal{I} \rightarrow \left[0, \tau\right]$ be given. 
-At the core of the model $\mathcal{M}$ is a prediction function $f : \left(\mathcal{A} \cup \left\lbrace \hat{a} \right\rbrace\right) \times \mathcal{I} \rightarrow \mathbb{R}$ that powers the decisions within the previously mentioned three steps.
-
-To be applicable to $\hat{a}$, we extend \emph{score} by replacing $\gamma\!\left(a, i\right)$ by the model's predictions $f\!: \left(\mathcal{A} \cup 
-\left\lbrace \hat{a} \right\rbrace\right) \times \mathcal{I} \rightarrow \left\lbrace 1, \dots, k \right\rbrace$.
+We denote the new solver to be ranked by $\hat a \not\in A$ and define $\hat A := A \cup \hat a$ to denote all solvers including the new solver. 
+At the core of the model $\mathcal{M}$ is a prediction function $f : \mathcal{\hat A} \times \mathcal{I} \rightarrow \mathbb{R}$ that powers the decisions within the previously mentioned three steps.
+
+To be applicable to $\hat{a}$, we extend \emph{score} by replacing $\gamma(a, e)$ by the model's predictions $f : \mathcal{\hat A} \times \mathcal{I} \rightarrow \left\lbrace 1, \dots, k \right\rbrace$.
 
@@
 We implement a model-uncertainty based approach and a model-based information-gain sampling strategy.
-%, and a voting-based committee-disagreement approach.
 
@@
 \begin{equation*}
-  f'\!: \left(\mathcal{A} \cup \left\lbrace \hat{a} \right\rbrace\right) \times \mathcal{I} \rightarrow \left[0, 1\right]^k
+  f' : \mathcal{\hat A} \times \mathcal{I} \rightarrow \left[0, 1\right]^k
 \end{equation*}
@@
 
+%JB: if this is info gain regarding runtimes, doesn't Eq.
+(4) mean that we would prefer instances with low entropy of runtime distribution after considering the new solver, i.e., instances that are not very discriminatory?
+%TF: I don't think so. We want to maximize the reduction in entropy of the runtime labels. Low entropy afterward -> we know that the solver behaves similarly to existing ones.
+
 %Finally, the voting-based committee-disagreement approach requires a committee of $c$ models.
@@
 \subsubsection{Wilcoxon Stopping Criterion}
-The Wilcoxon stopping criterion stops the iterative active-learning algorithm when we are on average certain enough that the predicted runtime labels are sufficiently different in comparison to the other solvers.
+The Wilcoxon stopping criterion stops the iterative active learning algorithm when we are on average certain enough that the predicted runtime labels are sufficiently different in comparison to the other solvers.
+
 To be more specific, we stop when
@@
 drops below a certain threshold.
-%JB: if this is info gain regarding runtimes, doesn't Eq.
-(4) mean that we would prefer instances with low entropy of runtime distribution after considering the new solver, i.e., instances that are not very discriminatory?
-%TF: I don't think so. We want to maximize the reduction in entropy of the runtime labels. Low entropy afterward -> we know that the solver behaves similarly to existing ones.
 Thereby, \emph{wilcoxon} denotes the $p$-value of a Wilcoxon signed-rank test of the runtime-label distributions and $\gamma\!\left(a, \mathcal{I}\right)$ and $f\!\left(\hat{a}, \mathcal{I}\right)$ are the natural set extensions on $\gamma$ and $f$ respectively.
-To improve the stability of this criterion, we use an exponential moving average
-
-\begin{equation}
-  W_{\exp}^{\left(m\right)} := \beta W_{\hat{a}} + \left(1 - \beta\right) W_{\exp}^{\left(m - 1\right)} \enspace \textrm{,}
-\end{equation}
-
-whereby $m$ denotes the count of iterations within Algorithm~\ref{algALBenchmark} and $W_{\exp}^{\left(0\right)} := 1$.
+To improve the stability of this criterion, we use an exponential moving average to smooth out outliers.
+
+%whereby $m$ denotes the count of iterations within Algorithm~\ref{algALBenchmark} and $W_{\exp}^{\left(0\right)} := 1$.
+%\begin{equation}
+%  W_{\exp}^{\left(m\right)} := \beta W_{\hat{a}} + \left(1 - \beta\right) W_{\exp}^{\left(m - 1\right)}
+%\end{equation}
 

### 2022-10-10 paper/paper.tex
Revise Section 3.2 (Solver Model)
E.g., make it a bit shorter and improve argumentation for discretization.

@@
 
-Our framework to score a new solver~$\hat{a}$ follows a simple three-step procedure, outlined in Algorithm~\ref{algALBenchmark}:
-While the stopping condition is not met (line~1), we select \emph{one} problem instance from the pool of instances with the help of a prediction model~$\mathcal{M}$ (line~2) and evaluate the new solver~$\hat{a}$ on this instance (line~3).
+Our framework to score a new solver~$\hat{a}$ follows a three-step procedure, outlined in Algorithm~\ref{algALBenchmark}:
+While the stopping condition is not met (line~1), we select \emph{one} SAT instance from the pool of instances with the help of a prediction model~$\mathcal{M}$ (line~2) and evaluate the new solver~$\hat{a}$ on this instance (line~3).
 We then use the acquired result to update the prediction model~$\mathcal{M}$ (line~4).
@@
 Let benchmark instances $\mathcal{I}$, SAT solvers $\mathcal{A}$, and runtimes $r : \mathcal{A} \times \mathcal{I} \rightarrow \left[0, \tau\right]$ be given. 
-We denote the new solver to be ranked by $\hat a \not\in A$ and define $\hat A := A \cup \hat a$ to denote all solvers including the new solver. 
-At the core of the model $\mathcal{M}$ is a prediction function $f : \mathcal{\hat A} \times \mathcal{I} \rightarrow \mathbb{R}$ that powers the decisions within the previously mentioned three steps.
-
-To be applicable to $\hat{a}$, we extend \emph{score} by replacing $\gamma(a, e)$ by the model's predictions $f : \mathcal{\hat A} \times \mathcal{I} \rightarrow \left\lbrace 1, \dots, k \right\rbrace$.
+We denote the new solver to be ranked by $\hat a \not\in A$ and define $\hat A := A \cup \lbrace \hat a \rbrace$ to denote all solvers including the new solver. 
+Model $\mathcal{M}$ provides a prediction function $f : \mathcal{\hat A} \times \mathcal{I} \rightarrow \mathbb{R}$ that powers the decisions within our framework.
+
+To be applicable to a new solver $\hat{a}$, whose runtimes are unknown, we extend \emph{score} by replacing $\gamma(a, e)$ with the model's predictions $f : \mathcal{\hat A} \times \mathcal{I} \rightarrow \left\lbrace 1, \dots, k \right\rbrace$.
+%JB: "score" and "gamma" aren't defined yet, so there is nothing to extend; score isn't even defined here, only "s", while later sections still use "score"
 
@@
 
-We transform the runtimes into discrete runtime classes which are determined on a per instance basis.
-For each instance $e \in \mathcal{I}$, we use a single-link hierarchial clustering algorithm to assign the runtimes in $\bigl\{ r(a, e) \mid a \in A \bigr\}$ to one of the $k$ clusters $C_1, \dots, C_k$ such that the fastest runtimes for instance $e$ are in cluster $C_1$ and the slowest are in cluster $C_k$.
+We transform the runtimes into discrete runtime labels on a per-instance basis.
+For each instance $e \in \mathcal{I}$, we use a single-link hierarchical clustering algorithm to assign the runtimes in $\bigl\{ r(a, e) \mid a \in A \bigr\}$ to one of the $k$ clusters $C_1, \dots, C_k$ such that the fastest runtimes for instance $e$ are in cluster $C_1$ and the slowest are in cluster $C_k$.
 For instances with many timeouts, we have to manually fix the clustering results such that all timeouts $\tau$ are in cluster $C_k$.
-The runtime transformation function $\gamma_k : {\mathcal{A} \times \mathcal{I}} \rightarrow \left\lbrace 1, \dots, k \right\rbrace$ is then specified as follows.
+%TF: maybe a bit unclear. We do not manually fix anything. There is always one cluster with only time-outs if there are any.
+%JB: was wondering about that as well; how do we exactly create the timeout cluster?
+do we only cluster with k-1, removing timeout instances, or is the timeout cluster an ordinary cluster (which might also contain on-timeout instances)?
+The runtime transformation function $\gamma_k : {\mathcal{A} \times \mathcal{I}} \rightarrow \left\lbrace 1, \dots, k \right\rbrace$ is then specified as follows:
 $$\gamma_k(a, e) = j ~\Leftrightarrow~ r(a, e) \in C_j$$
@@
 Given an instance $e \in \mathcal{I}$, a solver $a \in A$ belongs to the $\gamma_k(a, e)$-fastest solvers on instance $e$. 
-In our preliminary experiments, we achieved higher accuracy with prediction of such discrete runtime labels than with prediction of raw runtimes. 
-The mean squared error of raw runtime prediction was within the same magnitude as the values to be predicted. 
+In preliminary experiments, we achieved higher accuracy predicting such discrete runtime labels than predicting raw runtimes. 
+The mean squared error of raw runtime regression was within the same magnitude as the values to be predicted. 
 %MI: how much more accurate are we with discrete runtime labels
-Research in portfolio solvers has also shown that directly predicting runtimes without any transformation does not work well in practice~\cite{NgokoCT19,CollauttiMMO13}.
-
-To determine solver ranks, we use the transformed runtimes $\gamma_k(a, e)$ in the adapted scoring function $s_k : A \rightarrow [1, 2 \cdot k]$ as follows.
+Research on portfolio solvers has also shown that directly predicting runtimes without any transformation does not work well in practice~\cite{CollauttiMMO13,NgokoCT19}.
+%JB: maybe we should phrase this more defensively or just leave out this argument (might be dataset-dependend), (1) e.g., SATzilla 2008 does predict runtimes and (2) we could also have treid to predict normalized runtimes like Collautti et al.
+instead of discretizing (though prediction performance would probably still have been bad in our case)
+
+To determine solver ranks, we use the transformed runtimes $\gamma_k(a, e)$ in the adapted scoring function $s_k : A \rightarrow [1, 2 \cdot k]$ as follows:
+%
 \begin{align}
@@
 \end{align}
-
+%
+I.e., we apply a PAR-2 scoring, which is commonly used in SAT competitions~\cite{FroleyksHIJS21}, on the discrete labels.
 The scoring function $s_k$ induces a ranking among solvers.
-In preliminary experiments, we compared the ranking induced by $s_3$ to the ranking induced by the PAR-2 score which is commonly used in SAT competitions~\cite{FroleyksHIJS21}. 
-The Spearman correlation of $s_3$ and PAR-2 ranking is about \SI{0.988} which is very close to the optimal value of 1~\cite{de2016comparing}.
-
-That discretized runtimes are sufficient for discriminating solvers can also be shown by comparing the fraction of solver pairs whose runtimes are significantly different. 
-In preliminary experiments, we used a Wilcoxon signed rank test with $\alpha = 0.05$ to compare solver pairs by their runtime vectors.
-Within the SAT~Competition~2022 Anniversary track, about \SI{89.95}{\%} of solver pairs are significantly different according to that measure.
-The strongly discretized runtimes $\gamma_2$ are sufficient to significantly distinguish about \SI{87.04}{\%} of solver pairs, while $\gamma_3$ is sufficient to distinguish about \SI{87.83}{\%} of solver pairs.
-
-% MI: Enough justification, commented out the following
-%For the SAT Competition~2022 Anniversary track instances, this ranking approach is capable of correctly deciding for almost all pairs of solvers which one is faster (about \SI{97.45}{\%}; $\sigma = \SI{3.68}{\%}$).
+
+However, we need to ensure that discretized labels still discriminate solvers and align with the actual PAR-2 ranking.
+We analyzed the ranking induced by $s_3$ in preliminary experiments with the SAT Competition~2022 Anniversary Track~\cite{sat2022}.
+According to a Wilcoxon-signed-rank test with $\alpha = 0.05$, \SI{87.83}{\%} of solver pairs have significantly different scores after discretization, only a slight drop compared to \SI{89.95}{\%} before discretization.
+%While only time-out labels (solver times out or not on a particular instance) are sufficient to significantly distinguish about \SI{87.04}{\%} of solver pairs.
+%\footnote{Based on hierarchical clustering with a log-single-link criterion for $k = 3$.}
+Further, our ranking approach is capable of correctly deciding for almost all (about \SI{97.45}{\%}; $\sigma = \SI{3.68}{\%}$) solver pairs which solver is faster.
 %Moreover, it decides all solvers correctly if the difference in the cluster-label-ranking scores between two solvers is at least \SI{0.11}{}.
-
-This shows that monotonicity and score distances between ranks align quite well between both ranking methods.
-The correlation between $s_k$-induced and PAR-2 induced ranks, the discriminatory power of $\gamma_k$, as well as that it enables us to use classification, show the clear advantage of runtime discretization.
+%Refer to Table~\ref{tab:discreteranking} for a comparison with the PAR-2 ranking.
+In particular, the Spearman correlation of $s_3$ and PAR-2 ranking is about \SI{0.988}{}, which is very close to the optimal value of 1~\cite{de2016comparing}.
+All these results show that discretized runtimes are suitable for our framework.
 
@@
 We implement a model-uncertainty based approach and a model-based information-gain sampling strategy.
-
-These methods require the model's predictions to also include probabilities for the $k$ possible runtime labels, i.e., 
-\begin{equation*}
-  f' : \mathcal{\hat A} \times \mathcal{I} \rightarrow \left[0, 1\right]^k
-\end{equation*}
+These methods require the model's predictions to also include probabilities for the $k$ possible runtime labels denoted by $f' : \mathcal{\hat A} \times \mathcal{I} \rightarrow \left[0, 1\right]^k$.
 
@@
 
-The model-uncertainty-based approach simply selects the instance that is closest to the model's decision boundary and, of course, not already selected, i.e.,
+The model-uncertainty-based approach simply selects the instance that is closest to the model's decision boundary.
+
 \begin{equation*}
@@
 The model-based information-gain sampling strategy selects the instance with the highest expected entropy reduction.
-To be more specific, we select the instance $e$ that maximizes
+To be more specific, we select the instance $e$ that maximizes $IG(e)$ which is specified as follows.
+
 \begin{equation*}
@@
 \end{equation*}
-whereby $\operatorname{H}\!\left(i\right)$ denotes the entropy of the runtime labels $\gamma\!\left(a, i\right)$ for all $a \in \mathcal{A}$ and $\operatorname{H}\!\left(i, n\right)$ denotes the entropy of the runtime labels including the runtime label of $\hat{a}$ which is $n$.
-The term $\operatorname{H}\!\left(i, n\right)$ is computed for every possible runtime label $n$.
+
+Here $\operatorname{H}(e)$ denotes the entropy of the runtime labels $\gamma(a, e)$ for all $a \in \mathcal{A}$ and $\operatorname{H}(e, n)$ denotes the entropy of the runtime labels including the runtime label of $\hat{a}$ which is $n$.
+The term $\operatorname{H}(e, n)$ is computed for every possible runtime label $n \in \{1, \dots, k\}$.
 
@@
 \subsubsection{Wilcoxon Stopping Criterion}
+
 The Wilcoxon stopping criterion stops the iterative active learning algorithm when we are on average certain enough that the predicted runtime labels are sufficiently different in comparison to the other solvers.
 
-To be more specific, we stop when
-\begin{equation}
-  W_{\hat{a}} := \frac{1}{\left\lvert \mathcal{A} \right\rvert} \sum_{a \in \mathcal{A}} \operatorname{wilcoxon}_{\alpha}\!\left(\gamma\!\left(a, \mathcal{I}\right), f\!\left(\hat{a}, \mathcal{I}\right) \right)
-\end{equation}
-drops below a certain threshold.
-Thereby, \emph{wilcoxon} denotes the $p$-value of a Wilcoxon signed-rank test of the runtime-label distributions and $\gamma\!\left(a, \mathcal{I}\right)$ and $f\!\left(\hat{a}, \mathcal{I}\right)$ are the natural set extensions on $\gamma$ and $f$ respectively.
-To improve the stability of this criterion, we use an exponential moving average to smooth out outliers.
-
-%whereby $m$ denotes the count of iterations within Algorithm~\ref{algALBenchmark} and $W_{\exp}^{\left(0\right)} := 1$.
-%\begin{equation}
-%  W_{\exp}^{\left(m\right)} := \beta W_{\hat{a}} + \left(1 - \beta\right) W_{\exp}^{\left(m - 1\right)}
-%\end{equation}
+To be more specific, we use the average $p$-value $W_{\hat{a}}$ of a Wilcoxon signed-rank test $w(S,P)$ of the two runtime label distributions $S=\{ \gamma(a, e) \mid e \in \mathcal{I} \}$ and $P=\{ f(\hat a, e) \mid e \in \mathcal{I} \}$.
+
+\begin{equation*}
+  W_{\hat{a}} := \frac{1}{\lvert \mathcal{A} \rvert} \sum_{a \in \mathcal{A}} \operatorname{w}(S, P)
+\end{equation*}
+
+To improve the stability of this criterion, we use an exponential moving average to smooth out outliers as follows and stop as soon as $W^{(i)}_{\hat{a}}$ drops below a certain threshold.
+
+\begin{align*}
+	W_{\exp}^{\left(0\right)} &:= 1\\
+	W_{\exp}^{\left(i\right)} &:= \beta W_{\hat{a}} + \left(1 - \beta\right) W_{\exp}^{\left(i - 1\right)}
+\end{align*}
 
@@
 It simply stops the active-learning process if within the last $l$ iterations the ranking that is induced by the model's predictions (Equation~\ref{eq:rankingeq}) remains unchanged.
-Note that the concrete values of $s_k(a)$ might still change.
+Note that the concrete values of $s_k(\hat a)$ might still change.
 We are solemnly interested in the induced ranking in this case.
@@
 \section{Experimental Design}
+\label{sec:exdesign}
 
@@
     History size: consider the latest 1, 10, 20, 30, or 40 predictions within a voting approach for stability.
-    Given a particular ranking decision, the latest $x$ predictions vote on the winner.
-    %JB: confuses me a bit as it sounds we rank first and then vote on a winner; isn't it the other way around?
+    The latest $x$ predictions vote on the winner.
+    %JB: confuses me a bit as it sounds like we rank first and then vote on a winner; isn't it the other way around?
+    %TF: Removed confusing sentence
     \item
@@
     %JB: as clarification: one these two solvers has to be the new solver, right? if yes, would write it explicitly
+    %TF: Not necessarily
     As discussed in Section~\ref{sec:main-model}, ranking decisions are all correct if the difference in scores between two solvers exceeds a certain value.
@@
     %JB: yes, but (1) even combination with "minimum amount" can make stopping very easy (see my example) and (2) fraction for convergence could also be relative to absolute number of instances (which is constant over AL run) rather than fraction of currently sampled instances (which increases over AL run)
+    %TF: I'm sure that both would make sense. Isn't 1% convergence duration of 5k about ~50 instances?
   \end{itemize}
@@
 
-As a runtime-label prediction model, we only make use of one method that has proven to be successful in preliminary experiments (cf.~Section~\ref{sec:evalprel}).
+As runtime-label prediction model, we only make use of one fixed model since an exhaustive grid search would be infeasible (cf.~Section~\ref{sec:evalprel}).
 Our model of choice is an ensemble, stacking a quadratic-discriminant analysis\footnote{With a regularization parameter of zero and a singular value threshold of zero} onto a random forest model\footnote{With \emph{entropy} splitting criterion and \emph{balanced} class weights} using a decision tree\footnote{With \emph{Gini} impurity splitting criterion; choosing the best split for each branching decision and maximum tree depth of 5 levels} to weight its members.
@@
 \label{sec:eval}
-Now that we are all set, we want to, first, discuss preliminary results including the choice of the runtime-discretization method and the choice of the machine-learning predictor.
-Thereafter, we look at end-to-end results and instance-family importance.
+
+Before evaluating our approach end-to-end as described in Section~\ref{sec:exdesign}, we discuss the choice of the runtime-discretization method and the machine-learning predictor.
+Thereafter, we look at the results of our active-learning approach.
+Selection decisions of the aforementioned approach also reveal the importance of different problem-instance families to our model.
+Instance families comprise instances that are derived from the same application domain, e.g., planning, cryptography, etc., and are a useful tool for both analyzing solver performance and portfolio creation.
 %JB: instance families not really motivated till now, only appear in abstract and in text regarding stratification
-
-\subsection{Preliminary Results}
+%TF: Added a sentence for clarification. We could introduce more details in section 4.3.
+
+\subsection{Runtime Prediction}
 \label{sec:evalprel}
+
 %JB: if we dedicate a whole section to these results, we should probably rename this section (not "preliminary" in the typical sense); for TACAS, I guess this can be cut a bit
-An exhaustive grid search of all approaches outlined in Section~\ref{sec:main} is infeasible.
-We are, therefore, narrowing down the search by filtering based on preliminary results.
-Since all relevant decisions, i.e., \emph{selection}, \emph{stopping}, and \emph{ranking}, revolve around the choice of prediction model, we are first looking at the \emph{Matthews Correlation Coefficient} (MCC) scores~\cite{gorodkin2004comparing,matthews1975comparison} that a particular prediction model achieves.
+%TF: cut the section down a bit, renamed to "runtime prediction"
+An exhaustive grid search of all active-learning approaches outlined in Section~\ref{sec:main} and different runtime-prediction models is infeasible.
+Therefore, we use the runtime-prediction model that produced the best results for all further experiments.
+We compare runtime predictors by looking at the \emph{Matthews Correlation Coefficient} (MCC) scores~\cite{gorodkin2004comparing,matthews1975comparison} of the classification task.
 MCC scores are great in dealing with class imbalances in contrast to conventional metrics like accuracy.
 
-\begin{table}
-  \centering
-  \caption{
-  	Time-out prediction Matthews Correlation Coefficient performances.
-  	Results are obtained using cross-validation.
-  	We provide \SI{64}{\%} of ground-truth for training, use \SI{16}{\%} as a validation set, and \SI{20}{\%} as a test set (5-fold cross-validation).
-  }
-  \label{tab:mcc}
-  \begin{tabular}{lc}
-    Prediction Model & Avg. MCC (Std.) \\
-    \hline
-    Stacking ensemble (QDA, RF)           & 0.9527 (0.0292) \\
-    Quadratic-discriminant analysis (QDA) & 0.9290 (0.0339) \\
-    Random forest (RF)                    & 0.8530 (0.0479) \\
-    AdaBoost with decision trees          & 0.8384 (0.0444) \\
-    AdaBoost with logistic regression     & 0.8158 (0.0985) \\
-    Decision tree                         & 0.8059 (0.0707) \\
-    Logistic regression                   & 0.8052 (0.1018) \\
-    kNN classifier                        & 0.7885 (0.1521) \\
-    MLP classifier                        & 0.7760 (0.1408) \\
-    Support-vector classifier             & 0.7757 (0.2149) \\
-    Naive Bayes                           & 0.7306 (0.1394) \\
-    \hline
-  \end{tabular}
-\end{table}
+% \begin{table}
+%   \centering
+%   \caption{
+%   	Time-out prediction Matthews Correlation Coefficient performances.
+%   	Results are obtained using cross-validation.
+%   	We provide \SI{64}{\%} of ground-truth for training, use \SI{16}{\%} as a validation set, and \SI{20}{\%} as a test set (5-fold cross-validation).
+%   }
+%   \label{tab:mcc}
+%   \begin{tabular}{lc}
+%     Prediction Model & Avg. MCC (Std.) \\
+%     \hline
+%     Stacking ensemble (QDA, RF)           & 0.9527 (0.0292) \\
+%     Quadratic-discriminant analysis (QDA) & 0.9290 (0.0339) \\
+%     Random forest (RF)                    & 0.8530 (0.0479) \\
+%     AdaBoost with decision trees          & 0.8384 (0.0444) \\
+%     AdaBoost with logistic regression     & 0.8158 (0.0985) \\
+%     Decision tree                         & 0.8059 (0.0707) \\
+%     Logistic regression                   & 0.8052 (0.1018) \\
+%     kNN classifier                        & 0.7885 (0.1521) \\
+%     MLP classifier                        & 0.7760 (0.1408) \\
+%     Support-vector classifier             & 0.7757 (0.2149) \\
+%     Naive Bayes                           & 0.7306 (0.1394) \\
+%     \hline
+%   \end{tabular}
+% \end{table}
 
@@
 
-To obtain the results in Table~\ref{tab:mcc}, we perform cross-validation among all solvers: we assume that all other solver runtimes are known and provide \SI{64}{\%} of ground-truth runtime labels of the target solver for training, use \SI{16}{\%} as a validation set for hyper-parameter tuning, and \SI{20}{\%} as a test set (5-fold cross-validation).
+%To obtain the results in Table~\ref{tab:mcc},
+We perform cross-validation among all solvers: we assume that all other solver runtimes are known and provide \SI{64}{\%} of ground-truth runtime labels of the target solver for training, use \SI{16}{\%} as a validation set for hyper-parameter tuning of the runtime predictor, and \SI{20}{\%} as a test set (5-fold cross-validation).
 This is repeated for each solver once.
@@
 We report the average performances on the test sets.
-Since the differences in the MCC-performance distributions of the best and second-best approaches are statistically significant regarding a Wilcoxon signed-rank test with $\alpha = \SI{5}{\%}$, we only use the best approach for all further experiments as a prediction model.
-
-As a runtime discretization technique, we use hierarchical clustering with $k = 3$ using a log-single-link criterion for all further experiments.
+
+ML classification models that we have tried include random forests, decision trees, quadratic-discriminant analysis, AdaBoost, logistic regression, $k$-nearest neighbor, multi-layer perceptrons, support vector machines, and naive Bayes.
+A stacking ensemble consisting of quadratic-discriminant analysis and random forests performed best.
+It produces an MCC score of \SI{0.9527}{} for time-out prediction and \SI{0.8882}{} for runtime-label prediction.
+The results are significantly better than any other model that we have tried regarding a Wilcoxon signed-rank test with $\alpha = \SI{5}{\%}$.
+
+%Since the differences in the MCC-performance distributions of the best and second-best approaches are statistically significant regarding a Wilcoxon signed-rank test with $\alpha = \SI{5}{\%}$, we only use the best approach for all further experiments as a prediction model.
+
+For runtime discretization, we use hierarchical clustering with $k = 3$ using a log-single-link criterion for all further experiments.
 Each non-time-out runtime starts in a separate interval.
@@
 As already mentioned in Section~\ref{sec:main-model}, this approach has good properties regarding the predictiveness of the PAR-2 score ranking and its discriminatory power.
-Also, preliminary experiments showed that our stacking ensemble performed particularly well for this kind of discretization method.
+Also, experiments showed that our stacking ensemble performed particularly well for this kind of discretization method.
 Other clustering approaches that we have tried include hierarchical clustering with mean-, median- and complete-link criteria, as well as $k$-means and spectral clustering.
-Predicting runtime labels in a similar experiment as above produced an MCC score of on average \SI{88.82}{\%} ($\sigma = \SI{4.75}{\%}$; in comparison to \SI{95.27}{\%} for predicting time-outs).
+%Predicting runtime labels in a similar experiment as above produced an MCC score of on average \SI{88.82}{\%} ($\sigma = \SI{4.75}{\%}$; in comparison to \SI{95.27}{\%} for predicting time-outs).
 %JB: "similar experiment as above" - reference unclear; what do you want to say?
+%TF: Removed the sentence
 
@@
 \subsubsection{Tuning our algorithm.}
+
 Our main end-to-end experiments follow the evaluation framework that has been introduced in Section~\ref{sec:evalframe}.
 Fig.~\ref{fig:e2eallsolvers} shows the performance of the approaches listed in Section~\ref{sec:hyper} on $O_{\operatorname{rt}}$-$O_{\operatorname{acc}}$-diagrams for the smaller hyper-parameter-tuning dataset.
-The plotted lines represent the Pareto front for each possible choice of ranking approaches (Fig.~\ref{fig:annitraincolorranking}), selection approaches (Fig.~\ref{fig:annitraincolorselection}), and stopping criteria (Fig.~\ref{fig:annitraincolorstopping}).
+The plotted lines represent the best performing configurations per ranking approach (Fig.~\ref{fig:annitraincolorranking}), per selection approach (Fig.~\ref{fig:annitraincolorselection}), and per stopping criterion (Fig.~\ref{fig:annitraincolorstopping}).
 %JB: as far as I understand it, the lines only compare some top-level hyperparameters, but for each of these hyperparemeters, take the best (= Pareto front) out of some more fine-grained settings (the hyperparameters of the hyperparameters, so to speak), so it's no really "each possible choice" in the sense of actual, individual configurations, but rather a combination of configurations
@@
 
-\subsubsection{Instance-family importance.}
+\subsection{Instance-Family Importance}
+
 For all configurations of our approach that sample on average less than \SI{10}{\%} of all 5355 instances, we look at the particular instances and families that are selected by our active-learning approach.

### 2022-10-10 paper/paper.tex
Rewrite section 5.1

--- 2022-10-10 paper/paper.tex
@@
 \label{sec:eval}
-Now that we are all set, we want to, first, discuss preliminary results including the choice of the runtime-discretization method and the choice of the machine-learning predictor.
-Thereafter, we look at end-to-end results and instance-family importance.
+Before evaluating our approach end-to-end as described in Section~\ref{sec:exdesign}, we discuss the choice of the runtime-discretization method and the machine-learning predictor.
+Thereafter, we look at the results of our active-learning approach.
+Selection decisions of the aforementioned approach also reveal the importance of different problem-instance families to our model.
+Instance families comprise instances that are derived from the same application domain, e.g., planning, cryptography, etc., and are a useful tool for both analyzing solver performance and portfolio creation.
 %JB: instance families not really motivated till now, only appear in abstract and in text regarding stratification
-
-\subsection{Preliminary Results}
+%TF: Added a sentence for clarification. We could introduce more details in section 4.3.
+
+\subsection{Runtime Prediction}
 \label{sec:evalprel}
 %JB: if we dedicate a whole section to these results, we should probably rename this section (not "preliminary" in the typical sense); for TACAS, I guess this can be cut a bit
-An exhaustive grid search of all approaches outlined in Section~\ref{sec:main} is infeasible.
-We are, therefore, narrowing down the search by filtering based on preliminary results.
-Since all relevant decisions, i.e., \emph{selection}, \emph{stopping}, and \emph{ranking}, revolve around the choice of prediction model, we are first looking at the \emph{Matthews Correlation Coefficient} (MCC) scores~\cite{gorodkin2004comparing,matthews1975comparison} that a particular prediction model achieves.
+%TF: cut the section down a bit, renamed to "runtime prediction"
+An exhaustive grid search of all active-learning approaches outlined in Section~\ref{sec:main} and different runtime-prediction models is infeasible.
+Therefore, we use the runtime-prediction model that produced the best results for all further experiments.
+We compare runtime predictors by looking at the \emph{Matthews Correlation Coefficient} (MCC) scores~\cite{gorodkin2004comparing,matthews1975comparison} of the classification task.
 MCC scores are great in dealing with class imbalances in contrast to conventional metrics like accuracy.
 
-\begin{table}
-  \centering
-  \caption{
-  	Time-out prediction Matthews Correlation Coefficient performances.
-  	Results are obtained using cross-validation.
-  	We provide \SI{64}{\%} of ground-truth for training, use \SI{16}{\%} as a validation set, and \SI{20}{\%} as a test set (5-fold cross-validation).
-  }
-  \label{tab:mcc}
-  \begin{tabular}{lc}
-    Prediction Model & Avg. MCC (Std.) \\
-    \hline
-    Stacking ensemble (QDA, RF)           & 0.9527 (0.0292) \\
-    Quadratic-discriminant analysis (QDA) & 0.9290 (0.0339) \\
-    Random forest (RF)                    & 0.8530 (0.0479) \\
-    AdaBoost with decision trees          & 0.8384 (0.0444) \\
-    AdaBoost with logistic regression     & 0.8158 (0.0985) \\
-    Decision tree                         & 0.8059 (0.0707) \\
-    Logistic regression                   & 0.8052 (0.1018) \\
-    kNN classifier                        & 0.7885 (0.1521) \\
-    MLP classifier                        & 0.7760 (0.1408) \\
-    Support-vector classifier             & 0.7757 (0.2149) \\
-    Naive Bayes                           & 0.7306 (0.1394) \\
-    \hline
-  \end{tabular}
-\end{table}
+% \begin{table}
+%   \centering
+%   \caption{
+%   	Time-out prediction Matthews Correlation Coefficient performances.
+%   	Results are obtained using cross-validation.
+%   	We provide \SI{64}{\%} of ground-truth for training, use \SI{16}{\%} as a validation set, and \SI{20}{\%} as a test set (5-fold cross-validation).
+%   }
+%   \label{tab:mcc}
+%   \begin{tabular}{lc}
+%     Prediction Model & Avg. MCC (Std.) \\
+%     \hline
+%     Stacking ensemble (QDA, RF)           & 0.9527 (0.0292) \\
+%     Quadratic-discriminant analysis (QDA) & 0.9290 (0.0339) \\
+%     Random forest (RF)                    & 0.8530 (0.0479) \\
+%     AdaBoost with decision trees          & 0.8384 (0.0444) \\
+%     AdaBoost with logistic regression     & 0.8158 (0.0985) \\
+%     Decision tree                         & 0.8059 (0.0707) \\
+%     Logistic regression                   & 0.8052 (0.1018) \\
+%     kNN classifier                        & 0.7885 (0.1521) \\
+%     MLP classifier                        & 0.7760 (0.1408) \\
+%     Support-vector classifier             & 0.7757 (0.2149) \\
+%     Naive Bayes                           & 0.7306 (0.1394) \\
+%     \hline
+%   \end{tabular}
+% \end{table}
 

### 2022-10-10 paper/paper.tex
Address comments

@@
 For instances with many timeouts, we have to manually fix the clustering results such that all timeouts $\tau$ are in cluster $C_k$.
+%TF: maybe a bit unclear. We do not manually fix anything. There is always one cluster with only time-outs if there are any.
 The runtime transformation function $\gamma_k : {\mathcal{A} \times \mathcal{I}} \rightarrow \left\lbrace 1, \dots, k \right\rbrace$ is then specified as follows.
@@
 \label{sec:eval}
+
 Before evaluating our approach end-to-end as described in Section~\ref{sec:exdesign}, we discuss the choice of the runtime-discretization method and the machine-learning predictor.
@@
 \label{sec:evalprel}
+
 %JB: if we dedicate a whole section to these results, we should probably rename this section (not "preliminary" in the typical sense); for TACAS, I guess this can be cut a bit

### 2022-10-10 paper/paper.tex
partially improved readability of Algo 2

@@
 
-We transform the runtimes into discrete runtime classes which are determined on a per instance basis.
-For each instance $e \in \mathcal{I}$, we use a single-link hierarchial clustering algorithm to assign the runtimes in $\bigl\{ r(a, e) \mid a \in A \bigr\}$ to one of the $k$ clusters $C_1, \dots, C_k$ such that the fastest runtimes for instance $e$ are in cluster $C_1$ and the slowest are in cluster $C_k$.
+We transform the runtimes into discrete runtime labels on a per-instance basis.
+For each instance $e \in \mathcal{I}$, we use a single-link hierarchical clustering algorithm to assign the runtimes in $\bigl\{ r(a, e) \mid a \in A \bigr\}$ to one of the $k$ clusters $C_1, \dots, C_k$ such that the fastest runtimes for instance $e$ are in cluster $C_1$ and the slowest are in cluster $C_k$.
 For instances with many timeouts, we have to manually fix the clustering results such that all timeouts $\tau$ are in cluster $C_k$.
 %TF: maybe a bit unclear. We do not manually fix anything. There is always one cluster with only time-outs if there are any.
-The runtime transformation function $\gamma_k : {\mathcal{A} \times \mathcal{I}} \rightarrow \left\lbrace 1, \dots, k \right\rbrace$ is then specified as follows.
+%JB: was wondering about that as well; how do we exactly create the timeout cluster?
+do we only cluster with k-1, removing timeout instances, or is the timeout cluster an ordinary cluster (which might also contain on-timeout instances)?
+The runtime transformation function $\gamma_k : {\mathcal{A} \times \mathcal{I}} \rightarrow \left\lbrace 1, \dots, k \right\rbrace$ is then specified as follows:
 $$\gamma_k(a, e) = j ~\Leftrightarrow~ r(a, e) \in C_j$$
@@
 As already stated in the introductory section, this work addresses the \emph{New-Solver Problem} (cf.~Definition~\ref{def:new-solver-problem}).
-By running Algorithm~\ref{algALBenchmark}, we obtain a prediction model $\mathcal{M}$ that provides us with a scoring function \emph{score}.
-Also, it outputs the fraction of runtime that is needed to arrive at its conclusion.
+We create a prediction model $\mathcal{M}$ as described in Section~\ref{sec:main-model} that provides us with a scoring function $s_k$.
+% MI: Let this be the evaluation result:
+%Also, it outputs the fraction of runtime that is needed to arrive at its conclusion.
 First and foremost, our goal is to provide the engineer of new SAT solvers with an accurate ranking.
-We define the \emph{ranking accuracy} $O_{\operatorname{acc}} \in \left[0, 1\right]$ (higher is better) by the fraction of pairs $\left(\hat{a}, a\right)$ for all $a \in \mathcal{A}$ that are decided correctly by the given ranking (cf.
-Algorithm~\ref{alg:eval}).
+We define the \emph{ranking accuracy} $O_{\operatorname{acc}} \in \left[0, 1\right]$ (higher is better) by the fraction of pairs $\left(\hat{a}, a\right)$ for all $a \in \mathcal{A}$ that are decided correctly by the given ranking (cf.~Algorithm~\ref{alg:eval}).
 For now, we exclude the equality of solvers since data shows that the PAR-2 scores of all our solvers are different.
@@
 The ranking accuracy is affected by whether this decision is correctly made.
-Matricon et. al.~\cite{MatriconAFSH21} also evaluate their approach by the fraction of solver pairs that is correctly decided.
+%Matricon et. al.~\cite{MatriconAFSH21} also evaluate their approach by the fraction of solver pairs that is correctly decided.
 Second, we also have to optimize for runtime.
@@
 
-  \KwIn{A pool of instances $\mathcal{I}$, a set of solvers $\mathcal{S}$ with known runtimes $r\!: \mathcal{S} \times \mathcal{I} \rightarrow \left[0, \tau\right] \cup \left\lbrace \perp \right\rbrace$, and PAR-2 scores $par2\!: \mathcal{S} \rightarrow \mathbb{R}$ for each solver.}
+  %\KwIn{A pool of instances $\mathcal{I}$, a set of solvers $\mathcal{S}$ with known runtimes $r\!: \mathcal{S} \times \mathcal{I} \rightarrow \left[0, \tau\right] \cup \left\lbrace \perp \right\rbrace$, and PAR-2 scores $par2\!: \mathcal{S} \rightarrow \mathbb{R}$ for each solver.}
+  \KwIn{Solvers $\mathcal{A}$, Instaces $\mathcal{I}$, Runtimes $r : \mathcal{A} \times \mathcal{I} \rightarrow [0, \tau]$}
+  \KwOut{Average ranking accuracy $\bar{O}_{\operatorname{acc}}$, Average fraction of runtime $\bar{O}_{\operatorname{rt}}$}
+  \BlankLine
 
@@
 
-  \For{$\hat{a} \in \mathcal{S}$}{
-    $\mathcal{A} \leftarrow \mathcal{S} \setminus \left\lbrace \hat{a} \right\rbrace$
-
-    $\left( score,\, \tilde{\mathcal{I}} \right) \leftarrow \operatorname{runALAlgorithm}\!\left(\mathcal{I}, \mathcal{A}, \left.r\right|_{\mathcal{A} \times \mathcal{I}}, \hat{a}\right)$ \tcp*{Refer to Algorithm~\ref{algALBenchmark}}
-
-    $O_{\operatorname{acc}} \leftarrow \left(\sum_{a \in \mathcal{A}} \left[\operatorname{sign}\!\left(score\!\left(a\right) - score\!\left(\hat{a}\right)\right) = \operatorname{sign}\!\left(par2\!\left(a\right) - par2\!\left(\hat{a}\right)\right) \right]\right) / |\mathcal{A}|$
+  \For{$\hat{a} \in \mathcal{A}$}{
+    $\mathcal{A}' \leftarrow \mathcal{S} \setminus \left\lbrace \hat{a} \right\rbrace$
+
+    $\left( s, \tilde{\mathcal{I}} \right) \leftarrow \operatorname{runALAlgorithm}\!\left(\mathcal{I}, \mathcal{A}', r, \hat{a}\right)$ \tcp*{Refer to Algorithm~\ref{algALBenchmark}}
+
+    $O_{\operatorname{acc}} \leftarrow \frac{1}{|\mathcal{A}|} \sum\limits_{a \in \mathcal{A}} \left[\operatorname{sign}(s(a) - s(\hat{a})) = \operatorname{sign}(par2(a) - par2(\hat{a})) \right]$
 
@@
 
-    $O_{\operatorname{rt}} \leftarrow \left(\sum_{i \in \tilde{\mathcal{I}}} r\!\left(i, \hat{a}\right)\right) / \left(\sum_{i \in \mathcal{I}} r\!\left(i, \hat{a}\right)\right)$
+	$r' \leftarrow \sum\limits_{i \in \tilde{\mathcal{I}}} r(i, \hat{a})$\;
+	$r \leftarrow \sum\limits_{i \in \mathcal{I}} r(i, \hat{a})$\;
+    $O_{\operatorname{rt}} \leftarrow \frac{r'}{r}$
 
@@
 
-  $\left( \bar{O}_{\operatorname{acc}},\, \bar{O}_{\operatorname{rt}} \right) \leftarrow \operatorname{mean}\!\left( res \right)$
-
-  \KwOut{Return the average ranking accuracy $\bar{O}_{\operatorname{acc}}$ and the average fraction of runtime needed $\bar{O}_{\operatorname{rt}}$.}
+  $\left( \bar{O}_{\operatorname{acc}},\, \bar{O}_{\operatorname{rt}} \right) \leftarrow \operatorname{mean}\!\left( res \right)$ \;
+  \Return $\left( \bar{O}_{\operatorname{acc}},\, \bar{O}_{\operatorname{rt}} \right)$
 \end{algorithm}

### 2022-10-10 paper/paper.tex
Revise Section 3.3 (Instance Sampling)
Mostly language, slight change of information-gain formula.

@@
 
-\subsubsection{Static benchmark selection.}
+\subsubsection{Static Benchmark Selection.}
 
@@
 
-\subsubsection{Algorithm configuration.}
+\subsubsection{Algorithm Configuration.}
 
@@
 
-\subsubsection{Incremental benchmark selection.}
+\subsubsection{Incremental Benchmark Selection.}
 
@@
 
-\subsubsection{Active learning (AL).}
+\subsubsection{Active Learning (AL).}
 
@@
 %JB: "score" and "gamma" aren't defined yet, so there is nothing to extend; score isn't even defined here, only "s", while later sections still use "score"
+%JB: also, each model only predicts for one solver (there is no model covering multiple solvers), so I'm not sure whether introducing the solver as parameter makes sense
 
@@
 
-Selecting an instance based on the model is the core functionality of our method (cf. Algorithm~\ref{algALBenchmark}, Line 2). 
-In this section, we introduce our sampling strategies that make use of the model's label-prediction function $f$ which are inspired by existing work within the realms of active learning~\cite{settles2009active}.
-We implement a model-uncertainty based approach and a model-based information-gain sampling strategy.
-These methods require the model's predictions to also include probabilities for the $k$ possible runtime labels denoted by $f' : \mathcal{\hat A} \times \mathcal{I} \rightarrow \left[0, 1\right]^k$.
-
-\subsubsection{Model Uncertainty Based Sampling.}
-
-The model-uncertainty-based approach simply selects the instance that is closest to the model's decision boundary.
+Selecting an instance based on the model is a core functionality of our framework (cf. Algorithm~\ref{algALBenchmark}, Line 2).
+In this section, we introduce our sampling strategies, which make use of the model's label-prediction function $f$ and are inspired by existing work within the realms of active learning~\cite{settles2009active}.
+We implement a model-uncertainty and an information-gain sampling strategy.
+These methods require the model's predictions to also include probabilities for the $k$ possible runtime labels.
+Let $f' : \mathcal{\hat A} \times \mathcal{I} \rightarrow \left[0, 1\right]^k$ denote this modified prediction function.
+
+\subsubsection{Model Uncertainty Sampling.}
+
+The model-uncertainty sampling strategy simply selects the instance that is closest to the model's decision boundary.
 
@@
   \underset{e \in \mathcal{I} \setminus \tilde{\mathcal{I}}}{\arg\min} \left\lvert \frac{1}{k} - \max_{n \in \left\lbrace 1, \dots, k \right\rbrace} f'\!\left(\hat{a}, e\right)_{n} \right\rvert
+  %JB: \tilde I is undefined
 \end{equation*}
 
-\subsubsection{Information Gain Based Sampling.}
-
-The model-based information-gain sampling strategy selects the instance with the highest expected entropy reduction.
-To be more specific, we select the instance $e$ that maximizes $IG(e)$ which is specified as follows.
+\subsubsection{Information Gain Sampling.}
+
+The information-gain sampling strategy selects the instance with the highest expected entropy reduction regarding the runtime labels of the instance.
+To be more specific, we select the instance $e$ that maximizes $IG(e)$, which is specified as follows:
 
 \begin{equation*}
-  \operatorname{IG}(e) := \operatorname{H}(e) - \sum_{n = 1}^{k} f'(\hat{a}, e)_{n} \operatorname{H}(e, n)
+  \operatorname{IG}(e) := \operatorname{H}(e) - \sum_{n = 1}^{k} f'(\hat{a}, e)_{n} \operatorname{\hat H}_n(e)
 \end{equation*}
 
-Here $\operatorname{H}(e)$ denotes the entropy of the runtime labels $\gamma(a, e)$ for all $a \in \mathcal{A}$ and $\operatorname{H}(e, n)$ denotes the entropy of the runtime labels including the runtime label of $\hat{a}$ which is $n$.
-The term $\operatorname{H}(e, n)$ is computed for every possible runtime label $n \in \{1, \dots, k\}$.
+Here, $\operatorname{H}(e)$ denotes the entropy of the runtime labels $\gamma(a, e)$ over all $a \in \mathcal{A}$ and $\operatorname{H}(e, n)$ denotes the entropy of these labels plus $n$ as the runtime label for $\hat{a}$.
+The term $\operatorname{\hat H}_n(e)$ is computed for every possible runtime label $n \in \{1, \dots, k\}$.
 
@@
 %TF: I don't think so. We want to maximize the reduction in entropy of the runtime labels. Low entropy afterward -> we know that the solver behaves similarly to existing ones.
+%JB: but if we want to discriminate between solvers, what does it help to know that solver behaves similarly to other solvers?
 
@@
 %Bossek and Wagner~\cite{Bossek021a} introduce a custom ranking-score metric incorporating the number of \emph{good} pairs and \emph{bad} pairs.
-%JB: some selection methods, like this one, are not mentioned later anymore (in particular, not in experimental design), so we don't need to describe them in detail (maybe mention them shortly if used in preliminary experiments, but turned out to be not useful)
-%TF: I've cut the section down a bit
 %A good solver pair, regarding a particular instance, are solvers that perform according to their global ranking.

### 2022-10-10 paper/paper.tex
Revise Section 3.4 (Stopping Criteria)
Mostly language.

@@
 The runtime transformation function $\gamma_k : {\mathcal{A} \times \mathcal{I}} \rightarrow \left\lbrace 1, \dots, k \right\rbrace$ is then specified as follows:
+%
 $$\gamma_k(a, e) = j ~\Leftrightarrow~ r(a, e) \in C_j$$
@@
 The model-uncertainty sampling strategy simply selects the instance that is closest to the model's decision boundary.
-
+%
 \begin{equation*}
@@
 To be more specific, we select the instance $e$ that maximizes $IG(e)$, which is specified as follows:
-
+%
 \begin{equation*}
@@
 
-In this section, we present two dynamic stopping criteria, the Wilcoxon and the Model Uncertainty stopping criteria (cf. Algorithm~\ref{algALBenchmark}, Line~1).
+In this section, we present two dynamic stopping criteria, the Wilcoxon and the Model Uncertainty Stopping Criterion (cf. Algorithm~\ref{algALBenchmark}, Line~1).
 
@@
 
-The Wilcoxon stopping criterion stops the iterative active learning algorithm when we are on average certain enough that the predicted runtime labels are sufficiently different in comparison to the other solvers.
-
-To be more specific, we use the average $p$-value $W_{\hat{a}}$ of a Wilcoxon signed-rank test $w(S,P)$ of the two runtime label distributions $S=\{ \gamma(a, e) \mid e \in \mathcal{I} \}$ and $P=\{ f(\hat a, e) \mid e \in \mathcal{I} \}$.
-
+The Wilcoxon stopping criterion stops the active-learning process when we are certain enough that the predicted runtime labels of the new solver are sufficiently different from the labels of the existing solvers.
+We use the average $p$-value $W_{\hat{a}}$ of a Wilcoxon signed-rank test $w(S,P)$ of the two runtime label distributions $S=\{ \gamma(a, e) \mid e \in \mathcal{I} \}$ for an existing solver $a$ and $P=\{ f(\hat a, e) \mid e \in \mathcal{I} \}$ for the new solver $\hat{a}$:
+%
 \begin{equation*}
@@
 \end{equation*}
-
-To improve the stability of this criterion, we use an exponential moving average to smooth out outliers as follows and stop as soon as $W^{(i)}_{\hat{a}}$ drops below a certain threshold.
-
+%
+To improve the stability of this criterion, we use an exponential moving average to smooth out outliers and stop as soon as $W^{(i)}_{\exp}$ drops below a certain threshold:
+%
 \begin{align*}
@@
 The model-convergence stopping criterion is less sophisticated in comparison.
-It simply stops the active-learning process if within the last $l$ iterations the ranking that is induced by the model's predictions (Equation~\ref{eq:rankingeq}) remains unchanged.
+It stops the active-learning process if the ranking induced by the model's predictions (Equation~\ref{eq:rankingeq}) remains unchanged within the last $l$ iterations.
 Note that the concrete values of $s_k(\hat a)$ might still change.

### 2022-10-10 paper/paper.tex
Revise section 5

@@
 Selection decisions of the aforementioned approach also reveal the importance of different problem-instance families to our model.
-Instance families comprise instances that are derived from the same application domain, e.g., planning, cryptography, etc., and are a useful tool for both analyzing solver performance and portfolio creation.
+Instance families comprise instances that are derived from the same application domain, e.g., planning, cryptography, etc., and are a useful tool for both analyzing solver performance and portfolios.
 %JB: instance families not really motivated till now, only appear in abstract and in text regarding stratification
@@
 Fig.~\ref{fig:e2eallsolvers} shows the performance of the approaches listed in Section~\ref{sec:hyper} on $O_{\operatorname{rt}}$-$O_{\operatorname{acc}}$-diagrams for the smaller hyper-parameter-tuning dataset.
-The plotted lines represent the best performing configurations per ranking approach (Fig.~\ref{fig:annitraincolorranking}), per selection approach (Fig.~\ref{fig:annitraincolorselection}), and per stopping criterion (Fig.~\ref{fig:annitraincolorstopping}).
+Evaluating a particular configuration with Algorithm~\ref{alg:eval}, returns a point $\left(O_{\operatorname{rt}},\, O_{\operatorname{acc}}\right)$.
+The plotted lines represent the best performing configurations (convex hull) per top-level hyper-parameter choice, i.e., per ranking approach (Fig.~\ref{fig:annitraincolorranking}), per selection approach (Fig.~\ref{fig:annitraincolorselection}), and per stopping criterion (Fig.~\ref{fig:annitraincolorstopping}).
 %JB: as far as I understand it, the lines only compare some top-level hyperparameters, but for each of these hyperparemeters, take the best (= Pareto front) out of some more fine-grained settings (the hyperparameters of the hyperparameters, so to speak), so it's no really "each possible choice" in the sense of actual, individual configurations, but rather a combination of configurations
-Colors indicate the different hyper-parameter instantiations.
-To be more specific, they represent the Pareto front of all configurations with the respective instantiations.
-
-Regarding the ranking approach (Fig.~\ref{fig:annitraincolorranking}), using the ranking that is induced by our runtime-label predictions consistently performs better than the partially observed PAR-2 ranking for each possible value of $\delta$.
+%Colors indicate the different hyper-parameter instantiations.
+%To be more specific, they represent the Pareto front of all configurations with the respective instantiations.
+
+Regarding ranking approaches (Fig.~\ref{fig:annitraincolorranking}), using the $s_3$-induced ranking consistently outperforms the partially observed PAR-2 ranking for each possible value of $\delta$.
 %JB: as in previous comment: consistent outperformance not guaranteed in case the lines are not individual approaches, but Pareto fronts within the compared categories (which means that different parts of the line might be related to different hyperparameter settings)
-%This was expected since a possible bias in the selection decisions of the model is accounted for when using the model's predictions for ranking.
+%TF: The lines are in any case no individual approaches as individual approaches are only points. So, any lines in this plot are a "combination" of configs. It's only a question of how to combine.
 This is expected since selection decisions are not random.
@@
 However, the model-based information-gain sampling is beneficial if runtime is strongly favored (very small $\delta$; runtime fraction less than \SI{5}{\%}).
+Random sampling is outperformed by the model-based uncertainty strategy in all cases showing the benefit of active learning.
+By actively choosing the next experiment to run based on a model, we are more accurate in predicting the PAR-2 ranking of the solvers on the underlying dataset.
 %JB: we should also describe performance of random sampling, which is visible in plot as well
+%TF: added a few sentences
 
 Regarding the stopping criterion (Fig.~\ref{fig:annitraincolorstopping}), the ranking convergence criterion performs the most consistently well.
-If PAR-2 accuracy is strongly favored (very high $\delta$), the Wilcoxon stopping criterion performs better.
+If accuracy is strongly favored (very high $\delta$), the Wilcoxon stopping criterion performs better.
 
@@
   %JB: not sure if the color palette in the plots is grayscale-/color-blind-friendly; consider changing the palette (see https://colorbrewer2.org/#type=qualitative&scheme=Set3&n=3 ) and using different line types for each color
+  %TF: done
   \label{fig:e2eallsolvers}
@@
 
-Table~\ref{tab:paretotable} also shows the performance of particular hyper-parameter choices for different values of $\delta$.
-%JB: some plot for comparison would be better than a truncated table showing a variety hyperparameter configs
-Thereby, we observe that for small $\delta$ (runtime more important) our approaches preferably use easy instances to get a good sense of the PAR-2 ranking fast, e.g., row with $\delta = 0.15$.
-%JB: the "e.g." part needs more explanation (how can we see that easy instances used? -> fraction runtime vs. fraction instances)
-In contrast to that, harder instances are preferred for higher values of $\delta$, i.e., the sampled fraction of runtime is greater than the fraction of instances.
-In our opinion, the row with $\delta = 0.85$ represents the best trade-off between runtime and PAR-2 accuracy.
-%JB: this is a scientific article, not an opinion piece :-); all deltas represent different trade-offs, so I would not claim to have found a best trade-off (probably depends on user which kind of trade-off is most desired)
+Figure~\ref{fig:annitrainoptgoal} also shows an interesting difference between different optimization goals.
+It shows how different configurations select instances with different difficulties.
+If it is, on the one hand, desired to get a \emph{rough} estimate of a solver's performance fast (low $\delta$), corresponding approaches favor selecting many \emph{easy} problem instances (fraction of instances bigger than the fraction of runtime within $\tilde{\mathcal{I}}$).
+By having many observations, it is easier to build a model.
+If it is, on the other hand, desired to get a \emph{good} estimate of a solver's performance in a moderate amount of time (high $\delta$), corresponding approaches favor selecting fewer \emph{difficult} problem instances (fraction of instances smaller than the fraction of runtime within $\tilde{\mathcal{I}}$).
 
 \subsubsection{Full dataset evaluation.}
+
 Having selected the most promising hyper-parameters, we run our end-to-end active-learning experiments on the complete Anniversary Track dataset (5355 instances).
-The best-performing approach for $\delta = 0.85$ (since this was the most promising trade-off on the tuning set) uses runtime-label prediction ranking, model-based uncertainty sampling, and a ranking-convergence stopping criterion.
-Thereby, the sampled instances account for about \SI{5.24}{\%} of instances and about \SI{10.35}{\%} of runtime.
-Moreover, the PAR-2 accuracy is about \SI{92.33}{\%}, the average Spearman correlation between the PAR-2 ranking and the predicted ranking scores about \SI{0.9572}{}, and the weighted objective $O_{\delta} = 0.9193$.
+The best-performing approach for $\delta = 0.85$ uses runtime-label prediction ranking, model-based uncertainty sampling, and a ranking-convergence stopping criterion.
+It can predict a new solver's PAR-2 ranking with about \SI{92.33}{\%} accuracy ($O_{\operatorname{acc}}$), while only needing \SI{10.35}{\%} of the runtime that would be needed to evaluate the solver on the full dataset ($O_{\operatorname{rt}}$).
+This trade-off provides a better way of prototyping and evaluating new SAT solvers rather than evaluating them on a manually picked subset based on domain knowledge.
+The sampled instances account for about \SI{5.24}{\%} of instances indicating the preference for selecting more difficult instances as already discussed in the previous section.
+The average Spearman correlation between the PAR-2 ranking and the predicted ranking scores is about \SI{0.9572}{} indicating that the differences in PAR-2 scores are quite similar to $s_3$ scores.
+Because of that, the runtime-prediction model helps determine on what problem instances solvers perform similarly or are better/worse.
 %JB: many numbers, but some qualitative (and positively sounding, i.e., saying that our approach indeed is useful) interpretation would be nice
+%TF: revised this section
 
@@
 
-For all configurations of our approach that sample on average less than \SI{10}{\%} of all 5355 instances, we look at the particular instances and families that are selected by our active-learning approach.
-Thereby, the average benchmark set contains about \SI{192.34}{} instances.
-Table~\ref{tab:familiesa} lists the ten most frequent families that are chosen.
-For example, \emph{planning} instances account for on average \SI{19.14}{} of the \SI{192.34}{} benchmark instances.
-In contrast to that, Table~\ref{tab:familiesb} shows the families that are sampled the most often relative to their occurrences in the underlying dataset.
-There are, for example, only nine \emph{circuit-multiplier} instances in the complete dataset.
-However, we sample on average \SI{1.59}{} of them.
-%JB: there should be some (qualitative) interpretation (based on our results, is the sample interesting/special family-wise or not?) rather than just dropping a table into the paper and mainly explaining its structure
+Selection decisions of our approach also reveal the importance of different problem-instance families to our model.
+Instance families comprise instances that are derived from the same application domain, e.g., planning, cryptography, etc., and are a useful tool for both analyzing solver performance and portfolios.
+
+Figure~\ref{fig:annifinalfamilies} shows a scatter plot of problem-instance families regarding their occurrence within the dataset and the average occurrence in benchmarks created by active learning.
+We use the best-performing configurations for all $\delta \in \left[0, 1\right]$ and examine the selection decisions by the active-learning approach on the SAT Competition~2022 Anniversary Track benchmark set~\cite{sat2022}.
+While most families appear by the same fraction in the dataset and in active-learning-selected benchmarks, there are a few outliers that need some further discussion.
+Problem instances of the families \emph{fpga}, \emph{quasigroup-completion}, and \emph{planning} are especially helpful to our model in distinguishing solvers.
+Instances of these families are selected over-proportionally in comparison to their fraction in the dataset.
+In contrast to that, instances of the biggest family, i.e., \emph{hardware-verification}, roughly appear with the same fraction in the dataset and in active-learning-selected benchmarks.
+Finally, instances of the family \emph{cryptography} are less important in distinguishing solvers than their huge weight in the dataset suggests.
+A possible explanation is that \emph{cryptography} instances are too similar, which means that a smaller fraction of them is sufficient to determine a solver's performance on all of them.
+
+% For all configurations of our approach that sample on average less than \SI{10}{\%} of all 5355 instances, we look at the particular instances and families that are selected by our active-learning approach.
+% Thereby, the average benchmark set contains about \SI{192.34}{} instances.
+% Table~\ref{tab:familiesa} lists the ten most frequent families that are chosen.
+% For example, \emph{planning} instances account for on average \SI{19.14}{} of the \SI{192.34}{} benchmark instances.
+% In contrast to that, Table~\ref{tab:familiesb} shows the families that are sampled the most often relative to their occurrences in the underlying dataset.
+% There are, for example, only nine \emph{circuit-multiplier} instances in the complete dataset.
+% However, we sample on average \SI{1.59}{} of them.
+% %JB: there should be some (qualitative) interpretation (based on our results, is the sample interesting/special family-wise or not?) rather than just dropping a table into the paper and mainly explaining its structure
 
@@
   \caption{
-    Scatter plot showing the \emph{importance} of different instance families.
+    Scatter plot showing the \emph{importance} of different instance families to our model.
     The x-axis shows the fraction of instances within a particular family in the complete SAT Competition~2022 Anniversary Track benchmark set~\cite{sat2022}.
@@
 \section{Conclusion}
+
 In this work, we have discussed possible solutions to the \emph{New-Solver Problem}:
@@
 \subsection{Future Work}
+
 In future work, it may be of interest to compare further ranking algorithms, instance-selection approaches, and stopping criteria.

### 2022-10-10 paper/paper.tex
Revise section 6

@@
 Our approaches provide accurate ranking predictions while only needing a fraction of the runtime resources that a full evaluation on all benchmarking instances would need.
+We can determine a new solver's PAR-2 ranking with about \SI{92}{\%} accuracy, while only needing \SI{10}{\%} of the full-evaluation time.
 We use a runtime discretization technique as this enables us to transform the regression problem of directly predicting runtimes into the much simpler notion of classification.
@@
 We have evaluated several ranking algorithms, instance-selection approaches, and stopping criteria within our sequential active-learning process.
-A model-uncertainty-based selection approach in combination with runtime-label-prediction ranking and a ranking-convergence stopping criterion showed the consistently best results.
+%A model-uncertainty-based selection approach in combination with runtime-label-prediction ranking and a ranking-convergence stopping criterion showed the consistently best results in most cases.
 %JB: "consistently best" -- it's the best approach for one delta, but not sure regarding other deltas nor its sensitivity to hyperparameters
-We also took a brief look at which instance families are the most prevalent when sampling instances.
+%TF: commented out
+We also took a brief look at which problem-instance families are the most prevalent in selection decisions.
 
@@
 
-On a more general note, it would make sense to generalize this evaluation framework for arbitrary $\mathcal{NP}$-complete problems.
+On a more general note, it would make sense to generalize this evaluation framework for arbitrary $\mathcal{NP}$-complete problems as all discussed active-learning methods are agnostic to the underlying problem.
+Solely the runtime-prediction model might need some tweaking.
 %JB: is there some methodological generalization needed or can't we just apply the same framework (instantiating the instance-feature part differently) to these use cases?
+%TF: clarified
 Those problems share most of the relevant properties of SAT solving, i.e., there are established problem-instance features, a full benchmark run takes days, and creating new problem solvers traditionally requires expert knowledge to hand-select instances of interest.

### 2022-10-11 paper/paper.tex
Revise Evaluation
Rewording and minor cutting.

@@
 
-Before evaluating our approach end-to-end as described in Section~\ref{sec:exdesign}, we discuss the choice of the runtime-discretization method and the machine-learning predictor.
+Before evaluating our approach end-to-end as described in Section~\ref{sec:exdesign}, we discuss the choice of the machine-learning predictor and the runtime-discretization method.
 Thereafter, we look at the results of our active-learning approach.
-Selection decisions of the aforementioned approach also reveal the importance of different problem-instance families to our model.
-Instance families comprise instances that are derived from the same application domain, e.g., planning, cryptography, etc., and are a useful tool for both analyzing solver performance and portfolios.
+Finally, we also analyze the importance of different instance families to our framework.
+Instance families comprise instances derived from the same application domain, e.g., planning, cryptography, etc., and are a useful tool for both analyzing solver performance and portfolios.
 %JB: instance families not really motivated till now, only appear in abstract and in text regarding stratification
 %TF: Added a sentence for clarification. We could introduce more details in section 4.3.
+%JB: sounds like a good idea
 
@@
 
-%JB: if we dedicate a whole section to these results, we should probably rename this section (not "preliminary" in the typical sense); for TACAS, I guess this can be cut a bit
-%TF: cut the section down a bit, renamed to "runtime prediction"
+%JB: we should probably remove this section here and jump right into active learning;
+% - results of classifier comparison can shortly be described in Section 4 (new subsection) or 3.2 ("Solver Model", which currently mainly describes runtime discretization); in any case, classifier should be discussed earlier
+% - in current version, we already state in Section 3.2 that we use hierarchical clustering, so description of algorithm and some "performed well in preliminary experiments" could go there
 An exhaustive grid search of all active-learning approaches outlined in Section~\ref{sec:main} and different runtime-prediction models is infeasible.
-Therefore, we use the runtime-prediction model that produced the best results for all further experiments.
-We compare runtime predictors by looking at the \emph{Matthews Correlation Coefficient} (MCC) scores~\cite{gorodkin2004comparing,matthews1975comparison} of the classification task.
+Therefore, we compare predictors first and select one runtime-prediction model for all further experiments.
+We evaluate predictors with the \emph{Matthews Correlation Coefficient} (MCC) scores~\cite{gorodkin2004comparing,matthews1975comparison} of the classification task.
 MCC scores are great in dealing with class imbalances in contrast to conventional metrics like accuracy.
@@
 Note that we perform cross-validation on two levels, i.e., among the solvers and the target solver runtimes.
-We report the average performances on the test sets.
+%JB: we should state (were we introduce prediction model, i.e., either in 3.2 or a new subsection in 4) which features we use; I assume instance features + runtimes of other solvers?
+(else CV among solvers would not make sense for this experiment)
+We compare the average performances on the test sets.
 
@@
 A stacking ensemble consisting of quadratic-discriminant analysis and random forests performed best.
-It produces an MCC score of \SI{0.9527}{} for time-out prediction and \SI{0.8882}{} for runtime-label prediction.
+It produces an MCC score of \SI{0.8882}{} for predicting three-class runtime labels.
+% \SI{0.9527}{} for time-out prediction
 The results are significantly better than any other model that we have tried regarding a Wilcoxon signed-rank test with $\alpha = \SI{5}{\%}$.
@@
 
-For runtime discretization, we use hierarchical clustering with $k = 3$ using a log-single-link criterion for all further experiments.
+For runtime discretization, we use hierarchical clustering with $k = 3$ and a log-single-link criterion for all further experiments.
 Each non-time-out runtime starts in a separate interval.
@@
 
-\subsection{Experimental Results}
+\subsection{Active-Learning Results}
+
 \subsubsection{Tuning our algorithm.}
 
-Our main end-to-end experiments follow the evaluation framework that has been introduced in Section~\ref{sec:evalframe}.
-Fig.~\ref{fig:e2eallsolvers} shows the performance of the approaches listed in Section~\ref{sec:hyper} on $O_{\operatorname{rt}}$-$O_{\operatorname{acc}}$-diagrams for the smaller hyper-parameter-tuning dataset.
-Evaluating a particular configuration with Algorithm~\ref{alg:eval}, returns a point $\left(O_{\operatorname{rt}},\, O_{\operatorname{acc}}\right)$.
-The plotted lines represent the best performing configurations (convex hull) per top-level hyper-parameter choice, i.e., per ranking approach (Fig.~\ref{fig:annitraincolorranking}), per selection approach (Fig.~\ref{fig:annitraincolorselection}), and per stopping criterion (Fig.~\ref{fig:annitraincolorstopping}).
-%JB: as far as I understand it, the lines only compare some top-level hyperparameters, but for each of these hyperparemeters, take the best (= Pareto front) out of some more fine-grained settings (the hyperparameters of the hyperparameters, so to speak), so it's no really "each possible choice" in the sense of actual, individual configurations, but rather a combination of configurations
+Our end-to-end experiments follow the evaluation framework introduced in Section~\ref{sec:evalframe}.
+Fig.~\ref{fig:e2eallsolvers} shows the performance of the approaches from Section~\ref{sec:hyper} on $O_{\operatorname{rt}}$-$O_{\operatorname{acc}}$-diagrams for the smaller hyper-parameter-tuning dataset.
+Evaluating a particular configuration with Algorithm~\ref{alg:eval} returns a point $\left(O_{\operatorname{rt}},\, O_{\operatorname{acc}}\right)$.
+The plotted lines represent the best performing configurations (convex hull) per top-level hyper-parameter choice, i.e., per ranking approach (Fig.~\ref{fig:annitraincolorranking}), selection approach (Fig.~\ref{fig:annitraincolorselection}), and stopping criterion (Fig.~\ref{fig:annitraincolorstopping}).
 %Colors indicate the different hyper-parameter instantiations.
@@
 
-Regarding ranking approaches (Fig.~\ref{fig:annitraincolorranking}), using the $s_3$-induced ranking consistently outperforms the partially observed PAR-2 ranking for each possible value of $\delta$.
+Regarding ranking approaches (Fig.~\ref{fig:annitraincolorranking}), using the $s_3$-induced ranking consistently outperforms the partially observed PAR-2 ranking for each possible value of the runtime-accuracy trade-off parameter~$\delta$.
 %JB: as in previous comment: consistent outperformance not guaranteed in case the lines are not individual approaches, but Pareto fronts within the compared categories (which means that different parts of the line might be related to different hyperparameter settings)
 %TF: The lines are in any case no individual approaches as individual approaches are only points. So, any lines in this plot are a "combination" of configs. It's only a question of how to combine.
+%JB: I always made the implicit assumption that the plots are based on learning curves, showing the progress within individual AL runs, and then multiple of these curves were combined (per top-level configuration); so the curves actually draw a line between different final results of AL runs?
+should be clarified somewhere
 This is expected since selection decisions are not random.
@@
 Regarding the selection approaches (Fig.~\ref{fig:annitraincolorselection}), the model-based uncertainty strategy performs best in most cases.
+%JB: maybe we can drop "model-based" and use "uncertainty sampling" instead?
+I adapted this in Section 3, but section 4 + 5 + the plots would have to be adapted as well (else, we should revert my changes in Section 3)
 However, the model-based information-gain sampling is beneficial if runtime is strongly favored (very small $\delta$; runtime fraction less than \SI{5}{\%}).
-Random sampling is outperformed by the model-based uncertainty strategy in all cases showing the benefit of active learning.
-By actively choosing the next experiment to run based on a model, we are more accurate in predicting the PAR-2 ranking of the solvers on the underlying dataset.
-%JB: we should also describe performance of random sampling, which is visible in plot as well
-%TF: added a few sentences
+Random sampling performs quite well but is outperformed by the model-based uncertainty strategy in all cases, showing the benefit of actively selecting instances based on a prediction model.
 
@@
 If accuracy is strongly favored (very high $\delta$), the Wilcoxon stopping criterion performs better.
+%JB: plot says "Wilcoxon" without stopping; only minor point, but we should make naming of the components of our framework consistent
 
@@
     }
-    \caption{
-    	$O_{\operatorname{rt}}$-$O_{\operatorname{acc}}$-diagram showing Pareto fronts for configurations with the given ranking approaches.
-    	The lines show the best performances of approaches using runtime-label-prediction ranking and partially observed PAR-2 ranking respectively.
-    }
+    \caption{Ranking approaches}
     \label{fig:annitraincolorranking}
@@
     }
-    \caption{
-    	$O_{\operatorname{rt}}$-$O_{\operatorname{acc}}$-diagram showing Pareto fronts for configurations with the given instance-selection approaches.
-    	The lines show the best performances of approaches using uncertainty-based, information-gain-based, and random instance selection respectively.
-    }
+    \caption{Selection approaches}
     \label{fig:annitraincolorselection}
@@
     }
-    \caption{
-    	$O_{\operatorname{rt}}$-$O_{\operatorname{acc}}$-diagram showing Pareto fronts for configurations with the given stopping criteria.
-    	The lines show the best performances of approaches using the ranking-convergence criterion, Wilcoxon criterion, and a fixed subset size respectively.
-    }
+    \caption{Stopping criteria}
     \label{fig:annitraincolorstopping}
@@
   \caption{
-  	$O_{\operatorname{rt}}$-$O_{\operatorname{acc}}$-diagrams showing the performance of different hyper-parameter instantiations.
-  	The x-axis shows the fraction of runtime in comparison to the time needed to evaluate a solver on all instances.
-  	The y-axis shows the accuracy of the predicted rank in comparison to the true PAR-2 rank.
-  	Each line entails the front of Pareto-optimal configurations with the respective instantiation.
-  }
-  %JB: not sure if the color palette in the plots is grayscale-/color-blind-friendly; consider changing the palette (see https://colorbrewer2.org/#type=qualitative&scheme=Set3&n=3 ) and using different line types for each color
-  %TF: done
+  	$O_{\operatorname{rt}}$-$O_{\operatorname{acc}}$-diagrams comparing different hyper-parameter instantiations of our active-learning framework.
+  	The x-axis shows the ratio of total solver runtime on the sampled instances relative to all instances.
+  	The y-axis shows the ranking accuracy (cf. Section~\ref{sec:goal}).
+  	%JB: we should use "ranking accuracy" in plots, since we explicitly defined this metric earlier
+  	Each line entails the front of Pareto-optimal configurations for the respective hyper-parameter instantiation.
+  }
   \label{fig:e2eallsolvers}
@@
   \caption{
-    Scatter plot showing different instantiations of our active-learning strategy for different $\delta$ on our smaller hyper-parameter-tuning dataset.
-    The x-axis shows the fraction of runtime $O_{\operatorname{rt}}$ that a particular instantiation needs.
-    The y-axis shows the fraction of instances that a particular instantiation selects to arrive at its conclusion.
+    Scatter plot comparing different instantiations of $\delta$ for our active-learning strategy on the hyper-parameter-tuning dataset.
+    The x-axis shows the fraction of runtime $O_{\operatorname{rt}}$ of the sample, while the y-axis shows the fraction of instances sampled.
     The color indicates the weighting between different optimization goals $\delta \in \left[0, 1\right]$.
-    A small $\delta$ refers to favoring runtime over accuracy and a big $\delta$ refers to favoring accuracy over runtime.
+    The larger $\delta$, the more we favor accuracy over runtime.
   }
@@
 
-Figure~\ref{fig:annitrainoptgoal} also shows an interesting difference between different optimization goals.
-It shows how different configurations select instances with different difficulties.
-If it is, on the one hand, desired to get a \emph{rough} estimate of a solver's performance fast (low $\delta$), corresponding approaches favor selecting many \emph{easy} problem instances (fraction of instances bigger than the fraction of runtime within $\tilde{\mathcal{I}}$).
+Figure~\ref{fig:annitrainoptgoal} shows an interesting consequence of weighting our optimization goals:
+If we, on the one hand, desire to get a \emph{rough} estimate of a solver's performance fast (low $\delta$), approaches favor selecting many \emph{easy} instances.
+In particular, the fraction of sampled instances is bigger than the fraction of runtime.
 By having many observations, it is easier to build a model.
-If it is, on the other hand, desired to get a \emph{good} estimate of a solver's performance in a moderate amount of time (high $\delta$), corresponding approaches favor selecting fewer \emph{difficult} problem instances (fraction of instances smaller than the fraction of runtime within $\tilde{\mathcal{I}}$).
+If we, on the other hand, desire to get a \emph{good} estimate of a solver's performance in a moderate amount of time (high $\delta$), approaches favor selecting fewer \emph{difficult} instances.
+In particular, the fraction of instances is smaller than the fraction of runtime.
 
@@
 
-Having selected the most promising hyper-parameters, we run our end-to-end active-learning experiments on the complete Anniversary Track dataset (5355 instances).
+Having selected the most promising hyper-parame\-ters, we run our end-to-end active-learning experiments on the complete Anniversary Track dataset (5355 instances).
 The best-performing approach for $\delta = 0.85$ uses runtime-label prediction ranking, model-based uncertainty sampling, and a ranking-convergence stopping criterion.
-It can predict a new solver's PAR-2 ranking with about \SI{92.33}{\%} accuracy ($O_{\operatorname{acc}}$), while only needing \SI{10.35}{\%} of the runtime that would be needed to evaluate the solver on the full dataset ($O_{\operatorname{rt}}$).
-This trade-off provides a better way of prototyping and evaluating new SAT solvers rather than evaluating them on a manually picked subset based on domain knowledge.
-The sampled instances account for about \SI{5.24}{\%} of instances indicating the preference for selecting more difficult instances as already discussed in the previous section.
-The average Spearman correlation between the PAR-2 ranking and the predicted ranking scores is about \SI{0.9572}{} indicating that the differences in PAR-2 scores are quite similar to $s_3$ scores.
+%JB: we need to motivate choosing this \delta
+It can predict a new solver's PAR-2 ranking with about \SI{92.33}{\%} accuracy ($O_{\operatorname{acc}}$), while only needing \SI{10.35}{\%} of the time for running the solver on the full dataset ($O_{\operatorname{rt}}$).
+%JB: could we (with reasonable effort) integrate the learning curve of this particular approach as a plot?
+This trade-off provides a better way of prototyping and evaluating new SAT solvers rather than evaluating them on manually picked instance based on domain knowledge.
+The sampled instances account for about \SI{5.24}{\%} of instances, indicating a preference for selecting difficult instances.
+The average Spearman correlation between the PAR-2 ranking and the predicted ranking is about \SI{0.9572}{}, indicating a high similarity of these two rankings.
 Because of that, the runtime-prediction model helps determine on what problem instances solvers perform similarly or are better/worse.
-%JB: many numbers, but some qualitative (and positively sounding, i.e., saying that our approach indeed is useful) interpretation would be nice
-%TF: revised this section
+%JB: conclusion not straightforward from prior text; even if overall estimated score (which is an average over many instances) is good, estimate on individual instances might be quite erroneous
 
@@
 
-Selection decisions of our approach also reveal the importance of different problem-instance families to our model.
-Instance families comprise instances that are derived from the same application domain, e.g., planning, cryptography, etc., and are a useful tool for both analyzing solver performance and portfolios.
-
-Figure~\ref{fig:annifinalfamilies} shows a scatter plot of problem-instance families regarding their occurrence within the dataset and the average occurrence in benchmarks created by active learning.
-We use the best-performing configurations for all $\delta \in \left[0, 1\right]$ and examine the selection decisions by the active-learning approach on the SAT Competition~2022 Anniversary Track benchmark set~\cite{sat2022}.
-While most families appear by the same fraction in the dataset and in active-learning-selected benchmarks, there are a few outliers that need some further discussion.
-Problem instances of the families \emph{fpga}, \emph{quasigroup-completion}, and \emph{planning} are especially helpful to our model in distinguishing solvers.
-Instances of these families are selected over-proportionally in comparison to their fraction in the dataset.
-In contrast to that, instances of the biggest family, i.e., \emph{hardware-verification}, roughly appear with the same fraction in the dataset and in active-learning-selected benchmarks.
+Selection decisions of our approach also reveal the importance of different instance families to our framework.
+Figure~\ref{fig:annifinalfamilies} shows the occurrence of instance families within the dataset and the benchmarks created by active learning.
+We use the best-performing configurations for all $\delta \in \left[0, 1\right]$ and examine the selection decisions by the active-learning approach on the SAT Competition~2022 Anniversary Track dataset~\cite{sat2022}.
+While most families appear with the same fraction in the dataset and the sampled benchmarks, there are a few outliers that need further discussion.
+Problem instances of the families \emph{fpga}, \emph{quasigroup-completion}, and \emph{planning} are especially helpful to our framework in distinguishing solvers.
+Instances of these families are selected over-proportionally in comparison to the full dataset.
+In contrast to that, instances of the biggest family, i.e., \emph{hardware-verification}, roughly appear with the same fraction in the dataset and the sampled benchmarks.
 Finally, instances of the family \emph{cryptography} are less important in distinguishing solvers than their huge weight in the dataset suggests.
-A possible explanation is that \emph{cryptography} instances are too similar, which means that a smaller fraction of them is sufficient to determine a solver's performance on all of them.
+A possible explanation is that these instances are very similar, such that a small fraction of them is sufficient to estimate a solver's performance on all of them.
 
@@
 	\caption{
-    Scatter plot showing the \emph{importance} of different instance families to our model.
-    The x-axis shows the fraction of instances within a particular family in the complete SAT Competition~2022 Anniversary Track benchmark set~\cite{sat2022}.
-    The y-axis shows the average fraction of instances within a particular family that is selected by our active-learning strategies.
-    The dashed line represents families that are represented with the same fraction in, both, our dataset and the incremental benchmark sets $\tilde{\mathcal{I}}$ that are chosen by our approach.
+		Scatter plot showing the \emph{importance} of different instance families to our framework.
+		The x-axis shows the frequency of instance families in the full Anniversary Track dataset.
+		The y-axis shows the average frequency of instance families in the samples selected by active learning.
+		The dashed line represents families that are occur with the same frequency in both, dataset and samples.
 	}
@@
 %TF: commented out
-We also took a brief look at which problem-instance families are the most prevalent in selection decisions.
+We also took a brief look at which instance families are the most prevalent in selection decisions.
 
@@
 Solely the runtime-prediction model might need some tweaking.
-%JB: is there some methodological generalization needed or can't we just apply the same framework (instantiating the instance-feature part differently) to these use cases?
-%TF: clarified
 Those problems share most of the relevant properties of SAT solving, i.e., there are established problem-instance features, a full benchmark run takes days, and creating new problem solvers traditionally requires expert knowledge to hand-select instances of interest.

### 2022-10-11 paper/paper.tex
Harmonize section labels
Use two-level names consistently. No visible changes in compiled PDF.

@@
 In this section, we present the details of our approach.
-Section~\ref{sec:main-framework} describes the general framework.
-Section~\ref{sec:main-model} specifies the details of the underlying prediction model and describes how we may derive a solver ranking from that model.
-We discuss possible criteria for sampling instances in Section~\ref{sec:main-sampling}.
-Section~\ref{sec:main-stopping} concludes with possible stopping conditions.
+Section~\ref{sec:main:framework} describes the general framework.
+Section~\ref{sec:main:model} specifies the details of the underlying prediction model and describes how we may derive a solver ranking from that model.
+We discuss possible criteria for sampling instances in Section~\ref{sec:main:sampling}.
+Section~\ref{sec:main:stopping} concludes with possible stopping conditions.
 
 \subsection{Incremental Benchmarking Framework}
-\label{sec:main-framework}
+\label{sec:main:framework}
 
@@
 
-  \While(\tcp*[f]{cf. Section~\ref{sec:main-stopping}}){$\operatorname{not} \, \operatorname{stop}\!\left(\mathcal{M}\right)$}{
-    $s \leftarrow \operatorname{selectNextInstance}\!\left(\mathcal{M}\right)$ \tcp*{cf. Section~\ref{sec:main-sampling}}
+  \While(\tcp*[f]{cf. Section~\ref{sec:main:stopping}}){$\operatorname{not} \, \operatorname{stop}\!\left(\mathcal{M}\right)$}{
+    $s \leftarrow \operatorname{selectNextInstance}\!\left(\mathcal{M}\right)$ \tcp*{cf. Section~\ref{sec:main:sampling}}
 
@@
 
-    $\operatorname{updateModel}\!\left(\mathcal{M},\, s,\, t\right)$ \tcp*{cf. Section~\ref{sec:main-model}}
-  }
-
-  \Return $\operatorname{predictScore}\!\left(\mathcal{M}\right)$ \tcp*{cf. Section~\ref{sec:main-model}}
+    $\operatorname{updateModel}\!\left(\mathcal{M},\, s,\, t\right)$ \tcp*{cf. Section~\ref{sec:main:model}}
+  }
+
+  \Return $\operatorname{predictScore}\!\left(\mathcal{M}\right)$ \tcp*{cf. Section~\ref{sec:main:model}}
 
@@
 \subsection{Solver Model}
-\label{sec:main-model}
+\label{sec:main:model}
 
@@
 \subsection{Instance Sampling}
-\label{sec:main-sampling}
+\label{sec:main:sampling}
 
@@
 \subsection{Stopping Criteria}
-\label{sec:main-stopping}
+\label{sec:main:stopping}
 
@@
 \subsection{Optimization Goal}
-\label{sec:goal}
+\label{sec:exdesign:goal}
 
 As already stated in the introductory section, this work addresses the \emph{New-Solver Problem} (cf.~Definition~\ref{def:new-solver-problem}).
-We create a prediction model $\mathcal{M}$ as described in Section~\ref{sec:main-model} that provides us with a scoring function $s_k$.
+We create a prediction model $\mathcal{M}$ as described in Section~\ref{sec:main:model} that provides us with a scoring function $s_k$.
 % MI: Let this be the evaluation result:
@@
 \subsection{Evaluation Framework}
-\label{sec:evalframe}
+\label{sec:exdesign:evalframe}
 
@@
 \subsection{Hyper-parameters}
-\label{sec:hyper}
+\label{sec:exdesign:hyper}
 
@@
 Note that we are not considering all previously listed approaches since a grid search of all combinations would be infeasible.
-Rather, we filter approaches based on preliminary experimental results (cf. Section~\ref{sec:evalprel}) and do the main end-to-end experiments only with a subset.
+Rather, we filter approaches based on preliminary experimental results (cf. Section~\ref{sec:eval:prediction}) and do the main end-to-end experiments only with a subset.
 The end-to-end experiment configurations are given below.
@@
 
-Regarding \emph{ranking} (cf. Section~\ref{sec:main-model}), we experiment with the following approaches, including our used hyper-parameter values:
+Regarding \emph{ranking} (cf. Section~\ref{sec:main:model}), we experiment with the following approaches, including our used hyper-parameter values:
 
@@
     %TF: Not necessarily
-    As discussed in Section~\ref{sec:main-model}, ranking decisions are all correct if the difference in scores between two solvers exceeds a certain value.
+    As discussed in Section~\ref{sec:main:model}, ranking decisions are all correct if the difference in scores between two solvers exceeds a certain value.
   \end{itemize}
@@
 
-For \emph{selection} (cf. Section~\ref{sec:main-sampling}), we experiment with the following methods, including our used hyper-parameter values:
+For \emph{selection} (cf. Section~\ref{sec:main:sampling}), we experiment with the following methods, including our used hyper-parameter values:
 
@@
 
-For \emph{stopping} decisions (cf. Section~\ref{sec:main-stopping}), we experiment with the following criteria, including our used hyper-parameter values:
+For \emph{stopping} decisions (cf. Section~\ref{sec:main:stopping}), we experiment with the following criteria, including our used hyper-parameter values:
 
@@
 
-As runtime-label prediction model, we only make use of one fixed model since an exhaustive grid search would be infeasible (cf.~Section~\ref{sec:evalprel}).
+As runtime-label prediction model, we only make use of one fixed model since an exhaustive grid search would be infeasible (cf.~Section~\ref{sec:eval:prediction}).
 Our model of choice is an ensemble, stacking a quadratic-discriminant analysis\footnote{With a regularization parameter of zero and a singular value threshold of zero} onto a random forest model\footnote{With \emph{entropy} splitting criterion and \emph{balanced} class weights} using a decision tree\footnote{With \emph{Gini} impurity splitting criterion; choosing the best split for each branching decision and maximum tree depth of 5 levels} to weight its members.
@@
 \subsection{Runtime Prediction}
-\label{sec:evalprel}
+\label{sec:eval:prediction}
 
@@
 We then gradually merge intervals whose single-link logarithmic distance is the smallest until the desired number of partitions is reached.
-As already mentioned in Section~\ref{sec:main-model}, this approach has good properties regarding the predictiveness of the PAR-2 score ranking and its discriminatory power.
+As already mentioned in Section~\ref{sec:main:model}, this approach has good properties regarding the predictiveness of the PAR-2 score ranking and its discriminatory power.
 Also, experiments showed that our stacking ensemble performed particularly well for this kind of discretization method.
@@
 
-Our end-to-end experiments follow the evaluation framework introduced in Section~\ref{sec:evalframe}.
-Fig.~\ref{fig:e2eallsolvers} shows the performance of the approaches from Section~\ref{sec:hyper} on $O_{\operatorname{rt}}$-$O_{\operatorname{acc}}$-diagrams for the smaller hyper-parameter-tuning dataset.
+Our end-to-end experiments follow the evaluation framework introduced in Section~\ref{sec:exdesign:evalframe}.
+Fig.~\ref{fig:e2eallsolvers} shows the performance of the approaches from Section~\ref{sec:exdesign:hyper} on $O_{\operatorname{rt}}$-$O_{\operatorname{acc}}$-diagrams for the smaller hyper-parameter-tuning dataset.
 Evaluating a particular configuration with Algorithm~\ref{alg:eval} returns a point $\left(O_{\operatorname{rt}},\, O_{\operatorname{acc}}\right)$.
@@
   	The x-axis shows the ratio of total solver runtime on the sampled instances relative to all instances.
-  	The y-axis shows the ranking accuracy (cf. Section~\ref{sec:goal}).
+  	The y-axis shows the ranking accuracy (cf. Section~\ref{sec:exdesign:goal}).
   	%JB: we should use "ranking accuracy" in plots, since we explicitly defined this metric earlier

### 2022-10-11 paper/paper.tex
Revise Conclusion
Mainly language, text made a bit shorter

@@
 Given a new solver, we want to find its ranking amidst its competitors.
-Our approaches provide accurate ranking predictions while only needing a fraction of the runtime resources that a full evaluation on all benchmarking instances would need.
-We can determine a new solver's PAR-2 ranking with about \SI{92}{\%} accuracy, while only needing \SI{10}{\%} of the full-evaluation time.
-We use a runtime discretization technique as this enables us to transform the regression problem of directly predicting runtimes into the much simpler notion of classification.
-We have shown that, albeit being more simple, the classification of discrete runtime labels produces good results.
+Our approach provides accurate ranking predictions while needing significantly less runtime that a full evaluation on a given benchmark set.
+In particular, we can determine a new solver's PAR-2 ranking with about \SI{92}{\%} accuracy, while only needing \SI{10}{\%} of the full-evaluation time.
+We have found that our framework produces good results working with discrete runtime labels rather than regressing actual runtimes, albeit the former being more simple.
 We have evaluated several ranking algorithms, instance-selection approaches, and stopping criteria within our sequential active-learning process.
@@
 
-In future work, it may be of interest to compare further ranking algorithms, instance-selection approaches, and stopping criteria.
-Furthermore, it is possible to formulate the runtime discretization as an optimization problem.
-Given a dataset, optimization could select the discretization technique with the best discriminatory power rather than selecting one approach upfront by hand.
-
-A major shortcoming is the lack of parallelizability.
-Our current approach selects instances one at a time.
-Running benchmarks on a computing cluster with $n$ cores benefits from having a batch of $n$ instances at a time.
-This is, however, not trivial since, on the one hand, a higher $n$ leads to less \emph{active learning} (because of bigger batch sizes) and, on the other hand, it is not clear how to synchronize the model update and instance selection without a barrier lock, which wastes a lot of runtime resources.
-
-On a more general note, it would make sense to generalize this evaluation framework for arbitrary $\mathcal{NP}$-complete problems as all discussed active-learning methods are agnostic to the underlying problem.
+Future work may compare further ranking algorithms, instance-selection approaches, and stopping criteria.
+Furthermore, it is possible to formulate runtime discretization as an optimization problem, directly selecting the most discriminatory discretization technique rather than doing an empirical comparison.
+
+A major shortcoming of our current approach is the lack of parallelization, selecting instances one at a time.
+Benchmarking on a computing cluster with $n$ cores benefits from having batches of $n$ instances.
+However, bigger batch sizes $n$ impede \emph{active learning}.
+Also, it is unclear how to synchronize model update and instance selection without wasting too much runtime.
+
+On a more general note, one can apply our evaluation framework to arbitrary $\mathcal{NP}$-complete problems as all discussed active-learning methods are problem-agnostic.
 Solely the runtime-prediction model might need some tweaking.
-Those problems share most of the relevant properties of SAT solving, i.e., there are established problem-instance features, a full benchmark run takes days, and creating new problem solvers traditionally requires expert knowledge to hand-select instances of interest.
+%JB: other components of framework might need tweaking as well, or this there something special about prediction model?
+Those problems share most of the relevant properties of SAT solving, i.e., there are established problem-instance features, a full benchmark run takes days, and solver development traditionally requires expert knowledge to hand-select instances.
 

### 2022-10-11 paper/paper.tex
algo1

@@
 
-In this section, we present the details of our approach.
-Section~\ref{sec:main:framework} describes the general framework.
+The outline of our benchmarking framework is depicted in Algorithm~\ref{algALBenchmark}. 
+Given a set of known solvers $\mathcal{A}$, instances $\mathcal{I}$ and runtimes $r$, we first create a prediction model~$\mathcal{M}$ for the new solver $\hat a$ (Line~1). 
+The prediction model is used to repeatedly select an instance (Line~4) for benchmarking $\hat a$ (Line~5). 
+The acquired result is subsequently used to update the prediction model~$\mathcal{M}$ (Line~6). 
+When the stopping criterion is met (Line~3), we quit the benchmarking loop and predict the final score of $\hat a$ (Line~8). 
+Algorithm~\ref{algALBenchmark} returns the predicted score of $\hat a$ as well as the acquired runtime measurements (Line~9).
+
 Section~\ref{sec:main:model} specifies the details of the underlying prediction model and describes how we may derive a solver ranking from that model.
@@
 
-\subsection{Incremental Benchmarking Framework}
-\label{sec:main:framework}
-
-Our framework to score a new solver~$\hat{a}$ follows a three-step procedure, outlined in Algorithm~\ref{algALBenchmark}:
-While the stopping condition is not met (line~1), we select \emph{one} SAT instance from the pool of instances with the help of a prediction model~$\mathcal{M}$ (line~2) and evaluate the new solver~$\hat{a}$ on this instance (line~3).
-We then use the acquired result to update the prediction model~$\mathcal{M}$ (line~4).
-When the stopping criterion is met, we return the predicted score of the new solver~$\hat{a}$ (line~5).
-
+% MI: This belongs to discussion at the end since this also has substantial disadvantages:
 Since the potential runtime of experiments is by magnitudes larger than the model's update time, we only consider incrementing our benchmark by one instance at a time rather than using batches, which is also proposed in current active-learning advances~\cite{SinhaED19,2019gaal}.
@@
   %\KwIn{Instances $\mathcal{I}$, Solvers $\mathcal{A}$, Runtimes $r\!: \mathcal{A} \times \mathcal{I} \rightarrow \left[0, \tau\right]$}
-  \KwIn{Model $\mathcal{M}$,  Solver $\hat{a}$}
-  \KwOut{Score of $\hat{a}$}
+  \KwIn{Solvers $\mathcal{A}$, Instances $\mathcal{I}$, Runtimes $r : \mathcal{A} \times \mathcal{I} \rightarrow [0, \tau]$, Solver $\hat{a}$}
+  \KwOut{Predicted Score of $\hat{a}$, Measured Runtimes $\mathcal{R}$}
 
@@
 
-  %$\mathcal{M} \leftarrow \textrm{init model with $\mathcal{A}$, $\mathcal{I}$, $r$, and $\hat{a}$}$
-  %$\tilde{\mathcal{I}} \leftarrow \emptyset$
-
-  \While(\tcp*[f]{cf. Section~\ref{sec:main:stopping}}){$\operatorname{not} \, \operatorname{stop}\!\left(\mathcal{M}\right)$}{
-    $s \leftarrow \operatorname{selectNextInstance}\!\left(\mathcal{M}\right)$ \tcp*{cf. Section~\ref{sec:main:sampling}}
-
-    $t \leftarrow \operatorname{runExperiment}\!\left(\hat{a},\,  s\right)$  \tcp*{Runs $\hat{a}$ on $s$ with time-out $\tau$}
-
-    %$\tilde{\mathcal{I}} \leftarrow \tilde{\mathcal{I}} \cup \left\lbrace s \right\rbrace$
-
-    $\operatorname{updateModel}\!\left(\mathcal{M},\, s,\, t\right)$ \tcp*{cf. Section~\ref{sec:main:model}}
-  }
-
-  \Return $\operatorname{predictScore}\!\left(\mathcal{M}\right)$ \tcp*{cf. Section~\ref{sec:main:model}}
+  $\mathcal{M} \leftarrow \operatorname{initModel}(\mathcal{A}, \mathcal{I}, r, \hat{a})$ \;
+
+  \BlankLine
+  $\mathcal{R} \leftarrow \emptyset$ \;
+  \While(\tcp*[f]{cf. Section~\ref{sec:main:stopping}}){$\operatorname{not} \operatorname{stop}(\mathcal{M})$}{
+    $e \leftarrow \operatorname{selectNextInstance}\!\left(\mathcal{M}\right)$ \tcp*{cf. Section~\ref{sec:main:sampling}}
+
+    $t \leftarrow \operatorname{runExperiment}\!\left(\hat{a},\,  e\right)$  \tcp*{Runs $\hat{a}$ on $e$ with time-out $\tau$}
+
+    $\operatorname{updateModel}\bigr(\mathcal{M}, (e, t)\bigl)$ \tcp*{cf. Section~\ref{sec:main:model}}
+
+    \BlankLine
+    $\mathcal{R} \leftarrow \mathcal{R} \cup \bigl\{ (e, t) \bigr\}$
+  }
+  $s \leftarrow \operatorname{predictScore}(\mathcal{M})$ \tcp*{cf. Section~\ref{sec:main:model}}
+
+  \BlankLine
+  \Return $(s, \mathcal{R})$
 
@@
 
-\subsubsection{Wilcoxon Stopping Criterion}
+\subsubsection{Wilcoxon Stopping Criterion.}
 
@@
 
-\subsubsection{Model Convergence Stopping Criterion}
+\subsubsection{Model Convergence Stopping Criterion.}
 

### 2022-10-11 paper/paper.tex
algo2
table1

@@
   ~\\[1em]
-  \caption{Support for desired criteria from Section~\ref{sec:intro}.
-	The table compares a static benchmark-selection approach~\cite{HoosKSS13}, an algorithm configuration system~\cite{HutterHL11}, an existing active-learning approach~\cite{MatriconAFSH21}, as well as our approach.
+  \caption{Comparison of features of our benchmark selection approach, the static benchmark-selection approach by Hoos~et~al.~\cite{HoosKSS13}, the algorithm configuration system SMAC~\cite{HutterHL11}, and the active-learning approaches by Matricon et al.~\cite{MatriconAFSH21}.
   }
@@
   }
-  $s \leftarrow \operatorname{predictScore}(\mathcal{M})$ \tcp*{cf. Section~\ref{sec:main:model}}
+  $s_{\hat a} \leftarrow \operatorname{predictScore}(\mathcal{M})$ \tcp*{cf. Section~\ref{sec:main:model}}
 
   \BlankLine
-  \Return $(s, \mathcal{R})$
+  \Return $(s_{\hat a}, \mathcal{R})$
 
@@
   \KwIn{Solvers $\mathcal{A}$, Instaces $\mathcal{I}$, Runtimes $r : \mathcal{A} \times \mathcal{I} \rightarrow [0, \tau]$}
-  \KwOut{Average ranking accuracy $\bar{O}_{\operatorname{acc}}$, Average fraction of runtime $\bar{O}_{\operatorname{rt}}$}
+  \KwOut{Average Rank Accuracy $\bar{O}_{\operatorname{acc}}$, Average Fraction of Runtime $\bar{O}_{\operatorname{rt}}$}
   \BlankLine
 
-  $res \leftarrow \emptyset$
+  $O \leftarrow \emptyset$
+  \BlankLine
 
   \For{$\hat{a} \in \mathcal{A}$}{
-    $\mathcal{A}' \leftarrow \mathcal{S} \setminus \left\lbrace \hat{a} \right\rbrace$
-
-    $\left( s, \tilde{\mathcal{I}} \right) \leftarrow \operatorname{runALAlgorithm}\!\left(\mathcal{I}, \mathcal{A}', r, \hat{a}\right)$ \tcp*{Refer to Algorithm~\ref{algALBenchmark}}
-
-    $O_{\operatorname{acc}} \leftarrow \frac{1}{|\mathcal{A}|} \sum\limits_{a \in \mathcal{A}} \left[\operatorname{sign}(s(a) - s(\hat{a})) = \operatorname{sign}(par2(a) - par2(\hat{a})) \right]$
-
-    \tcp*{Iverson-bracket notation}
-    %JB: maybe we can simplify the expresson, since readers might be unfamiliar with that notation (I had to look it up)
-
-	$r' \leftarrow \sum\limits_{i \in \tilde{\mathcal{I}}} r(i, \hat{a})$\;
-	$r \leftarrow \sum\limits_{i \in \mathcal{I}} r(i, \hat{a})$\;
-    $O_{\operatorname{rt}} \leftarrow \frac{r'}{r}$
-
-    $res \leftarrow res \cup \left\lbrace \left( O_{\operatorname{acc}},\, O_{\operatorname{rt}} \right) \right\rbrace$
-  }
-
-  $\left( \bar{O}_{\operatorname{acc}},\, \bar{O}_{\operatorname{rt}} \right) \leftarrow \operatorname{mean}\!\left( res \right)$ \;
-  \Return $\left( \bar{O}_{\operatorname{acc}},\, \bar{O}_{\operatorname{rt}} \right)$
+    $\mathcal{A}' \leftarrow \mathcal{A} \setminus \left\lbrace \hat{a} \right\rbrace$ \;
+    $(s_{\hat a}, \mathcal{R}) \leftarrow \operatorname{runALAlgorithm}(\mathcal{A}', \mathcal{I}, r, \hat{a})$ \tcp*{Refer to Algorithm~\ref{algALBenchmark}}
+
+	\BlankLine
+	\tcp{Determine Rank Accuracy}
+    $O_{\operatorname{acc}} \leftarrow 0$ \;    
+    \For{$a \in \mathcal{A}$}{
+    	\If{$\bigl(s_k(a) - s_{\hat a}\bigr) \cdot \bigl(\operatorname{par_2}(a) - \operatorname{par_2}(\hat a)\bigr) > 0$}{
+   		    $O_{\operatorname{acc}} \leftarrow O_{\operatorname{acc}} + \frac{1}{|\mathcal{A}|}$ \;
+    	}
+    }
+
+	\BlankLine
+	\tcp{Determine Runtime Fraction}
+	$r \leftarrow \sum\limits_{e \in \mathcal{I}} r(\hat a, e)$\;
+	$O_{\operatorname{rt}} \leftarrow 0$\;
+    \For{$e \in \mathcal{I}$}{
+		\If{$\exists t, (e,t) \in \mathcal{R}$}{
+			$O_{\operatorname{rt}} \leftarrow O_{\operatorname{rt}} + \frac{t}{r}$\;
+		}
+	}
+
+	\BlankLine
+    $O \leftarrow O \cup \bigl\{ ( O_{\operatorname{acc}},\, O_{\operatorname{rt}} ) \bigr\}$
+  }
+
+  \BlankLine
+  $\bigl( \bar{O}_{\operatorname{acc}}, \bar{O}_{\operatorname{rt}} \bigr) \leftarrow \operatorname{average}(O)$ \;
+
+  \BlankLine
+  \Return $\bigl( \bar{O}_{\operatorname{acc}}, \bar{O}_{\operatorname{rt}} \bigr)$
 \end{algorithm}

### 2022-10-11 paper/paper.tex
remove backup of sentence
Revise different parts of text
- harmonize naming of components of AL framework (also in plots)
- revise dataset description, particularly instance families and input to
prediction
- resolve various other comments discussed in today's meeting

@@
 Balint et al. provide an overview of benchmark-selection criteria in different solver competitions~\cite{balint2015overview}.
-%JB: not sure if we should provide some details regarding SAT Competition 2022 \cite{sat2022}, previous competitions (if different procedure used; Related-Work section of Matricon et al.
-(2021) states that) or other competitions (e.g., those in https://www.floc2022.org/floc-olympic-games)
-%TF: @Markus maybe you are better at briefly discussing the  SAT Competition 2022 and previous competitions
+Froleyks et al. describe benchmark-selection for the SAT~Competition~\cite{FroleyksHIJS21}.
 Manthey and Möhle find that competition benchmarks might contain redundant instances, and propose a feature-based approach to remove redundancy~\cite{manthey2016better}.
@@
     Incremental/Non-Random & \xmark & \xmark & \cmark & \cmark \\
-    % \tablefootnote{While using a model-based selection strategy to select the next configuration to evaluate, the instance sampling is done randomly.}
+    % \tablefootnote{While using a model-based selection strategy to select the next configuration to evaluate, the instance selection is done randomly.}
     Scoring & \cmark & \xmark & \xmark & \cmark \\
@@
 First, an algorithm configurator cannot be used to \emph{rank} a new solver since it only looks at promising solvers/configurations rather than the overall average performance.
-Second, while using a model-based selection strategy to sample configurations, instance sampling is done \emph{randomly}, i.e., without building a model over instances.
+Second, while using a model-based selection strategy to sample configurations, instance selection is done \emph{randomly}, i.e., without building a model over instances.
 And third, only few configurations within their approach are \emph{scored} since algorithm configuration seeks to find solemnly the best performing configuration.
@@
 Most of the proposed strategies in~\cite{MatriconAFSH21} are instance-based, i.e., no model of the solver runtimes is built.
-Regarding our desired criteria of benchmark selection (Table~\ref{tab:requirements}), the aforementioned approach \emph{ranks} solvers, optimizes for \emph{runtime}, is \emph{feature-based}, and uses \emph{incremental non-random} sampling.
+Regarding our desired criteria of benchmark selection (Table~\ref{tab:requirements}), the aforementioned approach \emph{ranks} solvers, optimizes for \emph{runtime}, is \emph{feature-based}, and uses \emph{incremental non-random} instance selection.
 However, the approach only compares solvers binarily rather than providing a \emph{scoring}.
@@
 Section~\ref{sec:main:model} specifies the details of the underlying prediction model and describes how we may derive a solver ranking from that model.
-We discuss possible criteria for sampling instances in Section~\ref{sec:main:sampling}.
+We discuss possible criteria for selecting instances in Section~\ref{sec:main:selection}.
 Section~\ref{sec:main:stopping} concludes with possible stopping conditions.
@@
   \While(\tcp*[f]{cf. Section~\ref{sec:main:stopping}}){$\operatorname{not} \operatorname{stop}(\mathcal{M})$}{
-    $e \leftarrow \operatorname{selectNextInstance}\!\left(\mathcal{M}\right)$ \tcp*{cf. Section~\ref{sec:main:sampling}}
+    $e \leftarrow \operatorname{selectNextInstance}\!\left(\mathcal{M}\right)$ \tcp*{cf. Section~\ref{sec:main:selection}}
 
@@
 Model $\mathcal{M}$ provides a prediction function $f : \mathcal{\hat A} \times \mathcal{I} \rightarrow \mathbb{R}$ that powers the decisions within our framework.
-
-To be applicable to a new solver $\hat{a}$, whose runtimes are unknown, we extend \emph{score} by replacing $\gamma(a, e)$ with the model's predictions $f : \mathcal{\hat A} \times \mathcal{I} \rightarrow \left\lbrace 1, \dots, k \right\rbrace$.
-%JB: "score" and "gamma" aren't defined yet, so there is nothing to extend; score isn't even defined here, only "s", while later sections still use "score"
-%JB: also, each model only predicts for one solver (there is no model covering multiple solvers), so I'm not sure whether introducing the solver as parameter makes sense
 
@@
 
-\subsection{Instance Sampling}
-\label{sec:main:sampling}
+\subsection{Instance Selection}
+\label{sec:main:selection}
 
@@
 In this section, we introduce our sampling strategies, which make use of the model's label-prediction function $f$ and are inspired by existing work within the realms of active learning~\cite{settles2009active}.
-We implement a model-uncertainty and an information-gain sampling strategy.
+We implement an uncertainty and an information-gain sampling strategy.
 These methods require the model's predictions to also include probabilities for the $k$ possible runtime labels.
@@
 
-\subsubsection{Model Uncertainty Sampling.}
-
-The model-uncertainty sampling strategy simply selects the instance that is closest to the model's decision boundary.
+\subsubsection{Uncertainty Sampling.}
+
+The uncertainty sampling strategy simply selects the instance that is closest to the model's decision boundary.
 %
@@
 
-\subsubsection{Information Gain Sampling.}
+\subsubsection{Information-Gain Sampling.}
 
@@
 
-\subsubsection{Model Convergence Stopping Criterion.}
-
-The model-convergence stopping criterion is less sophisticated in comparison.
+\subsubsection{Ranking Stopping Criterion.}
+
+The ranking stopping criterion is less sophisticated in comparison.
 It stops the active-learning process if the ranking induced by the model's predictions (Equation~\ref{eq:rankingeq}) remains unchanged within the last $l$ iterations.
@@
 
-In our experiments, we work with the SAT~Competition~2022 Anniversary Track instances~\cite{sat2022}.
-The dataset consists of 5355 instances with respective runtime data.
-It has complete runtimes of 28 solvers.
-We use the GBD metadata database~\cite{IserS18} to provide us with the aforementioned instance features, problem instances, and solver runtimes.
-Thereby, all our approaches make use of the 56 base features that are maintained in the \emph{base} database of GBD~\cite{IserS18}.
-They are inspired by the SATzilla features~\cite{features}.
-Those features comprise general instance-size features and graph-representation features among others.
-All features are numeric and fortunately free of any missing values.
+In our experiments, we work with a dataset based on the SAT~Competition~2022 Anniversary Track~\cite{sat2022}, retrieved from the GBD metadata database~\cite{IserS18}.
+The dataset consists of 5355 instances with respective runtime data of 28 solvers.
+To characterize instances, we use 56 instance features from the \emph{base} database.
+They are inspired by the SATzilla~2012 features~\cite{features} and comprise instance-size features and graph-representation features among others.
+All features are numeric and free of missing values.
 We drop 10 out of 56 features because of zero variance.
+Overall, prediction models have access to 46 instances features and 27 runtime features, i.e., excluding the current new solver~ $\hat{a}$.
+Additionally, we retrieve instance-family information to evaluate the composition of our sampled benchmarks.
+Instance families comprise instances derived from the same application domain, e.g., planning, cryptography, etc., and are a useful tool for both analyzing solver performance and portfolios.
+
 For hyper-parameter tuning, we randomly sample \SI{10}{\%} of the complete set of 5355 instances with stratification regarding the instance's family.
@@
   \item Observed PAR-2 ranking of already sampled instances
-  \item Predicted ranking induced by runtime-label predictions
+  \item Predicted runtime-label ranking
   \begin{itemize}
@@
     The latest $x$ predictions vote on the winner.
-    %JB: confuses me a bit as it sounds like we rank first and then vote on a winner; isn't it the other way around?
-    %TF: Removed confusing sentence
+    %JB: we vote for the label of each instance separately, right? (rather than globally ensembling the overall solver score of the last x iteration)
     \item
@@
     %TF: Not necessarily
+    %JB: but how can the scores of two existing solvers change? and we only have partially observed scores for the new solver
     As discussed in Section~\ref{sec:main:model}, ranking decisions are all correct if the difference in scores between two solvers exceeds a certain value.
@@
 
-For \emph{selection} (cf. Section~\ref{sec:main:sampling}), we experiment with the following methods, including our used hyper-parameter values:
+For \emph{selection} (cf. Section~\ref{sec:main:selection}), we experiment with the following methods, including our used hyper-parameter values:
 
@@
   \item Random sampling 
-  \item Model-based uncertainty sampling
+  \item Uncertainty sampling
   \begin{itemize}
@@
 
-  \item Model-based information-gain sampling
+  \item Information-gain sampling
   \begin{itemize}
     \item Fallback threshold: use random sampling for the first \SI{0}{\%}, \SI{5}{\%}, \SI{10}{\%}, \SI{15}{\%}, or \SI{20}{\%} of instances to explore the instance space.
-    \item Runtime scaling: prefer instances with the greatest information-gain per average runtime (\texttt{True} or \texttt{False}).
+    \item Runtime scaling: prefer instances with the greatest information gain per average runtime (\texttt{True} or \texttt{False}).
   \end{itemize}
@@
   \item Neighborhood-aware random sampling (preliminary experiments only); randomly chooses among the instances with the highest amount of non-sampled $k$ neighbors
-  \item Ranking-based sampling~\cite{Bossek021a} (preliminary experiments only)
-  \item Discrimination-based sampling~\cite{GentHJKMNN14} (preliminary experiments only)
-  \item Variance-based sampling~\cite{MatriconAFSH21} (preliminary experiments only)
+  \item Ranking sampling~\cite{Bossek021a} (preliminary experiments only)
+  \item Discrimination sampling~\cite{GentHJKMNN14} (preliminary experiments only)
+  \item Variance sampling~\cite{MatriconAFSH21} (preliminary experiments only)
 \end{itemize}
@@
 \begin{itemize}
-  \item Fixed subset size of \SI{10}{\%} or \SI{20}{\%} of instances
-  \item Ranking convergence criterion
+  \item Subset-size stopping criterion, using \SI{10}{\%} or \SI{20}{\%} of instances
+  \item Ranking stopping criterion
   \begin{itemize}
     \item Minimum amount: sample at least \SI{2}{\%}, \SI{8}{\%}, \SI{10}{\%}, or \SI{12}{\%} of instances before applying the criterion.
-    \item Convergence duration: stop if the predicted ranking stays the same for the last \SI{1}{\%} or \SI{2}{\%} of sampled instances.
-    %JB: why this fraction and not an absolute number of instances?
-the definition here might make stopping in early iterations rather likely, as just very few consecutive instances are needed (which might happen by coincidence; e.g., for ~5k instances, 2% minimum amount-> ~100 instances, 1% convergence duration -> ~1 instance), while the criterion becomes harder in later iterations
-    %TF: for that reason, it is combined with the "minimum amount". Using fractions makes the approach more useful when adapting it to benchmarks of different sizes.
-    %JB: yes, but (1) even combination with "minimum amount" can make stopping very easy (see my example) and (2) fraction for convergence could also be relative to absolute number of instances (which is constant over AL run) rather than fraction of currently sampled instances (which increases over AL run)
-    %TF: I'm sure that both would make sense. Isn't 1% convergence duration of 5k about ~50 instances?
+    \item Convergence duration: stop if the predicted ranking stays the same for a number of sampled instances equal to \SI{1}{\%} or \SI{2}{\%} of all instances.
   \end{itemize}
 
-  \item Wilcoxon criterion
+  \item Wilcoxon stopping criterion
   \begin{itemize}
@@
 Finally, we also analyze the importance of different instance families to our framework.
-Instance families comprise instances derived from the same application domain, e.g., planning, cryptography, etc., and are a useful tool for both analyzing solver performance and portfolios.
-%JB: instance families not really motivated till now, only appear in abstract and in text regarding stratification
-%TF: Added a sentence for clarification. We could introduce more details in section 4.3.
-%JB: sounds like a good idea
 
@@
 Note that we perform cross-validation on two levels, i.e., among the solvers and the target solver runtimes.
-%JB: we should state (were we introduce prediction model, i.e., either in 3.2 or a new subsection in 4) which features we use; I assume instance features + runtimes of other solvers?
-(else CV among solvers would not make sense for this experiment)
 We compare the average performances on the test sets.
@@
 Evaluating a particular configuration with Algorithm~\ref{alg:eval} returns a point $\left(O_{\operatorname{rt}},\, O_{\operatorname{acc}}\right)$.
+In particular, we do not show intermediate results of the active-learning procedure but only final results after stopping.
 The plotted lines represent the best performing configurations (convex hull) per top-level hyper-parameter choice, i.e., per ranking approach (Fig.~\ref{fig:annitraincolorranking}), selection approach (Fig.~\ref{fig:annitraincolorselection}), and stopping criterion (Fig.~\ref{fig:annitraincolorstopping}).
@@
 
-Regarding ranking approaches (Fig.~\ref{fig:annitraincolorranking}), using the $s_3$-induced ranking consistently outperforms the partially observed PAR-2 ranking for each possible value of the runtime-accuracy trade-off parameter~$\delta$.
-%JB: as in previous comment: consistent outperformance not guaranteed in case the lines are not individual approaches, but Pareto fronts within the compared categories (which means that different parts of the line might be related to different hyperparameter settings)
-%TF: The lines are in any case no individual approaches as individual approaches are only points. So, any lines in this plot are a "combination" of configs. It's only a question of how to combine.
-%JB: I always made the implicit assumption that the plots are based on learning curves, showing the progress within individual AL runs, and then multiple of these curves were combined (per top-level configuration); so the curves actually draw a line between different final results of AL runs?
-should be clarified somewhere
+Regarding ranking approaches (Fig.~\ref{fig:annitraincolorranking}), using the $s_3$-induced runtime-label ranking consistently outperforms the partially observed PAR-2 ranking for each possible value of the runtime-accuracy trade-off parameter~$\delta$.
 This is expected since selection decisions are not random.
@@
 
-Regarding the selection approaches (Fig.~\ref{fig:annitraincolorselection}), the model-based uncertainty strategy performs best in most cases.
-%JB: maybe we can drop "model-based" and use "uncertainty sampling" instead?
-I adapted this in Section 3, but section 4 + 5 + the plots would have to be adapted as well (else, we should revert my changes in Section 3)
-However, the model-based information-gain sampling is beneficial if runtime is strongly favored (very small $\delta$; runtime fraction less than \SI{5}{\%}).
-Random sampling performs quite well but is outperformed by the model-based uncertainty strategy in all cases, showing the benefit of actively selecting instances based on a prediction model.
-
-Regarding the stopping criterion (Fig.~\ref{fig:annitraincolorstopping}), the ranking convergence criterion performs the most consistently well.
+Regarding the selection approaches (Fig.~\ref{fig:annitraincolorselection}), uncertainty sampling performs best in most cases.
+However, information-gain sampling is beneficial if runtime is strongly favored (very small $\delta$; runtime fraction less than \SI{5}{\%}).
+Random sampling performs quite well but is outperformed by uncertainty sampling in all cases, showing the benefit of actively selecting instances based on a prediction model.
+
+Regarding the stopping criterion (Fig.~\ref{fig:annitraincolorstopping}), the ranking stopping criterion performs the most consistently well.
 If accuracy is strongly favored (very high $\delta$), the Wilcoxon stopping criterion performs better.
-%JB: plot says "Wilcoxon" without stopping; only minor point, but we should make naming of the components of our framework consistent
+The subset-size stopping criterion performs reasonably well but does not improve beyond a certain accuracy.
 
@@
   	The y-axis shows the ranking accuracy (cf. Section~\ref{sec:exdesign:goal}).
-  	%JB: we should use "ranking accuracy" in plots, since we explicitly defined this metric earlier
   	Each line entails the front of Pareto-optimal configurations for the respective hyper-parameter instantiation.
@@
 
-% One of the main observations is that our model-uncertainty-based sampling method (blue points) consistently outperformed random selection (red points) for almost any value of $\delta$ within Eq.~\ref{eq:opt}.
-% Moreover, our ranking-convergence criterion (dot-shaped points) is particularly suited to achieve a good ranking accuracy fast.
-% In comparison, our Wilcoxon-based criterion is suited better for high-accuracy predictions.
+% One of the main observations is that uncertainty sampling (blue points) consistently outperformed random sampling (red points) for almost any value of $\delta$ within Eq.~\ref{eq:opt}.
+% Moreover, our ranking stopping criterion (dot-shaped points) is particularly suited to achieve a good ranking accuracy fast.
+% In comparison, our Wilcoxon stopping criterion is suited better for high-accuracy predictions.
 % This comes at a high runtime cost though.
@@
 %   	%JB: no change in what? all columns except delta?
-%   	Ranking approaches are either predicted cluster-label ranking (Pred.) or partially observed PAR-2 ranking (PAR-2).
-%   	Selection approaches are either uncertainty-based selection (Uncert.) or information-gain-based selection (IG).
-%   	Stopping criteria are either ranking convergence (Conv.) or Wilcoxon-based (Wilcoxon).
+%   	Ranking approaches are either predicted runtime-label ranking (Pred.) or observed PAR-2 ranking (PAR-2).
+%   	Selection approaches are either uncertainty sampling (Uncert.) or information-gain sampling (IG).
+%   	Stopping criteria are either ranking (Rank) or Wilcoxon (Wilcoxon).
 %   	%JB: How are the values for ranking/selection/stopping determined -- whatever is best for a particular delta?
@@
 %     \hline
-%     0.00 & PAR-2 & Uncert. & Conv. & 0.021 & 0.025 & 0.593 & 0.979 \\
-%     0.05 & Pred. & Uncert. & Conv. & 0.024 & 0.158 & 0.765 & 0.966 \\
-%     0.10 & Pred. & Uncert. & Conv. & 0.025 & 0.128 & 0.779 & 0.955 \\
-%     0.15 & Pred. & IG & Conv. & 0.028 & 0.112 & 0.802 & 0.946 \\
+%     0.00 & PAR-2 & Uncert. & Rank & 0.021 & 0.025 & 0.593 & 0.979 \\
+%     0.05 & Pred. & Uncert. & Rank & 0.024 & 0.158 & 0.765 & 0.966 \\
+%     0.10 & Pred. & Uncert. & Rank & 0.025 & 0.128 & 0.779 & 0.955 \\
+%     0.15 & Pred. & IG & Rank & 0.028 & 0.112 & 0.802 & 0.946 \\
 %     $\vdots$ & $\vdots$ & $\vdots$ & $\vdots$ & $\vdots$ & $\vdots$ & $\vdots$ & $\vdots$ \\
-%     0.50 & Pred. & IG & Conv. & 0.028 & 0.112 & 0.802 & 0.887 \\
-%     0.55 & Pred. & Uncert. & Conv. & 0.053 & 0.142 & 0.824 & 0.880 \\
-%     0.60 & Pred. & Uncert. & Conv. & 0.146 & 0.082 & 0.897 & 0.880 \\
+%     0.50 & Pred. & IG & Rank & 0.028 & 0.112 & 0.802 & 0.887 \\
+%     0.55 & Pred. & Uncert. & Rank & 0.053 & 0.142 & 0.824 & 0.880 \\
+%     0.60 & Pred. & Uncert. & Rank & 0.146 & 0.082 & 0.897 & 0.880 \\
 %     $\vdots$ & $\vdots$ & $\vdots$ & $\vdots$ & $\vdots$ & $\vdots$ & $\vdots$ & $\vdots$ \\
-%     0.75 & Pred. & Uncert. & Conv. & 0.146 & 0.082 & 0.897 & 0.886 \\
-%     0.80 & Pred. & Uncert. & Conv. & 0.219 & 0.128 & 0.917 & 0.890 \\
-%     \hline
-%     0.85 & Pred. & Uncert. & Conv. & 0.272 & 0.171 & 0.927 & 0.897 \\
+%     0.75 & Pred. & Uncert. & Rank & 0.146 & 0.082 & 0.897 & 0.886 \\
+%     0.80 & Pred. & Uncert. & Rank & 0.219 & 0.128 & 0.917 & 0.890 \\
+%     \hline
+%     0.85 & Pred. & Uncert. & Rank & 0.272 & 0.171 & 0.927 & 0.897 \\
 %     \hline
@@
 Having selected the most promising hyper-parame\-ters, we run our end-to-end active-learning experiments on the complete Anniversary Track dataset (5355 instances).
-The best-performing approach for $\delta = 0.85$ uses runtime-label prediction ranking, model-based uncertainty sampling, and a ranking-convergence stopping criterion.
+The best-performing approach for $\delta = 0.85$ uses the predicted runtime-label ranking, uncertainty sampling, and a ranking stopping criterion.
 %JB: we need to motivate choosing this \delta
@@
 		The y-axis shows the average frequency of instance families in the samples selected by active learning.
-		The dashed line represents families that are occur with the same frequency in both, dataset and samples.
+		The dashed line represents families that occur with the same frequency in both, dataset and samples.
 	}
@@
 We have evaluated several ranking algorithms, instance-selection approaches, and stopping criteria within our sequential active-learning process.
-%A model-uncertainty-based selection approach in combination with runtime-label-prediction ranking and a ranking-convergence stopping criterion showed the consistently best results in most cases.
-%JB: "consistently best" -- it's the best approach for one delta, but not sure regarding other deltas nor its sensitivity to hyperparameters
-%TF: commented out
 We also took a brief look at which instance families are the most prevalent in selection decisions.

### 2022-10-11 paper/paper.tex
Dissolve Section 5.1 (Runtime Prediction)
Move content to Section 4 (Experimental Design) and cut. Also moved the
runtime-discretization results from Section 3.1. Paper is roughly a page
shorter now (partly also because we integrated Algo 2 into the text
instead of having it on a separate page).

@@
 We transform the runtimes into discrete runtime labels on a per-instance basis.
-For each instance $e \in \mathcal{I}$, we use a single-link hierarchical clustering algorithm to assign the runtimes in $\bigl\{ r(a, e) \mid a \in A \bigr\}$ to one of the $k$ clusters $C_1, \dots, C_k$ such that the fastest runtimes for instance $e$ are in cluster $C_1$ and the slowest are in cluster $C_k$.
+For each instance $e \in \mathcal{I}$, we use clustering algorithm to assign the runtimes in $\bigl\{ r(a, e) \mid a \in A \bigr\}$ to one of the $k$ clusters $C_1, \dots, C_k$ such that the fastest runtimes for instance $e$ are in cluster $C_1$ and the slowest are in cluster $C_k$.
 For instances with many timeouts, we have to manually fix the clustering results such that all timeouts $\tau$ are in cluster $C_k$.
@@
 
-However, we need to ensure that discretized labels still discriminate solvers and align with the actual PAR-2 ranking.
+\subsection{Instance Selection}
+\label{sec:main:selection}
+
+Selecting an instance based on the model is a core functionality of our framework (cf. Algorithm~\ref{algALBenchmark}, Line 2).
+In this section, we introduce our sampling strategies, which make use of the model's label-prediction function $f$ and are inspired by existing work within the realms of active learning~\cite{settles2009active}.
+We implement an uncertainty and an information-gain sampling strategy.
+These methods require the model's predictions to also include probabilities for the $k$ possible runtime labels.
+Let $f' : \mathcal{\hat A} \times \mathcal{I} \rightarrow \left[0, 1\right]^k$ denote this modified prediction function.
+
+\subsubsection{Uncertainty Sampling.}
+
+The uncertainty sampling strategy simply selects the instance that is closest to the model's decision boundary.
+%
+\begin{equation*}
+  \underset{e \in \mathcal{I} \setminus \tilde{\mathcal{I}}}{\arg\min} \left\lvert \frac{1}{k} - \max_{n \in \left\lbrace 1, \dots, k \right\rbrace} f'\!\left(\hat{a}, e\right)_{n} \right\rvert
+  %JB: \tilde I is undefined
+\end{equation*}
+
+\subsubsection{Information-Gain Sampling.}
+
+The information-gain sampling strategy selects the instance with the highest expected entropy reduction regarding the runtime labels of the instance.
+To be more specific, we select the instance $e$ that maximizes $IG(e)$, which is specified as follows:
+%
+\begin{equation*}
+  \operatorname{IG}(e) := \operatorname{H}(e) - \sum_{n = 1}^{k} f'(\hat{a}, e)_{n} \operatorname{\hat H}_n(e)
+\end{equation*}
+
+Here, $\operatorname{H}(e)$ denotes the entropy of the runtime labels $\gamma(a, e)$ over all $a \in \mathcal{A}$ and $\operatorname{H}(e, n)$ denotes the entropy of these labels plus $n$ as the runtime label for $\hat{a}$.
+The term $\operatorname{\hat H}_n(e)$ is computed for every possible runtime label $n \in \{1, \dots, k\}$.
+
+%JB: if this is info gain regarding runtimes, doesn't Eq.
+(4) mean that we would prefer instances with low entropy of runtime distribution after considering the new solver, i.e., instances that are not very discriminatory?
+%TF: I don't think so. We want to maximize the reduction in entropy of the runtime labels. Low entropy afterward -> we know that the solver behaves similarly to existing ones.
+%JB: but if we want to discriminate between solvers, what does it help to know that solver behaves similarly to other solvers?
+
+%Finally, the voting-based committee-disagreement approach requires a committee of $c$ models.
+%Rather than having a single prediction function $f$, there is a separate function $f_j$ for all committee members $j \in \left\lbrace 1, \dots, c  \right\rbrace$.
+%We then select the instance $i$ whose entropy of runtime-label votes by the different ensemble members is the highest.
+
+% MI: The following belongs either to Related Work or to Experimental Design (commented out)
+%\subsubsection{Baselines and existing approaches.}
+%\label{sec:sampling1}
+%Apart from obvious baseline instance-selection strategies, i.e., random sampling and neighborhood-aware sampling (randomly choosing among the instances with the highest amount of non-sampled $k$ neighbors), we also implement existing instance-selection strategies.
+
+%Bossek and Wagner~\cite{Bossek021a} introduce a custom ranking-score metric incorporating the number of \emph{good} pairs and \emph{bad} pairs.
+%A good solver pair, regarding a particular instance, are solvers that perform according to their global ranking.
+%Otherwise, it is a \emph{bad} pair.
+%Based on this classification, they assign a custom score to each instance.
+%By selecting instances with the highest score, they try to favor instances that are predictive of the final ranking.
+
+%Gent~et~al.~\cite{GentHJKMNN14} make use of a concept known from racing for automated model selection.
+%Given a particular instance and a parameter $\rho$, a solver is $\rho$-dominated if it is at least $\rho$ times worse than the best solver on this problem instance.
+%By selecting instances with the highest amount of $\rho$-dominated solvers, they hope to sample instances that provide the best discriminatory power to distinguish the runtimes of a new contestant.
+
+%Finally, we also implement some of the selection approaches described by Matricon~et.al.~\cite{MatriconAFSH21}.
+%Besides uniformly random sampling and the aforementioned discrimination-based method by Gent~et~al.~\cite{GentHJKMNN14}, they also describe a variance-based selection approach.
+%The variance-based strategy selects instances with the highest runtime variance per average runtime.
+%We also implement a logarithmized version of this, i.e., the highest log-runtime variance per average log runtime.
+
+\subsection{Stopping Criteria}
+\label{sec:main:stopping}
+
+%Similar to the instance-selection strategies, we can also divide stopping criteria into baseline and active-learning-based approaches.
+%Also, refer to~Algorithm~\ref{algALBenchmark}, line~3.
+%While a fixed subset-size criterion, e.g., only sample \SI{20}{\%} of all available instances, is an example of the former, a stopping criterion based on a Wilcoxon signed-rank test or the convergence of the model's predictions is an example for the latter.
+
+In this section, we present two dynamic stopping criteria, the Wilcoxon and the Model Uncertainty Stopping Criterion (cf. Algorithm~\ref{algALBenchmark}, Line~1).
+
+\subsubsection{Wilcoxon Stopping Criterion.}
+
+The Wilcoxon stopping criterion stops the active-learning process when we are certain enough that the predicted runtime labels of the new solver are sufficiently different from the labels of the existing solvers.
+We use the average $p$-value $W_{\hat{a}}$ of a Wilcoxon signed-rank test $w(S,P)$ of the two runtime label distributions $S=\{ \gamma(a, e) \mid e \in \mathcal{I} \}$ for an existing solver $a$ and $P=\{ f(\hat a, e) \mid e \in \mathcal{I} \}$ for the new solver $\hat{a}$:
+%
+\begin{equation*}
+  W_{\hat{a}} := \frac{1}{\lvert \mathcal{A} \rvert} \sum_{a \in \mathcal{A}} \operatorname{w}(S, P)
+\end{equation*}
+%
+To improve the stability of this criterion, we use an exponential moving average to smooth out outliers and stop as soon as $W^{(i)}_{\exp}$ drops below a certain threshold:
+%
+\begin{align*}
+	W_{\exp}^{\left(0\right)} &:= 1\\
+	W_{\exp}^{\left(i\right)} &:= \beta W_{\hat{a}} + \left(1 - \beta\right) W_{\exp}^{\left(i - 1\right)}
+\end{align*}
+
+\subsubsection{Ranking Stopping Criterion.}
+
+The ranking stopping criterion is less sophisticated in comparison.
+It stops the active-learning process if the ranking induced by the model's predictions (Equation~\ref{eq:rankingeq}) remains unchanged within the last $l$ iterations.
+Note that the concrete values of $s_k(\hat a)$ might still change.
+We are solemnly interested in the induced ranking in this case.
+
+\section{Experimental Design}
+\label{sec:exdesign}
+
+Given all the previously presented instantiations for Algorithm~\ref{algALBenchmark}, this section briefly outlines our experimental design, including our optimization goal, evaluation framework, used data sets, hyper-parameter choices, as well as implementation details.
+% hyper-parameter values that we experiment with
+
+\subsection{Optimization Goal}
+\label{sec:exdesign:goal}
+
+As already stated in the introductory section, this work addresses the \emph{New-Solver Problem} (cf.~Definition~\ref{def:new-solver-problem}).
+We create a prediction model $\mathcal{M}$ as described in Section~\ref{sec:main:model} that provides us with a scoring function $s_k$.
+% MI: Let this be the evaluation result:
+%Also, it outputs the fraction of runtime that is needed to arrive at its conclusion.
+First and foremost, our goal is to provide the engineer of new SAT solvers with an accurate ranking.
+We define the \emph{ranking accuracy} $O_{\operatorname{acc}} \in \left[0, 1\right]$ (higher is better) by the fraction of pairs $\left(\hat{a}, a\right)$ for all $a \in \mathcal{A}$ that are decided correctly by the given ranking (cf.~Algorithm~\ref{alg:eval}).
+For now, we exclude the equality of solvers since data shows that the PAR-2 scores of all our solvers are different.
+So, possible ranking decisions may be, for example, solver $a$ is better than $b$ or $b$ is better than $a$.
+The ranking accuracy is affected by whether this decision is correctly made.
+%Matricon et. al.~\cite{MatriconAFSH21} also evaluate their approach by the fraction of solver pairs that is correctly decided.
+Second, we also have to optimize for runtime.
+The \emph{fraction of runtime} that the algorithm needs to arrive at its conclusion is denoted by $O_{\operatorname{rt}} \in \left[0, 1\right]$ (lower is better).
+This metric puts the runtime summed over the sampled instances in relation to the runtime summed over all instances in the dataset (cf. Algorithm~\ref{alg:eval}).
+Overall, we want to find an approach that maximizes
+%
+\begin{equation}
+  O_\delta := \delta O_{\operatorname{acc}} + \left(1 - \delta\right) \left(1 - O_{\operatorname{rt}}\right) \enspace \textrm{,}
+  \label{eq:opt}
+\end{equation} 
+%
+whereby $\delta \in \left[0, 1\right]$ allows for linear weighting between the two optimization goals $O_{\operatorname{acc}}$ and $O_{\operatorname{rt}}$.
+Plotting the approaches that maximize $O_\delta$ for all $\delta \in \left[0, 1\right]$ on a $O_{\operatorname{rt}}$-$O_{\operatorname{acc}}$-diagram provides us with a Pareto front of the best approaches for different optimization-goal weightings.
+
+\subsection{Evaluation Framework}
+\label{sec:exdesign:evalframe}
+
+To evaluate a concrete instantiation of Algorithm~\ref{algALBenchmark} (a concrete choice for all the sub-routines), we perform cross-validation on our set of solvers.
+That means that each solver once plays the role of the new solver.
+Algorithm~\ref{alg:eval} shows this.
+Note that the \emph{new} solver in each iteration is excluded from the set of solvers $\mathcal{A}$ to avoid data leakage.
+Also, the runtime function $r$ is restricted to $\mathcal{A} \times \mathcal{I}$.
+
+\begin{algorithm}[htb]
+  \caption{Evaluation Framework}
+  \label{alg:eval}
+
+  %\KwIn{A pool of instances $\mathcal{I}$, a set of solvers $\mathcal{S}$ with known runtimes $r\!: \mathcal{S} \times \mathcal{I} \rightarrow \left[0, \tau\right] \cup \left\lbrace \perp \right\rbrace$, and PAR-2 scores $par2\!: \mathcal{S} \rightarrow \mathbb{R}$ for each solver.}
+  \KwIn{Solvers $\mathcal{A}$, Instaces $\mathcal{I}$, Runtimes $r : \mathcal{A} \times \mathcal{I} \rightarrow [0, \tau]$}
+  \KwOut{Average Rank Accuracy $\bar{O}_{\operatorname{acc}}$, Average Fraction of Runtime $\bar{O}_{\operatorname{rt}}$}
+  \BlankLine
+
+  $O \leftarrow \emptyset$
+  \BlankLine
+
+  \For{$\hat{a} \in \mathcal{A}$}{
+    $\mathcal{A}' \leftarrow \mathcal{A} \setminus \left\lbrace \hat{a} \right\rbrace$ \;
+    $(s_{\hat a}, \mathcal{R}) \leftarrow \operatorname{runALAlgorithm}(\mathcal{A}', \mathcal{I}, r, \hat{a})$ \tcp*{Refer to Algorithm~\ref{algALBenchmark}}
+
+	\BlankLine
+	\tcp{Determine Rank Accuracy}
+    $O_{\operatorname{acc}} \leftarrow 0$ \;    
+    \For{$a \in \mathcal{A}$}{
+    	\If{$\bigl(s_k(a) - s_{\hat a}\bigr) \cdot \bigl(\operatorname{par_2}(a) - \operatorname{par_2}(\hat a)\bigr) > 0$}{
+   		    $O_{\operatorname{acc}} \leftarrow O_{\operatorname{acc}} + \frac{1}{|\mathcal{A}|}$ \;
+    	}
+    }
+
+	\BlankLine
+	\tcp{Determine Runtime Fraction}
+	$r \leftarrow \sum\limits_{e \in \mathcal{I}} r(\hat a, e)$\;
+	$O_{\operatorname{rt}} \leftarrow 0$\;
+    \For{$e \in \mathcal{I}$}{
+		\If{$\exists t, (e,t) \in \mathcal{R}$}{
+			$O_{\operatorname{rt}} \leftarrow O_{\operatorname{rt}} + \frac{t}{r}$\;
+		}
+	}
+
+	\BlankLine
+    $O \leftarrow O \cup \bigl\{ ( O_{\operatorname{acc}},\, O_{\operatorname{rt}} ) \bigr\}$
+  }
+
+  \BlankLine
+  $\bigl( \bar{O}_{\operatorname{acc}}, \bar{O}_{\operatorname{rt}} \bigr) \leftarrow \operatorname{average}(O)$ \;
+
+  \BlankLine
+  \Return $\bigl( \bar{O}_{\operatorname{acc}}, \bar{O}_{\operatorname{rt}} \bigr)$
+\end{algorithm}
+
+\subsection{Data}
+
+In our experiments, we work with a dataset based on the SAT~Competition~2022 Anniversary Track~\cite{sat2022}, retrieved from the GBD metadata database~\cite{IserS18}.
+The dataset consists of 5355 instances with respective runtime data of 28 solvers.
+To characterize instances, we use 56 instance features from the \emph{base} database.
+They are inspired by the SATzilla~2012 features~\cite{features} and comprise instance-size features and graph-representation features among others.
+All features are numeric and free of missing values.
+We drop 10 out of 56 features because of zero variance.
+Overall, prediction models have access to 46 instances features and 27 runtime features, i.e., excluding the current new solver~ $\hat{a}$.
+Additionally, we retrieve instance-family information to evaluate the composition of our sampled benchmarks.
+Instance families comprise instances derived from the same application domain, e.g., planning, cryptography, etc., and are a useful tool for both analyzing solver performance and portfolios.
+
+For hyper-parameter tuning, we randomly sample \SI{10}{\%} of the complete set of 5355 instances with stratification regarding the instance's family.
+All instance families that are too \emph{small}, i.e., \SI{10}{\%} of them corresponds to less than one instance, are put into one meta-family for stratification.
+This \emph{tuning dataset} allows for a more extensive exploration of hyper-parameter space.
+
+\subsection{Hyper-parameters}
+\label{sec:exdesign:hyper}
+
+Given Algorithm~\ref{algALBenchmark}, there are several possible instantiations for the three phases, i.e., \emph{selection}, \emph{stopping}, and \emph{ranking}.
+Also, there are different choices for the runtime-label prediction model and runtime discretization.
+Note that we are not considering all previously listed approaches since a grid search of all combinations would be infeasible.
+Rather, we filter approaches based on preliminary experimental results and do the end-to-end experiments only with a subset.
+The end-to-end experiment configurations are given below.
+
+\subsubsection{Ranking.}
+
+Regarding \emph{ranking} (cf. Section~\ref{sec:main:model}), we experiment with the following approaches, including our used hyper-parameter values:
+
+\begin{itemize}
+  \item Observed PAR-2 ranking of already sampled instances
+  \item Predicted runtime-label ranking
+  \begin{itemize}
+    \item
+    History size: consider the latest 1, 10, 20, 30, or 40 predictions within a voting approach for stability.
+    The latest $x$ predictions vote on the winner.
+    %JB: we vote for the label of each instance separately, right? (rather than globally ensembling the overall solver score of the last x iteration)
+    \item
+    Fallback threshold: if the difference of scores between two solvers drops below \SI{0.01}, \SI{0.05}, or \SI{0.1}, use the partially observed PAR-2 ranking as a tie-breaker.
+    %JB: as clarification: one these two solvers has to be the new solver, right? if yes, would write it explicitly
+    %TF: Not necessarily
+    %JB: but how can the scores of two existing solvers change? and we only have partially observed scores for the new solver
+    As discussed in Section~\ref{sec:main:model}, ranking decisions are all correct if the difference in scores between two solvers exceeds a certain value.
+  \end{itemize}
+\end{itemize}
+
+\subsubsection{Selection.}
+
+For \emph{selection} (cf. Section~\ref{sec:main:selection}), we experiment with the following methods, including our used hyper-parameter values:
+
+\begin{itemize}
+  \item Random sampling 
+  \item Uncertainty sampling
+  \begin{itemize}
+    \item Fallback threshold: use random sampling for the first \SI{0}{\%}, \SI{5}{\%}, \SI{10}{\%}, \SI{15}{\%}, or \SI{20}{\%} of instances to explore the instance space.
+    \item Runtime scaling: prefer instances with the greatest uncertainty per average (over all solvers) runtime (\texttt{True} or \texttt{False}).
+  \end{itemize}
+
+  \item Information-gain sampling
+  \begin{itemize}
+    \item Fallback threshold: use random sampling for the first \SI{0}{\%}, \SI{5}{\%}, \SI{10}{\%}, \SI{15}{\%}, or \SI{20}{\%} of instances to explore the instance space.
+    \item Runtime scaling: prefer instances with the greatest information gain per average runtime (\texttt{True} or \texttt{False}).
+  \end{itemize}
+
+  \item Neighborhood-aware random sampling (preliminary experiments only); randomly chooses among the instances with the highest amount of non-sampled $k$ neighbors
+  \item Ranking sampling~\cite{Bossek021a} (preliminary experiments only)
+  \item Discrimination sampling~\cite{GentHJKMNN14} (preliminary experiments only)
+  \item Variance sampling~\cite{MatriconAFSH21} (preliminary experiments only)
+\end{itemize}
+
+\subsubsection{Stopping.}
+
+For \emph{stopping} decisions (cf. Section~\ref{sec:main:stopping}), we experiment with the following criteria, including our used hyper-parameter values:
+
+\begin{itemize}
+  \item Subset-size stopping criterion, using \SI{10}{\%} or \SI{20}{\%} of instances
+  \item Ranking stopping criterion
+  \begin{itemize}
+    \item Minimum amount: sample at least \SI{2}{\%}, \SI{8}{\%}, \SI{10}{\%}, or \SI{12}{\%} of instances before applying the criterion.
+    \item Convergence duration: stop if the predicted ranking stays the same for a number of sampled instances equal to \SI{1}{\%} or \SI{2}{\%} of all instances.
+  \end{itemize}
+
+  \item Wilcoxon stopping criterion
+  \begin{itemize}
+    \item Minimum amount: sample at least \SI{2}{\%}, \SI{8}{\%}, \SI{10}{\%}, or \SI{12}{\%} of instances before applying the criterion.
+    \item Average of $p$-values to drop below: \SI{5}{\%}.
+    \item Exponential-moving average: incorporate previous significance values by using an EMA with $\beta = 0.1$ or $\beta = 0.7$.
+  \end{itemize}
+\end{itemize}
+
+\subsubsection{Prediction model.}
+
+For runtime-label prediction, we only use one model configuration in our experiments since an exhaustive grid search would be infeasible.
+In preliminary experiments, we compared a variety of model types from \emph{scikit-learn}~\cite{scikit-learn}.
+In particular, we conducted nested cross-validation, including hyper-parameter tuning, and used Matthews Correlation Coefficient~\cite{gorodkin2004comparing,matthews1975comparison} to assess the performance for predicting runtime labels.
+Our final choice is a stacking ensemble of a quadratic-discriminant analysis and a random forest, using a decision tree to combine its ensemble members.
+
+%Our model of choice is an ensemble, stacking a quadratic-discriminant analysis\footnote{With a regularization parameter of zero and a singular value threshold of zero} onto a random forest model\footnote{With \emph{entropy} splitting criterion and \emph{balanced} class weights} using a decision tree\footnote{With \emph{Gini} impurity splitting criterion; choosing the best split for each branching decision and maximum tree depth of 5 levels} to weight its members.
+%If not stated explicitly, all other model parameters default to the presets of \emph{scikit-learn}\footnote{\url{https://scikit-learn.org/stable/index.html}; Version 1.0.2}~\cite{scikit-learn}.
+%
+%
+%Therefore, we compare predictors first and select one runtime-prediction model for all further experiments.
+%We evaluate predictors with the \emph{Matthews Correlation Coefficient} (MCC) scores~\cite{gorodkin2004comparing,matthews1975comparison} of the classification task.
+%MCC scores are great in dealing with class imbalances in contrast to conventional metrics like accuracy.
+%
+%% \begin{table}
+%	%   \centering
+%	%   \caption{
+%		%   	Time-out prediction Matthews Correlation Coefficient performances.
+%		%   	Results are obtained using cross-validation.
+%		%   	We provide \SI{64}{\%} of ground-truth for training, use \SI{16}{\%} as a validation set, and \SI{20}{\%} as a test set (5-fold cross-validation).
+%		%   }
+%	%   \label{tab:mcc}
+%	%   \begin{tabular}{lc}
+%		%     Prediction Model & Avg. MCC (Std.) \\
+%		%     \hline
+%		%     Stacking ensemble (QDA, RF)           & 0.9527 (0.0292) \\
+%		%     Quadratic-discriminant analysis (QDA) & 0.9290 (0.0339) \\
+%		%     Random forest (RF)                    & 0.8530 (0.0479) \\
+%		%     AdaBoost with decision trees          & 0.8384 (0.0444) \\
+%		%     AdaBoost with logistic regression     & 0.8158 (0.0985) \\
+%		%     Decision tree                         & 0.8059 (0.0707) \\
+%		%     Logistic regression                   & 0.8052 (0.1018) \\
+%		%     kNN classifier                        & 0.7885 (0.1521) \\
+%		%     MLP classifier                        & 0.7760 (0.1408) \\
+%		%     Support-vector classifier             & 0.7757 (0.2149) \\
+%		%     Naive Bayes                           & 0.7306 (0.1394) \\
+%		%     \hline
+%		%   \end{tabular}
+%	% \end{table}
+%
+%% Stacking MCC
+%% 0.888221577313471
+%% 0.04753979056930393
+%
+%%To obtain the results in Table~\ref{tab:mcc},
+%We perform cross-validation among all solvers: we assume that all other solver runtimes are known and provide \SI{64}{\%} of ground-truth runtime labels of the target solver for training, use \SI{16}{\%} as a validation set for hyper-parameter tuning of the runtime predictor, and \SI{20}{\%} as a test set (5-fold cross-validation).
+%This is repeated for each solver once.
+%Note that we perform cross-validation on two levels, i.e., among the solvers and the target solver runtimes.
+%We compare the average performances on the test sets.
+%
+%ML classification models that we have tried include random forests, decision trees, quadratic-discriminant analysis, AdaBoost, logistic regression, $k$-nearest neighbor, multi-layer perceptrons, support vector machines, and naive Bayes.
+%A stacking ensemble consisting of quadratic-discriminant analysis and random forests performed best.
+%It produces an MCC score of \SI{0.8882}{} for predicting three-class runtime labels.
+%% \SI{0.9527}{} for time-out prediction
+%The results are significantly better than any other model that we have tried regarding a Wilcoxon signed-rank test with $\alpha = \SI{5}{\%}$.
+%%Since the differences in the MCC-performance distributions of the best and second-best approaches are statistically significant regarding a Wilcoxon signed-rank test with $\alpha = \SI{5}{\%}$, we only use the best approach for all further experiments as a prediction model.
+
+\subsubsection{Runtime discretization.}
+
+To define prediction targets, i.e., discretized runtime labels, we use hierarchical clustering with $k = 3$ and a log-single-link criterion, which produced the most \emph{useful} labels in preliminary experiments.
+We denote this adapted solver scoring function with $s_3$.
+Other clustering approaches that we have tried include hierarchical clustering with mean-, median- and complete-link criteria, as well as $k$-means and spectral clustering.
+%Predicting runtime labels in a similar experiment as above produced an MCC score of on average \SI{88.82}{\%} ($\sigma = \SI{4.75}{\%}$; in comparison to \SI{95.27}{\%} for predicting time-outs).
+%JB: "similar experiment as above" - reference unclear; what do you want to say?
+%TF: Removed the sentence
+In our chosen hierarchical procedure, each non-time-out runtime starts in a separate interval.
+We then gradually merge intervals whose single-link logarithmic distance is the smallest until the desired number of partitions is reached.
+
+To obtain \emph{useful} labels, we need to ensure that discretized labels still discriminate solvers and align with the actual PAR-2 ranking.
 We analyzed the ranking induced by $s_3$ in preliminary experiments with the SAT Competition~2022 Anniversary Track~\cite{sat2022}.
@@
 %\footnote{Based on hierarchical clustering with a log-single-link criterion for $k = 3$.}
-Further, our ranking approach is capable of correctly deciding for almost all (about \SI{97.45}{\%}; $\sigma = \SI{3.68}{\%}$) solver pairs which solver is faster.
+Further, our ranking approach correctly decides for almost all (about \SI{97.45}{\%}; $\sigma = \SI{3.68}{\%}$) solver pairs which solver is faster.
 %Moreover, it decides all solvers correctly if the difference in the cluster-label-ranking scores between two solvers is at least \SI{0.11}{}.
@@
 
-\subsection{Instance Selection}
-\label{sec:main:selection}
-
-Selecting an instance based on the model is a core functionality of our framework (cf. Algorithm~\ref{algALBenchmark}, Line 2).
-In this section, we introduce our sampling strategies, which make use of the model's label-prediction function $f$ and are inspired by existing work within the realms of active learning~\cite{settles2009active}.
-We implement an uncertainty and an information-gain sampling strategy.
-These methods require the model's predictions to also include probabilities for the $k$ possible runtime labels.
-Let $f' : \mathcal{\hat A} \times \mathcal{I} \rightarrow \left[0, 1\right]^k$ denote this modified prediction function.
-
-\subsubsection{Uncertainty Sampling.}
-
-The uncertainty sampling strategy simply selects the instance that is closest to the model's decision boundary.
-%
-\begin{equation*}
-  \underset{e \in \mathcal{I} \setminus \tilde{\mathcal{I}}}{\arg\min} \left\lvert \frac{1}{k} - \max_{n \in \left\lbrace 1, \dots, k \right\rbrace} f'\!\left(\hat{a}, e\right)_{n} \right\rvert
-  %JB: \tilde I is undefined
-\end{equation*}
-
-\subsubsection{Information-Gain Sampling.}
-
-The information-gain sampling strategy selects the instance with the highest expected entropy reduction regarding the runtime labels of the instance.
-To be more specific, we select the instance $e$ that maximizes $IG(e)$, which is specified as follows:
-%
-\begin{equation*}
-  \operatorname{IG}(e) := \operatorname{H}(e) - \sum_{n = 1}^{k} f'(\hat{a}, e)_{n} \operatorname{\hat H}_n(e)
-\end{equation*}
-
-Here, $\operatorname{H}(e)$ denotes the entropy of the runtime labels $\gamma(a, e)$ over all $a \in \mathcal{A}$ and $\operatorname{H}(e, n)$ denotes the entropy of these labels plus $n$ as the runtime label for $\hat{a}$.
-The term $\operatorname{\hat H}_n(e)$ is computed for every possible runtime label $n \in \{1, \dots, k\}$.
-
-%JB: if this is info gain regarding runtimes, doesn't Eq.
-(4) mean that we would prefer instances with low entropy of runtime distribution after considering the new solver, i.e., instances that are not very discriminatory?
-%TF: I don't think so. We want to maximize the reduction in entropy of the runtime labels. Low entropy afterward -> we know that the solver behaves similarly to existing ones.
-%JB: but if we want to discriminate between solvers, what does it help to know that solver behaves similarly to other solvers?
-
-%Finally, the voting-based committee-disagreement approach requires a committee of $c$ models.
-%Rather than having a single prediction function $f$, there is a separate function $f_j$ for all committee members $j \in \left\lbrace 1, \dots, c  \right\rbrace$.
-%We then select the instance $i$ whose entropy of runtime-label votes by the different ensemble members is the highest.
-
-% MI: The following belongs either to Related Work or to Experimental Design (commented out)
-%\subsubsection{Baselines and existing approaches.}
-%\label{sec:sampling1}
-%Apart from obvious baseline instance-selection strategies, i.e., random sampling and neighborhood-aware sampling (randomly choosing among the instances with the highest amount of non-sampled $k$ neighbors), we also implement existing instance-selection strategies.
-
-%Bossek and Wagner~\cite{Bossek021a} introduce a custom ranking-score metric incorporating the number of \emph{good} pairs and \emph{bad} pairs.
-%A good solver pair, regarding a particular instance, are solvers that perform according to their global ranking.
-%Otherwise, it is a \emph{bad} pair.
-%Based on this classification, they assign a custom score to each instance.
-%By selecting instances with the highest score, they try to favor instances that are predictive of the final ranking.
-
-%Gent~et~al.~\cite{GentHJKMNN14} make use of a concept known from racing for automated model selection.
-%Given a particular instance and a parameter $\rho$, a solver is $\rho$-dominated if it is at least $\rho$ times worse than the best solver on this problem instance.
-%By selecting instances with the highest amount of $\rho$-dominated solvers, they hope to sample instances that provide the best discriminatory power to distinguish the runtimes of a new contestant.
-
-%Finally, we also implement some of the selection approaches described by Matricon~et.al.~\cite{MatriconAFSH21}.
-%Besides uniformly random sampling and the aforementioned discrimination-based method by Gent~et~al.~\cite{GentHJKMNN14}, they also describe a variance-based selection approach.
-%The variance-based strategy selects instances with the highest runtime variance per average runtime.
-%We also implement a logarithmized version of this, i.e., the highest log-runtime variance per average log runtime.
-
-\subsection{Stopping Criteria}
-\label{sec:main:stopping}
-
-%Similar to the instance-selection strategies, we can also divide stopping criteria into baseline and active-learning-based approaches.
-%Also, refer to~Algorithm~\ref{algALBenchmark}, line~3.
-%While a fixed subset-size criterion, e.g., only sample \SI{20}{\%} of all available instances, is an example of the former, a stopping criterion based on a Wilcoxon signed-rank test or the convergence of the model's predictions is an example for the latter.
-
-In this section, we present two dynamic stopping criteria, the Wilcoxon and the Model Uncertainty Stopping Criterion (cf. Algorithm~\ref{algALBenchmark}, Line~1).
-
-\subsubsection{Wilcoxon Stopping Criterion.}
-
-The Wilcoxon stopping criterion stops the active-learning process when we are certain enough that the predicted runtime labels of the new solver are sufficiently different from the labels of the existing solvers.
-We use the average $p$-value $W_{\hat{a}}$ of a Wilcoxon signed-rank test $w(S,P)$ of the two runtime label distributions $S=\{ \gamma(a, e) \mid e \in \mathcal{I} \}$ for an existing solver $a$ and $P=\{ f(\hat a, e) \mid e \in \mathcal{I} \}$ for the new solver $\hat{a}$:
-%
-\begin{equation*}
-  W_{\hat{a}} := \frac{1}{\lvert \mathcal{A} \rvert} \sum_{a \in \mathcal{A}} \operatorname{w}(S, P)
-\end{equation*}
-%
-To improve the stability of this criterion, we use an exponential moving average to smooth out outliers and stop as soon as $W^{(i)}_{\exp}$ drops below a certain threshold:
-%
-\begin{align*}
-	W_{\exp}^{\left(0\right)} &:= 1\\
-	W_{\exp}^{\left(i\right)} &:= \beta W_{\hat{a}} + \left(1 - \beta\right) W_{\exp}^{\left(i - 1\right)}
-\end{align*}
-
-\subsubsection{Ranking Stopping Criterion.}
-
-The ranking stopping criterion is less sophisticated in comparison.
-It stops the active-learning process if the ranking induced by the model's predictions (Equation~\ref{eq:rankingeq}) remains unchanged within the last $l$ iterations.
-Note that the concrete values of $s_k(\hat a)$ might still change.
-We are solemnly interested in the induced ranking in this case.
-
-\section{Experimental Design}
-\label{sec:exdesign}
-
-Given all the previously presented instantiations for Algorithm~\ref{algALBenchmark}, this section briefly outlines our experimental design, including our optimization goal, evaluation framework, used data sets, hyper-parameter choices, as well as implementation details.
-% hyper-parameter values that we experiment with
-
-\subsection{Optimization Goal}
-\label{sec:exdesign:goal}
-
-As already stated in the introductory section, this work addresses the \emph{New-Solver Problem} (cf.~Definition~\ref{def:new-solver-problem}).
-We create a prediction model $\mathcal{M}$ as described in Section~\ref{sec:main:model} that provides us with a scoring function $s_k$.
-% MI: Let this be the evaluation result:
-%Also, it outputs the fraction of runtime that is needed to arrive at its conclusion.
-First and foremost, our goal is to provide the engineer of new SAT solvers with an accurate ranking.
-We define the \emph{ranking accuracy} $O_{\operatorname{acc}} \in \left[0, 1\right]$ (higher is better) by the fraction of pairs $\left(\hat{a}, a\right)$ for all $a \in \mathcal{A}$ that are decided correctly by the given ranking (cf.~Algorithm~\ref{alg:eval}).
-For now, we exclude the equality of solvers since data shows that the PAR-2 scores of all our solvers are different.
-So, possible ranking decisions may be, for example, solver $a$ is better than $b$ or $b$ is better than $a$.
-The ranking accuracy is affected by whether this decision is correctly made.
-%Matricon et. al.~\cite{MatriconAFSH21} also evaluate their approach by the fraction of solver pairs that is correctly decided.
-Second, we also have to optimize for runtime.
-The \emph{fraction of runtime} that the algorithm needs to arrive at its conclusion is denoted by $O_{\operatorname{rt}} \in \left[0, 1\right]$ (lower is better).
-This metric puts the runtime summed over the sampled instances in relation to the runtime summed over all instances in the dataset (cf. Algorithm~\ref{alg:eval}).
-Overall, we want to find an approach that maximizes
-%
-\begin{equation}
-  O_\delta := \delta O_{\operatorname{acc}} + \left(1 - \delta\right) \left(1 - O_{\operatorname{rt}}\right) \enspace \textrm{,}
-  \label{eq:opt}
-\end{equation} 
-%
-whereby $\delta \in \left[0, 1\right]$ allows for linear weighting between the two optimization goals $O_{\operatorname{acc}}$ and $O_{\operatorname{rt}}$.
-Plotting the approaches that maximize $O_\delta$ for all $\delta \in \left[0, 1\right]$ on a $O_{\operatorname{rt}}$-$O_{\operatorname{acc}}$-diagram provides us with a Pareto front of the best approaches for different optimization-goal weightings.
-
-\subsection{Evaluation Framework}
-\label{sec:exdesign:evalframe}
-
-To evaluate a concrete instantiation of Algorithm~\ref{algALBenchmark} (a concrete choice for all the sub-routines), we perform cross-validation on our set of solvers.
-That means that each solver once plays the role of the new solver.
-Algorithm~\ref{alg:eval} shows this.
-Note that the \emph{new} solver in each iteration is excluded from the set of solvers $\mathcal{A}$ to avoid data leakage.
-Also, the runtime function $r$ is restricted to $\mathcal{A} \times \mathcal{I}$.
-
-\begin{algorithm}[htbp]
-  \caption{Evaluation Framework}
-  \label{alg:eval}
-
-  %\KwIn{A pool of instances $\mathcal{I}$, a set of solvers $\mathcal{S}$ with known runtimes $r\!: \mathcal{S} \times \mathcal{I} \rightarrow \left[0, \tau\right] \cup \left\lbrace \perp \right\rbrace$, and PAR-2 scores $par2\!: \mathcal{S} \rightarrow \mathbb{R}$ for each solver.}
-  \KwIn{Solvers $\mathcal{A}$, Instaces $\mathcal{I}$, Runtimes $r : \mathcal{A} \times \mathcal{I} \rightarrow [0, \tau]$}
-  \KwOut{Average Rank Accuracy $\bar{O}_{\operatorname{acc}}$, Average Fraction of Runtime $\bar{O}_{\operatorname{rt}}$}
-  \BlankLine
-
-  $O \leftarrow \emptyset$
-  \BlankLine
-
-  \For{$\hat{a} \in \mathcal{A}$}{
-    $\mathcal{A}' \leftarrow \mathcal{A} \setminus \left\lbrace \hat{a} \right\rbrace$ \;
-    $(s_{\hat a}, \mathcal{R}) \leftarrow \operatorname{runALAlgorithm}(\mathcal{A}', \mathcal{I}, r, \hat{a})$ \tcp*{Refer to Algorithm~\ref{algALBenchmark}}
-
-	\BlankLine
-	\tcp{Determine Rank Accuracy}
-    $O_{\operatorname{acc}} \leftarrow 0$ \;    
-    \For{$a \in \mathcal{A}$}{
-    	\If{$\bigl(s_k(a) - s_{\hat a}\bigr) \cdot \bigl(\operatorname{par_2}(a) - \operatorname{par_2}(\hat a)\bigr) > 0$}{
-   		    $O_{\operatorname{acc}} \leftarrow O_{\operatorname{acc}} + \frac{1}{|\mathcal{A}|}$ \;
-    	}
-    }
-
-	\BlankLine
-	\tcp{Determine Runtime Fraction}
-	$r \leftarrow \sum\limits_{e \in \mathcal{I}} r(\hat a, e)$\;
-	$O_{\operatorname{rt}} \leftarrow 0$\;
-    \For{$e \in \mathcal{I}$}{
-		\If{$\exists t, (e,t) \in \mathcal{R}$}{
-			$O_{\operatorname{rt}} \leftarrow O_{\operatorname{rt}} + \frac{t}{r}$\;
-		}
-	}
-
-	\BlankLine
-    $O \leftarrow O \cup \bigl\{ ( O_{\operatorname{acc}},\, O_{\operatorname{rt}} ) \bigr\}$
-  }
-
-  \BlankLine
-  $\bigl( \bar{O}_{\operatorname{acc}}, \bar{O}_{\operatorname{rt}} \bigr) \leftarrow \operatorname{average}(O)$ \;
-
-  \BlankLine
-  \Return $\bigl( \bar{O}_{\operatorname{acc}}, \bar{O}_{\operatorname{rt}} \bigr)$
-\end{algorithm}
-
-\subsection{Data}
-
-In our experiments, we work with a dataset based on the SAT~Competition~2022 Anniversary Track~\cite{sat2022}, retrieved from the GBD metadata database~\cite{IserS18}.
-The dataset consists of 5355 instances with respective runtime data of 28 solvers.
-To characterize instances, we use 56 instance features from the \emph{base} database.
-They are inspired by the SATzilla~2012 features~\cite{features} and comprise instance-size features and graph-representation features among others.
-All features are numeric and free of missing values.
-We drop 10 out of 56 features because of zero variance.
-Overall, prediction models have access to 46 instances features and 27 runtime features, i.e., excluding the current new solver~ $\hat{a}$.
-Additionally, we retrieve instance-family information to evaluate the composition of our sampled benchmarks.
-Instance families comprise instances derived from the same application domain, e.g., planning, cryptography, etc., and are a useful tool for both analyzing solver performance and portfolios.
-
-For hyper-parameter tuning, we randomly sample \SI{10}{\%} of the complete set of 5355 instances with stratification regarding the instance's family.
-All instance families that are too \emph{small}, i.e., \SI{10}{\%} of them corresponds to less than one instance, are put into one meta-family for stratification.
-This smaller dataset allows for a more extensive exploration of hyper-parameter space.
-
-\subsection{Hyper-parameters}
-\label{sec:exdesign:hyper}
-
-Given Algorithm~\ref{algALBenchmark}, there are several possible instantiations for the three phases, i.e., \emph{selection}, \emph{stopping}, and \emph{ranking}.
-Also, there are different choices for the runtime-label prediction model.
-Note that we are not considering all previously listed approaches since a grid search of all combinations would be infeasible.
-Rather, we filter approaches based on preliminary experimental results (cf. Section~\ref{sec:eval:prediction}) and do the main end-to-end experiments only with a subset.
-The end-to-end experiment configurations are given below.
-
-\subsubsection{Ranking.}
-
-Regarding \emph{ranking} (cf. Section~\ref{sec:main:model}), we experiment with the following approaches, including our used hyper-parameter values:
-
-\begin{itemize}
-  \item Observed PAR-2 ranking of already sampled instances
-  \item Predicted runtime-label ranking
-  \begin{itemize}
-    \item
-    History size: consider the latest 1, 10, 20, 30, or 40 predictions within a voting approach for stability.
-    The latest $x$ predictions vote on the winner.
-    %JB: we vote for the label of each instance separately, right? (rather than globally ensembling the overall solver score of the last x iteration)
-    \item
-    Fallback threshold: if the difference of scores between two solvers drops below \SI{0.01}, \SI{0.05}, or \SI{0.1}, use the partially observed PAR-2 ranking as a tie-breaker.
-    %JB: as clarification: one these two solvers has to be the new solver, right? if yes, would write it explicitly
-    %TF: Not necessarily
-    %JB: but how can the scores of two existing solvers change? and we only have partially observed scores for the new solver
-    As discussed in Section~\ref{sec:main:model}, ranking decisions are all correct if the difference in scores between two solvers exceeds a certain value.
-  \end{itemize}
-\end{itemize}
-
-\subsubsection{Selection.}
-
-For \emph{selection} (cf. Section~\ref{sec:main:selection}), we experiment with the following methods, including our used hyper-parameter values:
-
-\begin{itemize}
-  \item Random sampling 
-  \item Uncertainty sampling
-  \begin{itemize}
-    \item Fallback threshold: use random sampling for the first \SI{0}{\%}, \SI{5}{\%}, \SI{10}{\%}, \SI{15}{\%}, or \SI{20}{\%} of instances to explore the instance space.
-    \item Runtime scaling: prefer instances with the greatest uncertainty per average (over all solvers) runtime (\texttt{True} or \texttt{False}).
-  \end{itemize}
-
-  \item Information-gain sampling
-  \begin{itemize}
-    \item Fallback threshold: use random sampling for the first \SI{0}{\%}, \SI{5}{\%}, \SI{10}{\%}, \SI{15}{\%}, or \SI{20}{\%} of instances to explore the instance space.
-    \item Runtime scaling: prefer instances with the greatest information gain per average runtime (\texttt{True} or \texttt{False}).
-  \end{itemize}
-
-  \item Neighborhood-aware random sampling (preliminary experiments only); randomly chooses among the instances with the highest amount of non-sampled $k$ neighbors
-  \item Ranking sampling~\cite{Bossek021a} (preliminary experiments only)
-  \item Discrimination sampling~\cite{GentHJKMNN14} (preliminary experiments only)
-  \item Variance sampling~\cite{MatriconAFSH21} (preliminary experiments only)
-\end{itemize}
-
-\subsubsection{Stopping.}
-
-For \emph{stopping} decisions (cf. Section~\ref{sec:main:stopping}), we experiment with the following criteria, including our used hyper-parameter values:
-
-\begin{itemize}
-  \item Subset-size stopping criterion, using \SI{10}{\%} or \SI{20}{\%} of instances
-  \item Ranking stopping criterion
-  \begin{itemize}
-    \item Minimum amount: sample at least \SI{2}{\%}, \SI{8}{\%}, \SI{10}{\%}, or \SI{12}{\%} of instances before applying the criterion.
-    \item Convergence duration: stop if the predicted ranking stays the same for a number of sampled instances equal to \SI{1}{\%} or \SI{2}{\%} of all instances.
-  \end{itemize}
-
-  \item Wilcoxon stopping criterion
-  \begin{itemize}
-    \item Minimum amount: sample at least \SI{2}{\%}, \SI{8}{\%}, \SI{10}{\%}, or \SI{12}{\%} of instances before applying the criterion.
-    \item Average of $p$-values to drop below: \SI{5}{\%}.
-    \item Exponential-moving average: incorporate previous significance values by using an EMA with $\beta = 0.1$ or $\beta = 0.7$.
-  \end{itemize}
-\end{itemize}
-
-\subsubsection{Runtime-label prediction model.}
-
-As runtime-label prediction model, we only make use of one fixed model since an exhaustive grid search would be infeasible (cf.~Section~\ref{sec:eval:prediction}).
-Our model of choice is an ensemble, stacking a quadratic-discriminant analysis\footnote{With a regularization parameter of zero and a singular value threshold of zero} onto a random forest model\footnote{With \emph{entropy} splitting criterion and \emph{balanced} class weights} using a decision tree\footnote{With \emph{Gini} impurity splitting criterion; choosing the best split for each branching decision and maximum tree depth of 5 levels} to weight its members.
-If not stated explicitly, all other model parameters default to the presets of \emph{scikit-learn}\footnote{\url{https://scikit-learn.org/stable/index.html}; Version 1.0.2}~\cite{scikit-learn}.
-
-%the proposed runtime label ranking, (2) the PAR-2 subsample ranking, and (3) the runtime label ranking that uses the PAR-2 subsample ranking as a fallback when performance differences drop below a threshold.
-%Possible values are \SI{0.01}{}, \SI{0.05}{}, and \SI{0.1}{}.
-%Also, we experiment with different lengths of the prediction history that is used for stabilizing the ranking.
-%Possible values are 10, 20, 30, 40, and 50.
-
 \subsection{Implementation Details}
@@
 
-Before evaluating our approach end-to-end as described in Section~\ref{sec:exdesign}, we discuss the choice of the machine-learning predictor and the runtime-discretization method.
-Thereafter, we look at the results of our active-learning approach.
-Finally, we also analyze the importance of different instance families to our framework.
-
-\subsection{Runtime Prediction}
-\label{sec:eval:prediction}
-
-%JB: we should probably remove this section here and jump right into active learning;
-% - results of classifier comparison can shortly be described in Section 4 (new subsection) or 3.2 ("Solver Model", which currently mainly describes runtime discretization); in any case, classifier should be discussed earlier
-% - in current version, we already state in Section 3.2 that we use hierarchical clustering, so description of algorithm and some "performed well in preliminary experiments" could go there
-An exhaustive grid search of all active-learning approaches outlined in Section~\ref{sec:main} and different runtime-prediction models is infeasible.
-Therefore, we compare predictors first and select one runtime-prediction model for all further experiments.
-We evaluate predictors with the \emph{Matthews Correlation Coefficient} (MCC) scores~\cite{gorodkin2004comparing,matthews1975comparison} of the classification task.
-MCC scores are great in dealing with class imbalances in contrast to conventional metrics like accuracy.
-
-% \begin{table}
-%   \centering
-%   \caption{
-%   	Time-out prediction Matthews Correlation Coefficient performances.
-%   	Results are obtained using cross-validation.
-%   	We provide \SI{64}{\%} of ground-truth for training, use \SI{16}{\%} as a validation set, and \SI{20}{\%} as a test set (5-fold cross-validation).
-%   }
-%   \label{tab:mcc}
-%   \begin{tabular}{lc}
-%     Prediction Model & Avg. MCC (Std.) \\
-%     \hline
-%     Stacking ensemble (QDA, RF)           & 0.9527 (0.0292) \\
-%     Quadratic-discriminant analysis (QDA) & 0.9290 (0.0339) \\
-%     Random forest (RF)                    & 0.8530 (0.0479) \\
-%     AdaBoost with decision trees          & 0.8384 (0.0444) \\
-%     AdaBoost with logistic regression     & 0.8158 (0.0985) \\
-%     Decision tree                         & 0.8059 (0.0707) \\
-%     Logistic regression                   & 0.8052 (0.1018) \\
-%     kNN classifier                        & 0.7885 (0.1521) \\
-%     MLP classifier                        & 0.7760 (0.1408) \\
-%     Support-vector classifier             & 0.7757 (0.2149) \\
-%     Naive Bayes                           & 0.7306 (0.1394) \\
-%     \hline
-%   \end{tabular}
-% \end{table}
-
-% Stacking MCC
-% 0.888221577313471
-% 0.04753979056930393
-
-%To obtain the results in Table~\ref{tab:mcc},
-We perform cross-validation among all solvers: we assume that all other solver runtimes are known and provide \SI{64}{\%} of ground-truth runtime labels of the target solver for training, use \SI{16}{\%} as a validation set for hyper-parameter tuning of the runtime predictor, and \SI{20}{\%} as a test set (5-fold cross-validation).
-This is repeated for each solver once.
-Note that we perform cross-validation on two levels, i.e., among the solvers and the target solver runtimes.
-We compare the average performances on the test sets.
-
-ML classification models that we have tried include random forests, decision trees, quadratic-discriminant analysis, AdaBoost, logistic regression, $k$-nearest neighbor, multi-layer perceptrons, support vector machines, and naive Bayes.
-A stacking ensemble consisting of quadratic-discriminant analysis and random forests performed best.
-It produces an MCC score of \SI{0.8882}{} for predicting three-class runtime labels.
-% \SI{0.9527}{} for time-out prediction
-The results are significantly better than any other model that we have tried regarding a Wilcoxon signed-rank test with $\alpha = \SI{5}{\%}$.
-
-%Since the differences in the MCC-performance distributions of the best and second-best approaches are statistically significant regarding a Wilcoxon signed-rank test with $\alpha = \SI{5}{\%}$, we only use the best approach for all further experiments as a prediction model.
-
-For runtime discretization, we use hierarchical clustering with $k = 3$ and a log-single-link criterion for all further experiments.
-Each non-time-out runtime starts in a separate interval.
-We then gradually merge intervals whose single-link logarithmic distance is the smallest until the desired number of partitions is reached.
-As already mentioned in Section~\ref{sec:main:model}, this approach has good properties regarding the predictiveness of the PAR-2 score ranking and its discriminatory power.
-Also, experiments showed that our stacking ensemble performed particularly well for this kind of discretization method.
-Other clustering approaches that we have tried include hierarchical clustering with mean-, median- and complete-link criteria, as well as $k$-means and spectral clustering.
-%Predicting runtime labels in a similar experiment as above produced an MCC score of on average \SI{88.82}{\%} ($\sigma = \SI{4.75}{\%}$; in comparison to \SI{95.27}{\%} for predicting time-outs).
-%JB: "similar experiment as above" - reference unclear; what do you want to say?
-%TF: Removed the sentence
-
-\subsection{Active-Learning Results}
-
-\subsubsection{Tuning our algorithm.}
+In this section, we evaluate our active-learning framework.
+First, we analyze and tune the different components of our framework on the tuning dataset.
+Next, we evaluate the best configuration with the full dataset.
+Finally, we analyze the importance of different instance families to our framework.
+
+\subsection{Hyper-Parameter Analysis}
 
@@
 
-\subsubsection{Full dataset evaluation.}
+\subsection{Full-Dataset Evaluation}
 

### 2022-10-11 paper/paper.tex
Slightly revise Algorithms 1 & 2
Review changes introduced in commits 78aaa40879 and 09d8648cd8.

@@
   ~\\[1em]
-  \caption{Comparison of features of our benchmark selection approach, the static benchmark-selection approach by Hoos~et~al.~\cite{HoosKSS13}, the algorithm configuration system SMAC~\cite{HutterHL11}, and the active-learning approaches by Matricon et al.~\cite{MatriconAFSH21}.
+  \caption{Comparison of features of our benchmark-selection approach, the static benchmark-selection approach by Hoos~et~al.~\cite{HoosKSS13}, the algorithm configuration system SMAC~\cite{HutterHL11}, and the active-learning approaches by Matricon et al.~\cite{MatriconAFSH21}.
   }
@@
 The outline of our benchmarking framework is depicted in Algorithm~\ref{algALBenchmark}. 
-Given a set of known solvers $\mathcal{A}$, instances $\mathcal{I}$ and runtimes $r$, we first create a prediction model~$\mathcal{M}$ for the new solver $\hat a$ (Line~1). 
-The prediction model is used to repeatedly select an instance (Line~4) for benchmarking $\hat a$ (Line~5). 
+Given a set of known solvers $\mathcal{A}$, instances $\mathcal{I}$ and runtimes $r$, we first initialize a prediction model~$\mathcal{M}$ for the new solver $\hat a$ (Line~1). 
+The prediction model~$\mathcal{M}$ is used to repeatedly select an instance (Line~4) for benchmarking $\hat a$ (Line~5). 
 The acquired result is subsequently used to update the prediction model~$\mathcal{M}$ (Line~6). 
 When the stopping criterion is met (Line~3), we quit the benchmarking loop and predict the final score of $\hat a$ (Line~8). 
-Algorithm~\ref{algALBenchmark} returns the predicted score of $\hat a$ as well as the acquired runtime measurements (Line~9).
-
-Section~\ref{sec:main:model} specifies the details of the underlying prediction model and describes how we may derive a solver ranking from that model.
-We discuss possible criteria for selecting instances in Section~\ref{sec:main:selection}.
+Algorithm~\ref{algALBenchmark} returns the predicted score of $\hat a$ as well as the acquired instances and runtime measurements (Line~9).
+
+Section~\ref{sec:main:model} describes the underlying prediction model~$\mathcal{M}$ and specifies how we may derive a solver ranking from it.
+We discuss criteria for selecting instances in Section~\ref{sec:main:selection}.
 Section~\ref{sec:main:stopping} concludes with possible stopping conditions.
@@
 
-  $\mathcal{M} \leftarrow \operatorname{initModel}(\mathcal{A}, \mathcal{I}, r, \hat{a})$ \;
+  $\mathcal{M} \leftarrow \operatorname{initModel}(\mathcal{A}, \mathcal{I}, r, \hat{a})$ \tcp*{cf. Section~\ref{sec:main:model}}
 
@@
   %\KwIn{A pool of instances $\mathcal{I}$, a set of solvers $\mathcal{S}$ with known runtimes $r\!: \mathcal{S} \times \mathcal{I} \rightarrow \left[0, \tau\right] \cup \left\lbrace \perp \right\rbrace$, and PAR-2 scores $par2\!: \mathcal{S} \rightarrow \mathbb{R}$ for each solver.}
-  \KwIn{Solvers $\mathcal{A}$, Instaces $\mathcal{I}$, Runtimes $r : \mathcal{A} \times \mathcal{I} \rightarrow [0, \tau]$}
+  \KwIn{Solvers $\mathcal{A}$, Instances $\mathcal{I}$, Runtimes $r : \mathcal{A} \times \mathcal{I} \rightarrow [0, \tau]$}
   \KwOut{Average Rank Accuracy $\bar{O}_{\operatorname{acc}}$, Average Fraction of Runtime $\bar{O}_{\operatorname{rt}}$}
@@
     	}
+    	%JB: why is "\hat a" subscript of s rather than argument?
+    	%JB: "par_2" needs to be defined in text
     }

### 2022-10-12 paper/paper.tex
Move two figures forward in text
No changes to the figures themselves.
Process feedback

@@
 \usepackage{subcaption}
-\usepackage{tablefootnote}
 \usepackage{tabularx}
@@
   This also applies to solutions of the SAT (propositional satisfiability) problem.
-  While the field of SAT solver benchmarking is well established, traditional benchmark selection approaches do not optimize for benchmark runtime.
   Benchmark selection chooses representative instances from a pool of instances such that they reliably discriminate SAT solvers based on their runtime.
@@
 This is also true for the pro\-po\-si\-tio\-nal sat\-is\-fia\-bi\-li\-ty problem (SAT), the archetypal $\mathcal{NP}$-complete problem.
-While state-of-the-art solvers embark upon more and more problem instances, SAT solver evaluation is steadily getting more costly.
-%JB: not sure about the "while" here -- is the message that (1) evaluation gets costly *since* solvers have to solve more instances or is it that (2) solvers are threoretically able to solve more instances, but evaluation gets costlier anyway
+Benchmarking is, however, quite expensive regarding the cost of experiments.
 Competitive SAT solvers are the result of extensive experimentation and a variety of ideas and considerations~\cite{FroleyksHIJS21,sat2022}.
@@
 
+% TODO: insert URLs
 All our source code\footnote{temporary, anonymized version for review: \url{xxx}} and and data\footnote{temporary, anonymized version for review: \url{xxx}} are available on GitHub.
@@
 In such competitions, the organizers define the rules for composing the corresponding benchmarks.
+These selection strategies are mostly static, i.e., they do not depend on particular solvers to distinguish.
 Balint et al. provide an overview of benchmark-selection criteria in different solver competitions~\cite{balint2015overview}.
-Froleyks et al. describe benchmark-selection for the SAT~Competition~\cite{FroleyksHIJS21}.
+Froleyks et al. describe benchmark selection for the SAT~Competition~\cite{FroleyksHIJS21}.
 Manthey and Möhle find that competition benchmarks might contain redundant instances, and propose a feature-based approach to remove redundancy~\cite{manthey2016better}.
 M{\i}s{\i}r presents a feature-based approach to reduce benchmarks by matrix factorization and clustering~\cite{misir2021benchmark}.
-
-Hoos et al. discuss which properties are most desirable when selecting SAT benchmark instances~\cite{HoosKSS13}.
-The selection strategy presented is static, i.e., it does not depend on particular solvers to distinguish.
+Hoos et al.~\cite{HoosKSS13} discuss which properties are most desirable when selecting SAT benchmark instances.
 Selection criteria are instance variety to avoid over-fitting, adapted instance hardness (not too easy but also not too hard), and avoiding duplicate instances.
 To filter too similar instances, they use a distance-based approach with the SATzilla features~\cite{XuHHL08,features}.
-Their approach \emph{ranks} solvers, is \emph{feature-based}, and \emph{scalable} in the sense that the evaluation time scales linearly with more instances and solvers.
-Also, it allows \emph{scoring} and thereby comparing solvers.
-However, the approach does not optimize for benchmark \emph{runtime} and selects instances \emph{randomly}, apart from constraints on the instance hardness and feature distance.
+The approach does, however, not optimize for benchmark \emph{runtime} and selects instances \emph{randomly}, apart from constraints on the instance hardness and feature distance.
 
@@
     \hline
-    Ranking & \cmark & \xmark & \cmark & \cmark \\
+    Ranking/Scoring & \cmark & \xmark & (\cmark) & \cmark \\
     Runtime Minimization & \xmark & \cmark & \cmark & \cmark \\
-    % Feature-based & \cmark & \cmark & \cmark & \cmark \\
     Incremental/Non-Random & \xmark & \xmark & \cmark & \cmark \\
-    % \tablefootnote{While using a model-based selection strategy to select the next configuration to evaluate, the instance selection is done randomly.}
-    Scoring & \cmark & \xmark & \xmark & \cmark \\
-    % \tablefootnote{While they provide a confidence measure, the strategy's output is still only binary, i.e., whether a solver A is better or worse than solver B.}
     Scalability & \cmark & \cmark & \xmark & \cmark \\
@@
 Further related work can be found within the field of algorithm configuration~\cite{HoosHL21,Stutzle0P22}, e.g., the configuration system SMAC~\cite{HutterHL11}.
-There, the goal is to tune SAT solvers for a given sub-domain of problem instances.
+Thereby, the goal is to tune SAT solvers for a given sub-domain of problem instances.
 Although this task is different from our goal, e.g., we do not need to navigate configuration space, there are similarities to our approach as well.
 For example, SMAC also employs an iterative, model-based selection procedure, though for configurations rather than instances.
-SMAC's configuration sampling has also proven to be better than random sampling in all tested scenarios and has even proven to be significantly better in most cases.
-While this approach considers the configuration's \emph{runtime}, is \emph{feature-based}, and \emph{scales} by having a fixed time budget, it does not satisfy the other three criteria listed in Table~\ref{tab:requirements}.
-First, an algorithm configurator cannot be used to \emph{rank} a new solver since it only looks at promising solvers/configurations rather than the overall average performance.
-Second, while using a model-based selection strategy to sample configurations, instance selection is done \emph{randomly}, i.e., without building a model over instances.
-And third, only few configurations within their approach are \emph{scored} since algorithm configuration seeks to find solemnly the best performing configuration.
+An algorithm configurator, however, cannot be used to \emph{rank/score} a new solver since algorithm configuration seeks to find solemnly the best performing configuration.
+Also, while using a model-based selection strategy to sample configurations, instance selection is done \emph{randomly}, i.e., without building a model over instances.
 
@@
 They calculate a scoring metric for all unselected instances, run the experiment with the highest score, and update the confidence.
-Most of the proposed strategies in~\cite{MatriconAFSH21} are instance-based, i.e., no model of the solver runtimes is built.
-Regarding our desired criteria of benchmark selection (Table~\ref{tab:requirements}), the aforementioned approach \emph{ranks} solvers, optimizes for \emph{runtime}, is \emph{feature-based}, and uses \emph{incremental non-random} instance selection.
+Their approach ticks off most of our desired features.
 However, the approach only compares solvers binarily rather than providing a \emph{scoring}.
@@
 
-% MI: move to approach if needed there:
-%For the scoring metric, they discuss (1) a random strategy as a baseline, (2) a discrimination strategy based on the work of Gent et al.~\cite{GentHJKMNN14}, (3) a variance-based strategy (select instance with the highest variance in its distribution of running times), (4) an information-gain-based method, and (5) a feature-distance-based strategy.
-%To update the confidence level after each experiment, they discuss using (1) a fixed subset size, (2) a Wilcoxon test, and (3) a distribution-based method.
-%Their experiments show that random instance selection with a Wilcoxon test as the stopping criterion performs consistently well across different data sets in the sense that it belongs to the Pareto front or is close to it in all cases.
-%JB: in our evaluation, we should state whether our results are consistent with this or not
-%TF: will do
-
 \subsubsection{Active Learning (AL).}
-
-%\begin{figure}[tbp!]
-%  \centering
-%  \begin{tabular}[c]{ccc}
-%  \begin{subfigure}[b]{0.25\textwidth}
-%  \centering
-%  \resizebox{!}{2.75cm}{
-%  \begin{tikzpicture}
-%  \draw[fill=black]  (-2.8,3.7) ellipse (0.1 and 0.1);
-%  \draw[fill=black]  (-2.6,3.4) arc (0:180:0.2);
-%  \node[right] at (-5.9,3.6) {\large Agent (Solver)};
-%  \draw[fill=white]  (-3.3,2.8) rectangle (-2.5,2.2);
-%  \draw[fill=white]  (-3.2,2.7) rectangle (-2.4,2.1);
-%  \draw[fill=white]  (-3.1,2.6) rectangle (-2.3,2);
-%  \node[right] at (-5.9,2.4) {\large Train. Data};
-%  \node (v1) at (-2.8,3.4) {};
-%  \node (v2) at (-2.8,2.8) {};
-%  \draw[->]  (v1) edge (v2);
-%  \node[right] at (-2.8,3.1) {\large $x, y$};
-%  \node (v3) at (-2.8,2) {};
-%  \node (v4) at (-2.8,1.4) {};
-%  \draw[->]  (v3) edge (v4);
-%  \draw  (-3.1,1.4) rectangle (-2.5,1);
-%  \draw  (-2.4,3.2) rectangle (-2.4,3.2);
-%  \node[right] at (-5.9,1.2) {\large Learning Alg.};
-%  \node (v5) at (-2.8,1) {};
-%  \node (v6) at (-2.8,0.4) {};
-%  \draw[->]  (v5) edge (v6);
-%  \node[right] at (-5.9,0.25) {\large Approx. Func.};
-%  \node at (-2.8,0.25) {\large $f$};
-%  \end{tikzpicture}
-%  }
-%  \caption{Passive}
-%  \label{fig:passive}
-%  \end{subfigure}
-%  &
-%  \begin{subfigure}[b]{0.35\textwidth}
-%  \centering
-%  \resizebox{!}{2.75cm}{
-%  \begin{tikzpicture}
-%  \draw[fill=black]  (-2.8,3.7) ellipse (0.1 and 0.1);
-%  \draw[fill=black]  (-2.6,3.4) arc (0:180:0.2);
-%  \node[right] at (-5.9,3.6) {\large Agent (Solver)};
-%  \draw[fill=white]  (-3.3,2.8) rectangle (-2.5,2.2);
-%  \draw[fill=white]  (-3.2,2.7) rectangle (-2.4,2.1);
-%  \draw[fill=white]  (-3.1,2.6) rectangle (-2.3,2);
-%  \node[right] at (-5.9,2.4) {\large Train. Data};
-%  \node (v1) at (-2.8,3.4) {};
-%  \node (v2) at (-2.8,2.8) {};
-%  \draw[->]  (v1) edge (v2);
-%  \node[right] at (-2.8,3.1) {\large $x, y$};
-%  \node (v3) at (-2.8,2) {};
-%  \node (v4) at (-2.8,1.4) {};
-%  \draw[->]  (v3) edge (v4);
-%  \draw  (-3.1,1.4) rectangle (-2.5,1);
-%  \draw  (-2.4,3.2) rectangle (-2.4,3.2);
-%  \node[right] at (-5.9,1.2) {\large Learning Alg.};
-%  \node (v5) at (-2.8,1) {};
-%  \node (v6) at (-2.8,0.4) {};
-%  \draw[->]  (v5) edge (v6);
-%  \node[right] at (-5.9,0.25) {\large Approx. Func.};
-%  \node at (-2.8,0.25) {\large $f$};
-%  \node[right] at (-1.2,2.4) {\large Pool};
-%  \draw[->, densely dashed] (-2.4,1.2) arc (-90:0:0.8);
-%  \draw[->, densely dashed] (-1.6,2.8) arc (0:90:0.8);
-%  \node[right] at (-1.6,3.135) {\large $x, ?$};
-%  \draw  (-1.6,2.4) ellipse (0.4 and 0.3);
-%  \draw[fill=black]  (-1.725,2.5) ellipse (0.07 and 0.07);
-%  \draw[fill=black]  (-1.625,2.25) ellipse (0.07 and 0.07);
-%  \draw[fill=black]  (-1.425,2.4) ellipse (0.07 and 0.07);
-%  \end{tikzpicture}
-%  }
-%  \caption{Active Pool-based}
-%  \label{fig:activepool}
-%  \end{subfigure}
-%  &
-%  \begin{subfigure}[b]{0.35\textwidth}
-%  \centering
-%  \resizebox{!}{2.75cm}{
-%  \begin{tikzpicture}
-%  \draw[fill=black]  (-2.8,3.7) ellipse (0.1 and 0.1);
-%  \draw[fill=black]  (-2.6,3.4) arc (0:180:0.2);
-%  \node[right] at (-5.9,3.6) {\large Agent (Solver)};
-%  \draw[fill=white]  (-3.3,2.8) rectangle (-2.5,2.2);
-%  \draw[fill=white]  (-3.2,2.7) rectangle (-2.4,2.1);
-%  \draw[fill=white]  (-3.1,2.6) rectangle (-2.3,2);
-%  \node[right] at (-5.9,2.4) {\large Train. Data};
-%  \node (v1) at (-2.8,3.4) {};
-%  \node (v2) at (-2.8,2.8) {};
-%  \draw[->]  (v1) edge (v2);
-%  \node[right] at (-2.8,3.1) {\large $x, y$};
-%  \node (v3) at (-2.8,2) {};
-%  \node (v4) at (-2.8,1.4) {};
-%  \draw[->]  (v3) edge (v4);
-%  \draw  (-3.1,1.4) rectangle (-2.5,1);
-%  \draw  (-2.4,3.2) rectangle (-2.4,3.2);
-%  \node[right] at (-5.9,1.2) {\large Learning Alg.};
-%  \node (v5) at (-2.8,1) {};
-%  \node (v6) at (-2.8,0.4) {};
-%  \draw[->]  (v5) edge (v6);
-%  \node[right] at (-5.9,0.25) {\large Approx. Func.};
-%  \node at (-2.8,0.25) {\large $f$};
-%  \node[right, label={[align=left]\large Gen.\\\large Model}] at (-0.6,1.85) {};
-%  \node[right] at (-1.3,3.135) {\large $x, ?$};
-%  \draw[->, densely dashed] (-2.4,1.2) arc (-90:90:1.2);
-%  \end{tikzpicture}
-%  }
-%  \caption{Active Synthesis-based}
-%  \label{fig:activesynth}
-%  \end{subfigure}
-%  \end{tabular}
-%  
-%  \caption{Types of Learning (Inspired by Rubens~et.al.~\cite{RubensESK15})}
-%  \label{fig:learning}
-%\end{figure}
 
@@
 In contrast to traditional passive machine-learning methods with given instance labels, active learning allows for selecting instances for which to acquire labels.
-AL algorithms can be categorized into \emph{synthesis-bas\-ed} \cite{0001AEMN22,GarzonMG22,2019gaal} and \emph{pool-bas\-ed} approaches \cite{GolbandiKL11,HarpaleY08,KorenBV09}.%SinhaED19,distribAL,TongK01,KapoorGUD07,KornerW06,MelvilleM04}.
+AL algorithms can be categorized into \emph{synthesis-bas\-ed} \cite{0001AEMN22,GarzonMG22,2019gaal} and \emph{pool-bas\-ed} approaches \cite{GolbandiKL11,HarpaleY08,KorenBV09}.
 While synthesis-based methods generate instances for labeling, pool-based methods rely on a fixed set of unlabeled instances from which to sample.
@@
 
-% MI: here we should limit this to the methods needed for presenting related work for AL in benchmarking:
-%JB: should probably just put the citation there where we explain the strategies (in Section 3)
-%Pool-based methods can be further refined into \emph{Error-based} \cite{GolbandiKL11,KorenBV09}, \emph{Un\-cer\-tain\-ty-based} \cite{HarpaleY08,KapoorGUD07,TongK01}, \emph{Distribution-based} methods \cite{distribAL,SinhaED19} and \emph{Ensembles} \cite{KornerW06,MelvilleM04}.
-%Error-based strategies pick instances that have the most influence on the model parameters.
-%Uncertainty-based methods sample instances close to their perceived decision boundary.
-%Distribution-based approaches sample instances whose features are unaccounted for amongst the already sampled instances.
-%Finally, ensembles are a meta-strategy that combines the predictions of several different active-learning models into a selection choice.
-
 \section{Active Learning for SAT Solver Benchmarking}
@@
 The prediction model~$\mathcal{M}$ is used to repeatedly select an instance (Line~4) for benchmarking $\hat a$ (Line~5). 
-The acquired result is subsequently used to update the prediction model~$\mathcal{M}$ (Line~6). 
+The acquired result is subsequently used to update the prediction model~$\mathcal{M}$ (Line~7). 
 When the stopping criterion is met (Line~3), we quit the benchmarking loop and predict the final score of $\hat a$ (Line~8). 
@@
 
-% MI: This belongs to discussion at the end since this also has substantial disadvantages:
-Since the potential runtime of experiments is by magnitudes larger than the model's update time, we only consider incrementing our benchmark by one instance at a time rather than using batches, which is also proposed in current active-learning advances~\cite{SinhaED19,2019gaal}.
-
 \begin{algorithm}
-% MI: comment in to rename to framework and remove numbering:
-%\SetAlgorithmName{Framework}{framework}{List of Frameworks}
-%\renewcommand{\thealgocf}{}
   \caption{Incremental Benchmarking Framework}
@@
 
-  %\KwIn{Instances $\mathcal{I}$, Solvers $\mathcal{A}$, Runtimes $r\!: \mathcal{A} \times \mathcal{I} \rightarrow \left[0, \tau\right]$}
   \KwIn{Solvers $\mathcal{A}$, Instances $\mathcal{I}$, Runtimes $r : \mathcal{A} \times \mathcal{I} \rightarrow [0, \tau]$, Solver $\hat{a}$}
@@
 
-  $\mathcal{M} \leftarrow \operatorname{initModel}(\mathcal{A}, \mathcal{I}, r, \hat{a})$ \tcp*{cf. Section~\ref{sec:main:model}}
+  $\mathcal{M} \leftarrow \operatorname{initModel}\left(\mathcal{A},\, \mathcal{I},\, r,\, \hat{a}\right)$ \tcp*{cf. Section~\ref{sec:main:model}}
 
@@
   $\mathcal{R} \leftarrow \emptyset$ \;
-  \While(\tcp*[f]{cf. Section~\ref{sec:main:stopping}}){$\operatorname{not} \operatorname{stop}(\mathcal{M})$}{
-    $e \leftarrow \operatorname{selectNextInstance}\!\left(\mathcal{M}\right)$ \tcp*{cf. Section~\ref{sec:main:selection}}
+  \While(\tcp*[f]{cf. Section~\ref{sec:main:stopping}}){$\operatorname{not} \operatorname{stop}\left(\mathcal{M}\right)$}{
+    $e \leftarrow \operatorname{selectNextInstance}\left(\mathcal{M}\right)$ \tcp*{cf. Section~\ref{sec:main:selection}}
 
-    $t \leftarrow \operatorname{runExperiment}\!\left(\hat{a},\,  e\right)$  \tcp*{Runs $\hat{a}$ on $e$ with time-out $\tau$}
+    $t \leftarrow \operatorname{runExperiment}\left(\hat{a},\,  e\right)$  \tcp*{Runs $\hat{a}$ on $e$ with time-out $\tau$}
 
-    $\operatorname{updateModel}\bigr(\mathcal{M}, (e, t)\bigl)$ \tcp*{cf. Section~\ref{sec:main:model}}
+    $\mathcal{R} \leftarrow \mathcal{R} \cup \left\lbrace (e,\, t) \right\rbrace$
 
     \BlankLine
-    $\mathcal{R} \leftarrow \mathcal{R} \cup \bigl\{ (e, t) \bigr\}$
+    $\operatorname{updateModel}\left(\mathcal{M},\, \mathcal{R}\right)$ \tcp*{cf. Section~\ref{sec:main:model}}
   }
@@
   \Return $(s_{\hat a}, \mathcal{R})$
-
-  %$\operatorname{score}\!\left(\right) \leftarrow \operatorname{predictScores}\!\left(\mathcal{M}\right)$
-  %\KwOut{Return tuple of scoring function $\operatorname{score}\!: \mathcal{A} \cup \left\lbrace \hat{a} \right\rbrace \rightarrow \mathbb{R}$ and the selected instances $\tilde{\mathcal{I}}$.}
 \end{algorithm}
@@
 
-% MI: describe initialization and update of model
-
 Let benchmark instances $\mathcal{I}$, SAT solvers $\mathcal{A}$, and runtimes $r : \mathcal{A} \times \mathcal{I} \rightarrow \left[0, \tau\right]$ be given. 
@@
 
+First, we initialize the model $\mathcal{M}$ in Line~1.
+This creates the machine-learning prediction model and also transforms runtimes as discussed in the subsequent section.
+We do not need any further steps here, since we have not obtained any labeled samples yet, so cannot train anything.
+The model update (Line~7) uses all acquired runtimes to train a new runtime-prediction function $f$ using common supervised machine-learning methodology.
+It has also access to global instance features that assign each instance a vector in $\mathbb{R}^n$.
+Note that we build a new prediction model in each iteration since running experiments (Line~5) dominates the runtime of Algorithm~\ref{algALBenchmark} by magnitudes.
+Finally, we determine the solver scores that are predicted in Line~8 by using the runtime-prediction function $f$ and the approach described in the following two sections.
+
 \subsubsection{Runtime Transformation.}
@@
 We transform the runtimes into discrete runtime labels on a per-instance basis.
-For each instance $e \in \mathcal{I}$, we use clustering algorithm to assign the runtimes in $\bigl\{ r(a, e) \mid a \in A \bigr\}$ to one of the $k$ clusters $C_1, \dots, C_k$ such that the fastest runtimes for instance $e$ are in cluster $C_1$ and the slowest are in cluster $C_k$.
-For instances with many timeouts, we have to manually fix the clustering results such that all timeouts $\tau$ are in cluster $C_k$.
-%TF: maybe a bit unclear. We do not manually fix anything. There is always one cluster with only time-outs if there are any.
-%JB: was wondering about that as well; how do we exactly create the timeout cluster?
-do we only cluster with k-1, removing timeout instances, or is the timeout cluster an ordinary cluster (which might also contain on-timeout instances)?
+For each instance $e \in \mathcal{I}$, we use a clustering algorithm to assign the runtimes in $\bigl\{ r(a, e) \mid a \in A \bigr\}$ to one of $k$ clusters $C_1, \dots, C_k$ such that the fastest runtimes for instance $e$ are in cluster $C_1$ and the slowest are in cluster $C_{k-1}$, excluding timeouts $\tau$.
+Those are always in cluster $C_{k}$.
 The runtime transformation function $\gamma_k : {\mathcal{A} \times \mathcal{I}} \rightarrow \left\lbrace 1, \dots, k \right\rbrace$ is then specified as follows:
@@
 $$\gamma_k(a, e) = j ~\Leftrightarrow~ r(a, e) \in C_j$$
+%
+Given an instance $e \in \mathcal{I}$, a solver $a \in A$ belongs to the $\gamma_k(a, e)$-fastest solvers on instance $e$. 
+In preliminary experiments, we achieved higher accuracy in predicting such discrete-runtime labels than in predicting raw runtimes.
+Research on portfolio solvers has also shown that discretization works well in practice~\cite{CollauttiMMO13,NgokoCT19}.
 
-Given an instance $e \in \mathcal{I}$, a solver $a \in A$ belongs to the $\gamma_k(a, e)$-fastest solvers on instance $e$. 
-In preliminary experiments, we achieved higher accuracy predicting such discrete runtime labels than predicting raw runtimes. 
-The mean squared error of raw runtime regression was within the same magnitude as the values to be predicted. 
-%MI: how much more accurate are we with discrete runtime labels
-Research on portfolio solvers has also shown that directly predicting runtimes without any transformation does not work well in practice~\cite{CollauttiMMO13,NgokoCT19}.
-%JB: maybe we should phrase this more defensively or just leave out this argument (might be dataset-dependend), (1) e.g., SATzilla 2008 does predict runtimes and (2) we could also have treid to predict normalized runtimes like Collautti et al.
-instead of discretizing (though prediction performance would probably still have been bad in our case)
+\subsubsection{Ranking Solvers.}
 
@@
 %
-I.e., we apply a PAR-2 scoring, which is commonly used in SAT competitions~\cite{FroleyksHIJS21}, on the discrete labels.
+I.e., we apply PAR-2 scoring, which is commonly used in SAT competitions~\cite{FroleyksHIJS21}, on the discrete labels.
 The scoring function $s_k$ induces a ranking among solvers.
@@
 
-Selecting an instance based on the model is a core functionality of our framework (cf. Algorithm~\ref{algALBenchmark}, Line 2).
+Selecting an instance based on the model is a core functionality of our framework (cf. Algorithm~\ref{algALBenchmark}, Line~2).
 In this section, we introduce our sampling strategies, which make use of the model's label-prediction function $f$ and are inspired by existing work within the realms of active learning~\cite{settles2009active}.
@@
 The uncertainty sampling strategy simply selects the instance that is closest to the model's decision boundary.
+The set $\tilde{\mathcal{I}} \subseteq \mathcal{I}$ denotes the instances that have already been sampled.
 %
@@
   \underset{e \in \mathcal{I} \setminus \tilde{\mathcal{I}}}{\arg\min} \left\lvert \frac{1}{k} - \max_{n \in \left\lbrace 1, \dots, k \right\rbrace} f'\!\left(\hat{a}, e\right)_{n} \right\rvert
-  %JB: \tilde I is undefined
 \end{equation*}
@@
 The information-gain sampling strategy selects the instance with the highest expected entropy reduction regarding the runtime labels of the instance.
-To be more specific, we select the instance $e$ that maximizes $IG(e)$, which is specified as follows:
+To be more specific, we select the instance $e \in \mathcal{I} \setminus \tilde{\mathcal{I}}$ that maximizes $IG(e)$, which is specified as follows:
 %
@@
 \end{equation*}
-
+%
 Here, $\operatorname{H}(e)$ denotes the entropy of the runtime labels $\gamma(a, e)$ over all $a \in \mathcal{A}$ and $\operatorname{H}(e, n)$ denotes the entropy of these labels plus $n$ as the runtime label for $\hat{a}$.
 The term $\operatorname{\hat H}_n(e)$ is computed for every possible runtime label $n \in \{1, \dots, k\}$.
-
-%JB: if this is info gain regarding runtimes, doesn't Eq.
-(4) mean that we would prefer instances with low entropy of runtime distribution after considering the new solver, i.e., instances that are not very discriminatory?
-%TF: I don't think so. We want to maximize the reduction in entropy of the runtime labels. Low entropy afterward -> we know that the solver behaves similarly to existing ones.
-%JB: but if we want to discriminate between solvers, what does it help to know that solver behaves similarly to other solvers?
-
-%Finally, the voting-based committee-disagreement approach requires a committee of $c$ models.
-%Rather than having a single prediction function $f$, there is a separate function $f_j$ for all committee members $j \in \left\lbrace 1, \dots, c  \right\rbrace$.
-%We then select the instance $i$ whose entropy of runtime-label votes by the different ensemble members is the highest.
-
-% MI: The following belongs either to Related Work or to Experimental Design (commented out)
-%\subsubsection{Baselines and existing approaches.}
-%\label{sec:sampling1}
-%Apart from obvious baseline instance-selection strategies, i.e., random sampling and neighborhood-aware sampling (randomly choosing among the instances with the highest amount of non-sampled $k$ neighbors), we also implement existing instance-selection strategies.
-
-%Bossek and Wagner~\cite{Bossek021a} introduce a custom ranking-score metric incorporating the number of \emph{good} pairs and \emph{bad} pairs.
-%A good solver pair, regarding a particular instance, are solvers that perform according to their global ranking.
-%Otherwise, it is a \emph{bad} pair.
-%Based on this classification, they assign a custom score to each instance.
-%By selecting instances with the highest score, they try to favor instances that are predictive of the final ranking.
-
-%Gent~et~al.~\cite{GentHJKMNN14} make use of a concept known from racing for automated model selection.
-%Given a particular instance and a parameter $\rho$, a solver is $\rho$-dominated if it is at least $\rho$ times worse than the best solver on this problem instance.
-%By selecting instances with the highest amount of $\rho$-dominated solvers, they hope to sample instances that provide the best discriminatory power to distinguish the runtimes of a new contestant.
-
-%Finally, we also implement some of the selection approaches described by Matricon~et.al.~\cite{MatriconAFSH21}.
-%Besides uniformly random sampling and the aforementioned discrimination-based method by Gent~et~al.~\cite{GentHJKMNN14}, they also describe a variance-based selection approach.
-%The variance-based strategy selects instances with the highest runtime variance per average runtime.
-%We also implement a logarithmized version of this, i.e., the highest log-runtime variance per average log runtime.
+By maximizing information gain, we select instances that identify solvers with similar behavior.
 
@@
 
-%Similar to the instance-selection strategies, we can also divide stopping criteria into baseline and active-learning-based approaches.
-%Also, refer to~Algorithm~\ref{algALBenchmark}, line~3.
-%While a fixed subset-size criterion, e.g., only sample \SI{20}{\%} of all available instances, is an example of the former, a stopping criterion based on a Wilcoxon signed-rank test or the convergence of the model's predictions is an example for the latter.
-
-In this section, we present two dynamic stopping criteria, the Wilcoxon and the Model Uncertainty Stopping Criterion (cf. Algorithm~\ref{algALBenchmark}, Line~1).
+In this section, we present two dynamic stopping criteria, the Wilcoxon and the ranking-stopping criterion (cf. Algorithm~\ref{algALBenchmark}, Line~1).
 
@@
 The Wilcoxon stopping criterion stops the active-learning process when we are certain enough that the predicted runtime labels of the new solver are sufficiently different from the labels of the existing solvers.
+This criterion is loosely inspired by Matricon et.~al.~\cite{MatriconAFSH21}.
 We use the average $p$-value $W_{\hat{a}}$ of a Wilcoxon signed-rank test $w(S,P)$ of the two runtime label distributions $S=\{ \gamma(a, e) \mid e \in \mathcal{I} \}$ for an existing solver $a$ and $P=\{ f(\hat a, e) \mid e \in \mathcal{I} \}$ for the new solver $\hat{a}$:
@@
 
-The ranking stopping criterion is less sophisticated in comparison.
+The ranking-stopping criterion is less sophisticated in comparison.
 It stops the active-learning process if the ranking induced by the model's predictions (Equation~\ref{eq:rankingeq}) remains unchanged within the last $l$ iterations.
@@
 
-Given all the previously presented instantiations for Algorithm~\ref{algALBenchmark}, this section briefly outlines our experimental design, including our optimization goal, evaluation framework, used data sets, hyper-parameter choices, as well as implementation details.
-% hyper-parameter values that we experiment with
-
-\subsection{Optimization Goal}
-\label{sec:exdesign:goal}
-
-As already stated in the introductory section, this work addresses the \emph{New-Solver Problem} (cf.~Definition~\ref{def:new-solver-problem}).
-We create a prediction model $\mathcal{M}$ as described in Section~\ref{sec:main:model} that provides us with a scoring function $s_k$.
-% MI: Let this be the evaluation result:
-%Also, it outputs the fraction of runtime that is needed to arrive at its conclusion.
-First and foremost, our goal is to provide the engineer of new SAT solvers with an accurate ranking.
-We define the \emph{ranking accuracy} $O_{\operatorname{acc}} \in \left[0, 1\right]$ (higher is better) by the fraction of pairs $\left(\hat{a}, a\right)$ for all $a \in \mathcal{A}$ that are decided correctly by the given ranking (cf.~Algorithm~\ref{alg:eval}).
-For now, we exclude the equality of solvers since data shows that the PAR-2 scores of all our solvers are different.
-So, possible ranking decisions may be, for example, solver $a$ is better than $b$ or $b$ is better than $a$.
-The ranking accuracy is affected by whether this decision is correctly made.
-%Matricon et. al.~\cite{MatriconAFSH21} also evaluate their approach by the fraction of solver pairs that is correctly decided.
-Second, we also have to optimize for runtime.
-The \emph{fraction of runtime} that the algorithm needs to arrive at its conclusion is denoted by $O_{\operatorname{rt}} \in \left[0, 1\right]$ (lower is better).
-This metric puts the runtime summed over the sampled instances in relation to the runtime summed over all instances in the dataset (cf. Algorithm~\ref{alg:eval}).
-Overall, we want to find an approach that maximizes
-%
-\begin{equation}
-  O_\delta := \delta O_{\operatorname{acc}} + \left(1 - \delta\right) \left(1 - O_{\operatorname{rt}}\right) \enspace \textrm{,}
-  \label{eq:opt}
-\end{equation} 
-%
-whereby $\delta \in \left[0, 1\right]$ allows for linear weighting between the two optimization goals $O_{\operatorname{acc}}$ and $O_{\operatorname{rt}}$.
-Plotting the approaches that maximize $O_\delta$ for all $\delta \in \left[0, 1\right]$ on a $O_{\operatorname{rt}}$-$O_{\operatorname{acc}}$-diagram provides us with a Pareto front of the best approaches for different optimization-goal weightings.
+Given all the previously presented instantiations for Algorithm~\ref{algALBenchmark}, this section briefly outlines our experimental design, including our evaluation framework, used data sets, hyper-parameter choices, as well as implementation details.
 
 \subsection{Evaluation Framework}
-\label{sec:exdesign:evalframe}
+\label{sec:exdesign:eval}
 
-To evaluate a concrete instantiation of Algorithm~\ref{algALBenchmark} (a concrete choice for all the sub-routines), we perform cross-validation on our set of solvers.
-That means that each solver once plays the role of the new solver.
-Algorithm~\ref{alg:eval} shows this.
-Note that the \emph{new} solver in each iteration is excluded from the set of solvers $\mathcal{A}$ to avoid data leakage.
-Also, the runtime function $r$ is restricted to $\mathcal{A} \times \mathcal{I}$.
-
-\begin{algorithm}[htb]
+\begin{algorithm}[tb]
   \caption{Evaluation Framework}
@@
 
-  %\KwIn{A pool of instances $\mathcal{I}$, a set of solvers $\mathcal{S}$ with known runtimes $r\!: \mathcal{S} \times \mathcal{I} \rightarrow \left[0, \tau\right] \cup \left\lbrace \perp \right\rbrace$, and PAR-2 scores $par2\!: \mathcal{S} \rightarrow \mathbb{R}$ for each solver.}
   \KwIn{Solvers $\mathcal{A}$, Instances $\mathcal{I}$, Runtimes $r : \mathcal{A} \times \mathcal{I} \rightarrow [0, \tau]$}
@@
     \For{$a \in \mathcal{A}$}{
-    	\If{$\bigl(s_k(a) - s_{\hat a}\bigr) \cdot \bigl(\operatorname{par_2}(a) - \operatorname{par_2}(\hat a)\bigr) > 0$}{
+      \If{$\bigl(s_k(a) - s_k(\hat a)\bigr) \cdot \bigl(\operatorname{par_2}(a) - \operatorname{par_2}(\hat a)\bigr) > 0$}{
            $O_{\operatorname{acc}} \leftarrow O_{\operatorname{acc}} + \frac{1}{|\mathcal{A}|}$ \;
       }
-    	%JB: why is "\hat a" subscript of s rather than argument?
-    	%JB: "par_2" needs to be defined in text
     }
@@
 
+As already stated in the introductory section, this work addresses the \emph{New-Solver Problem} (cf.~Definition~\ref{def:new-solver-problem}).
+We create a prediction model $\mathcal{M}$ as described in Section~\ref{sec:main:model} that provides us with a scoring function $s_k$.
+
+To evaluate a concrete instantiation of Algorithm~\ref{algALBenchmark} (a concrete choice for all the sub-routines), we perform cross-validation on our set of solvers.
+Algorithm~\ref{alg:eval} shows this.
+That means that each solver once plays the role of the new solver (Line~2).
+Note that the \emph{new} solver in each iteration is excluded from the set of solvers $\mathcal{A}$ to avoid data leakage (Line~3).
+After running our active-learning framework for a solver $\hat{a}$ in Line~4, we compute the value of both our optimization goals:
+First and foremost, we want to provide the engineer of new SAT solvers with an accurate ranking.
+We define the \emph{ranking accuracy} $O_{\operatorname{acc}} \in \left[0, 1\right]$ (higher is better) by the fraction of pairs $\left(\hat{a}, a\right)$ for all $a \in \mathcal{A}$ that are decided correctly by the given ranking regarding the PAR-2 scores $\operatorname{par_2}$ (Lines~5-8).
+Second, we also want to optimize for runtime.
+The \emph{fraction of runtime} that the algorithm needs to arrive at its conclusion is denoted by $O_{\operatorname{rt}} \in \left[0, 1\right]$ (lower is better).
+This metric puts the runtime summed over the sampled instances in relation to the runtime summed over all instances in the dataset (Lines~9-13).
+Finally, we compute averages of the output metrics in Line~15 after we have collected all cross-validation results in Line~14.
+Overall, we want to find an approach that maximizes
+%
+\begin{equation}
+  O_\delta := \delta O_{\operatorname{acc}} + \left(1 - \delta\right) \left(1 - O_{\operatorname{rt}}\right) \enspace \textrm{,}
+  \label{eq:opt}
+\end{equation} 
+%
+whereby $\delta \in \left[0, 1\right]$ allows for linear weighting between the two optimization goals $O_{\operatorname{acc}}$ and $O_{\operatorname{rt}}$.
+Plotting the approaches that maximize $O_\delta$ for all $\delta \in \left[0, 1\right]$ on a $O_{\operatorname{rt}}$-$O_{\operatorname{acc}}$-diagram provides us with a Pareto front of the best approaches for different optimization-goal weightings.
+
 \subsection{Data}
@@
 
-Given Algorithm~\ref{algALBenchmark}, there are several possible instantiations for the three phases, i.e., \emph{selection}, \emph{stopping}, and \emph{ranking}.
+Given Algorithm~\ref{algALBenchmark}, there are several possible instantiations for the three phases, i.e., \emph{ranking}, \emph{selection}, and \emph{stopping}.
 Also, there are different choices for the runtime-label prediction model and runtime discretization.
@@
     History size: consider the latest 1, 10, 20, 30, or 40 predictions within a voting approach for stability.
-    The latest $x$ predictions vote on the winner.
-    %JB: we vote for the label of each instance separately, right? (rather than globally ensembling the overall solver score of the last x iteration)
+    The latest $x$ predictions vote on the winner on a per-instance level.
     \item
-    Fallback threshold: if the difference of scores between two solvers drops below \SI{0.01}, \SI{0.05}, or \SI{0.1}, use the partially observed PAR-2 ranking as a tie-breaker.
-    %JB: as clarification: one these two solvers has to be the new solver, right? if yes, would write it explicitly
-    %TF: Not necessarily
-    %JB: but how can the scores of two existing solvers change? and we only have partially observed scores for the new solver
-    As discussed in Section~\ref{sec:main:model}, ranking decisions are all correct if the difference in scores between two solvers exceeds a certain value.
+    Fallback threshold: if the difference of scores between two solvers (including the new solver $\hat{a}$) drops below \SI{0.01}, \SI{0.05}, or \SI{0.1}, use the partially observed PAR-2 ranking as a tie-breaker.
   \end{itemize}
@@
 
-For \emph{selection} (cf. Section~\ref{sec:main:selection}), we experiment with the following methods, including our used hyper-parameter values:
+For \emph{selection} (cf. Section~\ref{sec:main:selection}), we experiment with the following methods, including our used hyper-parameter values.
+Since the potential runtime of experiments is by magnitudes larger than the model's update time, we only consider incrementing our benchmark by one instance at a time rather than using batches, which is also proposed in current active-learning advances~\cite{SinhaED19,2019gaal}.
+A drawback of this is the lack of parallelization.
 
@@
     \item Fallback threshold: use random sampling for the first \SI{0}{\%}, \SI{5}{\%}, \SI{10}{\%}, \SI{15}{\%}, or \SI{20}{\%} of instances to explore the instance space.
-    \item Runtime scaling: prefer instances with the greatest uncertainty per average (over all solvers) runtime (\texttt{True} or \texttt{False}).
+    \item Runtime scaling: prefer instances with the greatest uncertainty per average runtime (over all solvers; \texttt{True} or \texttt{False}).
   \end{itemize}
@@
   \end{itemize}
-
-  \item Neighborhood-aware random sampling (preliminary experiments only); randomly chooses among the instances with the highest amount of non-sampled $k$ neighbors
-  \item Ranking sampling~\cite{Bossek021a} (preliminary experiments only)
-  \item Discrimination sampling~\cite{GentHJKMNN14} (preliminary experiments only)
-  \item Variance sampling~\cite{MatriconAFSH21} (preliminary experiments only)
 \end{itemize}
@@
 
-%Our model of choice is an ensemble, stacking a quadratic-discriminant analysis\footnote{With a regularization parameter of zero and a singular value threshold of zero} onto a random forest model\footnote{With \emph{entropy} splitting criterion and \emph{balanced} class weights} using a decision tree\footnote{With \emph{Gini} impurity splitting criterion; choosing the best split for each branching decision and maximum tree depth of 5 levels} to weight its members.
-%If not stated explicitly, all other model parameters default to the presets of \emph{scikit-learn}\footnote{\url{https://scikit-learn.org/stable/index.html}; Version 1.0.2}~\cite{scikit-learn}.
-%
-%
-%Therefore, we compare predictors first and select one runtime-prediction model for all further experiments.
-%We evaluate predictors with the \emph{Matthews Correlation Coefficient} (MCC) scores~\cite{gorodkin2004comparing,matthews1975comparison} of the classification task.
-%MCC scores are great in dealing with class imbalances in contrast to conventional metrics like accuracy.
-%
-%% \begin{table}
-%	%   \centering
-%	%   \caption{
-%		%   	Time-out prediction Matthews Correlation Coefficient performances.
-%		%   	Results are obtained using cross-validation.
-%		%   	We provide \SI{64}{\%} of ground-truth for training, use \SI{16}{\%} as a validation set, and \SI{20}{\%} as a test set (5-fold cross-validation).
-%		%   }
-%	%   \label{tab:mcc}
-%	%   \begin{tabular}{lc}
-%		%     Prediction Model & Avg. MCC (Std.) \\
-%		%     \hline
-%		%     Stacking ensemble (QDA, RF)           & 0.9527 (0.0292) \\
-%		%     Quadratic-discriminant analysis (QDA) & 0.9290 (0.0339) \\
-%		%     Random forest (RF)                    & 0.8530 (0.0479) \\
-%		%     AdaBoost with decision trees          & 0.8384 (0.0444) \\
-%		%     AdaBoost with logistic regression     & 0.8158 (0.0985) \\
-%		%     Decision tree                         & 0.8059 (0.0707) \\
-%		%     Logistic regression                   & 0.8052 (0.1018) \\
-%		%     kNN classifier                        & 0.7885 (0.1521) \\
-%		%     MLP classifier                        & 0.7760 (0.1408) \\
-%		%     Support-vector classifier             & 0.7757 (0.2149) \\
-%		%     Naive Bayes                           & 0.7306 (0.1394) \\
-%		%     \hline
-%		%   \end{tabular}
-%	% \end{table}
-%
-%% Stacking MCC
-%% 0.888221577313471
-%% 0.04753979056930393
-%
-%%To obtain the results in Table~\ref{tab:mcc},
-%We perform cross-validation among all solvers: we assume that all other solver runtimes are known and provide \SI{64}{\%} of ground-truth runtime labels of the target solver for training, use \SI{16}{\%} as a validation set for hyper-parameter tuning of the runtime predictor, and \SI{20}{\%} as a test set (5-fold cross-validation).
-%This is repeated for each solver once.
-%Note that we perform cross-validation on two levels, i.e., among the solvers and the target solver runtimes.
-%We compare the average performances on the test sets.
-%
-%ML classification models that we have tried include random forests, decision trees, quadratic-discriminant analysis, AdaBoost, logistic regression, $k$-nearest neighbor, multi-layer perceptrons, support vector machines, and naive Bayes.
-%A stacking ensemble consisting of quadratic-discriminant analysis and random forests performed best.
-%It produces an MCC score of \SI{0.8882}{} for predicting three-class runtime labels.
-%% \SI{0.9527}{} for time-out prediction
-%The results are significantly better than any other model that we have tried regarding a Wilcoxon signed-rank test with $\alpha = \SI{5}{\%}$.
-%%Since the differences in the MCC-performance distributions of the best and second-best approaches are statistically significant regarding a Wilcoxon signed-rank test with $\alpha = \SI{5}{\%}$, we only use the best approach for all further experiments as a prediction model.
-
 \subsubsection{Runtime discretization.}
@@
 Other clustering approaches that we have tried include hierarchical clustering with mean-, median- and complete-link criteria, as well as $k$-means and spectral clustering.
-%Predicting runtime labels in a similar experiment as above produced an MCC score of on average \SI{88.82}{\%} ($\sigma = \SI{4.75}{\%}$; in comparison to \SI{95.27}{\%} for predicting time-outs).
-%JB: "similar experiment as above" - reference unclear; what do you want to say?
-%TF: Removed the sentence
 In our chosen hierarchical procedure, each non-time-out runtime starts in a separate interval.
@@
 According to a Wilcoxon-signed-rank test with $\alpha = 0.05$, \SI{87.83}{\%} of solver pairs have significantly different scores after discretization, only a slight drop compared to \SI{89.95}{\%} before discretization.
-%While only time-out labels (solver times out or not on a particular instance) are sufficient to significantly distinguish about \SI{87.04}{\%} of solver pairs.
-%\footnote{Based on hierarchical clustering with a log-single-link criterion for $k = 3$.}
 Further, our ranking approach correctly decides for almost all (about \SI{97.45}{\%}; $\sigma = \SI{3.68}{\%}$) solver pairs which solver is faster.
-%Moreover, it decides all solvers correctly if the difference in the cluster-label-ranking scores between two solvers is at least \SI{0.11}{}.
-%Refer to Table~\ref{tab:discreteranking} for a comparison with the PAR-2 ranking.
 In particular, the Spearman correlation of $s_3$ and PAR-2 ranking is about \SI{0.988}{}, which is very close to the optimal value of 1~\cite{de2016comparing}.
@@
 
-%\begin{table}[tbp]
-%  \centering
-%   \caption{
-	%     PAR-2 ranking vs. discretized-runtime ranking for the SAT Competition~2022 Anniversary Track instances~\cite{sat2022} ($k = 3$; hierarchical clustering).
-	%	  Differences are highlighted with a gray background.
-	%	}
-%  \label{tab:discreteranking}
-%  \begin{tabular}{
-		%    c
-		%    >{\centering\arraybackslash}p{1.8cm}p{3.6cm}
-		%    >{\centering\arraybackslash}p{1.8cm}p{3.6cm}
-		%  }
-	%    \hline
-	%    Rank & \multicolumn{2}{c}{PAR-2 Ranking} & \multicolumn{2}{c}{Cluster-Label Ranking} \\
-	%    \hline
-	%1   & 2806.400 & Kissat\_MAB\_ESA           & 1.172 & Kissat\_MAB\_ESA \\
-	%2   & 2810.959 & kissat-sc2022-bulky        & 1.183 & kissat-sc2022-bulky \\
-	%3   & 2830.143 & ekissat-mab-gb-db          & 1.185 & ekissat-mab-gb-db \\
-	%\cellcolor{gray!15} 4   & \cellcolor{gray!15} 2833.081 & \cellcolor{gray!15} Kissat\_MAB\_UCB           & \cellcolor{gray!15} 1.186 & \cellcolor{gray!15} kissat\_inc \\
-	%\cellcolor{gray!15} 5   & \cellcolor{gray!15} 2833.397 & \cellcolor{gray!15} kissat\_inc                & \cellcolor{gray!15} 1.186 & \cellcolor{gray!15} Kissat\_MAB\_UCB \\
-	%6   & 2841.734 & ekissat-mab-db-v1          & 1.192 & ekissat-mab-db-v1 \\
-	%7   & 2842.666 & Kissat\_MAB\_MOSS          & 1.192 & Kissat\_MAB\_MOSS \\
-	%8   & 2857.219 & Kissat\_MAB-HyWalk         & 1.195 & Kissat\_MAB-HyWalk \\
-	%9   & 2866.960 & kissat-sc2022-light        & 1.120 & kissat-sc2022-light \\
-	%10  & 2896.371 & kissat-els-v2              & 1.216 & kissat-els-v2 \\
-	%11  & 2950.179 & hKis-unsat                 & 1.229 & hKis-unsat \\
-	%12  & 2964.729 & Kissat\_adaptive\_restart  & 1.234 & Kissat\_adaptive\_restart \\
-	%13  & 2974.990 & SeqFROST-NoExtend          & 1.247 & SeqFROST-NoExtend \\
-	%14  & 3010.642 & Cadical\_ESA               & 1.248 & Cadical\_ESA \\
-	%\cellcolor{gray!15} 15  & \cellcolor{gray!15} 3012.152 & \cellcolor{gray!15} kissat-els-v1              & \cellcolor{gray!15} 1.258 & \cellcolor{gray!15} CadicalReorder \\
-	%\cellcolor{gray!15} 16  & \cellcolor{gray!15} 3028.922 & \cellcolor{gray!15} CadicalReorder             & \cellcolor{gray!15} 1.264 & \cellcolor{gray!15} kissat-els-v1 \\
-	%\cellcolor{gray!15} 17  & \cellcolor{gray!15} 3042.540 & \cellcolor{gray!15} cadical\_rel\_Scavel       & \cellcolor{gray!15} 1.278 & \cellcolor{gray!15} CaDiCaL\_DVDL\_V1 \\
-	%\cellcolor{gray!15} 18  & \cellcolor{gray!15} 3082.293 & \cellcolor{gray!15} kissat\_relaxed            & \cellcolor{gray!15} 1.283 & \cellcolor{gray!15} CaDiCaL\_DVDL\_V2 \\
-	%\cellcolor{gray!15} 19  & \cellcolor{gray!15} 3089.304 & \cellcolor{gray!15} CaDiCaL\_DVDL\_V1          & \cellcolor{gray!15} 1.298 & \cellcolor{gray!15} kissat\_relaxed \\
-	%\cellcolor{gray!15} 20  & \cellcolor{gray!15} 3095.118 & \cellcolor{gray!15} CaDiCaL\_DVDL\_V2          & \cellcolor{gray!15} 1.361 & \cellcolor{gray!15} cadical\_rel\_Scavel \\
-	%21  & 3272.674 & glucose-reboot             & 1.379 & glucose-reboot \\
-	%\cellcolor{gray!15} 22  & \cellcolor{gray!15} 3287.322 & \cellcolor{gray!15} LStech-Maple-HyWalk        & \cellcolor{gray!15} 1.414 & \cellcolor{gray!15} hCaD\_V1-psids \\
-	%23  & 3288.798 & LSTech\_Maple              & 1.446 & LSTech\_Maple \\
-	%\cellcolor{gray!15} 24  & \cellcolor{gray!15} 3399.281 & \cellcolor{gray!15} SLIME-SC-2022-beta         & \cellcolor{gray!15} 1.469 & \cellcolor{gray!15} LStech-Maple-HyWalk \\
-	%\cellcolor{gray!15} 25  & \cellcolor{gray!15} 3410.345 & \cellcolor{gray!15} SLIME-SC-2022              & \cellcolor{gray!15} 1.470 & \cellcolor{gray!15} SLIME-SC-2022-beta \\
-	%\cellcolor{gray!15} 26  & \cellcolor{gray!15} 3430.403 & \cellcolor{gray!15} hCaD\_V1-psids             & \cellcolor{gray!15} 1.523 & \cellcolor{gray!15} SLIME-SC-2022 \\
-	%27  & 3504.244 & MapleLCMDistChrBt-DL-v3    & 1.542 & MapleLCMDistChrBt-DL-v3 \\
-	%28  & 4750.808 & IsaSAT                     & 2.062 & IsaSAT \\
-	%    \hline
-	%  \end{tabular}
-%\end{table}
-
 \subsection{Implementation Details}
 
-For reproducibility, our source code, all experiments, and data are available on GitHub\footnote{\url{https://github.com/mathefuchs/al-for-SAT solver-benchmarking}}.
+% TODO: insert URLs
+For reproducibility, our source code and data are available on GitHub\footnote{temporary, anonymized version for review: \url{xxx}}.
 Our code is implemented in \textsc{Python} using \emph{scikit-learn}~\cite{scikit-learn} for making predictions and \emph{gbd-tools}~\cite{IserS18} for SAT-instance retrieval.
@@
 First, we analyze and tune the different components of our framework on the tuning dataset.
-Next, we evaluate the best configuration with the full dataset.
+Next, we evaluate the best configurations with the full dataset.
 Finally, we analyze the importance of different instance families to our framework.
@@
 \subsection{Hyper-Parameter Analysis}
-
-Our end-to-end experiments follow the evaluation framework introduced in Section~\ref{sec:exdesign:evalframe}.
-Fig.~\ref{fig:e2eallsolvers} shows the performance of the approaches from Section~\ref{sec:exdesign:hyper} on $O_{\operatorname{rt}}$-$O_{\operatorname{acc}}$-diagrams for the smaller hyper-parameter-tuning dataset.
-Evaluating a particular configuration with Algorithm~\ref{alg:eval} returns a point $\left(O_{\operatorname{rt}},\, O_{\operatorname{acc}}\right)$.
-In particular, we do not show intermediate results of the active-learning procedure but only final results after stopping.
-The plotted lines represent the best performing configurations (convex hull) per top-level hyper-parameter choice, i.e., per ranking approach (Fig.~\ref{fig:annitraincolorranking}), selection approach (Fig.~\ref{fig:annitraincolorselection}), and stopping criterion (Fig.~\ref{fig:annitraincolorstopping}).
-%Colors indicate the different hyper-parameter instantiations.
-%To be more specific, they represent the Pareto front of all configurations with the respective instantiations.
-
-Regarding ranking approaches (Fig.~\ref{fig:annitraincolorranking}), using the $s_3$-induced runtime-label ranking consistently outperforms the partially observed PAR-2 ranking for each possible value of the runtime-accuracy trade-off parameter~$\delta$.
-This is expected since selection decisions are not random.
-For example, we might sample more instances of one family if it benefits discrimination of solvers.
-While the partially observed PAR-2 score is skewed, the prediction model can account for this.
-
-Regarding the selection approaches (Fig.~\ref{fig:annitraincolorselection}), uncertainty sampling performs best in most cases.
-However, information-gain sampling is beneficial if runtime is strongly favored (very small $\delta$; runtime fraction less than \SI{5}{\%}).
-Random sampling performs quite well but is outperformed by uncertainty sampling in all cases, showing the benefit of actively selecting instances based on a prediction model.
-
-Regarding the stopping criterion (Fig.~\ref{fig:annitraincolorstopping}), the ranking stopping criterion performs the most consistently well.
-If accuracy is strongly favored (very high $\delta$), the Wilcoxon stopping criterion performs better.
-The subset-size stopping criterion performs reasonably well but does not improve beyond a certain accuracy.
 
@@
   \end{subfigure}
-
   \caption{
@@
     The x-axis shows the ratio of total solver runtime on the sampled instances relative to all instances.
-  	The y-axis shows the ranking accuracy (cf. Section~\ref{sec:exdesign:goal}).
+    The y-axis shows the ranking accuracy (cf. Section~\ref{sec:exdesign:eval}).
     Each line entails the front of Pareto-optimal configurations for the respective hyper-parameter instantiation.
@@
 
-% One of the main observations is that uncertainty sampling (blue points) consistently outperformed random sampling (red points) for almost any value of $\delta$ within Eq.~\ref{eq:opt}.
-% Moreover, our ranking stopping criterion (dot-shaped points) is particularly suited to achieve a good ranking accuracy fast.
-% In comparison, our Wilcoxon stopping criterion is suited better for high-accuracy predictions.
-% This comes at a high runtime cost though.
-% A good trade-off is achieved for $\delta = 0.85$ with an average PAR-2 ranking performance of \SI{91.83}{\%} and a runtime requirement of on average \SI{23.05}{\%} of the time it would take to evaluate the complete dataset.
+Our end-to-end experiments follow the evaluation framework introduced in Section~\ref{sec:exdesign:eval}.
+Fig.~\ref{fig:e2eallsolvers} shows the performance of the approaches from Section~\ref{sec:exdesign:hyper} on $O_{\operatorname{rt}}$-$O_{\operatorname{acc}}$-diagrams for the smaller hyper-parameter-tuning dataset.
+Evaluating a particular configuration with Algorithm~\ref{alg:eval} returns a point $\left(O_{\operatorname{rt}},\, O_{\operatorname{acc}}\right)$.
+In particular, we do not show intermediate results of the active-learning procedure but only the final results after stopping.
+The plotted lines represent the best performing configurations (convex hull) per top-level hyper-parameter choice, i.e., per ranking approach (Fig.~\ref{fig:annitraincolorranking}), selection approach (Fig.~\ref{fig:annitraincolorselection}), and stopping criterion (Fig.~\ref{fig:annitraincolorstopping}).
 
-% \begin{table}[tb!]
-%   %JB: really hard to extract take-aways by looking at this table (several hyperprameter settings and several metrics, the hyperparameter settings of our approach change from row to row)
-%   \centering
-%   \caption{
-%   	Pareto-optimal approaches for different choices of $\delta$ regarding our smaller hyper-parameter tuning dataset.
-%   	Rows with no changes in comparison to their predecessor are hidden.
-%   	%JB: no change in what? all columns except delta?
-%   	Ranking approaches are either predicted runtime-label ranking (Pred.) or observed PAR-2 ranking (PAR-2).
-%   	Selection approaches are either uncertainty sampling (Uncert.) or information-gain sampling (IG).
-%   	Stopping criteria are either ranking (Rank) or Wilcoxon (Wilcoxon).
-%   	%JB: How are the values for ranking/selection/stopping determined -- whatever is best for a particular delta?
-%   }
-%   \label{tab:paretotable}
-%   \vspace{0.2cm}
-%   \begin{tabular}{
-%     >{\centering\arraybackslash}m{0.6cm}
-%     >{\centering\arraybackslash}m{1.6cm}
-%     >{\centering\arraybackslash}m{1.6cm}
-%     >{\centering\arraybackslash}m{1.6cm}
-%     >{\centering\arraybackslash}m{1.5cm}
-%     >{\centering\arraybackslash}m{1.5cm}
-%     >{\centering\arraybackslash}m{1.5cm}
-%     >{\centering\arraybackslash}m{1.5cm}
-%   }
-%     \hline
-%     $\delta$ & Ranking & Selection & Stopping & Fraction Runtime ($O_{\operatorname{rt}}$) & Fraction Instances & $O_{\operatorname{acc}}$ & $O_{\delta}$ \\
-%     \hline
-%     0.00 & PAR-2 & Uncert. & Rank & 0.021 & 0.025 & 0.593 & 0.979 \\
-%     0.05 & Pred. & Uncert. & Rank & 0.024 & 0.158 & 0.765 & 0.966 \\
-%     0.10 & Pred. & Uncert. & Rank & 0.025 & 0.128 & 0.779 & 0.955 \\
-%     0.15 & Pred. & IG & Rank & 0.028 & 0.112 & 0.802 & 0.946 \\
-%     $\vdots$ & $\vdots$ & $\vdots$ & $\vdots$ & $\vdots$ & $\vdots$ & $\vdots$ & $\vdots$ \\
-%     0.50 & Pred. & IG & Rank & 0.028 & 0.112 & 0.802 & 0.887 \\
-%     0.55 & Pred. & Uncert. & Rank & 0.053 & 0.142 & 0.824 & 0.880 \\
-%     0.60 & Pred. & Uncert. & Rank & 0.146 & 0.082 & 0.897 & 0.880 \\
-%     $\vdots$ & $\vdots$ & $\vdots$ & $\vdots$ & $\vdots$ & $\vdots$ & $\vdots$ & $\vdots$ \\
-%     0.75 & Pred. & Uncert. & Rank & 0.146 & 0.082 & 0.897 & 0.886 \\
-%     0.80 & Pred. & Uncert. & Rank & 0.219 & 0.128 & 0.917 & 0.890 \\
-%     \hline
-%     0.85 & Pred. & Uncert. & Rank & 0.272 & 0.171 & 0.927 & 0.897 \\
-%     \hline
-%     0.90 & Pred. & Uncert. & Wilcoxon & 0.894 & 0.888 & 0.996 & 0.907 \\
-%     0.95 & Pred. & Uncert. & Wilcoxon & 0.931 & 0.924 & 1.000 & 0.953 \\
-%     1.00 & Pred. & Uncert. & Wilcoxon & 0.989 & 0.981 & 1.000 & 1.000 \\
-%     \hline
-%   \end{tabular}
-% \end{table}
+Regarding ranking approaches (Fig.~\ref{fig:annitraincolorranking}), using the $s_3$-induced runtime-label ranking consistently outperforms the partially observed PAR-2 ranking for each possible value of the runtime-accuracy trade-off parameter~$\delta$.
+This is expected since selection decisions are not random.
+For example, we might sample more instances of one family if it benefits discrimination of solvers.
+While the partially observed PAR-2 score is skewed, the prediction model can account for this.
+
+Regarding the selection approaches (Fig.~\ref{fig:annitraincolorselection}), uncertainty sampling performs best in most cases.
+However, information-gain sampling is beneficial if runtime is strongly favored (small $\delta$; runtime fraction less than \SI{5}{\%}).
+This also aligns with our expectations:
+Information-gain sampling selects instances that maximize the expected reduction in entropy.
+This means that we sample instances revealing similarities between solvers rather than differences.
+This helps in building a confident model fast.
+However, the method lacks the possibility of selecting helpful instances for distinguishing solvers later.
+Random sampling performs quite well but is outperformed by uncertainty sampling in all cases, showing the benefit of actively selecting instances based on a prediction model.
+
+Regarding the stopping criterion (Fig.~\ref{fig:annitraincolorstopping}), the ranking stopping criterion performs the most consistently well.
+If accuracy is strongly favored (very high $\delta$), the Wilcoxon stopping criterion performs better.
+The subset-size stopping criterion performs reasonably well but does not improve beyond a certain accuracy.
 
@@
   \centering
-  \resizebox{0.78\textwidth}{!}{
+  \begin{subfigure}{0.4775\textwidth}
+    \resizebox{\textwidth}{!}{
       \graphicspath{{../plots/}}
@@
     }
+    \caption{Runtime vs. Instances}
+    \label{fig:annitrainoptgoalruntime}
+  \end{subfigure}
+  \begin{subfigure}{0.5125\textwidth}
+    \resizebox{\textwidth}{!}{
+      \graphicspath{{../plots/}}
+      \input{../plots/anni_train_delta_acc.pgf}
+    }
+    \caption{Runtime vs. Accuracy}
+    \label{fig:annitrainoptgoalacc}
+  \end{subfigure}
+
   \caption{
     Scatter plot comparing different instantiations of $\delta$ for our active-learning strategy on the hyper-parameter-tuning dataset.
-    The x-axis shows the fraction of runtime $O_{\operatorname{rt}}$ of the sample, while the y-axis shows the fraction of instances sampled.
+    The x-axis shows the fraction of runtime $O_{\operatorname{rt}}$ of the sample, while the y-axes show the fraction of instances sampled and ranking accuracy respectively.
     The color indicates the weighting between different optimization goals $\delta \in \left[0, 1\right]$.
@@
 
-Figure~\ref{fig:annitrainoptgoal} shows an interesting consequence of weighting our optimization goals:
+Figure~\ref{fig:annitrainoptgoalruntime} shows an interesting consequence of weighting our optimization goals:
 If we, on the one hand, desire to get a \emph{rough} estimate of a solver's performance fast (low $\delta$), approaches favor selecting many \emph{easy} instances.
@@
 
+Furthermore, Figure~\ref{fig:annitrainoptgoalacc} reveals which values make the most sense for $\delta$.
+The range $\delta \in \left[0.2, 0.8\right]$, thereby, corresponds to the points between a runtime fraction of \SI{0.03} and \SI{0.22}.
+We consider this region to be most promising, analogous to the \emph{elbow} method in cluster analysis~\cite{kodinariya2013review}.
+
 \subsection{Full-Dataset Evaluation}
 
-Having selected the most promising hyper-parame\-ters, we run our end-to-end active-learning experiments on the complete Anniversary Track dataset (5355 instances).
-The best-performing approach for $\delta = 0.85$ uses the predicted runtime-label ranking, uncertainty sampling, and a ranking stopping criterion.
-%JB: we need to motivate choosing this \delta
-It can predict a new solver's PAR-2 ranking with about \SI{92.33}{\%} accuracy ($O_{\operatorname{acc}}$), while only needing \SI{10.35}{\%} of the time for running the solver on the full dataset ($O_{\operatorname{rt}}$).
-%JB: could we (with reasonable effort) integrate the learning curve of this particular approach as a plot?
-This trade-off provides a better way of prototyping and evaluating new SAT solvers rather than evaluating them on manually picked instance based on domain knowledge.
-The sampled instances account for about \SI{5.24}{\%} of instances, indicating a preference for selecting difficult instances.
-The average Spearman correlation between the PAR-2 ranking and the predicted ranking is about \SI{0.9572}{}, indicating a high similarity of these two rankings.
-Because of that, the runtime-prediction model helps determine on what problem instances solvers perform similarly or are better/worse.
-%JB: conclusion not straightforward from prior text; even if overall estimated score (which is an average over many instances) is good, estimate on individual instances might be quite erroneous
+Having selected the most promising hyper-parameters, we run our end-to-end active-learning experiments on the complete Anniversary Track dataset (5355 instances).
+The aforementioned range $\delta \in \left[0.2, 0.8\right]$ is covered by only two configurations.
+The best-performing approach for $\delta \in \left[0.2, 0.7\right]$ uses the predicted runtime-label ranking, information-gain sampling, and ranking-stopping criterion.
+It can predict a new solver's PAR-2 ranking with about \SI{90.48}{\%} accuracy ($O_{\operatorname{acc}}$), while only needing \SI{5.41}{\%} of the full evaluation time ($O_{\operatorname{rt}}$).
+The best-performing approach for $\delta \in (0.7, 0.8]$ uses the predicted runtime-label ranking, uncertainty sampling, and ranking-stopping criterion.
+It can predict a new solver's PAR-2 ranking with about \SI{92.33}{\%} accuracy ($O_{\operatorname{acc}}$), while only needing \SI{10.35}{\%} of the full evaluation time ($O_{\operatorname{rt}}$).
+These trade-offs provide a better way of prototyping and evaluating new SAT solvers rather than evaluating them on manually picked instances based on domain knowledge.
 
 \subsection{Instance-Family Importance}
-
-Selection decisions of our approach also reveal the importance of different instance families to our framework.
-Figure~\ref{fig:annifinalfamilies} shows the occurrence of instance families within the dataset and the benchmarks created by active learning.
-We use the best-performing configurations for all $\delta \in \left[0, 1\right]$ and examine the selection decisions by the active-learning approach on the SAT Competition~2022 Anniversary Track dataset~\cite{sat2022}.
-While most families appear with the same fraction in the dataset and the sampled benchmarks, there are a few outliers that need further discussion.
-Problem instances of the families \emph{fpga}, \emph{quasigroup-completion}, and \emph{planning} are especially helpful to our framework in distinguishing solvers.
-Instances of these families are selected over-proportionally in comparison to the full dataset.
-In contrast to that, instances of the biggest family, i.e., \emph{hardware-verification}, roughly appear with the same fraction in the dataset and the sampled benchmarks.
-Finally, instances of the family \emph{cryptography} are less important in distinguishing solvers than their huge weight in the dataset suggests.
-A possible explanation is that these instances are very similar, such that a small fraction of them is sufficient to estimate a solver's performance on all of them.
-
-% For all configurations of our approach that sample on average less than \SI{10}{\%} of all 5355 instances, we look at the particular instances and families that are selected by our active-learning approach.
-% Thereby, the average benchmark set contains about \SI{192.34}{} instances.
-% Table~\ref{tab:familiesa} lists the ten most frequent families that are chosen.
-% For example, \emph{planning} instances account for on average \SI{19.14}{} of the \SI{192.34}{} benchmark instances.
-% In contrast to that, Table~\ref{tab:familiesb} shows the families that are sampled the most often relative to their occurrences in the underlying dataset.
-% There are, for example, only nine \emph{circuit-multiplier} instances in the complete dataset.
-% However, we sample on average \SI{1.59}{} of them.
-% %JB: there should be some (qualitative) interpretation (based on our results, is the sample interesting/special family-wise or not?) rather than just dropping a table into the paper and mainly explaining its structure
-
-% \begin{table}[tb!]
-%   %JB: "fraction" here only draws family-specific comparison of sample vs.
-full dataset, but does not compare overall distribution of families in sample/dataset; might be interesting to have scatterplot with "(relative) frequency of family in sample" vs.
-"(relative) frequency of family in dataset" (does not allow to show all family labels (though we probably could manually label some families), but allows to show all families)
-%   %JB: could also correlate "(relative) frequency of family in sample" with "(relative) frequency of family in dataset"
-%   \centering
-%   \caption{Top-10 most \emph{important} families regarding the average occurrences and the fraction of selected occurrences regarding our AL approach}
-%   \label{tab:families}
-%   \begin{subtable}[t]{0.48\textwidth}
-%     \centering
-%     \caption{Sorted by the average number of occurrences in the sample}
-%     \label{tab:familiesa}
-
-%     \begin{tabular}{
-%       >{\centering\arraybackslash}m{1.8cm}
-%       >{\centering\arraybackslash}m{1.2cm}
-%       >{\centering\arraybackslash}m{1.2cm}
-%       >{\centering\arraybackslash}m{1.2cm}
-%     }
-%       \hline
-%       Family & In Sample & In Dataset & Fraction (\%) \\
-%       \hline
-%       planning              & 19.14 &  333 &  5.7 \\
-%       subgraph-isomorphism  & 16.27 &  175 &  9.3 \\
-%       crypto-graphy         & 13.11 &  311 &  4.2 \\
-%       pigeon-hole           &  7.55 &  151 &  5.0 \\
-%       fpga-routing          &  6.16 &  73  &  8.4 \\
-%       rbsat                 &  5.93 &  113 &  5.2 \\
-%       quasigroup-completion &  5.23 &  210 &  2.5 \\
-%       unknown               &  4.93 &  169 &  2.9 \\
-%       miter                 &  4.84 &  199 &  2.4 \\
-%       bitvector             &  4.73 &  214 &  2.2 \\
-%       % clique-coloring       &  4.21 &  64  &  0.066 \\
-%       % mosoi-289             &  3.75 &  31  &  0.121 \\
-%       % hardware-verification &  3.70 &  344 &  0.011 \\
-%       % hardware-bmc          &  2.77 &  78  &  0.035 \\
-%       % graph-based           &  2.77 &  134 &  0.021 \\
-%       % coloring              &  2.66 &  41  &  0.065 \\
-%       % waerden               &  2.55 &  43  &  0.059 \\
-%       % graph-coloring        &  2.50 &  87  &  0.029 \\
-%       % auto-correlation      &  2.39 &  46  &  0.052 \\
-%       % knights-problem       &  2.25 &  17  &  0.132 \\
-%       %\hline
-%       Remaining families    & 104.45 & 3407 & 3.1 \\
-%       \hline
-%       Sum                   & 192.34 & 5355 & 3.6 \\
-%       \hline
-%     \end{tabular}
-%   \end{subtable}
-%   \hfill
-%   \begin{subtable}[t]{0.48\textwidth}
-%     \centering
-%     \caption{Sorted by the fraction of sampled occurrences}
-%     \label{tab:familiesb}
-
-%     \begin{tabular}{
-%       >{\centering\arraybackslash}m{1.8cm}
-%       >{\centering\arraybackslash}m{1.2cm}
-%       >{\centering\arraybackslash}m{1.2cm}
-%       >{\centering\arraybackslash}m{1.2cm}
-%     }
-%       \hline
-%       Family & In Sample & In Dataset & Fraction (\%) \\
-%       \hline
-%       circuit-multiplier          & 1.59    &      9      &     17.7 \\
-%       subset-cardinality          & 1.55    &      9      &     17.3 \\
-%       ssp-0                       & 0.34    &      2      &     17.0 \\
-%       sliding-puzzle              & 1.52    &      11     &     13.8 \\
-%       knights-problem             & 2.25    &      17     &     13.2 \\
-%       genurq                      & 1.55    &      12     &     12.9 \\
-%       stone                       & 1.39    &      11     &     12.7 \\
-%       mosoi-289                   & 3.75    &      31     &     12.1 \\
-%       anti-bandwidth              & 1.52    &      13     &     11.7 \\
-%       random                      & 0.91    &      8      &     11.4 \\
-%       % spectrum-repacking          & 1.30    &      12     &     0.108631 \\
-%       % social-golfer               & 0.95    &      9      &     0.105159 \\
-%       % product-configuration       & 0.52    &      5      &     0.103571 \\
-%       % minimal-disagreement-parity & 1.54    &      16     &     0.095982 \\
-%       % subgraph-isomorphism        & 16.27   &       175   &     0.092959 \\
-%       % sgen-balanced               & 0.27    &      3      &     0.089286 \\
-%       % pythagorean-triples         & 1.82    &      21     &     0.086735 \\
-%       % fpga-routing                & 6.16    &      73     &     0.084393 \\
-%       % petrinet-concurrency        & 2.11    &      25     &     0.084286 \\
-%       % core-based-generator        & 1.00    &      13     &     0.076923 \\
-%       %\hline
-%       Remaining families    & 175.97 & 5232 & 3.4 \\
-%       \hline
-%       Sum                   & 192.34 & 5355 & 3.6 \\
-%       \hline
-%     \end{tabular}
-%   \end{subtable}
-% \end{table}
 
@@
     The y-axis shows the average frequency of instance families in the samples selected by active learning.
-		The dashed line represents families that occur with the same frequency in both, dataset and samples.
+    The dashed line represents families that occur with the same frequency in both, the dataset and samples.
   }
@@
 \end{figure}
+
+Selection decisions of our approach also reveal the importance of different instance families to our framework.
+Figure~\ref{fig:annifinalfamilies} shows the occurrence of instance families within the dataset and the benchmarks created by active learning.
+We use the best-performing configurations for all $\delta \in \left[0, 1\right]$ and examine the selection decisions by the active-learning approach on the SAT Competition~2022 Anniversary Track dataset~\cite{sat2022}.
+While most families appear with the same fraction in the dataset and the sampled benchmarks, there are a few outliers that need further discussion.
+Problem instances of the families \emph{fpga}, \emph{quasigroup-completion}, and \emph{planning} are especially helpful to our framework in distinguishing solvers.
+Instances of these families are selected over-proportionally in comparison to the full dataset.
+In contrast to that, instances of the biggest family, i.e., \emph{hardware-verification}, roughly appear with the same fraction in the dataset and the sampled benchmarks.
+Finally, instances of the family \emph{cryptography} are less important in distinguishing solvers than their huge weight in the dataset suggests.
+A possible explanation is that these instances are very similar, such that a small fraction of them is sufficient to estimate a solver's performance on all of them.
 
@@
 However, bigger batch sizes $n$ impede \emph{active learning}.
-Also, it is unclear how to synchronize model update and instance selection without wasting too much runtime.
+Also, it is unclear how to synchronize the model update and instance selection without wasting too much runtime.
 
 On a more general note, one can apply our evaluation framework to arbitrary $\mathcal{NP}$-complete problems as all discussed active-learning methods are problem-agnostic.
-Solely the runtime-prediction model might need some tweaking.
-%JB: other components of framework might need tweaking as well, or this there something special about prediction model?
-Those problems share most of the relevant properties of SAT solving, i.e., there are established problem-instance features, a full benchmark run takes days, and solver development traditionally requires expert knowledge to hand-select instances.
+Especially the runtime-prediction model would need some tweaking for this purpose though.
+Those problems share most of the relevant properties of SAT solving, i.e., there are established problem-instance features, a full benchmark run takes quite some time, and solver development traditionally requires expert knowledge to hand-select instances.
 

### 2022-10-12 paper/paper.tex
Minor revisions in whole text
... based on e47638aa7d.

@@
 Our approach satisfies several desirable criteria for benchmarking:
-It outputs a \emph{ranking} to compare the new solver against the existing set of solvers.
+Rather than outputting a binary classification, i.e, the new solver is worse than an existing solver or not, we provide a \emph{scoring} function that shows by which margin a solver is worse and how similar it is to existing solvers.
+In particular, our approach enables \emph{ranking} the new solver amidst a set of existing solvers.
 For this ranking, we do not even need to predict exact solver runtimes, which is trickier.
-We optimize the \emph{runtime} that our strategy needs to arrive at its conclusion.
+Further, we optimize the \emph{runtime} that our strategy needs to arrive at its conclusion.
 We use problem-instance and runtime \emph{features}.
@@
 By doing so, we can control the properties of the benchmarking approach such as its required runtime.
-Rather than solely outputting a binary classification, i.e, the new solver is worse than an existing solver, we provide a \emph{scoring} function that shows by which margin a solver is worse and how similar it is to existing solvers.
 Our approach is \emph{scalable} in that it ranks a new solver $\hat{a}$ among any number of known solvers $\mathcal{A}$.
@@
 In the following, we discuss related work from the areas of static benchmark selection, algorithm configuration, incremental benchmark selection, and active learning.
+We compare the most relevant approaches in Table~\ref{tab:requirements}.
 
@@
 They calculate a scoring metric for all unselected instances, run the experiment with the highest score, and update the confidence.
-Their approach ticks off most of our desired features.
+Their approach ticks off most of our desired features in Table~\ref{tab:requirements}.
 However, the approach only compares solvers binarily rather than providing a \emph{scoring}.
@@
 This creates the machine-learning prediction model and also transforms runtimes as discussed in the subsequent section.
-We do not need any further steps here, since we have not obtained any labeled samples yet, so cannot train anything.
+We do not need any further steps here, since we have not obtained any labeled samples yet, so we cannot train the model yet.
 The model update (Line~7) uses all acquired runtimes to train a new runtime-prediction function $f$ using common supervised machine-learning methodology.
@@
 We transform the runtimes into discrete runtime labels on a per-instance basis.
-For each instance $e \in \mathcal{I}$, we use a clustering algorithm to assign the runtimes in $\bigl\{ r(a, e) \mid a \in A \bigr\}$ to one of $k$ clusters $C_1, \dots, C_k$ such that the fastest runtimes for instance $e$ are in cluster $C_1$ and the slowest are in cluster $C_{k-1}$, excluding timeouts $\tau$.
-Those are always in cluster $C_{k}$.
+For each instance $e \in \mathcal{I}$, we use a clustering algorithm to assign the runtimes in $\bigl\{ r(a, e) \mid a \in A \bigr\}$ to one of $k$ clusters $C_1, \dots, C_k$ such that the fastest runtimes for instance $e$ are in cluster $C_1$ and the slowest are in cluster $C_{k-1}$.
+Timeouts $\tau$ always form a separate cluster $C_{k}$.
 The runtime transformation function $\gamma_k : {\mathcal{A} \times \mathcal{I}} \rightarrow \left\lbrace 1, \dots, k \right\rbrace$ is then specified as follows:
@@
 
-Selecting an instance based on the model is a core functionality of our framework (cf. Algorithm~\ref{algALBenchmark}, Line~2).
+Selecting an instance based on the model is a core functionality of our framework (cf. Algorithm~\ref{algALBenchmark}, Line~4).
 In this section, we introduce our sampling strategies, which make use of the model's label-prediction function $f$ and are inspired by existing work within the realms of active learning~\cite{settles2009active}.
@@
 
-In this section, we present two dynamic stopping criteria, the Wilcoxon and the ranking-stopping criterion (cf. Algorithm~\ref{algALBenchmark}, Line~1).
+In this section, we present two dynamic stopping criteria, the Wilcoxon and the ranking stopping criterion (cf. Algorithm~\ref{algALBenchmark}, Line~3).
 
@@
 
-The ranking-stopping criterion is less sophisticated in comparison.
+The ranking stopping criterion is less sophisticated in comparison.
 It stops the active-learning process if the ranking induced by the model's predictions (Equation~\ref{eq:rankingeq}) remains unchanged within the last $l$ iterations.
@@
     \For{$a \in \mathcal{A}$}{
-      \If{$\bigl(s_k(a) - s_k(\hat a)\bigr) \cdot \bigl(\operatorname{par_2}(a) - \operatorname{par_2}(\hat a)\bigr) > 0$}{
+      \If{$\bigl(s_k(a) - s_{\hat a}\bigr) \cdot \bigl(\operatorname{par_2}(a) - \operatorname{par_2}(\hat a)\bigr) > 0$}{
            $O_{\operatorname{acc}} \leftarrow O_{\operatorname{acc}} + \frac{1}{|\mathcal{A}|}$ \;
@@
 
-To evaluate a concrete instantiation of Algorithm~\ref{algALBenchmark} (a concrete choice for all the sub-routines), we perform cross-validation on our set of solvers.
+To evaluate a concrete instantiation of Algorithm~\ref{algALBenchmark}, i.e., a concrete choice for all the sub-routines, we perform cross-validation on our set of solvers.
 Algorithm~\ref{alg:eval} shows this.
-That means that each solver once plays the role of the new solver (Line~2).
+That means that each solver plays the role of the new solver once (Line~2).
 Note that the \emph{new} solver in each iteration is excluded from the set of solvers $\mathcal{A}$ to avoid data leakage (Line~3).
@@
     History size: consider the latest 1, 10, 20, 30, or 40 predictions within a voting approach for stability.
-    The latest $x$ predictions vote on the winner on a per-instance level.
+    The latest $x$ predictions for each instance vote on the instance's winning label.
     \item
-    Fallback threshold: if the difference of scores between two solvers (including the new solver $\hat{a}$) drops below \SI{0.01}, \SI{0.05}, or \SI{0.1}, use the partially observed PAR-2 ranking as a tie-breaker.
+    Fallback threshold: if the difference of scores between the new solver $\hat{a}$ and another solver drops below \SI{0.01}, \SI{0.05}, or \SI{0.1}, use the partially observed PAR-2 ranking as a tie-breaker.
   \end{itemize}
@@
 Furthermore, Figure~\ref{fig:annitrainoptgoalacc} reveals which values make the most sense for $\delta$.
-The range $\delta \in \left[0.2, 0.8\right]$, thereby, corresponds to the points between a runtime fraction of \SI{0.03} and \SI{0.22}.
+The range $\delta \in \left[0.2, 0.8\right]$, thereby, corresponds to the points with a runtime fraction between \SI{0.03} and \SI{0.22}.
 We consider this region to be most promising, analogous to the \emph{elbow} method in cluster analysis~\cite{kodinariya2013review}.
@@
 The aforementioned range $\delta \in \left[0.2, 0.8\right]$ is covered by only two configurations.
-The best-performing approach for $\delta \in \left[0.2, 0.7\right]$ uses the predicted runtime-label ranking, information-gain sampling, and ranking-stopping criterion.
+The best-performing approach for $\delta \in \left[0.2, 0.7\right]$ uses the predicted runtime-label ranking, information-gain sampling, and ranking stopping criterion.
 It can predict a new solver's PAR-2 ranking with about \SI{90.48}{\%} accuracy ($O_{\operatorname{acc}}$), while only needing \SI{5.41}{\%} of the full evaluation time ($O_{\operatorname{rt}}$).
-The best-performing approach for $\delta \in (0.7, 0.8]$ uses the predicted runtime-label ranking, uncertainty sampling, and ranking-stopping criterion.
+The best-performing approach for $\delta \in (0.7, 0.8]$ uses the predicted runtime-label ranking, uncertainty sampling, and ranking stopping criterion.
 It can predict a new solver's PAR-2 ranking with about \SI{92.33}{\%} accuracy ($O_{\operatorname{acc}}$), while only needing \SI{10.35}{\%} of the full evaluation time ($O_{\operatorname{rt}}$).

### 2022-10-12 paper/paper.tex
Address minor comments

@@
 Let benchmark instances $\mathcal{I}$, SAT solvers $\mathcal{A}$, and runtimes $r : \mathcal{A} \times \mathcal{I} \rightarrow \left[0, \tau\right]$ be given.
+Model initialization (Line~1) then creates the machine-learning prediction model as well as transforms runtimes as discussed in the subsequent section.
+No model training is needed at this point since we do not have any labeled data yet.
 We denote the new solver to be ranked by $\hat a \not\in A$ and define $\hat A := A \cup \lbrace \hat a \rbrace$ to denote all solvers including the new solver. 
 Model $\mathcal{M}$ provides a prediction function $f : \mathcal{\hat A} \times \mathcal{I} \rightarrow \mathbb{R}$ that powers the decisions within our framework.
-
-First, we initialize the model $\mathcal{M}$ in Line~1.
-This creates the machine-learning prediction model and also transforms runtimes as discussed in the subsequent section.
-We do not need any further steps here, since we have not obtained any labeled samples yet, so we cannot train the model yet.
 The model update (Line~7) uses all acquired runtimes to train a new runtime-prediction function $f$ using common supervised machine-learning methodology.
+Training features are problem-instance features and known runtimes $r$.
+After applying necessary transformations, as described in the subsequent section, we fit the machine-learning model using the transformed observations $\mathcal{R}$ as ground-truth labels.
 It has also access to global instance features that assign each instance a vector in $\mathbb{R}^n$.
@@
 Note that we are not considering all previously listed approaches since a grid search of all combinations would be infeasible.
-Rather, we filter approaches based on preliminary experimental results and do the end-to-end experiments only with a subset.
-The end-to-end experiment configurations are given below.
+Rather, we filter approaches based on preliminary experimental results and do the experiments only with a subset.
+The experiment configurations are given below.
 
@@
   \begin{itemize}
-    \item Fallback threshold: use random sampling for the first \SI{0}{\%}, \SI{5}{\%}, \SI{10}{\%}, \SI{15}{\%}, or \SI{20}{\%} of instances to explore the instance space.
-    \item Runtime scaling: prefer instances with the greatest uncertainty per average runtime (over all solvers; \texttt{True} or \texttt{False}).
+    \item Fallback threshold: Use random sampling for the first \SI{0}{\%}, \SI{5}{\%}, \SI{10}{\%}, \SI{15}{\%}, or \SI{20}{\%} of instances to explore the instance space.
+    \item Runtime scaling: Whether to normalize uncertainty scores of instances by the average runtime of solvers on it or use the absolute values.
   \end{itemize}
@@
   \begin{itemize}
-    \item Fallback threshold: use random sampling for the first \SI{0}{\%}, \SI{5}{\%}, \SI{10}{\%}, \SI{15}{\%}, or \SI{20}{\%} of instances to explore the instance space.
-    \item Runtime scaling: prefer instances with the greatest information gain per average runtime (\texttt{True} or \texttt{False}).
+    \item Fallback threshold: Use random sampling for the first \SI{0}{\%}, \SI{5}{\%}, \SI{10}{\%}, \SI{15}{\%}, or \SI{20}{\%} of instances to explore the instance space.
+    \item Runtime scaling: Whether to normalize information-gain scores of instances by the average runtime of solvers on it or use the absolute values.
   \end{itemize}
@@
   \begin{itemize}
-    \item Minimum amount: sample at least \SI{2}{\%}, \SI{8}{\%}, \SI{10}{\%}, or \SI{12}{\%} of instances before applying the criterion.
-    \item Convergence duration: stop if the predicted ranking stays the same for a number of sampled instances equal to \SI{1}{\%} or \SI{2}{\%} of all instances.
+    \item Minimum amount: Sample at least \SI{2}{\%}, \SI{8}{\%}, \SI{10}{\%}, or \SI{12}{\%} of instances before applying the criterion.
+    \item Convergence duration: Stop if the predicted ranking stays the same for a number of sampled instances equal to \SI{1}{\%} or \SI{2}{\%} of all instances.
   \end{itemize}
@@
   \begin{itemize}
-    \item Minimum amount: sample at least \SI{2}{\%}, \SI{8}{\%}, \SI{10}{\%}, or \SI{12}{\%} of instances before applying the criterion.
+    \item Minimum amount: Sample at least \SI{2}{\%}, \SI{8}{\%}, \SI{10}{\%}, or \SI{12}{\%} of instances before applying the criterion.
     \item Average of $p$-values to drop below: \SI{5}{\%}.
-    \item Exponential-moving average: incorporate previous significance values by using an EMA with $\beta = 0.1$ or $\beta = 0.7$.
+    \item Exponential-moving average: Incorporate previous significance values by using an EMA with $\beta = 0.1$ or $\beta = 0.7$.
   \end{itemize}
@@
 
-Our end-to-end experiments follow the evaluation framework introduced in Section~\ref{sec:exdesign:eval}.
+Our experiments follow the evaluation framework introduced in Section~\ref{sec:exdesign:eval}.
 Fig.~\ref{fig:e2eallsolvers} shows the performance of the approaches from Section~\ref{sec:exdesign:hyper} on $O_{\operatorname{rt}}$-$O_{\operatorname{acc}}$-diagrams for the smaller hyper-parameter-tuning dataset.
@@
 If accuracy is strongly favored (very high $\delta$), the Wilcoxon stopping criterion performs better.
-The subset-size stopping criterion performs reasonably well but does not improve beyond a certain accuracy.
+The subset-size stopping criterion performs reasonably well but does not improve beyond a certain accuracy because of sampling a fixed subset of instances.
 
@@
 
-Having selected the most promising hyper-parameters, we run our end-to-end active-learning experiments on the complete Anniversary Track dataset (5355 instances).
+Having selected the most promising hyper-parameters, we run our active-learning experiments on the complete Anniversary Track dataset (5355 instances).
 The aforementioned range $\delta \in \left[0.2, 0.8\right]$ is covered by only two configurations.

### 2022-10-13 paper/paper.tex
Minor language revisions
... over full text, based on Grammarly and the desire to cut lines
containing only 1-3 words. Also, slightly reformulate the "Solver Model"
part (improve logical flow).

@@
   Benchmarking is one of the most important phases when developing algorithms.
-  This also applies to solutions of the SAT (propositional satisfiability) problem.
-  Benchmark selection chooses representative instances from a pool of instances such that they reliably discriminate SAT solvers based on their runtime.
+  This also applies to solutions to the SAT (propositional satisfiability) problem.
+  Benchmark selection chooses representative instances from a pool of instances that reliably discriminate SAT solvers based on their runtime.
   In this paper, we present a dynamic benchmark selection approach based on active learning.
@@
 One of the main phases of algorithm engineering is benchmarking.
-This is also true for the pro\-po\-si\-tio\-nal sat\-is\-fia\-bi\-li\-ty problem (SAT), the archetypal $\mathcal{NP}$-complete problem.
+This is also true for the propositional satisfiability problem (SAT), the archetypal $\mathcal{NP}$-complete problem.
 Benchmarking is, however, quite expensive regarding the cost of experiments.
@@
 Instead, the primary goal of traditional approaches is to select a representative set of instances for ranking solvers with a scoring scheme~\cite{Gelder11,HoosKSS13}.
-SAT~Competitions typically use the widely adopted PAR-2 score, i.e, the average runtime with a penalty of $2 \tau$ for timeouts with time limit~$\tau$~\cite{FroleyksHIJS21}.
+SAT~Competitions typically use the PAR-2 score, i.e., the average runtime with a penalty of $2 \tau$ for timeouts with time limit~$\tau$~\cite{FroleyksHIJS21}.
 
 In this paper, we present a novel benchmark selection approach based on active learning.
-Our approach is able to predict the rank of a new solver with high accuracy in only a fraction of the time needed to evaluate the full benchmark.
+Our approach can predict the rank of a new solver with high accuracy in only a fraction of the time needed to evaluate the full benchmark.
 The problem we solve is specified in Definition~\ref{def:new-solver-problem}.
@@
 Our approach satisfies several desirable criteria for benchmarking:
-Rather than outputting a binary classification, i.e, the new solver is worse than an existing solver or not, we provide a \emph{scoring} function that shows by which margin a solver is worse and how similar it is to existing solvers.
+Rather than outputting a binary classification, i.e., whether the new solver is worse than an existing solver or not, we provide a \emph{scoring} function that shows by which margin a solver is worse and how similar it is to existing solvers.
 In particular, our approach enables \emph{ranking} the new solver amidst a set of existing solvers.
@@
 Further, we optimize the \emph{runtime} that our strategy needs to arrive at its conclusion.
-We use problem-instance and runtime \emph{features}.
+We use instance and runtime \emph{features}.
 Moreover, we select instances \emph{non-randomly} and \emph{incrementally}.
-In particular, we take runtime information from already done experiments into account when choosing the next.
-By doing so, we can control the properties of the benchmarking approach such as its required runtime.
+In particular, we consider runtime information from already done experiments when choosing the next.
+By doing so, we can control the properties of the benchmarking approach, such as its required runtime.
 Our approach is \emph{scalable} in that it ranks a new solver $\hat{a}$ among any number of known solvers $\mathcal{A}$.
 In particular, we only subsample the benchmark once instead of comparing pairwise against each other solver~\cite{MatriconAFSH21}.
-Since a new solver idea is rarely best on the first try, it is desired to obtain a solver ranking fast.
-In this way, a new solver idea can be discarded if it performs poorly across the board or may be further tweaked if it shows promising results at least in some cases.
+Since a new solver idea is rarely best on the first try, it is desired to rank it fast.
+In this way, one can discard the new solver idea if it performs poorly across the board or may tweak it further if it shows promising results at least in some cases.
 
-We evaluate our approach with the SAT Competition~2022 Anniversary Track benchmark~\cite{sat2022}, consisting of 5355~instances and full runtimes of 28~solvers.
+We evaluate our approach with the SAT Competition~2022 Anniversary Track benchmark~\cite{sat2022}, consisting of 5355~instances and complete runtimes of 28~solvers.
 We perform cross-validation by treating each solver as new once and predicting these solvers' PAR-2 rank.
@@
 Benchmarking is not only of high interest in many fields but also an active research area on its own.
-Recent studies show that benchmark selection is a difficult endeavor for multiple reasons.
+Recent studies show that benchmark selection is challenging for multiple reasons.
 Biased benchmarks can easily lead to fallacious interpretations~\cite{abs-2107-07002}.
 Also, benchmark selection has many movable parts, e.g., performance measures, aggregation, and handling of missing values.
-Questionable research practices might modify these elements a-posteriori to fit expectations, leading to biased results~\cite{NiesslHWCB22}.
+Questionable research practices might modify these elements a-posteriori to fit expectations, biasing results~\cite{NiesslHWCB22}.
 In the following, we discuss related work from the areas of static benchmark selection, algorithm configuration, incremental benchmark selection, and active learning.
-We compare the most relevant approaches in Table~\ref{tab:requirements}.
+Table~\ref{tab:requirements} compares the most relevant approaches.
 
@@
 
-Benchmark selection is an important issue for competitions, e.g., the SAT Competition.
-In such competitions, the organizers define the rules for composing the corresponding benchmarks.
+Benchmark selection is essential for competitions, e.g., the SAT Competition.
+In such competitions, the organizers define the rules for composing the benchmarks.
 These selection strategies are mostly static, i.e., they do not depend on particular solvers to distinguish.
@@
 Froleyks et al. describe benchmark selection for the SAT~Competition~\cite{FroleyksHIJS21}.
-Manthey and Möhle find that competition benchmarks might contain redundant instances, and propose a feature-based approach to remove redundancy~\cite{manthey2016better}.
+Manthey and Möhle find that competition benchmarks might contain redundant instances and propose a feature-based approach to remove redundancy~\cite{manthey2016better}.
 M{\i}s{\i}r presents a feature-based approach to reduce benchmarks by matrix factorization and clustering~\cite{misir2021benchmark}.
@@
 For example, SMAC also employs an iterative, model-based selection procedure, though for configurations rather than instances.
-An algorithm configurator, however, cannot be used to \emph{rank/score} a new solver since algorithm configuration seeks to find solemnly the best performing configuration.
-Also, while using a model-based selection strategy to sample configurations, instance selection is done \emph{randomly}, i.e., without building a model over instances.
+An algorithm configurator, however, cannot be used to \emph{rank/score} a new solver since algorithm configuration solemnly seeks to find the best-performing configuration.
+Also, while using a model-based selection strategy to sample configurations, instance selection is made \emph{randomly}, i.e., without building a model over instances.
 
@@
 Given a pair of SAT solvers, they iteratively select a subset of instances until the desired confidence level is reached to decide which of the two solvers is better.
-The selection of instances is dependent on the choice of the solvers to distinguish.
+The selection of instances depends on the choice of the solvers to distinguish.
 They calculate a scoring metric for all unselected instances, run the experiment with the highest score, and update the confidence.
@@
 However, the approach only compares solvers binarily rather than providing a \emph{scoring}.
-Thus, it is not clear how similar two given solvers are or on which instances they behave similarly.
-Moreover, a major shortcoming is the lacking \emph{scalability} with the number of solvers.
+Thus, it is unclear how similar two given solvers are or on which instances they behave similarly.
+Moreover, a significant shortcoming is the lacking \emph{scalability} with the number of solvers.
 Comparing only pairs of solvers, evaluating a new solver requires sampling several benchmarks.
@@
 This goal is, however, orthogonal to ours~\cite{0001AEMN22,GarzonMG22}.
-While those approaches want to generate problem instances on which a solver is good or bad, we want to predict whether a solver is good or bad on an existing benchmark.
+While those approaches want to generate instances on which a solver is good or bad, we want to predict whether a solver is good or bad on an existing benchmark.
 Volpato and Guangyan use pool-based AL to learn an instance-specific algorithm selector~\cite{volpato2019active}.
@@
 Let benchmark instances $\mathcal{I}$, SAT solvers $\mathcal{A}$, and runtimes $r : \mathcal{A} \times \mathcal{I} \rightarrow \left[0, \tau\right]$ be given.
-Model initialization (Line~1) then creates the machine-learning prediction model as well as transforms runtimes as discussed in the subsequent section.
+We denote the new solver to be ranked by $\hat a \not\in A$ and define $\hat A := A \cup \lbrace \hat a \rbrace$ to denote all solvers, including the new solver. 
+Model initialization (Line~1) creates the prediction model and transforms runtimes, as discussed in the subsequent section.
 No model training is needed at this point since we do not have any labeled data yet.
-We denote the new solver to be ranked by $\hat a \not\in A$ and define $\hat A := A \cup \lbrace \hat a \rbrace$ to denote all solvers including the new solver. 
-Model $\mathcal{M}$ provides a prediction function $f : \mathcal{\hat A} \times \mathcal{I} \rightarrow \mathbb{R}$ that powers the decisions within our framework.
-The model update (Line~7) uses all acquired runtimes to train a new runtime-prediction function $f$ using common supervised machine-learning methodology.
-Training features are problem-instance features and known runtimes $r$.
-After applying necessary transformations, as described in the subsequent section, we fit the machine-learning model using the transformed observations $\mathcal{R}$ as ground-truth labels.
-It has also access to global instance features that assign each instance a vector in $\mathbb{R}^n$.
+Later, model $\mathcal{M}$ provides a runtime-label prediction function $f : \mathcal{\hat A} \times \mathcal{I} \rightarrow \mathbb{R}$ that powers the decisions within our framework.
+In particular, the model update (Line~7) uses all acquired runtimes to train $f$ with a standard machine-learning methodology.
+Training features are instance features and known runtimes $r$.
+After applying the necessary transformations described in the subsequent section, we fit the prediction model using the transformed observations $\mathcal{R}$ as ground-truth labels.
+It also has access to global instance features that assign each instance a vector in $\mathbb{R}^n$.
 Note that we build a new prediction model in each iteration since running experiments (Line~5) dominates the runtime of Algorithm~\ref{algALBenchmark} by magnitudes.
-Finally, we determine the solver scores that are predicted in Line~8 by using the runtime-prediction function $f$ and the approach described in the following two sections.
+Finally, we predict the solver scores in Line~8 with the prediction function~$f$ and the approach described in the following.
 
@@
 Selecting an instance based on the model is a core functionality of our framework (cf. Algorithm~\ref{algALBenchmark}, Line~4).
-In this section, we introduce our sampling strategies, which make use of the model's label-prediction function $f$ and are inspired by existing work within the realms of active learning~\cite{settles2009active}.
+In this section, we introduce our sampling strategies, which use the model's label-prediction function $f$ and are inspired by existing work within the realms of active learning~\cite{settles2009active}.
 We implement an uncertainty and an information-gain sampling strategy.
-These methods require the model's predictions to also include probabilities for the $k$ possible runtime labels.
+These methods require the model's predictions to include probabilities for the $k$ possible runtime labels.
 Let $f' : \mathcal{\hat A} \times \mathcal{I} \rightarrow \left[0, 1\right]^k$ denote this modified prediction function.
@@
 
-The uncertainty sampling strategy simply selects the instance that is closest to the model's decision boundary.
+The uncertainty sampling strategy simply selects the instance closest to the model's decision boundary.
 The set $\tilde{\mathcal{I}} \subseteq \mathcal{I}$ denotes the instances that have already been sampled.
@@
 
-The Wilcoxon stopping criterion stops the active-learning process when we are certain enough that the predicted runtime labels of the new solver are sufficiently different from the labels of the existing solvers.
+The Wilcoxon stopping criterion stops the active-learning process when we are confident enough that the predicted runtime labels of the new solver are sufficiently different from the labels of the existing solvers.
 This criterion is loosely inspired by Matricon et.~al.~\cite{MatriconAFSH21}.
@@
 %
-To improve the stability of this criterion, we use an exponential moving average to smooth out outliers and stop as soon as $W^{(i)}_{\exp}$ drops below a certain threshold:
+To improve the stability of this criterion, we use an exponential moving average to smooth out outliers and stop as soon as $W^{(i)}_{\exp}$ drops below a fixed threshold:
 %
@@
 
-Given all the previously presented instantiations for Algorithm~\ref{algALBenchmark}, this section briefly outlines our experimental design, including our evaluation framework, used data sets, hyper-parameter choices, as well as implementation details.
+Given all the previously presented instantiations for Algorithm~\ref{algALBenchmark}, this section briefly outlines our experimental design, including our evaluation framework, used data sets, hyper-parameter choices, and implementation details.
 
@@
 The dataset consists of 5355 instances with respective runtime data of 28 solvers.
-To characterize instances, we use 56 instance features from the \emph{base} database.
-They are inspired by the SATzilla~2012 features~\cite{features} and comprise instance-size features and graph-representation features among others.
+We use 56 instance features from the \emph{base} database to characterize instances.
+They are inspired by the SATzilla~2012 features~\cite{features} and comprise instance-size features and graph-representation features, among others.
 All features are numeric and free of missing values.
@@
 Additionally, we retrieve instance-family information to evaluate the composition of our sampled benchmarks.
-Instance families comprise instances derived from the same application domain, e.g., planning, cryptography, etc., and are a useful tool for both analyzing solver performance and portfolios.
+Instance families comprise instances derived from the same application domain, e.g., planning, cryptography, etc., and are a valuable tool for analyzing solver performance and portfolios.
 
@@
 Note that we are not considering all previously listed approaches since a grid search of all combinations would be infeasible.
-Rather, we filter approaches based on preliminary experimental results and do the experiments only with a subset.
+Instead, we filter approaches based on preliminary experimental results and do the experiments only with a subset.
 The experiment configurations are given below.
@@
     \item Fallback threshold: Use random sampling for the first \SI{0}{\%}, \SI{5}{\%}, \SI{10}{\%}, \SI{15}{\%}, or \SI{20}{\%} of instances to explore the instance space.
-    \item Runtime scaling: Whether to normalize uncertainty scores of instances by the average runtime of solvers on it or use the absolute values.
+    \item Runtime scaling: Whether to normalize uncertainty scores per instance by the average runtime of solvers on it or use the absolute values.
   \end{itemize}
@@
     \item Fallback threshold: Use random sampling for the first \SI{0}{\%}, \SI{5}{\%}, \SI{10}{\%}, \SI{15}{\%}, or \SI{20}{\%} of instances to explore the instance space.
-    \item Runtime scaling: Whether to normalize information-gain scores of instances by the average runtime of solvers on it or use the absolute values.
+    \item Runtime scaling: Whether to normalize information-gain scores per instance by the average runtime of solvers on it or use the absolute values.
   \end{itemize}
@@
 
-For runtime-label prediction, we only use one model configuration in our experiments since an exhaustive grid search would be infeasible.
-In preliminary experiments, we compared a variety of model types from \emph{scikit-learn}~\cite{scikit-learn}.
+We only use one model configuration for runtime-label prediction in our experiments since an exhaustive grid search would be infeasible.
+In preliminary experiments, we compared various model types from \emph{scikit-learn}~\cite{scikit-learn}.
 In particular, we conducted nested cross-validation, including hyper-parameter tuning, and used Matthews Correlation Coefficient~\cite{gorodkin2004comparing,matthews1975comparison} to assess the performance for predicting runtime labels.
@@
 Our experiments follow the evaluation framework introduced in Section~\ref{sec:exdesign:eval}.
-Fig.~\ref{fig:e2eallsolvers} shows the performance of the approaches from Section~\ref{sec:exdesign:hyper} on $O_{\operatorname{rt}}$-$O_{\operatorname{acc}}$-diagrams for the smaller hyper-parameter-tuning dataset.
+Fig.~\ref{fig:e2eallsolvers} shows the performance of the approaches from Section~\ref{sec:exdesign:hyper} on $O_{\operatorname{rt}}$-$O_{\operatorname{acc}}$-diagrams for the hyper-parameter-tuning dataset.
 Evaluating a particular configuration with Algorithm~\ref{alg:eval} returns a point $\left(O_{\operatorname{rt}},\, O_{\operatorname{acc}}\right)$.
@@
 
-Regarding ranking approaches (Fig.~\ref{fig:annitraincolorranking}), using the $s_3$-induced runtime-label ranking consistently outperforms the partially observed PAR-2 ranking for each possible value of the runtime-accuracy trade-off parameter~$\delta$.
-This is expected since selection decisions are not random.
+Regarding ranking approaches (Fig.~\ref{fig:annitraincolorranking}), using the $s_3$-induced runtime-label ranking consistently outperforms the partially observed PAR-2 ranking for each possible value of the trade-off parameter~$\delta$.
+This outcome is expected since selection decisions are not random.
 For example, we might sample more instances of one family if it benefits discrimination of solvers.
@@
 However, information-gain sampling is beneficial if runtime is strongly favored (small $\delta$; runtime fraction less than \SI{5}{\%}).
-This also aligns with our expectations:
+This result aligns with our expectations:
 Information-gain sampling selects instances that maximize the expected reduction in entropy.
-This means that we sample instances revealing similarities between solvers rather than differences.
-This helps in building a confident model fast.
+This means we sample instances revealing similarities between solvers rather than differences, which helps to build a confident model fast.
 However, the method lacks the possibility of selecting helpful instances for distinguishing solvers later.
-Random sampling performs quite well but is outperformed by uncertainty sampling in all cases, showing the benefit of actively selecting instances based on a prediction model.
+Random sampling performs reasonably well but is outperformed by uncertainty sampling in all cases, showing the benefit of actively selecting instances based on a prediction model.
 
@@
     Scatter plot comparing different instantiations of $\delta$ for our active-learning strategy on the hyper-parameter-tuning dataset.
-    The x-axis shows the fraction of runtime $O_{\operatorname{rt}}$ of the sample, while the y-axes show the fraction of instances sampled and ranking accuracy respectively.
+    The x-axis shows the fraction of runtime $O_{\operatorname{rt}}$ of the sample, while the y-axes show the fraction of instances sampled and ranking accuracy, respectively.
     The color indicates the weighting between different optimization goals $\delta \in \left[0, 1\right]$.
@@
 If we, on the one hand, desire to get a \emph{rough} estimate of a solver's performance fast (low $\delta$), approaches favor selecting many \emph{easy} instances.
-In particular, the fraction of sampled instances is bigger than the fraction of runtime.
+In particular, the fraction of sampled instances is larger than the fraction of runtime.
 By having many observations, it is easier to build a model.
@@
 We use the best-performing configurations for all $\delta \in \left[0, 1\right]$ and examine the selection decisions by the active-learning approach on the SAT Competition~2022 Anniversary Track dataset~\cite{sat2022}.
-While most families appear with the same fraction in the dataset and the sampled benchmarks, there are a few outliers that need further discussion.
+While most families appear with the same fraction in the dataset and the sampled benchmarks, a few outliers need further discussion.
 Problem instances of the families \emph{fpga}, \emph{quasigroup-completion}, and \emph{planning} are especially helpful to our framework in distinguishing solvers.
 Instances of these families are selected over-proportionally in comparison to the full dataset.
-In contrast to that, instances of the biggest family, i.e., \emph{hardware-verification}, roughly appear with the same fraction in the dataset and the sampled benchmarks.
-Finally, instances of the family \emph{cryptography} are less important in distinguishing solvers than their huge weight in the dataset suggests.
+In contrast, instances of the largest family, i.e., \emph{hardware-verification}, roughly appear with the same fraction in the dataset and the sampled benchmarks.
+Finally, instances of the family \emph{cryptography} are less important in distinguishing solvers than their vast weight in the dataset suggests.
 A possible explanation is that these instances are very similar, such that a small fraction of them is sufficient to estimate a solver's performance on all of them.
@@
 Given a new solver, we want to find its ranking amidst its competitors.
-Our approach provides accurate ranking predictions while needing significantly less runtime that a full evaluation on a given benchmark set.
-In particular, we can determine a new solver's PAR-2 ranking with about \SI{92}{\%} accuracy, while only needing \SI{10}{\%} of the full-evaluation time.
-We have found that our framework produces good results working with discrete runtime labels rather than regressing actual runtimes, albeit the former being more simple.
+Our approach provides accurate ranking predictions while needing significantly less runtime than a full evaluation on a given benchmark set.
+In particular, we can determine a new solver's PAR-2 ranking with about \SI{92}{\%} accuracy while only needing \SI{10}{\%} of the full-evaluation time.
+Our framework produces good results working with discrete runtime labels rather than regressing actual runtimes, albeit the former is more straightforward.
 We have evaluated several ranking algorithms, instance-selection approaches, and stopping criteria within our sequential active-learning process.
@@
 Future work may compare further ranking algorithms, instance-selection approaches, and stopping criteria.
-Furthermore, it is possible to formulate runtime discretization as an optimization problem, directly selecting the most discriminatory discretization technique rather than doing an empirical comparison.
+Furthermore, it is possible to formulate runtime discretization as an optimization problem, directly selecting the most discriminatory discretization technique rather than making an empirical comparison.
 
@@
 
-On a more general note, one can apply our evaluation framework to arbitrary $\mathcal{NP}$-complete problems as all discussed active-learning methods are problem-agnostic.
-Especially the runtime-prediction model would need some tweaking for this purpose though.
-Those problems share most of the relevant properties of SAT solving, i.e., there are established problem-instance features, a full benchmark run takes quite some time, and solver development traditionally requires expert knowledge to hand-select instances.
+On a more general note, one can apply our evaluation framework to arbitrary $\mathcal{NP}$-complete problems, as all discussed active-learning methods are problem-agnostic.
+Especially the runtime-prediction model would need some tweaking for this purpose, though.
+Those problems share most of the relevant properties of SAT solving, i.e., there are established instance features, a complete benchmark run takes quite some time, and solver development traditionally requires expert knowledge to hand-select instances.
 

### 2022-10-13 src/al_experiments.py
Minor code changes

@@ -421,7 +421,7 @@ if __name__ == "__main__":
         if DEBUG:
             solver_results = [
                 run_e2e_experiment(i, experiment, target_solver)
-                for i, target_solver in enumerate([column_list[0]])
+                for i, target_solver in enumerate(column_list)
             ]
         else:
             solver_results = Parallel(n_jobs=-1)(

### 2022-10-13 src/al_experiments/final_experiments.py

@@ -4,6 +4,7 @@ from al_experiments.experiment import Experiment
 from al_experiments.selection import (
     RandomSampling,
     SelectionFunctor,
+    MostFrequent,
     ModelUncertaintySampling,
     ModelBasedInformationGainSampling,
 )
@@ -169,3 +170,17 @@ all_experiments.append(Experiment.custom(
     WilcoxonSignificance(0.08, 0.05, 0.1),
     PartialObservationBasedRanking(),
 ))
+
+# all_experiments.append(Experiment.custom(
+#     "anni_final", False, 1,
+#     RandomSampling(1),
+#     FixedSubsetSize(0.103498, True),
+#     PartialObservationBasedRanking(),
+# ))
+
+# all_experiments.append(Experiment.custom(
+#     "anni_final", False, 1,
+#     MostFrequent(),
+#     FixedSubsetSize(0.103498, True),
+#     PartialObservationBasedRanking(),
+# ))

### 2022-10-13 src/al_experiments/selection.py

@@ -82,6 +82,48 @@ class RandomSampling(SelectionFunctor):
             y_sampled[sampled_index] = 1
 
 
+class MostFrequent(SelectionFunctor):
+    """ Select most frequently selected instances. """
+
+    def __init__(self) -> None:
+        super().__init__(0.0)
+
+    def get_params(self) -> str:
+        return f"most_freq"
+
+    def __call__(
+        self, y_sampled: np.ndarray, runtimes_df: pd.DataFrame,
+        clustering: pd.DataFrame, cluster_boundaries: pd.DataFrame,
+        timeout_clf: StackingClassifier, label_clf: StackingClassifier,
+        x: np.ndarray, y: np.ndarray, target_solver: str, experiment
+    ) -> None:
+        """ Selects another instance in 'y_sampled'.
+
+        Args:
+            y_sampled (np.ndarray): The already sampled instances.
+            runtimes_df (pd.DataFrame): The runtimes.
+            clustering (pd.DataFrame): The labels of known solvers.
+            cluster_boundaries (pd.DataFrame): The cluster boundaries.
+            timeout_clf (StackingClassifier): The time-out predictor.
+            label_clf (StackingClassifier): The non-time-out label predictor.
+            x (np.ndarray): The input features.
+            y (np.ndarray): The target labels.
+            target_solver (str): The target solver.
+            experiment (Experiment): The experiment configuration.
+        """
+
+        if np.count_nonzero(y_sampled) == y_sampled.shape[0]:
+            return
+
+        most_freq_hashes = pd.read_csv(
+            "../al-for-sat-solver-benchmarking-data/pickled-data/active_learning_instance_sampling_frequencies.csv",
+            index_col="hash"
+        ).index
+        hash_to_select = most_freq_hashes[np.count_nonzero(y_sampled)]
+        pos_to_select = np.where(runtimes_df.index == hash_to_select)[0][0]
+        y_sampled[pos_to_select] = 1
+
+
 class FastRuntimeSampling(SelectionFunctor):
     """ Select-fast-instances-first sampling. """
 

### 2022-10-13 src/al_experiments/stopping.py

@@ -46,14 +46,16 @@ class StoppingPredicate(ABC):
 class FixedSubsetSize(StoppingPredicate):
     """ Stop after a fixed amount of instances is in the subset. """
 
-    def __init__(self, subset_size: float) -> None:
+    def __init__(self, subset_size: float, use_fraction_runtime: bool = False) -> None:
         """ Initializes this stopping predicate.
 
         Args:
             subset_size (float): The desired subset size.
+            use_fraction_runtime (bool): Whether to sample the given amount of runtime.
         """
 
         self.subset_size = subset_size
+        self.use_fraction_runtime = use_fraction_runtime
 
     def get_params(self) -> str:
         return f"fixed_{self.subset_size:.2f}"
@@ -83,7 +85,18 @@ class FixedSubsetSize(StoppingPredicate):
             bool: Whether to stop adding more instances.
         """
 
-        return np.count_nonzero(y_sampled) / y_sampled.shape[0] >= self.subset_size
+        if not self.use_fraction_runtime:
+            return np.count_nonzero(y_sampled) / y_sampled.shape[0] >= self.subset_size
+        else:
+            # Data leakage, only for evaluation
+            amt = (
+                np.sum(runtimes_df.loc[
+                    y_sampled == 1,
+                    target_solver
+                ].replace(np.inf, 5000))
+                / np.sum(runtimes_df.loc[:, target_solver].replace(np.inf, 5000))
+            )
+            return amt >= self.subset_size
 
 
 class RankingConverged(StoppingPredicate):
@@ -136,7 +149,7 @@ class RankingConverged(StoppingPredicate):
             return True
 
 
-class WilcoxonSignificance(StoppingPredicate):  # TODO
+class WilcoxonSignificance(StoppingPredicate):
     """ Stop after we are confident enough to distinguish solvers. """
 
     def __init__(

### 2022-10-13 paper/paper.tex
section4.2

@@
 
-In our experiments, we work with a dataset based on the SAT~Competition~2022 Anniversary Track~\cite{sat2022}, retrieved from the GBD metadata database~\cite{IserS18}.
-The dataset consists of 5355 instances with respective runtime data of 28 solvers.
-We use 56 instance features from the \emph{base} database to characterize instances.
-They are inspired by the SATzilla~2012 features~\cite{features} and comprise instance-size features and graph-representation features, among others.
+In our experiments, we work with the dataset of the SAT~Competition~2022 Anniversary Track~\cite{sat2022}.
+The dataset consists of 5355 instances with respective runtime data of 28 sequential SAT solvers.
+We also use a database of 56 instance features\footnote{\url{https://gbd.iti.kit.edu/getdatabase/base_db}} which are provided in the GBD Benchmark Database~(GBD) by Iser~et~al~\cite{IserS18}.
+They comprise instance size features and node distribution statistics for several graph representations of SAT instances among others and are largely inspired by the SATzilla~2012 features~\cite{features}.
 All features are numeric and free of missing values.
@@
 Overall, prediction models have access to 46 instances features and 27 runtime features, i.e., excluding the current new solver~ $\hat{a}$.
-Additionally, we retrieve instance-family information to evaluate the composition of our sampled benchmarks.
+Additionally, we retrieve instance-family information\footnote{\url{https://gbd.iti.kit.edu/getdatabase/meta_db}} to evaluate the composition of our sampled benchmarks.
 Instance families comprise instances derived from the same application domain, e.g., planning, cryptography, etc., and are a valuable tool for analyzing solver performance and portfolios.

### 2022-10-13 paper/paper.tex
sweep over paper, many minor changes

@@
 This is also true for the propositional satisfiability problem (SAT), the archetypal $\mathcal{NP}$-complete problem.
-Benchmarking is, however, quite expensive regarding the cost of experiments.
-Competitive SAT solvers are the result of extensive experimentation and a variety of ideas and considerations~\cite{FroleyksHIJS21,sat2022}.
+Benchmarking is, however, quite expensive regarding the runtime of experiments.
+Competitive SAT solvers are the result of extensive experimentation with a variety of ideas and considerations~\cite{FroleyksHIJS21,sat2022}.
 While the field of SAT solver benchmarking is well established, traditional benchmark selection approaches do not optimize for benchmark runtime.
 Instead, the primary goal of traditional approaches is to select a representative set of instances for ranking solvers with a scoring scheme~\cite{Gelder11,HoosKSS13}.
-SAT~Competitions typically use the PAR-2 score, i.e., the average runtime with a penalty of $2 \tau$ for timeouts with time limit~$\tau$~\cite{FroleyksHIJS21}.
+SAT~Competitions typically use the \mbox{PAR-2} score, i.e., the average runtime with a penalty of $2 \tau$ for timeouts with time-limit~$\tau$~\cite{FroleyksHIJS21}.
 
@@
 \begin{definition}[New-Solver Problem]
-  Given a pool of instances~$\mathcal{I}$, solvers~$\mathcal{A}$, and runtimes~$r\!: \mathcal{A} \times \mathcal{I} \rightarrow \left[0, \tau\right]$ with time limit $\tau$, maximize the confidence in predicting the rank of a new solver $\hat{a} \notin \mathcal{A}$ while minimizing the total benchmark runtime by incrementally selecting instances in $\mathcal{I}$ to evaluate $\hat{a}$.
+  Given solvers~$\mathcal{A}$, instances~$\mathcal{I}$, runtimes~$r\!: \mathcal{A} \times \mathcal{I} \rightarrow \left[0, \tau\right]$ with time-limit $\tau$, and a new solver $\hat{a} \notin \mathcal{A}$, incrementally select benchmark instances in $\mathcal{I}$ to maximize the confidence in predicting the rank of $\hat{a}$ while minimizing the total benchmark runtime.
   \label{def:new-solver-problem}
@@
 Balint et al. provide an overview of benchmark-selection criteria in different solver competitions~\cite{balint2015overview}.
-Froleyks et al. describe benchmark selection for the SAT~Competition~\cite{FroleyksHIJS21}.
+Froleyks et al. describe benchmark selection as it is done in recent SAT~competitions~\cite{FroleyksHIJS21}.
 Manthey and Möhle find that competition benchmarks might contain redundant instances and propose a feature-based approach to remove redundancy~\cite{manthey2016better}.
 M{\i}s{\i}r presents a feature-based approach to reduce benchmarks by matrix factorization and clustering~\cite{misir2021benchmark}.
+
 Hoos et al.~\cite{HoosKSS13} discuss which properties are most desirable when selecting SAT benchmark instances.
-Selection criteria are instance variety to avoid over-fitting, adapted instance hardness (not too easy but also not too hard), and avoiding duplicate instances.
+Selection criteria are instance variety to avoid over-fitting, adapted instance hardness (not too easy but also not too hard), and avoidance of duplicate instances.
 To filter too similar instances, they use a distance-based approach with the SATzilla features~\cite{XuHHL08,features}.
@@
 Moreover, a significant shortcoming is the lacking \emph{scalability} with the number of solvers.
-Comparing only pairs of solvers, evaluating a new solver requires sampling several benchmarks.
+Comparing only pairs of solvers, evaluating a new solver requires sampling several benchmarks for each pair.
 In contrast, our approach allows comparing a new solver against a set of existing solvers with one benchmark.
 
-\subsubsection{Active Learning (AL).}
+\subsubsection{Active Learning.}
 
-The posed \emph{New-Solver Problem} has stark similarities to the well-studied field of active learning AL within recommender systems, especially the \emph{new-user problem}~\cite{RubensESK15}.
-On the one hand, we want to maximize the utility an instance provides to our model and, on the other hand, minimize the cost (CPU time) that is associated with its acquisition.
+The posed \emph{New-Solver Problem} has stark similarities to the well-studied field of active learning (AL) within recommender systems, in particular the \emph{new-user problem}~\cite{RubensESK15}.
+On the one hand, we want to maximize the utility an instance provides to our model, i.e., rank prediction accuracy, and on the other hand, minimize the cost, i.e., runtime, that is associated with its acquisition.
 In contrast to traditional passive machine-learning methods with given instance labels, active learning allows for selecting instances for which to acquire labels.
@@
 Volpato and Guangyan use pool-based AL to learn an instance-specific algorithm selector~\cite{volpato2019active}.
-Rather than benchmarking a solver's overall performance, the goal is to recommend the best solver out of a set of solvers for each SAT instance.
+Rather than benchmarking a solver's overall performance, their goal is to recommend the best solver out of a set of solvers for each SAT instance.
 
@@
 
-Let benchmark instances $\mathcal{I}$, SAT solvers $\mathcal{A}$, and runtimes $r : \mathcal{A} \times \mathcal{I} \rightarrow \left[0, \tau\right]$ be given.
+Let SAT solvers $\mathcal{A}$, benchmark instances $\mathcal{I}$, and runtimes $r : \mathcal{A} \times \mathcal{I} \rightarrow \left[0, \tau\right]$ be given.
 We denote the new solver to be ranked by $\hat a \not\in A$ and define $\hat A := A \cup \lbrace \hat a \rbrace$ to denote all solvers, including the new solver. 
@@
 Later, model $\mathcal{M}$ provides a runtime-label prediction function $f : \mathcal{\hat A} \times \mathcal{I} \rightarrow \mathbb{R}$ that powers the decisions within our framework.
-In particular, the model update (Line~7) uses all acquired runtimes to train $f$ with a standard machine-learning methodology.
+In particular, the model update (Line~7) uses all acquired runtimes to train $f$ with a standard supervised machine-learning methodology.
 Training features are instance features and known runtimes $r$.
 After applying the necessary transformations described in the subsequent section, we fit the prediction model using the transformed observations $\mathcal{R}$ as ground-truth labels.
-It also has access to global instance features that assign each instance a vector in $\mathbb{R}^n$.
 Note that we build a new prediction model in each iteration since running experiments (Line~5) dominates the runtime of Algorithm~\ref{algALBenchmark} by magnitudes.
-Finally, we predict the solver scores in Line~8 with the prediction function~$f$ and the approach described in the following.
+Finally, we predict the score of the new solver $\hat a$ in Line~8 with the prediction function~$f$.
 
@@
 We implement an uncertainty and an information-gain sampling strategy.
-These methods require the model's predictions to include probabilities for the $k$ possible runtime labels.
-Let $f' : \mathcal{\hat A} \times \mathcal{I} \rightarrow \left[0, 1\right]^k$ denote this modified prediction function.
+These methods require the model's predictions to include probabilities for the $k$ discrete runtime labels induced by the clustering described in the preceeding section.
+Let \mbox{$f' : \mathcal{\hat A} \times \mathcal{I} \rightarrow \left[0, 1\right]^k$} denote this modified prediction function.
 
@@
 This criterion is loosely inspired by Matricon et.~al.~\cite{MatriconAFSH21}.
-We use the average $p$-value $W_{\hat{a}}$ of a Wilcoxon signed-rank test $w(S,P)$ of the two runtime label distributions $S=\{ \gamma(a, e) \mid e \in \mathcal{I} \}$ for an existing solver $a$ and $P=\{ f(\hat a, e) \mid e \in \mathcal{I} \}$ for the new solver $\hat{a}$:
+We use the average $p$-value $W_{\hat{a}}$ of a Wilcoxon signed-rank test $w(S,P)$ of the two runtime label distributions $S=\{ \gamma(a, e) \mid e \in \mathcal{I} \}$ for an existing solver $a$ and \mbox{$P=\{ f(\hat a, e) \mid e \in \mathcal{I} \}$} for the new solver $\hat{a}$:
 %
@@
 It stops the active-learning process if the ranking induced by the model's predictions (Equation~\ref{eq:rankingeq}) remains unchanged within the last $l$ iterations.
-Note that the concrete values of $s_k(\hat a)$ might still change.
+Note that the concrete values of the predicted score $s_{\hat a}$ might still change.
 We are solemnly interested in the induced ranking in this case.
@@
 Also, there are different choices for the runtime-label prediction model and runtime discretization.
-Note that we are not considering all previously listed approaches since a grid search of all combinations would be infeasible.
-Instead, we filter approaches based on preliminary experimental results and do the experiments only with a subset.
 The experiment configurations are given below.
@@
 Since the potential runtime of experiments is by magnitudes larger than the model's update time, we only consider incrementing our benchmark by one instance at a time rather than using batches, which is also proposed in current active-learning advances~\cite{SinhaED19,2019gaal}.
-A drawback of this is the lack of parallelization.
+A drawback of this is the lack of parallel execution of runtime experiment.
 
@@
 Evaluating a particular configuration with Algorithm~\ref{alg:eval} returns a point $\left(O_{\operatorname{rt}},\, O_{\operatorname{acc}}\right)$.
-In particular, we do not show intermediate results of the active-learning procedure but only the final results after stopping.
+We do not show intermediate results of the active-learning procedure but only the final results after stopping.
 The plotted lines represent the best performing configurations (convex hull) per top-level hyper-parameter choice, i.e., per ranking approach (Fig.~\ref{fig:annitraincolorranking}), selection approach (Fig.~\ref{fig:annitraincolorselection}), and stopping criterion (Fig.~\ref{fig:annitraincolorstopping}).
@@
 In particular, we can determine a new solver's PAR-2 ranking with about \SI{92}{\%} accuracy while only needing \SI{10}{\%} of the full-evaluation time.
-Our framework produces good results working with discrete runtime labels rather than regressing actual runtimes, albeit the former is more straightforward.
+Our framework produces good results working with discrete runtime labels rather than regressing actual runtimes.
 We have evaluated several ranking algorithms, instance-selection approaches, and stopping criteria within our sequential active-learning process.

### 2022-10-13 paper/paper.tex
Extend evaluation in 5.2

@@
 It can predict a new solver's PAR-2 ranking with about \SI{92.33}{\%} accuracy ($O_{\operatorname{acc}}$), while only needing \SI{10.35}{\%} of the full evaluation time ($O_{\operatorname{rt}}$).
-These trade-offs provide a better way of prototyping and evaluating new SAT solvers rather than evaluating them on manually picked instances based on domain knowledge.
+
+\vspace{-0.5cm}
+\begin{table}[h]
+  \centering
+  \caption{
+      Performance comparison of the best-performing active-learning approaches (\emph{AL}), random sampling of the same runtime fraction with 1000 repetitions (\emph{Random}) and statically selecting the instances that are most frequently sampled by active-learning approaches (\emph{Most Freq.})
+    }
+  \label{tab:fulldataset}
+
+  \begin{subfigure}{\textwidth}
+    \centering
+    \caption{Best-performing AL approach for $\delta \in \left[0.2, 0.7\right]$}
+    \begin{tabular}{
+      >{\arraybackslash}p{0.4\textwidth}
+      S[table-format=2.2,table-column-width=0.15\textwidth]
+      S[table-format=2.2,table-column-width=0.15\textwidth]
+      S[table-format=2.2,table-column-width=0.15\textwidth]
+    }
+      \hline
+        & {AL} & {Random} & {Most Freq.} \\
+      \hline
+      Sampled Runtime Fraction (\%) & 5.41 & 5.43 & 5.44 \\
+      Sampled Instance Fraction (\%) & 26.53 & 5.43 & 27.75 \\
+      Ranking Accuracy (\%) & 90.48 & 88.54 & 81.08 \\
+      \hline
+    \end{tabular}
+  \end{subfigure}
+  \\
+  \begin{subfigure}{\textwidth}
+    \centering
+    \caption{Best-performing AL approach for $\delta \in (0.7, 0.8]$}
+    \begin{tabular}{
+      >{\arraybackslash}p{0.4\textwidth}
+      S[table-format=2.2,table-column-width=0.15\textwidth]
+      S[table-format=2.2,table-column-width=0.15\textwidth]
+      S[table-format=2.2,table-column-width=0.15\textwidth]
+    }
+      \hline
+        & {AL} & {Random} & {Most Freq.} \\
+      \hline
+      Sampled Runtime Fraction (\%) & 10.35 & 10.37 & 10.37 \\
+      Sampled Instance Fraction (\%) & 5.24 & 10.37 & 36.96 \\
+      Ranking Accuracy (\%) & 92.33 & 91.61 & 84.52 \\
+      \hline
+    \end{tabular}
+  \end{subfigure}
+\end{table}
+
+Table~\ref{tab:fulldataset} shows how both active-learning approaches compare against static baselines.
+The first column shows the two aforementioned AL configurations.
+The random baseline samples instances until roughly the same fraction of runtime is in the benchmark set.
+We repeat this 1000 times and report average results.
+Our approaches perform slightly better in both cases.
+However, differences are not significant regarding a Wilcoxon signed-rank test with $\alpha = 0.05$.
+Finally, we create a static benchmark set with the most frequently selected instances by active learning.
+This benchmark performs poorly, though, in comparison to active learning and random sampling.
+While active-learning results are less strong on the full dataset than on the smaller tuning dataset, it still shows the benefit of making benchmark selection dependent on the solvers to distinguish.
 

### 2022-10-13 src/al_experiments.py

@@ -426,10 +426,10 @@ if __name__ == "__main__":
         else:
             solver_results = Parallel(n_jobs=-1)(
                 delayed(run_e2e_experiment)(
-                    i, experiment, target_solver
+                    i * experiment.repetitions + j, experiment, target_solver
                 )
                 for i, target_solver in enumerate(column_list)
-                for _ in range(experiment.repetitions)
+                for j in range(experiment.repetitions)
             )
 
         # Store results

### 2022-10-13 src/al_experiments/final_experiments.py

@@ -172,7 +172,21 @@ all_experiments.append(Experiment.custom(
 ))
 
 # all_experiments.append(Experiment.custom(
+#     "anni_final", False, 1000,
+#     RandomSampling(1),
+#     FixedSubsetSize(0.054083, True),
+#     PartialObservationBasedRanking(),
+# ))
+
+# all_experiments.append(Experiment.custom(
 #     "anni_final", False, 1,
+#     MostFrequent(),
+#     FixedSubsetSize(0.054083, True),
+#     PartialObservationBasedRanking(),
+# ))
+
+# all_experiments.append(Experiment.custom(
+#     "anni_final", False, 1000,
 #     RandomSampling(1),
 #     FixedSubsetSize(0.103498, True),
 #     PartialObservationBasedRanking(),

### 2022-10-13 paper/paper.tex
Regenerate pdf
Minor changes
... after reviewing commits a69316d265 and c6d06e8d51.

@@
 Moreover, a significant shortcoming is the lacking \emph{scalability} with the number of solvers.
-Comparing only pairs of solvers, evaluating a new solver requires sampling several benchmarks for each pair.
-In contrast, our approach allows comparing a new solver against a set of existing solvers with one benchmark.
+Comparing only pairs of solvers, evaluating a new solver requires sampling a separate benchmark for each existing solver.
+In contrast, our approach allows comparing a new solver against a set of existing solvers by sampling only one benchmark.
 
@@
 We implement an uncertainty and an information-gain sampling strategy.
-These methods require the model's predictions to include probabilities for the $k$ discrete runtime labels induced by the clustering described in the preceeding section.
+These methods require the model's predictions to include probabilities for the $k$ discrete runtime labels induced by the clustering described in the preceding section.
 Let \mbox{$f' : \mathcal{\hat A} \times \mathcal{I} \rightarrow \left[0, 1\right]^k$} denote this modified prediction function.
@@
 The dataset consists of 5355 instances with respective runtime data of 28 sequential SAT solvers.
-We also use a database of 56 instance features\footnote{\url{https://gbd.iti.kit.edu/getdatabase/base_db}} which are provided in the GBD Benchmark Database~(GBD) by Iser~et~al~\cite{IserS18}.
-They comprise instance size features and node distribution statistics for several graph representations of SAT instances among others and are largely inspired by the SATzilla~2012 features~\cite{features}.
+We also use a database of 56 instance features\footnote{\url{https://gbd.iti.kit.edu/getdatabase/base_db}} provided in the GBD Benchmark Database~(GBD) by Iser~et~al.~\cite{IserS18}.
+They comprise instance size features and node distribution statistics for several graph representations of SAT instances, among others, and are primarily inspired by the SATzilla~2012 features~\cite{features}.
 All features are numeric and free of missing values.
@@
 
-\vspace{-0.5cm}
 \begin{table}[h]
@@
 
-\begin{figure}[tb!]
+Selection decisions of our approach also reveal the importance of different instance families to our framework.
+Figure~\ref{fig:annifinalfamilies} shows the occurrence of instance families within the dataset and the benchmarks created by active learning.
+We use the best-performing configurations for all $\delta \in \left[0, 1\right]$ and examine the selection decisions by the active-learning approach on the SAT Competition~2022 Anniversary Track dataset~\cite{sat2022}.
+While most families appear with the same fraction in the dataset and the sampled benchmarks, a few outliers need further discussion.
+Problem instances of the families \emph{fpga}, \emph{quasigroup-completion}, and \emph{planning} are especially helpful to our framework in distinguishing solvers.
+Instances of these families are selected over-proportionally in comparison to the full dataset.
+In contrast, instances of the largest family, i.e., \emph{hardware-verification}, roughly appear with the same fraction in the dataset and the sampled benchmarks.
+Finally, instances of the family \emph{cryptography} are less important in distinguishing solvers than their vast weight in the dataset suggests.
+A possible explanation is that these instances are very similar, such that a small fraction of them is sufficient to estimate a solver's performance on all of them.
+
+\begin{figure}[htb!]
   \centering
@@
 \end{figure}
-
-Selection decisions of our approach also reveal the importance of different instance families to our framework.
-Figure~\ref{fig:annifinalfamilies} shows the occurrence of instance families within the dataset and the benchmarks created by active learning.
-We use the best-performing configurations for all $\delta \in \left[0, 1\right]$ and examine the selection decisions by the active-learning approach on the SAT Competition~2022 Anniversary Track dataset~\cite{sat2022}.
-While most families appear with the same fraction in the dataset and the sampled benchmarks, a few outliers need further discussion.
-Problem instances of the families \emph{fpga}, \emph{quasigroup-completion}, and \emph{planning} are especially helpful to our framework in distinguishing solvers.
-Instances of these families are selected over-proportionally in comparison to the full dataset.
-In contrast, instances of the largest family, i.e., \emph{hardware-verification}, roughly appear with the same fraction in the dataset and the sampled benchmarks.
-Finally, instances of the family \emph{cryptography} are less important in distinguishing solvers than their vast weight in the dataset suggests.
-A possible explanation is that these instances are very similar, such that a small fraction of them is sufficient to estimate a solver's performance on all of them.
 

### 2022-10-13 paper/paper.tex
Revise 5.2

@@
   \centering
+  \caption{Comparison of features of our benchmark-selection approach, the static benchmark-selection approach by Hoos~et~al.~\cite{HoosKSS13}, the algorithm configuration system SMAC~\cite{HutterHL11}, and the active-learning approaches by Matricon et al.~\cite{MatriconAFSH21}.
+  }
+  \label{tab:requirements}
+  ~\\[1em]
   \begin{tabular}{
@@
   \end{tabular}
-  ~\\[1em]
-  \caption{Comparison of features of our benchmark-selection approach, the static benchmark-selection approach by Hoos~et~al.~\cite{HoosKSS13}, the algorithm configuration system SMAC~\cite{HutterHL11}, and the active-learning approaches by Matricon et al.~\cite{MatriconAFSH21}.
-  }
-  \label{tab:requirements}
 \end{table}
@@
 
-\begin{table}[h]
+\begin{table}[t]
   \centering
@@
 The first column shows the two aforementioned AL configurations.
-The random baseline samples instances until roughly the same fraction of runtime is in the benchmark set.
+We compare it against a random baseline and a static benchmark set consisting of the most frequently sampled instances by active learning.
+The random baseline (\emph{random}) samples instances until roughly the same fraction of runtime is in the benchmark set.
 We repeat this 1000 times and report average results.
-Our approaches perform slightly better in both cases.
+The benchmark set of the most frequently AL-sampled instances (\emph{Most Freq.}) uses the average sampling frequencies over all solvers using all Pareto-optimal active-learning approaches.
+
+Our AL approaches perform slightly better compared to random sampling.
 However, differences are not significant regarding a Wilcoxon signed-rank test with $\alpha = 0.05$.
-Finally, we create a static benchmark set with the most frequently selected instances by active learning.
-This benchmark performs poorly, though, in comparison to active learning and random sampling.
-While active-learning results are less strong on the full dataset than on the smaller tuning dataset, it still shows the benefit of making benchmark selection dependent on the solvers to distinguish.
+Note that the performance difference between random and non-random sampling changes for different fractions of sampled runtime (cf. Figure~\ref{fig:annitraincolorselection}).
+The static benchmark using the most frequently AL-sampled instances performs poorly, though, in comparison to active learning and random sampling.
+This is somewhat expected since it does not reflect the right balance of instance families:
+Families whose instances are uniform-randomly selected by AL appear less often in this benchmark than families where some instances are sampled more often than others.
+While the active-learning results are less strong on the full dataset than on the smaller tuning dataset, it still shows the benefit of making benchmark selection dependent on the solvers to distinguish.
 
@@
 
-Selection decisions of our approach also reveal the importance of different instance families to our framework.
-Figure~\ref{fig:annifinalfamilies} shows the occurrence of instance families within the dataset and the benchmarks created by active learning.
-We use the best-performing configurations for all $\delta \in \left[0, 1\right]$ and examine the selection decisions by the active-learning approach on the SAT Competition~2022 Anniversary Track dataset~\cite{sat2022}.
-While most families appear with the same fraction in the dataset and the sampled benchmarks, a few outliers need further discussion.
-Problem instances of the families \emph{fpga}, \emph{quasigroup-completion}, and \emph{planning} are especially helpful to our framework in distinguishing solvers.
-Instances of these families are selected over-proportionally in comparison to the full dataset.
-In contrast, instances of the largest family, i.e., \emph{hardware-verification}, roughly appear with the same fraction in the dataset and the sampled benchmarks.
-Finally, instances of the family \emph{cryptography} are less important in distinguishing solvers than their vast weight in the dataset suggests.
-A possible explanation is that these instances are very similar, such that a small fraction of them is sufficient to estimate a solver's performance on all of them.
-
-\begin{figure}[htb!]
+\begin{figure}[tb]
   \centering
@@
 \end{figure}
+
+Selection decisions of our approach also reveal the importance of different instance families to our framework.
+Figure~\ref{fig:annifinalfamilies} shows the occurrence of instance families within the dataset and the benchmarks created by active learning.
+We use the best-performing configurations for all $\delta \in \left[0, 1\right]$ and examine the selection decisions by the active-learning approach on the SAT Competition~2022 Anniversary Track dataset~\cite{sat2022}.
+While most families appear with the same fraction in the dataset and the sampled benchmarks, a few outliers need further discussion.
+Problem instances of the families \emph{fpga}, \emph{quasigroup-completion}, and \emph{planning} are especially helpful to our framework in distinguishing solvers.
+Instances of these families are selected over-proportionally in comparison to the full dataset.
+In contrast, instances of the largest family, i.e., \emph{hardware-verification}, roughly appear with the same fraction in the dataset and the sampled benchmarks.
+Finally, instances of the family \emph{cryptography} are less important in distinguishing solvers than their vast weight in the dataset suggests.
+A possible explanation is that these instances are very similar, such that a small fraction of them is sufficient to estimate a solver's performance on all of them.
 

### 2022-10-13 paper/paper.tex
Anonymize paper
Revert after review.

@@
 \author{
-  Tobias Fuchs\orcidID{0000-0001-9727-2878} (\Letter) \and \\
-  Jakob Bach\orcidID{0000-0003-0301-2798} \and \\
-  Markus Iser\orcidID{0000-0003-2904-232X}
+  Authors anonymized for review
+  \vspace{0.4cm}
 }
@@
 \institute{
-  Karlsruhe Institute of Technology (KIT), Karlsruhe, Germany \\
-  \email{info@tobiasfuchs.de, jakob.bach@kit.edu, markus.iser@kit.edu}
+  Institute anonymized for review
+  \vspace{0.4cm}
 }
 
-\authorrunning{T. Fuchs et al.} % according to sample paper of template: first names are abbreviated; if there are more than two authors, 'et al.' is used.
+\authorrunning{Authors anonymized for review} % according to sample paper of template: first names are abbreviated; if there are more than two authors, 'et al.' is used.
 
@@
 
-% TODO: insert URLs
-All our source code\footnote{temporary, anonymized version for review: \url{xxx}} and and data\footnote{temporary, anonymized version for review: \url{xxx}} are available on GitHub.
+All our source code\footnote{temporary reference, anonymized version for review: \url{https://github.com/dgnr7vexskfoc3a67jdtf96kd9opz/al-for-sat-solver-benchmarking}} and and data\footnote{temporary reference, anonymized version for review: \url{https://github.com/dgnr7vexskfoc3a67jdtf96kd9opz/al-for-sat-solver-benchmarking-data}} are available on GitHub.
 
@@
 
-\begin{algorithm}
+\begin{algorithm}[t]
   \caption{Incremental Benchmarking Framework}
@@
 
-% TODO: insert URLs
-For reproducibility, our source code and data are available on GitHub\footnote{temporary, anonymized version for review: \url{xxx}}.
+For reproducibility, our source code and data are available on GitHub (cf. footnotes in Section~\ref{sec:intro}).
 Our code is implemented in \textsc{Python} using \emph{scikit-learn}~\cite{scikit-learn} for making predictions and \emph{gbd-tools}~\cite{IserS18} for SAT-instance retrieval.

### 2022-10-14 paper/paper.tex
Revise Section 5.2
... based on commit ba8ab2dc54. Adapt language.

@@
   \caption{
-      Performance comparison of the best-performing active-learning approaches (\emph{AL}), random sampling of the same runtime fraction with 1000 repetitions (\emph{Random}) and statically selecting the instances that are most frequently sampled by active-learning approaches (\emph{Most Freq.})
+      Performance comparison of the best-performing active-learning approaches (\emph{AL}), random sampling of the same runtime fraction with 1000 repetitions (\emph{Random}), and statically selecting the instances most frequently sampled by active-learning approaches (\emph{Most Freq.})
     }
@@
 The first column shows the two aforementioned AL configurations.
-We compare it against a random baseline and a static benchmark set consisting of the most frequently sampled instances by active learning.
-The random baseline (\emph{random}) samples instances until roughly the same fraction of runtime is in the benchmark set.
-We repeat this 1000 times and report average results.
-The benchmark set of the most frequently AL-sampled instances (\emph{Most Freq.}) uses the average sampling frequencies over all solvers using all Pareto-optimal active-learning approaches.
+We compare them against a random baseline and a static benchmark set consisting of the instances most frequently sampled by active learning.
+The random baseline (\emph{Random}) samples instances until it reaches roughly the same fraction of runtime as the AL benchmark sets.
+We repeat this procedure 1000 times and report average results.
+The benchmark set of the most frequently AL-sampled instances (\emph{Most Freq.}) uses the average sampling frequencies of instances over all solvers and all Pareto-optimal active-learning approaches.
 
-Our AL approaches perform slightly better compared to random sampling.
+Our two AL approaches both perform better than random sampling.
 However, differences are not significant regarding a Wilcoxon signed-rank test with $\alpha = 0.05$.
 Note that the performance difference between random and non-random sampling changes for different fractions of sampled runtime (cf. Figure~\ref{fig:annitraincolorselection}).
-The static benchmark using the most frequently AL-sampled instances performs poorly, though, in comparison to active learning and random sampling.
-This is somewhat expected since it does not reflect the right balance of instance families:
-Families whose instances are uniform-randomly selected by AL appear less often in this benchmark than families where some instances are sampled more often than others.
-While the active-learning results are less strong on the full dataset than on the smaller tuning dataset, it still shows the benefit of making benchmark selection dependent on the solvers to distinguish.
+The static benchmark using the most frequently AL-sampled instances performs poorly, though, compared to active learning and random sampling.
+This outcome is somewhat expected since the static benchmark does not reflect the right balance of instance families:
+Families whose instances are uniform-randomly selected by AL, e.g., for different solvers, appear less often in this benchmark than families where some instances are sampled more often than others.
+While the active-learning results are less strong on the full dataset than on the smaller tuning dataset, they still show the benefit of making benchmark selection dependent on the solvers to distinguish.
 

### 2022-10-14 paper/paper.tex
Minor language revisions in whole text
Mostly in abstract, intro, and conclusion. Problem definition appears on
1st page of paper now. Also, fixed a URL field in the bibliography.

@@
 \author{
-  Authors anonymized for review
+  Author(s) anonymized for review
   \vspace{0.4cm}
@@
 \institute{
-  Institute anonymized for review
+  Institute(s) anonymized for review
   \vspace{0.4cm}
@@
 
-\authorrunning{Authors anonymized for review} % according to sample paper of template: first names are abbreviated; if there are more than two authors, 'et al.' is used.
+\authorrunning{Author(s) anonymized for review} % according to sample paper of template: first names are abbreviated; if there are more than two authors, 'et al.' is used.
 
@@
   Benchmarking is one of the most important phases when developing algorithms.
-  This also applies to solutions to the SAT (propositional satisfiability) problem.
-  Benchmark selection chooses representative instances from a pool of instances that reliably discriminate SAT solvers based on their runtime.
+  This also applies to solvers for the SAT (propositional satisfiability) problem.
+  Benchmark selection chooses representative problem instances that reliably discriminate solvers based on their runtime.
   In this paper, we present a dynamic benchmark selection approach based on active learning.
-  Our approach predicts the rank of a new solver among its competitors with minimum running time and maximum rank prediction accuracy.
-  We evaluated this approach using the Anniversary Track dataset from the 2022 SAT Competition.
+  Our approach predicts the rank of a new solver among its competitors with minimum runtime and maximum rank prediction accuracy.
+  We evaluated this approach on the Anniversary Track dataset from the 2022 SAT Competition.
   Our selection approach can predict the rank of a new solver after about \SI{10}{\%} of the time it would take to run the solver on all instances of this dataset, with a prediction accuracy of about \SI{92}{\%}.
-  In this paper, we also discuss the importance of instance families in the selection process. 
-  Overall, our tool provides a reliable way for SAT solver engineers to efficiently determine the performance of a new SAT solver.
+  We also discuss the importance of instance families in the selection process. 
+  Overall, our tool provides a reliable way for solver engineers to efficiently determine the performance of a new solver.
 
@@
 While the field of SAT solver benchmarking is well established, traditional benchmark selection approaches do not optimize for benchmark runtime.
-Instead, the primary goal of traditional approaches is to select a representative set of instances for ranking solvers with a scoring scheme~\cite{Gelder11,HoosKSS13}.
+Instead, traditional approaches focus on selecting a representative set of instances for ranking solvers with a scoring scheme~\cite{Gelder11,HoosKSS13}.
 SAT~Competitions typically use the \mbox{PAR-2} score, i.e., the average runtime with a penalty of $2 \tau$ for timeouts with time-limit~$\tau$~\cite{FroleyksHIJS21}.
@@
 \begin{definition}[New-Solver Problem]
-  Given solvers~$\mathcal{A}$, instances~$\mathcal{I}$, runtimes~$r\!: \mathcal{A} \times \mathcal{I} \rightarrow \left[0, \tau\right]$ with time-limit $\tau$, and a new solver $\hat{a} \notin \mathcal{A}$, incrementally select benchmark instances in $\mathcal{I}$ to maximize the confidence in predicting the rank of $\hat{a}$ while minimizing the total benchmark runtime.
+  Given solvers~$\mathcal{A}$, instances~$\mathcal{I}$, runtimes~$r\!: \mathcal{A} \times \mathcal{I} \rightarrow \left[0, \tau\right]$ with time-limit $\tau$, and a new solver $\hat{a} \notin \mathcal{A}$, incrementally select benchmark instances from $\mathcal{I}$ to maximize the confidence in predicting the rank of $\hat{a}$ while minimizing the total benchmark runtime.
   \label{def:new-solver-problem}
@@
 
-Note that our scenario assumes that we know the runtimes of all solvers, except the new one, on all instances.
+Note that our scenario assumes knowing the runtimes of all solvers, except the new one, on all instances.
 One could also imagine a collaborative filtering scenario, where runtimes are only partially known~\cite{misir2017data,misir2017alors}.
@@
 
-We evaluate our approach with the SAT Competition~2022 Anniversary Track benchmark~\cite{sat2022}, consisting of 5355~instances and complete runtimes of 28~solvers.
+We evaluate our approach with the SAT Competition~2022 Anniversary Track dataset~\cite{sat2022}, consisting of 5355~instances and complete runtimes of 28~solvers.
 We perform cross-validation by treating each solver as new once and predicting these solvers' PAR-2 rank.
@@
 
-All our source code\footnote{temporary reference, anonymized version for review: \url{https://github.com/dgnr7vexskfoc3a67jdtf96kd9opz/al-for-sat-solver-benchmarking}} and and data\footnote{temporary reference, anonymized version for review: \url{https://github.com/dgnr7vexskfoc3a67jdtf96kd9opz/al-for-sat-solver-benchmarking-data}} are available on GitHub.
+All our source code\footnote{temporary link, anonymized version for review: \url{https://github.com/dgnr7vexskfoc3a67jdtf96kd9opz/al-for-sat-solver-benchmarking}} and and data\footnote{temporary link, anonymized version for review: \url{https://github.com/dgnr7vexskfoc3a67jdtf96kd9opz/al-for-sat-solver-benchmarking-data}} are available on GitHub.
 
@@
 Hoos et al.~\cite{HoosKSS13} discuss which properties are most desirable when selecting SAT benchmark instances.
-Selection criteria are instance variety to avoid over-fitting, adapted instance hardness (not too easy but also not too hard), and avoidance of duplicate instances.
+The selection criteria are instance variety to avoid over-fitting, adapted instance hardness (not too easy but also not too hard), and avoidance of duplicate instances.
 To filter too similar instances, they use a distance-based approach with the SATzilla features~\cite{XuHHL08,features}.
@@
 Thereby, the goal is to tune SAT solvers for a given sub-domain of problem instances.
-Although this task is different from our goal, e.g., we do not need to navigate configuration space, there are similarities to our approach as well.
+Although this task is different from our goal, e.g., we do not need to navigate the configuration space, there are similarities to our approach as well.
 For example, SMAC also employs an iterative, model-based selection procedure, though for configurations rather than instances.
@@
 
-Recent synthesis-based methods within the field of SAT solving show how to generate problem instances with desired properties.
-This goal is, however, orthogonal to ours~\cite{0001AEMN22,GarzonMG22}.
+Recent synthesis-based methods within the field of SAT solving show how to generate problem instances with desired properties~\cite{0001AEMN22,GarzonMG22}.
+This goal is, however, orthogonal to ours.
 While those approaches want to generate instances on which a solver is good or bad, we want to predict whether a solver is good or bad on an existing benchmark.
@@
 
-The outline of our benchmarking framework is depicted in Algorithm~\ref{algALBenchmark}. 
-Given a set of known solvers $\mathcal{A}$, instances $\mathcal{I}$ and runtimes $r$, we first initialize a prediction model~$\mathcal{M}$ for the new solver $\hat a$ (Line~1).
+Algorithm~\ref{algALBenchmark} outlines our benchmarking framework. 
+Given a set of solvers~$\mathcal{A}$, instances~$\mathcal{I}$ and runtimes~$r$, we first initialize a prediction model~$\mathcal{M}$ for the new solver $\hat a$ (Line~1).
 The prediction model~$\mathcal{M}$ is used to repeatedly select an instance (Line~4) for benchmarking $\hat a$ (Line~5). 
@@
 In particular, the model update (Line~7) uses all acquired runtimes to train $f$ with a standard supervised machine-learning methodology.
-Training features are instance features and known runtimes $r$.
-After applying the necessary transformations described in the subsequent section, we fit the prediction model using the transformed observations $\mathcal{R}$ as ground-truth labels.
+Model training uses instance features and known runtimes $r$.
+The transformed runtime observations~$\mathcal{R}$ serve as ground-truth labels.
 Note that we build a new prediction model in each iteration since running experiments (Line~5) dominates the runtime of Algorithm~\ref{algALBenchmark} by magnitudes.
@@
 Given an instance $e \in \mathcal{I}$, a solver $a \in A$ belongs to the $\gamma_k(a, e)$-fastest solvers on instance $e$. 
-In preliminary experiments, we achieved higher accuracy in predicting such discrete-runtime labels than in predicting raw runtimes.
+In preliminary experiments, we achieved higher accuracy for predicting such discrete-runtime labels than for predicting raw runtimes.
 Research on portfolio solvers has also shown that discretization works well in practice~\cite{CollauttiMMO13,NgokoCT19}.
@@
 We implement an uncertainty and an information-gain sampling strategy.
-These methods require the model's predictions to include probabilities for the $k$ discrete runtime labels induced by the clustering described in the preceding section.
+These methods require the model's predictions to include probabilities for the $k$ discrete runtime labels.
 Let \mbox{$f' : \mathcal{\hat A} \times \mathcal{I} \rightarrow \left[0, 1\right]^k$} denote this modified prediction function.
@@
 
-The uncertainty sampling strategy simply selects the instance closest to the model's decision boundary.
+The uncertainty sampling strategy selects the instance closest to the model's decision boundary.
 The set $\tilde{\mathcal{I}} \subseteq \mathcal{I}$ denotes the instances that have already been sampled.
@@
 It stops the active-learning process if the ranking induced by the model's predictions (Equation~\ref{eq:rankingeq}) remains unchanged within the last $l$ iterations.
-Note that the concrete values of the predicted score $s_{\hat a}$ might still change.
+However, the concrete values of the predicted score $s_{\hat a}$ might still change.
 We are solemnly interested in the induced ranking in this case.
@@
 
-Given all the previously presented instantiations for Algorithm~\ref{algALBenchmark}, this section briefly outlines our experimental design, including our evaluation framework, used data sets, hyper-parameter choices, and implementation details.
+Given all the previously presented instantiations for Algorithm~\ref{algALBenchmark}, this section outlines our experimental design, including our evaluation framework, used data sets, hyper-parameter choices, and implementation details.
 
@@
 As already stated in the introductory section, this work addresses the \emph{New-Solver Problem} (cf.~Definition~\ref{def:new-solver-problem}).
-We create a prediction model $\mathcal{M}$ as described in Section~\ref{sec:main:model} that provides us with a scoring function $s_k$.
+As described in Section~\ref{sec:main:model}, a prediction model $\mathcal{M}$ provides us with an estimated scoring~$s_{\hat a}$ for the new solver~$\hat{a}$.
 
@@
 Algorithm~\ref{alg:eval} shows this.
-That means that each solver plays the role of the new solver once (Line~2).
+That means that each solver plays the role of the new solver~$\hat{a}$ once (Line~2).
 Note that the \emph{new} solver in each iteration is excluded from the set of solvers $\mathcal{A}$ to avoid data leakage (Line~3).
@@
 whereby $\delta \in \left[0, 1\right]$ allows for linear weighting between the two optimization goals $O_{\operatorname{acc}}$ and $O_{\operatorname{rt}}$.
-Plotting the approaches that maximize $O_\delta$ for all $\delta \in \left[0, 1\right]$ on a $O_{\operatorname{rt}}$-$O_{\operatorname{acc}}$-diagram provides us with a Pareto front of the best approaches for different optimization-goal weightings.
+Plotting the approaches that maximize $O_\delta$ for all $\delta \in \left[0, 1\right]$ on an $O_{\operatorname{rt}}$-$O_{\operatorname{acc}}$-diagram provides us with a Pareto front of the best approaches for different optimization-goal weightings.
 
@@
 
-For hyper-parameter tuning, we randomly sample \SI{10}{\%} of the complete set of 5355 instances with stratification regarding the instance's family.
+For hyper-parameter tuning, we randomly sample \SI{10}{\%} of the complete set of 5355 instances with stratification regarding the instances' family.
 All instance families that are too \emph{small}, i.e., \SI{10}{\%} of them corresponds to less than one instance, are put into one meta-family for stratification.
-This \emph{tuning dataset} allows for a more extensive exploration of hyper-parameter space.
+This \emph{tuning dataset} allows for a more extensive exploration of the hyper-parameter space.
 
@@
 Since the potential runtime of experiments is by magnitudes larger than the model's update time, we only consider incrementing our benchmark by one instance at a time rather than using batches, which is also proposed in current active-learning advances~\cite{SinhaED19,2019gaal}.
-A drawback of this is the lack of parallel execution of runtime experiment.
+A drawback of this is the lack of parallel execution of runtime experiments.
 
@@
 Information-gain sampling selects instances that maximize the expected reduction in entropy.
-This means we sample instances revealing similarities between solvers rather than differences, which helps to build a confident model fast.
+This means we sample instances revealing similarities between solvers rather than differences, which helps to build a confident model quickly.
 However, the method lacks the possibility of selecting helpful instances for distinguishing solvers later.
@@
 Having selected the most promising hyper-parameters, we run our active-learning experiments on the complete Anniversary Track dataset (5355 instances).
-The aforementioned range $\delta \in \left[0.2, 0.8\right]$ is covered by only two configurations.
+The aforementioned range $\delta \in \left[0.2, 0.8\right]$ only results in two distinct configurations.
 The best-performing approach for $\delta \in \left[0.2, 0.7\right]$ uses the predicted runtime-label ranking, information-gain sampling, and ranking stopping criterion.
@@
 However, differences are not significant regarding a Wilcoxon signed-rank test with $\alpha = 0.05$.
-Note that the performance difference between random and non-random sampling changes for different fractions of sampled runtime (cf. Figure~\ref{fig:annitraincolorselection}).
+This performance difference between random and non-random sampling also changes for different fractions of sampled runtime (cf. Figure~\ref{fig:annitraincolorselection}).
 The static benchmark using the most frequently AL-sampled instances performs poorly, though, compared to active learning and random sampling.
@@
     The y-axis shows the average frequency of instance families in the samples selected by active learning.
-    The dashed line represents families that occur with the same frequency in both, the dataset and samples.
+    The dashed line represents families that occur with the same frequency in the dataset and samples.
   }
@@
 Given a new solver, we want to find its ranking amidst its competitors.
-Our approach provides accurate ranking predictions while needing significantly less runtime than a full evaluation on a given benchmark set.
-In particular, we can determine a new solver's PAR-2 ranking with about \SI{92}{\%} accuracy while only needing \SI{10}{\%} of the full-evaluation time.
-Our framework produces good results working with discrete runtime labels rather than regressing actual runtimes.
-We have evaluated several ranking algorithms, instance-selection approaches, and stopping criteria within our sequential active-learning process.
+Our approach provides accurate ranking predictions while needing significantly less runtime than a complete evaluation on all instances of a given benchmark set.
+On data from the SAT~Competition~2022 Anniversary Track, we can determine a new solver's PAR-2 ranking with about \SI{92}{\%} accuracy while only needing \SI{10}{\%} of the full-evaluation time.
+We have evaluated several ranking algorithms, instance-selection approaches, and stopping criteria within our sequential active-learning framework.
 We also took a brief look at which instance families are the most prevalent in selection decisions.
@@
 However, bigger batch sizes $n$ impede \emph{active learning}.
-Also, it is unclear how to synchronize the model update and instance selection without wasting too much runtime.
+Also, it is unclear how to synchronize instance selection and updates of the prediction model without wasting too much runtime.
 
 On a more general note, one can apply our evaluation framework to arbitrary $\mathcal{NP}$-complete problems, as all discussed active-learning methods are problem-agnostic.
-Especially the runtime-prediction model would need some tweaking for this purpose, though.
-Those problems share most of the relevant properties of SAT solving, i.e., there are established instance features, a complete benchmark run takes quite some time, and solver development traditionally requires expert knowledge to hand-select instances.
+These problems share most of the relevant properties of SAT solving, i.e., there are established instance features, a complete benchmark run takes a considerable amount of time, and solver benchmarking traditionally requires expert knowledge to hand-select instances.
 

### 2022-10-14 paper/paper.tex
Regenerate pdf

@@
   Author(s) anonymized for review
-  \vspace{0.4cm}
+  \vspace{1.0cm}
 }
@@
   Institute(s) anonymized for review
-  \vspace{0.4cm}
+  \vspace{0.8cm}
 }
@@
 
-The posed \emph{New-Solver Problem} has stark similarities to the well-studied field of active learning (AL) within recommender systems, in particular the \emph{new-user problem}~\cite{RubensESK15}.
+The posed \emph{New-Solver Problem} has stark similarities to the well-studied field of active learning (AL) within recommender systems, in particular, the \emph{new-user problem}~\cite{RubensESK15}.
 On the one hand, we want to maximize the utility an instance provides to our model, i.e., rank prediction accuracy, and on the other hand, minimize the cost, i.e., runtime, that is associated with its acquisition.

### 2023-01-06 paper/paper.tex
Add todos for reviewer comments
... and some minor issues I detected after submission.
De-anonymize authors
Revert (part of) changes from 6600dd240f. We still need to link the actual
code and data.

@@
 \author{
-  Author(s) anonymized for review
-  \vspace{1.0cm}
+  Tobias Fuchs\orcidID{0000-0001-9727-2878} (\Letter) \and \\
+  Jakob Bach\orcidID{0000-0003-0301-2798} \and \\
+  Markus Iser\orcidID{0000-0003-2904-232X}
+  %JB: according to camera-ready mail, we might need to include ORCID logo and link ourselves; there is a package called "orcidlink" for that purpose
 }
@@
 \institute{
-  Institute(s) anonymized for review
-  \vspace{0.8cm}
+  Karlsruhe Institute of Technology (KIT), Karlsruhe, Germany \\
+  \email{info@tobiasfuchs.de, jakob.bach@kit.edu, markus.iser@kit.edu}
 }
 
-\authorrunning{Author(s) anonymized for review} % according to sample paper of template: first names are abbreviated; if there are more than two authors, 'et al.' is used.
+\authorrunning{T. Fuchs et al.} % according to sample paper of template: first names are abbreviated; if there are more than two authors, 'et al.' is used.
 
@@
 Competitive SAT solvers are the result of extensive experimentation with a variety of ideas and considerations~\cite{FroleyksHIJS21,sat2022}.
+%JB: TODO: clarify that simply running a solver on all benchmarks is not a good approach in this scenario, i.e., motivate our approach better -> see arguments in rebuttal (comment of Reviewer 3)
 While the field of SAT solver benchmarking is well established, traditional benchmark selection approaches do not optimize for benchmark runtime.
@@
 Table~\ref{tab:requirements} compares the most relevant approaches.
+%JB: TODO: clarify that competitors in table have different goals and thus our approach is not a general improvement on them, but targets a different problem (comment of Reviewer 2)
 
@@
 \subsubsection{Active Learning.}
+%JB: TODO: explain AL for non-ML audience (comment of Reviewer 2)
+%JB: TODO: (re-)add figure for AL cycle or explain the latter better (comment of Reviewer 2); could also be placed in next section
 
@@
 We denote the new solver to be ranked by $\hat a \not\in A$ and define $\hat A := A \cup \lbrace \hat a \rbrace$ to denote all solvers, including the new solver. 
+%JB: we should harmonize use of \mathcal for A; sometimes \mathcal is used (e.g., two sentences ago), sometimes not (e.g., in previous sentence)
 Model initialization (Line~1) creates the prediction model and transforms runtimes, as discussed in the subsequent section.
@@
 %
+%JB: this function cannot be applied to new solvers, as the latter are not in $A$ (only in $\hat A$); instead of "a", could also use actual runtime as first input to \gamma, but that might lead to less straigthforward notation later in text (e.g., when defining ranking and stopping criteria)
 Given an instance $e \in \mathcal{I}$, a solver $a \in A$ belongs to the $\gamma_k(a, e)$-fastest solvers on instance $e$. 
@@
 The scoring function $s_k$ induces a ranking among solvers.
+%JB: slight inconsistency that we use number of clusters "k" as subscript (and the solver "a" as argument) here and at some other places, while we also use "\hat a" (i.e., the new solver) as subscript (and not as argument) in further places
+%JB: at the moment, s_k(a) could not be applied to new solver anyway, since it depends on \gamma, and \gamma does not work for new solvers (see comment above)
 
@@
 They comprise instance size features and node distribution statistics for several graph representations of SAT instances, among others, and are primarily inspired by the SATzilla~2012 features~\cite{features}.
+%JB: TODO: explain features in more detail or clarify where to find this information (comment of Reviewer 2)
 All features are numeric and free of missing values.
@@
 Our final choice is a stacking ensemble of a quadratic-discriminant analysis and a random forest, using a decision tree to combine its ensemble members.
+%JB: TODO: explain "stacking ensemble" (comment of Reviewer 2)
 
@@
   }
+  %JB: TODO: make figures grayscale friendly (comment of Reviewer 2) by showing line types in legend (instead of only filled circles)
   \label{fig:e2eallsolvers}
@@
 The plotted lines represent the best performing configurations (convex hull) per top-level hyper-parameter choice, i.e., per ranking approach (Fig.~\ref{fig:annitraincolorranking}), selection approach (Fig.~\ref{fig:annitraincolorselection}), and stopping criterion (Fig.~\ref{fig:annitraincolorstopping}).
+%JB: TODO: clarify how the hyperparameter analysis is conducted here, i.e., Pareto front / maximum over all other (not displayed) hyperparameters (comment of Reviewer 2)
 
@@
 Our two AL approaches both perform better than random sampling.
+%JB: TODO: explain table better, particularly implications of our approach only slightly beating random sampling (comment of Reviewer 2)
 However, differences are not significant regarding a Wilcoxon signed-rank test with $\alpha = 0.05$.
@@
 On a more general note, one can apply our evaluation framework to arbitrary $\mathcal{NP}$-complete problems, as all discussed active-learning methods are problem-agnostic.
+%JB: TODO: clarify that our approach works for any computation-intensive problem, not only NP-complete (comment of Reviewer 3)
 These problems share most of the relevant properties of SAT solving, i.e., there are established instance features, a complete benchmark run takes a considerable amount of time, and solver benchmarking traditionally requires expert knowledge to hand-select instances.

### 2023-01-11 src/al_experiments.py
Remove random file suffix

@@ -396,15 +396,14 @@ def run_e2e_experiment(
 if __name__ == "__main__":
     # Notify experiment start
     push_notification("Starting experiments.")
-    rand_int = random.randint(1, 100)
 
     for i_exp, experiment in enumerate(all_experiments):
         # Skip if present
         time.sleep(random.random() * 5)
-        if os.path.exists(f"{experiment.results_location}_{rand_int:03}.csv"):
+        if os.path.exists(f"{experiment.results_location}.csv"):
             continue
         else:
-            with open(f"{experiment.results_location}_{rand_int:03}.csv", "w") as lock_file:
+            with open(f"{experiment.results_location}.csv", "w") as lock_file:
                 lock_file.write("lock")
 
         # Retrieve column list
@@ -442,7 +441,7 @@ if __name__ == "__main__":
             ],
             index="solver",
         )
-        res_df.to_csv(f"{experiment.results_location}_{rand_int:03}.csv")
+        res_df.to_csv(f"{experiment.results_location}.csv")
 
         # Store sampled instances
         for i, (_, y_sampled) in enumerate(solver_results):

### 2023-01-12 paper/paper.tex
Update paper's links to code/data

@@
 
-All our source code\footnote{temporary link, anonymized version for review: \url{https://github.com/dgnr7vexskfoc3a67jdtf96kd9opz/al-for-sat-solver-benchmarking}} and and data\footnote{temporary link, anonymized version for review: \url{https://github.com/dgnr7vexskfoc3a67jdtf96kd9opz/al-for-sat-solver-benchmarking-data}} are available on GitHub.
+All our source code\footnote{\url{https://github.com/mathefuchs/al-for-sat-solver-benchmarking}} and and data\footnote{\url{https://github.com/mathefuchs/al-for-sat-solver-benchmarking-data}} are available on GitHub.
 
@@
   }
-  %JB: TODO: make figures grayscale friendly (comment of Reviewer 2) by showing line types in legend (instead of only filled circles)
   \label{fig:e2eallsolvers}

### 2023-01-12 paper/paper.tex
Include ORCID icons
According to camera-ready instructions, authors should do this themselves.

@@
 \usepackage[table]{xcolor}
+\usepackage{orcidlink} % ORCID icon and link; has to be loaded after "xcolor" here, else "option clash"; class "llncs" also has command "\orcidID" (which only displays full ID), but camera-ready instructions state you should add icon and link yourself
 \usepackage{tikz} % has to be loaded after "xcolor" here, else "option clash"
@@
 \author{
-  Tobias Fuchs\orcidID{0000-0001-9727-2878} (\Letter) \and \\
-  Jakob Bach\orcidID{0000-0003-0301-2798} \and \\
-  Markus Iser\orcidID{0000-0003-2904-232X}
-  %JB: according to camera-ready mail, we might need to include ORCID logo and link ourselves; there is a package called "orcidlink" for that purpose
+  Tobias Fuchs~\orcidlink{0000-0001-9727-2878} (\Letter) \and
+  Jakob Bach~\orcidlink{0000-0003-0301-2798} \and
+  Markus Iser~\orcidlink{0000-0003-2904-232X}
 }

### 2023-01-12 paper/paper.tex
Fix R3 comment on Future Work

@@
 
-On a more general note, one can apply our evaluation framework to arbitrary $\mathcal{NP}$-complete problems, as all discussed active-learning methods are problem-agnostic.
-%JB: TODO: clarify that our approach works for any computation-intensive problem, not only NP-complete (comment of Reviewer 3)
+On a more general note, one can apply our evaluation framework to arbitrary computation-intensive problems, e.g., other $\mathcal{NP}$-complete problems than SAT, as all discussed active-learning methods are problem-agnostic.
 These problems share most of the relevant properties of SAT solving, i.e., there are established instance features, a complete benchmark run takes a considerable amount of time, and solver benchmarking traditionally requires expert knowledge to hand-select instances.

### 2023-01-12 paper/paper.tex
Harmonize use of \mathcal
Was not used consistently for algorithms "A".

@@
 Let SAT solvers $\mathcal{A}$, benchmark instances $\mathcal{I}$, and runtimes $r : \mathcal{A} \times \mathcal{I} \rightarrow \left[0, \tau\right]$ be given.
-We denote the new solver to be ranked by $\hat a \not\in A$ and define $\hat A := A \cup \lbrace \hat a \rbrace$ to denote all solvers, including the new solver. 
-%JB: we should harmonize use of \mathcal for A; sometimes \mathcal is used (e.g., two sentences ago), sometimes not (e.g., in previous sentence)
+We denote the new solver to be ranked by $\hat a \not\in \mathcal{A}$ and define $\mathcal{\hat A} := \mathcal{A} \cup \lbrace \hat a \rbrace$ to denote all solvers, including the new solver.
 Model initialization (Line~1) creates the prediction model and transforms runtimes, as discussed in the subsequent section.
@@
 We transform the runtimes into discrete runtime labels on a per-instance basis.
-For each instance $e \in \mathcal{I}$, we use a clustering algorithm to assign the runtimes in $\bigl\{ r(a, e) \mid a \in A \bigr\}$ to one of $k$ clusters $C_1, \dots, C_k$ such that the fastest runtimes for instance $e$ are in cluster $C_1$ and the slowest are in cluster $C_{k-1}$.
+For each instance $e \in \mathcal{I}$, we use a clustering algorithm to assign the runtimes in $\bigl\{ r(a, e) \mid a \in \mathcal{A} \bigr\}$ to one of $k$ clusters $C_1, \dots, C_k$ such that the fastest runtimes for instance $e$ are in cluster $C_1$ and the slowest are in cluster $C_{k-1}$.
 Timeouts $\tau$ always form a separate cluster $C_{k}$.
@@
 %
-%JB: this function cannot be applied to new solvers, as the latter are not in $A$ (only in $\hat A$); instead of "a", could also use actual runtime as first input to \gamma, but that might lead to less straigthforward notation later in text (e.g., when defining ranking and stopping criteria)
-Given an instance $e \in \mathcal{I}$, a solver $a \in A$ belongs to the $\gamma_k(a, e)$-fastest solvers on instance $e$. 
+%JB: this function cannot be applied to new solvers, as the latter are not in $\mathcal{A}$ (only in $\mathcal{\hat A}$); instead of "a", could also use actual runtime as first input to \gamma, but that might lead to less straigthforward notation later in text (e.g., when defining ranking and stopping criteria)
+Given an instance $e \in \mathcal{I}$, a solver $a \in \mathcal{A}$ belongs to the $\gamma_k(a, e)$-fastest solvers on instance $e$. 
 In preliminary experiments, we achieved higher accuracy for predicting such discrete-runtime labels than for predicting raw runtimes.
@@
 
-To determine solver ranks, we use the transformed runtimes $\gamma_k(a, e)$ in the adapted scoring function $s_k : A \rightarrow [1, 2 \cdot k]$ as follows:
+To determine solver ranks, we use the transformed runtimes $\gamma_k(a, e)$ in the adapted scoring function $s_k : \mathcal{A} \rightarrow [1, 2 \cdot k]$ as follows:
 %

### 2023-01-12 paper/paper.tex
Add first batch of revisions

@@
 Competitive SAT solvers are the result of extensive experimentation with a variety of ideas and considerations~\cite{FroleyksHIJS21,sat2022}.
-%JB: TODO: clarify that simply running a solver on all benchmarks is not a good approach in this scenario, i.e., motivate our approach better -> see arguments in rebuttal (comment of Reviewer 3)
-While the field of SAT solver benchmarking is well established, traditional benchmark selection approaches do not optimize for benchmark runtime.
+While benchmarking a single SAT solver is not too expensive, trying out a multitude of different approaches and configurations steadily increases the evaluation cost.
+The field of SAT solver benchmarking is well established.
+However, traditional benchmark selection approaches do not optimize for benchmark runtime.
 Instead, traditional approaches focus on selecting a representative set of instances for ranking solvers with a scoring scheme~\cite{Gelder11,HoosKSS13}.
@@
 Table~\ref{tab:requirements} compares the most relevant approaches.
-%JB: TODO: clarify that competitors in table have different goals and thus our approach is not a general improvement on them, but targets a different problem (comment of Reviewer 2)
+Note that our approach is \emph{not} a general improvement over all the shown approaches.
+Rather, they all have slightly different goals, which do not fully align with Definition~\ref{def:new-solver-problem}.
 
@@
 \subsubsection{Active Learning.}
-%JB: TODO: explain AL for non-ML audience (comment of Reviewer 2)
-%JB: TODO: (re-)add figure for AL cycle or explain the latter better (comment of Reviewer 2); could also be placed in next section
 
-The posed \emph{New-Solver Problem} has stark similarities to the well-studied field of active learning (AL) within recommender systems, in particular, the \emph{new-user problem}~\cite{RubensESK15}.
-On the one hand, we want to maximize the utility an instance provides to our model, i.e., rank prediction accuracy, and on the other hand, minimize the cost, i.e., runtime, that is associated with its acquisition.
-In contrast to traditional passive machine-learning methods with given instance labels, active learning allows for selecting instances for which to acquire labels.
+In contrast to traditional passive machine-learning methods with given instance labels (cf. Figure~\ref{fig:passive}), active learning (AL) starts with no or little labeled data.
+Then, it repeatedly selects interesting instances for which to acquire labels based on a gradually improving model of the agent (cf. Figure~\ref{fig:activepool}).
+Figure~\ref{fig:learning} illustrates the difference between both types of learning.
+AL methods are especially useful if acquiring labels is computationally expensive.
+Without active-learning methods, it would not be obvious which instances to label and which not to.
+On the one hand, we want to maximize the utility an instance provides to our model, i.e., rank prediction accuracy, and on the other hand, minimize the cost, i.e., predicted runtime, that is associated with its acquisition.
+In that way, we obtain an accurate picture of the agent without the need of labeling every data point.
+Rubens et.~al.~\cite{RubensESK15} provide a survey of recent active-learning advances.
+
+\begin{figure}[tbp!]
+  \centering
+  \begin{tabular}[c]{ccc}
+  \begin{subfigure}[b]{0.26\textwidth}
+  \centering
+  \resizebox{!}{2.85cm}{
+  \begin{tikzpicture}
+  \draw[fill=black]  (-2.8,3.7) ellipse (0.1 and 0.1);
+  \draw[fill=black]  (-2.6,3.4) arc (0:180:0.2);
+  \node[right] at (-5.9,3.6) {Agent (Solver)};
+  \draw[fill=white]  (-3.3,2.8) rectangle (-2.5,2.2);
+  \draw[fill=white]  (-3.2,2.7) rectangle (-2.4,2.1);
+  \draw[fill=white]  (-3.1,2.6) rectangle (-2.3,2);
+  \node[right] at (-5.9,2.4) {Training Data};
+  \node (v1) at (-2.8,3.4) {};
+  \node (v2) at (-2.8,2.8) {};
+  \draw[->]  (v1) edge (v2);
+  \node[right] at (-2.8,3.1) {$x, y$};
+  \node (v3) at (-2.8,2) {};
+  \node (v4) at (-2.8,1.4) {};
+  \draw[->]  (v3) edge (v4);
+  \draw  (-3.1,1.4) rectangle (-2.5,1);
+  \draw  (-2.4,3.2) rectangle (-2.4,3.2);
+  \node[right] at (-5.9,1.2) {Learning Alg.};
+  \node (v5) at (-2.8,1) {};
+  \node (v6) at (-2.8,0.4) {};
+  \draw[->]  (v5) edge (v6);
+  \node[right] at (-5.9,0.25) {Approx. Function};
+  \node at (-2.8,0.25) {$f$};
+  \end{tikzpicture}
+  }
+  \caption{Passive.}
+  \label{fig:passive}
+  \end{subfigure}
+  &
+  \begin{subfigure}[b]{0.36\textwidth}
+  \centering
+  \resizebox{!}{2.85cm}{
+  \begin{tikzpicture}
+  \draw[fill=black]  (-2.8,3.7) ellipse (0.1 and 0.1);
+  \draw[fill=black]  (-2.6,3.4) arc (0:180:0.2);
+  \node[right] at (-5.9,3.6) {Agent (Solver)};
+  \draw[fill=white]  (-3.3,2.8) rectangle (-2.5,2.2);
+  \draw[fill=white]  (-3.2,2.7) rectangle (-2.4,2.1);
+  \draw[fill=white]  (-3.1,2.6) rectangle (-2.3,2);
+  \node[right] at (-5.9,2.4) {Training Data};
+  \node (v1) at (-2.8,3.4) {};
+  \node (v2) at (-2.8,2.8) {};
+  \draw[->]  (v1) edge (v2);
+  \node[right] at (-2.8,3.1) {$x, y$};
+  \node (v3) at (-2.8,2) {};
+  \node (v4) at (-2.8,1.4) {};
+  \draw[->]  (v3) edge (v4);
+  \draw  (-3.1,1.4) rectangle (-2.5,1);
+  \draw  (-2.4,3.2) rectangle (-2.4,3.2);
+  \node[right] at (-5.9,1.2) {Learning Alg.};
+  \node (v5) at (-2.8,1) {};
+  \node (v6) at (-2.8,0.4) {};
+  \draw[->]  (v5) edge (v6);
+  \node[right] at (-5.9,0.25) {Approx. Function};
+  \node at (-2.8,0.25) {$f$};
+  \node[right] at (-1.2,2.4) {Pool};
+  \draw[->, densely dashed] (-2.4,1.2) arc (-90:0:0.8);
+  \draw[->, densely dashed] (-1.6,2.8) arc (0:90:0.8);
+  \node[right] at (-1.6,3.135) {$x, ?$};
+  \draw  (-1.6,2.4) ellipse (0.4 and 0.3);
+  \draw[fill=black]  (-1.725,2.5) ellipse (0.07 and 0.07);
+  \draw[fill=black]  (-1.625,2.25) ellipse (0.07 and 0.07);
+  \draw[fill=black]  (-1.425,2.4) ellipse (0.07 and 0.07);
+  \end{tikzpicture}
+  }
+  \caption{Active Pool-based.}
+  \label{fig:activepool}
+  \end{subfigure}
+  &
+  \begin{subfigure}[b]{0.36\textwidth}
+  \centering
+  \resizebox{!}{2.85cm}{
+  \begin{tikzpicture}
+  \draw[fill=black]  (-2.8,3.7) ellipse (0.1 and 0.1);
+  \draw[fill=black]  (-2.6,3.4) arc (0:180:0.2);
+  \node[right] at (-5.9,3.6) {Agent (Solver)};
+  \draw[fill=white]  (-3.3,2.8) rectangle (-2.5,2.2);
+  \draw[fill=white]  (-3.2,2.7) rectangle (-2.4,2.1);
+  \draw[fill=white]  (-3.1,2.6) rectangle (-2.3,2);
+  \node[right] at (-5.9,2.4) {Training Data};
+  \node (v1) at (-2.8,3.4) {};
+  \node (v2) at (-2.8,2.8) {};
+  \draw[->]  (v1) edge (v2);
+  \node[right] at (-2.8,3.1) {$x, y$};
+  \node (v3) at (-2.8,2) {};
+  \node (v4) at (-2.8,1.4) {};
+  \draw[->]  (v3) edge (v4);
+  \draw  (-3.1,1.4) rectangle (-2.5,1);
+  \draw  (-2.4,3.2) rectangle (-2.4,3.2);
+  \node[right] at (-5.9,1.2) {Learning Alg.};
+  \node (v5) at (-2.8,1) {};
+  \node (v6) at (-2.8,0.4) {};
+  \draw[->]  (v5) edge (v6);
+  \node[right] at (-5.9,0.25) {Approx. Function};
+  \node at (-2.8,0.25) {$f$};
+  \node[right, label={[align=left]Gen.\\Model}] at (-0.75,1.88) {};
+  \node[right] at (-1.3,3.135) {$x, ?$};
+  \draw[->, densely dashed] (-2.4,1.2) arc (-90:90:1.2);
+  \end{tikzpicture}
+  }
+  \caption{Active Synthesis-based.}
+  \label{fig:activesynth}
+  \end{subfigure}
+  \end{tabular}
+
+  \caption{Types of Learning (depiction inspired by Rubens~et.al.~\cite{RubensESK15}).}
+  \label{fig:learning}
+\end{figure}
+
 AL algorithms can be categorized into \emph{synthesis-bas\-ed} \cite{0001AEMN22,GarzonMG22,2019gaal} and \emph{pool-bas\-ed} approaches \cite{GolbandiKL11,HarpaleY08,KorenBV09}.
 While synthesis-based methods generate instances for labeling, pool-based methods rely on a fixed set of unlabeled instances from which to sample.
-
 Recent synthesis-based methods within the field of SAT solving show how to generate problem instances with desired properties~\cite{0001AEMN22,GarzonMG22}.
@@
 They comprise instance size features and node distribution statistics for several graph representations of SAT instances, among others, and are primarily inspired by the SATzilla~2012 features~\cite{features}.
-%JB: TODO: explain features in more detail or clarify where to find this information (comment of Reviewer 2)
+A technical report by Xu et.~.al.~\cite{features} describes them in depth.
 All features are numeric and free of missing values.
 We drop 10 out of 56 features because of zero variance.
-Overall, prediction models have access to 46 instances features and 27 runtime features, i.e., excluding the current new solver~ $\hat{a}$.
+Overall, prediction models have access to 46 instance features and 27 runtime features, i.e., excluding the current new solver~$\hat{a}$.
 Additionally, we retrieve instance-family information\footnote{\url{https://gbd.iti.kit.edu/getdatabase/meta_db}} to evaluate the composition of our sampled benchmarks.
@@
 In particular, we conducted nested cross-validation, including hyper-parameter tuning, and used Matthews Correlation Coefficient~\cite{gorodkin2004comparing,matthews1975comparison} to assess the performance for predicting runtime labels.
-Our final choice is a stacking ensemble of a quadratic-discriminant analysis and a random forest, using a decision tree to combine its ensemble members.
-%JB: TODO: explain "stacking ensemble" (comment of Reviewer 2)
+Our final choice is a stacking ensemble of a quadratic-discriminant analysis and a random forest.
+A stacking ensemble uses a meta-model to decide which ensemble member makes the prediction on which instance.
+To do so, it uses a simple decision tree.
 
@@
 We do not show intermediate results of the active-learning procedure but only the final results after stopping.
-The plotted lines represent the best performing configurations (convex hull) per top-level hyper-parameter choice, i.e., per ranking approach (Fig.~\ref{fig:annitraincolorranking}), selection approach (Fig.~\ref{fig:annitraincolorselection}), and stopping criterion (Fig.~\ref{fig:annitraincolorstopping}).
-%JB: TODO: clarify how the hyperparameter analysis is conducted here, i.e., Pareto front / maximum over all other (not displayed) hyperparameters (comment of Reviewer 2)
+The plotted lines represent the best-performing configurations (convex hull) per top-level hyper-parameter choice, i.e., per ranking approach (Fig.~\ref{fig:annitraincolorranking}), selection approach (Fig.~\ref{fig:annitraincolorselection}), and stopping criterion (Fig.~\ref{fig:annitraincolorstopping}).
+So to speak, it is the maximum over the final results for each hyper-parameter choice that shares the plotted attributes.
 
@@
 Our two AL approaches both perform better than random sampling.
-%JB: TODO: explain table better, particularly implications of our approach only slightly beating random sampling (comment of Reviewer 2)
 However, differences are not significant regarding a Wilcoxon signed-rank test with $\alpha = 0.05$.
 This performance difference between random and non-random sampling also changes for different fractions of sampled runtime (cf. Figure~\ref{fig:annitraincolorselection}).
-The static benchmark using the most frequently AL-sampled instances performs poorly, though, compared to active learning and random sampling.
+A clear advantage of our approach is, though, that it indicates when to stop adding further instances.
+Otherwise, it would not be clear why the shown runtime fractions would yield the best trade-off at all.
+While the active-learning results are less strong on the full dataset than on the smaller tuning dataset, they still show the benefit of making benchmark selection dependent on the solvers to distinguish.
+
+A static benchmark using the most frequently AL-sampled instances performs poorly, though, compared to active learning and random sampling.
 This outcome is somewhat expected since the static benchmark does not reflect the right balance of instance families:
 Families whose instances are uniform-randomly selected by AL, e.g., for different solvers, appear less often in this benchmark than families where some instances are sampled more often than others.
-While the active-learning results are less strong on the full dataset than on the smaller tuning dataset, they still show the benefit of making benchmark selection dependent on the solvers to distinguish.
 

### 2023-01-13 paper/paper.tex
Review recent camera-ready revisions
... from commit d3f82ed117. Extend some explanations, cut text at other
places. Fix overfull boxes. We are within the page limit again.

@@
 One of the main phases of algorithm engineering is benchmarking.
-This is also true for the propositional satisfiability problem (SAT), the archetypal $\mathcal{NP}$-complete problem.
+This also applies to propositional satisfiability (SAT), the archetypal $\mathcal{NP}$-complete problem.
 Benchmarking is, however, quite expensive regarding the runtime of experiments.
-Competitive SAT solvers are the result of extensive experimentation with a variety of ideas and considerations~\cite{FroleyksHIJS21,sat2022}.
-While benchmarking a single SAT solver is not too expensive, trying out a multitude of different approaches and configurations steadily increases the evaluation cost.
-The field of SAT solver benchmarking is well established.
-However, traditional benchmark selection approaches do not optimize for benchmark runtime.
-Instead, traditional approaches focus on selecting a representative set of instances for ranking solvers with a scoring scheme~\cite{Gelder11,HoosKSS13}.
-SAT~Competitions typically use the \mbox{PAR-2} score, i.e., the average runtime with a penalty of $2 \tau$ for timeouts with time-limit~$\tau$~\cite{FroleyksHIJS21}.
+While benchmarking a single SAT solver might still be feasible, developing new, competitive SAT solvers requires extensive experimentation with a variety of configurations~\cite{FroleyksHIJS21,sat2022}.
+Thus, it is highly desirable to reduce benchmarking time and discard unpromising configurations early, which allows testing more approaches.
+The field of SAT solver benchmarking is well established, but traditional benchmark selection approaches do not optimize benchmark runtime.
+Instead, they focus on selecting a representative set of instances for scoring solvers~\cite{Gelder11,HoosKSS13}.
+For the latter, SAT~Competitions typically employ the \mbox{PAR-2} score, i.e., the average runtime with a penalty of $2 \tau$ for timeouts with time-limit~$\tau$~\cite{FroleyksHIJS21}.
 
@@
 In the following, we discuss related work from the areas of static benchmark selection, algorithm configuration, incremental benchmark selection, and active learning.
-Table~\ref{tab:requirements} compares the most relevant approaches.
-Note that our approach is \emph{not} a general improvement over all the shown approaches.
-Rather, they all have slightly different goals, which do not fully align with Definition~\ref{def:new-solver-problem}.
+Table~\ref{tab:requirements} compares the most relevant approaches, which all pursue slightly different goals.
+Our approach is \emph{not} a general improvement over the others but the only one fully aligned with Definition~\ref{def:new-solver-problem}.
 
@@
 
-In contrast to traditional passive machine-learning methods with given instance labels (cf. Figure~\ref{fig:passive}), active learning (AL) starts with no or little labeled data.
-Then, it repeatedly selects interesting instances for which to acquire labels based on a gradually improving model of the agent (cf. Figure~\ref{fig:activepool}).
-Figure~\ref{fig:learning} illustrates the difference between both types of learning.
-AL methods are especially useful if acquiring labels is computationally expensive.
-Without active-learning methods, it would not be obvious which instances to label and which not to.
-On the one hand, we want to maximize the utility an instance provides to our model, i.e., rank prediction accuracy, and on the other hand, minimize the cost, i.e., predicted runtime, that is associated with its acquisition.
-In that way, we obtain an accurate picture of the agent without the need of labeling every data point.
-Rubens et.~al.~\cite{RubensESK15} provide a survey of recent active-learning advances.
+Prediction models in passive machine learning are trained on datasets with given instance labels (cf.~Figure~\ref{fig:passive}).
+In contrast, active learning (AL) starts with no or little labeled data.
+It repeatedly selects interesting problem instances for which to acquire labels, aiming to gradually improve the prediction model (cf.~Figure~\ref{fig:activepool}).
+AL methods are especially useful if acquiring labels is computationally expensive, like obtaining solver runtimes.
+Without AL methods, it is not obvious which instances to label and which not.
+On the one hand, we want to maximize the utility an instance provides to our model, i.e., rank prediction accuracy, and on the other hand, minimize the cost, i.e., predicted runtime, associated with the instance's acquisition.
+Thus, we obtain an accurate prediction model without having labeled every data point.
 
@@
   \begin{tabular}[c]{ccc}
-  \begin{subfigure}[b]{0.26\textwidth}
+  \begin{subfigure}[b]{0.25\textwidth}
   \centering
-  \resizebox{!}{2.85cm}{
+  \resizebox{!}{2.75cm}{
   \begin{tikzpicture}
@@
   \draw[->]  (v5) edge (v6);
-  \node[right] at (-5.9,0.25) {Approx. Function};
+  \node[right] at (-5.9,0.25) {Predic. Function};
   \node at (-2.8,0.25) {$f$};
@@
   &
-  \begin{subfigure}[b]{0.36\textwidth}
+  \begin{subfigure}[b]{0.35\textwidth}
   \centering
-  \resizebox{!}{2.85cm}{
+  \resizebox{!}{2.75cm}{
   \begin{tikzpicture}
@@
 
-  \caption{Types of Learning (depiction inspired by Rubens~et.al.~\cite{RubensESK15}).}
+  \caption{Types of machine learning (depiction inspired by Rubens~et.al.~\cite{RubensESK15}).}
   \label{fig:learning}
@@
 
-AL algorithms can be categorized into \emph{synthesis-bas\-ed} \cite{0001AEMN22,GarzonMG22,2019gaal} and \emph{pool-bas\-ed} approaches \cite{GolbandiKL11,HarpaleY08,KorenBV09}.
-While synthesis-based methods generate instances for labeling, pool-based methods rely on a fixed set of unlabeled instances from which to sample.
+Rubens et.~al.~\cite{RubensESK15} survey active-learning advances.
+While synthesis-based AL methods~\cite{0001AEMN22,GarzonMG22,2019gaal} generate instances for labeling, pool-based methods~\cite{GolbandiKL11,HarpaleY08,KorenBV09} rely on a fixed set of unlabeled instances to sample from.
 Recent synthesis-based methods within the field of SAT solving show how to generate problem instances with desired properties~\cite{0001AEMN22,GarzonMG22}.
@@
 In particular, the model update (Line~7) uses all acquired runtimes to train $f$ with a standard supervised machine-learning methodology.
-Model training uses instance features and known runtimes $r$.
+Model training uses instance features and known runtimes~$r$.
 The transformed runtime observations~$\mathcal{R}$ serve as ground-truth labels.
@@
 Given an instance $e \in \mathcal{I}$, a solver $a \in \mathcal{A}$ belongs to the $\gamma_k(a, e)$-fastest solvers on instance $e$. 
-In preliminary experiments, we achieved higher accuracy for predicting such discrete-runtime labels than for predicting raw runtimes.
+In preliminary experiments, we achieved higher accuracy for predicting such discrete runtime labels than for predicting raw runtimes.
 Research on portfolio solvers has also shown that discretization works well in practice~\cite{CollauttiMMO13,NgokoCT19}.
@@
 We also use a database of 56 instance features\footnote{\url{https://gbd.iti.kit.edu/getdatabase/base_db}} provided in the GBD Benchmark Database~(GBD) by Iser~et~al.~\cite{IserS18}.
-They comprise instance size features and node distribution statistics for several graph representations of SAT instances, among others, and are primarily inspired by the SATzilla~2012 features~\cite{features}.
-A technical report by Xu et.~.al.~\cite{features} describes them in depth.
+They comprise instance size features and node distribution statistics for several graph representations of SAT instances, among others, and are primarily inspired by the SATzilla~2012 features described in~\cite{features}.
 All features are numeric and free of missing values.
@@
 Additionally, we retrieve instance-family information\footnote{\url{https://gbd.iti.kit.edu/getdatabase/meta_db}} to evaluate the composition of our sampled benchmarks.
-Instance families comprise instances derived from the same application domain, e.g., planning, cryptography, etc., and are a valuable tool for analyzing solver performance and portfolios.
+Instance families comprise instances from the same application domain, e.g., planning, cryptography, etc., and are a valuable tool for analyzing solver performance.
 
@@
 In particular, we conducted nested cross-validation, including hyper-parameter tuning, and used Matthews Correlation Coefficient~\cite{gorodkin2004comparing,matthews1975comparison} to assess the performance for predicting runtime labels.
-Our final choice is a stacking ensemble of a quadratic-discriminant analysis and a random forest.
-A stacking ensemble uses a meta-model to decide which ensemble member makes the prediction on which instance.
-To do so, it uses a simple decision tree.
+Our final choice is a stacking ensemble~\cite{wolpert1992stacked} of two prediction models, a quadratic-discriminant analysis~\cite{tharwat2016linear} and a random forest~\cite{breiman2001random}.
+Both these models can learn non-linear relationships between the instance features and the runtime labels.
+Stacking means that another prediction model, in our case a simple decision tree, decides which of the two ensemble members makes the prediction on which instance.
 
@@
 
-To define prediction targets, i.e., discretized runtime labels, we use hierarchical clustering with $k = 3$ and a log-single-link criterion, which produced the most \emph{useful} labels in preliminary experiments.
+To define prediction targets, i.e., discrete runtime labels, we use hierarchical clustering with $k = 3$ and a log-single-link criterion, which produced the most \emph{useful} labels in preliminary experiments.
 We denote this adapted solver scoring function with $s_3$.
@@
 We do not show intermediate results of the active-learning procedure but only the final results after stopping.
-The plotted lines represent the best-performing configurations (convex hull) per top-level hyper-parameter choice, i.e., per ranking approach (Fig.~\ref{fig:annitraincolorranking}), selection approach (Fig.~\ref{fig:annitraincolorselection}), and stopping criterion (Fig.~\ref{fig:annitraincolorstopping}).
-So to speak, it is the maximum over the final results for each hyper-parameter choice that shares the plotted attributes.
+The plotted lines represent the best-performing configurations per ranking approach (Fig.~\ref{fig:annitraincolorranking}), selection approach (Fig.~\ref{fig:annitraincolorselection}), and stopping criterion (Fig.~\ref{fig:annitraincolorstopping}).
+In particular, we show the Pareto front, i.e., of all configurations that share a particular value of the plotted hyper-parameter, we take the maximum ranking accuracy over all remaining hyper-parameters \emph{not} displayed in the corresponding plot.
 
@@
 This performance difference between random and non-random sampling also changes for different fractions of sampled runtime (cf. Figure~\ref{fig:annitraincolorselection}).
-A clear advantage of our approach is, though, that it indicates when to stop adding further instances.
-Otherwise, it would not be clear why the shown runtime fractions would yield the best trade-off at all.
+A clear advantage of our approach is, though, that it indicates when to stop adding further instances, depending on the trade-off parameter~$\delta$.
 While the active-learning results are less strong on the full dataset than on the smaller tuning dataset, they still show the benefit of making benchmark selection dependent on the solvers to distinguish.
@@
 
-\section{Conclusion}
+\section{Conclusions and Future Work}
 
@@
 
-\subsection{Future Work}
+Future work may compare further subroutines for ranking, instance selection, and stopping.
+Additionally, one can apply our evaluation framework to arbitrary computation-intensive problems, e.g., other $\mathcal{NP}$-complete problems than SAT, as all discussed active-learning methods are problem-agnostic.
+Such problems share most of the relevant properties of SAT solving, i.e., there are established instance features, a complete benchmark is expensive, and benchmark selection traditionally requires expert knowledge.
 
-Future work may compare further ranking algorithms, instance-selection approaches, and stopping criteria.
-Furthermore, it is possible to formulate runtime discretization as an optimization problem, directly selecting the most discriminatory discretization technique rather than making an empirical comparison.
-
-A major shortcoming of our current approach is the lack of parallelization, selecting instances one at a time.
+From the technical perspective, one could formulate runtime discretization as an optimization problem rather than addressing it empirically.
+Further, a major shortcoming of our current approach is the lack of parallelization, selecting instances one at a time.
 Benchmarking on a computing cluster with $n$ cores benefits from having batches of $n$ instances.
@@
 Also, it is unclear how to synchronize instance selection and updates of the prediction model without wasting too much runtime.
-
-On a more general note, one can apply our evaluation framework to arbitrary computation-intensive problems, e.g., other $\mathcal{NP}$-complete problems than SAT, as all discussed active-learning methods are problem-agnostic.
-These problems share most of the relevant properties of SAT solving, i.e., there are established instance features, a complete benchmark run takes a considerable amount of time, and solver benchmarking traditionally requires expert knowledge to hand-select instances.
 

### 2023-01-13 paper/paper.tex
Read and revise full paper
Several language revisions, though most of them were not strictly required.

@@
 Benchmarking is, however, quite expensive regarding the runtime of experiments.
-While benchmarking a single SAT solver might still be feasible, developing new, competitive SAT solvers requires extensive experimentation with a variety of configurations~\cite{FroleyksHIJS21,sat2022}.
-Thus, it is highly desirable to reduce benchmarking time and discard unpromising configurations early, which allows testing more approaches.
+While benchmarking a single SAT solver might still be feasible, developing new, competitive SAT solvers requires extensive experimentation with a variety of ideas~\cite{FroleyksHIJS21,sat2022}.
+In particular, a new solver idea is rarely best on the first try.
+Thus, it is highly desirable to reduce benchmarking time and discard unpromising ideas early, thereby testing more approaches or spending more time on the promising ones.
 The field of SAT solver benchmarking is well established, but traditional benchmark selection approaches do not optimize benchmark runtime.
@@
 Our approach can predict the rank of a new solver with high accuracy in only a fraction of the time needed to evaluate the full benchmark.
-The problem we solve is specified in Definition~\ref{def:new-solver-problem}.
+Definition~\ref{def:new-solver-problem} specifies the problem we address.
 
@@
 In particular, we only subsample the benchmark once instead of comparing pairwise against each other solver~\cite{MatriconAFSH21}.
-Since a new solver idea is rarely best on the first try, it is desired to rank it fast.
-In this way, one can discard the new solver idea if it performs poorly across the board or may tweak it further if it shows promising results at least in some cases.
 
-We evaluate our approach with the SAT Competition~2022 Anniversary Track dataset~\cite{sat2022}, consisting of 5355~instances and complete runtimes of 28~solvers.
-We perform cross-validation by treating each solver as new once and predicting these solvers' PAR-2 rank.
-On average, our predictions reach about \SI{92}{\%} accuracy with only about \SI{10}{\%} of the runtime that would be needed to evaluate these solvers on the complete set of instances.
+We evaluate our approach with the SAT Competition~2022 Anniversary Track dataset~\cite{sat2022}, consisting of 5355~instances and runtimes of 28~solvers.
+We perform cross-validation by treating each solver as the new one once and predicting these solvers' PAR-2 rank.
+On average, our predictions reach about \SI{92}{\%} accuracy with only about \SI{10}{\%} of the runtime required to evaluate these solvers on the complete set of instances.
 
-All our source code\footnote{\url{https://github.com/mathefuchs/al-for-sat-solver-benchmarking}} and and data\footnote{\url{https://github.com/mathefuchs/al-for-sat-solver-benchmarking-data}} are available on GitHub.
+All our source code\footnote{\url{https://github.com/mathefuchs/al-for-sat-solver-benchmarking}} and data\footnote{\url{https://github.com/mathefuchs/al-for-sat-solver-benchmarking-data}} are available on GitHub.
 
@@
 Table~\ref{tab:requirements} compares the most relevant approaches, which all pursue slightly different goals.
-Our approach is \emph{not} a general improvement over the others but the only one fully aligned with Definition~\ref{def:new-solver-problem}.
+Thus, our approach is \emph{not} a general improvement over the others but the only one fully aligned with Definition~\ref{def:new-solver-problem}.
 
@@
 
-Prediction models in passive machine learning are trained on datasets with given instance labels (cf.~Figure~\ref{fig:passive}).
+Prediction models in passive machine learning are trained on datasets with given instance labels (cf.~Fig.~\ref{fig:passive}).
 In contrast, active learning (AL) starts with no or little labeled data.
-It repeatedly selects interesting problem instances for which to acquire labels, aiming to gradually improve the prediction model (cf.~Figure~\ref{fig:activepool}).
+It repeatedly selects interesting problem instances for which to acquire labels, aiming to gradually improve the prediction model (cf.~Fig.~\ref{fig:activepool}).
 AL methods are especially useful if acquiring labels is computationally expensive, like obtaining solver runtimes.
@@
 On the one hand, we want to maximize the utility an instance provides to our model, i.e., rank prediction accuracy, and on the other hand, minimize the cost, i.e., predicted runtime, associated with the instance's acquisition.
-Thus, we obtain an accurate prediction model without having labeled every data point.
+Thus, we strive for an accurate prediction model without having to label every data point.
 
@@
 
-    $t \leftarrow \operatorname{runExperiment}\left(\hat{a},\,  e\right)$  \tcp*{Runs $\hat{a}$ on $e$ with time-out $\tau$}
+    $t \leftarrow \operatorname{runExperiment}\left(\hat{a},\,  e\right)$  \tcp*{Runs $\hat{a}$ on $e$ with timeout $\tau$}
 
@@
 Let SAT solvers $\mathcal{A}$, benchmark instances $\mathcal{I}$, and runtimes $r : \mathcal{A} \times \mathcal{I} \rightarrow \left[0, \tau\right]$ be given.
-We denote the new solver to be ranked by $\hat a \not\in \mathcal{A}$ and define $\mathcal{\hat A} := \mathcal{A} \cup \lbrace \hat a \rbrace$ to denote all solvers, including the new solver.
-Model initialization (Line~1) creates the prediction model and transforms runtimes, as discussed in the subsequent section.
+We denote the new solver to be ranked by $\hat a \not\in \mathcal{A}$ and define $\mathcal{\hat A} := \mathcal{A} \cup \lbrace \hat a \rbrace$ to denote all solvers, including the new one.
+Model initialization (Line~1) creates the prediction model and transforms runtimes, as discussed later.
 No model training is needed at this point since we do not have any labeled data yet.
@@
 The transformed runtime observations~$\mathcal{R}$ serve as ground-truth labels.
-Note that we build a new prediction model in each iteration since running experiments (Line~5) dominates the runtime of Algorithm~\ref{algALBenchmark} by magnitudes.
-Finally, we predict the score of the new solver $\hat a$ in Line~8 with the prediction function~$f$.
+Note that we build a new prediction model in each iteration since running experiments (Line~5) dominates the runtime of model training by magnitudes.
+Finally, we predict the score of the new solver $\hat a$ with the prediction function~$f$ (Line~8).
 
@@
 %
-%JB: this function cannot be applied to new solvers, as the latter are not in $\mathcal{A}$ (only in $\mathcal{\hat A}$); instead of "a", could also use actual runtime as first input to \gamma, but that might lead to less straigthforward notation later in text (e.g., when defining ranking and stopping criteria)
 Given an instance $e \in \mathcal{I}$, a solver $a \in \mathcal{A}$ belongs to the $\gamma_k(a, e)$-fastest solvers on instance $e$. 
@@
 The scoring function $s_k$ induces a ranking among solvers.
-%JB: slight inconsistency that we use number of clusters "k" as subscript (and the solver "a" as argument) here and at some other places, while we also use "\hat a" (i.e., the new solver) as subscript (and not as argument) in further places
-%JB: at the moment, s_k(a) could not be applied to new solver anyway, since it depends on \gamma, and \gamma does not work for new solvers (see comment above)
 
@@
 The ranking stopping criterion is less sophisticated in comparison.
-It stops the active-learning process if the ranking induced by the model's predictions (Equation~\ref{eq:rankingeq}) remains unchanged within the last $l$ iterations.
+It stops the active-learning process if the ranking induced by the model's predictions (Equation~\ref{eq:rankingeq}) remained unchanged within the last~$l$~iterations.
 However, the concrete values of the predicted score $s_{\hat a}$ might still change.
@@
   \KwIn{Solvers $\mathcal{A}$, Instances $\mathcal{I}$, Runtimes $r : \mathcal{A} \times \mathcal{I} \rightarrow [0, \tau]$}
-  \KwOut{Average Rank Accuracy $\bar{O}_{\operatorname{acc}}$, Average Fraction of Runtime $\bar{O}_{\operatorname{rt}}$}
+  \KwOut{Average Ranking Accuracy $\bar{O}_{\operatorname{acc}}$, Average Fraction of Runtime $\bar{O}_{\operatorname{rt}}$}
   \BlankLine
@@
   \BlankLine
-  \tcp{Determine Rank Accuracy}
+  \tcp{Determine Ranking Accuracy}
     $O_{\operatorname{acc}} \leftarrow 0$ \;    
@@
 
-As already stated in the introductory section, this work addresses the \emph{New-Solver Problem} (cf.~Definition~\ref{def:new-solver-problem}).
+As stated in the Introduction, this work addresses the \emph{New-Solver Problem} (cf.~Definition~\ref{def:new-solver-problem}).
 As described in Section~\ref{sec:main:model}, a prediction model $\mathcal{M}$ provides us with an estimated scoring~$s_{\hat a}$ for the new solver~$\hat{a}$.
@@
 Note that the \emph{new} solver in each iteration is excluded from the set of solvers $\mathcal{A}$ to avoid data leakage (Line~3).
-After running our active-learning framework for a solver $\hat{a}$ in Line~4, we compute the value of both our optimization goals:
+After running our active-learning framework for solver~$\hat{a}$ (Line~4), we compute the value of both our optimization goals:
 First and foremost, we want to provide the engineer of new SAT solvers with an accurate ranking.
@@
 The dataset consists of 5355 instances with respective runtime data of 28 sequential SAT solvers.
-We also use a database of 56 instance features\footnote{\url{https://gbd.iti.kit.edu/getdatabase/base_db}} provided in the GBD Benchmark Database~(GBD) by Iser~et~al.~\cite{IserS18}.
+We also use a database of 56 instance features\footnote{\url{https://gbd.iti.kit.edu/getdatabase/base_db}} from the Global Benchmark Database~(GBD) by Iser~et~al.~\cite{IserS18}.
 They comprise instance size features and node distribution statistics for several graph representations of SAT instances, among others, and are primarily inspired by the SATzilla~2012 features described in~\cite{features}.
@@
 
-Given Algorithm~\ref{algALBenchmark}, there are several possible instantiations for the three phases, i.e., \emph{ranking}, \emph{selection}, and \emph{stopping}.
+Given Algorithm~\ref{algALBenchmark}, there are several possible instantiations for the three sub-routines, i.e., \emph{ranking}, \emph{selection}, and \emph{stopping}.
 Also, there are different choices for the runtime-label prediction model and runtime discretization.
@@
     \item
-    History size: consider the latest 1, 10, 20, 30, or 40 predictions within a voting approach for stability.
+    History size: Consider the latest 1, 10, 20, 30, or 40 predictions within a voting approach for stability.
     The latest $x$ predictions for each instance vote on the instance's winning label.
     \item
-    Fallback threshold: if the difference of scores between the new solver $\hat{a}$ and another solver drops below \SI{0.01}, \SI{0.05}, or \SI{0.1}, use the partially observed PAR-2 ranking as a tie-breaker.
+    Fallback threshold: If the difference of scores between the new solver~$\hat{a}$ and another solver drops below \SI{0.01}, \SI{0.05}, or \SI{0.1}, use the partially observed PAR-2 ranking as a tie-breaker.
   \end{itemize}
@@
 We denote this adapted solver scoring function with $s_3$.
-Other clustering approaches that we have tried include hierarchical clustering with mean-, median- and complete-link criteria, as well as $k$-means and spectral clustering.
-In our chosen hierarchical procedure, each non-time-out runtime starts in a separate interval.
+In our chosen hierarchical procedure, each non-timeout runtime starts in a separate interval.
 We then gradually merge intervals whose single-link logarithmic distance is the smallest until the desired number of partitions is reached.
+Other clustering approaches that we tried include hierarchical clustering with mean-, median-, and complete-link criterion, as well as $k$-means and spectral clustering.
 
@@
 In this section, we evaluate our active-learning framework.
-First, we analyze and tune the different components of our framework on the tuning dataset.
+First, we analyze and tune the different sub-routines of our framework on the tuning dataset.
 Next, we evaluate the best configurations with the full dataset.
@@
   \caption{
-    $O_{\operatorname{rt}}$-$O_{\operatorname{acc}}$-diagrams comparing different hyper-parameter instantiations of our active-learning framework.
+    $O_{\operatorname{rt}}$-$O_{\operatorname{acc}}$-diagrams comparing different hyper-parameter instantiations of our active-learning framework on the hyper-parameter-tuning dataset.
     The x-axis shows the ratio of total solver runtime on the sampled instances relative to all instances.
@@
 
-Regarding the stopping criterion (Fig.~\ref{fig:annitraincolorstopping}), the ranking stopping criterion performs the most consistently well.
+Regarding the stopping criteria (Fig.~\ref{fig:annitraincolorstopping}), the ranking stopping criterion performs the most consistently well.
 If accuracy is strongly favored (very high $\delta$), the Wilcoxon stopping criterion performs better.
@@
   \caption{
-    Scatter plot comparing different instantiations of $\delta$ for our active-learning strategy on the hyper-parameter-tuning dataset.
+    Scatter plot comparing different instantiations of trade-off parameter~$\delta$ for our active-learning framework on the hyper-parameter-tuning dataset.
     The x-axis shows the fraction of runtime $O_{\operatorname{rt}}$ of the sample, while the y-axes show the fraction of instances sampled and ranking accuracy, respectively.
@@
 
-Figure~\ref{fig:annitrainoptgoalruntime} shows an interesting consequence of weighting our optimization goals:
+Fig.~\ref{fig:annitrainoptgoalruntime} shows an interesting consequence of weighting our optimization goals:
 If we, on the one hand, desire to get a \emph{rough} estimate of a solver's performance fast (low $\delta$), approaches favor selecting many \emph{easy} instances.
@@
 By having many observations, it is easier to build a model.
-If we, on the other hand, desire to get a \emph{good} estimate of a solver's performance in a moderate amount of time (high $\delta$), approaches favor selecting fewer \emph{difficult} instances.
+If we, on the other hand, desire to get a \emph{good} estimate of a solver's performance in a moderate amount of time (high $\delta$), approaches favor selecting few, \emph{difficult} instances.
 In particular, the fraction of instances is smaller than the fraction of runtime.
 
-Furthermore, Figure~\ref{fig:annitrainoptgoalacc} reveals which values make the most sense for $\delta$.
+Furthermore, Fig.~\ref{fig:annitrainoptgoalacc} reveals which values make the most sense for $\delta$.
 The range $\delta \in \left[0.2, 0.8\right]$, thereby, corresponds to the points with a runtime fraction between \SI{0.03} and \SI{0.22}.
@@
   \caption{
-      Performance comparison of the best-performing active-learning approaches (\emph{AL}), random sampling of the same runtime fraction with 1000 repetitions (\emph{Random}), and statically selecting the instances most frequently sampled by active-learning approaches (\emph{Most Freq.})
+      Performance comparison (on the full dataset) of the best-performing active-learning approaches (\emph{AL}), random sampling of the same runtime fraction with 1000 repetitions (\emph{Random}), and statically selecting the instances most frequently sampled by active-learning approaches (\emph{Most Freq.})
     }
@@
 However, differences are not significant regarding a Wilcoxon signed-rank test with $\alpha = 0.05$.
-This performance difference between random and non-random sampling also changes for different fractions of sampled runtime (cf. Figure~\ref{fig:annitraincolorselection}).
+This performance difference between random and non-random sampling also changes for different fractions of sampled runtime (cf. Fig.~\ref{fig:annitraincolorselection}).
 A clear advantage of our approach is, though, that it indicates when to stop adding further instances, depending on the trade-off parameter~$\delta$.
@@
   \caption{
-    Scatter plot showing the \emph{importance} of different instance families to our framework.
-    The x-axis shows the frequency of instance families in the full Anniversary Track dataset.
+    Scatter plot showing the \emph{importance} of different instance families to our framework on the full dataset.
+    The x-axis shows the frequency of instance families in the dataset.
     The y-axis shows the average frequency of instance families in the samples selected by active learning.
@@
 Selection decisions of our approach also reveal the importance of different instance families to our framework.
-Figure~\ref{fig:annifinalfamilies} shows the occurrence of instance families within the dataset and the benchmarks created by active learning.
+Fig.~\ref{fig:annifinalfamilies} shows the occurrence of instance families within the dataset and the benchmarks created by active learning.
 We use the best-performing configurations for all $\delta \in \left[0, 1\right]$ and examine the selection decisions by the active-learning approach on the SAT Competition~2022 Anniversary Track dataset~\cite{sat2022}.
@@
 
-In this work, we have discussed possible solutions to the \emph{New-Solver Problem}:
-Given a new solver, we want to find its ranking amidst its competitors.
-Our approach provides accurate ranking predictions while needing significantly less runtime than a complete evaluation on all instances of a given benchmark set.
+In this work, we have addressed to the \emph{New-Solver Problem}:
+Given a new solver, we want to find its ranking amidst competitors.
+Our approach provides accurate ranking predictions while needing significantly less runtime than a complete evaluation on a given benchmark set.
 On data from the SAT~Competition~2022 Anniversary Track, we can determine a new solver's PAR-2 ranking with about \SI{92}{\%} accuracy while only needing \SI{10}{\%} of the full-evaluation time.
@@
 
-Future work may compare further subroutines for ranking, instance selection, and stopping.
+Future work may compare further sub-routines for ranking, instance selection, and stopping.
 Additionally, one can apply our evaluation framework to arbitrary computation-intensive problems, e.g., other $\mathcal{NP}$-complete problems than SAT, as all discussed active-learning methods are problem-agnostic.

### 2023-01-13 paper/paper.tex
Further minor language revisions
Pasting the text into Word and Grammarly.

@@
 \begin{abstract}
-  Benchmarking is one of the most important phases when developing algorithms.
+  Benchmarking is a crucial phase when developing algorithms.
   This also applies to solvers for the SAT (propositional satisfiability) problem.
@@
   We also discuss the importance of instance families in the selection process. 
-  Overall, our tool provides a reliable way for solver engineers to efficiently determine the performance of a new solver.
+  Overall, our tool provides a reliable way for solver engineers to determine a new solver's performance efficiently.
 
@@
 In particular, a new solver idea is rarely best on the first try.
-Thus, it is highly desirable to reduce benchmarking time and discard unpromising ideas early, thereby testing more approaches or spending more time on the promising ones.
+Thus, it is highly desirable to reduce benchmarking time and discard unpromising ideas early, allowing to test more approaches or spend more time on promising ones.
 The field of SAT solver benchmarking is well established, but traditional benchmark selection approaches do not optimize benchmark runtime.
@@
 In this paper, we present a novel benchmark selection approach based on active learning.
-Our approach can predict the rank of a new solver with high accuracy in only a fraction of the time needed to evaluate the full benchmark.
+Our approach can predict the rank of a new solver with high accuracy in only a fraction of the time needed to evaluate the complete benchmark.
 Definition~\ref{def:new-solver-problem} specifies the problem we address.
@@
 
-All our source code\footnote{\url{https://github.com/mathefuchs/al-for-sat-solver-benchmarking}} and data\footnote{\url{https://github.com/mathefuchs/al-for-sat-solver-benchmarking-data}} are available on GitHub.
+Our entire source code\footnote{\url{https://github.com/mathefuchs/al-for-sat-solver-benchmarking}} and experimental data\footnote{\url{https://github.com/mathefuchs/al-for-sat-solver-benchmarking-data}} are available on GitHub.
 
@@
 In such competitions, the organizers define the rules for composing the benchmarks.
-These selection strategies are mostly static, i.e., they do not depend on particular solvers to distinguish.
+These selection strategies are primarily static, i.e., they do not depend on particular solvers to distinguish.
 Balint et al. provide an overview of benchmark-selection criteria in different solver competitions~\cite{balint2015overview}.
-Froleyks et al. describe benchmark selection as it is done in recent SAT~competitions~\cite{FroleyksHIJS21}.
+Froleyks et al. describe benchmark selection in recent SAT~competitions~\cite{FroleyksHIJS21}.
 Manthey and Möhle find that competition benchmarks might contain redundant instances and propose a feature-based approach to remove redundancy~\cite{manthey2016better}.
@@
 Hoos et al.~\cite{HoosKSS13} discuss which properties are most desirable when selecting SAT benchmark instances.
-The selection criteria are instance variety to avoid over-fitting, adapted instance hardness (not too easy but also not too hard), and avoidance of duplicate instances.
+The selection criteria are instance variety to avoid over-fitting, adapted instance hardness (not too easy but also not too hard), and avoiding duplicate instances.
 To filter too similar instances, they use a distance-based approach with the SATzilla features~\cite{XuHHL08,features}.
@@
 It repeatedly selects interesting problem instances for which to acquire labels, aiming to gradually improve the prediction model (cf.~Fig.~\ref{fig:activepool}).
-AL methods are especially useful if acquiring labels is computationally expensive, like obtaining solver runtimes.
+AL methods are especially beneficial if acquiring labels is computationally expensive, like obtaining solver runtimes.
 Without AL methods, it is not obvious which instances to label and which not.
@@
 
-In this section, we present two dynamic stopping criteria, the Wilcoxon and the ranking stopping criterion (cf. Algorithm~\ref{algALBenchmark}, Line~3).
+In this section, we present the two dynamic stopping criteria in our experiments, the Wilcoxon and the ranking stopping criterion (cf. Algorithm~\ref{algALBenchmark}, Line~3).
 
@@
 
-The Wilcoxon stopping criterion stops the active-learning process when we are confident enough that the predicted runtime labels of the new solver are sufficiently different from the labels of the existing solvers.
+The Wilcoxon stopping criterion stops the active-learning process when we are confident enough that the predicted runtime labels of the new solver are sufficiently different from existing solvers.
 This criterion is loosely inspired by Matricon et.~al.~\cite{MatriconAFSH21}.
@@
 Algorithm~\ref{alg:eval} shows this.
-That means that each solver plays the role of the new solver~$\hat{a}$ once (Line~2).
+That means each solver plays the role of the new solver~$\hat{a}$ once (Line~2).
 Note that the \emph{new} solver in each iteration is excluded from the set of solvers $\mathcal{A}$ to avoid data leakage (Line~3).
@@
 Also, there are different choices for the runtime-label prediction model and runtime discretization.
-The experiment configurations are given below.
+We describe these experimental configurations in the following.
 
@@
 
-We only use one model configuration for runtime-label prediction in our experiments since an exhaustive grid search would be infeasible.
+Our experiments only use one model configuration for runtime-label prediction since an exhaustive grid search would be infeasible.
 In preliminary experiments, we compared various model types from \emph{scikit-learn}~\cite{scikit-learn}.
@@
 This means we sample instances revealing similarities between solvers rather than differences, which helps to build a confident model quickly.
-However, the method lacks the possibility of selecting helpful instances for distinguishing solvers later.
+However, the method cannot select helpful instances for distinguishing solvers later.
 Random sampling performs reasonably well but is outperformed by uncertainty sampling in all cases, showing the benefit of actively selecting instances based on a prediction model.
@@
 
-In this work, we have addressed to the \emph{New-Solver Problem}:
+In this work, we have addressed the \emph{New-Solver Problem}:
 Given a new solver, we want to find its ranking amidst competitors.

### 2023-01-23 paper/paper.tex
Add Acknowledgments
Re-phrase evaluation a bit (without deleting content) to make space.

@@
 The best-performing approach for $\delta \in \left[0.2, 0.7\right]$ uses the predicted runtime-label ranking, information-gain sampling, and ranking stopping criterion.
-It can predict a new solver's PAR-2 ranking with about \SI{90.48}{\%} accuracy ($O_{\operatorname{acc}}$), while only needing \SI{5.41}{\%} of the full evaluation time ($O_{\operatorname{rt}}$).
+It can predict a new solver's PAR-2 ranking with \SI{90.48}{\%} accuracy ($O_{\operatorname{acc}}$) in only \SI{5.41}{\%} of the full evaluation time ($O_{\operatorname{rt}}$).
 The best-performing approach for $\delta \in (0.7, 0.8]$ uses the predicted runtime-label ranking, uncertainty sampling, and ranking stopping criterion.
-It can predict a new solver's PAR-2 ranking with about \SI{92.33}{\%} accuracy ($O_{\operatorname{acc}}$), while only needing \SI{10.35}{\%} of the full evaluation time ($O_{\operatorname{rt}}$).
+It can predict a new solver's PAR-2 ranking with \SI{92.33}{\%} accuracy ($O_{\operatorname{acc}}$) in only \SI{10.35}{\%} of the full evaluation time ($O_{\operatorname{rt}}$).
 
@@
 
-Table~\ref{tab:fulldataset} shows how both active-learning approaches compare against static baselines.
-The first column shows the two aforementioned AL configurations.
-We compare them against a random baseline and a static benchmark set consisting of the instances most frequently sampled by active learning.
-The random baseline (\emph{Random}) samples instances until it reaches roughly the same fraction of runtime as the AL benchmark sets.
-We repeat this procedure 1000 times and report average results.
-The benchmark set of the most frequently AL-sampled instances (\emph{Most Freq.}) uses the average sampling frequencies of instances over all solvers and all Pareto-optimal active-learning approaches.
+Table~\ref{tab:fulldataset} shows how both active-learning approaches (column \emph{AL}) compare against two static baselines:
+\emph{Random} samples instances until it reaches roughly the same fraction of runtime as the AL benchmark sets.
+We repeat sampling 1000 times and report average results.
+\emph{Most Freq.} uses the instances most frequently sampled by active learning.
+In particular, we consider the average sampling frequencies over all solvers and Pareto-optimal active-learning approaches.
 
-Our two AL approaches both perform better than random sampling.
-However, differences are not significant regarding a Wilcoxon signed-rank test with $\alpha = 0.05$.
-This performance difference between random and non-random sampling also changes for different fractions of sampled runtime (cf. Fig.~\ref{fig:annitraincolorselection}).
+Both our AL approaches perform better than random sampling.
+However, the performance differences are not significant regarding a Wilcoxon signed-rank test with $\alpha = 0.05$ and also depend on the fraction of sampled runtime (cf.
+Fig.~\ref{fig:annitraincolorselection}).
 A clear advantage of our approach is, though, that it indicates when to stop adding further instances, depending on the trade-off parameter~$\delta$.
@@
 
+\subsubsection*{Acknowledgments.}
+
+This work was supported by the Ministry of Science, Research and the Arts Baden-Württemberg, project \emph{Algorithm Engineering for the Scalability Challenge (AESC)}.
+
 %

### 2023-01-24 paper/paper.tex
language: wording and phrases

@@
   This also applies to solvers for the SAT (propositional satisfiability) problem.
-  Benchmark selection chooses representative problem instances that reliably discriminate solvers based on their runtime.
+  Benchmark selection is about choosing representative problem instances that reliably discriminate solvers based on their runtime.
   In this paper, we present a dynamic benchmark selection approach based on active learning.
@@
 We evaluate our approach with the SAT Competition~2022 Anniversary Track dataset~\cite{sat2022}, consisting of 5355~instances and runtimes of 28~solvers.
-We perform cross-validation by treating each solver as the new one once and predicting these solvers' PAR-2 rank.
+We perform cross-validation by treating each solver once as the new solver and learning to predict the PAR-2 rank of that solver.
 On average, our predictions reach about \SI{92}{\%} accuracy with only about \SI{10}{\%} of the runtime required to evaluate these solvers on the complete set of instances.
@@
 Biased benchmarks can easily lead to fallacious interpretations~\cite{abs-2107-07002}.
-Also, benchmark selection has many movable parts, e.g., performance measures, aggregation, and handling of missing values.
-Questionable research practices might modify these elements a-posteriori to fit expectations, biasing results~\cite{NiesslHWCB22}.
+Benchmarking also has many interchangeable parts, such as the performance measures used, including how measurement points are aggregated and how missing values are handled.
+Questionable research practices could alter these elements a-posteriori to meet expectations, thereby skewing the results~\cite{NiesslHWCB22}.
 In the following, we discuss related work from the areas of static benchmark selection, algorithm configuration, incremental benchmark selection, and active learning.
@@
 Hoos et al.~\cite{HoosKSS13} discuss which properties are most desirable when selecting SAT benchmark instances.
-The selection criteria are instance variety to avoid over-fitting, adapted instance hardness (not too easy but also not too hard), and avoiding duplicate instances.
+The selection criteria are instance \emph{variety} to avoid over-fitting, adapted instance \emph{hardness} (not too easy but also not too hard), and avoiding \emph{duplicate} instances.
 To filter too similar instances, they use a distance-based approach with the SATzilla features~\cite{XuHHL08,features}.
@@
 Selecting an instance based on the model is a core functionality of our framework (cf. Algorithm~\ref{algALBenchmark}, Line~4).
-In this section, we introduce our sampling strategies, which use the model's label-prediction function $f$ and are inspired by existing work within the realms of active learning~\cite{settles2009active}.
-We implement an uncertainty and an information-gain sampling strategy.
+In this section, we two instance sampling strategies, one that minimizes uncertainty and one that maximizes information-gain.
+Both strategies use the model's label-prediction function $f$ and are inspired by existing work within the realms of active learning~\cite{settles2009active}.
 These methods require the model's predictions to include probabilities for the $k$ discrete runtime labels.
 Let \mbox{$f' : \mathcal{\hat A} \times \mathcal{I} \rightarrow \left[0, 1\right]^k$} denote this modified prediction function.
+In the following, the set $\tilde{\mathcal{I}} \subseteq \mathcal{I}$ denotes the instances that have already been sampled.
 
@@
 
-The uncertainty sampling strategy selects the instance closest to the model's decision boundary.
-The set $\tilde{\mathcal{I}} \subseteq \mathcal{I}$ denotes the instances that have already been sampled.
+The uncertainty sampling strategy selects the instance closest to the model's decision boundary, i.e., 
+we select the instance $e \in \mathcal{I} \setminus \tilde{\mathcal{I}}$ that minimizes $U(e)$, which is specified as follows:
 %
 \begin{equation*}
-  \underset{e \in \mathcal{I} \setminus \tilde{\mathcal{I}}}{\arg\min} \left\lvert \frac{1}{k} - \max_{n \in \left\lbrace 1, \dots, k \right\rbrace} f'\!\left(\hat{a}, e\right)_{n} \right\rvert
+  \operatorname{U}(e) := \left\lvert \frac{1}{k} - \max_{n \in \left\lbrace 1, \dots, k \right\rbrace} f'\!\left(\hat{a}, e\right)_{n} \right\rvert
 \end{equation*}
@@
 Note that the \emph{new} solver in each iteration is excluded from the set of solvers $\mathcal{A}$ to avoid data leakage (Line~3).
-After running our active-learning framework for solver~$\hat{a}$ (Line~4), we compute the value of both our optimization goals:
-First and foremost, we want to provide the engineer of new SAT solvers with an accurate ranking.
-We define the \emph{ranking accuracy} $O_{\operatorname{acc}} \in \left[0, 1\right]$ (higher is better) by the fraction of pairs $\left(\hat{a}, a\right)$ for all $a \in \mathcal{A}$ that are decided correctly by the given ranking regarding the PAR-2 scores $\operatorname{par_2}$ (Lines~5-8).
-Second, we also want to optimize for runtime.
+After running our active-learning framework for solver~$\hat{a}$ (Line~4), we compute the value of both our optimization goals, i.e., ranking accuracy and runtime.
+We define the \emph{ranking accuracy} $O_{\operatorname{acc}} \in \left[0, 1\right]$ (higher is better) by the fraction of pairs $\left(\hat{a}, a\right)$ for all $a \in \mathcal{A}$ that are decided correctly by the ranking function $\operatorname{par_2}$ (Lines~5-8).
 The \emph{fraction of runtime} that the algorithm needs to arrive at its conclusion is denoted by $O_{\operatorname{rt}} \in \left[0, 1\right]$ (lower is better).
@@
 The dataset consists of 5355 instances with respective runtime data of 28 sequential SAT solvers.
-We also use a database of 56 instance features\footnote{\url{https://gbd.iti.kit.edu/getdatabase/base_db}} from the Global Benchmark Database~(GBD) by Iser~et~al.~\cite{IserS18}.
+We also use a database of 56 instance features\footnote{\url{https://benchmark-database.de/getdatabase/base_db}} from the Global Benchmark Database~(GBD) by Iser~et~al.~\cite{IserS18}.
 They comprise instance size features and node distribution statistics for several graph representations of SAT instances, among others, and are primarily inspired by the SATzilla~2012 features described in~\cite{features}.
@@
 Overall, prediction models have access to 46 instance features and 27 runtime features, i.e., excluding the current new solver~$\hat{a}$.
-Additionally, we retrieve instance-family information\footnote{\url{https://gbd.iti.kit.edu/getdatabase/meta_db}} to evaluate the composition of our sampled benchmarks.
+
+Additionally, we retrieve instance-family information\footnote{\url{https://benchmark-database.de/getdatabase/meta_db}} to evaluate the composition of our sampled benchmarks.
 Instance families comprise instances from the same application domain, e.g., planning, cryptography, etc., and are a valuable tool for analyzing solver performance.
-
 For hyper-parameter tuning, we randomly sample \SI{10}{\%} of the complete set of 5355 instances with stratification regarding the instances' family.
@@
 
-Regarding \emph{ranking} (cf. Section~\ref{sec:main:model}), we experiment with the following approaches, including our used hyper-parameter values:
+Regarding \emph{ranking} (cf. Section~\ref{sec:main:model}), we experiment with the following approaches and hyper-parameter values:
 
@@
 
-For \emph{selection} (cf. Section~\ref{sec:main:selection}), we experiment with the following methods, including our used hyper-parameter values.
+For \emph{selection} (cf. Section~\ref{sec:main:selection}), we experiment with the following methods and hyper-parameter values.
 Since the potential runtime of experiments is by magnitudes larger than the model's update time, we only consider incrementing our benchmark by one instance at a time rather than using batches, which is also proposed in current active-learning advances~\cite{SinhaED19,2019gaal}.
@@
 
-For \emph{stopping} decisions (cf. Section~\ref{sec:main:stopping}), we experiment with the following criteria, including our used hyper-parameter values:
+For \emph{stopping} decisions (cf. Section~\ref{sec:main:stopping}), we experiment with the following criteria and hyper-parameter values:
 
@@
 
-Regarding ranking approaches (Fig.~\ref{fig:annitraincolorranking}), using the $s_3$-induced runtime-label ranking consistently outperforms the partially observed PAR-2 ranking for each possible value of the trade-off parameter~$\delta$.
+Regarding ranking approaches (Fig.~\ref{fig:annitraincolorranking}), using the predicted $s_3$-induced runtime-label ranking consistently outperforms the partially observed PAR-2 ranking for each possible value of the trade-off parameter~$\delta$.
 This outcome is expected since selection decisions are not random.
@@
 
-Regarding the stopping criteria (Fig.~\ref{fig:annitraincolorstopping}), the ranking stopping criterion performs the most consistently well.
+Regarding the stopping criteria (Fig.~\ref{fig:annitraincolorstopping}), the ranking stopping criterion performs well most consistently.
 If accuracy is strongly favored (very high $\delta$), the Wilcoxon stopping criterion performs better.
@@
 We repeat sampling 1000 times and report average results.
-\emph{Most Freq.} uses the instances most frequently sampled by active learning.
-In particular, we consider the average sampling frequencies over all solvers and Pareto-optimal active-learning approaches.
+\emph{Most Freq.} uses a static benchmark set consisting of those instances most frequently sampled by our active learning approach.
+In particular, we consider the average sampling frequency over all solvers and Pareto-optimal active-learning approaches.
 
@@
 Additionally, one can apply our evaluation framework to arbitrary computation-intensive problems, e.g., other $\mathcal{NP}$-complete problems than SAT, as all discussed active-learning methods are problem-agnostic.
-Such problems share most of the relevant properties of SAT solving, i.e., there are established instance features, a complete benchmark is expensive, and benchmark selection traditionally requires expert knowledge.
+Such problems share most of the relevant properties of SAT solving, i.e., there are established instance features, a complete benchmark is expensive, and traditional benchmark selection requires expert knowledge.
 

### 2023-01-24 paper/paper.tex
3.1 model

@@
 Algorithm~\ref{algALBenchmark} outlines our benchmarking framework. 
-Given a set of solvers~$\mathcal{A}$, instances~$\mathcal{I}$ and runtimes~$r$, we first initialize a prediction model~$\mathcal{M}$ for the new solver $\hat a$ (Line~1).
+Given a set of solvers~$\mathcal{A}$, instances~$\mathcal{I}$ and runtimes~$r$, we first initialize a prediction model~$\mathcal{M}$ for the new solver $\hat a \not\in \mathcal{A}$ (Line~1).
 The prediction model~$\mathcal{M}$ is used to repeatedly select an instance (Line~4) for benchmarking $\hat a$ (Line~5). 
@@
 
-Let SAT solvers $\mathcal{A}$, benchmark instances $\mathcal{I}$, and runtimes $r : \mathcal{A} \times \mathcal{I} \rightarrow \left[0, \tau\right]$ be given.
-We denote the new solver to be ranked by $\hat a \not\in \mathcal{A}$ and define $\mathcal{\hat A} := \mathcal{A} \cup \lbrace \hat a \rbrace$ to denote all solvers, including the new one.
-Model initialization (Line~1) creates the prediction model and transforms runtimes, as discussed later.
-No model training is needed at this point since we do not have any labeled data yet.
-Later, model $\mathcal{M}$ provides a runtime-label prediction function $f : \mathcal{\hat A} \times \mathcal{I} \rightarrow \mathbb{R}$ that powers the decisions within our framework.
-In particular, the model update (Line~7) uses all acquired runtimes to train $f$ with a standard supervised machine-learning methodology.
-Model training uses instance features and known runtimes~$r$.
-The transformed runtime observations~$\mathcal{R}$ serve as ground-truth labels.
+The model $M$ maintains runtime transformations which we describe in the subsequent section. 
+Additionally, $M$ provides a runtime-label prediction function $f : \mathcal{\hat A} \times \mathcal{I} \rightarrow \mathbb{R}$ for all solvers $\mathcal{\hat A} := \mathcal{A} \cup \lbrace \hat a \rbrace$.
+The prediction function $f$ powers instance selection as described in Section~\ref{sec:main:selection}.
+During model update (Algorithm~\ref{algALBenchmark},~Line~7), $f$ is trained to predict the known and acquired transformed runtimes based on instance features described in Section~\ref{sec:exdesign:data}.
 Note that we build a new prediction model in each iteration since running experiments (Line~5) dominates the runtime of model training by magnitudes.
@@
 \subsection{Data}
+\label{sec:exdesign:data}
 

### 2023-01-24 paper/paper.tex
minor

@@
 This criterion is loosely inspired by Matricon et.~al.~\cite{MatriconAFSH21}.
-We use the average $p$-value $W_{\hat{a}}$ of a Wilcoxon signed-rank test $w(S,P)$ of the two runtime label distributions $S=\{ \gamma(a, e) \mid e \in \mathcal{I} \}$ for an existing solver $a$ and \mbox{$P=\{ f(\hat a, e) \mid e \in \mathcal{I} \}$} for the new solver $\hat{a}$:
+We use the average $p$-value $W_{\hat{a}}$ of a Wilcoxon signed-rank test $\operatorname{w}(S,P)$ of the two runtime label distributions $S=\{ \gamma(a, e) \mid e \in \mathcal{I} \}$ for an existing solver $a$ and \mbox{$P=\{ f(\hat a, e) \mid e \in \mathcal{I} \}$} for the new solver $\hat{a}$:
 %

### 2023-01-24 paper/paper.tex
Fix minor issues

@@
 Hoos et al.~\cite{HoosKSS13} discuss which properties are most desirable when selecting SAT benchmark instances.
-The selection criteria are instance \emph{variety} to avoid over-fitting, adapted instance \emph{hardness} (not too easy but also not too hard), and avoiding \emph{duplicate} instances.
+The selection criteria are instance variety to avoid over-fitting, adapted instance hardness (not too easy but also not too hard), and avoiding duplicate instances.
 To filter too similar instances, they use a distance-based approach with the SATzilla features~\cite{XuHHL08,features}.
@@
 Selecting an instance based on the model is a core functionality of our framework (cf. Algorithm~\ref{algALBenchmark}, Line~4).
-In this section, we two instance sampling strategies, one that minimizes uncertainty and one that maximizes information-gain.
+In this section, we introduce two instance sampling strategies, one that minimizes uncertainty and one that maximizes information-gain.
 Both strategies use the model's label-prediction function $f$ and are inspired by existing work within the realms of active learning~\cite{settles2009active}.
@@
 
-Regarding ranking approaches (Fig.~\ref{fig:annitraincolorranking}), using the predicted $s_3$-induced runtime-label ranking consistently outperforms the partially observed PAR-2 ranking for each possible value of the trade-off parameter~$\delta$.
+Regarding ranking approaches (Fig.~\ref{fig:annitraincolorranking}), using the predicted $s_3$-induced run\-time-label rank\-ing consistently outperforms the partially observed PAR-2 ranking for each possible value of the trade-off parameter~$\delta$.
 This outcome is expected since selection decisions are not random.

### 2024-04-16 paper_jar/paper.tex
Add bonus tables and figures

--- 2023-01-24 paper/paper.tex
@@
 \usepackage{tabularx}
+\usepackage{booktabs}
 \usepackage[table]{xcolor}
@@
 Biased benchmarks can easily lead to fallacious interpretations~\cite{abs-2107-07002}.
-Benchmarking also has many interchangeable parts, such as the performance measures used, including how measurement points are aggregated and how missing values are handled.
+Benchmarking also has many interchangeable parts, such as the performance measures used, how measurement points are aggregated, and how missing values are handled.
 Questionable research practices could alter these elements a-posteriori to meet expectations, thereby skewing the results~\cite{NiesslHWCB22}.
@@
   }
-    \hline
+    \toprule
     Feature & Hoos~\cite{HoosKSS13} & SMAC~\cite{HutterHL11} & Matricon~\cite{MatriconAFSH21} & Our approach \\
-    \hline
+    \midrule
     Ranking/Scoring & \cmark & \xmark & (\cmark) & \cmark \\
@@
     Scalability & \cmark & \cmark & \xmark & \cmark \\
-    \hline
+    \bottomrule
   \end{tabular}
@@
 
-The model $M$ maintains runtime transformations which we describe in the subsequent section. 
-Additionally, $M$ provides a runtime-label prediction function $f : \mathcal{\hat A} \times \mathcal{I} \rightarrow \mathbb{R}$ for all solvers $\mathcal{\hat A} := \mathcal{A} \cup \lbrace \hat a \rbrace$.
-The prediction function $f$ powers instance selection as described in Section~\ref{sec:main:selection}.
-During model update (Algorithm~\ref{algALBenchmark},~Line~7), $f$ is trained to predict the known and acquired transformed runtimes based on instance features described in Section~\ref{sec:exdesign:data}.
-Note that we build a new prediction model in each iteration since running experiments (Line~5) dominates the runtime of model training by magnitudes.
+The model $M$ provides a runtime-label prediction function $f : \mathcal{\hat A} \times \mathcal{I} \rightarrow \mathbb{R}$ for all solvers $\mathcal{\hat A} := \mathcal{A} \cup \lbrace \hat a \rbrace$.
+This prediction function powers instance selection as described in Section~\ref{sec:main:selection}.
+During model updates (Algorithm~\ref{algALBenchmark},~Line~7), $f$ is trained to predict a transformed version of the acquired runtimes~$\mathcal{R}$.
+We describe the runtime transformation in the subsequent section.
+The features described in Section~\ref{sec:exdesign:data} serve as the input to the model.
+Further, note that we build a new prediction model in each iteration since running experiments (Line~5) dominates the runtime of model training by magnitudes.
 Finally, we predict the score of the new solver $\hat a$ with the prediction function~$f$ (Line~8).
@@
 
-We transform the runtimes into discrete runtime labels on a per-instance basis.
-For each instance $e \in \mathcal{I}$, we use a clustering algorithm to assign the runtimes in $\bigl\{ r(a, e) \mid a \in \mathcal{A} \bigr\}$ to one of $k$ clusters $C_1, \dots, C_k$ such that the fastest runtimes for instance $e$ are in cluster $C_1$ and the slowest are in cluster $C_{k-1}$.
+For the prediction model~$M$, we transform the real-valued runtimes into discrete runtime labels on a per-instance basis.
+For each instance $e \in \mathcal{I}$, we use a clustering algorithm to assign the runtimes in $\bigl\{ r(a, e) \mid a \in \mathcal{A} \bigr\}$ to one of $k$ clusters $C_1, \dots, C_k$ such that the fastest runtimes for the instance $e$ are in cluster $C_1$ and the slowest are in cluster $C_{k-1}$.
 Timeouts $\tau$ always form a separate cluster $C_{k}$.
@@
 Selecting an instance based on the model is a core functionality of our framework (cf. Algorithm~\ref{algALBenchmark}, Line~4).
-In this section, we introduce two instance sampling strategies, one that minimizes uncertainty and one that maximizes information-gain.
+In this section, we introduce two instance sampling strategies, one that minimizes uncertainty and one that maximizes information gain.
 Both strategies use the model's label-prediction function $f$ and are inspired by existing work within the realms of active learning~\cite{settles2009active}.
@@
 After running our active-learning framework for solver~$\hat{a}$ (Line~4), we compute the value of both our optimization goals, i.e., ranking accuracy and runtime.
-We define the \emph{ranking accuracy} $O_{\operatorname{acc}} \in \left[0, 1\right]$ (higher is better) by the fraction of pairs $\left(\hat{a}, a\right)$ for all $a \in \mathcal{A}$ that are decided correctly by the ranking function $\operatorname{par_2}$ (Lines~5-8).
+We define the \emph{ranking accuracy} $O_{\operatorname{acc}} \in \left[0, 1\right]$ (higher is better) by the fraction of pairs $\left(\hat{a}, a\right)$ for all $a \in \mathcal{A}$ that are decided correctly regarding the ground-truth scoring~$\operatorname{par_2}$ (Lines~5-8).
 The \emph{fraction of runtime} that the algorithm needs to arrive at its conclusion is denoted by $O_{\operatorname{rt}} \in \left[0, 1\right]$ (lower is better).
@@
 Instance families comprise instances from the same application domain, e.g., planning, cryptography, etc., and are a valuable tool for analyzing solver performance.
+
 For hyper-parameter tuning, we randomly sample \SI{10}{\%} of the complete set of 5355 instances with stratification regarding the instances' family.
@@
 Given Algorithm~\ref{algALBenchmark}, there are several possible instantiations for the three sub-routines, i.e., \emph{ranking}, \emph{selection}, and \emph{stopping}.
-Also, there are different choices for the runtime-label prediction model and runtime discretization.
 We describe these experimental configurations in the following.
@@
 
-\subsubsection{Prediction model.}
+\subsection{Runtime Discretization and Prediction}
 
-Our experiments only use one model configuration for runtime-label prediction since an exhaustive grid search would be infeasible.
-In preliminary experiments, we compared various model types from \emph{scikit-learn}~\cite{scikit-learn}.
-In particular, we conducted nested cross-validation, including hyper-parameter tuning, and used Matthews Correlation Coefficient~\cite{gorodkin2004comparing,matthews1975comparison} to assess the performance for predicting runtime labels.
-Our final choice is a stacking ensemble~\cite{wolpert1992stacked} of two prediction models, a quadratic-discriminant analysis~\cite{tharwat2016linear} and a random forest~\cite{breiman2001random}.
-Both these models can learn non-linear relationships between the instance features and the runtime labels.
-Stacking means that another prediction model, in our case a simple decision tree, decides which of the two ensemble members makes the prediction on which instance.
+% TODO tf: Why not regression? Bad performance
+\begin{table}[htbp]
+  \centering
+  \caption{Regression performance of models trained on 46 instance and 27 runtime features. We report the RMSE and MAE metrics on the test sets.}
+  ~\\[1em]
+  \begin{tabular}{
+    >{\raggedleft\arraybackslash}m{0.37\textwidth}
+    >{\centering\arraybackslash}m{0.01\textwidth}
+    >{\centering\arraybackslash}m{0.28\textwidth}
+    >{\centering\arraybackslash}m{0.28\textwidth}
+  }
+    \toprule
+    {Regression Models} & & {Avg. RMSE ($\pm$ std)} & {Avg. MAE ($\pm$ std)} \\
+    \midrule
+    Random-Forest-Regressor & & \SI{1929.84}{s} ($\pm$ \SI{1161.15}{s}) & \phantom{0}\SI{957.36}{s} ($\pm$ \SI{297.21}{s}) \\[0.4ex]
+    MLP-Regressor           & & \SI{2275.20}{s} ($\pm$ \SI{1545.01}{s}) & \SI{1284.30}{s} ($\pm$ \SI{390.57}{s}) \\
+    \bottomrule
+  \end{tabular}
+\end{table}
 
-\subsubsection{Runtime discretization.}
+% TODO tf: Timeout prediction performance
+\begin{table}[htbp]
+  \centering
+  \caption{MCC performance of models predicting timeouts. The models are trained 28 times with each solver as a target once.}
+  ~\\[1em]
+  \begin{tabular}{
+    >{\raggedleft\arraybackslash}m{0.37\textwidth}
+    >{\centering\arraybackslash}m{0.01\textwidth}
+    >{\centering\arraybackslash}m{0.28\textwidth}
+    >{\centering\arraybackslash}m{0.28\textwidth}
+  }
+    \toprule
+    {Timeout Prediction Models} & & {46 Instance Features Avg. MCC ($\pm$ std)} & {46\,+\,27 Runtime Feat. Avg. MCC ($\pm$ std)} \\
+    \midrule
+    Stacking (QDA\,+\,RF)                 & & 0.6513 ($\pm$ 0.0458) & 0.9527 ($\pm$ 0.0292) \\[0.4ex]
+    Quadratic-DA (QDA)                    & & 0.2593 ($\pm$ 0.1310) & 0.9290 ($\pm$ 0.0339) \\[0.4ex]
+    Random-Forest-Classifier (RF)         & & 0.6607 ($\pm$ 0.0547) & 0.8530 ($\pm$ 0.0479) \\[0.4ex]
+    Ada-Boost                             & & 0.5412 ($\pm$ 0.0612) & 0.8384 ($\pm$ 0.0444) \\[0.4ex]
+    Decision-Tree                         & & 0.5980 ($\pm$ 0.0423) & 0.8059 ($\pm$ 0.0707) \\[0.4ex]
+    Logistic-Regression                   & & 0.2031 ($\pm$ 0.0728) & 0.8052 ($\pm$ 0.1018) \\[0.4ex]
+    $k$NN-Classifier                      & & 0.5108 ($\pm$ 0.0387) & 0.7885 ($\pm$ 0.1521) \\[0.4ex]
+    MLP-Classifier                        & & 0.1293 ($\pm$ 0.0718) & 0.7760 ($\pm$ 0.1408) \\[0.4ex]
+    Support-Vector-Machines               & & 0.0595 ($\pm$ 0.0554) & 0.7757 ($\pm$ 0.2149) \\[0.4ex]
+    Naive-Bayes                           & & 0.1173 ($\pm$ 0.0872) & 0.7306 ($\pm$ 0.1394) \\
+    \bottomrule
+  \end{tabular}
+\end{table}
 
+% TODO tf: Cluster prediction performance
+\begin{table}[htbp]
+  \centering
+  \caption{MCC performance of the best models predicting the $C_k$ clusterings.}
+  \begin{subtable}{\textwidth}
+    \centering
+    \caption{Predicting $C_3$ clusters.}
+    \begin{tabular}{
+      >{\raggedleft\arraybackslash}m{0.37\textwidth}
+      >{\centering\arraybackslash}m{0.01\textwidth}
+      >{\centering\arraybackslash}m{0.28\textwidth}
+      >{\centering\arraybackslash}m{0.28\textwidth}
+    }
+      \toprule
+      {$C_3$ Cluster Prediction Models} & & {46\,+\,27 Runtime Feat. Avg. MCC ($\pm$ std)} & {46\,+\,27\,+\,27 Clus. Feat. Avg. MCC ($\pm$ std)} \\
+      \midrule
+      Stacking (QDA\,+\,RF)                 & & 0.7464 ($\pm$ 0.0497) & 0.8380 ($\pm$ 0.0634) \\[0.4ex]
+      Quadratic-DA (QDA)                    & & 0.6903 ($\pm$ 0.0607) & 0.7738 ($\pm$ 0.0459) \\[0.4ex]
+      Random-Forest-Classifier (RF)         & & 0.7116 ($\pm$ 0.0385) & 0.7841 ($\pm$ 0.0469) \\
+      \bottomrule
+    \end{tabular}
+  \end{subtable}
+  \begin{subtable}{\textwidth}
+    \centering
+    \caption{Predicting $C_4$ clusters.}
+    \begin{tabular}{
+      >{\raggedleft\arraybackslash}m{0.37\textwidth}
+      >{\centering\arraybackslash}m{0.01\textwidth}
+      >{\centering\arraybackslash}m{0.28\textwidth}
+      >{\centering\arraybackslash}m{0.28\textwidth}
+    }
+      \toprule
+      {$C_4$ Cluster Prediction Models} & & {46\,+\,27 Runtime Feat. Avg. MCC ($\pm$ std)} & {46\,+\,27\,+\,27 Clus. Feat. Avg. MCC ($\pm$ std)} \\
+      \midrule
+      Stacking (QDA\,+\,RF)                 & & 0.6562 ($\pm$ 0.0538) & 0.7366 ($\pm$ 0.0377) \\[0.4ex]
+      Quadratic-DA (QDA)                    & & 0.6007 ($\pm$ 0.0765) & 0.6872 ($\pm$ 0.0469) \\[0.4ex]
+      Random-Forest-Classifier (RF)         & & 0.6033 ($\pm$ 0.0455) & 0.6828 ($\pm$ 0.0527) \\
+      \bottomrule
+    \end{tabular}
+  \end{subtable}
+  \begin{subtable}{\textwidth}
+    \centering
+    \caption{Predicting $C_5$ clusters.}
+    \begin{tabular}{
+      >{\raggedleft\arraybackslash}m{0.37\textwidth}
+      >{\centering\arraybackslash}m{0.01\textwidth}
+      >{\centering\arraybackslash}m{0.28\textwidth}
+      >{\centering\arraybackslash}m{0.28\textwidth}
+    }
+      \toprule
+      {$C_5$ Cluster Prediction Models} & & {46\,+\,27 Runtime Feat. Avg. MCC ($\pm$ std)} & {46\,+\,27\,+\,27 Clus. Feat. Avg. MCC ($\pm$ std)} \\
+      \midrule
+      Stacking (QDA\,+\,RF)                 & & 0.5835 ($\pm$ 0.0615) & 0.6396 ($\pm$ 0.0553) \\[0.4ex]
+      Quadratic-DA (QDA)                    & & 0.5508 ($\pm$ 0.0669) & 0.5881 ($\pm$ 0.0912) \\[0.4ex]
+      Random-Forest-Classifier (RF)         & & 0.5420 ($\pm$ 0.0459) & 0.5895 ($\pm$ 0.0475) \\
+      \bottomrule
+    \end{tabular}
+  \end{subtable}
+\end{table}
+
+% TODO tf: Confusion matrices
+\begin{figure}[tbp]
+  \centering
+  \begin{subfigure}{0.45\textwidth}
+    \centering
+    \includegraphics[height=4.123cm]{../plots/cm2labels.pdf}
+    \caption{Timeout prediction.}
+  \end{subfigure}
+  \begin{subfigure}{0.45\textwidth}
+    \centering
+    \includegraphics[height=4.228cm]{../plots/cm3labels.pdf}
+    \caption{$C_3$ cluster prediction.}
+  \end{subfigure}
+  \caption{Confusion matrices for predicting timeouts and $C_3$ clusters.}
+\end{figure}
+
+% TODO tf: Actual rankings with discretization
+% * Correct ranking pairs in table below: 97.45%
+% * Max. difference between wrong pairs 0.11
+% * Spearman correlation of PAR2 and Discrete rank: 0.9862
+\begin{table}[htbp]
+  \centering
+  \caption{Ranking of Anniversary-Track solvers with standard PAR-2 runtime and our discretized $s_3$ ranking.}
+  \label{tab:ranking}
+  ~\\[1em]
+  \begin{tabular}{
+    >{\centering\arraybackslash}m{0.07\textwidth}
+    >{\centering\arraybackslash}m{0.12\textwidth}
+    >{\centering\arraybackslash}m{0.07\textwidth}
+    >{\centering\arraybackslash}m{0.12\textwidth}
+    >{\raggedright\arraybackslash}m{0.01\textwidth}
+    >{\raggedright\arraybackslash}m{0.32\textwidth}
+  }
+    \toprule
+    \multicolumn{2}{c}{PAR-2} & \multicolumn{2}{c}{$s_3$} \\
+    \cmidrule(r){1-2}\cmidrule(r){3-4}
+    Rank &   Score & Rank &  Score & & Solver \\
+    \midrule
+       1 & 2808.13 &    1 & 1.1717 & & kissat-mab-esa \\
+       2 & 2812.93 &    2 & 1.1832 & & kissat-sc2022-bulky \\
+       3 & 2835.25 &    3 & 1.1862 & & ekissat-mab-gb-db \\
+       4 & 2835.59 &    4 & 1.1868 & & kissat-mab-ucb \\
+       5 & 2836.92 &    5 & 1.1868 & & kissat-inc \\
+       6 & 2845.19 &    6 & 1.1926 & & ekissat-mab-db-v1 \\
+       7 & 2846.73 &    7 & 1.1930 & & kissat-mab-moss \\
+       8 & 2857.67 &    8 & 1.1947 & & kissat-mab-hywalk \\
+       9 & 2869.45 &    9 & 1.1998 & & kissat-sc2022-light \\
+      10 & 2899.70 &   10 & 1.2164 & & kissat-els-v2 \\
+      11 & 2953.59 &   11 & 1.2290 & & hkis-unsat \\
+      12 & 2967.53 &   12 & 1.2347 & & kissat-adaptive-restart \\
+      13 & 2976.56 &   13 & 1.2475 & & seqfrost-noextend \\
+      14 & 3014.40 &   16 & 1.2645 & & kissat-els-v1 \\
+      15 & 3017.73 &   14 & 1.2509 & & cadical-esa \\
+      16 & 3036.83 &   15 & 1.2613 & & cadical-reorder \\
+      17 & 3049.90 &   20 & 1.3648 & & cadical-rel-scavel \\
+      18 & 3080.66 &   19 & 1.2965 & & kissat-relaxed \\
+      19 & 3095.73 &   17 & 1.2815 & & cadical-dvdl-v1 \\
+      20 & 3101.12 &   18 & 1.2856 & & cadical-dvdl-v2 \\
+      21 & 3273.95 &   21 & 1.3786 & & glucose-reboot \\
+      22 & 3290.90 &   25 & 1.4707 & & lstech-maple-hywalk \\
+      23 & 3292.68 &   23 & 1.4478 & & lstech-maple \\
+      24 & 3400.72 &   24 & 1.4693 & & slime-sc-2022-beta \\
+      25 & 3412.11 &   26 & 1.5237 & & slime-sc-2022 \\
+      26 & 3436.07 &   22 & 1.4161 & & hcad-v1-psids \\
+      27 & 3506.32 &   27 & 1.5433 & & maple-lcmdistchrbt-dl-v3 \\
+      28 & 4741.50 &   28 & 2.0581 & & isasat \\
+    \bottomrule
+  \end{tabular}
+\end{table}
+
+% TODO tfu: Amount of significant solver pairs (Wilcoxon signed-rank, 5% alpha), 
+% Amount of significant pairs:
+% PAR-2:   684/756 = 0.9048
+% timeout: 660/756 = 0.8730
+% k=3:     662/756 = 0.8757
+% k=4:     662/756 = 0.8757
+% k=5:     662/756 = 0.8757
+\begin{figure}[tbp]
+  \centering
+  \begin{subfigure}{0.32\textwidth}
+    \centering
+    \includegraphics[width=3.9cm]{../plots/par2sigdiff.pdf}
+    \caption{PAR-2 scores.}
+  \end{subfigure}
+  \begin{subfigure}{0.32\textwidth}
+    \centering
+    \includegraphics[width=3.9cm]{../plots/timeoutsigdiff.pdf}
+    \caption{Timeouts.}
+  \end{subfigure}
+  \begin{subfigure}{0.32\textwidth}
+    \centering
+    \includegraphics[width=3.9cm]{../plots/s3sigdiff.pdf}
+    \caption{$s_3$ scores.}
+  \end{subfigure}
+  \caption{Solver pairs with significant differences regarding the respective scorings are colored white.
+Non-significant differences are colored black.
+Solvers are ordered indentically to Tab.~\ref{tab:ranking}.}
+\end{figure}
+
+% TODO tf: Describe Table 2
+% TODO tf: Describe Table 3
+% TODO tf: Describe Figure 2
+% TODO tf: Describe Table 4
+% TODO tf: Describe Figure 3
+% TODO tf: Describe Table 5
+
+% TODO tf: extend, describe plots
 To define prediction targets, i.e., discrete runtime labels, we use hierarchical clustering with $k = 3$ and a log-single-link criterion, which produced the most \emph{useful} labels in preliminary experiments.
@@
 All these results show that discretized runtimes are suitable for our framework.
+
+% \subsection{Runtime Prediction Model}
+
+% TODO tf: extend, describe plots
+Our experiments only use one model configuration for runtime-label prediction since an exhaustive grid search would be infeasible.
+In preliminary experiments, we compared various model types from \emph{scikit-learn}~\cite{scikit-learn}.
+In particular, we conducted nested cross-validation, including hyper-parameter tuning, and used Matthews Correlation Coefficient~\cite{gorodkin2004comparing,matthews1975comparison} to assess the performance for predicting runtime labels.
+Our final choice is a stacking ensemble~\cite{wolpert1992stacked} of two prediction models, a quadratic-discriminant analysis~\cite{tharwat2016linear} and a random forest~\cite{breiman2001random}.
+Both these models can learn non-linear relationships between the instance features and the runtime labels.
+Stacking means that another prediction model, in our case a simple decision tree, decides which of the two ensemble members makes the prediction on which instance.
 
@@
 
-Regarding the stopping criteria (Fig.~\ref{fig:annitraincolorstopping}), the ranking stopping criterion performs well most consistently.
+Regarding the stopping criteria (Fig.~\ref{fig:annitraincolorstopping}), the ranking stopping criterion performs most consistently well.
 If accuracy is strongly favored (very high $\delta$), the Wilcoxon stopping criterion performs better.
@@
     }
-      \hline
+      \toprule
         & {AL} & {Random} & {Most Freq.} \\
-      \hline
+      \midrule
       Sampled Runtime Fraction (\%) & 5.41 & 5.43 & 5.44 \\
@@
       Ranking Accuracy (\%) & 90.48 & 88.54 & 81.08 \\
-      \hline
+      \bottomrule
     \end{tabular}
@@
     }
-      \hline
+      \toprule
         & {AL} & {Random} & {Most Freq.} \\
-      \hline
+      \midrule
       Sampled Runtime Fraction (\%) & 10.35 & 10.37 & 10.37 \\
@@
       Ranking Accuracy (\%) & 92.33 & 91.61 & 84.52 \\
-      \hline
+      \bottomrule
     \end{tabular}
@@
 
+\subsection{Instance Frequency in Selected Benchmarks}
+% TODO tf: Frequency of instances
+\begin{figure}
+  \centering
+  \includegraphics[width=0.6\textwidth]{../plots/instoccs.pdf}
+  \caption{Frequency of how often individual problem instances are selected within all active-learning benchmarks.
+Instances that are choosen more frequently are intuitively more important to distinguish solver runtimes.}
+\end{figure}
+
+% TODO tf: Describe Figure 7
+
+Text.
+
 \section{Conclusions and Future Work}

### 2024-04-16 paper_jar/paper.tex
Use Springer journal template

@@
-\documentclass[runningheads]{llncs}
+\documentclass{sn-jnl}
 
 \usepackage[ruled,vlined,linesnumbered,norelsize]{algorithm2e}
-\usepackage{amsmath,amssymb}
+\usepackage{amsmath,amssymb,amsfonts}
+\usepackage{amsthm}
 \usepackage[T1]{fontenc} % font encoding; recommended by LNCS template
@@
 \usepackage{booktabs}
+\usepackage{multirow}
 \usepackage[table]{xcolor}
@@
 
+%% as per the requirement new theorem styles can be included as shown below
+\theoremstyle{thmstyleone}%
+\newtheorem{theorem}{Theorem}%  meant for continuous numbers
+%%\newtheorem{theorem}{Theorem}[section]% meant for sectionwise numbers
+%% optional argument [theorem] produces theorem numbering sequence instead of independent numbers for Proposition
+\newtheorem{proposition}[theorem]{Proposition}% 
+%%\newtheorem{proposition}{Proposition}% to get separate numbers for theorem and proposition etc.
+
+\theoremstyle{thmstyletwo}%
+\newtheorem{example}{Example}%
+\newtheorem{remark}{Remark}%
+
+\theoremstyle{thmstylethree}%
+\newtheorem{definition}{Definition}%
+
+\raggedbottom
+
 \begin{document}
 
-\title{Active Learning for SAT Solver Benchmarking}
-
-\author{
-  Tobias Fuchs~\orcidlink{0000-0001-9727-2878} (\Letter) \and
-  Jakob Bach~\orcidlink{0000-0003-0301-2798} \and
-  Markus Iser~\orcidlink{0000-0003-2904-232X}
-}
-
-\institute{
-  Karlsruhe Institute of Technology (KIT), Karlsruhe, Germany \\
-  \email{info@tobiasfuchs.de, jakob.bach@kit.edu, markus.iser@kit.edu}
-}
-
-\authorrunning{T. Fuchs et al.} % according to sample paper of template: first names are abbreviated; if there are more than two authors, 'et al.' is used.
-
-\maketitle
-
-\begin{abstract}
+\title[Active Learning for SAT Solver Benchmarking]{Active Learning for SAT Solver Benchmarking}
+
+\author{\fnm{Tobias} \sur{Fuchs}}\email{tobias.fuchs@kit.edu}
+\author{\fnm{Jakob} \sur{Bach}}\email{jakob.bach@kit.edu}
+\author{\fnm{Markus} \sur{Iser}}\email{markus.iser@kit.edu}
+
+\affil{\orgname{Karlsruhe Institute of Technology}, \orgaddress{\city{Karlsruhe}, \country{Germany}}}
+
+\abstract{
   Benchmarking is a crucial phase when developing algorithms.
@@
   Overall, our tool provides a reliable way for solver engineers to determine a new solver's performance efficiently.
-
-  \keywords{Propositional satisfiability \and Benchmarking \and Active learning}
-\end{abstract}
+}
+
+\keywords{Propositional satisfiability, Benchmarking, Active learning}
+
+\maketitle
 
@@
   \label{tab:requirements}
-  ~\\[1em]
   \begin{tabular}{
-    m{0.32\textwidth}
-    >{\centering\arraybackslash}m{0.16\textwidth}
-    >{\centering\arraybackslash}m{0.16\textwidth}
-    >{\centering\arraybackslash}m{0.16\textwidth}
-    >{\centering\arraybackslash}m{0.16\textwidth}
+    m{0.2752\textwidth}
+    >{\centering\arraybackslash}m{0.1376\textwidth}
+    >{\centering\arraybackslash}m{0.1376\textwidth}
+    >{\centering\arraybackslash}m{0.1376\textwidth}
+    >{\centering\arraybackslash}m{0.1376\textwidth}
   }
@@
   \begin{tabular}[c]{ccc}
-  \begin{subfigure}[b]{0.25\textwidth}
+  \begin{subfigure}[b]{0.236\textwidth}
   \centering
@@
   &
-  \begin{subfigure}[b]{0.35\textwidth}
+  \begin{subfigure}[b]{0.33\textwidth}
   \centering
@@
   \caption{Regression performance of models trained on 46 instance and 27 runtime features. We report the RMSE and MAE metrics on the test sets.}
-  ~\\[1em]
   \begin{tabular}{
-    >{\raggedleft\arraybackslash}m{0.37\textwidth}
+    >{\raggedleft\arraybackslash}m{0.33855\textwidth}
     >{\centering\arraybackslash}m{0.01\textwidth}
-    >{\centering\arraybackslash}m{0.28\textwidth}
-    >{\centering\arraybackslash}m{0.28\textwidth}
+    >{\centering\arraybackslash}m{0.2562\textwidth}
+    >{\centering\arraybackslash}m{0.2562\textwidth}
   }
@@
   \begin{tabular}{
-    >{\raggedleft\arraybackslash}m{0.37\textwidth}
+    >{\raggedleft\arraybackslash}m{0.33855\textwidth}
     >{\centering\arraybackslash}m{0.01\textwidth}
-    >{\centering\arraybackslash}m{0.28\textwidth}
-    >{\centering\arraybackslash}m{0.28\textwidth}
+    >{\centering\arraybackslash}m{0.2562\textwidth}
+    >{\centering\arraybackslash}m{0.2562\textwidth}
   }
@@
   \caption{MCC performance of the best models predicting the $C_k$ clusterings.}
-  \begin{subtable}{\textwidth}
-    \centering
-    \caption{Predicting $C_3$ clusters.}
   \begin{tabular}{
-      >{\raggedleft\arraybackslash}m{0.37\textwidth}
+    >{\raggedleft\arraybackslash}m{0.33855\textwidth}
     >{\centering\arraybackslash}m{0.01\textwidth}
-      >{\centering\arraybackslash}m{0.28\textwidth}
-      >{\centering\arraybackslash}m{0.28\textwidth}
+    >{\centering\arraybackslash}m{0.2562\textwidth}
+    >{\centering\arraybackslash}m{0.2562\textwidth}
   }
@@
     Quadratic-DA (QDA)                    & & 0.6903 ($\pm$ 0.0607) & 0.7738 ($\pm$ 0.0459) \\[0.4ex]
-      Random-Forest-Classifier (RF)         & & 0.7116 ($\pm$ 0.0385) & 0.7841 ($\pm$ 0.0469) \\
-      \bottomrule
-    \end{tabular}
-  \end{subtable}
-  \begin{subtable}{\textwidth}
-    \centering
-    \caption{Predicting $C_4$ clusters.}
-    \begin{tabular}{
-      >{\raggedleft\arraybackslash}m{0.37\textwidth}
-      >{\centering\arraybackslash}m{0.01\textwidth}
-      >{\centering\arraybackslash}m{0.28\textwidth}
-      >{\centering\arraybackslash}m{0.28\textwidth}
-    }
-      \toprule
+    Random-Forest-Classifier (RF)         & & 0.7116 ($\pm$ 0.0385) & 0.7841 ($\pm$ 0.0469) \\[0.4ex]
+    \midrule
     {$C_4$ Cluster Prediction Models} & & {46\,+\,27 Runtime Feat. Avg. MCC ($\pm$ std)} & {46\,+\,27\,+\,27 Clus. Feat. Avg. MCC ($\pm$ std)} \\
@@
     Quadratic-DA (QDA)                    & & 0.6007 ($\pm$ 0.0765) & 0.6872 ($\pm$ 0.0469) \\[0.4ex]
-      Random-Forest-Classifier (RF)         & & 0.6033 ($\pm$ 0.0455) & 0.6828 ($\pm$ 0.0527) \\
-      \bottomrule
-    \end{tabular}
-  \end{subtable}
-  \begin{subtable}{\textwidth}
-    \centering
-    \caption{Predicting $C_5$ clusters.}
-    \begin{tabular}{
-      >{\raggedleft\arraybackslash}m{0.37\textwidth}
-      >{\centering\arraybackslash}m{0.01\textwidth}
-      >{\centering\arraybackslash}m{0.28\textwidth}
-      >{\centering\arraybackslash}m{0.28\textwidth}
-    }
-      \toprule
+    Random-Forest-Classifier (RF)         & & 0.6033 ($\pm$ 0.0455) & 0.6828 ($\pm$ 0.0527) \\[0.4ex]
+    \midrule
     {$C_5$ Cluster Prediction Models} & & {46\,+\,27 Runtime Feat. Avg. MCC ($\pm$ std)} & {46\,+\,27\,+\,27 Clus. Feat. Avg. MCC ($\pm$ std)} \\
@@
   \end{tabular}
-  \end{subtable}
 \end{table}
@@
   \begin{tabular}{
-    >{\centering\arraybackslash}m{0.07\textwidth}
-    >{\centering\arraybackslash}m{0.12\textwidth}
-    >{\centering\arraybackslash}m{0.07\textwidth}
-    >{\centering\arraybackslash}m{0.12\textwidth}
+    >{\centering\arraybackslash}m{0.09\textwidth}
+    >{\centering\arraybackslash}m{0.15\textwidth}
+    >{\centering\arraybackslash}m{0.09\textwidth}
+    >{\centering\arraybackslash}m{0.15\textwidth}
     >{\raggedright\arraybackslash}m{0.01\textwidth}
-    >{\raggedright\arraybackslash}m{0.32\textwidth}
+    >{\raggedright\arraybackslash}m{0.314\textwidth}
   }
     \toprule
-    \multicolumn{2}{c}{PAR-2} & \multicolumn{2}{c}{$s_3$} \\
+    \multicolumn{2}{c}{PAR-2} & \multicolumn{2}{c}{$s_3$} & & \multirow[c]{2}{*}{\vspace{-0.15cm}\hspace{-0.02cm}Solver} \\
     \cmidrule(r){1-2}\cmidrule(r){3-4}
-    Rank &   Score & Rank &  Score & & Solver \\
+    Rank &   Score & Rank &  Score & & \\
     \midrule
@@
 
-\begin{table}[t]
+\begin{table}[tp]
   \centering
   \caption{
-      Performance comparison (on the full dataset) of the best-performing active-learning approaches (\emph{AL}), random sampling of the same runtime fraction with 1000 repetitions (\emph{Random}), and statically selecting the instances most frequently sampled by active-learning approaches (\emph{Most Freq.})
+    Performance comparison (on the full dataset) of the best-performing AL approach (\emph{AL}), random sampling of the same runtime fraction with 1000 repetitions (\emph{Random}), and statically selecting the instances most frequently sampled by active-learning approaches (\emph{Most Frequent})
   }
   \label{tab:fulldataset}
-
-  \begin{subfigure}{\textwidth}
-    \centering
-    \caption{Best-performing AL approach for $\delta \in \left[0.2, 0.7\right]$}
   \begin{tabular}{
@@
     \toprule
-        & {AL} & {Random} & {Most Freq.} \\
+    {For $\delta \in \left[0.2, 0.7\right]$} & {AL} & {Random} & {Most Frequent} \\
     \midrule
@@
     Ranking Accuracy (\%) & 90.48 & 88.54 & 81.08 \\
-      \bottomrule
-    \end{tabular}
-  \end{subfigure}
-  \\
-  \begin{subfigure}{\textwidth}
-    \centering
-    \caption{Best-performing AL approach for $\delta \in (0.7, 0.8]$}
-    \begin{tabular}{
-      >{\arraybackslash}p{0.4\textwidth}
-      S[table-format=2.2,table-column-width=0.15\textwidth]
-      S[table-format=2.2,table-column-width=0.15\textwidth]
-      S[table-format=2.2,table-column-width=0.15\textwidth]
-    }
-      \toprule
-        & {AL} & {Random} & {Most Freq.} \\
+    \midrule
+    {For $\delta \in (0.7, 0.8]$}  & {AL} & {Random} & {Most Frequent} \\
     \midrule
@@
   \end{tabular}
-  \end{subfigure}
 \end{table}

### 2024-04-22 paper_jar/paper.tex
Add a few To Dos
Minor adaptations to SN journal template
- remove unused packages
- some headings become separate-line instead of inline; remove full stop
  accordingly
- change bibstyle
- remove unused commands in preamble
- mark corresponding author as intended (instead of using custom symbol)
- add declarations

@@
-\documentclass{sn-jnl}
+\documentclass[sn-basic, Numbered]{sn-jnl} % may use "lineno" for line numbers
+% template from https://www.springernature.com/gp/authors/campaigns/latex-author-support
 
@@
 \usepackage{amsthm}
-\usepackage[T1]{fontenc} % font encoding; recommended by LNCS template
+\usepackage{booktabs}
 \usepackage{graphicx}
 \usepackage{hyperref}
-\usepackage[misc]{ifsym} % letter symbol to mark corresponding author
+\usepackage{multirow}
 \usepackage{pifont}
@@
 \usepackage{tabularx}
-\usepackage{booktabs}
-\usepackage{multirow}
 \usepackage[table]{xcolor}
-\usepackage{orcidlink} % ORCID icon and link; has to be loaded after "xcolor" here, else "option clash"; class "llncs" also has command "\orcidID" (which only displays full ID), but camera-ready instructions state you should add icon and link yourself
 \usepackage{tikz} % has to be loaded after "xcolor" here, else "option clash"
@@
 
-\renewcommand\UrlFont{\color{blue}\rmfamily} % link coloring according to LNCS template
-
-%% as per the requirement new theorem styles can be included as shown below
-\theoremstyle{thmstyleone}%
-\newtheorem{theorem}{Theorem}%  meant for continuous numbers
-%%\newtheorem{theorem}{Theorem}[section]% meant for sectionwise numbers
-%% optional argument [theorem] produces theorem numbering sequence instead of independent numbers for Proposition
-\newtheorem{proposition}[theorem]{Proposition}% 
-%%\newtheorem{proposition}{Proposition}% to get separate numbers for theorem and proposition etc.
-
-\theoremstyle{thmstyletwo}%
-\newtheorem{example}{Example}%
-\newtheorem{remark}{Remark}%
-
-\theoremstyle{thmstylethree}%
-\newtheorem{definition}{Definition}%
+\newtheorem{definition}{Definition} % only kind of theorems we use are definitions (well, one)
 
@@
 
-\title[Active Learning for SAT Solver Benchmarking]{Active Learning for SAT Solver Benchmarking}
-
-\author{\fnm{Tobias} \sur{Fuchs}}\email{tobias.fuchs@kit.edu}
+\title[Active Learning for SAT Solver Benchmarking]{Active Learning for SAT Solver Benchmarking} % short and long title
+%JB: not sure if we should use a new title or not; in past special issues of journal, both options (re-use conference-version title or choose new title) were used
+
+\author*{\fnm{Tobias} \sur{Fuchs}}\email{tobias.fuchs@kit.edu}
 \author{\fnm{Jakob} \sur{Bach}}\email{jakob.bach@kit.edu}
@@
 
-\affil{\orgname{Karlsruhe Institute of Technology}, \orgaddress{\city{Karlsruhe}, \country{Germany}}}
+\affil{\orgname{Karlsruhe Institute of Technology (KIT)}, \orgaddress{\city{Karlsruhe}, \country{Germany}}}
 
@@
 \label{sec:intro}
+
+% TODO: somewhere in introduction, refer to conference version ("fuchs2023active") and state what has changed in journal version
+% TODO: if appropriate, extend with motivation/results regarding runtime discretization and prediction models
 
@@
 
+% TODO: if appropriate, extend with related work regarding runtime discretization and prediction models
+
 Benchmarking is not only of high interest in many fields but also an active research area on its own.
@@
 
-\subsubsection{Static Benchmark Selection.}
+\subsection{Static Benchmark Selection}
 
@@
 
-\subsubsection{Algorithm Configuration.}
+\subsection{Algorithm Configuration}
 
@@
 
-\subsubsection{Incremental Benchmark Selection.}
+\subsection{Incremental Benchmark Selection}
 
@@
 
-\subsubsection{Active Learning.}
+\subsection{Active Learning}
 
@@
 
-\subsubsection{Runtime Transformation.}
+\subsubsection{Runtime Transformation}
 
@@
 
-\subsubsection{Ranking Solvers.}
+\subsubsection{Ranking Solvers}
 
@@
 
-\subsubsection{Uncertainty Sampling.}
+\subsubsection{Uncertainty Sampling}
 
@@
 
-\subsubsection{Information-Gain Sampling.}
+\subsubsection{Information-Gain Sampling}
 
@@
 
-\subsubsection{Wilcoxon Stopping Criterion.}
+\subsubsection{Wilcoxon Stopping Criterion}
 
@@
 
-\subsubsection{Ranking Stopping Criterion.}
+\subsubsection{Ranking Stopping Criterion}
 
@@
 
-\subsubsection{Ranking.}
+\subsubsection{Ranking}
 
@@
 
-\subsubsection{Selection.}
+\subsubsection{Selection}
 
@@
 
-\subsubsection{Stopping.}
+\subsubsection{Stopping}
 
@@
 We repeat sampling 1000 times and report average results.
-\emph{Most Freq.} uses a static benchmark set consisting of those instances most frequently sampled by our active learning approach.
+\emph{Most Frequent} uses a static benchmark set consisting of those instances most frequently sampled by our active learning approach.
 In particular, we consider the average sampling frequency over all solvers and Pareto-optimal active-learning approaches.
@@
 
+% TODO: if appropriate, extend with results regarding runtime discretization and prediction models
+
 In this work, we have addressed the \emph{New-Solver Problem}:
@@
 
-\subsubsection*{Acknowledgments.}
+\backmatter
+
+\section*{Declarations}
+
+\bmhead{Funding}
 
@@
 
-%
-% ---- Bibliography ----
-%
-\bibliographystyle{splncs04}
+\bmhead{Competing interests}
+
+The authors have no competing interests to declare that are relevant to the content of this article.
+
+\bmhead{Availability of data and materials}
+
+All experimental data are available online at \url{https://github.com/mathefuchs/al-for-sat-solver-benchmarking-data}.
+% TODO: does this also include all the data for the newly added results?
+
+\bmhead{Code availability}
+
+The code is available online at \url{https://github.com/mathefuchs/al-for-sat-solver-benchmarking}.
+%TODO: README in code repo should also describe how to reproduce the experiments for the newly added results (at least running the corresponding plot notebook is necessary; currently, README only refers to "anni_results.ipynb")
+
 \bibliography{literature}

### 2024-04-22 paper_jar/paper.tex
Refer to conference version in Intro

@@
 \title[Active Learning for SAT Solver Benchmarking]{Active Learning for SAT Solver Benchmarking} % short and long title
-%JB: not sure if we should use a new title or not; in past special issues of journal, both options (re-use conference-version title or choose new title) were used
 
@@
 \label{sec:intro}
-
-% TODO: somewhere in introduction, refer to conference version ("fuchs2023active") and state what has changed in journal version
 % TODO: if appropriate, extend with motivation/results regarding runtime discretization and prediction models
@@
 On average, our predictions reach about \SI{92}{\%} accuracy with only about \SI{10}{\%} of the runtime required to evaluate these solvers on the complete set of instances.
-
 Our entire source code\footnote{\url{https://github.com/mathefuchs/al-for-sat-solver-benchmarking}} and experimental data\footnote{\url{https://github.com/mathefuchs/al-for-sat-solver-benchmarking-data}} are available on GitHub.
+
+This article is an extended version of a conference paper with the same name~\cite{fuchs2023active}.
+% TODO: not sure if we should choose a new title or not; in past special issues of journal, both options (re-use conference-version title or choose new title) were used
+In particular, we significantly extended the analysis of runtime-prediction approaches (cf.~Section~\ref{sec:exdesign:disc-pred}) and added an evaluation of how often problem instances were selected in benchmarks (cf.~Section~\ref{sec:eval:inst-freq}).
+% TODO: before submission, we should check which other changes we made and list them here accordingly
 
@@
 \subsection{Runtime Discretization and Prediction}
+\label{sec:exdesign:disc-pred}
 
@@
 \subsection{Instance Frequency in Selected Benchmarks}
+\label{sec:eval:inst-freq}
+
 % TODO tf: Frequency of instances

### 2024-05-22 paper_jar/paper.tex
draft of tables and figures descriptions

@@
 
-% TODO tf: Why not regression? Bad performance
+% MI: table with regression performance:
+One of the most crucial parts of our framework is the choice of the surrogate model to predict the runtime labels of the new solver from the instance feature vector.
+The instance feature vector consists of 46 instance features and 27 known solver runtimes.
+Table~\ref{tab:regression} shows the prediction performance of two regression models, a Random-Forest-Regressor and an MLP-Regressor, on the test set.
+As can be seen, both models perform poorly, which is why we decided to model the runtimes as discrete labels.
+
 \begin{table}[htbp]
@@
   \caption{Regression performance of models trained on 46 instance and 27 runtime features. We report the RMSE and MAE metrics on the test sets.}
+  \label{tab:regression}
   \begin{tabular}{
@@
 
-% TODO tf: Timeout prediction performance
+% MI: table with timeout prediction performance: (merged with existing text)
+Table~\ref{tab:timeout-prediction} shows the performance of different models in predicting timeouts.
+While the second column shows the performance of models trained on the 46 instance features, the third column shows the performance of models trained on the 46 instance features and the 27 runtime features.
+While the Quadratic-DA model performs poorly on the 46 instance features, it performs well on the 46 instance features and 27 runtime features.
+The best prediction performance however is achieved by the stacking ensemble~\cite{wolpert1992stacked}, which combines quadratic-discriminant analysis~\cite{tharwat2016linear} and a random forest~\cite{breiman2001random}.
+Stacking means that another prediction model, in our case a simple decision tree, decides which of the two ensemble members makes the prediction on which instance.
+
 \begin{table}[htbp]
@@
   \caption{MCC performance of models predicting timeouts. The models are trained 28 times with each solver as a target once.}
+  \label{tab:timeout-prediction}
   ~\\[1em]
@@
 
-% TODO tf: Cluster prediction performance
+% MI: table with cluster prediction performance: (merged with existing text)
+Runtime prediction however is a very rough discretization of the runtimes which is not very informative.
+To define prediction targets, i.e., discrete runtime labels, we use hierarchical clustering with $k$ clusters and a log-single-link criterion, which produced the most \emph{useful} labels in preliminary experiments.
+We denote this adapted solver scoring function with $s_k$.
+In our chosen hierarchical procedure, each non-timeout runtime starts in a separate interval.
+We then gradually merge intervals whose single-link logarithmic distance is the smallest until the desired number of partitions is reached.
+Other clustering approaches that we tried include hierarchical clustering with mean-, median-, and complete-link criterion, as well as $k$-means and spectral clustering.
+
+% MI: table with cluster prediction performance:
+We evaluated the performance of the best models from Table~\ref{tab:timeout-prediction} in predicting the such determined runtime clusters.
+Table~\ref{tab:cluster-prediction} shows their performance in predicting the $C_k$ clusterings for $k = 3, 4, 5$.
+While the second column shows the prediction performance of the models trained on the 46 instance features including the 27 known solver runtimes, the third column shows the prediction performance of the models trained on those features plus 27 features representing the known $C_k$ clustering labels of the known solver runtimes.
+We observe that the models perform better when the clustering labels are included as features.
+Moreover, the prediction performance degradades rapidly with an increasing number of clusters.
+In all cases, the stacking ensemble performs best.
+
 \begin{table}[htbp]
@@
   \caption{MCC performance of the best models predicting the $C_k$ clusterings.}
+  \label{tab:cluster-prediction}
   \begin{tabular}{
@@
 
-% TODO tf: Confusion matrices
-\begin{figure}[tbp]
+% MI: table with runtime prediction performance: (merged with existing text)
+Our final choice for the prediction model is the stacking ensemble of a quadratic-discriminant analysis and a random forest.
+Both these models can learn non-linear relationships between the instance features and the runtime labels.
+Figure~\ref{fig:confusion-matrices} shows the confusion matrices of the stacking ensemble for predicting timeouts and $C_3$ clusters.
+Both perform reasonably well, with the $C_3$ cluster prediction model performing slightly worse than the timeout prediction model.
+
+\begin{figure}[htbp]
   \centering
@@
   \caption{Confusion matrices for predicting timeouts and $C_3$ clusters.}
+  \label{fig:confusion-matrices}
 \end{figure}
+
+% MI: table with rankings after discretization: (merged with existing text)
+But to obtain \emph{useful} labels, we also need to ensure that discretized labels still discriminate solvers and align with the actual PAR-2 ranking.
+Table~\ref{tab:ranking} shows the ranking of the solvers in the SAT Competition~2022 Anniversary Track~\cite{sat2022} according to the standard PAR-2 runtime score and according to our discretized $s_3$ score.
+We observe that our ranking approach correctly decides for almost all (about \SI{97.45}{\%}; $\sigma = \SI{3.68}{\%}$) solver pairs which solver is faster.
+In particular, the Spearman correlation of $s_3$ and PAR-2 ranking is about \SI{0.988}{}, which is very close to the optimal value of 1~\cite{de2016comparing}.
 
@@
 
+% MI: table with solver confusion matrices:
+We further analyzed how runtime discretization affects the distinguishability of solver pairs.
+Figure~\ref{fig:solver-confusion-matrices} shows the confusion matrices of solver pairs with significant differences in their PAR-2 scores, timeouts, and $s_3$ scores.
+According to a Wilcoxon-signed-rank test with $\alpha = 0.05$, \SI{87.83}{\%} of solver pairs have significantly different scores after discretization, only a slight drop compared to \SI{89.95}{\%} before discretization.
+All these results show that discretized runtimes are suitable for our framework.
+
 % TODO tfu: Amount of significant solver pairs (Wilcoxon signed-rank, 5% alpha), 
@@
 Solvers are ordered indentically to Tab.~\ref{tab:ranking}.}
+  \label{fig:solver-confusion-matrices}
 \end{figure}
 
-% TODO tf: Describe Table 2
-% TODO tf: Describe Table 3
-% TODO tf: Describe Figure 2
-% TODO tf: Describe Table 4
-% TODO tf: Describe Figure 3
-% TODO tf: Describe Table 5
-
 % TODO tf: extend, describe plots
-To define prediction targets, i.e., discrete runtime labels, we use hierarchical clustering with $k = 3$ and a log-single-link criterion, which produced the most \emph{useful} labels in preliminary experiments.
-We denote this adapted solver scoring function with $s_3$.
-In our chosen hierarchical procedure, each non-timeout runtime starts in a separate interval.
-We then gradually merge intervals whose single-link logarithmic distance is the smallest until the desired number of partitions is reached.
-Other clustering approaches that we tried include hierarchical clustering with mean-, median-, and complete-link criterion, as well as $k$-means and spectral clustering.
-
-To obtain \emph{useful} labels, we need to ensure that discretized labels still discriminate solvers and align with the actual PAR-2 ranking.
-We analyzed the ranking induced by $s_3$ in preliminary experiments with the SAT Competition~2022 Anniversary Track~\cite{sat2022}.
-According to a Wilcoxon-signed-rank test with $\alpha = 0.05$, \SI{87.83}{\%} of solver pairs have significantly different scores after discretization, only a slight drop compared to \SI{89.95}{\%} before discretization.
-Further, our ranking approach correctly decides for almost all (about \SI{97.45}{\%}; $\sigma = \SI{3.68}{\%}$) solver pairs which solver is faster.
-In particular, the Spearman correlation of $s_3$ and PAR-2 ranking is about \SI{0.988}{}, which is very close to the optimal value of 1~\cite{de2016comparing}.
-All these results show that discretized runtimes are suitable for our framework.
-
-% \subsection{Runtime Prediction Model}
-
-% TODO tf: extend, describe plots
-Our experiments only use one model configuration for runtime-label prediction since an exhaustive grid search would be infeasible.
-In preliminary experiments, we compared various model types from \emph{scikit-learn}~\cite{scikit-learn}.
-In particular, we conducted nested cross-validation, including hyper-parameter tuning, and used Matthews Correlation Coefficient~\cite{gorodkin2004comparing,matthews1975comparison} to assess the performance for predicting runtime labels.
-Our final choice is a stacking ensemble~\cite{wolpert1992stacked} of two prediction models, a quadratic-discriminant analysis~\cite{tharwat2016linear} and a random forest~\cite{breiman2001random}.
-Both these models can learn non-linear relationships between the instance features and the runtime labels.
-Stacking means that another prediction model, in our case a simple decision tree, decides which of the two ensemble members makes the prediction on which instance.
+%Our experiments only use one model configuration for runtime-label prediction since an exhaustive grid search would be infeasible.
+%In preliminary experiments, we compared various model types from \emph{scikit-learn}~\cite{scikit-learn}.
+%In particular, we conducted nested cross-validation, including hyper-parameter tuning, and used Matthews Correlation Coefficient~\cite{gorodkin2004comparing,matthews1975comparison} to assess the performance for predicting runtime labels.
 
@@
 
-\subsection{Instance-Family Importance}
+\subsection{Instance Frequency and Instance-Family Importance}
+
+% TODO tf: Frequency of instances
+\begin{figure}[htbp]
+  \centering
+  \includegraphics[width=0.6\textwidth]{../plots/instoccs.pdf}
+  \caption{Frequency of how often individual problem instances are selected within all active-learning benchmarks.
+Instances that are choosen more frequently are intuitively more important to distinguish solver runtimes.}
+  \label{fig:instoccs}
+\end{figure}
+
+% MI: instance frequency (not so much to say, TODO: maybe add some more text)
+We analyze the frequency of individual instances being selected by our active-learning approach.
+Fig.~\ref{fig:instoccs} shows the distribution of how often individual instances are selected within all active-learning benchmarks.
+We observe that a small fraction of instances is selected very often, while most instances are selected only a few times.
 
@@
 
-\subsection{Instance Frequency in Selected Benchmarks}
-\label{sec:eval:inst-freq}
-
-% TODO tf: Frequency of instances
-\begin{figure}
-  \centering
-  \includegraphics[width=0.6\textwidth]{../plots/instoccs.pdf}
-  \caption{Frequency of how often individual problem instances are selected within all active-learning benchmarks.
-Instances that are choosen more frequently are intuitively more important to distinguish solver runtimes.}
-\end{figure}
-
-% TODO tf: Describe Figure 7
-
-Text.
-
 \section{Conclusions and Future Work}

### 2024-05-23 paper_jar/paper.tex
Revise text

@@
 This article is an extended version of a conference paper with the same name~\cite{fuchs2023active}.
-% TODO: not sure if we should choose a new title or not; in past special issues of journal, both options (re-use conference-version title or choose new title) were used
 In particular, we significantly extended the analysis of runtime-prediction approaches (cf.~Section~\ref{sec:exdesign:disc-pred}) and added an evaluation of how often problem instances were selected in benchmarks (cf.~Section~\ref{sec:eval:inst-freq}).
@@
 \section{Related Work}
-
-% TODO: if appropriate, extend with related work regarding runtime discretization and prediction models
 
@@
 \label{sec:exdesign:disc-pred}
+One of the most crucial parts of our framework is the choice of the surrogate model to predict runtime labels and inform active-learning selection decisions.
+In this section, we detail why we do \emph{not} directly predict absolute solver runtimes, which machine-learning models work best in predicting the time-outs of a solver, how clustering solver runtimes can lead to more nuanced runtime labels, and how ML models perform on those cluster labels.
 
 % MI: table with regression performance:
-One of the most crucial parts of our framework is the choice of the surrogate model to predict the runtime labels of the new solver from the instance feature vector.
+Table~\ref{tab:regression} shows the prediction performance of two regression models, a Random-Forest-Regressor and an MLP-Regressor, on the test set.
+Both models have been chosen because of their robustness and capability of capturing non-linear dependencies~\cite{breiman2001random}.
 The instance feature vector consists of 46 instance features and 27 known solver runtimes.
-Table~\ref{tab:regression} shows the prediction performance of two regression models, a Random-Forest-Regressor and an MLP-Regressor, on the test set.
-As can be seen, both models perform poorly, which is why we decided to model the runtimes as discrete labels.
-
-\begin{table}[htbp]
+Both models perform poorly, which is evident from the magnitudes of the average regression errors.
+An RMSE error of about \SI{2000}{s} is quite poor compared to the solver runtimes which range from 0 to \SI{5000}{s}.
+Therefore, we decided to discretize runtimes.
+
+\begin{table}[tbp]
   \centering
@@
 While the Quadratic-DA model performs poorly on the 46 instance features, it performs well on the 46 instance features and 27 runtime features.
-The best prediction performance however is achieved by the stacking ensemble~\cite{wolpert1992stacked}, which combines quadratic-discriminant analysis~\cite{tharwat2016linear} and a random forest~\cite{breiman2001random}.
+The best prediction performance however is achieved by a stacking ensemble~\cite{wolpert1992stacked}, which combines quadratic-discriminant analysis~\cite{tharwat2016linear} and a random-forest classifier~\cite{breiman2001random}.
 Stacking means that another prediction model, in our case a simple decision tree, decides which of the two ensemble members makes the prediction on which instance.
 
-\begin{table}[htbp]
+\begin{table}[tbp]
   \centering
@@
 % MI: table with cluster prediction performance: (merged with existing text)
-Runtime prediction however is a very rough discretization of the runtimes which is not very informative.
-To define prediction targets, i.e., discrete runtime labels, we use hierarchical clustering with $k$ clusters and a log-single-link criterion, which produced the most \emph{useful} labels in preliminary experiments.
-We denote this adapted solver scoring function with $s_k$.
+As already mentioned, runtime prediction is quite challenging.
+Instead, we discretize runtimes.
+To define prediction targets, we use hierarchical clustering with $k$ clusters and a log-single-link criterion, which produced the most \emph{useful} labels in preliminary experiments.
+We denote this adapted solver scoring function with $s_k$ as defined in~\eqref{eq:rankingeq}.
 In our chosen hierarchical procedure, each non-timeout runtime starts in a separate interval.
 We then gradually merge intervals whose single-link logarithmic distance is the smallest until the desired number of partitions is reached.
+The clustering is instance-specific, i.e., the boundaries are chosen such that it produces groups of solvers with similar performances for each instance.
 Other clustering approaches that we tried include hierarchical clustering with mean-, median-, and complete-link criterion, as well as $k$-means and spectral clustering.
@@
 % MI: table with cluster prediction performance:
-We evaluated the performance of the best models from Table~\ref{tab:timeout-prediction} in predicting the such determined runtime clusters.
-Table~\ref{tab:cluster-prediction} shows their performance in predicting the $C_k$ clusterings for $k = 3, 4, 5$.
+We evaluated the performance of the best models from Table~\ref{tab:timeout-prediction} in predicting these runtime clusters as mentioned above.
+Table~\ref{tab:cluster-prediction} shows their performance in predicting the $C_k$ clusterings for $k \in \lbrace 3, 4, 5 \rbrace$.
 While the second column shows the prediction performance of the models trained on the 46 instance features including the 27 known solver runtimes, the third column shows the prediction performance of the models trained on those features plus 27 features representing the known $C_k$ clustering labels of the known solver runtimes.
 We observe that the models perform better when the clustering labels are included as features.
-Moreover, the prediction performance degradades rapidly with an increasing number of clusters.
+Moreover, the prediction performance degrades rapidly with an increasing number of clusters.
 In all cases, the stacking ensemble performs best.
 
-\begin{table}[htbp]
+\begin{table}[tbp]
   \centering
@@
 Both perform reasonably well, with the $C_3$ cluster prediction model performing slightly worse than the timeout prediction model.
-
-\begin{figure}[htbp]
+Notably, for the $C_3$ prediction model, there is less confusion between the non-timeout labels (labels 1 and 2) and the timeout label (label 3) than between the non-timeout labels (labels 1 and 2).
+This indicates that deciding timeouts is easier than distinguishing different groups of solvers.
+
+\begin{figure}[tbp]
   \centering
@@
 % MI: table with rankings after discretization: (merged with existing text)
-But to obtain \emph{useful} labels, we also need to ensure that discretized labels still discriminate solvers and align with the actual PAR-2 ranking.
+To obtain \emph{useful} labels, we also need to ensure that the discretized labels still discriminate solvers and align with the actual PAR-2 ranking.
 Table~\ref{tab:ranking} shows the ranking of the solvers in the SAT Competition~2022 Anniversary Track~\cite{sat2022} according to the standard PAR-2 runtime score and according to our discretized $s_3$ score.
-We observe that our ranking approach correctly decides for almost all (about \SI{97.45}{\%}; $\sigma = \SI{3.68}{\%}$) solver pairs which solver is faster.
-In particular, the Spearman correlation of $s_3$ and PAR-2 ranking is about \SI{0.988}{}, which is very close to the optimal value of 1~\cite{de2016comparing}.
+We observe that our ranking approach correctly decides almost all (about \SI{97.45}{\%}; $\sigma = \SI{3.68}{\%}$) solver pairs, i.e., which solver is faster among two given solvers.
+In particular, the Spearman correlation of $s_3$ and PAR-2 ranking is about \SI{0.988}{}, which is very close to the optimal value of~1~\cite{de2016comparing}.
 
@@
 % * Spearman correlation of PAR2 and Discrete rank: 0.9862
-\begin{table}[htbp]
+\begin{table}[tbp]
   \centering
@@
 Figure~\ref{fig:solver-confusion-matrices} shows the confusion matrices of solver pairs with significant differences in their PAR-2 scores, timeouts, and $s_3$ scores.
-According to a Wilcoxon-signed-rank test with $\alpha = 0.05$, \SI{87.83}{\%} of solver pairs have significantly different scores after discretization, only a slight drop compared to \SI{89.95}{\%} before discretization.
-All these results show that discretized runtimes are suitable for our framework.
+According to a Wilcoxon signed-rank test with level $\alpha = 0.05$, \SI{87.83}{\%} of solver pairs have significantly different scores after discretization, only a slight drop compared to \SI{89.95}{\%} before discretization.
+These results underpin the usefulness of discretized runtimes for our framework.
 
@@
 
-% TODO tf: extend, describe plots
-%Our experiments only use one model configuration for runtime-label prediction since an exhaustive grid search would be infeasible.
-%In preliminary experiments, we compared various model types from \emph{scikit-learn}~\cite{scikit-learn}.
-%In particular, we conducted nested cross-validation, including hyper-parameter tuning, and used Matthews Correlation Coefficient~\cite{gorodkin2004comparing,matthews1975comparison} to assess the performance for predicting runtime labels.
-
 \subsection{Implementation Details}
@@
 
-% TODO tf: Frequency of instances
-\begin{figure}[htbp]
+\begin{figure}[tbp]
   \centering
@@
   \caption{Frequency of how often individual problem instances are selected within all active-learning benchmarks.
-Instances that are choosen more frequently are intuitively more important to distinguish solver runtimes.}
+Instances that are chosen more frequently are intuitively more important to distinguish solver runtimes.}
   \label{fig:instoccs}
 \end{figure}
-
-% MI: instance frequency (not so much to say, TODO: maybe add some more text)
-We analyze the frequency of individual instances being selected by our active-learning approach.
-Fig.~\ref{fig:instoccs} shows the distribution of how often individual instances are selected within all active-learning benchmarks.
-We observe that a small fraction of instances is selected very often, while most instances are selected only a few times.
 
@@
 \end{figure}
+
+% MI: instance frequency (not so much to say, TODO: maybe add some more text)
+We analyze the frequency of individual instances being selected by our active-learning approach.
+Fig.~\ref{fig:instoccs} shows the distribution of how often individual instances are selected within all active-learning benchmarks.
+We observe that a small fraction of instances is selected very often, while most instances are selected only a few times.
+There is no fixed subset of instances that can perfectly distinguish all solvers, which shows the usefulness of employing an active-learning-based strategy.
 
@@
 
-% TODO: if appropriate, extend with results regarding runtime discretization and prediction models
-
 In this work, we have addressed the \emph{New-Solver Problem}:
@@
 All experimental data are available online at \url{https://github.com/mathefuchs/al-for-sat-solver-benchmarking-data}.
-% TODO: does this also include all the data for the newly added results?
 
@@
 The code is available online at \url{https://github.com/mathefuchs/al-for-sat-solver-benchmarking}.
-%TODO: README in code repo should also describe how to reproduce the experiments for the newly added results (at least running the corresponding plot notebook is necessary; currently, README only refers to "anni_results.ipynb")
 

### 2024-05-24 paper_jar/paper.tex
section 1

@@
 One of the main phases of algorithm engineering is benchmarking.
-This also applies to propositional satisfiability (SAT), the archetypal $\mathcal{NP}$-complete problem.
+This also applies to propositional satisfiability (SAT), the canonical $\mathcal{NP}$-complete problem.
 Benchmarking is, however, quite expensive regarding the runtime of experiments.
-While benchmarking a single SAT solver might still be feasible, developing new, competitive SAT solvers requires extensive experimentation with a variety of ideas~\cite{FroleyksHIJS21,sat2022}.
+While it is still feasible to benchmark a single or fixed amount of SAT solvers, developing new, competitive SAT solvers requires extensive experimentation with a variety of ideas, which often results in a combinatorial explosion of the configuration space of each algorithm~\cite{HutterHL11}.
 In particular, a new solver idea is rarely best on the first try.
 Thus, it is highly desirable to reduce benchmarking time and discard unpromising ideas early, allowing to test more approaches or spend more time on promising ones.
-The field of SAT solver benchmarking is well established, but traditional benchmark selection approaches do not optimize benchmark runtime.
-Instead, they focus on selecting a representative set of instances for scoring solvers~\cite{Gelder11,HoosKSS13}.
-For the latter, SAT~Competitions typically employ the \mbox{PAR-2} score, i.e., the average runtime with a penalty of $2 \tau$ for timeouts with time-limit~$\tau$~\cite{FroleyksHIJS21}.
+The field of SAT solver benchmarking is well established, yet traditional benchmark selection approaches do not optimize benchmark runtime.
+Instead, they focus on selecting a representative set of instances for ranking solvers~\cite{Gelder11,HoosKSS13}.
+With regard to the ranking of solvers, SAT Competitions typically employ the \mbox{PAR-2} score, which, given a time limit of $\tau$, is the average runtime with a penalty of $2 \tau$ for timeouts~\cite{FroleyksHIJS21}.
 
@@
 Definition~\ref{def:new-solver-problem} specifies the problem we address.
+Note that our scenario assumes knowing the runtimes of all solvers, except the new one, on all instances.
+One could also imagine a collaborative filtering scenario, where runtimes are only partially known~\cite{misir2017data,misir2017alors}.
 
 \begin{definition}[New-Solver Problem]
-  Given solvers~$\mathcal{A}$, instances~$\mathcal{I}$, runtimes~$r\!: \mathcal{A} \times \mathcal{I} \rightarrow \left[0, \tau\right]$ with time-limit $\tau$, and a new solver $\hat{a} \notin \mathcal{A}$, incrementally select benchmark instances from $\mathcal{I}$ to maximize the confidence in predicting the rank of $\hat{a}$ while minimizing the total benchmark runtime.
+  Let us consider a set of solvers~$\mathcal{A}$, instances~$\mathcal{I}$, and runtimes~${r\!: \mathcal{A} \times \mathcal{I} \rightarrow \left[0, \tau\right]}$ with each solver having a runtime limit of $\tau$.
+  For a new solver $\hat{a} \notin \mathcal{A}$, incrementally select benchmark instances from $\mathcal{I}$ to maximize the confidence in predicting the rank of $\hat{a}$ while minimizing the total benchmark runtime.
   \label{def:new-solver-problem}
@@
 
-Note that our scenario assumes knowing the runtimes of all solvers, except the new one, on all instances.
-One could also imagine a collaborative filtering scenario, where runtimes are only partially known~\cite{misir2017data,misir2017alors}.
-
-Our approach satisfies several desirable criteria for benchmarking:
-Rather than outputting a binary classification, i.e., whether the new solver is worse than an existing solver or not, we provide a \emph{scoring} function that shows by which margin a solver is worse and how similar it is to existing solvers.
-In particular, our approach enables \emph{ranking} the new solver amidst a set of existing solvers.
-For this ranking, we do not even need to predict exact solver runtimes, which is trickier.
-Further, we optimize the \emph{runtime} that our strategy needs to arrive at its conclusion.
-We use instance and runtime \emph{features}.
+The approach presented in this paper meets several criteria for benchmarking that are considered desirable.
+Rather than outputting a binary classification, namely whether the new solver is worse than an existing solver or not, we provide a scoring function that shows by which margin a solver is worse and how similar it is to existing solvers.
+In particular, our approach enables ranking the new solver amidst a set of existing solvers.
+For this ranking, it is not necessary to predict the exact runtime of the solvers, which would be a more challenging task.
+This is demonstrated in Section~\ref{sec:exdesign:disc-pred}.
+%
+Furthermore, we optimize the \emph{runtime} required for our strategy to arrive at its conclusion.
+To this end, we use instance \emph{features} and known solver \emph{runtimes}.
 Moreover, we select instances \emph{non-randomly} and \emph{incrementally}.
@@
 
-We evaluate our approach with the SAT Competition~2022 Anniversary Track dataset~\cite{sat2022}, consisting of 5355~instances and runtimes of 28~solvers.
-We perform cross-validation by treating each solver once as the new solver and learning to predict the PAR-2 rank of that solver.
-On average, our predictions reach about \SI{92}{\%} accuracy with only about \SI{10}{\%} of the runtime required to evaluate these solvers on the complete set of instances.
+We evaluated our approach using the SAT Competition~2022 Anniversary Track dataset~\cite{sat2022}, consisting of 5355~instances and runtimes of 28~solvers.
+Cross-validation was performed by treating each solver once as the new solver and learning to predict the PAR-2 rank of that solver.
+On average, our predictions achieve an accuracy of approximately \SI{92}{\%} with only approximately \SI{10}{\%} of the runtime required to evaluate these solvers on the complete set of instances.
 Our entire source code\footnote{\url{https://github.com/mathefuchs/al-for-sat-solver-benchmarking}} and experimental data\footnote{\url{https://github.com/mathefuchs/al-for-sat-solver-benchmarking-data}} are available on GitHub.
 
-This article is an extended version of a conference paper with the same name~\cite{fuchs2023active}.
-In particular, we significantly extended the analysis of runtime-prediction approaches (cf.~Section~\ref{sec:exdesign:disc-pred}) and added an evaluation of how often problem instances were selected in benchmarks (cf.~Section~\ref{sec:eval:inst-freq}).
+This article is an extended and refined version of our conference paper with the same title, ``Active Learning for SAT Solver Benchmarking''~\cite{fuchs2023active}.
+In particular, we significantly extended the analysis of runtime-prediction approaches, which led to our choice of a solver runtime model for use in benchmark instance selection (cf.~Section~\ref{sec:exdesign:disc-pred}).
+Furthermore, we expanded the evaluation of which benchmark instances are selected by our approach (cf.~Section~\ref{sec:eval:instance}).
 % TODO: before submission, we should check which other changes we made and list them here accordingly
@@
 One of the most crucial parts of our framework is the choice of the surrogate model to predict runtime labels and inform active-learning selection decisions.
-In this section, we detail why we do \emph{not} directly predict absolute solver runtimes, which machine-learning models work best in predicting the time-outs of a solver, how clustering solver runtimes can lead to more nuanced runtime labels, and how ML models perform on those cluster labels.
+In this section, we detail why we do \emph{not} directly predict absolute solver runtimes, which machine-learning models work best in predicting the timeouts of a solver, how clustering solver runtimes can lead to more nuanced runtime labels, and how ML models perform on those cluster labels.
 
@@
 \subsection{Instance Frequency and Instance-Family Importance}
+\label{sec:eval:instance}
 
@@
 
-% MI: instance frequency (not so much to say, TODO: maybe add some more text)
 We analyze the frequency of individual instances being selected by our active-learning approach.

### 2024-05-24 paper_jar/paper.tex
section 2

@@
 
-Benchmarking is not only of high interest in many fields but also an active research area on its own.
-Recent studies show that benchmark selection is challenging for multiple reasons.
-Biased benchmarks can easily lead to fallacious interpretations~\cite{abs-2107-07002}.
-Benchmarking also has many interchangeable parts, such as the performance measures used, how measurement points are aggregated, and how missing values are handled.
-Questionable research practices could alter these elements a-posteriori to meet expectations, thereby skewing the results~\cite{NiesslHWCB22}.
-In the following, we discuss related work from the areas of static benchmark selection, algorithm configuration, incremental benchmark selection, and active learning.
+Benchmarking is of great interest in numerous research areas and represents an active field of research in its own right.
+Studies have shown that the compilation of benchmark instances poses a number of challenges.
+The use of biased benchmarks can easily lead to fallacious interpretations~\cite{abs-2107-07002}.
+Benchmarking also involves a number of interchangeable elements, including the performance measures used, the way the measures are aggregated, and the way missing values are handled.
+It is possible that questionable research practices could alter these elements a-posteriori in order to meet expectations, thereby skewing the results~\cite{NiesslHWCB22}.
+
+The following sections present a discussion of related work from the areas of static benchmark selection, algorithm configuration, incremental benchmark selection, and active learning.
 Table~\ref{tab:requirements} compares the most relevant approaches, which all pursue slightly different goals.
@@
 
-Benchmark selection is essential for competitions, e.g., the SAT Competition.
+Benchmark selection is essential for competitions such as the SAT Competition.
 In such competitions, the organizers define the rules for composing the benchmarks.
-These selection strategies are primarily static, i.e., they do not depend on particular solvers to distinguish.
+These selection strategies are primarily static, in that they do not depend on particular solvers to distinguish.
 Balint et al. provide an overview of benchmark-selection criteria in different solver competitions~\cite{balint2015overview}.
 Froleyks et al. describe benchmark selection in recent SAT~competitions~\cite{FroleyksHIJS21}.
-Manthey and Möhle find that competition benchmarks might contain redundant instances and propose a feature-based approach to remove redundancy~\cite{manthey2016better}.
-M{\i}s{\i}r presents a feature-based approach to reduce benchmarks by matrix factorization and clustering~\cite{misir2021benchmark}.
-
-Hoos et al.~\cite{HoosKSS13} discuss which properties are most desirable when selecting SAT benchmark instances.
-The selection criteria are instance variety to avoid over-fitting, adapted instance hardness (not too easy but also not too hard), and avoiding duplicate instances.
-To filter too similar instances, they use a distance-based approach with the SATzilla features~\cite{XuHHL08,features}.
-The approach does, however, not optimize for benchmark \emph{runtime} and selects instances \emph{randomly}, apart from constraints on the instance hardness and feature distance.
+Manthey and Möhle propose a feature-based approach to remove redundancy from competition benchmarks~\cite{manthey2016better}, while M{\i}s{\i}r presents a similar feature-based approach to reduce benchmarks by matrix factorization and clustering~\cite{misir2021benchmark}.
+
+In their study, Hoos et al.~\cite{HoosKSS13} discuss the desirable properties of SAT benchmark instances.
+They identify three key selection criteria: instance variety to avoid over-fitting, adapted instance hardness (not too easy but also not too hard), and the avoidance of duplicate instances. 
+To filter instances that are too similar, they employ a distance-based approach with the SATzilla features~\cite{XuHHL08,features}.
+It should be noted, however, that the approach does not optimize for benchmark runtime.
+Instead, instances are selected at random, with the exception of constraints on instance hardness and feature distance.
 
@@
 
-Further related work can be found within the field of algorithm configuration~\cite{HoosHL21,Stutzle0P22}, e.g., the configuration system SMAC~\cite{HutterHL11}.
+Further related work can be found within the field of algorithm configuration~\cite{HoosHL21,Stutzle0P22}, as exemplified by the configuration system SMAC~\cite{HutterHL11}.
 Thereby, the goal is to tune SAT solvers for a given sub-domain of problem instances.
-Although this task is different from our goal, e.g., we do not need to navigate the configuration space, there are similarities to our approach as well.
-For example, SMAC also employs an iterative, model-based selection procedure, though for configurations rather than instances.
-An algorithm configurator, however, cannot be used to \emph{rank/score} a new solver since algorithm configuration solemnly seeks to find the best-performing configuration.
-Also, while using a model-based selection strategy to sample configurations, instance selection is made \emph{randomly}, i.e., without building a model over instances.
+While this task differs from our goal, for instance, in that we do not need to navigate the configuration space, there are similarities to our approach as well.
+For example, SMAC employs an iterative, model-based selection procedure similar to our approach, but this is for selecting configurations rather than instances.
+However, an algorithm configurator cannot be used to \emph{rank} or \emph{score} a new solver since algorithm configuration is designed to identify the optimal configuration.
+Furthermore, while a model-based selection strategy is employed to sample configurations, instance selection is made \emph{randomly}, i.e., without building a model over instances.
 
@@
 Matricon~et~al. present an incremental benchmark selection approach~\cite{MatriconAFSH21}.
-Their \emph{per-set efficient algorithm selection problem} (PSEAS) is similar to our \emph{New-Solver Problem} (cf.~Definition~\ref{def:new-solver-problem}).
-Given a pair of SAT solvers, they iteratively select a subset of instances until the desired confidence level is reached to decide which of the two solvers is better.
-The selection of instances depends on the choice of the solvers to distinguish.
-They calculate a scoring metric for all unselected instances, run the experiment with the highest score, and update the confidence.
-Their approach ticks off most of our desired features in Table~\ref{tab:requirements}.
+Their \emph{per-set efficient algorithm selection problem} (PSEAS) is similar to our \emph{New-Solver Problem}, as given in Definition~\ref{def:new-solver-problem}.
+To differentiate between \emph{a pair of} SAT solvers, the authors propose an iterative approach to select a solver-specific subset of instances until a desired confidence level is reached.
+This is achieved by calculating a scoring metric for all unselected instances, running the experiment with the highest score, and updating the confidence.
+Their approach ticks off most of the desired features in Table~\ref{tab:requirements}.
 However, the approach only compares solvers binarily rather than providing a \emph{scoring}.
-Thus, it is unclear how similar two given solvers are or on which instances they behave similarly.
-Moreover, a significant shortcoming is the lacking \emph{scalability} with the number of solvers.
+Consequently, it is unclear how similar two given solvers are or on which instances they behave similarly.
+Furthermore, a significant shortcoming is the lack of \emph{scalability} with the number of solvers.
 Comparing only pairs of solvers, evaluating a new solver requires sampling a separate benchmark for each existing solver.
 In contrast, our approach allows comparing a new solver against a set of existing solvers by sampling only one benchmark.
+In contrast to our approach, which allows for the comparison of a new solver against a set of existing solvers by sampling only one benchmark, their pairwise method requires the evaluation of a new solver to be conducted on a separate benchmark for each existing solver.
 
@@
 
-Prediction models in passive machine learning are trained on datasets with given instance labels (cf.~Fig.~\ref{fig:passive}).
-In contrast, active learning (AL) starts with no or little labeled data.
-It repeatedly selects interesting problem instances for which to acquire labels, aiming to gradually improve the prediction model (cf.~Fig.~\ref{fig:activepool}).
-AL methods are especially beneficial if acquiring labels is computationally expensive, like obtaining solver runtimes.
-Without AL methods, it is not obvious which instances to label and which not.
-On the one hand, we want to maximize the utility an instance provides to our model, i.e., rank prediction accuracy, and on the other hand, minimize the cost, i.e., predicted runtime, associated with the instance's acquisition.
-Thus, we strive for an accurate prediction model without having to label every data point.
+In passive machine learning, prediction models are trained on datasets with given instance labels (cf.~Fig.~\ref{fig:passive}).
+In contrast, active learning (AL) begins with no or little labeled data.
+AL then repeatedly selects interesting problem instances for which to acquire labels, gradually improving the prediction model (cf.~Fig.~\ref{fig:activepool}).
+AL methods are particularly advantageous when acquiring labels is computationally expensive, such as obtaining solver runtimes.
+Without AL methods, it is not evident which instances to label and which not.
+The objective is to maximize the utility of instances to our model, which is measured by the rank prediction accuracy.
+Concurrently, the objective is to minimize the cost of label acquisition, which is measured by the instance's predicted runtime.
+Consequently, the objective is to develop an accurate prediction model without the necessity of labeling each data point.
+
+Rubens et.~al.~\cite{RubensESK15} survey active-learning advances.
+While synthesis-based AL methods~\cite{0001AEMN22,GarzonMG22,2019gaal} generate instances for labeling, pool-based methods~\cite{GolbandiKL11,HarpaleY08,KorenBV09} rely on a fixed set of unlabeled instances to sample from.
+Recent synthesis-based methods within the field of SAT solving demonstrate the generation of problem instances with desired properties~\cite{0001AEMN22,GarzonMG22}.
+However, this objective is distinct from ours.
+While those approaches aim to generate instances on which a solver is either effective or ineffective, our objective is to predict whether a solver is effective or ineffective on an \emph{existing} benchmark.
+In their work, Volpato and Guangyan employ pool-based AL to train an instance-specific algorithm selector~\cite{volpato2019active}.
+Rather than benchmarking a solver's overall performance, their objective is to recommend the best solver for each SAT instance from a set of available solvers.
 
@@
 \end{figure}
-
-Rubens et.~al.~\cite{RubensESK15} survey active-learning advances.
-While synthesis-based AL methods~\cite{0001AEMN22,GarzonMG22,2019gaal} generate instances for labeling, pool-based methods~\cite{GolbandiKL11,HarpaleY08,KorenBV09} rely on a fixed set of unlabeled instances to sample from.
-Recent synthesis-based methods within the field of SAT solving show how to generate problem instances with desired properties~\cite{0001AEMN22,GarzonMG22}.
-This goal is, however, orthogonal to ours.
-While those approaches want to generate instances on which a solver is good or bad, we want to predict whether a solver is good or bad on an existing benchmark.
-Volpato and Guangyan use pool-based AL to learn an instance-specific algorithm selector~\cite{volpato2019active}.
-Rather than benchmarking a solver's overall performance, their goal is to recommend the best solver out of a set of solvers for each SAT instance.
 

### 2024-05-24 paper_jar/paper.tex
section 3

@@
   \begin{tabular}[c]{ccc}
-  \begin{subfigure}[b]{0.236\textwidth}
+  \begin{subfigure}[b]{0.24\textwidth}
   \centering
@@
 Given a set of solvers~$\mathcal{A}$, instances~$\mathcal{I}$ and runtimes~$r$, we first initialize a prediction model~$\mathcal{M}$ for the new solver $\hat a \not\in \mathcal{A}$ (Line~1).
-The prediction model~$\mathcal{M}$ is used to repeatedly select an instance (Line~4) for benchmarking $\hat a$ (Line~5). 
-The acquired result is subsequently used to update the prediction model~$\mathcal{M}$ (Line~7). 
-When the stopping criterion is met (Line~3), we quit the benchmarking loop and predict the final score of $\hat a$ (Line~8). 
-Algorithm~\ref{algALBenchmark} returns the predicted score of $\hat a$ as well as the acquired instances and runtime measurements (Line~9).
-
-Section~\ref{sec:main:model} describes the underlying prediction model~$\mathcal{M}$ and specifies how we may derive a solver ranking from it.
-We discuss criteria for selecting instances in Section~\ref{sec:main:selection}.
-Section~\ref{sec:main:stopping} concludes with possible stopping conditions.
+The prediction model~$\mathcal{M}$ is used to select instances repeatedly (Line~4) for benchmarking the new solver $\hat a$ (Line~5). 
+The acquired result is subsequently employed to update the prediction model, denoted by~$\mathcal{M}$ (Line~7). 
+When the stopping criterion is met (Line~3), the benchmarking loop is terminated and the final score of $\hat{a}$ is estimated (Line~8). 
+This score, along with the acquired instances and runtime measurements, is returned by Algorithm~\ref{algALBenchmark}.
+
+Section~\ref{sec:main:model} provides an overview of the underlying prediction model~$\mathcal{M}$ and outlines the methodology for deriving a solver ranking from it.
+Section~\ref{sec:main:selection} presents the criteria for selecting instances.
+Finally, Section~\ref{sec:main:stopping} presents potential stopping conditions.
 
@@
 This prediction function powers instance selection as described in Section~\ref{sec:main:selection}.
-During model updates (Algorithm~\ref{algALBenchmark},~Line~7), $f$ is trained to predict a transformed version of the acquired runtimes~$\mathcal{R}$.
-We describe the runtime transformation in the subsequent section.
+During model updates (Algorithm~\ref{algALBenchmark},~Line~7), the prediction function $f$ is trained to predict a transformed version of the acquired runtime data set~$\mathcal{R}$.
+The transformation of the runtime data set is described in the subsequent section.
 The features described in Section~\ref{sec:exdesign:data} serve as the input to the model.
-Further, note that we build a new prediction model in each iteration since running experiments (Line~5) dominates the runtime of model training by magnitudes.
-Finally, we predict the score of the new solver $\hat a$ with the prediction function~$f$ (Line~8).
+Furthermore, it should be noted that a new prediction model is constructed in each iteration, as the runtime of the experiments (Line~5) dominates the training time of the model by a significant margin.
+Finally, the score of the new solver, denoted by $\hat a$, is predicted with the prediction function~$f$ (Line~8).
 
@@
 %
-Given an instance $e \in \mathcal{I}$, a solver $a \in \mathcal{A}$ belongs to the $\gamma_k(a, e)$-fastest solvers on instance $e$. 
-In preliminary experiments, we achieved higher accuracy for predicting such discrete runtime labels than for predicting raw runtimes.
-Research on portfolio solvers has also shown that discretization works well in practice~\cite{CollauttiMMO13,NgokoCT19}.
+In the context of such a clustering, for an instance $e \in \mathcal{I}$, a solver $a \in \mathcal{A}$ is defined as belonging to the $\gamma_k(a, e)$-fastest solvers on instance $e$. 
+%
+Empirical studies on portfolio solvers have demonstrated that discretization is an effective approach in practice~\cite{CollauttiMMO13,NgokoCT19}.
+Section~\ref{sec:exdesign:disc-pred} presents a detailed analysis of the runtime transformation we designed for our approach.
+The analysis demonstrates that our approach achieves higher accuracy for discrete runtime labels than for raw runtimes.
 
@@
 
-To determine solver ranks, we use the transformed runtimes $\gamma_k(a, e)$ in the adapted scoring function $s_k : \mathcal{A} \rightarrow [1, 2 \cdot k]$ as follows:
+To determine the rank of solvers, we apply PAR-2 scoring using the discrete runtime labels $\gamma_k(a, e)$.
+The adapted scoring function $s_k : \mathcal{A} \rightarrow [1, 2 \cdot k]$ induces a ranking among solvers and is depicted in Equation~\eqref{eq:rankingeq}.
 %
@@
 %
-I.e., we apply PAR-2 scoring, which is commonly used in SAT competitions~\cite{FroleyksHIJS21}, on the discrete labels.
-The scoring function $s_k$ induces a ranking among solvers.
-
 \subsection{Instance Selection}
@@
 
-Selecting an instance based on the model is a core functionality of our framework (cf. Algorithm~\ref{algALBenchmark}, Line~4).
+Selecting an instance based on the model is a core functionality of our framework (cf.~Algorithm~\ref{algALBenchmark}, Line~4).
 In this section, we introduce two instance sampling strategies, one that minimizes uncertainty and one that maximizes information gain.
@@
 
-\subsubsection{Uncertainty Sampling}
-
-The uncertainty sampling strategy selects the instance closest to the model's decision boundary, i.e., 
-we select the instance $e \in \mathcal{I} \setminus \tilde{\mathcal{I}}$ that minimizes $U(e)$, which is specified as follows:
+\paragraph{Uncertainty Sampling}
+The uncertainty sampling strategy selects the instance that is closest to the model's decision boundary.
+This is achieved by selecting the instance $e \in \mathcal{I} \setminus \tilde{\mathcal{I}}$ that minimizes $U(e)$, which is specified by the following equation.
 %
@@
 
-\subsubsection{Information-Gain Sampling}
-
+\paragraph{Information-Gain Sampling}
 The information-gain sampling strategy selects the instance with the highest expected entropy reduction regarding the runtime labels of the instance.
-To be more specific, we select the instance $e \in \mathcal{I} \setminus \tilde{\mathcal{I}}$ that maximizes $IG(e)$, which is specified as follows:
+To be more precise, we select the instance $e \in \mathcal{I} \setminus \tilde{\mathcal{I}}$ that maximizes $IG(e)$, which is specified in the following equation.
 %
@@
 %
-Here, $\operatorname{H}(e)$ denotes the entropy of the runtime labels $\gamma(a, e)$ over all $a \in \mathcal{A}$ and $\operatorname{H}(e, n)$ denotes the entropy of these labels plus $n$ as the runtime label for $\hat{a}$.
+In the equation, $\operatorname{H}(e)$ denotes the entropy of the runtime labels $\gamma(a, e)$ over all $a \in \mathcal{A}$ and $\operatorname{H}(e, n)$ denotes the entropy of these labels plus $n$ as the runtime label for $\hat{a}$.
 The term $\operatorname{\hat H}_n(e)$ is computed for every possible runtime label $n \in \{1, \dots, k\}$.
-By maximizing information gain, we select instances that identify solvers with similar behavior.
+By maximizing the information gain, instances are selected that identify solvers with similar behavior.
 
@@
 
-In this section, we present the two dynamic stopping criteria in our experiments, the Wilcoxon and the ranking stopping criterion (cf. Algorithm~\ref{algALBenchmark}, Line~3).
-
-\subsubsection{Wilcoxon Stopping Criterion}
-
-The Wilcoxon stopping criterion stops the active-learning process when we are confident enough that the predicted runtime labels of the new solver are sufficiently different from existing solvers.
+In this section, we present the two dynamic stopping criteria employed in our experiments: the Wilcoxon and the ranking stopping criterion (cf. Algorithm~\ref{algALBenchmark}, Line~3).
+
+\paragraph{Wilcoxon Stopping Criterion}
+The Wilcoxon stopping criterion is employed to halt the active-learning process based on the confidence that the predicted runtime labels of the new solver are sufficiently different from those of existing solvers.
 This criterion is loosely inspired by Matricon et.~al.~\cite{MatriconAFSH21}.
-We use the average $p$-value $W_{\hat{a}}$ of a Wilcoxon signed-rank test $\operatorname{w}(S,P)$ of the two runtime label distributions $S=\{ \gamma(a, e) \mid e \in \mathcal{I} \}$ for an existing solver $a$ and \mbox{$P=\{ f(\hat a, e) \mid e \in \mathcal{I} \}$} for the new solver $\hat{a}$:
+The average $p$-value $W_{\hat{a}}$ of a Wilcoxon signed-rank test $\operatorname{w}(S,P)$, as given by the following equation, of the two runtime label distributions ${S=\{ \gamma(a, e) \mid e \in \mathcal{I} \}}$ for an existing solver $a$ and \mbox{$P=\{ f(\hat a, e) \mid e \in \mathcal{I} \}$} of the new solver $\hat{a}$ is used to assess the statistical significance of the predicted runtime labels.
 %
@@
 %
-To improve the stability of this criterion, we use an exponential moving average to smooth out outliers and stop as soon as $W^{(i)}_{\exp}$ drops below a fixed threshold:
+In order to improve the stability of this criterion, an exponential moving average is employed to mitigate the impact of outliers, with the process terminated when the value of $W^{(i)}_{\exp}$ drops below a fixed threshold.
 %
@@
 
-\subsubsection{Ranking Stopping Criterion}
-
-The ranking stopping criterion is less sophisticated in comparison.
-It stops the active-learning process if the ranking induced by the model's predictions (Equation~\ref{eq:rankingeq}) remained unchanged within the last~$l$~iterations.
-However, the concrete values of the predicted score $s_{\hat a}$ might still change.
-We are solemnly interested in the induced ranking in this case.
+\paragraph{Ranking Stopping Criterion}
+In contrast, the stopping criterion for the ranking process is less sophisticated than the previous one.
+In this case, the active learning process is terminated if the ranking induced by the model's predictions (Equation~\ref{eq:rankingeq}) remains unchanged within the last~$l$~iterations.
+However, it should be noted that the concrete values of the predicted score $s_{\hat a}$ may still change.
+Nevertheless, in this case, the induced ranking is of sole interest.
 

### 2024-05-24 paper_jar/paper.tex
section 4

@@
 
-Given all the previously presented instantiations for Algorithm~\ref{algALBenchmark}, this section outlines our experimental design, including our evaluation framework, used data sets, hyper-parameter choices, and implementation details.
+In light of the preceding instantiations of Algorithm~\ref{algALBenchmark}, this section outlines the experimental design employed in the presented study.
+This includes an overview of the evaluation framework, the data sets employed, the hyper-parameter selections, and the implementation details.
 
@@
   $O \leftarrow \emptyset$
-  \BlankLine
+  % \BlankLine
 
@@
 
-  \BlankLine
+  % \BlankLine
     $O \leftarrow O \cup \bigl\{ ( O_{\operatorname{acc}},\, O_{\operatorname{rt}} ) \bigr\}$
@@
 
-  \BlankLine
+  % \BlankLine
   $\bigl( \bar{O}_{\operatorname{acc}}, \bar{O}_{\operatorname{rt}} \bigr) \leftarrow \operatorname{average}(O)$ \;
 
-  \BlankLine
+  % \BlankLine
   \Return $\bigl( \bar{O}_{\operatorname{acc}}, \bar{O}_{\operatorname{rt}} \bigr)$
@@
 As stated in the Introduction, this work addresses the \emph{New-Solver Problem} (cf.~Definition~\ref{def:new-solver-problem}).
-As described in Section~\ref{sec:main:model}, a prediction model $\mathcal{M}$ provides us with an estimated scoring~$s_{\hat a}$ for the new solver~$\hat{a}$.
-
-To evaluate a concrete instantiation of Algorithm~\ref{algALBenchmark}, i.e., a concrete choice for all the sub-routines, we perform cross-validation on our set of solvers.
-Algorithm~\ref{alg:eval} shows this.
-That means each solver plays the role of the new solver~$\hat{a}$ once (Line~2).
-Note that the \emph{new} solver in each iteration is excluded from the set of solvers $\mathcal{A}$ to avoid data leakage (Line~3).
+As described in Section~\ref{sec:main:model}, a prediction model provides us with an estimated scoring value for the new solver~$\hat{a}$.
+%
+To evaluate the efficacy of a concrete instantiation of Algorithm~\ref{algALBenchmark}, which entails a specific choice for all subroutines, we perform cross-validation on our set of solvers.
+
+This is demonstrated in Algorithm~\ref{alg:eval}.
+Each solver assumes the role of the new solver~$\hat{a}$ once (Line~2) and is consequently excluded from the set of solvers $\mathcal{A}$ (Line~3).
 After running our active-learning framework for solver~$\hat{a}$ (Line~4), we compute the value of both our optimization goals, i.e., ranking accuracy and runtime.
-We define the \emph{ranking accuracy} $O_{\operatorname{acc}} \in \left[0, 1\right]$ (higher is better) by the fraction of pairs $\left(\hat{a}, a\right)$ for all $a \in \mathcal{A}$ that are decided correctly regarding the ground-truth scoring~$\operatorname{par_2}$ (Lines~5-8).
-The \emph{fraction of runtime} that the algorithm needs to arrive at its conclusion is denoted by $O_{\operatorname{rt}} \in \left[0, 1\right]$ (lower is better).
-This metric puts the runtime summed over the sampled instances in relation to the runtime summed over all instances in the dataset (Lines~9-13).
-Finally, we compute averages of the output metrics in Line~15 after we have collected all cross-validation results in Line~14.
-Overall, we want to find an approach that maximizes
+%
+The \emph{ranking accuracy}, expressed in the interval $[0,1]$ (higher is better), is defined as the fraction of correctly ranked pairs of elements, namely, $\left(\hat{a}, a\right)$, where $a$ belongs to the set of all solvers $\mathcal{A}$ ranked by the ground-truth scoring function $\operatorname{par_2}$ (Lines~5-8).
+%
+The \emph{fraction of runtime} required by the algorithm to reach its conclusion is represented by $O_{\operatorname{rt}}$, which lies within the interval $[0, 1]$ (lower is better).
+This metric compares the accumulated runtimes over the sampled instances to the accumulated runtimes over all instances in the dataset (Lines~9-13).
+Once all cross-validation results have been collected (Line~14), the output metrics are averaged (Line~15).
+
+The procedure is designed to identify an approach that maximizes Equation~\eqref{eq:opt}, in which $\delta \in \left[0, 1\right]$ is used for linear weighting, with weighting factor $\delta \in \left[0, 1\right]$, between the two optimization goals $O_{\operatorname{acc}}$ and $O_{\operatorname{rt}}$.
+The plotting of the approaches that maximize $O_\delta$ for all $\delta \in \left[0, 1\right]$ on an $O_{\operatorname{rt}}$-$O_{\operatorname{acc}}$-diagram provides us with a Pareto front of the best approaches for different optimization-goal weightings.
 %
@@
 %
-whereby $\delta \in \left[0, 1\right]$ allows for linear weighting between the two optimization goals $O_{\operatorname{acc}}$ and $O_{\operatorname{rt}}$.
-Plotting the approaches that maximize $O_\delta$ for all $\delta \in \left[0, 1\right]$ on an $O_{\operatorname{rt}}$-$O_{\operatorname{acc}}$-diagram provides us with a Pareto front of the best approaches for different optimization-goal weightings.
-
+%
 \subsection{Data}
@@
 
-In our experiments, we work with the dataset of the SAT~Competition~2022 Anniversary Track~\cite{sat2022}.
-The dataset consists of 5355 instances with respective runtime data of 28 sequential SAT solvers.
+In our experiments we work with the dataset of the SAT~Competition~2022 Anniversary Track~\cite{sat2022}.
+The dataset consists of 5355 instances with corresponding runtime data of 28 sequential SAT solvers.
 We also use a database of 56 instance features\footnote{\url{https://benchmark-database.de/getdatabase/base_db}} from the Global Benchmark Database~(GBD) by Iser~et~al.~\cite{IserS18}.
-They comprise instance size features and node distribution statistics for several graph representations of SAT instances, among others, and are primarily inspired by the SATzilla~2012 features described in~\cite{features}.
-All features are numeric and free of missing values.
-We drop 10 out of 56 features because of zero variance.
-Overall, prediction models have access to 46 instance features and 27 runtime features, i.e., excluding the current new solver~$\hat{a}$.
-
-Additionally, we retrieve instance-family information\footnote{\url{https://benchmark-database.de/getdatabase/meta_db}} to evaluate the composition of our sampled benchmarks.
-Instance families comprise instances from the same application domain, e.g., planning, cryptography, etc., and are a valuable tool for analyzing solver performance.
-
-For hyper-parameter tuning, we randomly sample \SI{10}{\%} of the complete set of 5355 instances with stratification regarding the instances' family.
-All instance families that are too \emph{small}, i.e., \SI{10}{\%} of them corresponds to less than one instance, are put into one meta-family for stratification.
-This \emph{tuning dataset} allows for a more extensive exploration of the hyper-parameter space.
-
-\subsection{Hyper-parameters}
+They include instance size features and node distribution statistics for several graph representations of SAT instances, and are primarily inspired by the SATzilla~2012 features described in~\cite{features}.
+All features are numeric and have no missing values.
+We drop 10 of the 56 features due to zero variance.
+In total, the prediction models have access to 46 instance features and 27 out of 28 runtime features, excluding the respective new solver~$\hat{a}$.
+
+In addition, we retrieve instance family information\footnote{\url{https://benchmark-database.de/getdatabase/meta_db}} to evaluate the composition of our sampled benchmarks.
+Instance families consist of instances from the same application domain, such as planning, cryptography, etc., and are a valuable tool for analyzing solver performance.
+%
+For hyperparameter tuning, we randomly sample \SI{10}{\%} of the complete set of 5355 instances with stratification regarding the instance family.
+All instance families that are too \emph{small}, i.e., \SI{10}{\%} of them correspond to less than one instance, are put into a meta-family for stratification.
+This \emph{tuning dataset} allows a more extensive exploration of the hyperparameter space.
+
+\subsection{Hyperparameters}
 \label{sec:exdesign:hyper}
 
-Given Algorithm~\ref{algALBenchmark}, there are several possible instantiations for the three sub-routines, i.e., \emph{ranking}, \emph{selection}, and \emph{stopping}.
-We describe these experimental configurations in the following.
+Given the Algorithm~\ref{algALBenchmark}, there are several possible instantiations for the three subroutines \emph{ranking}, \emph{selection}, and \emph{stopping}.
+These experimental configurations are described in the following.
 

### 2024-05-24 paper_jar/paper.tex
sections 4 and 5

@@
 In light of the preceding instantiations of Algorithm~\ref{algALBenchmark}, this section outlines the experimental design employed in the presented study.
-This includes an overview of the evaluation framework, the data sets employed, the hyper-parameter selections, and the implementation details.
+This includes an overview of the evaluation framework, the data sets employed, the hyperparameter selections, and the implementation details.
 
@@
 
-\subsubsection{Ranking}
-
-Regarding \emph{ranking} (cf. Section~\ref{sec:main:model}), we experiment with the following approaches and hyper-parameter values:
-
-\begin{itemize}
+\subparagraph{Ranking}
+Regarding \emph{ranking}, we experiment with the following instantiations.
+
+\begin{itemize}\setlength{\itemsep}{1pt}
   \item Observed PAR-2 ranking of already sampled instances
   \item Predicted runtime-label ranking
-  \begin{itemize}
+  \vspace*{-1ex}
+  \begin{itemize}\setlength{\itemsep}{1pt}
     \item
@@
 
-\subsubsection{Selection}
-
-For \emph{selection} (cf. Section~\ref{sec:main:selection}), we experiment with the following methods and hyper-parameter values.
-Since the potential runtime of experiments is by magnitudes larger than the model's update time, we only consider incrementing our benchmark by one instance at a time rather than using batches, which is also proposed in current active-learning advances~\cite{SinhaED19,2019gaal}.
+\subparagraph{Selection}
+We experiment with the following instantiations of instance \emph{selection}.
+We experiment with the following instantiations of the instance \emph{selection} criteria.
+Since the potential runtime of the experiments is orders of magnitude larger than the model update time, we consider incrementing our benchmark one instance at a time rather than using batches, which is also proposed in recent advances in active learning~\cite{SinhaED19,2019gaal}.
 A drawback of this is the lack of parallel execution of runtime experiments.
 
-\begin{itemize}
+\begin{itemize}\setlength{\itemsep}{1pt}
   \item Random sampling 
   \item Uncertainty sampling
-  \begin{itemize}
+  \vspace*{-1ex}
+  \begin{itemize}\setlength{\itemsep}{1pt}
     \item Fallback threshold: Use random sampling for the first \SI{0}{\%}, \SI{5}{\%}, \SI{10}{\%}, \SI{15}{\%}, or \SI{20}{\%} of instances to explore the instance space.
@@
   \item Information-gain sampling
+  \vspace*{-1ex}
   \begin{itemize}
@@
 
-\subsubsection{Stopping}
-
-For \emph{stopping} decisions (cf. Section~\ref{sec:main:stopping}), we experiment with the following criteria and hyper-parameter values:
-
-\begin{itemize}
+\subparagraph{Stopping}
+We evaluate the following \emph{stopping} criteria.
+
+\begin{itemize}\setlength{\itemsep}{1pt}
   \item Subset-size stopping criterion, using \SI{10}{\%} or \SI{20}{\%} of instances
   \item Ranking stopping criterion
-  \begin{itemize}
+  \vspace*{-1ex}
+  \begin{itemize}\setlength{\itemsep}{1pt}
     \item Minimum amount: Sample at least \SI{2}{\%}, \SI{8}{\%}, \SI{10}{\%}, or \SI{12}{\%} of instances before applying the criterion.
@@
   \end{itemize}
-
   \item Wilcoxon stopping criterion
-  \begin{itemize}
+  \vspace*{-1ex}
+  \begin{itemize}\setlength{\itemsep}{1pt}
     \item Minimum amount: Sample at least \SI{2}{\%}, \SI{8}{\%}, \SI{10}{\%}, or \SI{12}{\%} of instances before applying the criterion.
@@
 \label{sec:exdesign:disc-pred}
-One of the most crucial parts of our framework is the choice of the surrogate model to predict runtime labels and inform active-learning selection decisions.
-In this section, we detail why we do \emph{not} directly predict absolute solver runtimes, which machine-learning models work best in predicting the timeouts of a solver, how clustering solver runtimes can lead to more nuanced runtime labels, and how ML models perform on those cluster labels.
-
-% MI: table with regression performance:
-Table~\ref{tab:regression} shows the prediction performance of two regression models, a Random-Forest-Regressor and an MLP-Regressor, on the test set.
-Both models have been chosen because of their robustness and capability of capturing non-linear dependencies~\cite{breiman2001random}.
-The instance feature vector consists of 46 instance features and 27 known solver runtimes.
-Both models perform poorly, which is evident from the magnitudes of the average regression errors.
-An RMSE error of about \SI{2000}{s} is quite poor compared to the solver runtimes which range from 0 to \SI{5000}{s}.
-Therefore, we decided to discretize runtimes.
-
-\begin{table}[tbp]
+One of the most important parts of our framework is the choice of solver runtime model to inform active learning selection decisions.
+In this section, we detail how clustering solver runtimes can lead to nuanced runtime labels, and which machine learning models are best at predicting them.
+
+As a baseline for improvement, Table~\ref{tab:regression} shows the runtime prediction performance of two regression models, a random forest regressor and an MLP regressor, using 46 instance features and 27 known solver runtimes.
+Both models were chosen for their robustness and ability to capture nonlinear dependencies~\cite{breiman2001random}, and both perform poorly, as evidenced by the magnitudes of the mean regression errors.
+An RMSE error of about \SI{2000}{s} is quite poor compared to solver runtimes that range from 0 to \SI{5000}{s}.
+Therefore, we decided to discretize the runtimes.
+%
+\begin{table}[b]
   \centering
@@
 
-% MI: table with timeout prediction performance: (merged with existing text)
 Table~\ref{tab:timeout-prediction} shows the performance of different models in predicting timeouts.
 While the second column shows the performance of models trained on the 46 instance features, the third column shows the performance of models trained on the 46 instance features and the 27 runtime features.
-While the Quadratic-DA model performs poorly on the 46 instance features, it performs well on the 46 instance features and 27 runtime features.
-The best prediction performance however is achieved by a stacking ensemble~\cite{wolpert1992stacked}, which combines quadratic-discriminant analysis~\cite{tharwat2016linear} and a random-forest classifier~\cite{breiman2001random}.
-Stacking means that another prediction model, in our case a simple decision tree, decides which of the two ensemble members makes the prediction on which instance.
+While the Quadratic DA model performs poorly on the 46 instance features, it performs well on the 46 instance features and the 27 runtime features.
+However, the best prediction performance is achieved by a stacking ensemble~\cite{wolpert1992stacked}, which combines quadratic discriminant analysis~\cite{tharwat2016linear} and a random forest classifier~\cite{breiman2001random}.
+Stacking means that another prediction model, in our case a simple decision tree, decides which of the two ensemble members makes the prediction about which instance.
 
@@
 % MI: table with cluster prediction performance: (merged with existing text)
-As already mentioned, runtime prediction is quite challenging.
-Instead, we discretize runtimes.
-To define prediction targets, we use hierarchical clustering with $k$ clusters and a log-single-link criterion, which produced the most \emph{useful} labels in preliminary experiments.
-We denote this adapted solver scoring function with $s_k$ as defined in~\eqref{eq:rankingeq}.
+Since runtime prediction is quite challenging, we discretize the runtimes.
+To define prediction targets for each instance, we cluster the solver runtimes into $k$ clusters.
+We denote this adapted solver scoring function which assigns a solver to its cluster for each instance with $s_k$ as defined in Equation~\eqref{eq:rankingeq}.
+%
+In preliminary experiments, the most \emph{useful} labels were produced with hierarchical clustering and a log-single-link criterion.
 In our chosen hierarchical procedure, each non-timeout runtime starts in a separate interval.
-We then gradually merge intervals whose single-link logarithmic distance is the smallest until the desired number of partitions is reached.
-The clustering is instance-specific, i.e., the boundaries are chosen such that it produces groups of solvers with similar performances for each instance.
-Other clustering approaches that we tried include hierarchical clustering with mean-, median-, and complete-link criterion, as well as $k$-means and spectral clustering.
-
-% MI: table with cluster prediction performance:
-We evaluated the performance of the best models from Table~\ref{tab:timeout-prediction} in predicting these runtime clusters as mentioned above.
-Table~\ref{tab:cluster-prediction} shows their performance in predicting the $C_k$ clusterings for $k \in \lbrace 3, 4, 5 \rbrace$.
-While the second column shows the prediction performance of the models trained on the 46 instance features including the 27 known solver runtimes, the third column shows the prediction performance of the models trained on those features plus 27 features representing the known $C_k$ clustering labels of the known solver runtimes.
+We then gradually merge intervals with the smallest single-link logarithmic distance until the desired number of partitions is reached.
+The clustering is instance-specific, i.e., the bounds are chosen to produce groups of solvers with similar performance for each instance.
+Other clustering approaches we have tried include hierarchical clustering with mean, median, and full-link criteria, as well as $k$-means and spectral clustering.
+
+We evaluated the performance of the best models from Table~\ref{tab:timeout-prediction} in predicting the such determined $s_k$.
+Table~\ref{tab:cluster-prediction} shows their performance in predicting the $C_k$ clusters for $k \in \lbrace 3, 4, 5 \rbrace$.
+While the second column shows the prediction performance of the models trained on the 46 instance features including the 27 known solver runtimes, the third column shows the prediction performance of the models trained on these features plus 27 features representing the known $C_k$ clustering labels of the known solver runtimes.
 We observe that the models perform better when the clustering labels are included as features.
-Moreover, the prediction performance degrades rapidly with an increasing number of clusters.
+In addition, the prediction performance degrades rapidly as the number of clusters increases.
 In all cases, the stacking ensemble performs best.
@@
 
-% MI: table with runtime prediction performance: (merged with existing text)
-Our final choice for the prediction model is the stacking ensemble of a quadratic-discriminant analysis and a random forest.
-Both these models can learn non-linear relationships between the instance features and the runtime labels.
+Our final choice for the prediction model is the stacking ensemble of a quadratic discriminant analysis and a random forest.
+Both of these models can learn non-linear relationships between the instance features and the runtime labels.
 Figure~\ref{fig:confusion-matrices} shows the confusion matrices of the stacking ensemble for predicting timeouts and $C_3$ clusters.
@@
 Notably, for the $C_3$ prediction model, there is less confusion between the non-timeout labels (labels 1 and 2) and the timeout label (label 3) than between the non-timeout labels (labels 1 and 2).
-This indicates that deciding timeouts is easier than distinguishing different groups of solvers.
+This suggests that deciding on timeouts is easier than distinguishing between different groups of solvers.
 
@@
 
-% MI: table with rankings after discretization: (merged with existing text)
-To obtain \emph{useful} labels, we also need to ensure that the discretized labels still discriminate solvers and align with the actual PAR-2 ranking.
+To ensure that we obtained \emph{useful} labels, we evaluated whether the discretized labels still discriminate solvers and agree with the actual PAR-2 ranking.
 Table~\ref{tab:ranking} shows the ranking of the solvers in the SAT Competition~2022 Anniversary Track~\cite{sat2022} according to the standard PAR-2 runtime score and according to our discretized $s_3$ score.
-We observe that our ranking approach correctly decides almost all (about \SI{97.45}{\%}; $\sigma = \SI{3.68}{\%}$) solver pairs, i.e., which solver is faster among two given solvers.
+We observe that our ranking approach correctly decides almost all (about \SI{97.45}{\%}; $\sigma = \SI{3.68}{\%}$) solver pairs, i.e. which solver is faster among two given solvers.
 In particular, the Spearman correlation of $s_3$ and PAR-2 ranking is about \SI{0.988}{}, which is very close to the optimal value of~1~\cite{de2016comparing}.
@@
 
-% MI: table with solver confusion matrices:
-We further analyzed how runtime discretization affects the distinguishability of solver pairs.
+We also analyzed how runtime discretization affects the distinguishability of solver pairs.
 Figure~\ref{fig:solver-confusion-matrices} shows the confusion matrices of solver pairs with significant differences in their PAR-2 scores, timeouts, and $s_3$ scores.
-According to a Wilcoxon signed-rank test with level $\alpha = 0.05$, \SI{87.83}{\%} of solver pairs have significantly different scores after discretization, only a slight drop compared to \SI{89.95}{\%} before discretization.
-These results underpin the usefulness of discretized runtimes for our framework.
+According to a Wilcoxon signed-rank test with a level of $\alpha = 0.05$, \SI{87.83}{\%} of the solver pairs have significantly different scores after discretization, only a small decrease compared to \SI{89.95}{\%} before discretization.
+These results support the usefulness of discretized runtimes for our framework.
 
@@
 
-For reproducibility, our source code and data are available on GitHub (cf. footnotes in Section~\ref{sec:intro}).
-Our code is implemented in \textsc{Python} using \emph{scikit-learn}~\cite{scikit-learn} for making predictions and \emph{gbd-tools}~\cite{IserS18} for SAT-instance retrieval.
+For reproducibility, our source code and data are available on GitHub (see footnotes in Section~\ref{sec:intro}).
+Our code is implemented in \textsc{Python} using \emph{scikit-learn}~\cite{scikit-learn} for making predictions and \emph{gbd-tools}~\cite{IserS18} for retrieving SAT instances.
 
@@
 
-\subsection{Hyper-Parameter Analysis}
+\subsection{Hyperparameter Analysis}
 
@@
   \caption{
-    $O_{\operatorname{rt}}$-$O_{\operatorname{acc}}$-diagrams comparing different hyper-parameter instantiations of our active-learning framework on the hyper-parameter-tuning dataset.
+    $O_{\operatorname{rt}}$-$O_{\operatorname{acc}}$-diagrams comparing different hyperparameter instantiations of our active-learning framework on the hyperparameter-tuning dataset.
     The x-axis shows the ratio of total solver runtime on the sampled instances relative to all instances.
     The y-axis shows the ranking accuracy (cf. Section~\ref{sec:exdesign:eval}).
-    Each line entails the front of Pareto-optimal configurations for the respective hyper-parameter instantiation.
+    Each line entails the front of Pareto-optimal configurations for the respective hyperparameter instantiation.
   }
@@
 Our experiments follow the evaluation framework introduced in Section~\ref{sec:exdesign:eval}.
-Fig.~\ref{fig:e2eallsolvers} shows the performance of the approaches from Section~\ref{sec:exdesign:hyper} on $O_{\operatorname{rt}}$-$O_{\operatorname{acc}}$-diagrams for the hyper-parameter-tuning dataset.
+Fig.~\ref{fig:e2eallsolvers} shows the performance of the approaches from Section~\ref{sec:exdesign:hyper} on $O_{\operatorname{rt}}$-$O_{\operatorname{acc}}$-diagrams for the hyperparameter-tuning dataset.
 Evaluating a particular configuration with Algorithm~\ref{alg:eval} returns a point $\left(O_{\operatorname{rt}},\, O_{\operatorname{acc}}\right)$.
@@
 The plotted lines represent the best-performing configurations per ranking approach (Fig.~\ref{fig:annitraincolorranking}), selection approach (Fig.~\ref{fig:annitraincolorselection}), and stopping criterion (Fig.~\ref{fig:annitraincolorstopping}).
-In particular, we show the Pareto front, i.e., of all configurations that share a particular value of the plotted hyper-parameter, we take the maximum ranking accuracy over all remaining hyper-parameters \emph{not} displayed in the corresponding plot.
+In particular, we show the Pareto front, i.e., of all configurations that share a particular value of the plotted hyperparameter, we take the maximum ranking accuracy over all remaining hyperparameters \emph{not} displayed in the corresponding plot.
 
@@
   \caption{
-    Scatter plot comparing different instantiations of trade-off parameter~$\delta$ for our active-learning framework on the hyper-parameter-tuning dataset.
+    Scatter plot comparing different instantiations of trade-off parameter~$\delta$ for our active-learning framework on the hyperparameter-tuning dataset.
     The x-axis shows the fraction of runtime $O_{\operatorname{rt}}$ of the sample, while the y-axes show the fraction of instances sampled and ranking accuracy, respectively.
@@
 
-Having selected the most promising hyper-parameters, we run our active-learning experiments on the complete Anniversary Track dataset (5355 instances).
+Having selected the most promising hyperparameters, we run our active-learning experiments on the complete Anniversary Track dataset (5355 instances).
 The aforementioned range $\delta \in \left[0.2, 0.8\right]$ only results in two distinct configurations.

### 2024-05-24 paper_jar/paper.tex
section 5

@@
 
-In this section, we evaluate our active-learning framework.
-First, we analyze and tune the different sub-routines of our framework on the tuning dataset.
-Next, we evaluate the best configurations with the full dataset.
-Finally, we analyze the importance of different instance families to our framework.
+In this section, we evaluate our active learning framework.
+First, we analyze and tune the different subroutines of our framework on the tuning dataset.
+Next, we evaluate the best configurations on the full dataset.
+Finally, we analyze the importance of different instance families for our framework.
 
@@
 Our experiments follow the evaluation framework introduced in Section~\ref{sec:exdesign:eval}.
-Fig.~\ref{fig:e2eallsolvers} shows the performance of the approaches from Section~\ref{sec:exdesign:hyper} on $O_{\operatorname{rt}}$-$O_{\operatorname{acc}}$-diagrams for the hyperparameter-tuning dataset.
-Evaluating a particular configuration with Algorithm~\ref{alg:eval} returns a point $\left(O_{\operatorname{rt}},\, O_{\operatorname{acc}}\right)$.
-We do not show intermediate results of the active-learning procedure but only the final results after stopping.
-The plotted lines represent the best-performing configurations per ranking approach (Fig.~\ref{fig:annitraincolorranking}), selection approach (Fig.~\ref{fig:annitraincolorselection}), and stopping criterion (Fig.~\ref{fig:annitraincolorstopping}).
-In particular, we show the Pareto front, i.e., of all configurations that share a particular value of the plotted hyperparameter, we take the maximum ranking accuracy over all remaining hyperparameters \emph{not} displayed in the corresponding plot.
-
-Regarding ranking approaches (Fig.~\ref{fig:annitraincolorranking}), using the predicted $s_3$-induced run\-time-label rank\-ing consistently outperforms the partially observed PAR-2 ranking for each possible value of the trade-off parameter~$\delta$.
-This outcome is expected since selection decisions are not random.
-For example, we might sample more instances of one family if it benefits discrimination of solvers.
-While the partially observed PAR-2 score is skewed, the prediction model can account for this.
+Fig.~\ref{fig:e2eallsolvers} shows the performance of the approaches from Section~\ref{sec:exdesign:hyper} on $O_{\operatorname{rt}}$-$O_{\operatorname{acc}}$ plots for the hyperparameter tuning dataset.
+Evaluating a given configuration with the algorithm~\ref{alg:eval} returns a point $\left(O_{\operatorname{rt}},\, O_{\operatorname{acc}}\right)$.
+We do not show intermediate results of the active learning procedure, but only the final results after stopping.
+The plotted lines represent the best performing configurations per ranking approach (Fig.~\ref{fig:annitraincolorranking}), selection approach (Fig.~\ref{fig:annitraincolorselection}), and stopping criterion (Fig.~\ref{fig:annitraincolorstopping}).
+In particular, we show the Pareto front, i.e., from all configurations that share a particular value of the plotted hyperparameter, we take the maximum ranking accuracy over all remaining hyperparameters \emph{not} shown in the corresponding plot.
+
+Regarding ranking approaches (Fig.~\ref{fig:annitraincolorranking}), using the predicted $s_3$-induced run\-time-label ranking consistently outperforms the partially observed PAR-2 ranking for every possible value of the tradeoff parameter~$\delta$.
+This result is expected, since selection decisions are not random.
+For example, we might sample more instances of a family if it benefits the discrimination of solvers.
+While the partially observed PAR-2 score is biased, the prediction model can account for this.
 
 Regarding the selection approaches (Fig.~\ref{fig:annitraincolorselection}), uncertainty sampling performs best in most cases.
-However, information-gain sampling is beneficial if runtime is strongly favored (small $\delta$; runtime fraction less than \SI{5}{\%}).
-This result aligns with our expectations:
+However, information-gain sampling is advantageous when runtime is strongly favored (small $\delta$; runtime fraction less than \SI{5}{\%}).
+This result is consistent with our expectations:
 Information-gain sampling selects instances that maximize the expected reduction in entropy.
-This means we sample instances revealing similarities between solvers rather than differences, which helps to build a confident model quickly.
-However, the method cannot select helpful instances for distinguishing solvers later.
-Random sampling performs reasonably well but is outperformed by uncertainty sampling in all cases, showing the benefit of actively selecting instances based on a prediction model.
+This means that we sample instances that reveal similarities between solvers rather than differences, which helps to build a confident model quickly.
+However, the method cannot select instances that are useful for distinguishing solvers later.
+Random sampling performs reasonably well, but is outperformed by uncertainty sampling in all cases, demonstrating the benefit of actively selecting instances based on a prediction model.
 
 Regarding the stopping criteria (Fig.~\ref{fig:annitraincolorstopping}), the ranking stopping criterion performs most consistently well.
-If accuracy is strongly favored (very high $\delta$), the Wilcoxon stopping criterion performs better.
-The subset-size stopping criterion performs reasonably well but does not improve beyond a certain accuracy because of sampling a fixed subset of instances.
+When accuracy is strongly favored (very high $\delta$), the Wilcoxon stopping criterion performs better.
+The subset-size stopping criterion performs reasonably well, but does not improve beyond a certain accuracy because it samples a fixed subset of instances.
 
@@
   \caption{
-    Scatter plot comparing different instantiations of trade-off parameter~$\delta$ for our active-learning framework on the hyperparameter-tuning dataset.
-    The x-axis shows the fraction of runtime $O_{\operatorname{rt}}$ of the sample, while the y-axes show the fraction of instances sampled and ranking accuracy, respectively.
-    The color indicates the weighting between different optimization goals $\delta \in \left[0, 1\right]$.
+    Scatterplot comparing different instantiations of the tradeoff parameter~$\delta$ for our active learning framework on the hyperparameter tuning dataset.
+    The x-axis shows the fraction of runtime $O_{\operatorname{rt}}$ of the sample, while the y-axes show the fraction of sampled instances and the ranking accuracy, respectively.
+    The color indicates the weighting between different optimization objectives $\delta \in \left[0, 1\right]$.
     The larger $\delta$, the more we favor accuracy over runtime.
@@
 Fig.~\ref{fig:annitrainoptgoalruntime} shows an interesting consequence of weighting our optimization goals:
-If we, on the one hand, desire to get a \emph{rough} estimate of a solver's performance fast (low $\delta$), approaches favor selecting many \emph{easy} instances.
+On the one hand, if we want to get a \emph{rough} estimate of a solver's performance quickly (low~$\delta$), the approaches favor selecting many \emph{easy} instances.
 In particular, the fraction of sampled instances is larger than the fraction of runtime.
 By having many observations, it is easier to build a model.
-If we, on the other hand, desire to get a \emph{good} estimate of a solver's performance in a moderate amount of time (high $\delta$), approaches favor selecting few, \emph{difficult} instances.
+On the other hand, if we want to get a \emph{good} estimate of a solver's performance in a moderate amount of time (high~$\delta$), approaches favor selecting few, \emph{difficult} instances.
 In particular, the fraction of instances is smaller than the fraction of runtime.
-
-Furthermore, Fig.~\ref{fig:annitrainoptgoalacc} reveals which values make the most sense for $\delta$.
-The range $\delta \in \left[0.2, 0.8\right]$, thereby, corresponds to the points with a runtime fraction between \SI{0.03} and \SI{0.22}.
-We consider this region to be most promising, analogous to the \emph{elbow} method in cluster analysis~\cite{kodinariya2013review}.
+Furthermore, Fig.~\ref{fig:annitrainoptgoalacc} shows which values of $\delta$ make the most sense.
+The range $\delta \in \left[0.2, 0.8\right]$ corresponds to the points with a runtime fraction between \SI{0.03} and \SI{0.22}.
+We consider this region to be the most promising, analogous to the \emph{elbow} method in cluster analysis~\cite{kodinariya2013review}.
 
@@
 
-Having selected the most promising hyperparameters, we run our active-learning experiments on the complete Anniversary Track dataset (5355 instances).
-The aforementioned range $\delta \in \left[0.2, 0.8\right]$ only results in two distinct configurations.
-The best-performing approach for $\delta \in \left[0.2, 0.7\right]$ uses the predicted runtime-label ranking, information-gain sampling, and ranking stopping criterion.
-It can predict a new solver's PAR-2 ranking with \SI{90.48}{\%} accuracy ($O_{\operatorname{acc}}$) in only \SI{5.41}{\%} of the full evaluation time ($O_{\operatorname{rt}}$).
-The best-performing approach for $\delta \in (0.7, 0.8]$ uses the predicted runtime-label ranking, uncertainty sampling, and ranking stopping criterion.
-It can predict a new solver's PAR-2 ranking with \SI{92.33}{\%} accuracy ($O_{\operatorname{acc}}$) in only \SI{10.35}{\%} of the full evaluation time ($O_{\operatorname{rt}}$).
+After selecting the most promising hyperparameters, we run our active learning experiments on the complete Anniversary Track dataset (5355 instances).
+The above range $\delta \in \left[0.2, 0.8\right]$ results in only two different configurations.
+The best performing approach for $\delta \in \left[0.2, 0.7\right]$ uses the predicted runtime label ranking, information-gain sampling, and ranking stopping criterion.
+It can predict the PAR-2 ranking of a new solver with \SI{90.48}{\%} accuracy ($O_{\operatorname{acc}}$) in only \SI{5.41}{\%} of the total evaluation time ($O_{\operatorname{rt}}$).
+The best performing approach for $\delta \in (0.7, 0.8)$ uses the predicted runtime label ranking, uncertainty sampling, and ranking stopping criterion.
+It can predict the PAR-2 ranking of a new solver with \SI{92.33}{\%} accuracy~($O_{\operatorname{acc}}$) in only \SI{10.35}{\%} of the total evaluation time~($O_{\operatorname{rt}}$).
 
@@
   \caption{
-    Performance comparison (on the full dataset) of the best-performing AL approach (\emph{AL}), random sampling of the same runtime fraction with 1000 repetitions (\emph{Random}), and statically selecting the instances most frequently sampled by active-learning approaches (\emph{Most Frequent})
+    Performance comparison (on the full dataset) of the best performing AL approach (\emph{AL}), random sampling of the same runtime fraction with 1000 repetitions (\emph{Random}), and static selection of the instances most frequently sampled by active learning approaches (\emph{Most Frequent}).
   }
@@
 
-Table~\ref{tab:fulldataset} shows how both active-learning approaches (column \emph{AL}) compare against two static baselines:
+Table~\ref{tab:fulldataset} shows how the two active learning approaches (column \emph{AL}) compare to two static baselines:
 \emph{Random} samples instances until it reaches roughly the same fraction of runtime as the AL benchmark sets.
-We repeat sampling 1000 times and report average results.
-\emph{Most Frequent} uses a static benchmark set consisting of those instances most frequently sampled by our active learning approach.
-In particular, we consider the average sampling frequency over all solvers and Pareto-optimal active-learning approaches.
-
-Both our AL approaches perform better than random sampling.
-However, the performance differences are not significant regarding a Wilcoxon signed-rank test with $\alpha = 0.05$ and also depend on the fraction of sampled runtime (cf.
-Fig.~\ref{fig:annitraincolorselection}).
-A clear advantage of our approach is, though, that it indicates when to stop adding further instances, depending on the trade-off parameter~$\delta$.
-While the active-learning results are less strong on the full dataset than on the smaller tuning dataset, they still show the benefit of making benchmark selection dependent on the solvers to distinguish.
-
-A static benchmark using the most frequently AL-sampled instances performs poorly, though, compared to active learning and random sampling.
-This outcome is somewhat expected since the static benchmark does not reflect the right balance of instance families:
-Families whose instances are uniform-randomly selected by AL, e.g., for different solvers, appear less often in this benchmark than families where some instances are sampled more often than others.
+We repeat the sampling 1000 times and report the average results.
+\emph{Most Frequent} uses a static benchmark set consisting of the instances most frequently sampled by our active learning approach.
+In particular, we consider the average sampling frequency across all solvers and Pareto-optimal active learning approaches.
+
+Both of our AL approaches outperform random sampling.
+However, the performance differences are not significant with respect to a Wilcoxon signed-rank test with $\alpha = 0.05$, and also depend on the fraction of the runtime sampled (see Fig.~\ref{fig:annitraincolorselection}).
+However, a clear advantage of our approach is that it indicates when to stop adding new instances, depending on the tradeoff parameter~$\delta$.
+While the active learning results on the full dataset are not as strong as on the smaller tuning dataset, they still show the benefit of making the benchmark selection dependent on the solvers to be distinguished.
+
+A static benchmark utilizing the most frequently AL-sampled instances exhibits suboptimal performance when compared to active learning and random sampling.
+This outcome is somewhat anticipated, given that the static benchmark fails to reflect an appropriate balance of instance families.
+For instance, families whose instances are uniformly randomly selected by AL, e.g., for different solvers, appear less frequently in this benchmark than families where some instances are sampled more often than others.
+% MI: What? I don't understand this sentence.
 
@@
   \includegraphics[width=0.6\textwidth]{../plots/instoccs.pdf}
-  \caption{Frequency of how often individual problem instances are selected within all active-learning benchmarks.
-Instances that are chosen more frequently are intuitively more important to distinguish solver runtimes.}
+  \caption{The frequency with which individual problem instances are selected within all active-learning benchmarks provides insight into the relative importance of these instances.
+Instances that are chosen more frequently are intuitively more important for distinguishing solver runtimes.}
   \label{fig:instoccs}
@@
   \caption{
-    Scatter plot showing the \emph{importance} of different instance families to our framework on the full dataset.
-    The x-axis shows the frequency of instance families in the dataset.
-    The y-axis shows the average frequency of instance families in the samples selected by active learning.
+    The scatter plot illustrates the relative importance of different instance families for ranking.
+The x-axis depicts the frequency of instance families in the dataset, while the y-axis represents the average frequency of instance families in the samples selected by active learning.
 The dashed line represents families that occur with the same frequency in the dataset and samples.
@@
 
-We analyze the frequency of individual instances being selected by our active-learning approach.
-Fig.~\ref{fig:instoccs} shows the distribution of how often individual instances are selected within all active-learning benchmarks.
-We observe that a small fraction of instances is selected very often, while most instances are selected only a few times.
-There is no fixed subset of instances that can perfectly distinguish all solvers, which shows the usefulness of employing an active-learning-based strategy.
-
-Selection decisions of our approach also reveal the importance of different instance families to our framework.
-Fig.~\ref{fig:annifinalfamilies} shows the occurrence of instance families within the dataset and the benchmarks created by active learning.
-We use the best-performing configurations for all $\delta \in \left[0, 1\right]$ and examine the selection decisions by the active-learning approach on the SAT Competition~2022 Anniversary Track dataset~\cite{sat2022}.
-While most families appear with the same fraction in the dataset and the sampled benchmarks, a few outliers need further discussion.
-Problem instances of the families \emph{fpga}, \emph{quasigroup-completion}, and \emph{planning} are especially helpful to our framework in distinguishing solvers.
-Instances of these families are selected over-proportionally in comparison to the full dataset.
-In contrast, instances of the largest family, i.e., \emph{hardware-verification}, roughly appear with the same fraction in the dataset and the sampled benchmarks.
-Finally, instances of the family \emph{cryptography} are less important in distinguishing solvers than their vast weight in the dataset suggests.
-A possible explanation is that these instances are very similar, such that a small fraction of them is sufficient to estimate a solver's performance on all of them.
+Figure~\ref{fig:instoccs} illustrates the distribution of the frequency with which individual instances are selected across all active learning benchmarks.
+It can be observed that a small proportion of instances is selected with great frequency, while the majority of instances are selected on only a few occasions.
+This indicates that there is no fixed subset of instances that can perfectly distinguish all solvers, which highlights the value of employing an active learning-based strategy.
+
+The selection decisions of our approach also reveal the importance of different instance families to our framework.
+Fig.~\ref{fig:annifinalfamilies} illustrates the occurrence of instance families within the dataset and the benchmarks created by active learning.
+We utilize the most effective configurations for all values of $\delta \in \left[0, 1\right]$ and examine the selection decisions made by the active learning approach on the SAT Competition~2022 Anniversary Track dataset~\cite{sat2022}.
+While the majority of families exhibit the same fraction in the dataset and the sampled benchmarks, a few outliers require further discussion.
+Instances of the families \emph{fpga}, \emph{quasigroup-completion}, and \emph{planning} are particularly useful for distinguishing solvers within our framework.
+These instances are selected in greater proportion than in the full dataset.
+In contrast, instances of the largest family, i.e., \emph{hardware-verification}, appear with approximately the same fraction in the dataset and the sampled benchmarks.
+Finally, instances of the family \emph{cryptography} are less important in distinguishing solvers than their substantial representation in the dataset would suggest.
+One possible explanation for this discrepancy is that these instances are highly similar, such that a small fraction of them is sufficient to estimate a solver's performance on all of them.
 

### 2024-05-24 paper_jar/paper.tex
section 6

@@
 In this work, we have addressed the \emph{New-Solver Problem}:
-Given a new solver, we want to find its ranking amidst competitors.
-Our approach provides accurate ranking predictions while needing significantly less runtime than a complete evaluation on a given benchmark set.
-On data from the SAT~Competition~2022 Anniversary Track, we can determine a new solver's PAR-2 ranking with about \SI{92}{\%} accuracy while only needing \SI{10}{\%} of the full-evaluation time.
-We have evaluated several ranking algorithms, instance-selection approaches, and stopping criteria within our sequential active-learning framework.
-We also took a brief look at which instance families are the most prevalent in selection decisions.
-
-Future work may compare further sub-routines for ranking, instance selection, and stopping.
-Additionally, one can apply our evaluation framework to arbitrary computation-intensive problems, e.g., other $\mathcal{NP}$-complete problems than SAT, as all discussed active-learning methods are problem-agnostic.
-Such problems share most of the relevant properties of SAT solving, i.e., there are established instance features, a complete benchmark is expensive, and traditional benchmark selection requires expert knowledge.
-
-From the technical perspective, one could formulate runtime discretization as an optimization problem rather than addressing it empirically.
+Given a new solver, we aim to determine its ranking relative to competitors.
+Our approach provides accurate ranking predictions while requiring significantly less runtime than a complete evaluation on a given benchmark set.
+On data from the SAT~Competition~2022 Anniversary Track, we can determine a new solver's PAR-2 ranking with about \SI{92}{\%} accuracy while only requiring \SI{10}{\%} of the full-evaluation time.
+We have evaluated several ranking algorithms, instance-selection approaches, and stopping criteria within our sequential active learning framework.
+We have also looked at which instance families are the most prevalent in selection decisions.
+
+Future work could compare more sub-routines for ranking, instance selection, and stopping.
+Our evaluation framework can be used for other computation-intensive problems, like other $\mathcal{NP}$-complete problems than SAT.
+Such problems share most of the relevant properties of SAT, like established instance features, a complete benchmark is expensive, and traditional benchmark selection requires expert knowledge.
+
+From a technical perspective, we could formulate the problem of runtime discretization as an optimization problem, rather than addressing it empirically.
 Further, a major shortcoming of our current approach is the lack of parallelization, selecting instances one at a time.
@@
 
+In order to perform benchmarking on a computer cluster with multiple cores, it is necessary to have batches of multiple instances.
+This application demonstrates one of the principal limitations of our approach: it does not permit the concurrent execution of multiple instances, but rather necessitates the benchmarking of a single instance at a time.
+In future work, we intend to investigate models that learn from partial labels, which can be extracted from larger batch sizes in a parallel approach that synchronizes instance selection and model updates.
+
 \backmatter

### 2024-05-24 paper_jar/paper.tex
abstract

@@
 \abstract{
-  Benchmarking is a crucial phase when developing algorithms.
-  This also applies to solvers for the SAT (propositional satisfiability) problem.
+Benchmarking is a crucial phase in the development of algorithms.
+This also applies to SAT solvers for the problem of propositional satisfiability.
 Benchmark selection is about choosing representative problem instances that reliably discriminate solvers based on their runtime.
@@
 Our approach predicts the rank of a new solver among its competitors with minimum runtime and maximum rank prediction accuracy.
-  We evaluated this approach on the Anniversary Track dataset from the 2022 SAT Competition.
-  Our selection approach can predict the rank of a new solver after about \SI{10}{\%} of the time it would take to run the solver on all instances of this dataset, with a prediction accuracy of about \SI{92}{\%}.
-  We also discuss the importance of instance families in the selection process. 
-  Overall, our tool provides a reliable way for solver engineers to determine a new solver's performance efficiently.
+This approach was evaluated on the Anniversary Track dataset from the 2022 SAT Competition.
+Our selection approach can predict the rank of a new solver after approximately \SI{10}{\%} of the time it would take to run the solver on all instances of this dataset, with a prediction accuracy of approximately \SI{92}{\%}.
+Additionally, the importance of instance families in the selection process is discussed.  
+In conclusion, our tool offers a reliable method for solver engineers to assess a new solver's performance in an efficient manner.
 }

### 2024-05-24 paper_jar/paper.tex
intro

@@
 Our approach predicts the rank of a new solver among its competitors with minimum runtime and maximum rank prediction accuracy.
-This approach was evaluated on the Anniversary Track dataset from the 2022 SAT Competition.
+This approach was evaluated on the Anniversary track dataset from the 2022 SAT Competition.
 Our selection approach can predict the rank of a new solver after approximately \SI{10}{\%} of the time it would take to run the solver on all instances of this dataset, with a prediction accuracy of approximately \SI{92}{\%}.
@@
 In particular, our approach enables ranking the new solver amidst a set of existing solvers.
-For this ranking, it is not necessary to predict the exact runtime of the solvers, which would be a more challenging task.
-This is demonstrated in Section~\ref{sec:exdesign:disc-pred}.
-%
+We show that for ranking, it is not necessary to predict the exact per-instance runtime of the solvers, which would be a more challenging task.
 Furthermore, we optimize the \emph{runtime} required for our strategy to arrive at its conclusion.

### 2024-05-24 paper_jar/paper.tex
title

@@
 
-\title[Active Learning for SAT Solver Benchmarking]{Active Learning for SAT Solver Benchmarking} % short and long title
+\title[Active Learning for SAT Solver Benchmarking]{Active Learning for SAT Solver Benchmarking {\large (extended and revised version)}} % short and long title
 
@@
 
-This article is an extended and refined version of our conference paper with the same title, ``Active Learning for SAT Solver Benchmarking''~\cite{fuchs2023active}.
+This article is an extended and refined version of our conference paper with the title ``Active Learning for SAT Solver Benchmarking''~\cite{fuchs2023active}.
 In particular, we significantly extended the analysis of runtime-prediction approaches, which led to our choice of a solver runtime model for use in benchmark instance selection (cf.~Section~\ref{sec:exdesign:disc-pred}).

### 2024-05-25 paper_jar/paper.tex
Adapt names of prediction models
- remove hyphens
- spell out abbreviations

@@
 
-As a baseline for improvement, Table~\ref{tab:regression} shows the runtime prediction performance of two regression models, a random forest regressor and an MLP regressor, using 46 instance features and 27 known solver runtimes.
+As a baseline for improvement, Table~\ref{tab:regression} shows the runtime prediction performance of two regression models, a random forest and a multilayer perceptron, using 46 instance features and 27 known solver runtimes.
 Both models were chosen for their robustness and ability to capture nonlinear dependencies~\cite{breiman2001random}, and both perform poorly, as evidenced by the magnitudes of the mean regression errors.
@@
     \midrule
-    Random-Forest-Regressor & & \SI{1929.84}{s} ($\pm$ \SI{1161.15}{s}) & \phantom{0}\SI{957.36}{s} ($\pm$ \SI{297.21}{s}) \\[0.4ex]
-    MLP-Regressor           & & \SI{2275.20}{s} ($\pm$ \SI{1545.01}{s}) & \SI{1284.30}{s} ($\pm$ \SI{390.57}{s}) \\
+    Random Forest & & \SI{1929.84}{s} ($\pm$ \SI{1161.15}{s}) & \phantom{0}\SI{957.36}{s} ($\pm$ \SI{297.21}{s}) \\[0.4ex]
+    Multilayer Perceptron & & \SI{2275.20}{s} ($\pm$ \SI{1545.01}{s}) & \SI{1284.30}{s} ($\pm$ \SI{390.57}{s}) \\
     \bottomrule
@@
   \begin{tabular}{
-    >{\raggedleft\arraybackslash}m{0.33855\textwidth}
-    >{\centering\arraybackslash}m{0.01\textwidth}
-    >{\centering\arraybackslash}m{0.2562\textwidth}
-    >{\centering\arraybackslash}m{0.2562\textwidth}
+    >{\raggedleft\arraybackslash}m{0.4\textwidth}
+    >{\centering\arraybackslash}m{0.25\textwidth}
+    >{\centering\arraybackslash}m{0.25\textwidth}
   }
     \toprule
-    {Timeout Prediction Models} & & {46 Instance Features Avg. MCC ($\pm$ std)} & {46\,+\,27 Runtime Feat. Avg. MCC ($\pm$ std)} \\
-    \midrule
-    Stacking (QDA\,+\,RF)                 & & 0.6513 ($\pm$ 0.0458) & 0.9527 ($\pm$ 0.0292) \\[0.4ex]
-    Quadratic-DA (QDA)                    & & 0.2593 ($\pm$ 0.1310) & 0.9290 ($\pm$ 0.0339) \\[0.4ex]
-    Random-Forest-Classifier (RF)         & & 0.6607 ($\pm$ 0.0547) & 0.8530 ($\pm$ 0.0479) \\[0.4ex]
-    Ada-Boost                             & & 0.5412 ($\pm$ 0.0612) & 0.8384 ($\pm$ 0.0444) \\[0.4ex]
-    Decision-Tree                         & & 0.5980 ($\pm$ 0.0423) & 0.8059 ($\pm$ 0.0707) \\[0.4ex]
-    Logistic-Regression                   & & 0.2031 ($\pm$ 0.0728) & 0.8052 ($\pm$ 0.1018) \\[0.4ex]
-    $k$NN-Classifier                      & & 0.5108 ($\pm$ 0.0387) & 0.7885 ($\pm$ 0.1521) \\[0.4ex]
-    MLP-Classifier                        & & 0.1293 ($\pm$ 0.0718) & 0.7760 ($\pm$ 0.1408) \\[0.4ex]
-    Support-Vector-Machines               & & 0.0595 ($\pm$ 0.0554) & 0.7757 ($\pm$ 0.2149) \\[0.4ex]
-    Naive-Bayes                           & & 0.1173 ($\pm$ 0.0872) & 0.7306 ($\pm$ 0.1394) \\
+    {Timeout Prediction Models} & {46 Instance Features Avg. MCC ($\pm$ std)} & {46\,+\,27 Runtime Feat. Avg. MCC ($\pm$ std)} \\
+    \midrule
+    Stacking (QDA\,+\,RF)                 & 0.6513 ($\pm$ 0.0458) & 0.9527 ($\pm$ 0.0292) \\[0.4ex]
+    Quadratic Discriminant Analysis (QDA)                    & 0.2593 ($\pm$ 0.1310) & 0.9290 ($\pm$ 0.0339) \\[0.4ex]
+    Random Forest (RF)         & 0.6607 ($\pm$ 0.0547) & 0.8530 ($\pm$ 0.0479) \\[0.4ex]
+    AdaBoost                             & 0.5412 ($\pm$ 0.0612) & 0.8384 ($\pm$ 0.0444) \\[0.4ex]
+    Decision Tree                         & 0.5980 ($\pm$ 0.0423) & 0.8059 ($\pm$ 0.0707) \\[0.4ex]
+    Logistic Regression                   & 0.2031 ($\pm$ 0.0728) & 0.8052 ($\pm$ 0.1018) \\[0.4ex]
+    $k$NN                   & 0.5108 ($\pm$ 0.0387) & 0.7885 ($\pm$ 0.1521) \\[0.4ex]
+    Multilayer Perceptron                      & 0.1293 ($\pm$ 0.0718) & 0.7760 ($\pm$ 0.1408) \\[0.4ex]
+    Support Vector Machine               & 0.0595 ($\pm$ 0.0554) & 0.7757 ($\pm$ 0.2149) \\[0.4ex]
+    Naive Bayes                           & 0.1173 ($\pm$ 0.0872) & 0.7306 ($\pm$ 0.1394) \\
     \bottomrule
@@
   \begin{tabular}{
-    >{\raggedleft\arraybackslash}m{0.33855\textwidth}
-    >{\centering\arraybackslash}m{0.01\textwidth}
-    >{\centering\arraybackslash}m{0.2562\textwidth}
-    >{\centering\arraybackslash}m{0.2562\textwidth}
+    >{\raggedleft\arraybackslash}m{0.4\textwidth}
+    >{\centering\arraybackslash}m{0.25\textwidth}
+    >{\centering\arraybackslash}m{0.25\textwidth}
   }
     \toprule
-    {$C_3$ Cluster Prediction Models} & & {46\,+\,27 Runtime Feat. Avg. MCC ($\pm$ std)} & {46\,+\,27\,+\,27 Clus. Feat. Avg. MCC ($\pm$ std)} \\
-    \midrule
-    Stacking (QDA\,+\,RF)                 & & 0.7464 ($\pm$ 0.0497) & 0.8380 ($\pm$ 0.0634) \\[0.4ex]
-    Quadratic-DA (QDA)                    & & 0.6903 ($\pm$ 0.0607) & 0.7738 ($\pm$ 0.0459) \\[0.4ex]
-    Random-Forest-Classifier (RF)         & & 0.7116 ($\pm$ 0.0385) & 0.7841 ($\pm$ 0.0469) \\[0.4ex]
-    \midrule
-    {$C_4$ Cluster Prediction Models} & & {46\,+\,27 Runtime Feat. Avg. MCC ($\pm$ std)} & {46\,+\,27\,+\,27 Clus. Feat. Avg. MCC ($\pm$ std)} \\
-    \midrule
-    Stacking (QDA\,+\,RF)                 & & 0.6562 ($\pm$ 0.0538) & 0.7366 ($\pm$ 0.0377) \\[0.4ex]
-    Quadratic-DA (QDA)                    & & 0.6007 ($\pm$ 0.0765) & 0.6872 ($\pm$ 0.0469) \\[0.4ex]
-    Random-Forest-Classifier (RF)         & & 0.6033 ($\pm$ 0.0455) & 0.6828 ($\pm$ 0.0527) \\[0.4ex]
-    \midrule
-    {$C_5$ Cluster Prediction Models} & & {46\,+\,27 Runtime Feat. Avg. MCC ($\pm$ std)} & {46\,+\,27\,+\,27 Clus. Feat. Avg. MCC ($\pm$ std)} \\
-    \midrule
-    Stacking (QDA\,+\,RF)                 & & 0.5835 ($\pm$ 0.0615) & 0.6396 ($\pm$ 0.0553) \\[0.4ex]
-    Quadratic-DA (QDA)                    & & 0.5508 ($\pm$ 0.0669) & 0.5881 ($\pm$ 0.0912) \\[0.4ex]
-    Random-Forest-Classifier (RF)         & & 0.5420 ($\pm$ 0.0459) & 0.5895 ($\pm$ 0.0475) \\
+    {$C_3$ Cluster Prediction Models} & {46\,+\,27 Runtime Feat. Avg. MCC ($\pm$ std)} & {46\,+\,27\,+\,27 Clus. Feat. Avg. MCC ($\pm$ std)} \\
+    \midrule
+    Stacking (QDA\,+\,RF)                 & 0.7464 ($\pm$ 0.0497) & 0.8380 ($\pm$ 0.0634) \\[0.4ex]
+    Quadratic Discriminant Analysis (QDA)                     & 0.6903 ($\pm$ 0.0607) & 0.7738 ($\pm$ 0.0459) \\[0.4ex]
+    Random Forest (RF)         & 0.7116 ($\pm$ 0.0385) & 0.7841 ($\pm$ 0.0469) \\[0.4ex]
+    \midrule
+    {$C_4$ Cluster Prediction Models} & {46\,+\,27 Runtime Feat. Avg. MCC ($\pm$ std)} & {46\,+\,27\,+\,27 Clus. Feat. Avg. MCC ($\pm$ std)} \\
+    \midrule
+    Stacking (QDA\,+\,RF)                 & 0.6562 ($\pm$ 0.0538) & 0.7366 ($\pm$ 0.0377) \\[0.4ex]
+    Quadratic Discriminant Analysis (QDA)                     & 0.6007 ($\pm$ 0.0765) & 0.6872 ($\pm$ 0.0469) \\[0.4ex]
+    Random Forest (RF)         & 0.6033 ($\pm$ 0.0455) & 0.6828 ($\pm$ 0.0527) \\[0.4ex]
+    \midrule
+    {$C_5$ Cluster Prediction Models} & {46\,+\,27 Runtime Feat. Avg. MCC ($\pm$ std)} & {46\,+\,27\,+\,27 Clus. Feat. Avg. MCC ($\pm$ std)} \\
+    \midrule
+    Stacking (QDA\,+\,RF)                 & 0.5835 ($\pm$ 0.0615) & 0.6396 ($\pm$ 0.0553) \\[0.4ex]
+    Quadratic Discriminant Analysis (QDA)                      & 0.5508 ($\pm$ 0.0669) & 0.5881 ($\pm$ 0.0912) \\[0.4ex]
+    Random Forest (RF)         & 0.5420 ($\pm$ 0.0459) & 0.5895 ($\pm$ 0.0475) \\
     \bottomrule

### 2024-05-25 paper_jar/paper.tex
Review changes in non-extended parts of JAR paper
I.e., excluding the two subsection containing new material. Revert or fix
some changes.

@@
 Benchmarking is a crucial phase in the development of algorithms.
-This also applies to SAT solvers for the problem of propositional satisfiability.
+This also applies to solvers for the propositional satisfiability (SAT) problem.
 Benchmark selection is about choosing representative problem instances that reliably discriminate solvers based on their runtime.
@@
 Our approach predicts the rank of a new solver among its competitors with minimum runtime and maximum rank prediction accuracy.
-This approach was evaluated on the Anniversary track dataset from the 2022 SAT Competition.
+We evaluated this approach on the Anniversary Track dataset from the SAT Competition 2022.
 Our selection approach can predict the rank of a new solver after approximately \SI{10}{\%} of the time it would take to run the solver on all instances of this dataset, with a prediction accuracy of approximately \SI{92}{\%}.
-Additionally, the importance of instance families in the selection process is discussed.  
-In conclusion, our tool offers a reliable method for solver engineers to assess a new solver's performance in an efficient manner.
+Additionally, we discuss the importance of instance families in the selection process.
+In conclusion, our tool offers a reliable method for solver engineers to assess a new solver's performance efficiently.
 }
@@
 \label{sec:intro}
-% TODO: if appropriate, extend with motivation/results regarding runtime discretization and prediction models
 
@@
 Benchmarking is, however, quite expensive regarding the runtime of experiments.
-While it is still feasible to benchmark a single or fixed amount of SAT solvers, developing new, competitive SAT solvers requires extensive experimentation with a variety of ideas, which often results in a combinatorial explosion of the configuration space of each algorithm~\cite{HutterHL11}.
+While it is still feasible to benchmark a single or small number of SAT solvers, developing new, competitive SAT solvers requires extensive experimentation with a variety of ideas.
+The latter often results in a combinatorial explosion of the configuration space of each algorithm~\cite{HutterHL11}.
 In particular, a new solver idea is rarely best on the first try.
@@
 Instead, they focus on selecting a representative set of instances for ranking solvers~\cite{Gelder11,HoosKSS13}.
-With regard to the ranking of solvers, SAT Competitions typically employ the \mbox{PAR-2} score, which, given a time limit of $\tau$, is the average runtime with a penalty of $2 \tau$ for timeouts~\cite{FroleyksHIJS21}.
+In this regard, SAT Competitions typically employ the \mbox{PAR-2} score, which, given a time limit of $\tau$, is the average runtime with a penalty of $2 \tau$ for timeouts~\cite{FroleyksHIJS21}.
 
@@
 
-The approach presented in this paper meets several criteria for benchmarking that are considered desirable.
-Rather than outputting a binary classification, namely whether the new solver is worse than an existing solver or not, we provide a scoring function that shows by which margin a solver is worse and how similar it is to existing solvers.
-In particular, our approach enables ranking the new solver amidst a set of existing solvers.
+The approach presented in this paper meets several criteria for benchmarking that are considered desirable (cf.~Table~\ref{tab:requirements}).
+Rather than outputting a binary classification, namely whether the new solver is worse than an existing solver or not, we provide a \emph{scoring} function that shows by which margin a solver is worse and how similar it is to existing solvers.
+In particular, our approach enables \emph{ranking} the new solver amidst a set of existing solvers.
 We show that for ranking, it is not necessary to predict the exact per-instance runtime of the solvers, which would be a more challenging task.
@@
 This article is an extended and refined version of our conference paper with the title ``Active Learning for SAT Solver Benchmarking''~\cite{fuchs2023active}.
-In particular, we significantly extended the analysis of runtime-prediction approaches, which led to our choice of a solver runtime model for use in benchmark instance selection (cf.~Section~\ref{sec:exdesign:disc-pred}).
+In particular, we significantly extended the analysis of runtime-prediction approaches, which determined our chosen solver runtime model for instance selection (cf.~Section~\ref{sec:exdesign:disc-pred}).
 Furthermore, we expanded the evaluation of which benchmark instances are selected by our approach (cf.~Section~\ref{sec:eval:instance}).
-% TODO: before submission, we should check which other changes we made and list them here accordingly
 
@@
 Benchmarking also involves a number of interchangeable elements, including the performance measures used, the way the measures are aggregated, and the way missing values are handled.
-It is possible that questionable research practices could alter these elements a-posteriori in order to meet expectations, thereby skewing the results~\cite{NiesslHWCB22}.
+Questionable research practices could alter these elements a-posteriori in order to meet expectations, thereby skewing the results~\cite{NiesslHWCB22}.
 
@@
 Froleyks et al. describe benchmark selection in recent SAT~competitions~\cite{FroleyksHIJS21}.
-Manthey and Möhle propose a feature-based approach to remove redundancy from competition benchmarks~\cite{manthey2016better}, while M{\i}s{\i}r presents a similar feature-based approach to reduce benchmarks by matrix factorization and clustering~\cite{misir2021benchmark}.
-
-In their study, Hoos et al.~\cite{HoosKSS13} discuss the desirable properties of SAT benchmark instances.
+Manthey and Möhle propose an approach to remove redundancy from competition benchmarks~\cite{manthey2016better}, considering feature equivalence of formulas.
+M{\i}s{\i}r also presents a feature-based approach to reduce benchmarks, using matrix factorization and clustering~\cite{misir2021benchmark}.
+
+Hoos et al.~\cite{HoosKSS13} discuss desirable properties of SAT benchmark instances.
 They identify three key selection criteria: instance variety to avoid over-fitting, adapted instance hardness (not too easy but also not too hard), and the avoidance of duplicate instances. 
 To filter instances that are too similar, they employ a distance-based approach with the SATzilla features~\cite{XuHHL08,features}.
-It should be noted, however, that the approach does not optimize for benchmark runtime.
-Instead, instances are selected at random, with the exception of constraints on instance hardness and feature distance.
+It should be noted, however, that the approach does not optimize for benchmark \emph{runtime}.
+Instead, instances are selected \emph{at random}, with the exception of constraints on instance hardness and feature distance.
 
@@
 Furthermore, a significant shortcoming is the lack of \emph{scalability} with the number of solvers.
-Comparing only pairs of solvers, evaluating a new solver requires sampling a separate benchmark for each existing solver.
+Comparing only pairs of solvers, evaluating a new solver with their method requires sampling a separate benchmark for each existing solver.
 In contrast, our approach allows comparing a new solver against a set of existing solvers by sampling only one benchmark.
-In contrast to our approach, which allows for the comparison of a new solver against a set of existing solvers by sampling only one benchmark, their pairwise method requires the evaluation of a new solver to be conducted on a separate benchmark for each existing solver.
 
@@
 Without AL methods, it is not evident which instances to label and which not.
-The objective is to maximize the utility of instances to our model, which is measured by the rank prediction accuracy.
-Concurrently, the objective is to minimize the cost of label acquisition, which is measured by the instance's predicted runtime.
-Consequently, the objective is to develop an accurate prediction model without the necessity of labeling each data point.
+The first goal is to maximize the utility of instances to our model, which is measured by the rank prediction accuracy.
+The second goal is to minimize the cost of label acquisition, which is measured by the instance's predicted runtime.
+Consequently, the overall objective is to develop an accurate prediction model without the necessity of labeling each data point.
 
@@
 When the stopping criterion is met (Line~3), the benchmarking loop is terminated and the final score of $\hat{a}$ is estimated (Line~8). 
-This score, along with the acquired instances and runtime measurements, is returned by Algorithm~\ref{algALBenchmark}.
+Algorithm~\ref{algALBenchmark} returns this score, along with the acquired instances and runtime measurements.
 
@@
 This prediction function powers instance selection as described in Section~\ref{sec:main:selection}.
-During model updates (Algorithm~\ref{algALBenchmark},~Line~7), the prediction function $f$ is trained to predict a transformed version of the acquired runtime data set~$\mathcal{R}$.
-The transformation of the runtime data set is described in the subsequent section.
+During model updates (Algorithm~\ref{algALBenchmark},~Line~7), the prediction function $f$ is trained to predict a transformed version of the acquired runtime dataset~$\mathcal{R}$.
+We describe the transformation of the runtime dataset in the subsequent section.
 The features described in Section~\ref{sec:exdesign:data} serve as the input to the model.
@@
 \paragraph{Uncertainty Sampling}
+
 The uncertainty sampling strategy selects the instance that is closest to the model's decision boundary.
@@
 \paragraph{Information-Gain Sampling}
+
 The information-gain sampling strategy selects the instance with the highest expected entropy reduction regarding the runtime labels of the instance.
@@
 \paragraph{Wilcoxon Stopping Criterion}
+
 The Wilcoxon stopping criterion is employed to halt the active-learning process based on the confidence that the predicted runtime labels of the new solver are sufficiently different from those of existing solvers.
@@
 \paragraph{Ranking Stopping Criterion}
-In contrast, the stopping criterion for the ranking process is less sophisticated than the previous one.
+
+In contrast, the ranking stopping criterion is less sophisticated than the previous one.
 In this case, the active learning process is terminated if the ranking induced by the model's predictions (Equation~\ref{eq:rankingeq}) remains unchanged within the last~$l$~iterations.
@@
 In light of the preceding instantiations of Algorithm~\ref{algALBenchmark}, this section outlines the experimental design employed in the presented study.
-This includes an overview of the evaluation framework, the data sets employed, the hyperparameter selections, and the implementation details.
+This includes an overview of the evaluation framework, the datasets employed, the hyperparameter selections, and the implementation details.
 
@@
 As described in Section~\ref{sec:main:model}, a prediction model provides us with an estimated scoring value for the new solver~$\hat{a}$.
-%
 To evaluate the efficacy of a concrete instantiation of Algorithm~\ref{algALBenchmark}, which entails a specific choice for all subroutines, we perform cross-validation on our set of solvers.
@@
 After running our active-learning framework for solver~$\hat{a}$ (Line~4), we compute the value of both our optimization goals, i.e., ranking accuracy and runtime.
-%
-The \emph{ranking accuracy}, expressed in the interval $[0,1]$ (higher is better), is defined as the fraction of correctly ranked pairs of elements, namely, $\left(\hat{a}, a\right)$, where $a$ belongs to the set of all solvers $\mathcal{A}$ ranked by the ground-truth scoring function $\operatorname{par_2}$ (Lines~5-8).
-%
+The \emph{ranking accuracy} $O_{\operatorname{acc}} \in \left[0, 1\right]$ (higher is better) is defined as the fraction of correctly ranked pairs of elements, namely, $\left(\hat{a}, a\right)$, regarding the ground-truth scoring function $\operatorname{par_2}$, where $a$ belongs to the set of all solvers $\mathcal{A}$ (Lines~5-8).
 The \emph{fraction of runtime} required by the algorithm to reach its conclusion is represented by $O_{\operatorname{rt}}$, which lies within the interval $[0, 1]$ (lower is better).
@@
 %
-%
+
 \subsection{Data}
@@
 
-In our experiments we work with the dataset of the SAT~Competition~2022 Anniversary Track~\cite{sat2022}.
+In our experiments, we work with the dataset of the SAT~Competition~2022 Anniversary Track~\cite{sat2022}.
 The dataset consists of 5355 instances with corresponding runtime data of 28 sequential SAT solvers.
@@
 \subparagraph{Ranking}
-Regarding \emph{ranking}, we experiment with the following instantiations.
+
+Regarding \emph{ranking} (cf.~Section~\ref{sec:main:model}), we experiment with the following instantiations.
 
@@
 \subparagraph{Selection}
-We experiment with the following instantiations of instance \emph{selection}.
-We experiment with the following instantiations of the instance \emph{selection} criteria.
+
+We experiment with the following instantiations of the instance \emph{selection} (cf. Section~\ref{sec:main:selection}).
 Since the potential runtime of the experiments is orders of magnitude larger than the model update time, we consider incrementing our benchmark one instance at a time rather than using batches, which is also proposed in recent advances in active learning~\cite{SinhaED19,2019gaal}.
@@
 \subparagraph{Stopping}
-We evaluate the following \emph{stopping} criteria.
+
+We evaluate the following \emph{stopping} (cf. Section~\ref{sec:main:stopping}) criteria.
 
@@
 \label{sec:exdesign:disc-pred}
+
 One of the most important parts of our framework is the choice of solver runtime model to inform active learning selection decisions.
@@
 We denote this adapted solver scoring function which assigns a solver to its cluster for each instance with $s_k$ as defined in Equation~\eqref{eq:rankingeq}.
-%
 In preliminary experiments, the most \emph{useful} labels were produced with hierarchical clustering and a log-single-link criterion.
@@
 Both of these models can learn non-linear relationships between the instance features and the runtime labels.
-Figure~\ref{fig:confusion-matrices} shows the confusion matrices of the stacking ensemble for predicting timeouts and $C_3$ clusters.
+Fig.~\ref{fig:confusion-matrices} shows the confusion matrices of the stacking ensemble for predicting timeouts and $C_3$ clusters.
 Both perform reasonably well, with the $C_3$ cluster prediction model performing slightly worse than the timeout prediction model.
@@
 Table~\ref{tab:ranking} shows the ranking of the solvers in the SAT Competition~2022 Anniversary Track~\cite{sat2022} according to the standard PAR-2 runtime score and according to our discretized $s_3$ score.
-We observe that our ranking approach correctly decides almost all (about \SI{97.45}{\%}; $\sigma = \SI{3.68}{\%}$) solver pairs, i.e. which solver is faster among two given solvers.
+We observe that our ranking approach correctly decides almost all (about \SI{97.45}{\%}; $\sigma = \SI{3.68}{\%}$) solver pairs, i.e., which solver is faster among two given solvers.
 In particular, the Spearman correlation of $s_3$ and PAR-2 ranking is about \SI{0.988}{}, which is very close to the optimal value of~1~\cite{de2016comparing}.
@@
 We also analyzed how runtime discretization affects the distinguishability of solver pairs.
-Figure~\ref{fig:solver-confusion-matrices} shows the confusion matrices of solver pairs with significant differences in their PAR-2 scores, timeouts, and $s_3$ scores.
+Fig.~\ref{fig:solver-confusion-matrices} shows the confusion matrices of solver pairs with significant differences in their PAR-2 scores, timeouts, and $s_3$ scores.
 According to a Wilcoxon signed-rank test with a level of $\alpha = 0.05$, \SI{87.83}{\%} of the solver pairs have significantly different scores after discretization, only a small decrease compared to \SI{89.95}{\%} before discretization.
@@
 Our experiments follow the evaluation framework introduced in Section~\ref{sec:exdesign:eval}.
-Fig.~\ref{fig:e2eallsolvers} shows the performance of the approaches from Section~\ref{sec:exdesign:hyper} on $O_{\operatorname{rt}}$-$O_{\operatorname{acc}}$ plots for the hyperparameter tuning dataset.
-Evaluating a given configuration with the algorithm~\ref{alg:eval} returns a point $\left(O_{\operatorname{rt}},\, O_{\operatorname{acc}}\right)$.
+Fig.~\ref{fig:e2eallsolvers} shows the performance of the approaches from Section~\ref{sec:exdesign:hyper} on $O_{\operatorname{rt}}$-$O_{\operatorname{acc}}$ plots for the hyperparameter-tuning dataset.
+Evaluating a given configuration with Algorithm~\ref{alg:eval} returns a point $\left(O_{\operatorname{rt}},\, O_{\operatorname{acc}}\right)$.
 We do not show intermediate results of the active learning procedure, but only the final results after stopping.
@@
   \caption{
-    Scatterplot comparing different instantiations of the tradeoff parameter~$\delta$ for our active learning framework on the hyperparameter tuning dataset.
+    Scatter plot comparing different instantiations of the tradeoff parameter~$\delta$ for our active learning framework on the hyperparameter-tuning dataset.
     The x-axis shows the fraction of runtime $O_{\operatorname{rt}}$ of the sample, while the y-axes show the fraction of sampled instances and the ranking accuracy, respectively.
@@
 It can predict the PAR-2 ranking of a new solver with \SI{90.48}{\%} accuracy ($O_{\operatorname{acc}}$) in only \SI{5.41}{\%} of the total evaluation time ($O_{\operatorname{rt}}$).
-The best performing approach for $\delta \in (0.7, 0.8)$ uses the predicted runtime label ranking, uncertainty sampling, and ranking stopping criterion.
+The best performing approach for $\delta \in (0.7, 0.8]$ uses the predicted runtime label ranking, uncertainty sampling, and ranking stopping criterion.
 It can predict the PAR-2 ranking of a new solver with \SI{92.33}{\%} accuracy~($O_{\operatorname{acc}}$) in only \SI{10.35}{\%} of the total evaluation time~($O_{\operatorname{rt}}$).
@@
 Both of our AL approaches outperform random sampling.
-However, the performance differences are not significant with respect to a Wilcoxon signed-rank test with $\alpha = 0.05$, and also depend on the fraction of the runtime sampled (see Fig.~\ref{fig:annitraincolorselection}).
+However, the performance differences are not significant with respect to a Wilcoxon signed-rank test with $\alpha = 0.05$, and also depend on the fraction of the runtime sampled (cf.~Fig.~\ref{fig:annitraincolorselection}).
 However, a clear advantage of our approach is that it indicates when to stop adding new instances, depending on the tradeoff parameter~$\delta$.
@@
   \includegraphics[width=0.6\textwidth]{../plots/instoccs.pdf}
-  \caption{The frequency with which individual problem instances are selected within all active-learning benchmarks provides insight into the relative importance of these instances.
+  \caption{Frequency with which individual problem instances are selected within all active-learning benchmarks.
 Instances that are chosen more frequently are intuitively more important for distinguishing solver runtimes.}
@@
   \caption{
-    The scatter plot illustrates the relative importance of different instance families for ranking.
+    Scatter plot illustrating the relative importance of different instance families for ranking.
 The x-axis depicts the frequency of instance families in the dataset, while the y-axis represents the average frequency of instance families in the samples selected by active learning.
@@
 
-Figure~\ref{fig:instoccs} illustrates the distribution of the frequency with which individual instances are selected across all active learning benchmarks.
-It can be observed that a small proportion of instances is selected with great frequency, while the majority of instances are selected on only a few occasions.
-This indicates that there is no fixed subset of instances that can perfectly distinguish all solvers, which highlights the value of employing an active learning-based strategy.
+Fig.~\ref{fig:instoccs} illustrates the distribution of the frequency with which individual instances are selected across all active learning benchmarks.
+It can be observed that a small proportion of instances is selected with great frequency, while most instances are selected on only a few occasions.
+This indicates that there is no fixed subset of instances that can perfectly distinguish all solvers, which highlights the value of employing an active-learning-based strategy.
 
@@
 Fig.~\ref{fig:annifinalfamilies} illustrates the occurrence of instance families within the dataset and the benchmarks created by active learning.
-We utilize the most effective configurations for all values of $\delta \in \left[0, 1\right]$ and examine the selection decisions made by the active learning approach on the SAT Competition~2022 Anniversary Track dataset~\cite{sat2022}.
-While the majority of families exhibit the same fraction in the dataset and the sampled benchmarks, a few outliers require further discussion.
+We use the best performing configurations for all values of $\delta \in \left[0, 1\right]$ and examine the selection decisions made by the active learning approach on the SAT Competition~2022 Anniversary Track dataset~\cite{sat2022}.
+While most families exhibit the same fraction in the dataset and the sampled benchmarks, a few outliers require further discussion.
 Instances of the families \emph{fpga}, \emph{quasigroup-completion}, and \emph{planning} are particularly useful for distinguishing solvers within our framework.
@@
 Future work could compare more sub-routines for ranking, instance selection, and stopping.
-Our evaluation framework can be used for other computation-intensive problems, like other $\mathcal{NP}$-complete problems than SAT.
+Further, our evaluation framework can be used for other computation-intensive problems, like other $\mathcal{NP}$-complete problems than SAT.
 Such problems share most of the relevant properties of SAT, like established instance features, a complete benchmark is expensive, and traditional benchmark selection requires expert knowledge.
@@
 
-In order to perform benchmarking on a computer cluster with multiple cores, it is necessary to have batches of multiple instances.
-This application demonstrates one of the principal limitations of our approach: it does not permit the concurrent execution of multiple instances, but rather necessitates the benchmarking of a single instance at a time.
+To perform benchmarking on a computer cluster with multiple cores, it is beneficial to work with batches of multiple instances.
+In contrast, our approach is limited to benchmarking with a single instance at a time.
 In future work, we intend to investigate models that learn from partial labels, which can be extracted from larger batch sizes in a parallel approach that synchronizes instance selection and model updates.

### 2024-05-25 paper_jar/paper.tex
Add more section references
We have enough space in the journal version.

@@
 
+\subparagraph{Motivation}
+
 One of the main phases of algorithm engineering is benchmarking.
@@
 In this regard, SAT Competitions typically employ the \mbox{PAR-2} score, which, given a time limit of $\tau$, is the average runtime with a penalty of $2 \tau$ for timeouts~\cite{FroleyksHIJS21}.
+
+\subparagraph{Problem statement}
 
@@
 
+\subparagraph{Experiments}
+
 We evaluated our approach using the SAT Competition~2022 Anniversary Track dataset~\cite{sat2022}, consisting of 5355~instances and runtimes of 28~solvers.
@@
 
+\subparagraph{Outline}
+
+The remainder of this paper is organized as follows:
+Section~\ref{sec:related} reviews related work.
+Section~\ref{sec:main} describes our approach.
+Section~\ref{sec:exdesign} introduces the experimental design.
+Section~\ref{sec:eval} evaluates the experimental results.
+Section~\ref{sec:conclusion} concludes.
+
+\subparagraph{Disclaimer}
+
 This article is an extended and refined version of our conference paper with the title ``Active Learning for SAT Solver Benchmarking''~\cite{fuchs2023active}.
@@
 \section{Related Work}
+\label{sec:related}
 
@@
 
-The following sections present a discussion of related work from the areas of static benchmark selection, algorithm configuration, incremental benchmark selection, and active learning.
+The following sections present a discussion of related work from the areas of static benchmark selection (cf.~Section~\ref{sec:related:static-selection}), algorithm configuration (cf.~Section~\ref{sec:related:algo-configuration}), incremental benchmark selection (cf.~Section~\ref{sec:related:incremental-selection}), and active learning (cf.~Section~\ref{sec:related:active-learning}).
 Table~\ref{tab:requirements} compares the most relevant approaches, which all pursue slightly different goals.
@@
 \subsection{Static Benchmark Selection}
+\label{sec:related:static-selection}
 
@@
 \subsection{Algorithm Configuration}
+\label{sec:related:algo-configuration}
 
@@
 \subsection{Incremental Benchmark Selection}
+\label{sec:related:incremental-selection}
 
@@
 \subsection{Active Learning}
+\label{sec:related:active-learning}
 
@@
   \end{tabular}
-
   \caption{Types of machine learning (depiction inspired by Rubens~et.al.~\cite{RubensESK15}).}
@@
 In contrast, the ranking stopping criterion is less sophisticated than the previous one.
-In this case, the active learning process is terminated if the ranking induced by the model's predictions (Equation~\ref{eq:rankingeq}) remains unchanged within the last~$l$~iterations.
+In this case, the active learning process is terminated if the ranking induced by the model's predictions (Equation~\eqref{eq:rankingeq}) remains unchanged within the last~$l$~iterations.
 However, it should be noted that the concrete values of the predicted score $s_{\hat a}$ may still change.
@@
 In light of the preceding instantiations of Algorithm~\ref{algALBenchmark}, this section outlines the experimental design employed in the presented study.
-This includes an overview of the evaluation framework, the datasets employed, the hyperparameter selections, and the implementation details.
+This includes an overview of the evaluation framework (cf.~Section~\ref{sec:exdesign:eval}), the datasets employed (cf.~Section~\ref{sec:exdesign:data}), the hyperparameter selections (cf.~Section~\ref{sec:exdesign:hyper}), the runtime discretization (cf.~Section~\ref{sec:exdesign:disc-pred}), and the implementation (cf.~Section~\ref{sec:exdesign:impl}).
 
@@
 \subsection{Implementation Details}
+\label{sec:exdesign:impl}
 
@@
 In this section, we evaluate our active learning framework.
-First, we analyze and tune the different subroutines of our framework on the tuning dataset.
-Next, we evaluate the best configurations on the full dataset.
-Finally, we analyze the importance of different instance families for our framework.
+First, we analyze and tune the different subroutines of our framework on the tuning dataset (cf.~Section~\ref{sec:eval:hyper}).
+Next, we evaluate the best configurations on the full dataset (cf.~Section~\ref{sec:eval:full}).
+Finally, we analyze the importance of instance families for our framework (cf.~Section~\ref{sec:eval:instance}).
 
 \subsection{Hyperparameter Analysis}
+\label{sec:eval:hyper}
 
@@
 \subsection{Full-Dataset Evaluation}
+\label{sec:eval:full}
 
@@
 \section{Conclusions and Future Work}
+\label{sec:conclusion}
 

### 2024-05-25 paper_jar/paper.tex
Review changes in extended part of JAR paper
I.e., Sections 4.4 and 5.3. Not many changes in 5.3 but lots in 4.4:

- use (sub)paragraphs to structure this long subsection
- improve flow between paragraphs
- explain ML concepts more
- interpret figures and tables more
- improve table formatting (e.g., proper two-line headers in some tables
  rather than hoping that line break will appear at expected position)

@@
 \subsubsection{Runtime Transformation}
+\label{sec:main:model:transformation}
 
@@
 \subsubsection{Ranking Solvers}
+\label{sec:main:model:ranking}
 
@@
 
-One of the most important parts of our framework is the choice of solver runtime model to inform active learning selection decisions.
-In this section, we detail how clustering solver runtimes can lead to nuanced runtime labels, and which machine learning models are best at predicting them.
-
-As a baseline for improvement, Table~\ref{tab:regression} shows the runtime prediction performance of two regression models, a random forest and a multilayer perceptron, using 46 instance features and 27 known solver runtimes.
-Both models were chosen for their robustness and ability to capture nonlinear dependencies~\cite{breiman2001random}, and both perform poorly, as evidenced by the magnitudes of the mean regression errors.
-An RMSE error of about \SI{2000}{s} is quite poor compared to solver runtimes that range from 0 to \SI{5000}{s}.
-Therefore, we decided to discretize the runtimes.
-%
-\begin{table}[b]
-  \centering
-  \caption{Regression performance of models trained on 46 instance and 27 runtime features. We report the RMSE and MAE metrics on the test sets.}
+One of the most important parts of our framework is the solver runtime model (cf.~Section~\ref{sec:main:model}), which influences instance selection in active learning.
+In this section, we detail how clustering solver runtimes can lead to informative runtime labels, and which machine learning models are best at predicting them.
+To this end, we analyze prediction performance for estimating runtimes, timeouts, and discretized runtimes.
+
+\paragraph{Regressing runtime}
+
+\begin{table}[tb]
+	\centering
+	\caption{Test-set prediction performance for predicting solver runtime (a continuous target) with two prediction models.}
 	\label{tab:regression}
-  \begin{tabular}{
-    >{\raggedleft\arraybackslash}m{0.33855\textwidth}
-    >{\centering\arraybackslash}m{0.01\textwidth}
-    >{\centering\arraybackslash}m{0.2562\textwidth}
-    >{\centering\arraybackslash}m{0.2562\textwidth}
-  }
+	\begin{tabular}{lcc}
 		\toprule
-    {Regression Models} & & {Avg. RMSE ($\pm$ std)} & {Avg. MAE ($\pm$ std)} \\
+		{Regression Models} & {Avg. RMSE ($\pm$ std)} & {Avg. MAE ($\pm$ std)} \\
 		\midrule
-    Random Forest & & \SI{1929.84}{s} ($\pm$ \SI{1161.15}{s}) & \phantom{0}\SI{957.36}{s} ($\pm$ \SI{297.21}{s}) \\[0.4ex]
-    Multilayer Perceptron & & \SI{2275.20}{s} ($\pm$ \SI{1545.01}{s}) & \SI{1284.30}{s} ($\pm$ \SI{390.57}{s}) \\
+		Random Forest & \SI{1929.84}{s} ($\pm$ \SI{1161.15}{s}) & \phantom{0}\SI{957.36}{s} ($\pm$ \SI{297.21}{s}) \\[0.4ex]
+		Multilayer Perceptron & \SI{2275.20}{s} ($\pm$ \SI{1545.01}{s}) & \SI{1284.30}{s} ($\pm$ \SI{390.57}{s}) \\
 		\bottomrule
@@
 
-Table~\ref{tab:timeout-prediction} shows the performance of different models in predicting timeouts.
-While the second column shows the performance of models trained on the 46 instance features, the third column shows the performance of models trained on the 46 instance features and the 27 runtime features.
-While the Quadratic DA model performs poorly on the 46 instance features, it performs well on the 46 instance features and the 27 runtime features.
-However, the best prediction performance is achieved by a stacking ensemble~\cite{wolpert1992stacked}, which combines quadratic discriminant analysis~\cite{tharwat2016linear} and a random forest classifier~\cite{breiman2001random}.
-Stacking means that another prediction model, in our case a simple decision tree, decides which of the two ensemble members makes the prediction about which instance.
-
-\begin{table}[tbp]
-  \centering
-  \caption{MCC performance of models predicting timeouts. The models are trained 28 times with each solver as a target once.}
+As a baseline, Table~\ref{tab:regression} shows the prediction performance of two regression models, i.e., a random forest and a multilayer perceptron.
+We employ these models to directly estimate a new solver's runtime as the prediction target.
+To this end, we use 46 instance features and 27 known solver runtimes as inputs for the predictions.
+We repeat this evaluation for each solver as prediction target and report the average and standard deviation of prediction performance over the 28 solvers.
+
+We chose both types of prediction models for their robustness and ability to capture nonlinear dependencies~\cite{breiman2001random}.
+However, both still perform poorly, as evidenced by the magnitudes of the mean regression errors in Table~\ref{tab:regression}.
+In particular, a root mean squared error (RMSE) of about \SI{2000}{s} is quite high for solver runtimes that range from 0 to \SI{5000}{s}.
+Therefore, we decided to discretize the runtimes for prediction, moving from a regression scenario to a classification scenario.
+
+\paragraph{Classifying timeouts}
+
+\begin{table}[tb]
+	\centering
+	\caption{Test-set prediction performance for predicting solver timeouts (a binary target) with different prediction models.}
 	\label{tab:timeout-prediction}
 	~\\[1em]
-  \begin{tabular}{
-    >{\raggedleft\arraybackslash}m{0.4\textwidth}
-    >{\centering\arraybackslash}m{0.25\textwidth}
-    >{\centering\arraybackslash}m{0.25\textwidth}
-  }
+	\begin{tabular}{lcc}
 		\toprule
-    {Timeout Prediction Models} & {46 Instance Features Avg. MCC ($\pm$ std)} & {46\,+\,27 Runtime Feat. Avg. MCC ($\pm$ std)} \\
+		\multirow[c]{2}{*}{Timeout Prediction Models} & \multicolumn{2}{c}{Average MCC ($\pm$ std)} \\
+		\cmidrule(lr){2-3}
+		& {Without Runtime Feat.} & {With Runtime Feat.} \\
 		\midrule
@@
 
-% MI: table with cluster prediction performance: (merged with existing text)
-Since runtime prediction is quite challenging, we discretize the runtimes.
-To define prediction targets for each instance, we cluster the solver runtimes into $k$ clusters.
-We denote this adapted solver scoring function which assigns a solver to its cluster for each instance with $s_k$ as defined in Equation~\eqref{eq:rankingeq}.
-In preliminary experiments, the most \emph{useful} labels were produced with hierarchical clustering and a log-single-link criterion.
-In our chosen hierarchical procedure, each non-timeout runtime starts in a separate interval.
-We then gradually merge intervals with the smallest single-link logarithmic distance until the desired number of partitions is reached.
-The clustering is instance-specific, i.e., the bounds are chosen to produce groups of solvers with similar performance for each instance.
-Other clustering approaches we have tried include hierarchical clustering with mean, median, and full-link criteria, as well as $k$-means and spectral clustering.
-
-We evaluated the performance of the best models from Table~\ref{tab:timeout-prediction} in predicting the such determined $s_k$.
-Table~\ref{tab:cluster-prediction} shows their performance in predicting the $C_k$ clusters for $k \in \lbrace 3, 4, 5 \rbrace$.
-While the second column shows the prediction performance of the models trained on the 46 instance features including the 27 known solver runtimes, the third column shows the prediction performance of the models trained on these features plus 27 features representing the known $C_k$ clustering labels of the known solver runtimes.
-We observe that the models perform better when the clustering labels are included as features.
-In addition, the prediction performance degrades rapidly as the number of clusters increases.
-In all cases, the stacking ensemble performs best.
-
-\begin{table}[tbp]
-  \centering
-  \caption{MCC performance of the best models predicting the $C_k$ clusterings.}
+Table~\ref{tab:timeout-prediction} shows the prediction performance for estimating whether a solver times out or not.
+This binary prediction target represents a very strong discretization of runtimes.
+We compare a variety of prediction models, which pursue different learning paradigms, each of them trained with either instance features only (second column) or instance and runtime features (third column).
+As the evaluation metric for prediction performance, we use Matthews Correlation Coefficient (MCC)~\cite{gorodkin2004comparing, matthews1975comparison}, which reaches its maximum of~1 for a perfect prediction, is~0 for random guessing, and has a minimum of~-1.
+
+Training the classification models with instance features only, a random forest performs best.
+In contrast, a quadratic discriminant analysis performs poorly in this scenario but performs very well if runtime features are included.
+However, the best prediction performance with all features is achieved by a stacking ensemble~\cite{wolpert1992stacked} out of the quadratic discriminant analysis~\cite{tharwat2016linear} and the random forest~\cite{breiman2001random}.
+Stacking means that another model, in our case a simple decision tree~\cite{breiman1984classification}, makes a prediction based on the predictions of the ensembled models.
+
+\paragraph{Classifying discretized runtime labels}
+
+Building on the success of binary timeout prediction, we investigate a classification scenario where the target is solver runtime discretized into multiple categories.
+
+\subparagraph{Label definition}
+
+To define this prediction target, we partition the solver runtimes into $k$~clusters.
+We conduct this clustering step per instance, i.e., the solvers forming a particular cluster may differ from instance to instance.
+This clustering forms the basis of the adapted solver scoring function~$s_k$ (cf.~Equation~\eqref{eq:rankingeq}).
+In preliminary experiments, the most \emph{useful} labels were produced with hierarchical clustering and a log-single-linkage criterion; we will discuss our notion of labels usefulness later.
+Other clustering approaches we have tried include hierarchical clustering with mean-, median-, and full-linkage criteria, as well as $k$-means and spectral clustering.
+%JB: What is "full" -- "complete"?
+
+In our chosen hierarchical clustering procedure, each non-timeout runtime starts in a separate cluster.
+We then gradually merge the closest clusters until the desired number of $k-1$ partitions is reached.
+By definition, the $k$-th cluster always contains the timeouts and no other runtimes (cf.~Section~\ref{sec:main:model:transformation}).
+To quantify the distance between two clusters, we consider the minimum pairwise difference of logarithmic runtimes between solvers in the two clusters.
+
+\subparagraph{Label prediction}
+
+\begin{table}[tb]
+  \centering
+  \caption{Test-set prediction performance for predicting the $C_k$ cluster labels (a discrete target) with the best performing models.}
   \label{tab:cluster-prediction}
-  \begin{tabular}{
-    >{\raggedleft\arraybackslash}m{0.4\textwidth}
-    >{\centering\arraybackslash}m{0.25\textwidth}
-    >{\centering\arraybackslash}m{0.25\textwidth}
-  }
+  \begin{tabular}{lcc}
     \toprule
-    {$C_3$ Cluster Prediction Models} & {46\,+\,27 Runtime Feat. Avg. MCC ($\pm$ std)} & {46\,+\,27\,+\,27 Clus. Feat. Avg. MCC ($\pm$ std)} \\
+    \multirow[c]{2}{*}{$C_3$ Cluster Prediction Models} & \multicolumn{2}{c}{Average MCC ($\pm$ std)} \\
+    \cmidrule(lr){2-3}
+    & {Without Clustering Feat.} & {With Clustering Feat.} \\
     \midrule
@@
     \midrule
-    {$C_4$ Cluster Prediction Models} & {46\,+\,27 Runtime Feat. Avg. MCC ($\pm$ std)} & {46\,+\,27\,+\,27 Clus. Feat. Avg. MCC ($\pm$ std)} \\
+    \multirow[c]{2}{*}{$C_4$ Cluster Prediction Models} & \multicolumn{2}{c}{Average MCC ($\pm$ std)} \\
+    \cmidrule(lr){2-3}
+    & {Without Clustering Feat.} & {With Clustering Feat.} \\
     \midrule
@@
     \midrule
-    {$C_5$ Cluster Prediction Models} & {46\,+\,27 Runtime Feat. Avg. MCC ($\pm$ std)} & {46\,+\,27\,+\,27 Clus. Feat. Avg. MCC ($\pm$ std)} \\
+    \multirow[c]{2}{*}{$C_5$ Cluster Prediction Models} & \multicolumn{2}{c}{Average MCC ($\pm$ std)} \\
+    \cmidrule(lr){2-3}
+    & {Without Clustering Feat.} & {With Clustering Feat.} \\
     \midrule
@@
 
-Our final choice for the prediction model is the stacking ensemble of a quadratic discriminant analysis and a random forest.
-Both of these models can learn non-linear relationships between the instance features and the runtime labels.
-Fig.~\ref{fig:confusion-matrices} shows the confusion matrices of the stacking ensemble for predicting timeouts and $C_3$ clusters.
-Both perform reasonably well, with the $C_3$ cluster prediction model performing slightly worse than the timeout prediction model.
-Notably, for the $C_3$ prediction model, there is less confusion between the non-timeout labels (labels 1 and 2) and the timeout label (label 3) than between the non-timeout labels (labels 1 and 2).
-This suggests that deciding on timeouts is easier than distinguishing between different groups of solvers.
-
-\begin{figure}[tbp]
+Given this discretization procedure, we evaluate the performance of the best classification models from Table~\ref{tab:timeout-prediction} in predicting these runtime labels.
+Table~\ref{tab:cluster-prediction} shows the prediction performance for $k \in \lbrace 3, 4, 5 \rbrace$ clusters.
+We train the prediction models either with 46 instance features and 27 known solver runtimes (second column) or additionally including 27 features representing the known $C_k$ cluster labels of other solvers (third column).
+For all prediction models and values of~$k$, the prediction performance is better when the cluster labels are included as features.
+In addition, the prediction performance degrades rapidly as the number of clusters increases.
+In all cases, the stacking ensemble performs best, so we use it in all subsequent experiments.
+
+\begin{figure}[tb]
 	\centering
@@
 	\end{subfigure}
-  \caption{Confusion matrices for predicting timeouts and $C_3$ clusters.}
+	\caption{Confusion matrices for predicting timeouts and $C_3$ cluster labels.}
 	\label{fig:confusion-matrices}
@@
 
-To ensure that we obtained \emph{useful} labels, we evaluated whether the discretized labels still discriminate solvers and agree with the actual PAR-2 ranking.
-Table~\ref{tab:ranking} shows the ranking of the solvers in the SAT Competition~2022 Anniversary Track~\cite{sat2022} according to the standard PAR-2 runtime score and according to our discretized $s_3$ score.
-We observe that our ranking approach correctly decides almost all (about \SI{97.45}{\%}; $\sigma = \SI{3.68}{\%}$) solver pairs, i.e., which solver is faster among two given solvers.
-In particular, the Spearman correlation of $s_3$ and PAR-2 ranking is about \SI{0.988}{}, which is very close to the optimal value of~1~\cite{de2016comparing}.
+To analyze the stacking ensemble in more depth, Fig.~\ref{fig:confusion-matrices} shows the confusion matrices for predicting timeouts and $C_3$ cluster labels.
+Both scenarios admit relatively good prediction performance, with the model for $C_3$ performing slightly worse than the model for timeouts.
+Notably, for $C_3$, there is less confusion between the non-timeout labels (Labels~1 and~2) and the timeout label (Label~3) than between the non-timeout labels (Labels~1 and~2).
+This observation suggests that deciding on timeouts is easier than distinguishing between discretized non-timeout runtimes.
+
+\subparagraph{Label usefulness}
 
@@
 % * Spearman correlation of PAR2 and Discrete rank: 0.9862
-\begin{table}[tbp]
-  \centering
-  \caption{Ranking of Anniversary-Track solvers with standard PAR-2 runtime and our discretized $s_3$ ranking.}
+\begin{table}[tb]
+	\centering
+	\caption{Ranking of solvers on the Anniversary Track for the PAR-2 score and our discretized $s_3$ score.}
 	\label{tab:ranking}
@@
 
-We also analyzed how runtime discretization affects the distinguishability of solver pairs.
-Fig.~\ref{fig:solver-confusion-matrices} shows the confusion matrices of solver pairs with significant differences in their PAR-2 scores, timeouts, and $s_3$ scores.
-According to a Wilcoxon signed-rank test with a level of $\alpha = 0.05$, \SI{87.83}{\%} of the solver pairs have significantly different scores after discretization, only a small decrease compared to \SI{89.95}{\%} before discretization.
-These results support the usefulness of discretized runtimes for our framework.
+To ensure that the runtime labels cannot only be predicted well but are \emph{useful} for active learning, we evaluate whether the labels still discriminate solvers and agree with the actual PAR-2 ranking.
+Table~\ref{tab:ranking} shows the ranking of the solvers in the SAT Competition~2022 Anniversary Track~\cite{sat2022} according to the standard PAR-2 score and according to our discretized $s_3$ score.
+We observe that ranking with our score correctly decides for almost all (about \SI{97.45}{\%}; $\sigma = \SI{3.68}{\%}$) solver pairs which solver is faster.
+In particular, the Spearman correlation of $s_3$ and the PAR-2 score over all solvers is about \SI{0.988}{}, which is very close to the optimal value of~1~\cite{de2016comparing}.
 
@@
 % k=5:     662/756 = 0.8757
-\begin{figure}[tbp]
+\begin{figure}[tb]
 	\centering
@@
 	\end{subfigure}
-  \caption{Solver pairs with significant differences regarding the respective scorings are colored white.
-Non-significant differences are colored black.
-Solvers are ordered indentically to Tab.~\ref{tab:ranking}.}
+	\caption{Solver pairs with significant differences regarding their scorings are colored white, non-significant ones black. Solvers are ordered indentically to Table~\ref{tab:ranking}.}
 	\label{fig:solver-confusion-matrices}
 \end{figure}
+
+We also analyze how runtime discretization affects the distinguishability of solver pairs.
+The matrices in Fig.~\ref{fig:solver-confusion-matrices} visualize which solver pairs exhibit significant differences in their PAR-2 scores, timeouts, and $s_3$ scores.
+According to a Wilcoxon signed-rank test with a significance level of $\alpha = 0.05$, \SI{87.83}{\%} of the solver pairs have significantly different scores after discretization, which is only a small decrease compared to \SI{89.95}{\%} before discretization.
+%JB: numbers in notebook "bonus_results" differ
+%JB: does "after discretization" mean s_3 or timeout? and what is then the number for the other one?
+These results support the usefulness of discretized runtimes for our framework.
 
@@
 Fig.~\ref{fig:instoccs} illustrates the distribution of the frequency with which individual instances are selected across all active learning benchmarks.
+%JB: on the full dataset (complete Anniversary Track), I suppose? and what is "all" aggregating over? these information may also go into the figure's caption
 It can be observed that a small proportion of instances is selected with great frequency, while most instances are selected on only a few occasions.
 This indicates that there is no fixed subset of instances that can perfectly distinguish all solvers, which highlights the value of employing an active-learning-based strategy.
-
-The selection decisions of our approach also reveal the importance of different instance families to our framework.
+%JB: could also indicate that (1) there is such a set but AL cannot always find it or (2) there are multiple such sets and different AL runs return different sets
+
+The selection decisions of our approach also reveal the importance of instance families to our framework.
 Fig.~\ref{fig:annifinalfamilies} illustrates the occurrence of instance families within the dataset and the benchmarks created by active learning.

### 2024-05-26 paper_jar/paper.tex
Revise Abstract and Intro
Small clarifications and extensions

@@
 
-\author*{\fnm{Tobias} \sur{Fuchs}}\email{tobias.fuchs@kit.edu}
-\author{\fnm{Jakob} \sur{Bach}}\email{jakob.bach@kit.edu}
-\author{\fnm{Markus} \sur{Iser}}\email{markus.iser@kit.edu}
-
-\affil{\orgname{Karlsruhe Institute of Technology (KIT)}, \orgaddress{\city{Karlsruhe}, \country{Germany}}}
+\author*[1]{\fnm{Tobias} \sur{Fuchs}}\email{tobias.fuchs@kit.edu}
+\author[1]{\fnm{Jakob} \sur{Bach}}\email{jakob.bach@kit.edu}
+\author[1]{\fnm{Markus} \sur{Iser}}\email{markus.iser@kit.edu}
+
+\affil[1]{\orgname{Karlsruhe Institute of Technology (KIT)}, \orgaddress{\city{Karlsruhe}, \country{Germany}}}
 
 \abstract{
-Benchmarking is a crucial phase in the development of algorithms.
+Benchmarking is crucial for developing new algorithms.
 This also applies to solvers for the propositional satisfiability (SAT) problem.
@@
 In this paper, we present a dynamic benchmark selection approach based on active learning.
-Our approach predicts the rank of a new solver among its competitors with minimum runtime and maximum rank prediction accuracy.
+Our approach estimates the rank of a new solver among its competitors, striving to minimize benchmarking runtime but maximize rank prediction accuracy.
+Instead of using real-valued solver runtimes, our approach works with discretized runtime labels, which yielded better solver rank predictions.
 We evaluated this approach on the Anniversary Track dataset from the SAT Competition 2022.
-Our selection approach can predict the rank of a new solver after approximately \SI{10}{\%} of the time it would take to run the solver on all instances of this dataset, with a prediction accuracy of approximately \SI{92}{\%}.
+Our benchmark selection approach can predict the rank of a new solver after approximately \SI{10}{\%} of the time it would take to run the solver on all instances of this dataset, with a prediction accuracy of approximately \SI{92}{\%}.
 Additionally, we discuss the importance of instance families in the selection process.
@@
 
-\keywords{Propositional satisfiability, Benchmarking, Active learning}
+\keywords{Propositional satisfiability, Benchmark selection, Active learning}
 
@@
 One of the main phases of algorithm engineering is benchmarking.
-This also applies to propositional satisfiability (SAT), the canonical $\mathcal{NP}$-complete problem.
+This also applies to solvers for propositional satisfiability (SAT), the canonical $\mathcal{NP}$-complete problem.
 Benchmarking is, however, quite expensive regarding the runtime of experiments.
-While it is still feasible to benchmark a single or small number of SAT solvers, developing new, competitive SAT solvers requires extensive experimentation with a variety of ideas.
-The latter often results in a combinatorial explosion of the configuration space of each algorithm~\cite{HutterHL11}.
+While it is still feasible to benchmark a single or a small number of given SAT solvers, developing new, competitive SAT solvers requires extensive experimentation with a variety of ideas.
+The latter often results in a combinatorial explosion of the configuration space~\cite{HutterHL11}.
 In particular, a new solver idea is rarely best on the first try.
@@
 Instead, they focus on selecting a representative set of instances for ranking solvers~\cite{Gelder11,HoosKSS13}.
-In this regard, SAT Competitions typically employ the \mbox{PAR-2} score, which, given a time limit of $\tau$, is the average runtime with a penalty of $2 \tau$ for timeouts~\cite{FroleyksHIJS21}.
+In this regard, SAT Competitions typically employ the \mbox{PAR-2} score, which, given a time limit of $\tau$, is the average solver runtime with a penalty of $2 \tau$ for timeouts~\cite{FroleyksHIJS21}.
 
@@
 \begin{definition}[New-Solver Problem]
-  Let us consider a set of solvers~$\mathcal{A}$, instances~$\mathcal{I}$, and runtimes~${r\!: \mathcal{A} \times \mathcal{I} \rightarrow \left[0, \tau\right]}$ with each solver having a runtime limit of $\tau$.
-  For a new solver $\hat{a} \notin \mathcal{A}$, incrementally select benchmark instances from $\mathcal{I}$ to maximize the confidence in predicting the rank of $\hat{a}$ while minimizing the total benchmark runtime.
+  Let us consider a set of solvers~$\mathcal{A}$, instances~$\mathcal{I}$, and runtimes~${r\!: \mathcal{A} \times \mathcal{I} \rightarrow \left[0, \tau\right]}$, with each solver having a runtime limit of $\tau$.
+  The \emph{New-Solver Problem} is about incrementally selecting benchmark instances from $\mathcal{I}$ to maximize the confidence in predicting the rank of a new solver $\hat{a} \notin \mathcal{A}$, while minimizing the total benchmark runtime.
   \label{def:new-solver-problem}
@@
 The approach presented in this paper meets several criteria for benchmarking that are considered desirable (cf.~Table~\ref{tab:requirements}).
-Rather than outputting a binary classification, namely whether the new solver is worse than an existing solver or not, we provide a \emph{scoring} function that shows by which margin a solver is worse and how similar it is to existing solvers.
-In particular, our approach enables \emph{ranking} the new solver amidst a set of existing solvers.
+Rather than outputting a binary classification, namely whether the new solver is worse than an existing solver or not, we provide a \emph{scoring} function that shows by which margin a solver is worse and how similar its performance is to existing solvers.
+In particular, our approach enables \emph{ranking} the new solver among a set of existing solvers.
 We show that for ranking, it is not necessary to predict the exact per-instance runtime of the solvers, which would be a more challenging task.
-Furthermore, we optimize the \emph{runtime} required for our strategy to arrive at its conclusion.
-To this end, we use instance \emph{features} and known solver \emph{runtimes}.
-Moreover, we select instances \emph{non-randomly} and \emph{incrementally}.
-In particular, we consider runtime information from already done experiments when choosing the next.
+Instead, we work with discretized runtime labels.
+We use instance features and known solver runtimes for predicting these labels.
+Furthermore, we minimize the \emph{runtime} required for our approach to arrive at its conclusion.
+Moreover, we select benchmark instances \emph{non-randomly} and \emph{incrementally}.
+In particular, we consider runtime information from already done experiments when choosing the next instance.
 By doing so, we can control the properties of the benchmarking approach, such as its required runtime.
@@
 We evaluated our approach using the SAT Competition~2022 Anniversary Track dataset~\cite{sat2022}, consisting of 5355~instances and runtimes of 28~solvers.
-Cross-validation was performed by treating each solver once as the new solver and learning to predict the PAR-2 rank of that solver.
-On average, our predictions achieve an accuracy of approximately \SI{92}{\%} with only approximately \SI{10}{\%} of the runtime required to evaluate these solvers on the complete set of instances.
+Cross-validation was performed by treating each solver as the new solver once and learning to predict the PAR-2 rank of that solver.
+On average, our predictions achieve a ranking accuracy of approximately \SI{92}{\%} with only approximately \SI{10}{\%} of the runtime required to evaluate these solvers on the complete set of instances.
 Our entire source code\footnote{\url{https://github.com/mathefuchs/al-for-sat-solver-benchmarking}} and experimental data\footnote{\url{https://github.com/mathefuchs/al-for-sat-solver-benchmarking-data}} are available on GitHub.
+%JB: data repo not updated since TACAS; does it also contain all the results for the new experiments? as a minimum, it should mention the journal version of the paper (see README of code repo)
+%JB: code repo currently has all JAR changes in a branch; probably best to push these changes to "main" for review (in case reviewers have a look at repo but don't switch branches)
 
@@
 Section~\ref{sec:related} reviews related work.
-Section~\ref{sec:main} describes our approach.
+Section~\ref{sec:main} describes our active learning approach.
 Section~\ref{sec:exdesign} introduces the experimental design.
@@
 This article is an extended and refined version of our conference paper with the title ``Active Learning for SAT Solver Benchmarking''~\cite{fuchs2023active}.
-In particular, we significantly extended the analysis of runtime-prediction approaches, which determined our chosen solver runtime model for instance selection (cf.~Section~\ref{sec:exdesign:disc-pred}).
+In particular, we significantly extended the analysis of runtime-prediction and runtime-discretization approaches (cf.~Section~\ref{sec:exdesign:disc-pred}), which determined our chosen solver runtime model for instance selection.
 Furthermore, we expanded the evaluation of which benchmark instances are selected by our approach (cf.~Section~\ref{sec:eval:instance}).

### 2024-05-26 paper_jar/paper.tex
Revise Related Work

@@
 
-The following sections present a discussion of related work from the areas of static benchmark selection (cf.~Section~\ref{sec:related:static-selection}), algorithm configuration (cf.~Section~\ref{sec:related:algo-configuration}), incremental benchmark selection (cf.~Section~\ref{sec:related:incremental-selection}), and active learning (cf.~Section~\ref{sec:related:active-learning}).
+The following sections discusses related work from the areas of static benchmark selection (cf.~Section~\ref{sec:related:static-selection}), algorithm configuration (cf.~Section~\ref{sec:related:algo-configuration}), incremental benchmark selection (cf.~Section~\ref{sec:related:incremental-selection}), and active learning (cf.~Section~\ref{sec:related:active-learning}).
 Table~\ref{tab:requirements} compares the most relevant approaches, which all pursue slightly different goals.
-Thus, our approach is \emph{not} a general improvement over the others but the only one fully aligned with Definition~\ref{def:new-solver-problem}.
+In particular, our approach is \emph{not} a general improvement over the others but the only one fully aligned with Definition~\ref{def:new-solver-problem}.
 
@@
 
-Benchmark selection is essential for competitions such as the SAT Competition.
+Benchmark selection is essential for solver competitions like the SAT Competition.
 In such competitions, the organizers define the rules for composing the benchmarks.
@@
 Froleyks et al. describe benchmark selection in recent SAT~competitions~\cite{FroleyksHIJS21}.
-Manthey and Möhle propose an approach to remove redundancy from competition benchmarks~\cite{manthey2016better}, considering feature equivalence of formulas.
-M{\i}s{\i}r also presents a feature-based approach to reduce benchmarks, using matrix factorization and clustering~\cite{misir2021benchmark}.
+Manthey and Möhle propose an approach to remove redundancy from competition benchmarks by considering feature equivalence of formulas~\cite{manthey2016better}.
+M{\i}s{\i}r also presents a feature-based approach to reduce benchmarks by using matrix factorization and clustering~\cite{misir2021benchmark}.
 
@@
 For example, SMAC employs an iterative, model-based selection procedure similar to our approach, but this is for selecting configurations rather than instances.
-However, an algorithm configurator cannot be used to \emph{rank} or \emph{score} a new solver since algorithm configuration is designed to identify the optimal configuration.
-Furthermore, while a model-based selection strategy is employed to sample configurations, instance selection is made \emph{randomly}, i.e., without building a model over instances.
+In contrast, instance selection is made \emph{randomly} in SMAC, i.e., without building a model over instances.
+Furthermore, an algorithm configurator cannot be used to \emph{rank} or \emph{score} a new solver relative to others since algorithm configuration is designed to identify the optimal configuration.
 
@@
 Their \emph{per-set efficient algorithm selection problem} (PSEAS) is similar to our \emph{New-Solver Problem}, as given in Definition~\ref{def:new-solver-problem}.
-To differentiate between \emph{a pair of} SAT solvers, the authors propose an iterative approach to select a solver-specific subset of instances until a desired confidence level is reached.
+To compare \emph{a pair of} SAT solvers, the authors propose an iterative approach to select a solver-specific subset of instances until a desired confidence level is reached.
 This is achieved by calculating a scoring metric for all unselected instances, running the experiment with the highest score, and updating the confidence.
 Their approach ticks off most of the desired features in Table~\ref{tab:requirements}.
-However, the approach only compares solvers binarily rather than providing a \emph{scoring}.
+However, the approach only compares solvers binarily rather than providing a multi-solver \emph{scoring} or \emph{ranking}.
 Consequently, it is unclear how similar two given solvers are or on which instances they behave similarly.
 Furthermore, a significant shortcoming is the lack of \emph{scalability} with the number of solvers.
-Comparing only pairs of solvers, evaluating a new solver with their method requires sampling a separate benchmark for each existing solver.
+Due to comparing only pairs of solvers, evaluating a new solver with their method requires sampling a separate benchmark for each existing solver.
 In contrast, our approach allows comparing a new solver against a set of existing solvers by sampling only one benchmark.
@@
 
+Our benchmark selection approach uses active learning.
 In passive machine learning, prediction models are trained on datasets with given instance labels (cf.~Fig.~\ref{fig:passive}).
@@
 Without AL methods, it is not evident which instances to label and which not.
-The first goal is to maximize the utility of instances to our model, which is measured by the rank prediction accuracy.
-The second goal is to minimize the cost of label acquisition, which is measured by the instance's predicted runtime.
-Consequently, the overall objective is to develop an accurate prediction model without the necessity of labeling each data point.
+In our scenario, the first goal is to maximize the utility of instances to our prediction model, which is measured by the rank prediction accuracy.
+Our second goal is to minimize the cost of label acquisition, which is measured by the instance's predicted runtime.
+Consequently, our overall objective is to develop an accurate prediction model without the necessity of labeling each data point.
 
@@
 While those approaches aim to generate instances on which a solver is either effective or ineffective, our objective is to predict whether a solver is effective or ineffective on an \emph{existing} benchmark.
-In their work, Volpato and Guangyan employ pool-based AL to train an instance-specific algorithm selector~\cite{volpato2019active}.
-Rather than benchmarking a solver's overall performance, their objective is to recommend the best solver for each SAT instance from a set of available solvers.
+In the latter direction, Volpato and Guangyan employ pool-based AL to train an instance-specific algorithm selector~\cite{volpato2019active}.
+However, their objective is not benchmarking a solver's overall performance but recommending the best solver for each SAT instance from a set of available solvers.
 

### 2024-05-26 paper_jar/paper.tex
Revise Approach

@@
 Algorithm~\ref{algALBenchmark} outlines our benchmarking framework. 
-Given a set of solvers~$\mathcal{A}$, instances~$\mathcal{I}$ and runtimes~$r$, we first initialize a prediction model~$\mathcal{M}$ for the new solver $\hat a \not\in \mathcal{A}$ (Line~1).
+Given a set of solvers~$\mathcal{A}$, instances~$\mathcal{I}$, and runtimes~$r$, we first initialize a prediction model~$\mathcal{M}$ for the new solver $\hat a \not\in \mathcal{A}$ (Line~1).
 The prediction model~$\mathcal{M}$ is used to select instances repeatedly (Line~4) for benchmarking the new solver $\hat a$ (Line~5). 
-The acquired result is subsequently employed to update the prediction model, denoted by~$\mathcal{M}$ (Line~7). 
-When the stopping criterion is met (Line~3), the benchmarking loop is terminated and the final score of $\hat{a}$ is estimated (Line~8). 
-Algorithm~\ref{algALBenchmark} returns this score, along with the acquired instances and runtime measurements.
-
-Section~\ref{sec:main:model} provides an overview of the underlying prediction model~$\mathcal{M}$ and outlines the methodology for deriving a solver ranking from it.
+The acquired result is subsequently employed to update the prediction model (Line~7). 
+When the stopping criterion is met (Line~3), the benchmarking loop terminates and the final score of $\hat{a}$ is estimated (Line~8). 
+Algorithm~\ref{algALBenchmark} returns this score, along with the acquired benchmark instances and runtime measurements.
+
+Section~\ref{sec:main:model} describes the underlying prediction model~$\mathcal{M}$ and outlines the methodology for deriving a solver ranking from it.
 Section~\ref{sec:main:selection} presents the criteria for selecting instances.
@@
   $\mathcal{M} \leftarrow \operatorname{initModel}\left(\mathcal{A},\, \mathcal{I},\, r,\, \hat{a}\right)$ \tcp*{cf. Section~\ref{sec:main:model}}
+  %JB: do we actually need \hat{a} here? (does not provide any information we could store/use here)
 
@@
 
-The model $M$ provides a runtime-label prediction function $f : \mathcal{\hat A} \times \mathcal{I} \rightarrow \mathbb{R}$ for all solvers $\mathcal{\hat A} := \mathcal{A} \cup \lbrace \hat a \rbrace$.
+The model $M$ provides a prediction function $f : \mathcal{\hat A} \times \mathcal{I} \rightarrow \mathbb{R}$ for all solvers $\mathcal{\hat A} := \mathcal{A} \cup \lbrace \hat a \rbrace$.
+%JB: shouldn't we drop the algorithm as an argument (for $f$ and later for $f'$), since we only train and predict for one solver ($\hat{a}$) anyway?
+(what would the predictions for the other solvers be -- looking up the exact runtime values from $r$?); instead, could use $\hat{a}$ as a subscript of $f$
 This prediction function powers instance selection as described in Section~\ref{sec:main:selection}.
-During model updates (Algorithm~\ref{algALBenchmark},~Line~7), the prediction function $f$ is trained to predict a transformed version of the acquired runtime dataset~$\mathcal{R}$.
-We describe the transformation of the runtime dataset in the subsequent section.
-The features described in Section~\ref{sec:exdesign:data} serve as the input to the model.
-Furthermore, it should be noted that a new prediction model is constructed in each iteration, as the runtime of the experiments (Line~5) dominates the training time of the model by a significant margin.
-Finally, the score of the new solver, denoted by $\hat a$, is predicted with the prediction function~$f$ (Line~8).
+In the update step (Algorithm~\ref{algALBenchmark},~Line~7), the model is trained to predict a transformed version of solver runtime (cf.~Section~\ref{sec:main:model:transformation}) for the new solver~$\hat{a}$, using the features described in Section~\ref{sec:exdesign:data} and the previously acquired runtimes~$\mathcal{R}$.
+Each iteration trains a new prediction model from scratch, since the time for running the solver~$\hat{a}$ on one instance~$e$ (Line~5) dominates the training time of the model~$M$ on all acquired instances~$\mathcal{R}$ by a significant margin.
+Once the benchmarking loop terminates, the score of the new solver~$\hat{a}$ is predicted and then returned (Lines~8--9).
 
@@
 %
-In the context of such a clustering, for an instance $e \in \mathcal{I}$, a solver $a \in \mathcal{A}$ is defined as belonging to the $\gamma_k(a, e)$-fastest solvers on instance $e$. 
-%
 Empirical studies on portfolio solvers have demonstrated that discretization is an effective approach in practice~\cite{CollauttiMMO13,NgokoCT19}.
-Section~\ref{sec:exdesign:disc-pred} presents a detailed analysis of the runtime transformation we designed for our approach.
-The analysis demonstrates that our approach achieves higher accuracy for discrete runtime labels than for raw runtimes.
+Section~\ref{sec:exdesign:disc-pred} presents a detailed analysis of the runtime transformation in our scenario.
+In particular, the analysis demonstrates that our benchmarking approach achieves higher rank prediction accuracy when working with discrete runtime labels rather than raw runtimes.
 
@@
 
-To determine the rank of solvers, we apply PAR-2 scoring using the discrete runtime labels $\gamma_k(a, e)$.
-The adapted scoring function $s_k : \mathcal{A} \rightarrow [1, 2 \cdot k]$ induces a ranking among solvers and is depicted in Equation~\eqref{eq:rankingeq}.
+To determine the rank of solvers, we apply PAR-2 scoring to the discrete runtime labels $\gamma_k(a, e)$ and obtain the adapted scoring function $s_k : \mathcal{A} \rightarrow [1, 2 \cdot k]$:
 %
@@
 %
+In particular, timeouts are still penalized.
+
 \subsection{Instance Selection}
@@
 
-Selecting an instance based on the model is a core functionality of our framework (cf.~Algorithm~\ref{algALBenchmark}, Line~4).
+Selecting an instance based on the prediction model is a core functionality of our framework (cf.~Algorithm~\ref{algALBenchmark}, Line~4).
 In this section, we introduce two instance sampling strategies, one that minimizes uncertainty and one that maximizes information gain.
-Both strategies use the model's label-prediction function $f$ and are inspired by existing work within the realms of active learning~\cite{settles2009active}.
+Both strategies use the model's prediction function~$f$ and are inspired by existing work on active learning~\cite{settles2009active}.
 These methods require the model's predictions to include probabilities for the $k$ discrete runtime labels.
@@
 The uncertainty sampling strategy selects the instance that is closest to the model's decision boundary.
-This is achieved by selecting the instance $e \in \mathcal{I} \setminus \tilde{\mathcal{I}}$ that minimizes $U(e)$, which is specified by the following equation.
+This is achieved by selecting the instance $e \in \mathcal{I} \setminus \tilde{\mathcal{I}}$ that minimizes the certainty~$U(e)$, which is specified by the following equation:
 %
@@
 
-The information-gain sampling strategy selects the instance with the highest expected entropy reduction regarding the runtime labels of the instance.
-To be more precise, we select the instance $e \in \mathcal{I} \setminus \tilde{\mathcal{I}}$ that maximizes $IG(e)$, which is specified in the following equation.
+The information-gain sampling strategy selects the instance with the highest expected entropy reduction regarding the runtime labels.
+To be more precise, we select the instance $e \in \mathcal{I} \setminus \tilde{\mathcal{I}}$ that maximizes $IG(e)$, which is specified in the following equation:
 %
@@
 By maximizing the information gain, instances are selected that identify solvers with similar behavior.
+%JB: "similar behavior" is a bit vague; strategy favors instances where the distribution of solvers over clusters is skewed rather than uniform, right?
+and the new solver only has little influence here, since its predicted labels is only one of |A| + 1 labels to determine entropy, right?
 
@@
 
-In this section, we present the two dynamic stopping criteria employed in our experiments: the Wilcoxon and the ranking stopping criterion (cf. Algorithm~\ref{algALBenchmark}, Line~3).
+In this section, we present the two dynamic stopping criteria (cf.~Algorithm~\ref{algALBenchmark}, Line~3) employed in our experiments: the Wilcoxon and the ranking stopping criterion.
 
@@
 
-The Wilcoxon stopping criterion is employed to halt the active-learning process based on the confidence that the predicted runtime labels of the new solver are sufficiently different from those of existing solvers.
+The Wilcoxon stopping criterion terminates the active-learning process based on the confidence that the predicted runtime labels of the new solver are sufficiently different from those of existing solvers.
 This criterion is loosely inspired by Matricon et.~al.~\cite{MatriconAFSH21}.
-The average $p$-value $W_{\hat{a}}$ of a Wilcoxon signed-rank test $\operatorname{w}(S,P)$, as given by the following equation, of the two runtime label distributions ${S=\{ \gamma(a, e) \mid e \in \mathcal{I} \}}$ for an existing solver $a$ and \mbox{$P=\{ f(\hat a, e) \mid e \in \mathcal{I} \}$} of the new solver $\hat{a}$ is used to assess the statistical significance of the predicted runtime labels.
+To assess the statistical significance of the predicted runtime labels, the criterion uses the average $p$-value $W_{\hat{a}}$ of a Wilcoxon signed-rank test $\operatorname{w}(S,P)$ of the runtime label distributions ${S=\{ \gamma(a, e) \mid e \in \mathcal{I} \}}$ for an existing solver $a$ and \mbox{$P=\{ f(\hat a, e) \mid e \in \mathcal{I} \}$} of the new solver $\hat{a}$:
 %
@@
 %
-In order to improve the stability of this criterion, an exponential moving average is employed to mitigate the impact of outliers, with the process terminated when the value of $W^{(i)}_{\exp}$ drops below a fixed threshold.
+To improve the stability of this criterion and mitigate the impact of outliers, we employ an exponential moving average~$W^{(i)}_{\exp}$:
 %
@@
 \end{align*}
+%
+Active learning terminates when the value of $W^{(i)}_{\exp}$ drops below a fixed threshold.
 
@@
 In contrast, the ranking stopping criterion is less sophisticated than the previous one.
-In this case, the active learning process is terminated if the ranking induced by the model's predictions (Equation~\eqref{eq:rankingeq}) remains unchanged within the last~$l$~iterations.
-However, it should be noted that the concrete values of the predicted score $s_{\hat a}$ may still change.
-Nevertheless, in this case, the induced ranking is of sole interest.
+In this case, active learning terminates when the ranking induced by the model's predictions (Equation~\eqref{eq:rankingeq}) has remained unchanged within the last~$l$~iterations.
+While the concrete values of the predicted score $s_{\hat a}$ may still change, the induced ranking is of sole interest in this criterion.
 

### 2024-05-26 paper_jar/paper.tex
Revise Experimental Design

@@
 In this paper, we present a dynamic benchmark selection approach based on active learning.
-Our approach estimates the rank of a new solver among its competitors, striving to minimize benchmarking runtime but maximize rank prediction accuracy.
+Our approach estimates the rank of a new solver among its competitors, striving to minimize benchmarking runtime but maximize ranking accuracy.
 Instead of using real-valued solver runtimes, our approach works with discretized runtime labels, which yielded better solver rank predictions.
@@
 Without AL methods, it is not evident which instances to label and which not.
-In our scenario, the first goal is to maximize the utility of instances to our prediction model, which is measured by the rank prediction accuracy.
+In our scenario, the first goal is to maximize the utility of instances to our prediction model, which is measured by the ranking accuracy.
 Our second goal is to minimize the cost of label acquisition, which is measured by the instance's predicted runtime.
@@
 Section~\ref{sec:exdesign:disc-pred} presents a detailed analysis of the runtime transformation in our scenario.
-In particular, the analysis demonstrates that our benchmarking approach achieves higher rank prediction accuracy when working with discrete runtime labels rather than raw runtimes.
+In particular, the analysis demonstrates that our benchmarking approach achieves higher ranking accuracy when working with discrete runtime labels rather than raw runtimes.
 
@@
 
-In light of the preceding instantiations of Algorithm~\ref{algALBenchmark}, this section outlines the experimental design employed in the presented study.
-This includes an overview of the evaluation framework (cf.~Section~\ref{sec:exdesign:eval}), the datasets employed (cf.~Section~\ref{sec:exdesign:data}), the hyperparameter selections (cf.~Section~\ref{sec:exdesign:hyper}), the runtime discretization (cf.~Section~\ref{sec:exdesign:disc-pred}), and the implementation (cf.~Section~\ref{sec:exdesign:impl}).
+This section outlines the experimental design employed in the presented study.
+This includes an overview of the evaluation framework (cf.~Section~\ref{sec:exdesign:eval}), the datasets employed (cf.~Section~\ref{sec:exdesign:data}), the hyperparameter configuration (cf.~Section~\ref{sec:exdesign:hyper}), the runtime discretization (cf.~Section~\ref{sec:exdesign:disc-pred}), and the implementation (cf.~Section~\ref{sec:exdesign:impl}).
 
@@
     \For{$a \in \mathcal{A}$}{
+      %JB: mabye Lines 6 and 8 should use A' instead of A (at the moment, solver is also compared to itself but this does not increase ranking accuracy since second factor in Line 7 is 0)
       \If{$\bigl(s_k(a) - s_{\hat a}\bigr) \cdot \bigl(\operatorname{par_2}(a) - \operatorname{par_2}(\hat a)\bigr) > 0$}{
@@
   \tcp{Determine Runtime Fraction}
-  $r \leftarrow \sum\limits_{e \in \mathcal{I}} r(\hat a, e)$\;
+  $t_{\text{total}} \leftarrow \sum\limits_{e \in \mathcal{I}} r(\hat a, e)$\;
   $O_{\operatorname{rt}} \leftarrow 0$\;
@@
     \If{$\exists t, (e,t) \in \mathcal{R}$}{
-      $O_{\operatorname{rt}} \leftarrow O_{\operatorname{rt}} + \frac{t}{r}$\;
+      $O_{\operatorname{rt}} \leftarrow O_{\operatorname{rt}} + \frac{t}{t_{\text{total}}}$\;
     }
@@
 
-As stated in the Introduction, this work addresses the \emph{New-Solver Problem} (cf.~Definition~\ref{def:new-solver-problem}).
-As described in Section~\ref{sec:main:model}, a prediction model provides us with an estimated scoring value for the new solver~$\hat{a}$.
-To evaluate the efficacy of a concrete instantiation of Algorithm~\ref{algALBenchmark}, which entails a specific choice for all subroutines, we perform cross-validation on our set of solvers.
-
-This is demonstrated in Algorithm~\ref{alg:eval}.
-Each solver assumes the role of the new solver~$\hat{a}$ once (Line~2) and is consequently excluded from the set of solvers $\mathcal{A}$ (Line~3).
-After running our active-learning framework for solver~$\hat{a}$ (Line~4), we compute the value of both our optimization goals, i.e., ranking accuracy and runtime.
-The \emph{ranking accuracy} $O_{\operatorname{acc}} \in \left[0, 1\right]$ (higher is better) is defined as the fraction of correctly ranked pairs of elements, namely, $\left(\hat{a}, a\right)$, regarding the ground-truth scoring function $\operatorname{par_2}$, where $a$ belongs to the set of all solvers $\mathcal{A}$ (Lines~5-8).
-The \emph{fraction of runtime} required by the algorithm to reach its conclusion is represented by $O_{\operatorname{rt}}$, which lies within the interval $[0, 1]$ (lower is better).
-This metric compares the accumulated runtimes over the sampled instances to the accumulated runtimes over all instances in the dataset (Lines~9-13).
-Once all cross-validation results have been collected (Line~14), the output metrics are averaged (Line~15).
-
-The procedure is designed to identify an approach that maximizes Equation~\eqref{eq:opt}, in which $\delta \in \left[0, 1\right]$ is used for linear weighting, with weighting factor $\delta \in \left[0, 1\right]$, between the two optimization goals $O_{\operatorname{acc}}$ and $O_{\operatorname{rt}}$.
-The plotting of the approaches that maximize $O_\delta$ for all $\delta \in \left[0, 1\right]$ on an $O_{\operatorname{rt}}$-$O_{\operatorname{acc}}$-diagram provides us with a Pareto front of the best approaches for different optimization-goal weightings.
+To assess our active-learning framework, we perform cross-validation over a set of solvers (cf.~Algorithm~\ref{alg:eval}).
+Each solver assumes the role of the new solver~$\hat{a}$ once (Line~2) and is therefore excluded from the set of solvers $\mathcal{A}$ in that iteration (Line~3).
+After running active learning for solver~$\hat{a}$ (Line~4), we compute the value of two optimization goals, i.e., ranking accuracy and runtime.
+The \emph{ranking accuracy} $O_{\operatorname{acc}} \in \left[0, 1\right]$ (higher is better) is defined as the fraction of correctly ranked solver pairs~$\left(\hat{a}, a\right)$ regarding the PAR-2 scoring, where $a$ belongs to the set of all solvers $\mathcal{A}$ (Lines~5--8).
+The \emph{fraction of runtime} required by the framework to reach its conclusion is represented by $O_{\operatorname{rt}} \in [0, 1]$ (lower is better).
+This metric compares the accumulated runtimes over the sampled instances to the accumulated runtimes over all instances in the dataset (Lines~9--13).
+Once all cross-validation results have been collected (Line~14), the output metrics are averaged and returned (Lines~15--16).
+
+To compare different instantiations of our active-learning framework holistically, we insert the cross-validation results into Equation~\eqref{eq:opt}:
 %
@@
 %
+$\delta \in \left[0, 1\right]$ is a factor for weighting the two optimization goals $O_{\operatorname{acc}}$ and $O_{\operatorname{rt}}$.
+Plotting the approaches that maximize $O_\delta$ for all $\delta \in \left[0, 1\right]$ on an $O_{\operatorname{rt}}$-$O_{\operatorname{acc}}$-diagram provides us with a Pareto front of the best approaches for different optimization-goal weightings.
 
@@
 In our experiments, we work with the dataset of the SAT~Competition~2022 Anniversary Track~\cite{sat2022}.
-The dataset consists of 5355 instances with corresponding runtime data of 28 sequential SAT solvers.
-We also use a database of 56 instance features\footnote{\url{https://benchmark-database.de/getdatabase/base_db}} from the Global Benchmark Database~(GBD) by Iser~et~al.~\cite{IserS18}.
+This dataset consists of 5355 instances with corresponding runtimes of 28 sequential SAT solvers.
+For predictions within our framework, we also use a database of 56 instance features\footnote{\url{https://benchmark-database.de/getdatabase/base_db}} from the Global Benchmark Database~(GBD) by Iser~et~al.~\cite{IserS18}.
 They include instance size features and node distribution statistics for several graph representations of SAT instances, and are primarily inspired by the SATzilla~2012 features described in~\cite{features}.
@@
 For hyperparameter tuning, we randomly sample \SI{10}{\%} of the complete set of 5355 instances with stratification regarding the instance family.
-All instance families that are too \emph{small}, i.e., \SI{10}{\%} of them correspond to less than one instance, are put into a meta-family for stratification.
+All instance families that are too \emph{small}, i.e., \SI{10}{\%} of them correspond to less than one instance, are put into one meta-family for stratification.
+%JB: i.e., families with less than 10 instances? could be phrased more directly
 This \emph{tuning dataset} allows a more extensive exploration of the hyperparameter space.
@@
 
-Given the Algorithm~\ref{algALBenchmark}, there are several possible instantiations for the three subroutines \emph{ranking}, \emph{selection}, and \emph{stopping}.
-These experimental configurations are described in the following.
+There are several possible instantiations for the three subroutines \emph{ranking}, \emph{selection}, and \emph{stopping} in Algorithm~\ref{algALBenchmark}.
+We describe these experimental configurations next.
 
@@
 
-Regarding \emph{ranking} (cf.~Section~\ref{sec:main:model}), we experiment with the following instantiations.
+Regarding \emph{ranking} (cf.~Section~\ref{sec:main:model}), we experiment with the following instantiations:
 
 \begin{itemize}\setlength{\itemsep}{1pt}
-  \item Observed PAR-2 ranking of already sampled instances
+  \item Observed PAR-2 ranking of already sampled instances, without prediction
   \item Predicted runtime-label ranking
@@
 We experiment with the following instantiations of the instance \emph{selection} (cf. Section~\ref{sec:main:selection}).
-Since the potential runtime of the experiments is orders of magnitude larger than the model update time, we consider incrementing our benchmark one instance at a time rather than using batches, which is also proposed in recent advances in active learning~\cite{SinhaED19,2019gaal}.
+Since the potential runtime of the solver experiments is orders of magnitude larger than the model update time, we increment our benchmark one instance at a time rather than using batches of multiple instances, which is also proposed in recent advances in active learning~\cite{SinhaED19,2019gaal}.
+%JB: not entirely clear if one-at-a-time "is also propsed" or batching "is also proposed"
 A drawback of this is the lack of parallel execution of runtime experiments.
@@
     \item Runtime scaling: Whether to normalize uncertainty scores per instance by the average runtime of solvers on it or use the absolute values.
+    %JB: but isn't the uncertainty score a difference of probabilities rather than directly runtime-related anyway?
+so probably not a normalization in the strict sense (maybe use term "divide" instead)
   \end{itemize}
@@
 
-We evaluate the following \emph{stopping} (cf. Section~\ref{sec:main:stopping}) criteria.
+We evaluate the following \emph{stopping} (cf. Section~\ref{sec:main:stopping}) criteria:
 
 \begin{itemize}\setlength{\itemsep}{1pt}
-  \item Subset-size stopping criterion, using \SI{10}{\%} or \SI{20}{\%} of instances
+  \item Subset-size stopping criterion: Stop after sampling \SI{10}{\%} or \SI{20}{\%} of instances.
   \item Ranking stopping criterion
@@
     \item Average of $p$-values to drop below: \SI{5}{\%}.
-    \item Exponential-moving average: Incorporate previous significance values by using an EMA with $\beta = 0.1$ or $\beta = 0.7$.
+    \item Exponential moving average: Incorporate previous significance values by using an EMA with $\beta = 0.1$ or $\beta = 0.7$.
   \end{itemize}
@@
 
-One of the most important parts of our framework is the solver runtime model (cf.~Section~\ref{sec:main:model}), which influences instance selection in active learning.
+One of the most important parts of our framework is the solver runtime model (cf.~Section~\ref{sec:main:model}).
 In this section, we detail how clustering solver runtimes can lead to informative runtime labels, and which machine learning models are best at predicting them.
@@
 
-\paragraph{Regressing runtime}
+\paragraph{Regressing Runtime}
 
@@
 To this end, we use 46 instance features and 27 known solver runtimes as inputs for the predictions.
-We repeat this evaluation for each solver as prediction target and report the average and standard deviation of prediction performance over the 28 solvers.
+We repeat this evaluation for each solver as the prediction target and report the average and standard deviation of prediction performance over the 28 solvers.
 
@@
 In particular, a root mean squared error (RMSE) of about \SI{2000}{s} is quite high for solver runtimes that range from 0 to \SI{5000}{s}.
+Such unreliable runtime estimates could affect active learning negatively.
 Therefore, we decided to discretize the runtimes for prediction, moving from a regression scenario to a classification scenario.
 
-\paragraph{Classifying timeouts}
+\paragraph{Classifying Timeouts}
 
@@
 
-\paragraph{Classifying discretized runtime labels}
+\paragraph{Classifying Discretized Runtime Labels}
 
@@
 We conduct this clustering step per instance, i.e., the solvers forming a particular cluster may differ from instance to instance.
-This clustering forms the basis of the adapted solver scoring function~$s_k$ (cf.~Equation~\eqref{eq:rankingeq}).
+The adapted solver scoring function~$s_k$ (cf.~Equation~\eqref{eq:rankingeq}) subsequently uses the cluster labels.
 In preliminary experiments, the most \emph{useful} labels were produced with hierarchical clustering and a log-single-linkage criterion; we will discuss our notion of labels usefulness later.
@@
 By definition, the $k$-th cluster always contains the timeouts and no other runtimes (cf.~Section~\ref{sec:main:model:transformation}).
-To quantify the distance between two clusters, we consider the minimum pairwise difference of logarithmic runtimes between solvers in the two clusters.
+To quantify the distance between two clusters, we consider the minimum pairwise difference of logarithmic runtimes between solvers from the two clusters.
 
@@
 
-Given this discretization procedure, we evaluate the performance of the best classification models from Table~\ref{tab:timeout-prediction} in predicting these runtime labels.
-Table~\ref{tab:cluster-prediction} shows the prediction performance for $k \in \lbrace 3, 4, 5 \rbrace$ clusters.
+Next, we use the runtimes labels from this discretization procedure as prediction targets.
+Table~\ref{tab:cluster-prediction} shows the prediction performance for $k \in \lbrace 3, 4, 5 \rbrace$ clusters and using the best classification models from Table~\ref{tab:timeout-prediction}.
 We train the prediction models either with 46 instance features and 27 known solver runtimes (second column) or additionally including 27 features representing the known $C_k$ cluster labels of other solvers (third column).
+
 For all prediction models and values of~$k$, the prediction performance is better when the cluster labels are included as features.
@@
 	\centering
-	\caption{Ranking of solvers on the Anniversary Track for the PAR-2 score and our discretized $s_3$ score.}
+	\caption{Ranking of solvers for the PAR-2 score and our discretized $s_3$ score.}
 	\label{tab:ranking}
@@
 
-To ensure that the runtime labels cannot only be predicted well but are \emph{useful} for active learning, we evaluate whether the labels still discriminate solvers and agree with the actual PAR-2 ranking.
+To ensure that the runtime labels also are \emph{useful} for active learning, we evaluate whether the labels still discriminate solvers and agree with the actual PAR-2 ranking.
 Table~\ref{tab:ranking} shows the ranking of the solvers in the SAT Competition~2022 Anniversary Track~\cite{sat2022} according to the standard PAR-2 score and according to our discretized $s_3$ score.
@@
 %JB: does "after discretization" mean s_3 or timeout? and what is then the number for the other one?
-These results support the usefulness of discretized runtimes for our framework.
+These results support the usefulness of discretized runtimes for our active learning framework.
 

### 2024-05-26 paper_jar/paper.tex
Revise Evaluation

@@
     }
-    \caption{Ranking approaches}
+    \caption{Ranking approaches.}
     \label{fig:annitraincolorranking}
@@
     }
-    \caption{Selection approaches}
+    \caption{Selection approaches.}
     \label{fig:annitraincolorselection}
@@
     }
-    \caption{Stopping criteria}
+    \caption{Stopping criteria.}
     \label{fig:annitraincolorstopping}
@@
   \caption{
-    $O_{\operatorname{rt}}$-$O_{\operatorname{acc}}$-diagrams comparing different hyperparameter instantiations of our active-learning framework on the hyperparameter-tuning dataset.
+    $O_{\operatorname{rt}}$-$O_{\operatorname{acc}}$-diagrams (cf.
+Section~\ref{sec:exdesign:eval}) comparing different hyperparameter instantiations of our active-learning framework on the hyperparameter-tuning dataset.
     The x-axis shows the ratio of total solver runtime on the sampled instances relative to all instances.
-    The y-axis shows the ranking accuracy (cf. Section~\ref{sec:exdesign:eval}).
-    Each line entails the front of Pareto-optimal configurations for the respective hyperparameter instantiation.
+    The y-axis shows the ranking accuracy.
+    Each line represents the front of Pareto-optimal configurations for the respective hyperparameter instantiation.
   }
@@
 
-Our experiments follow the evaluation framework introduced in Section~\ref{sec:exdesign:eval}.
-Fig.~\ref{fig:e2eallsolvers} shows the performance of the approaches from Section~\ref{sec:exdesign:hyper} on $O_{\operatorname{rt}}$-$O_{\operatorname{acc}}$ plots for the hyperparameter-tuning dataset.
-Evaluating a given configuration with Algorithm~\ref{alg:eval} returns a point $\left(O_{\operatorname{rt}},\, O_{\operatorname{acc}}\right)$.
+\subparagraph{Methodology}
+
+Fig.~\ref{fig:e2eallsolvers} shows the performance of the hyperparameter configurations from Section~\ref{sec:exdesign:hyper} on $O_{\operatorname{rt}}$-$O_{\operatorname{acc}}$ plots (cf.~Section~\ref{sec:exdesign:eval}) for the hyperparameter-tuning dataset.
+Evaluating one configuration with Algorithm~\ref{alg:eval} returns one point $\left(O_{\operatorname{rt}},\, O_{\operatorname{acc}}\right)$.
 We do not show intermediate results of the active learning procedure, but only the final results after stopping.
@@
 
-Regarding ranking approaches (Fig.~\ref{fig:annitraincolorranking}), using the predicted $s_3$-induced run\-time-label ranking consistently outperforms the partially observed PAR-2 ranking for every possible value of the tradeoff parameter~$\delta$.
-This result is expected, since selection decisions are not random.
-For example, we might sample more instances of a family if it benefits the discrimination of solvers.
-While the partially observed PAR-2 score is biased, the prediction model can account for this.
+\subparagraph{Ranking}
+
+Regarding ranking approaches (Fig.~\ref{fig:annitraincolorranking}), the predicted runtime-label ranking consistently outperforms the partially observed PAR-2 ranking for every possible value of the tradeoff parameter~$\delta$.
+This result is expected, since selection decisions are usually not random.
+For example, we might sample more instances of a particular family if it benefits the discrimination of solvers.
+While the partially observed PAR-2 score is biased by such a sampling, the prediction model can account for this.
+
+\subparagraph{Selection}
 
@@
 
+\subparagraph{Stopping}
+
 Regarding the stopping criteria (Fig.~\ref{fig:annitraincolorstopping}), the ranking stopping criterion performs most consistently well.
 When accuracy is strongly favored (very high $\delta$), the Wilcoxon stopping criterion performs better.
-The subset-size stopping criterion performs reasonably well, but does not improve beyond a certain accuracy because it samples a fixed subset of instances.
+The subset-size stopping criterion performs reasonably well, but does not improve beyond a certain accuracy because it samples a fixed subset of instances, independent from prediction performance.
 
@@
     Scatter plot comparing different instantiations of the tradeoff parameter~$\delta$ for our active learning framework on the hyperparameter-tuning dataset.
-    The x-axis shows the fraction of runtime $O_{\operatorname{rt}}$ of the sample, while the y-axes show the fraction of sampled instances and the ranking accuracy, respectively.
+    The x-axes show the fraction of runtime $O_{\operatorname{rt}}$ of the sample, while the y-axes show the fraction of sampled instances and the ranking accuracy, respectively.
     The color indicates the weighting between different optimization objectives $\delta \in \left[0, 1\right]$.
@@
 
+\subparagraph{Objective weighting}
+
 Fig.~\ref{fig:annitrainoptgoalruntime} shows an interesting consequence of weighting our optimization goals:
-On the one hand, if we want to get a \emph{rough} estimate of a solver's performance quickly (low~$\delta$), the approaches favor selecting many \emph{easy} instances.
+On the one hand, if we want to get a \emph{rough} estimate of a solver's performance quickly (low~$\delta$), our framework favors selecting many \emph{easy} instances.
 In particular, the fraction of sampled instances is larger than the fraction of runtime.
 By having many observations, it is easier to build a model.
-On the other hand, if we want to get a \emph{good} estimate of a solver's performance in a moderate amount of time (high~$\delta$), approaches favor selecting few, \emph{difficult} instances.
+On the other hand, if we want to get a \emph{good} estimate of a solver's performance in a moderate amount of time (high~$\delta$), our framework favors selecting few, \emph{difficult} instances.
 In particular, the fraction of instances is smaller than the fraction of runtime.
@@
 After selecting the most promising hyperparameters, we run our active learning experiments on the complete Anniversary Track dataset (5355 instances).
-The above range $\delta \in \left[0.2, 0.8\right]$ results in only two different configurations.
+The above range $\delta \in \left[0.2, 0.8\right]$ results in only two different hyperparameter configurations.
 The best performing approach for $\delta \in \left[0.2, 0.7\right]$ uses the predicted runtime label ranking, information-gain sampling, and ranking stopping criterion.
@@
 
-Table~\ref{tab:fulldataset} shows how the two active learning approaches (column \emph{AL}) compare to two static baselines:
+Table~\ref{tab:fulldataset} shows how these two active learning approaches (column \emph{AL}) compare to two static baselines:
 \emph{Random} samples instances until it reaches roughly the same fraction of runtime as the AL benchmark sets.
@@
 However, the performance differences are not significant with respect to a Wilcoxon signed-rank test with $\alpha = 0.05$, and also depend on the fraction of the runtime sampled (cf.~Fig.~\ref{fig:annitraincolorselection}).
-However, a clear advantage of our approach is that it indicates when to stop adding new instances, depending on the tradeoff parameter~$\delta$.
+However, a clear advantage of our approach is that it indicates when to stop adding new instances.
 While the active learning results on the full dataset are not as strong as on the smaller tuning dataset, they still show the benefit of making the benchmark selection dependent on the solvers to be distinguished.
@@
 This outcome is somewhat anticipated, given that the static benchmark fails to reflect an appropriate balance of instance families.
-For instance, families whose instances are uniformly randomly selected by AL, e.g., for different solvers, appear less frequently in this benchmark than families where some instances are sampled more often than others.
+For example, families whose instances are uniformly randomly selected by AL, e.g., for different solvers, appear less frequently in this benchmark than families where some instances are sampled more often than others.
 % MI: What? I don't understand this sentence.
@@
   \caption{
-    Scatter plot illustrating the relative importance of different instance families for ranking.
+    Scatter plot illustrating the relative importance of different instance families for ranking on the full dataset.
 The x-axis depicts the frequency of instance families in the dataset, while the y-axis represents the average frequency of instance families in the samples selected by active learning.
@@
 Fig.~\ref{fig:instoccs} illustrates the distribution of the frequency with which individual instances are selected across all active learning benchmarks.
-%JB: on the full dataset (complete Anniversary Track), I suppose? and what is "all" aggregating over? these information may also go into the figure's caption
+%JB: on the full dataset, I suppose (since tuning dataset does not contain all instances)? and what is "all" aggregating over? these information may also go into the figure's caption
 It can be observed that a small proportion of instances is selected with great frequency, while most instances are selected on only a few occasions.
@@
 
-The selection decisions of our approach also reveal the importance of instance families to our framework.
+The selection decisions of our approach also reveal the importance of instance families for our framework.
 Fig.~\ref{fig:annifinalfamilies} illustrates the occurrence of instance families within the dataset and the benchmarks created by active learning.
-We use the best performing configurations for all values of $\delta \in \left[0, 1\right]$ and examine the selection decisions made by the active learning approach on the SAT Competition~2022 Anniversary Track dataset~\cite{sat2022}.
+We use the best performing configurations for all values of $\delta \in \left[0, 1\right]$.
 While most families exhibit the same fraction in the dataset and the sampled benchmarks, a few outliers require further discussion.

### 2024-05-26 paper_jar/paper.tex
Revise Conclusion

@@
 On data from the SAT~Competition~2022 Anniversary Track, we can determine a new solver's PAR-2 ranking with about \SI{92}{\%} accuracy while only requiring \SI{10}{\%} of the full-evaluation time.
-We have evaluated several ranking algorithms, instance-selection approaches, and stopping criteria within our sequential active learning framework.
+We have evaluated several ranking methods, instance-selection approaches, and stopping criteria within our sequential active learning framework.
 We have also looked at which instance families are the most prevalent in selection decisions.
@@
 Future work could compare more sub-routines for ranking, instance selection, and stopping.
-Further, our evaluation framework can be used for other computation-intensive problems, like other $\mathcal{NP}$-complete problems than SAT.
-Such problems share most of the relevant properties of SAT, like established instance features, a complete benchmark is expensive, and traditional benchmark selection requires expert knowledge.
+Further, our evaluation framework can be used for other computation-intensive problems.
+In particular, many $\mathcal{NP}$-complete problems share most of the relevant properties of SAT, like established instance features, a complete benchmark is expensive, and traditional benchmark selection requires expert knowledge.
 
@@
 
-To perform benchmarking on a computer cluster with multiple cores, it is beneficial to work with batches of multiple instances.
-In contrast, our approach is limited to benchmarking with a single instance at a time.
-In future work, we intend to investigate models that learn from partial labels, which can be extracted from larger batch sizes in a parallel approach that synchronizes instance selection and model updates.
-
 \backmatter

### 2024-05-26 paper_jar/paper.tex
Run Grammarly over text

@@
 In particular, our approach enables \emph{ranking} the new solver among a set of existing solvers.
-We show that for ranking, it is not necessary to predict the exact per-instance runtime of the solvers, which would be a more challenging task.
+We show that predicting the exact per-instance runtime of the solvers, which would be a more challenging task, is not necessary for ranking.
 Instead, we work with discretized runtime labels.
-We use instance features and known solver runtimes for predicting these labels.
+We use instance features and known solver runtimes to predict these labels.
 Furthermore, we minimize the \emph{runtime} required for our approach to arrive at its conclusion.
 Moreover, we select benchmark instances \emph{non-randomly} and \emph{incrementally}.
-In particular, we consider runtime information from already done experiments when choosing the next instance.
+In particular, we consider runtime information from experiments already done when choosing the next instance.
 By doing so, we can control the properties of the benchmarking approach, such as its required runtime.
@@
 We evaluated our approach using the SAT Competition~2022 Anniversary Track dataset~\cite{sat2022}, consisting of 5355~instances and runtimes of 28~solvers.
-Cross-validation was performed by treating each solver as the new solver once and learning to predict the PAR-2 rank of that solver.
+Cross-validation was performed by treating each solver as the new solver once and learning to predict its PAR-2 rank.
 On average, our predictions achieve a ranking accuracy of approximately \SI{92}{\%} with only approximately \SI{10}{\%} of the runtime required to evaluate these solvers on the complete set of instances.
@@
 
-This article is an extended and refined version of our conference paper with the title ``Active Learning for SAT Solver Benchmarking''~\cite{fuchs2023active}.
+This article is an extended and refined version of our conference paper titled ``Active Learning for SAT Solver Benchmarking''~\cite{fuchs2023active}.
 In particular, we significantly extended the analysis of runtime-prediction and runtime-discretization approaches (cf.~Section~\ref{sec:exdesign:disc-pred}), which determined our chosen solver runtime model for instance selection.
@@
 Benchmarking is of great interest in numerous research areas and represents an active field of research in its own right.
-Studies have shown that the compilation of benchmark instances poses a number of challenges.
-The use of biased benchmarks can easily lead to fallacious interpretations~\cite{abs-2107-07002}.
-Benchmarking also involves a number of interchangeable elements, including the performance measures used, the way the measures are aggregated, and the way missing values are handled.
-Questionable research practices could alter these elements a-posteriori in order to meet expectations, thereby skewing the results~\cite{NiesslHWCB22}.
-
-The following sections discusses related work from the areas of static benchmark selection (cf.~Section~\ref{sec:related:static-selection}), algorithm configuration (cf.~Section~\ref{sec:related:algo-configuration}), incremental benchmark selection (cf.~Section~\ref{sec:related:incremental-selection}), and active learning (cf.~Section~\ref{sec:related:active-learning}).
+Studies have shown that the compilation of benchmark instances poses several challenges.
+Using biased benchmarks can easily lead to fallacious interpretations~\cite{abs-2107-07002}.
+Benchmarking also involves several interchangeable elements, including the performance measures used, how the measures are aggregated, and how missing values are handled.
+Questionable research practices could alter these elements a-posteriori to meet expectations, thereby skewing the results~\cite{NiesslHWCB22}.
+
+The following sections discuss related work from the areas of static benchmark selection (cf.~Section~\ref{sec:related:static-selection}), algorithm configuration (cf.~Section~\ref{sec:related:algo-configuration}), incremental benchmark selection (cf.~Section~\ref{sec:related:incremental-selection}), and active learning (cf.~Section~\ref{sec:related:active-learning}).
 Table~\ref{tab:requirements} compares the most relevant approaches, which all pursue slightly different goals.
@@
 In such competitions, the organizers define the rules for composing the benchmarks.
-These selection strategies are primarily static, in that they do not depend on particular solvers to distinguish.
+These selection strategies are primarily static since they do not depend on particular solvers to distinguish.
 Balint et al. provide an overview of benchmark-selection criteria in different solver competitions~\cite{balint2015overview}.
@@
 Hoos et al.~\cite{HoosKSS13} discuss desirable properties of SAT benchmark instances.
-They identify three key selection criteria: instance variety to avoid over-fitting, adapted instance hardness (not too easy but also not too hard), and the avoidance of duplicate instances. 
+They identify three key selection criteria: instance variety to avoid over-fitting, adapted instance hardness (neither too easy nor too hard), and avoiding duplicate instances. 
 To filter instances that are too similar, they employ a distance-based approach with the SATzilla features~\cite{XuHHL08,features}.
 It should be noted, however, that the approach does not optimize for benchmark \emph{runtime}.
-Instead, instances are selected \emph{at random}, with the exception of constraints on instance hardness and feature distance.
+Instead, instances are selected \emph{at random} except considering constraints on instance hardness and feature distance.
 
@@
 In contrast, instance selection is made \emph{randomly} in SMAC, i.e., without building a model over instances.
-Furthermore, an algorithm configurator cannot be used to \emph{rank} or \emph{score} a new solver relative to others since algorithm configuration is designed to identify the optimal configuration.
+Furthermore, since algorithm configuration is designed to identify the optimal configuration, an algorithm configurator cannot be used to \emph{rank} or \emph{score} a new solver relative to others.
 
@@
 In passive machine learning, prediction models are trained on datasets with given instance labels (cf.~Fig.~\ref{fig:passive}).
-In contrast, active learning (AL) begins with no or little labeled data.
+In contrast, active learning (AL) begins with little or no labeled data.
 AL then repeatedly selects interesting problem instances for which to acquire labels, gradually improving the prediction model (cf.~Fig.~\ref{fig:activepool}).
@@
 Our second goal is to minimize the cost of label acquisition, which is measured by the instance's predicted runtime.
-Consequently, our overall objective is to develop an accurate prediction model without the necessity of labeling each data point.
-
-Rubens et.~al.~\cite{RubensESK15} survey active-learning advances.
+Consequently, our overall objective is to develop an accurate prediction model without labeling each data point.
+
+Rubens et~al.~\cite{RubensESK15} survey active-learning advances.
 While synthesis-based AL methods~\cite{0001AEMN22,GarzonMG22,2019gaal} generate instances for labeling, pool-based methods~\cite{GolbandiKL11,HarpaleY08,KorenBV09} rely on a fixed set of unlabeled instances to sample from.
@@
 In the latter direction, Volpato and Guangyan employ pool-based AL to train an instance-specific algorithm selector~\cite{volpato2019active}.
-However, their objective is not benchmarking a solver's overall performance but recommending the best solver for each SAT instance from a set of available solvers.
+However, their objective is not to benchmark a solver's overall performance but to recommend the best solver for each SAT instance from a set of available solvers.
 
@@
   \end{tabular}
-  \caption{Types of machine learning (depiction inspired by Rubens~et.al.~\cite{RubensESK15}).}
+  \caption{Types of machine learning (depiction inspired by Rubens~et~al.~\cite{RubensESK15}).}
   \label{fig:learning}
@@
 The acquired result is subsequently employed to update the prediction model (Line~7). 
-When the stopping criterion is met (Line~3), the benchmarking loop terminates and the final score of $\hat{a}$ is estimated (Line~8). 
+When the stopping criterion is met (Line~3), the benchmarking loop terminates, and the final score of $\hat{a}$ is estimated (Line~8). 
 Algorithm~\ref{algALBenchmark} returns this score, along with the acquired benchmark instances and runtime measurements.
@@
 In the update step (Algorithm~\ref{algALBenchmark},~Line~7), the model is trained to predict a transformed version of solver runtime (cf.~Section~\ref{sec:main:model:transformation}) for the new solver~$\hat{a}$, using the features described in Section~\ref{sec:exdesign:data} and the previously acquired runtimes~$\mathcal{R}$.
-Each iteration trains a new prediction model from scratch, since the time for running the solver~$\hat{a}$ on one instance~$e$ (Line~5) dominates the training time of the model~$M$ on all acquired instances~$\mathcal{R}$ by a significant margin.
+Each iteration trains a new prediction model from scratch since the time for running the solver~$\hat{a}$ on one instance~$e$ (Line~5) dominates the training time of the model~$M$ on all acquired instances~$\mathcal{R}$ by a significant margin.
 Once the benchmarking loop terminates, the score of the new solver~$\hat{a}$ is predicted and then returned (Lines~8--9).
@@
 For the prediction model~$M$, we transform the real-valued runtimes into discrete runtime labels on a per-instance basis.
-For each instance $e \in \mathcal{I}$, we use a clustering algorithm to assign the runtimes in $\bigl\{ r(a, e) \mid a \in \mathcal{A} \bigr\}$ to one of $k$ clusters $C_1, \dots, C_k$ such that the fastest runtimes for the instance $e$ are in cluster $C_1$ and the slowest are in cluster $C_{k-1}$.
+For each instance $e \in \mathcal{I}$, we use a clustering algorithm to assign the runtimes in $\bigl\{ r(a, e) \mid a \in \mathcal{A} \bigr\}$ to one of $k$ clusters $C_1, \dots, C_k$ such that the fastest runtimes for the benchmark instance $e$ are in cluster $C_1$ and the slowest are in cluster $C_{k-1}$.
 Timeouts $\tau$ always form a separate cluster $C_{k}$.
@@
 
-The uncertainty sampling strategy selects the instance that is closest to the model's decision boundary.
+The uncertainty sampling strategy selects the instance closest to the model's decision boundary.
 This is achieved by selecting the instance $e \in \mathcal{I} \setminus \tilde{\mathcal{I}}$ that minimizes the certainty~$U(e)$, which is specified by the following equation:
@@
 The Wilcoxon stopping criterion terminates the active-learning process based on the confidence that the predicted runtime labels of the new solver are sufficiently different from those of existing solvers.
-This criterion is loosely inspired by Matricon et.~al.~\cite{MatriconAFSH21}.
+This criterion is loosely inspired by Matricon et~al.~\cite{MatriconAFSH21}.
 To assess the statistical significance of the predicted runtime labels, the criterion uses the average $p$-value $W_{\hat{a}}$ of a Wilcoxon signed-rank test $\operatorname{w}(S,P)$ of the runtime label distributions ${S=\{ \gamma(a, e) \mid e \in \mathcal{I} \}}$ for an existing solver $a$ and \mbox{$P=\{ f(\hat a, e) \mid e \in \mathcal{I} \}$} of the new solver $\hat{a}$:
@@
 In addition, we retrieve instance family information\footnote{\url{https://benchmark-database.de/getdatabase/meta_db}} to evaluate the composition of our sampled benchmarks.
-Instance families consist of instances from the same application domain, such as planning, cryptography, etc., and are a valuable tool for analyzing solver performance.
+Instance families consist of instances from the same application domain, such as planning, cryptography, etc., and are valuable for analyzing solver performance.
 %
@@
 \begin{itemize}\setlength{\itemsep}{1pt}
-  \item Observed PAR-2 ranking of already sampled instances, without prediction
+  \item Observed PAR-2 ranking of already sampled instances, without using prediction
   \item Predicted runtime-label ranking
@@
     \item
-    History size: Consider the latest 1, 10, 20, 30, or 40 predictions within a voting approach for stability.
+    History size: For stability, consider the latest 1, 10, 20, 30, or 40 predictions within a voting approach.
     The latest $x$ predictions for each instance vote on the instance's winning label.
@@
 One of the most important parts of our framework is the solver runtime model (cf.~Section~\ref{sec:main:model}).
-In this section, we detail how clustering solver runtimes can lead to informative runtime labels, and which machine learning models are best at predicting them.
+In this section, we detail how clustering solver runtimes can lead to informative runtime labels and which machine learning models are best at predicting them.
 To this end, we analyze prediction performance for estimating runtimes, timeouts, and discretized runtimes.
@@
 As a baseline, Table~\ref{tab:regression} shows the prediction performance of two regression models, i.e., a random forest and a multilayer perceptron.
-We employ these models to directly estimate a new solver's runtime as the prediction target.
+We employ these models to estimate a new solver's raw runtime as the prediction target.
 To this end, we use 46 instance features and 27 known solver runtimes as inputs for the predictions.
@@
 However, both still perform poorly, as evidenced by the magnitudes of the mean regression errors in Table~\ref{tab:regression}.
-In particular, a root mean squared error (RMSE) of about \SI{2000}{s} is quite high for solver runtimes that range from 0 to \SI{5000}{s}.
+In particular, a root mean squared error (RMSE) of about \SI{2000}{s} is relatively high for solver runtimes that range from 0 to \SI{5000}{s}.
 Such unreliable runtime estimates could affect active learning negatively.
@@
 This binary prediction target represents a very strong discretization of runtimes.
-We compare a variety of prediction models, which pursue different learning paradigms, each of them trained with either instance features only (second column) or instance and runtime features (third column).
+We compare a variety of prediction models that pursue different learning paradigms.
+Each is trained with either instance features only (second column) or instance and runtime features (third column).
 As the evaluation metric for prediction performance, we use Matthews Correlation Coefficient (MCC)~\cite{gorodkin2004comparing, matthews1975comparison}, which reaches its maximum of~1 for a perfect prediction, is~0 for random guessing, and has a minimum of~-1.
@@
 The adapted solver scoring function~$s_k$ (cf.~Equation~\eqref{eq:rankingeq}) subsequently uses the cluster labels.
-In preliminary experiments, the most \emph{useful} labels were produced with hierarchical clustering and a log-single-linkage criterion; we will discuss our notion of labels usefulness later.
+In preliminary experiments, the most \emph{useful} labels were produced with hierarchical clustering and a log-single-linkage criterion; we will discuss our notion of label usefulness later.
 Other clustering approaches we have tried include hierarchical clustering with mean-, median-, and full-linkage criteria, as well as $k$-means and spectral clustering.
@@
   \centering
-  \caption{Test-set prediction performance for predicting the $C_k$ cluster labels (a discrete target) with the best performing models.}
+  \caption{Test-set prediction performance for predicting the $C_k$ cluster labels (a discrete target) with the best-performing models.}
   \label{tab:cluster-prediction}
@@
 Next, we use the runtimes labels from this discretization procedure as prediction targets.
-Table~\ref{tab:cluster-prediction} shows the prediction performance for $k \in \lbrace 3, 4, 5 \rbrace$ clusters and using the best classification models from Table~\ref{tab:timeout-prediction}.
+Table~\ref{tab:cluster-prediction} shows the prediction performance for the best classification models from Table~\ref{tab:timeout-prediction} and $k \in \lbrace 3, 4, 5 \rbrace$ clusters.
 We train the prediction models either with 46 instance features and 27 known solver runtimes (second column) or additionally including 27 features representing the known $C_k$ cluster labels of other solvers (third column).
@@
 In addition, the prediction performance degrades rapidly as the number of clusters increases.
-In all cases, the stacking ensemble performs best, so we use it in all subsequent experiments.
+The stacking ensemble performs best in all cases, so we use it in all subsequent experiments.
 
@@
 	\end{subfigure}
-	\caption{Solver pairs with significant differences regarding their scorings are colored white, non-significant ones black. Solvers are ordered indentically to Table~\ref{tab:ranking}.}
+	\caption{Solver pairs with significant differences regarding their scorings are colored white, non-significant ones black. Solvers are ordered identically to Table~\ref{tab:ranking}.}
 	\label{fig:solver-confusion-matrices}
@@
 The matrices in Fig.~\ref{fig:solver-confusion-matrices} visualize which solver pairs exhibit significant differences in their PAR-2 scores, timeouts, and $s_3$ scores.
-According to a Wilcoxon signed-rank test with a significance level of $\alpha = 0.05$, \SI{87.83}{\%} of the solver pairs have significantly different scores after discretization, which is only a small decrease compared to \SI{89.95}{\%} before discretization.
+According to a Wilcoxon signed-rank test with a significance level of $\alpha = 0.05$, \SI{87.83}{\%} of the solver pairs have significantly different scores after discretization, which is only a slight decrease compared to \SI{89.95}{\%} before discretization.
 %JB: numbers in notebook "bonus_results" differ
@@
 For reproducibility, our source code and data are available on GitHub (see footnotes in Section~\ref{sec:intro}).
-Our code is implemented in \textsc{Python} using \emph{scikit-learn}~\cite{scikit-learn} for making predictions and \emph{gbd-tools}~\cite{IserS18} for retrieving SAT instances.
+Our code is implemented in \textsc{Python} using \emph{scikit-learn}~\cite{scikit-learn} for making predictions, and \emph{gbd-tools}~\cite{IserS18} for retrieving SAT instances.
 
@@
 We do not show intermediate results of the active learning procedure, but only the final results after stopping.
-The plotted lines represent the best performing configurations per ranking approach (Fig.~\ref{fig:annitraincolorranking}), selection approach (Fig.~\ref{fig:annitraincolorselection}), and stopping criterion (Fig.~\ref{fig:annitraincolorstopping}).
+The plotted lines represent the best-performing configurations per ranking approach (Fig.~\ref{fig:annitraincolorranking}), selection approach (Fig.~\ref{fig:annitraincolorselection}), and stopping criterion (Fig.~\ref{fig:annitraincolorstopping}).
 In particular, we show the Pareto front, i.e., from all configurations that share a particular value of the plotted hyperparameter, we take the maximum ranking accuracy over all remaining hyperparameters \emph{not} shown in the corresponding plot.
@@
 
-Regarding ranking approaches (Fig.~\ref{fig:annitraincolorranking}), the predicted runtime-label ranking consistently outperforms the partially observed PAR-2 ranking for every possible value of the tradeoff parameter~$\delta$.
-This result is expected, since selection decisions are usually not random.
+Regarding ranking approaches (Fig.~\ref{fig:annitraincolorranking}), the predicted runtime-label ranking outperforms the partially observed PAR-2 ranking consistently for every possible value of the tradeoff parameter~$\delta$.
+This result is expected since selection decisions are usually not random.
 For example, we might sample more instances of a particular family if it benefits the discrimination of solvers.
-While the partially observed PAR-2 score is biased by such a sampling, the prediction model can account for this.
+While such a sampling biases the partially observed PAR-2 score, the prediction model can account for this.
 
@@
 Information-gain sampling selects instances that maximize the expected reduction in entropy.
-This means that we sample instances that reveal similarities between solvers rather than differences, which helps to build a confident model quickly.
+This means that we sample instances that reveal similarities rather than differences between solvers, which helps to build a confident model quickly.
 However, the method cannot select instances that are useful for distinguishing solvers later.
-Random sampling performs reasonably well, but is outperformed by uncertainty sampling in all cases, demonstrating the benefit of actively selecting instances based on a prediction model.
+Random sampling performs reasonably well but is outperformed by uncertainty sampling in all cases, demonstrating the benefit of actively selecting instances based on a prediction model.
 
@@
 When accuracy is strongly favored (very high $\delta$), the Wilcoxon stopping criterion performs better.
-The subset-size stopping criterion performs reasonably well, but does not improve beyond a certain accuracy because it samples a fixed subset of instances, independent from prediction performance.
+The subset-size stopping criterion performs reasonably well, but does not improve beyond a certain accuracy because it samples a fixed subset of instances, independent of prediction performance.
 
@@
 The above range $\delta \in \left[0.2, 0.8\right]$ results in only two different hyperparameter configurations.
-The best performing approach for $\delta \in \left[0.2, 0.7\right]$ uses the predicted runtime label ranking, information-gain sampling, and ranking stopping criterion.
+The best-performing approach for $\delta \in \left[0.2, 0.7\right]$ uses the predicted runtime label ranking, information-gain sampling, and ranking stopping criterion.
 It can predict the PAR-2 ranking of a new solver with \SI{90.48}{\%} accuracy ($O_{\operatorname{acc}}$) in only \SI{5.41}{\%} of the total evaluation time ($O_{\operatorname{rt}}$).
-The best performing approach for $\delta \in (0.7, 0.8]$ uses the predicted runtime label ranking, uncertainty sampling, and ranking stopping criterion.
+The best-performing approach for $\delta \in (0.7, 0.8]$ uses the predicted runtime label ranking, uncertainty sampling, and ranking stopping criterion.
 It can predict the PAR-2 ranking of a new solver with \SI{92.33}{\%} accuracy~($O_{\operatorname{acc}}$) in only \SI{10.35}{\%} of the total evaluation time~($O_{\operatorname{rt}}$).
@@
   \caption{
-    Performance comparison (on the full dataset) of the best performing AL approach (\emph{AL}), random sampling of the same runtime fraction with 1000 repetitions (\emph{Random}), and static selection of the instances most frequently sampled by active learning approaches (\emph{Most Frequent}).
+    Performance comparison (on the full dataset) of the best-performing AL approach (\emph{AL}), random sampling of the same runtime fraction with 1000 repetitions (\emph{Random}), and static selection of the instances most frequently sampled by active learning approaches (\emph{Most Frequent}).
   }
@@
 Both of our AL approaches outperform random sampling.
-However, the performance differences are not significant with respect to a Wilcoxon signed-rank test with $\alpha = 0.05$, and also depend on the fraction of the runtime sampled (cf.~Fig.~\ref{fig:annitraincolorselection}).
+However, the performance differences are not significant for a Wilcoxon signed-rank test with $\alpha = 0.05$ and also depend on the fraction of the runtime sampled (cf.~Fig.~\ref{fig:annitraincolorselection}).
 However, a clear advantage of our approach is that it indicates when to stop adding new instances.
@@
 
-A static benchmark utilizing the most frequently AL-sampled instances exhibits suboptimal performance when compared to active learning and random sampling.
-This outcome is somewhat anticipated, given that the static benchmark fails to reflect an appropriate balance of instance families.
+A static benchmark utilizing the most frequently AL-sampled instances exhibits suboptimal performance compared to active learning and random sampling.
+This outcome is anticipated since the static benchmark fails to reflect an appropriate balance of instance families.
 For example, families whose instances are uniformly randomly selected by AL, e.g., for different solvers, appear less frequently in this benchmark than families where some instances are sampled more often than others.
@@
     Scatter plot illustrating the relative importance of different instance families for ranking on the full dataset.
-The x-axis depicts the frequency of instance families in the dataset, while the y-axis represents the average frequency of instance families in the samples selected by active learning.
+The x-axis depicts the frequency of instance families in the dataset, while the y-axis represents the average frequency in the samples selected by active learning.
 The dashed line represents families that occur with the same frequency in the dataset and samples.
@@
 
-Fig.~\ref{fig:instoccs} illustrates the distribution of the frequency with which individual instances are selected across all active learning benchmarks.
+Fig.~\ref{fig:instoccs} illustrates the frequency distribution with which individual instances are selected across all active learning benchmarks.
 %JB: on the full dataset, I suppose (since tuning dataset does not contain all instances)? and what is "all" aggregating over? these information may also go into the figure's caption
-It can be observed that a small proportion of instances is selected with great frequency, while most instances are selected on only a few occasions.
+A small proportion of instances is selected with great frequency, while most instances are selected on only a few occasions.
 This indicates that there is no fixed subset of instances that can perfectly distinguish all solvers, which highlights the value of employing an active-learning-based strategy.
@@
 Fig.~\ref{fig:annifinalfamilies} illustrates the occurrence of instance families within the dataset and the benchmarks created by active learning.
-We use the best performing configurations for all values of $\delta \in \left[0, 1\right]$.
+We use the best-performing configurations for all values of $\delta \in \left[0, 1\right]$.
 While most families exhibit the same fraction in the dataset and the sampled benchmarks, a few outliers require further discussion.
@@
 
-From a technical perspective, we could formulate the problem of runtime discretization as an optimization problem, rather than addressing it empirically.
+From a technical perspective, we could formulate the problem of runtime discretization as an optimization problem rather than addressing it empirically.
 Further, a major shortcoming of our current approach is the lack of parallelization, selecting instances one at a time.
@@
 However, bigger batch sizes $n$ impede \emph{active learning}.
-Also, it is unclear how to synchronize instance selection and updates of the prediction model without wasting too much runtime.
+Also, it needs to be clarified how to synchronize instance selection and updates of the prediction model without wasting too much runtime.
 

### 2024-05-27 paper_jar/paper.tex
Add comments based on notebook "bonus_results"
Quick comparison of figures and plots in notebook with paper.
Revise paper

@@
 Our entire source code\footnote{\url{https://github.com/mathefuchs/al-for-sat-solver-benchmarking}} and experimental data\footnote{\url{https://github.com/mathefuchs/al-for-sat-solver-benchmarking-data}} are available on GitHub.
-%JB: data repo not updated since TACAS; does it also contain all the results for the new experiments? as a minimum, it should mention the journal version of the paper (see README of code repo)
-%JB: code repo currently has all JAR changes in a branch; probably best to push these changes to "main" for review (in case reviewers have a look at repo but don't switch branches)
 
@@
 
-  $\mathcal{M} \leftarrow \operatorname{initModel}\left(\mathcal{A},\, \mathcal{I},\, r,\, \hat{a}\right)$ \tcp*{cf. Section~\ref{sec:main:model}}
-  %JB: do we actually need \hat{a} here? (does not provide any information we could store/use here)
+  $\mathcal{M} \leftarrow \operatorname{initModel}\left(\mathcal{A},\, \mathcal{I},\, r\right)$ \tcp*{cf. Section~\ref{sec:main:model}}
 
@@
 
-The model $M$ provides a prediction function $f : \mathcal{\hat A} \times \mathcal{I} \rightarrow \mathbb{R}$ for all solvers $\mathcal{\hat A} := \mathcal{A} \cup \lbrace \hat a \rbrace$.
-%JB: shouldn't we drop the algorithm as an argument (for $f$ and later for $f'$), since we only train and predict for one solver ($\hat{a}$) anyway?
-(what would the predictions for the other solvers be -- looking up the exact runtime values from $r$?); instead, could use $\hat{a}$ as a subscript of $f$
+The model $M$ provides a prediction function $f_{\hat a} : \mathcal{I} \rightarrow \mathbb{R}$ for solver $\hat a$.
 This prediction function powers instance selection as described in Section~\ref{sec:main:selection}.
@@
 These methods require the model's predictions to include probabilities for the $k$ discrete runtime labels.
-Let \mbox{$f' : \mathcal{\hat A} \times \mathcal{I} \rightarrow \left[0, 1\right]^k$} denote this modified prediction function.
+Let $f_{\hat a}' : \mathcal{I} \rightarrow \left[0, 1\right]^k$ denote this modified prediction function.
 In the following, the set $\tilde{\mathcal{I}} \subseteq \mathcal{I}$ denotes the instances that have already been sampled.
@@
 \begin{equation*}
-  \operatorname{U}(e) := \left\lvert \frac{1}{k} - \max_{n \in \left\lbrace 1, \dots, k \right\rbrace} f'\!\left(\hat{a}, e\right)_{n} \right\rvert
+  \operatorname{U}(e) := \left\lvert \frac{1}{k} - \max_{n \in \left\lbrace 1, \dots, k \right\rbrace} f_{\hat{a}}'(e)_{n} \right\rvert
 \end{equation*}
@@
 \begin{equation*}
-  \operatorname{IG}(e) := \operatorname{H}(e) - \sum_{n = 1}^{k} f'(\hat{a}, e)_{n} \operatorname{\hat H}_n(e)
+  \operatorname{IG}(e) := \operatorname{H}(e) - \sum_{n = 1}^{k} f_{\hat{a}}'(e)_{n} \operatorname{\hat H}(e)_n
 \end{equation*}
 %
-In the equation, $\operatorname{H}(e)$ denotes the entropy of the runtime labels $\gamma(a, e)$ over all $a \in \mathcal{A}$ and $\operatorname{H}(e, n)$ denotes the entropy of these labels plus $n$ as the runtime label for $\hat{a}$.
-The term $\operatorname{\hat H}_n(e)$ is computed for every possible runtime label $n \in \{1, \dots, k\}$.
-By maximizing the information gain, instances are selected that identify solvers with similar behavior.
-%JB: "similar behavior" is a bit vague; strategy favors instances where the distribution of solvers over clusters is skewed rather than uniform, right?
-and the new solver only has little influence here, since its predicted labels is only one of |A| + 1 labels to determine entropy, right?
+In the equation, $\operatorname{H}(e)$ denotes the entropy of the runtime labels $\gamma(a, e)$ over all $a \in \mathcal{A}$ and $\operatorname{\hat H}(e)$ denotes the entropy of these labels plus the runtime label for $\hat{a}$.
+The term $\operatorname{\hat H}(e)_n$ is computed for every possible runtime label $n \in \{1, \dots, k\}$.
+By maximizing the information gain, instances are selected that identify solvers with similar runtime cluster labels, i.e., the new solver ranks similarly to the solvers in the same clusters.
 
@@
 This criterion is loosely inspired by Matricon et~al.~\cite{MatriconAFSH21}.
-To assess the statistical significance of the predicted runtime labels, the criterion uses the average $p$-value $W_{\hat{a}}$ of a Wilcoxon signed-rank test $\operatorname{w}(S,P)$ of the runtime label distributions ${S=\{ \gamma(a, e) \mid e \in \mathcal{I} \}}$ for an existing solver $a$ and \mbox{$P=\{ f(\hat a, e) \mid e \in \mathcal{I} \}$} of the new solver $\hat{a}$:
+To assess the statistical significance of the predicted runtime labels, the criterion uses the average $p$-value $W_{\hat{a}}$ of a Wilcoxon signed-rank test $\operatorname{w}(S,P)$ of the runtime label distributions ${S=\{ \gamma(a, e) \mid e \in \mathcal{I} \}}$ for an existing solver $a$ and \mbox{$P=\{ f_{\hat a}(e) \mid e \in \mathcal{I} \}$} of the new solver $\hat{a}$:
 %
@@
     $O_{\operatorname{acc}} \leftarrow 0$ \;    
-    \For{$a \in \mathcal{A}$}{
-      %JB: mabye Lines 6 and 8 should use A' instead of A (at the moment, solver is also compared to itself but this does not increase ranking accuracy since second factor in Line 7 is 0)
+    \For{$a \in \mathcal{A}'$}{
       \If{$\bigl(s_k(a) - s_{\hat a}\bigr) \cdot \bigl(\operatorname{par_2}(a) - \operatorname{par_2}(\hat a)\bigr) > 0$}{
-           $O_{\operatorname{acc}} \leftarrow O_{\operatorname{acc}} + \frac{1}{|\mathcal{A}|}$ \;
+           $O_{\operatorname{acc}} \leftarrow O_{\operatorname{acc}} + \frac{1}{|\mathcal{A}'|}$ \;
       }
@@
 For hyperparameter tuning, we randomly sample \SI{10}{\%} of the complete set of 5355 instances with stratification regarding the instance family.
-All instance families that are too \emph{small}, i.e., \SI{10}{\%} of them correspond to less than one instance, are put into one meta-family for stratification.
-%JB: i.e., families with less than 10 instances? could be phrased more directly
+All instance families that are too \emph{small}, i.e., families with less than 10 instances, are put into one meta-family for stratification.
 This \emph{tuning dataset} allows a more extensive exploration of the hyperparameter space.
@@
 We experiment with the following instantiations of the instance \emph{selection} (cf. Section~\ref{sec:main:selection}).
-Since the potential runtime of the solver experiments is orders of magnitude larger than the model update time, we increment our benchmark one instance at a time rather than using batches of multiple instances, which is also proposed in recent advances in active learning~\cite{SinhaED19,2019gaal}.
-%JB: not entirely clear if one-at-a-time "is also propsed" or batching "is also proposed"
+Since the potential runtime of the solver experiments is orders of magnitude larger than the model update time, we increment our benchmark one instance at a time rather than using batches of multiple instances.
+Both are used in the state-of-the-art in active learning~\cite{SinhaED19,2019gaal}.
 A drawback of this is the lack of parallel execution of runtime experiments.
@@
     \item Fallback threshold: Use random sampling for the first \SI{0}{\%}, \SI{5}{\%}, \SI{10}{\%}, \SI{15}{\%}, or \SI{20}{\%} of instances to explore the instance space.
-    \item Runtime scaling: Whether to normalize uncertainty scores per instance by the average runtime of solvers on it or use the absolute values.
-    %JB: but isn't the uncertainty score a difference of probabilities rather than directly runtime-related anyway?
-so probably not a normalization in the strict sense (maybe use term "divide" instead)
+    \item Runtime scaling: Whether to scale the uncertainty scores by the average runtime of solvers or use $\operatorname{U}(e)$ as is.
   \end{itemize}
@@
 In preliminary experiments, the most \emph{useful} labels were produced with hierarchical clustering and a log-single-linkage criterion; we will discuss our notion of label usefulness later.
-Other clustering approaches we have tried include hierarchical clustering with mean-, median-, and full-linkage criteria, as well as $k$-means and spectral clustering.
-%JB: What is "full" -- "complete"?
+Other clustering approaches we have tried include hierarchical clustering with mean-, median-, and complete-linkage criteria, as well as $k$-means and spectral clustering.
 
@@
 
-% TODO tfu: Amount of significant solver pairs (Wilcoxon signed-rank, 5% alpha), 
-% Amount of significant pairs:
+% Amount of significant solver pairs (Wilcoxon signed-rank, 5% alpha):
 % PAR-2:   684/756 = 0.9048
@@
 The matrices in Fig.~\ref{fig:solver-confusion-matrices} visualize which solver pairs exhibit significant differences in their PAR-2 scores, timeouts, and $s_3$ scores.
-According to a Wilcoxon signed-rank test with a significance level of $\alpha = 0.05$, \SI{87.83}{\%} of the solver pairs have significantly different scores after discretization, which is only a slight decrease compared to \SI{89.95}{\%} before discretization.
-%JB: numbers in notebook "bonus_results" differ
-%JB: does "after discretization" mean s_3 or timeout? and what is then the number for the other one?
+According to a Wilcoxon signed-rank test with a significance level of $\alpha = 0.05$, \SI{87.57}{\%} of the solver pairs have significantly different scores after $\gamma_3$-discretization, which is only a slight decrease compared to \SI{90.48}{\%} before discretization.
+$\gamma_2$-discretization significantly distinguishes \SI{87.30}{\%} of solver pairs.
 These results support the usefulness of discretized runtimes for our active learning framework.
@@
 This outcome is anticipated since the static benchmark fails to reflect an appropriate balance of instance families.
-For example, families whose instances are uniformly randomly selected by AL, e.g., for different solvers, appear less frequently in this benchmark than families where some instances are sampled more often than others.
-% MI: What? I don't understand this sentence.
 
@@
   \caption{Frequency with which individual problem instances are selected within all active-learning benchmarks.
-Instances that are chosen more frequently are intuitively more important for distinguishing solver runtimes.}
+Instances that are chosen more frequently are intuitively more important for distinguishing solver runtimes.
+Results are on the full dataset with cross-validation over all solvers.}
   \label{fig:instoccs}
@@
 The dashed line represents families that occur with the same frequency in the dataset and samples.
+Results are on the full dataset with cross-validation over all solvers.
   }
@@
 
-Fig.~\ref{fig:instoccs} illustrates the frequency distribution with which individual instances are selected across all active learning benchmarks.
-%JB: on the full dataset, I suppose (since tuning dataset does not contain all instances)? and what is "all" aggregating over? these information may also go into the figure's caption
+Fig.~\ref{fig:instoccs} illustrates the frequency distribution with which individual instances are selected across all active learning benchmarks regarding the full dataset with cross-validation over all solvers.
 A small proportion of instances is selected with great frequency, while most instances are selected on only a few occasions.
 This indicates that there is no fixed subset of instances that can perfectly distinguish all solvers, which highlights the value of employing an active-learning-based strategy.
-%JB: could also indicate that (1) there is such a set but AL cannot always find it or (2) there are multiple such sets and different AL runs return different sets
 

### 2024-05-27 paper_jar/paper.tex
minor changes

@@
 In contrast, the ranking stopping criterion is less sophisticated than the previous one.
-In this case, active learning terminates when the ranking induced by the model's predictions (Equation~\eqref{eq:rankingeq}) has remained unchanged within the last~$l$~iterations.
+In this case, active learning terminates when the ranking induced by the model's predictions (cf.~Equation~\eqref{eq:rankingeq}) has remained unchanged within the last~$l$~iterations.
 While the concrete values of the predicted score $s_{\hat a}$ may still change, the induced ranking is of sole interest in this criterion.
@@
 
-In our experiments, we work with the dataset of the SAT~Competition~2022 Anniversary Track~\cite{sat2022}.
+In our experiments, we work with the dataset of the SAT~Competition~2022 Anniversary Track~\cite{sat2022,zenodoanniversarytrack}.
 This dataset consists of 5355 instances with corresponding runtimes of 28 sequential SAT solvers.
-For predictions within our framework, we also use a database of 56 instance features\footnote{\url{https://benchmark-database.de/getdatabase/base_db}} from the Global Benchmark Database~(GBD) by Iser~et~al.~\cite{IserS18}.
+For predictions within our framework, we also use a database of 56 instance features\footnote{\url{https://benchmark-database.de/getdatabase/base_db}} from the Global Benchmark Database~(GBD)~\cite{GBD24paper}.
 They include instance size features and node distribution statistics for several graph representations of SAT instances, and are primarily inspired by the SATzilla~2012 features described in~\cite{features}.
@@
 For reproducibility, our source code and data are available on GitHub (see footnotes in Section~\ref{sec:intro}).
-Our code is implemented in \textsc{Python} using \emph{scikit-learn}~\cite{scikit-learn} for making predictions, and \emph{gbd-tools}~\cite{IserS18} for retrieving SAT instances.
+Our code is implemented in \textsc{Python} using \emph{scikit-learn}~\cite{scikit-learn} for making predictions, and \emph{gbd-tools}~\cite{GBD24paper} for retrieving SAT instances.
 

### 2024-05-27 paper_jar/paper.tex
Fix plot path

@@
 		\centering
-		\includegraphics[height=4.123cm]{../plots/cm2labels.pdf}
+		\includegraphics[height=4.123cm]{plots/cm2labels.pdf}
 		\caption{Timeout prediction.}
@@
 		\centering
-		\includegraphics[height=4.228cm]{../plots/cm3labels.pdf}
+		\includegraphics[height=4.228cm]{plots/cm3labels.pdf}
 		\caption{$C_3$ cluster prediction.}
@@
 		\centering
-		\includegraphics[width=3.9cm]{../plots/par2sigdiff.pdf}
+		\includegraphics[width=3.9cm]{plots/par2sigdiff.pdf}
 		\caption{PAR-2 scores.}
@@
 		\centering
-		\includegraphics[width=3.9cm]{../plots/timeoutsigdiff.pdf}
+		\includegraphics[width=3.9cm]{plots/timeoutsigdiff.pdf}
 		\caption{Timeouts.}
@@
 		\centering
-		\includegraphics[width=3.9cm]{../plots/s3sigdiff.pdf}
+		\includegraphics[width=3.9cm]{plots/s3sigdiff.pdf}
 		\caption{$s_3$ scores.}
@@
     \resizebox{\textwidth}{!}{
-      \graphicspath{{../plots/}}
-      \input{../plots/anni_train_color_ranking.pgf}
+      \graphicspath{{plots/}}
+      \input{plots/anni_train_color_ranking.pgf}
     }
@@
     \resizebox{\textwidth}{!}{
-      \graphicspath{{../plots/}}
-      \input{../plots/anni_train_color_selection.pgf}
+      \graphicspath{{plots/}}
+      \input{plots/anni_train_color_selection.pgf}
     }
@@
     \resizebox{\textwidth}{!}{
-      \graphicspath{{../plots/}}
-      \input{../plots/anni_train_color_stopping.pgf}
+      \graphicspath{{plots/}}
+      \input{plots/anni_train_color_stopping.pgf}
     }
@@
     \resizebox{\textwidth}{!}{
-      \graphicspath{{../plots/}}
-      \input{../plots/anni_train_optimization_goal.pgf}
+      \graphicspath{{plots/}}
+      \input{plots/anni_train_optimization_goal.pgf}
     }
@@
     \resizebox{\textwidth}{!}{
-      \graphicspath{{../plots/}}
-      \input{../plots/anni_train_delta_acc.pgf}
+      \graphicspath{{plots/}}
+      \input{plots/anni_train_delta_acc.pgf}
     }
@@
   \centering
-  \includegraphics[width=0.6\textwidth]{../plots/instoccs.pdf}
+  \includegraphics[width=0.6\textwidth]{plots/instoccs.pdf}
   \caption{Frequency with which individual problem instances are selected within all active-learning benchmarks.
@@
   \resizebox{0.85\textwidth}{!}{
-    \graphicspath{{../plots/}}
-    \input{../plots/anni_final_families.pgf}
+    \graphicspath{{plots/}}
+    \input{plots/anni_final_families.pgf}
   }

### 2024-05-27 paper_jar/paper.tex
Minor changes based on previous commits today

@@
 
-The model $M$ provides a prediction function $f_{\hat a} : \mathcal{I} \rightarrow \mathbb{R}$ for solver $\hat a$.
+The model $M$ provides a prediction function $f_{\hat a} : \mathcal{I} \rightarrow \mathbb{R}$ for the new solver $\hat a$.
 This prediction function powers instance selection as described in Section~\ref{sec:main:selection}.
-In the update step (Algorithm~\ref{algALBenchmark},~Line~7), the model is trained to predict a transformed version of solver runtime (cf.~Section~\ref{sec:main:model:transformation}) for the new solver~$\hat{a}$, using the features described in Section~\ref{sec:exdesign:data} and the previously acquired runtimes~$\mathcal{R}$.
+In the update step (Algorithm~\ref{algALBenchmark},~Line~7), the model is trained to predict a transformed version of solver runtime (cf.~Section~\ref{sec:main:model:transformation}) for the solver~$\hat{a}$, using the features described in Section~\ref{sec:exdesign:data} and the previously acquired runtimes~$\mathcal{R}$.
 Each iteration trains a new prediction model from scratch since the time for running the solver~$\hat{a}$ on one instance~$e$ (Line~5) dominates the training time of the model~$M$ on all acquired instances~$\mathcal{R}$ by a significant margin.
-Once the benchmarking loop terminates, the score of the new solver~$\hat{a}$ is predicted and then returned (Lines~8--9).
+Once the benchmarking loop terminates, the score of the solver~$\hat{a}$ is predicted and then returned (Lines~8--9).
 
@@
 Since the potential runtime of the solver experiments is orders of magnitude larger than the model update time, we increment our benchmark one instance at a time rather than using batches of multiple instances.
-Both are used in the state-of-the-art in active learning~\cite{SinhaED19,2019gaal}.
-A drawback of this is the lack of parallel execution of runtime experiments.
+Both these strategies are used in the state-of-the-art in active learning~\cite{SinhaED19,2019gaal}.
+A drawback of selecting instances one at a time is the lack of parallel execution of runtime experiments.
 
@@
     \item Fallback threshold: Use random sampling for the first \SI{0}{\%}, \SI{5}{\%}, \SI{10}{\%}, \SI{15}{\%}, or \SI{20}{\%} of instances to explore the instance space.
-    \item Runtime scaling: Whether to scale the uncertainty scores by the average runtime of solvers or use $\operatorname{U}(e)$ as is.
+    \item Runtime scaling: Whether to scale the uncertainty scores by the average runtime of solvers per instance or use $\operatorname{U}(e)$ as is.
   \end{itemize}
@@
     \item Fallback threshold: Use random sampling for the first \SI{0}{\%}, \SI{5}{\%}, \SI{10}{\%}, \SI{15}{\%}, or \SI{20}{\%} of instances to explore the instance space.
-    \item Runtime scaling: Whether to normalize information-gain scores per instance by the average runtime of solvers on it or use the absolute values.
+    \item Runtime scaling: Whether to scale the information-gain scores by the average runtime of solvers per instance or use $\operatorname{IG}(e)$ as is.
   \end{itemize}

### 2024-05-28 paper_jar/paper.tex
Minor changes

@@
 The term $\operatorname{\hat H}(e)_n$ is computed for every possible runtime label $n \in \{1, \dots, k\}$.
-By maximizing the information gain, instances are selected that identify solvers with similar runtime cluster labels, i.e., the new solver ranks similarly to the solvers in the same clusters.
+By maximizing the information gain, one favors instances whose most likely cluster labels, as determined by $f_{\hat a}'$, are similar to a great number of existing solvers.
+This decreases the entropy among the cluster labels of the given instance and we know that the new solver performs similarly to a number of solvers.
 

### 2024-05-28 paper_jar/paper.tex
Minor changes

@@
 		\toprule
-		\multirow[c]{2}{*}{Timeout Prediction Models} & \multicolumn{2}{c}{Average MCC ($\pm$ std)} \\
+		\multirow[c]{2}{*}[-0.27em]{Timeout Prediction Models} & \multicolumn{2}{c}{Average MCC ($\pm$ std)} \\
 		\cmidrule(lr){2-3}
@@
     \toprule
-    \multirow[c]{2}{*}{$C_3$ Cluster Prediction Models} & \multicolumn{2}{c}{Average MCC ($\pm$ std)} \\
+    \multirow[c]{2}{*}[-0.27em]{$C_3$ Cluster Prediction Models} & \multicolumn{2}{c}{Average MCC ($\pm$ std)} \\
     \cmidrule(lr){2-3}
@@
     \midrule
-    \multirow[c]{2}{*}{$C_4$ Cluster Prediction Models} & \multicolumn{2}{c}{Average MCC ($\pm$ std)} \\
+    \multirow[c]{2}{*}[-0.27em]{$C_4$ Cluster Prediction Models} & \multicolumn{2}{c}{Average MCC ($\pm$ std)} \\
     \cmidrule(lr){2-3}
@@
     \midrule
-    \multirow[c]{2}{*}{$C_5$ Cluster Prediction Models} & \multicolumn{2}{c}{Average MCC ($\pm$ std)} \\
+    \multirow[c]{2}{*}[-0.27em]{$C_5$ Cluster Prediction Models} & \multicolumn{2}{c}{Average MCC ($\pm$ std)} \\
     \cmidrule(lr){2-3}
@@
   \centering
-  \resizebox{0.85\textwidth}{!}{
+  \resizebox{0.75\textwidth}{!}{
     \graphicspath{{plots/}}

### 2025-02-10 paper_jar/paper.tex
Address eleven minor points of R2
- definition of new-solver problem
- ranking via binary classification
- "binarily" -> "pairwise"
- separate benchmark
- "measured" -> "estimated"
- \mathcal{M}
- clustering algorithm
- "particular" -> "this way"
- stopping criteria
- product > 0
- "ensembled models"

@@
 
+% Highlight changes for re-submission (might be removed later)
+\usepackage{soul}
+\soulregister\cite7 % some hacking to make highlighting work in special situations, e.g., if text contains references: https://tex.stackexchange.com/q/139463
+\soulregister\ref7
+\soulregister\pageref7
+\newcommand{\showchanges}{1} % set to 0 or 1
+\newcommand{\change}[1]{\ifnum \showchanges > 0 \sethlcolor{yellow}\hl{#1} \else #1 \fi}
+
 \raggedbottom
@@
   Let us consider a set of solvers~$\mathcal{A}$, instances~$\mathcal{I}$, and runtimes~${r\!: \mathcal{A} \times \mathcal{I} \rightarrow \left[0, \tau\right]}$, with each solver having a runtime limit of $\tau$.
-  The \emph{New-Solver Problem} is about incrementally selecting benchmark instances from $\mathcal{I}$ to maximize the confidence in predicting the rank of a new solver $\hat{a} \notin \mathcal{A}$, while minimizing the total benchmark runtime.
+  The \emph{New-Solver Problem} is \change{about selecting} benchmark instances from $\mathcal{I}$ to maximize the confidence in predicting the rank of a new solver $\hat{a} \notin \mathcal{A}$, while minimizing the \change{summed runtime of~$\hat{a}$ over all selected instances}.
   \label{def:new-solver-problem}
@@
 
+\change{
+The two objectives of this problem are conflicting, i.e., the ranking's confidence typically benefits from more instances while the runtime criterion encourages less or at least shorter evaluations.
+The crux is finding a good trade-off.
+}
+
 The approach presented in this paper meets several criteria for benchmarking that are considered desirable (cf.~Table~\ref{tab:requirements}).
 Rather than outputting a binary classification, namely whether the new solver is worse than an existing solver or not, we provide a \emph{scoring} function that shows by which margin a solver is worse and how similar its performance is to existing solvers.
-In particular, our approach enables \emph{ranking} the new solver among a set of existing solvers.
+In particular, our approach enables \emph{ranking} the new solver among a set of existing solvers \change{in one benchmarking run}.
 We show that predicting the exact per-instance runtime of the solvers, which would be a more challenging task, is not necessary for ranking.
@@
 Their approach ticks off most of the desired features in Table~\ref{tab:requirements}.
-However, the approach only compares solvers binarily rather than providing a multi-solver \emph{scoring} or \emph{ranking}.
+However, the approach only compares solvers \change{pairwise} rather than providing a multi-solver \emph{scoring} or \emph{ranking}.
 Consequently, it is unclear how similar two given solvers are or on which instances they behave similarly.
 Furthermore, a significant shortcoming is the lack of \emph{scalability} with the number of solvers.
-Due to comparing only pairs of solvers, evaluating a new solver with their method requires sampling a separate benchmark for each existing solver.
-In contrast, our approach allows comparing a new solver against a set of existing solvers by sampling only one benchmark.
+Due to comparing only pairs of solvers, evaluating a new solver with their method requires sampling \change{an individual set of benchmark instances} for each existing solver.
+\change{In particular, comparing against one existing solver will typically require a different set of instances than comparing against another one.}
+In contrast, our approach allows comparing a new solver against a set of existing solvers by sampling only one \change{set of} benchmark \change{instances}.
 
@@
 In our scenario, the first goal is to maximize the utility of instances to our prediction model, which is measured by the ranking accuracy.
-Our second goal is to minimize the cost of label acquisition, which is measured by the instance's predicted runtime.
+Our second goal is to minimize the cost of label acquisition, which is \change{estimated} by the instance's predicted runtime.
 Consequently, our overall objective is to develop an accurate prediction model without labeling each data point.
@@
 
-The model $M$ provides a prediction function $f_{\hat a} : \mathcal{I} \rightarrow \mathbb{R}$ for the new solver $\hat a$.
+The model~\change{$\mathcal{M}$} provides a prediction function $f_{\hat a} : \mathcal{I} \rightarrow \mathbb{R}$ for the new solver $\hat a$.
 This prediction function powers instance selection as described in Section~\ref{sec:main:selection}.
 In the update step (Algorithm~\ref{algALBenchmark},~Line~7), the model is trained to predict a transformed version of solver runtime (cf.~Section~\ref{sec:main:model:transformation}) for the solver~$\hat{a}$, using the features described in Section~\ref{sec:exdesign:data} and the previously acquired runtimes~$\mathcal{R}$.
-Each iteration trains a new prediction model from scratch since the time for running the solver~$\hat{a}$ on one instance~$e$ (Line~5) dominates the training time of the model~$M$ on all acquired instances~$\mathcal{R}$ by a significant margin.
+Each iteration trains a new prediction model from scratch since the time for running the solver~$\hat{a}$ on one instance~$e$ (Line~5) dominates the training time of the model~\change{$\mathcal{M}$} on all acquired instances~$\mathcal{R}$ by a significant margin.
 Once the benchmarking loop terminates, the score of the solver~$\hat{a}$ is predicted and then returned (Lines~8--9).
@@
 
-For the prediction model~$M$, we transform the real-valued runtimes into discrete runtime labels on a per-instance basis.
+For the prediction model~\change{$\mathcal{M}$}, we transform the real-valued runtimes into discrete runtime labels on a per-instance basis.
 For each instance $e \in \mathcal{I}$, we use a clustering algorithm to assign the runtimes in $\bigl\{ r(a, e) \mid a \in \mathcal{A} \bigr\}$ to one of $k$ clusters $C_1, \dots, C_k$ such that the fastest runtimes for the benchmark instance $e$ are in cluster $C_1$ and the slowest are in cluster $C_{k-1}$.
@@
 Empirical studies on portfolio solvers have demonstrated that discretization is an effective approach in practice~\cite{CollauttiMMO13,NgokoCT19}.
-Section~\ref{sec:exdesign:disc-pred} presents a detailed analysis of the runtime transformation in our scenario.
+Section~\ref{sec:exdesign:disc-pred} presents a detailed analysis of the runtime transformation in our scenario\change{, including the clustering approach}.
 In particular, the analysis demonstrates that our benchmarking approach achieves higher ranking accuracy when working with discrete runtime labels rather than raw runtimes.
@@
 %
-In particular, timeouts are still penalized.
+In \change{this way}, timeouts are still penalized.
 
@@
 In this section, we present the two dynamic stopping criteria (cf.~Algorithm~\ref{algALBenchmark}, Line~3) employed in our experiments: the Wilcoxon and the ranking stopping criterion.
+\change{
+In the worst case, both criteria sample all instances, but typically they should terminate active learning considerably earlier.
+How early they terminate depends on how easy it is to determine the rank of the new solver among the existing ones.
+}
 
@@
     \For{$a \in \mathcal{A}'$}{
-      \If{$\bigl(s_k(a) - s_{\hat a}\bigr) \cdot \bigl(\operatorname{par_2}(a) - \operatorname{par_2}(\hat a)\bigr) > 0$}{
+      \If(\tcp*[f]{\change{same sign/order}}){\change{$\operatorname{sgn}(s_k(a) - s_{\hat a}) = \operatorname{sgn}(\operatorname{par_2}(a) - \operatorname{par_2}(\hat a))$}}{
            $O_{\operatorname{acc}} \leftarrow O_{\operatorname{acc}} + \frac{1}{|\mathcal{A}'|}$ \;
@@
 However, the best prediction performance with all features is achieved by a stacking ensemble~\cite{wolpert1992stacked} out of the quadratic discriminant analysis~\cite{tharwat2016linear} and the random forest~\cite{breiman2001random}.
-Stacking means that another model, in our case a simple decision tree~\cite{breiman1984classification}, makes a prediction based on the predictions of the ensembled models.
+Stacking means that another model, in our case a simple decision tree~\cite{breiman1984classification}, \change{acts as a meta-model:
+The prediction outputs of the ensemble members, i.e., quadratic discriminant analysis and random forests, become the prediction inputs of the meta-model, i.e., decision tree.
+}
 

### 2025-02-12 paper_jar/paper.tex
Fix figure alignment
Address comment regarding similar runtimes

@@
 \newcommand{\showchanges}{1} % set to 0 or 1
-\newcommand{\change}[1]{\ifnum \showchanges > 0 \sethlcolor{yellow}\hl{#1} \else #1 \fi}
+\newcommand{\change}[1]{\ifnum \showchanges > 0 \sethlcolor{yellow}\hl{#1}\else#1\fi}
 
@@
 
-\subparagraph{Motivation}
-
 One of the main phases of algorithm engineering is benchmarking.
@@
 Rather than outputting a binary classification, namely whether the new solver is worse than an existing solver or not, we provide a \emph{scoring} function that shows by which margin a solver is worse and how similar its performance is to existing solvers.
+\change{This helps to identify the solvers that have the most similar runtimes more quickly.}
 In particular, our approach enables \emph{ranking} the new solver among a set of existing solvers \change{in one benchmarking run}.
@@
 
-\subparagraph{Experiments}
-
-We evaluated our approach using the SAT Competition~2022 Anniversary Track dataset~\cite{sat2022}, consisting of 5355~instances and runtimes of 28~solvers.
-Cross-validation was performed by treating each solver as the new solver once and learning to predict its PAR-2 rank.
-On average, our predictions achieve a ranking accuracy of approximately \SI{92}{\%} with only approximately \SI{10}{\%} of the runtime required to evaluate these solvers on the complete set of instances.
-Our entire source code\footnote{\url{https://github.com/mathefuchs/al-for-sat-solver-benchmarking}} and experimental data\footnote{\url{https://github.com/mathefuchs/al-for-sat-solver-benchmarking-data}} are available on GitHub.
-
-\subparagraph{Outline}
-
-The remainder of this paper is organized as follows:
-Section~\ref{sec:related} reviews related work.
-Section~\ref{sec:main} describes our active learning approach.
-Section~\ref{sec:exdesign} introduces the experimental design.
-Section~\ref{sec:eval} evaluates the experimental results.
-Section~\ref{sec:conclusion} concludes.
-
-\subparagraph{Disclaimer}
-
-This article is an extended and refined version of our conference paper titled ``Active Learning for SAT Solver Benchmarking''~\cite{fuchs2023active}.
-In particular, we significantly extended the analysis of runtime-prediction and runtime-discretization approaches (cf.~Section~\ref{sec:exdesign:disc-pred}), which determined our chosen solver runtime model for instance selection.
-Furthermore, we expanded the evaluation of which benchmark instances are selected by our approach (cf.~Section~\ref{sec:eval:instance}).
-
-\section{Related Work}
-\label{sec:related}
-
-Benchmarking is of great interest in numerous research areas and represents an active field of research in its own right.
-Studies have shown that the compilation of benchmark instances poses several challenges.
-Using biased benchmarks can easily lead to fallacious interpretations~\cite{abs-2107-07002}.
-Benchmarking also involves several interchangeable elements, including the performance measures used, how the measures are aggregated, and how missing values are handled.
-Questionable research practices could alter these elements a-posteriori to meet expectations, thereby skewing the results~\cite{NiesslHWCB22}.
-
-The following sections discuss related work from the areas of static benchmark selection (cf.~Section~\ref{sec:related:static-selection}), algorithm configuration (cf.~Section~\ref{sec:related:algo-configuration}), incremental benchmark selection (cf.~Section~\ref{sec:related:incremental-selection}), and active learning (cf.~Section~\ref{sec:related:active-learning}).
-Table~\ref{tab:requirements} compares the most relevant approaches, which all pursue slightly different goals.
-In particular, our approach is \emph{not} a general improvement over the others but the only one fully aligned with Definition~\ref{def:new-solver-problem}.
-
-\subsection{Static Benchmark Selection}
-\label{sec:related:static-selection}
-
-Benchmark selection is essential for solver competitions like the SAT Competition.
-In such competitions, the organizers define the rules for composing the benchmarks.
-These selection strategies are primarily static since they do not depend on particular solvers to distinguish.
-Balint et al. provide an overview of benchmark-selection criteria in different solver competitions~\cite{balint2015overview}.
-Froleyks et al. describe benchmark selection in recent SAT~competitions~\cite{FroleyksHIJS21}.
-Manthey and Möhle propose an approach to remove redundancy from competition benchmarks by considering feature equivalence of formulas~\cite{manthey2016better}.
-M{\i}s{\i}r also presents a feature-based approach to reduce benchmarks by using matrix factorization and clustering~\cite{misir2021benchmark}.
-
-Hoos et al.~\cite{HoosKSS13} discuss desirable properties of SAT benchmark instances.
-They identify three key selection criteria: instance variety to avoid over-fitting, adapted instance hardness (neither too easy nor too hard), and avoiding duplicate instances. 
-To filter instances that are too similar, they employ a distance-based approach with the SATzilla features~\cite{XuHHL08,features}.
-It should be noted, however, that the approach does not optimize for benchmark \emph{runtime}.
-Instead, instances are selected \emph{at random} except considering constraints on instance hardness and feature distance.
-
 \begin{table}[tbp]
@@
 \end{table}
+
+\subparagraph{Experiments}
+
+We evaluated our approach using the SAT Competition~2022 Anniversary Track dataset~\cite{sat2022}, consisting of 5355~instances and runtimes of 28~solvers.
+Cross-validation was performed by treating each solver as the new solver once and learning to predict its PAR-2 rank.
+On average, our predictions achieve a ranking accuracy of approximately \SI{92}{\%} with only approximately \SI{10}{\%} of the runtime required to evaluate these solvers on the complete set of instances.
+Our entire source code\footnote{\url{https://github.com/mathefuchs/al-for-sat-solver-benchmarking}} and experimental data\footnote{\url{https://github.com/mathefuchs/al-for-sat-solver-benchmarking-data}} are available on GitHub.
+
+\subparagraph{Outline}
+
+The remainder of this paper is organized as follows:
+Section~\ref{sec:related} reviews related work.
+Section~\ref{sec:main} describes our active learning approach.
+Section~\ref{sec:exdesign} introduces the experimental design.
+Section~\ref{sec:eval} evaluates the experimental results.
+Section~\ref{sec:conclusion} concludes.
+
+\subparagraph{Disclaimer}
+
+This article is an extended and refined version of our conference paper titled ``Active Learning for SAT Solver Benchmarking''~\cite{fuchs2023active}.
+In particular, we significantly extended the analysis of runtime-prediction and runtime-discretization approaches (cf.~Section~\ref{sec:exdesign:disc-pred}), which determined our chosen solver runtime model for instance selection.
+Furthermore, we expanded the evaluation of which benchmark instances are selected by our approach (cf.~Section~\ref{sec:eval:instance}).
+
+\section{Related Work}
+\label{sec:related}
+
+Benchmarking is of great interest in numerous research areas and represents an active field of research in its own right.
+Studies have shown that the compilation of benchmark instances poses several challenges.
+Using biased benchmarks can easily lead to fallacious interpretations~\cite{abs-2107-07002}.
+Benchmarking also involves several interchangeable elements, including the performance measures used, how the measures are aggregated, and how missing values are handled.
+Questionable research practices could alter these elements a-posteriori to meet expectations, thereby skewing the results~\cite{NiesslHWCB22}.
+
+The following sections discuss related work from the areas of static benchmark selection (cf.~Section~\ref{sec:related:static-selection}), algorithm configuration (cf.~Section~\ref{sec:related:algo-configuration}), incremental benchmark selection (cf.~Section~\ref{sec:related:incremental-selection}), and active learning (cf.~Section~\ref{sec:related:active-learning}).
+Table~\ref{tab:requirements} compares the most relevant approaches, which all pursue slightly different goals.
+In particular, our approach is \emph{not} a general improvement over the others but the only one fully aligned with Definition~\ref{def:new-solver-problem}.
+
+\subsection{Static Benchmark Selection}
+\label{sec:related:static-selection}
+
+Benchmark selection is essential for solver competitions like the SAT Competition.
+In such competitions, the organizers define the rules for composing the benchmarks.
+These selection strategies are primarily static since they do not depend on particular solvers to distinguish.
+Balint et al. provide an overview of benchmark-selection criteria in different solver competitions~\cite{balint2015overview}.
+Froleyks et al. describe benchmark selection in recent SAT~competitions~\cite{FroleyksHIJS21}.
+Manthey and Möhle propose an approach to remove redundancy from competition benchmarks by considering feature equivalence of formulas~\cite{manthey2016better}.
+M{\i}s{\i}r also presents a feature-based approach to reduce benchmarks by using matrix factorization and clustering~\cite{misir2021benchmark}.
+
+Hoos et al.~\cite{HoosKSS13} discuss desirable properties of SAT benchmark instances.
+They identify three key selection criteria: instance variety to avoid over-fitting, adapted instance hardness (neither too easy nor too hard), and avoiding duplicate instances. 
+To filter instances that are too similar, they employ a distance-based approach with the SATzilla features~\cite{XuHHL08,features}.
+It should be noted, however, that the approach does not optimize for benchmark \emph{runtime}.
+Instead, instances are selected \emph{at random} except considering constraints on instance hardness and feature distance.
 
@@
 \label{sec:related:active-learning}
-
-Our benchmark selection approach uses active learning.
-In passive machine learning, prediction models are trained on datasets with given instance labels (cf.~Fig.~\ref{fig:passive}).
-In contrast, active learning (AL) begins with little or no labeled data.
-AL then repeatedly selects interesting problem instances for which to acquire labels, gradually improving the prediction model (cf.~Fig.~\ref{fig:activepool}).
-AL methods are particularly advantageous when acquiring labels is computationally expensive, such as obtaining solver runtimes.
-Without AL methods, it is not evident which instances to label and which not.
-In our scenario, the first goal is to maximize the utility of instances to our prediction model, which is measured by the ranking accuracy.
-Our second goal is to minimize the cost of label acquisition, which is \change{estimated} by the instance's predicted runtime.
-Consequently, our overall objective is to develop an accurate prediction model without labeling each data point.
-
-Rubens et~al.~\cite{RubensESK15} survey active-learning advances.
-While synthesis-based AL methods~\cite{0001AEMN22,GarzonMG22,2019gaal} generate instances for labeling, pool-based methods~\cite{GolbandiKL11,HarpaleY08,KorenBV09} rely on a fixed set of unlabeled instances to sample from.
-Recent synthesis-based methods within the field of SAT solving demonstrate the generation of problem instances with desired properties~\cite{0001AEMN22,GarzonMG22}.
-However, this objective is distinct from ours.
-While those approaches aim to generate instances on which a solver is either effective or ineffective, our objective is to predict whether a solver is effective or ineffective on an \emph{existing} benchmark.
-In the latter direction, Volpato and Guangyan employ pool-based AL to train an instance-specific algorithm selector~\cite{volpato2019active}.
-However, their objective is not to benchmark a solver's overall performance but to recommend the best solver for each SAT instance from a set of available solvers.
 
@@
 
+Our benchmark selection approach uses active learning.
+In passive machine learning, prediction models are trained on datasets with given instance labels (cf.~Fig.~\ref{fig:passive}).
+In contrast, active learning (AL) begins with little or no labeled data.
+AL then repeatedly selects interesting problem instances for which to acquire labels, gradually improving the prediction model (cf.~Fig.~\ref{fig:activepool}).
+AL methods are particularly advantageous when acquiring labels is computationally expensive, such as obtaining solver runtimes.
+Without AL methods, it is not evident which instances to label and which not.
+In our scenario, the first goal is to maximize the utility of instances to our prediction model, which is measured by the ranking accuracy.
+Our second goal is to minimize the cost of label acquisition, which is \change{estimated} by the instance's predicted runtime.
+Consequently, our overall objective is to develop an accurate prediction model without labeling each data point.
+
+Rubens et~al.~\cite{RubensESK15} survey active-learning advances.
+While synthesis-based AL methods~\cite{0001AEMN22,GarzonMG22,2019gaal} generate instances for labeling, pool-based methods~\cite{GolbandiKL11,HarpaleY08,KorenBV09} rely on a fixed set of unlabeled instances to sample from.
+Recent synthesis-based methods within the field of SAT solving demonstrate the generation of problem instances with desired properties~\cite{0001AEMN22,GarzonMG22}.
+However, this objective is distinct from ours.
+While those approaches aim to generate instances on which a solver is either effective or ineffective, our objective is to predict whether a solver is effective or ineffective on an \emph{existing} benchmark.
+In the latter direction, Volpato and Guangyan employ pool-based AL to train an instance-specific algorithm selector~\cite{volpato2019active}.
+However, their objective is not to benchmark a solver's overall performance but to recommend the best solver for each SAT instance from a set of available solvers.
+
 \section{Active Learning for SAT Solver Benchmarking}
@@
 \change{
-In the worst case, both criteria sample all instances, but typically they should terminate active learning considerably earlier.
+In the worst case, both criteria sample all instances, but typically they terminate earlier.
 How early they terminate depends on how easy it is to determine the rank of the new solver among the existing ones.
@@
 % * Spearman correlation of PAR2 and Discrete rank: 0.9862
-\begin{table}[tb]
+\begin{table}[p]
 	\centering
@@
 % k=5:     662/756 = 0.8757
-\begin{figure}[tb]
+\begin{figure}[p]
 	\centering

### 2025-02-12 paper_jar/paper.tex
Add appendix

@@
 This dataset consists of 5355 instances with corresponding runtimes of 28 sequential SAT solvers.
-For predictions within our framework, we also use a database of 56 instance features\footnote{\url{https://benchmark-database.de/getdatabase/base_db}} from the Global Benchmark Database~(GBD)~\cite{GBD24paper}.
+For predictions within our framework, we also use a database of 56 instance features from the Global Benchmark Database~(GBD)~\cite{IserJ24}.
+\change{Refer to Appendix~\ref{sec:appendix-features} for a full list of all 56 features.}
 They include instance size features and node distribution statistics for several graph representations of SAT instances, and are primarily inspired by the SATzilla~2012 features described in~\cite{features}.
@@
 
-In addition, we retrieve instance family information\footnote{\url{https://benchmark-database.de/getdatabase/meta_db}} to evaluate the composition of our sampled benchmarks.
+In addition, we retrieve instance family information\footnote{\url{https://benchmark-database.de/getdatabase/meta}} to evaluate the composition of our sampled benchmarks.
 Instance families consist of instances from the same application domain, such as planning, cryptography, etc., and are valuable for analyzing solver performance.
-%
 For hyperparameter tuning, we randomly sample \SI{10}{\%} of the complete set of 5355 instances with stratification regarding the instance family.
@@
 We employ these models to estimate a new solver's raw runtime as the prediction target.
-To this end, we use 46 instance features and 27 known solver runtimes as inputs for the predictions.
+To this end, we use 46 instance features \change{(Section~\ref{sec:exdesign:data})} and 27 known solver runtimes as inputs for the predictions.
 We repeat this evaluation for each solver as the prediction target and report the average and standard deviation of prediction performance over the 28 solvers.
@@
 Table~\ref{tab:cluster-prediction} shows the prediction performance for the best classification models from Table~\ref{tab:timeout-prediction} and $k \in \lbrace 3, 4, 5 \rbrace$ clusters.
-We train the prediction models either with 46 instance features and 27 known solver runtimes (second column) or additionally including 27 features representing the known $C_k$ cluster labels of other solvers (third column).
+We train the prediction models either with 46 instance features \change{(Section~\ref{sec:exdesign:data})} and 27 known solver runtimes (second column) or additionally including 27 features representing the known $C_k$ cluster labels of other solvers (third column).
 
@@
 For reproducibility, our source code and data are available on GitHub (see footnotes in Section~\ref{sec:intro}).
-Our code is implemented in \textsc{Python} using \emph{scikit-learn}~\cite{scikit-learn} for making predictions, and \emph{gbd-tools}~\cite{GBD24paper} for retrieving SAT instances.
+Our code is implemented in \textsc{Python} using \emph{scikit-learn}~\cite{scikit-learn} for making predictions, and \emph{gbd-tools}~\cite{IserJ24} for retrieving SAT instances.
 
@@
 
+%
+% ---- Appendix ----
+%
+
+\newpage
+\appendix
+\section{Instance Features}
+\label{sec:appendix-features}
+
+\change{For predictions within our framework, we use a database of 56 instance features\footnote{\url{https://udopia.github.io/gbdc/doc/extractors/CNFBaseFeatures.html}} from the Global Benchmark Database~(GBD)~\cite{IserJ24}.
+Table~\ref{tab:features} briefly summarizes all 56 features.
+The placeholder \texttt{<n>} can take values within the range 1 to 9 and the placeholder \texttt{<dist>} is one of \texttt{mean}, \texttt{variance}, \texttt{min}, \texttt{max}, or \texttt{entropy}.
+Accounting for all possible instantiations of \texttt{<n>} and \texttt{<dist>}, there are 56 features in total.}
+
+\begin{table}[h]
+    \centering
+    \caption{
+      \change{List of all 56 instance features, where \texttt{<dist>} is one of \texttt{mean}, \texttt{variance}, \texttt{min}, \texttt{max}, or \texttt{entropy}.}
+    }
+    \label{tab:features}
+    \begin{tabular}{
+      >{\arraybackslash}p{0.22\textwidth}
+      >{\arraybackslash}p{0.21\textwidth}
+      >{\arraybackslash}p{0.46\textwidth}
+    }
+      \toprule
+      Category & Feature Name & Description \\
+      \midrule
+      Basic statistics & \verb|clauses| & Number of clauses \\
+      & \verb|variables| & Number of variables \\
+      \midrule
+      Clause types & \verb|cls<n>| & Number of clauses with \verb|<n>| literals for \verb|<n>| in 1--9 \\
+      & \verb|cls10p| & Number of clauses with 10 or more literals \\
+      & \verb|horn| & Number of Horn clauses \\
+      & \verb|invhorn| & Number of inverted Horn clauses \\
+      & \verb|positive| & Number of positive clauses \\
+      & \verb|negative| & Number of negative clauses \\
+      \midrule
+      Horn proximity & \verb|hornvars_<dist>| & Distribution of variable occurrence in horn clauses \\
+      & \verb|invhornvars_<dist>| & Distribution of variable occurrence in inverted horn clauses \\
+      \midrule
+      Polarity balance & \verb|balancecls_<dist>| & Distributions of fractions of number of positive/negative literals in clauses \\
+      & \verb|balancevars_<dist>| & Distributions of fractions of number of positive/negative literals of variables \\
+      \midrule
+      Variable-clause graph & \verb|vcg_vdegree_<dist>| & Variable degree distribution \\
+      & \verb|vcg_cdegree_<dist>| & Clause degree distribution \\
+      \midrule
+      Variable graph & \verb|vg_degree_<dist>| & Degree distribution \\
+      \midrule
+      Clause graph & \verb|cg_degree_<dist>| & Degree distribution \\
+      \bottomrule
+    \end{tabular}
+\end{table}
+
 \end{document}

### 2025-02-12 paper_jar/paper.tex
Describe sampling strategies

@@
 Timeouts $\tau$ always form a separate cluster $C_{k}$.
+\change{The clustering is performed on a per-instance level as the best runtimes may be seconds on one instance and minutes on another.
+After clustering, the best runtimes are all in the same cluster.}
 The runtime transformation function $\gamma_k : {\mathcal{A} \times \mathcal{I}} \rightarrow \left\lbrace 1, \dots, k \right\rbrace$ is then specified as follows:
@@
 \begin{equation*}
-  \operatorname{U}(e) := \left\lvert \frac{1}{k} - \max_{n \in \left\lbrace 1, \dots, k \right\rbrace} f_{\hat{a}}'(e)_{n} \right\rvert
+  \operatorname{U}(e) := \left\lvert \frac{1}{k} - \max_{n \in \left\lbrace 1, \dots, k \right\rbrace} f_{\hat{a}}'(e)_{n} \right\rvert \text{.}
 \end{equation*}
+\change{Intuitively, uncertainty sampling selects the instances where it is most unclear in which runtime cluster the new solver resides.
+For example, if the new solver behaves similarly to two known solvers on the already run experiments, it is likely that this sampling strategy selects instances where these two solvers have runtimes that are not within the same cluster.}
 
@@
 \begin{equation*}
-  \operatorname{IG}(e) := \operatorname{H}(e) - \sum_{n = 1}^{k} f_{\hat{a}}'(e)_{n} \operatorname{\hat H}(e)_n
+  \operatorname{IG}(e) := \operatorname{H}(e) - \sum_{n = 1}^{k} f_{\hat{a}}'(e)_{n} \operatorname{\hat H}(e)_n \text{.}
 \end{equation*}
@@
 The term $\operatorname{\hat H}(e)_n$ is computed for every possible runtime label $n \in \{1, \dots, k\}$.
-By maximizing the information gain, one favors instances whose most likely cluster labels, as determined by $f_{\hat a}'$, are similar to a great number of existing solvers.
-This decreases the entropy among the cluster labels of the given instance and we know that the new solver performs similarly to a number of solvers.
+\change{Intuitively, this strategy favors instances where the most likely predicted cluster labels (as determined by $f_{\hat a}'$) would provide the most insight regarding similarities to known solvers.
+For example, on a particular instance, one group of solvers might run into timeouts and another group of solvers might return a solution.
+In this case, the information-gain strategy will likely select this instance as it provides insight about the group of solvers the new solver is similar to.}
 
@@
     \For{$a \in \mathcal{A}'$}{
-      \If(\tcp*[f]{\change{same sign/order}}){\change{$\operatorname{sgn}(s_k(a) - s_{\hat a}) = \operatorname{sgn}(\operatorname{par_2}(a) - \operatorname{par_2}(\hat a))$}}{
+      \If(\tcp*[f]{\change{Same sign/order}}){\change{$\operatorname{sgn}(s_k(a) - s_{\hat a}) = \operatorname{sgn}(\operatorname{par_2}(a) - \operatorname{par_2}(\hat a))$}}{
            $O_{\operatorname{acc}} \leftarrow O_{\operatorname{acc}} + \frac{1}{|\mathcal{A}'|}$ \;

### 2025-02-12 paper_jar/paper.tex
Add clarification

@@
 To this end, we use 46 instance features \change{(Section~\ref{sec:exdesign:data})} and 27 known solver runtimes as inputs for the predictions.
-We repeat this evaluation for each solver as the prediction target and report the average and standard deviation of prediction performance over the 28 solvers.
+\change{For this experiment, the data is split into a training set (80\% of all data) on which the target runtime is known and a test set (20\% of all data) on which we predict the runtimes of the new solver.}
+We repeat this evaluation for each solver as the prediction target and report the average and standard deviation of prediction performance over the 28 solvers \change{in Table~\ref{tab:regression}}.
 
@@
 Each is trained with either instance features only (second column) or instance and runtime features (third column).
+\change{For this experiment, the data is split into a training set (80\% of all data) on which the target label is known and a test set (20\% of all data) on which we predict the runtime label of the new solver.}
 As the evaluation metric for prediction performance, we use Matthews Correlation Coefficient (MCC)~\cite{gorodkin2004comparing, matthews1975comparison}, which reaches its maximum of~1 for a perfect prediction, is~0 for random guessing, and has a minimum of~-1.
 
-Training the classification models with instance features only, a random forest performs best.
+Training the classification models with instance features only \change{(as well as the target labels on the training set)}, a random forest performs best in predicting the target labels on the test set.
 In contrast, a quadratic discriminant analysis performs poorly in this scenario but performs very well if runtime features are included.
-However, the best prediction performance with all features is achieved by a stacking ensemble~\cite{wolpert1992stacked} out of the quadratic discriminant analysis~\cite{tharwat2016linear} and the random forest~\cite{breiman2001random}.
+However, the best prediction performance with all features \change{(as well as the target labels on the training set)} is achieved by a stacking ensemble~\cite{wolpert1992stacked} consisting of the quadratic discriminant analysis~\cite{tharwat2016linear} and the random forest~\cite{breiman2001random}.
 Stacking means that another model, in our case a simple decision tree~\cite{breiman1984classification}, \change{acts as a meta-model:
-The prediction outputs of the ensemble members, i.e., quadratic discriminant analysis and random forests, become the prediction inputs of the meta-model, i.e., decision tree.
-}
+The prediction outputs of the ensemble members, i.e., quadratic discriminant analysis and random forests, become the prediction inputs of the meta-model, i.e., decision tree.}
 
@@
 We train the prediction models either with 46 instance features \change{(Section~\ref{sec:exdesign:data})} and 27 known solver runtimes (second column) or additionally including 27 features representing the known $C_k$ cluster labels of other solvers (third column).
+\change{For this experiment, the data is split into a training set (80\% of all data) on which the target cluster label is known and a test set (20\% of all data) on which we predict the cluster label of the new solver.}
 
@@
 In this section, we evaluate our active learning framework.
+\change{While we split the data into a fixed train and test set to evaluate performance in Section~\ref{sec:exdesign:disc-pred}, in this section, the target runtime labels of the new solver are only known for the experiments that have already been selected by active-learning.}
 First, we analyze and tune the different subroutines of our framework on the tuning dataset (cf.~Section~\ref{sec:eval:hyper}).

### 2025-02-13 paper_jar/paper.tex
Revise

@@
 However, the performance differences are not significant for a Wilcoxon signed-rank test with $\alpha = 0.05$ and also depend on the fraction of the runtime sampled (cf.~Fig.~\ref{fig:annitraincolorselection}).
-However, a clear advantage of our approach is that it indicates when to stop adding new instances.
-While the active learning results on the full dataset are not as strong as on the smaller tuning dataset, they still show the benefit of making the benchmark selection dependent on the solvers to be distinguished.
+\change{However, an advantage of our approach is that it indicates when to stop adding new instances.
+Recall that Figure~\ref{fig:annitrainoptgoalacc} shows the trade-offs of the amount of sampled instances and the ranking accuracy.
+Active-learning helps to identify the optimal region as sampling too many instances only provides marginal improvements compared to its cost.}
 

### 2025-02-14 paper_jar/paper.tex
Disable tracked changes
Minor language changes
For consistency or other reasons.

@@
 \soulregister\pageref7
-\newcommand{\showchanges}{1} % set to 0 or 1
+\newcommand{\showchanges}{0} % set to 0 or 1
 \newcommand{\change}[1]{\ifnum \showchanges > 0 \sethlcolor{yellow}\hl{#1}\else#1\fi}
@@
 \change{Intuitively, uncertainty sampling selects the instances where it is most unclear in which runtime cluster the new solver resides.
-For example, if the new solver behaves similarly to two known solvers on the already run experiments, it is likely that this sampling strategy selects instances where these two solvers have runtimes that are not within the same cluster.}
+For example, if the new solver behaves similarly to two known solvers in the already run experiments, it is likely that this sampling strategy selects instances where these two solvers have runtimes that are not within the same cluster.}
 
@@
 We employ these models to estimate a new solver's raw runtime as the prediction target.
-To this end, we use 46 instance features \change{(Section~\ref{sec:exdesign:data})} and 27 known solver runtimes as inputs for the predictions.
-\change{For this experiment, the data is split into a training set (80\% of all data) on which the target runtime is known and a test set (20\% of all data) on which we predict the runtimes of the new solver.}
+To this end, we use 46 instance features \change{(cf.~Section~\ref{sec:exdesign:data})} and 27 known solver runtimes as inputs for the predictions.
+\change{For this experiment, we split the data into a training set (80\% of all data) on which the target runtime is known and a test set (20\% of all data) on which we predict the runtimes of the new solver.}
 We repeat this evaluation for each solver as the prediction target and report the average and standard deviation of prediction performance over the 28 solvers \change{in Table~\ref{tab:regression}}.
@@
 Each is trained with either instance features only (second column) or instance and runtime features (third column).
-\change{For this experiment, the data is split into a training set (80\% of all data) on which the target label is known and a test set (20\% of all data) on which we predict the runtime label of the new solver.}
+\change{For this experiment, we split the data into a training set (80\% of all data) on which the target label is known and a test set (20\% of all data) on which we predict the runtime label of the new solver.}
 As the evaluation metric for prediction performance, we use Matthews Correlation Coefficient (MCC)~\cite{gorodkin2004comparing, matthews1975comparison}, which reaches its maximum of~1 for a perfect prediction, is~0 for random guessing, and has a minimum of~-1.
@@
 Table~\ref{tab:cluster-prediction} shows the prediction performance for the best classification models from Table~\ref{tab:timeout-prediction} and $k \in \lbrace 3, 4, 5 \rbrace$ clusters.
-We train the prediction models either with 46 instance features \change{(Section~\ref{sec:exdesign:data})} and 27 known solver runtimes (second column) or additionally including 27 features representing the known $C_k$ cluster labels of other solvers (third column).
-\change{For this experiment, the data is split into a training set (80\% of all data) on which the target cluster label is known and a test set (20\% of all data) on which we predict the cluster label of the new solver.}
+We train the prediction models either with 46 instance features \change{(cf.~Section~\ref{sec:exdesign:data})} and 27 known solver runtimes (second column) or additionally including 27 features representing the known $C_k$ cluster labels of other solvers (third column).
+\change{For this experiment, we split the data into a training set (80\% of all data) on which the target cluster label is known and a test set (20\% of all data) on which we predict the cluster label of the new solver.}
 
@@
 In this section, we evaluate our active learning framework.
-\change{While we split the data into a fixed train and test set to evaluate performance in Section~\ref{sec:exdesign:disc-pred}, in this section, the target runtime labels of the new solver are only known for the experiments that have already been selected by active-learning.}
+\change{While we split the data into a fixed train and test set to evaluate performance in Section~\ref{sec:exdesign:disc-pred}, in this section, the target runtime labels of the new solver are only known for the experiments that have already been selected by active learning.}
 First, we analyze and tune the different subroutines of our framework on the tuning dataset (cf.~Section~\ref{sec:eval:hyper}).
@@
 \change{However, an advantage of our approach is that it indicates when to stop adding new instances.
-Recall that Figure~\ref{fig:annitrainoptgoalacc} shows the trade-offs of the amount of sampled instances and the ranking accuracy.
-Active-learning helps to identify the optimal region as sampling too many instances only provides marginal improvements compared to its cost.}
+Recall that Fig.~\ref{fig:annitrainoptgoalacc} shows the trade-off of the number of sampled instances and the ranking accuracy.
+Active learning helps to identify the optimal region as sampling too many instances only provides marginal improvements compared to its cost.}
 
@@
 Table~\ref{tab:features} briefly summarizes all 56 features.
-The placeholder \texttt{<n>} can take values within the range 1 to 9 and the placeholder \texttt{<dist>} is one of \texttt{mean}, \texttt{variance}, \texttt{min}, \texttt{max}, or \texttt{entropy}.
+The placeholder \texttt{<n>} can take values within the range from 1 to 9, and the placeholder \texttt{<dist>} is one of \texttt{mean}, \texttt{variance}, \texttt{min}, \texttt{max}, or \texttt{entropy}.
 Accounting for all possible instantiations of \texttt{<n>} and \texttt{<dist>}, there are 56 features in total.}
@@
       Horn proximity & \verb|hornvars_<dist>| & Distribution of variable occurrence in horn clauses \\
-      & \verb|invhornvars_<dist>| & Distribution of variable occurrence in inverted horn clauses \\
+      & \verb|invhornvars_<dist>| & Distribution of variable occurrence in inverted Horn clauses \\
       \midrule

### 2025-05-17 paper_jar/paper.tex
Remove change markers
Not visible anyway, but there is a small probability that they mess up
typesetting somehow, so remove them altogether.

@@
 
-% Highlight changes for re-submission (might be removed later)
-\usepackage{soul}
-\soulregister\cite7 % some hacking to make highlighting work in special situations, e.g., if text contains references: https://tex.stackexchange.com/q/139463
-\soulregister\ref7
-\soulregister\pageref7
-\newcommand{\showchanges}{0} % set to 0 or 1
-\newcommand{\change}[1]{\ifnum \showchanges > 0 \sethlcolor{yellow}\hl{#1}\else#1\fi}
-
 \raggedbottom
@@
   Let us consider a set of solvers~$\mathcal{A}$, instances~$\mathcal{I}$, and runtimes~${r\!: \mathcal{A} \times \mathcal{I} \rightarrow \left[0, \tau\right]}$, with each solver having a runtime limit of $\tau$.
-  The \emph{New-Solver Problem} is \change{about selecting} benchmark instances from $\mathcal{I}$ to maximize the confidence in predicting the rank of a new solver $\hat{a} \notin \mathcal{A}$, while minimizing the \change{summed runtime of~$\hat{a}$ over all selected instances}.
+  The \emph{New-Solver Problem} is about selecting benchmark instances from $\mathcal{I}$ to maximize the confidence in predicting the rank of a new solver $\hat{a} \notin \mathcal{A}$, while minimizing the summed runtime of~$\hat{a}$ over all selected instances.
   \label{def:new-solver-problem}
@@
 
-\change{
 The two objectives of this problem are conflicting, i.e., the ranking's confidence typically benefits from more instances while the runtime criterion encourages less or at least shorter evaluations.
 The crux is finding a good trade-off.
-}
 
@@
 Rather than outputting a binary classification, namely whether the new solver is worse than an existing solver or not, we provide a \emph{scoring} function that shows by which margin a solver is worse and how similar its performance is to existing solvers.
-\change{This helps to identify the solvers that have the most similar runtimes more quickly.}
-In particular, our approach enables \emph{ranking} the new solver among a set of existing solvers \change{in one benchmarking run}.
+This helps to identify the solvers that have the most similar runtimes more quickly.
+In particular, our approach enables \emph{ranking} the new solver among a set of existing solvers in one benchmarking run.
 We show that predicting the exact per-instance runtime of the solvers, which would be a more challenging task, is not necessary for ranking.
@@
 Their approach ticks off most of the desired features in Table~\ref{tab:requirements}.
-However, the approach only compares solvers \change{pairwise} rather than providing a multi-solver \emph{scoring} or \emph{ranking}.
+However, the approach only compares solvers pairwise rather than providing a multi-solver \emph{scoring} or \emph{ranking}.
 Consequently, it is unclear how similar two given solvers are or on which instances they behave similarly.
 Furthermore, a significant shortcoming is the lack of \emph{scalability} with the number of solvers.
-Due to comparing only pairs of solvers, evaluating a new solver with their method requires sampling \change{an individual set of benchmark instances} for each existing solver.
-\change{In particular, comparing against one existing solver will typically require a different set of instances than comparing against another one.}
-In contrast, our approach allows comparing a new solver against a set of existing solvers by sampling only one \change{set of} benchmark \change{instances}.
+Due to comparing only pairs of solvers, evaluating a new solver with their method requires sampling an individual set of benchmark instances for each existing solver.
+In particular, comparing against one existing solver will typically require a different set of instances than comparing against another one.
+In contrast, our approach allows comparing a new solver against a set of existing solvers by sampling only one set of benchmark instances.
 
@@
 In our scenario, the first goal is to maximize the utility of instances to our prediction model, which is measured by the ranking accuracy.
-Our second goal is to minimize the cost of label acquisition, which is \change{estimated} by the instance's predicted runtime.
+Our second goal is to minimize the cost of label acquisition, which is estimated by the instance's predicted runtime.
 Consequently, our overall objective is to develop an accurate prediction model without labeling each data point.
@@
 
-The model~\change{$\mathcal{M}$} provides a prediction function $f_{\hat a} : \mathcal{I} \rightarrow \mathbb{R}$ for the new solver $\hat a$.
+The model~$\mathcal{M}$ provides a prediction function $f_{\hat a} : \mathcal{I} \rightarrow \mathbb{R}$ for the new solver $\hat a$.
 This prediction function powers instance selection as described in Section~\ref{sec:main:selection}.
 In the update step (Algorithm~\ref{algALBenchmark},~Line~7), the model is trained to predict a transformed version of solver runtime (cf.~Section~\ref{sec:main:model:transformation}) for the solver~$\hat{a}$, using the features described in Section~\ref{sec:exdesign:data} and the previously acquired runtimes~$\mathcal{R}$.
-Each iteration trains a new prediction model from scratch since the time for running the solver~$\hat{a}$ on one instance~$e$ (Line~5) dominates the training time of the model~\change{$\mathcal{M}$} on all acquired instances~$\mathcal{R}$ by a significant margin.
+Each iteration trains a new prediction model from scratch since the time for running the solver~$\hat{a}$ on one instance~$e$ (Line~5) dominates the training time of the model~$\mathcal{M}$ on all acquired instances~$\mathcal{R}$ by a significant margin.
 Once the benchmarking loop terminates, the score of the solver~$\hat{a}$ is predicted and then returned (Lines~8--9).
@@
 
-For the prediction model~\change{$\mathcal{M}$}, we transform the real-valued runtimes into discrete runtime labels on a per-instance basis.
+For the prediction model~$\mathcal{M}$, we transform the real-valued runtimes into discrete runtime labels on a per-instance basis.
 For each instance $e \in \mathcal{I}$, we use a clustering algorithm to assign the runtimes in $\bigl\{ r(a, e) \mid a \in \mathcal{A} \bigr\}$ to one of $k$ clusters $C_1, \dots, C_k$ such that the fastest runtimes for the benchmark instance $e$ are in cluster $C_1$ and the slowest are in cluster $C_{k-1}$.
 Timeouts $\tau$ always form a separate cluster $C_{k}$.
-\change{The clustering is performed on a per-instance level as the best runtimes may be seconds on one instance and minutes on another.
-After clustering, the best runtimes are all in the same cluster.}
+The clustering is performed on a per-instance level as the best runtimes may be seconds on one instance and minutes on another.
+After clustering, the best runtimes are all in the same cluster.
 The runtime transformation function $\gamma_k : {\mathcal{A} \times \mathcal{I}} \rightarrow \left\lbrace 1, \dots, k \right\rbrace$ is then specified as follows:
@@
 Empirical studies on portfolio solvers have demonstrated that discretization is an effective approach in practice~\cite{CollauttiMMO13,NgokoCT19}.
-Section~\ref{sec:exdesign:disc-pred} presents a detailed analysis of the runtime transformation in our scenario\change{, including the clustering approach}.
+Section~\ref{sec:exdesign:disc-pred} presents a detailed analysis of the runtime transformation in our scenario, including the clustering approach.
 In particular, the analysis demonstrates that our benchmarking approach achieves higher ranking accuracy when working with discrete runtime labels rather than raw runtimes.
@@
 %
-In \change{this way}, timeouts are still penalized.
+In this way, timeouts are still penalized.
 
@@
 \end{equation*}
-\change{Intuitively, uncertainty sampling selects the instances where it is most unclear in which runtime cluster the new solver resides.
-For example, if the new solver behaves similarly to two known solvers in the already run experiments, it is likely that this sampling strategy selects instances where these two solvers have runtimes that are not within the same cluster.}
+%
+Intuitively, uncertainty sampling selects the instances where it is most unclear in which runtime cluster the new solver resides.
+For example, if the new solver behaves similarly to two known solvers in the already run experiments, it is likely that this sampling strategy selects instances where these two solvers have runtimes that are not within the same cluster.
 
@@
 The term $\operatorname{\hat H}(e)_n$ is computed for every possible runtime label $n \in \{1, \dots, k\}$.
-\change{Intuitively, this strategy favors instances where the most likely predicted cluster labels (as determined by $f_{\hat a}'$) would provide the most insight regarding similarities to known solvers.
+Intuitively, this strategy favors instances where the most likely predicted cluster labels (as determined by $f_{\hat a}'$) would provide the most insight regarding similarities to known solvers.
 For example, on a particular instance, one group of solvers might run into timeouts and another group of solvers might return a solution.
-In this case, the information-gain strategy will likely select this instance as it provides insight about the group of solvers the new solver is similar to.}
+In this case, the information-gain strategy will likely select this instance as it provides insight about the group of solvers the new solver is similar to.
 
@@
 In this section, we present the two dynamic stopping criteria (cf.~Algorithm~\ref{algALBenchmark}, Line~3) employed in our experiments: the Wilcoxon and the ranking stopping criterion.
-\change{
 In the worst case, both criteria sample all instances, but typically they terminate earlier.
 How early they terminate depends on how easy it is to determine the rank of the new solver among the existing ones.
-}
 
@@
     \For{$a \in \mathcal{A}'$}{
-      \If(\tcp*[f]{\change{Same sign/order}}){\change{$\operatorname{sgn}(s_k(a) - s_{\hat a}) = \operatorname{sgn}(\operatorname{par_2}(a) - \operatorname{par_2}(\hat a))$}}{
+      \If(\tcp*[f]{Same sign/order}){$\operatorname{sgn}(s_k(a) - s_{\hat a}) = \operatorname{sgn}(\operatorname{par_2}(a) - \operatorname{par_2}(\hat a))$}{
            $O_{\operatorname{acc}} \leftarrow O_{\operatorname{acc}} + \frac{1}{|\mathcal{A}'|}$ \;
@@
 For predictions within our framework, we also use a database of 56 instance features from the Global Benchmark Database~(GBD)~\cite{IserJ24}.
-\change{Refer to Appendix~\ref{sec:appendix-features} for a full list of all 56 features.}
+Refer to Appendix~\ref{sec:appendix-features} for a full list of all 56 features.
 They include instance size features and node distribution statistics for several graph representations of SAT instances, and are primarily inspired by the SATzilla~2012 features described in~\cite{features}.
@@
 We employ these models to estimate a new solver's raw runtime as the prediction target.
-To this end, we use 46 instance features \change{(cf.~Section~\ref{sec:exdesign:data})} and 27 known solver runtimes as inputs for the predictions.
-\change{For this experiment, we split the data into a training set (80\% of all data) on which the target runtime is known and a test set (20\% of all data) on which we predict the runtimes of the new solver.}
-We repeat this evaluation for each solver as the prediction target and report the average and standard deviation of prediction performance over the 28 solvers \change{in Table~\ref{tab:regression}}.
+To this end, we use 46 instance features (cf.~Section~\ref{sec:exdesign:data}) and 27 known solver runtimes as inputs for the predictions.
+For this experiment, we split the data into a training set (80\% of all data) on which the target runtime is known and a test set (20\% of all data) on which we predict the runtimes of the new solver.
+We repeat this evaluation for each solver as the prediction target and report the average and standard deviation of prediction performance over the 28 solvers in Table~\ref{tab:regression}.
 
@@
 Each is trained with either instance features only (second column) or instance and runtime features (third column).
-\change{For this experiment, we split the data into a training set (80\% of all data) on which the target label is known and a test set (20\% of all data) on which we predict the runtime label of the new solver.}
+For this experiment, we split the data into a training set (80\% of all data) on which the target label is known and a test set (20\% of all data) on which we predict the runtime label of the new solver.
 As the evaluation metric for prediction performance, we use Matthews Correlation Coefficient (MCC)~\cite{gorodkin2004comparing, matthews1975comparison}, which reaches its maximum of~1 for a perfect prediction, is~0 for random guessing, and has a minimum of~-1.
 
-Training the classification models with instance features only \change{(as well as the target labels on the training set)}, a random forest performs best in predicting the target labels on the test set.
+Training the classification models with instance features only (as well as the target labels on the training set), a random forest performs best in predicting the target labels on the test set.
 In contrast, a quadratic discriminant analysis performs poorly in this scenario but performs very well if runtime features are included.
-However, the best prediction performance with all features \change{(as well as the target labels on the training set)} is achieved by a stacking ensemble~\cite{wolpert1992stacked} consisting of the quadratic discriminant analysis~\cite{tharwat2016linear} and the random forest~\cite{breiman2001random}.
-Stacking means that another model, in our case a simple decision tree~\cite{breiman1984classification}, \change{acts as a meta-model:
-The prediction outputs of the ensemble members, i.e., quadratic discriminant analysis and random forests, become the prediction inputs of the meta-model, i.e., decision tree.}
+However, the best prediction performance with all features (as well as the target labels on the training set) is achieved by a stacking ensemble~\cite{wolpert1992stacked} consisting of the quadratic discriminant analysis~\cite{tharwat2016linear} and the random forest~\cite{breiman2001random}.
+Stacking means that another model, in our case a simple decision tree~\cite{breiman1984classification}, acts as a meta-model:
+The prediction outputs of the ensemble members, i.e., quadratic discriminant analysis and random forests, become the prediction inputs of the meta-model, i.e., decision tree.
 
@@
 Table~\ref{tab:cluster-prediction} shows the prediction performance for the best classification models from Table~\ref{tab:timeout-prediction} and $k \in \lbrace 3, 4, 5 \rbrace$ clusters.
-We train the prediction models either with 46 instance features \change{(cf.~Section~\ref{sec:exdesign:data})} and 27 known solver runtimes (second column) or additionally including 27 features representing the known $C_k$ cluster labels of other solvers (third column).
-\change{For this experiment, we split the data into a training set (80\% of all data) on which the target cluster label is known and a test set (20\% of all data) on which we predict the cluster label of the new solver.}
+We train the prediction models either with 46 instance features (cf.~Section~\ref{sec:exdesign:data}) and 27 known solver runtimes (second column) or additionally including 27 features representing the known $C_k$ cluster labels of other solvers (third column).
+For this experiment, we split the data into a training set (80\% of all data) on which the target cluster label is known and a test set (20\% of all data) on which we predict the cluster label of the new solver.
 
@@
 In this section, we evaluate our active learning framework.
-\change{While we split the data into a fixed train and test set to evaluate performance in Section~\ref{sec:exdesign:disc-pred}, in this section, the target runtime labels of the new solver are only known for the experiments that have already been selected by active learning.}
+While we split the data into a fixed train and test set to evaluate performance in Section~\ref{sec:exdesign:disc-pred}, in this section, the target runtime labels of the new solver are only known for the experiments that have already been selected by active learning.
 First, we analyze and tune the different subroutines of our framework on the tuning dataset (cf.~Section~\ref{sec:eval:hyper}).
@@
 However, the performance differences are not significant for a Wilcoxon signed-rank test with $\alpha = 0.05$ and also depend on the fraction of the runtime sampled (cf.~Fig.~\ref{fig:annitraincolorselection}).
-\change{However, an advantage of our approach is that it indicates when to stop adding new instances.
+However, an advantage of our approach is that it indicates when to stop adding new instances.
 Recall that Fig.~\ref{fig:annitrainoptgoalacc} shows the trade-off of the number of sampled instances and the ranking accuracy.
-Active learning helps to identify the optimal region as sampling too many instances only provides marginal improvements compared to its cost.}
+Active learning helps to identify the optimal region as sampling too many instances only provides marginal improvements compared to its cost.
 
@@
 
-\change{For predictions within our framework, we use a database of 56 instance features\footnote{\url{https://udopia.github.io/gbdc/doc/extractors/CNFBaseFeatures.html}} from the Global Benchmark Database~(GBD)~\cite{IserJ24}.
+For predictions within our framework, we use a database of 56 instance features\footnote{\url{https://udopia.github.io/gbdc/doc/extractors/CNFBaseFeatures.html}} from the Global Benchmark Database~(GBD)~\cite{IserJ24}.
 Table~\ref{tab:features} briefly summarizes all 56 features.
 The placeholder \texttt{<n>} can take values within the range from 1 to 9, and the placeholder \texttt{<dist>} is one of \texttt{mean}, \texttt{variance}, \texttt{min}, \texttt{max}, or \texttt{entropy}.
-Accounting for all possible instantiations of \texttt{<n>} and \texttt{<dist>}, there are 56 features in total.}
+Accounting for all possible instantiations of \texttt{<n>} and \texttt{<dist>}, there are 56 features in total.
 
@@
     \caption{
-      \change{List of all 56 instance features, where \texttt{<dist>} is one of \texttt{mean}, \texttt{variance}, \texttt{min}, \texttt{max}, or \texttt{entropy}.}
+      List of all 56 instance features, where \texttt{<dist>} is one of \texttt{mean}, \texttt{variance}, \texttt{min}, \texttt{max}, or \texttt{entropy}.
     }
@@
       \midrule
-      Horn proximity & \verb|hornvars_<dist>| & Distribution of variable occurrence in horn clauses \\
+      Horn proximity & \verb|hornvars_<dist>| & Distribution of variable occurrence in Horn clauses \\
       & \verb|invhornvars_<dist>| & Distribution of variable occurrence in inverted Horn clauses \\
